### 2019-04-25 TrueBioTest.Rmd
initialize multiple atlantis models

---
title: "Testing atlantisom: true biomass output from Atlantis"
author: "Sarah Gaichas and Christine Stawitz"
date: "April 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.

The purpose of atlantisom is to use existing Atlantis model output to generate input datasets for a variety of models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics. Atlantis models can be run using different climate forcing, fishing, and other scenarios. Users of atlantisom will be able to specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes for any Atlantis scenario. Internally consistent multispecies and ecosystem datasets with known observation error characteristics will be the atlantisom outputs, for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against "true" Atlantis outputs.

Initial testing was conducted by S. Gaichas using R scripts in the R folder of this repository that are titled "PoseidonTest_[whatwastested].R". Initial tests are expanded and documented in more detail in these pages. C. Stawitz improved and streamlined the setup and intialization sections.

## Setup

First, you will want to load packages and initialize the correct file and working directory names.

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
```

If you want to load the local of the atlantisom R package, use devtools. The R package here [@`citation(here)'] is helpful for installing from your local atlantisom directory, as it avoids hardcoding the specific location. The code below is not running from our local atlantisom directory and is not evaluated:

```{r eval=FALSE, message=FALSE, warning=FALSE}
require(devtools)
package.dir <- here()
devtools::load_all(package.dir)
library(atlantisom)
```

Or you can install directly from the Github repository.

```{r eval=FALSE, message=FALSE, warning=FALSE}
devtools::install_github("r4atlantis\atlantisom")
```

## Initializing input files and directories

You will first need to tell `atlantisom` to know where to look for the output and input files from your atlantis model run. Here, we will give the directory where the Atlantis inputs and outputs are stored `d.name`, the location of the functional groups file `functional.group.file` (.csv), the biomass pools file `biomass.pools.file` (.nc), the box locations file `box.file` (.bgm), an `initial.conditions.file` (.nc), the biology .prm file `biol.prm.file` (.prm), and the run .prm file `run.prm.file` (.prm). You will also need to specify a scenario name, which will be used to define the output files (i.e. output is stored in a number of netCDF files of the format: output<scenario><value>.nc). All of these files should be stored in `d.name`.

In general, Atlantis model output files that are sufficiently detailed to mimic fishery sampling in space and time are too large to store on GitHub, so are not included as examples with the atlantisom code.

In this example, we have a local folder "atlantisoutput" with three subdirectories:
  * "CalCurrentSummitScenario1": California Current Atlantis model output
  * "NEUStest20160303": older (broken) Northeast US Atlantis model output
  * "NOBACERESGlobalSustainability": Norwegian-Barents Sea Atlantis model output

Our directory structure is set up to take advantage of `here()' to allow setup on a different computer.

```{r initialize}

initCCA <- TRUE
initNEUS <- FALSE
initNOBA <- FALSE

if(initCCA){
  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
  functional.groups.file <- "CalCurrentV3Groups.csv"
  biomass.pools.file <- "DIVCalCurrentV3_BIOL.nc"
  biol.prm.file <- "CalCurrentV3_Biol.prm"
  box.file <- "CalCurrentV3_utm.bgm"
  initial.conditions.file <- "DIVCalCurrentV3_BIOL.nc"
  run.prm.file <- "CalCurrentV3_run.xml"
  scenario.name <- "CCV3"
}

if(initNEUS){
  d.name <- here("atlantisoutput","NEUStest20160303")
  functional.groups.file <- "NeusGroups.csv" 
  biomass.pools.file <- ""
  biol.prm.file <- "at_biol_neus_v15_DE.prm"
  box.file <- "neus30_2006.bgm"
  initial.conditions.file <- "inneus_2012.nc"
  run.prm.file <- "at_run_neus_v15_DE.xml"
  scenario.name <- "neusDynEffort_Test1_"
}

if(initNOBA){
  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
  functional.groups.file <- "nordic_groups_v04.csv" 
  biomass.pools.file <- "nordic_biol_v23.nc"
  biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
  box.file <- "Nordic02.bgm"
  initial.conditions.file <- ".nc"
  run.prm.file <- "nordic_run_v01.prm"
  scenario.name <- "nordic_runresults_01"
}

```

## Getting the "true" operating model values

There are a number of functions in the package that begin with the prefix `load` that load various files. See documentation if you'd only like to load one file. The `atlantisom::run_truth()` function uses the above file definitions and calls a number of the `load` functions to read in all of the atlantis output. Note: this call reads in a number of large .nc files, so it will take a few minutes to return.

```{r get_truth, message=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

#Store all loaded results into an R object
results <- run_truth(scenario = scenario.name,
          dir = d.name,
          file_fgs = functional.groups.file,
          file_bgm = box.file,
          select_groups = funct.group.names,
          file_init = initial.conditions.file,
          file_biolprm = biol.prm.file,
          file_runprm = run.prm.file
)

if(initCCA) CCAresults <- results
if(initNEUS) NEUSresults <- results
if(initNOBA) NOBAresults <- results
```

Now the R object `results` with the comprehensive results from the Atlantis model has been read in.

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2019-04-25 TrueBioTest.Rmd
add detail on run_truth output storage
adding surveys, in progress, not working

@@
 
-First, you will want to load packages and initialize the correct file and working directory names.
-
-```{r message=FALSE, warning=FALSE}
-library(tidyr)
-require(dplyr)
-library(ggplot2)
-library(data.table)
-library(here)
-```
+First, you will want to install atlantisom if you haven't already.
 
@@
 devtools::load_all(package.dir)
-library(atlantisom)
 ```
@@
 devtools::install_github("r4atlantis\atlantisom")
+```
+
+Next, load packages.
+
+```{r message=FALSE, warning=FALSE}
+library(tidyr)
+require(dplyr)
+library(ggplot2)
+library(data.table)
+library(here)
+library(atlantisom)
 ```
@@
 
-initCCA <- TRUE
-initNEUS <- FALSE
+initCCA <- FALSE
+initNEUS <- TRUE
 initNOBA <- FALSE
@@
 
-```{r get_truth, message=FALSE}
+```{r get_truth, message=FALSE, warning=FALSE, eval=FALSE}
 #Load functional groups
@@
 Now the R object `results` with the comprehensive results from the Atlantis model has been read in.
+It is also saved as an .RData file titled "output[scenario.name]run_truth.RData" in the directory with the model output, so that later analyses can use `load()' instead of taking the time to rerun `atlantisom::run_truth()'.
 
-```{r cars}
-summary(cars)
+## Simulate a survey part 1: census to compare with truth
+
+This section tests the `atlantisom::create_survey()' and `atlantisom::sample_survey_biomass()' functions by comparing a census (survey sampling everything) with the results generated by `atlantisom::run_truth()' above.
+
+To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficieny ("q"), and the selectivity at age for each species.
+
+The following settings should achieve a survey that samples all Atlantis model output timesteps, all species, and all model polygons, with perfect efficiency and full selectivity for all ages:
+
+```{r census-spec, message=FALSE, warning=FALSE}
+
+# should return a perfectly scaled survey 
+effic1 <- data.frame(species=funct.group.names,
+                     efficiency=rep(1.0,length(funct.group.names)))
+
+# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
+selex1 <- data.frame(species=rep(funct.group.names, each=10),
+                     agecl=rep(c(1:10),length(funct.group.names)),
+                     selex=rep(1.0,length(funct.group.names)*10))
+
+# should return all model areas; need to generalize
+if(initCCA)
+if(initNEUS) boxall <- c(0:29)
+if(initNOBA)
+
+# should return all model output timesteps; need to generalize
+if(initCCA)
+if(initNEUS) timeall <- c(0:251)
+if(initNOBA)
+
+# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
+# for ecosystem indicator work test all species
+# to keep plots simpler, currently hardcoded for vertebrate groups
+if(initCCA)
+if(initNEUS) survspp <- funct.group.names[1:21]
+if(initNOBA)
+
 ```
 
-## Including Plots
+Because the results of `run_truth' provide both biomass at age and numbers at age, we can use `create_survey' on both with some modifications.
 
-You can also embed plots, for example:
+```{r surveyBbased}
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+# this uses result$biomass_ages to sample biomass directly
+
+if(initCCA) datB <- CCAresults$biomass_ages
+if(initNEUS) datB <- NEUSresults$biomass_ages
+if(initNOBA) datB <- NOBAresults$biomass_ages
+
+survey_testBall <- create_survey(dat = datB,
+                                 time = timeall,
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
 ```
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+```{r surveyNbased}
+
+# this uses result$nums so will need a weight at age to get biomass
+
+if(initCCA) datN <- CCAresults$nums
+if(initNEUS) datN <- NEUSresults$nums
+if(initNOBA) datN <- NOBAresults$nums
+
+survey_testNall <- create_survey(dat = datN,
+                                 time = timeall,
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
+```

### 2019-04-26 TrueBioTest.Rmd
add NOBA files, separate functions

@@
   box.file <- "Nordic02.bgm"
-  initial.conditions.file <- ".nc"
-  run.prm.file <- "nordic_run_v01.prm"
+  initial.conditions.file <- "nordic_biol_v23.nc"
+  run.prm.file <- "nordic_run_v01.xml"
   scenario.name <- "nordic_runresults_01"
@@
 
-```{r get_truth, message=FALSE, warning=FALSE, eval=FALSE}
+```{r get_names, message=FALSE, warning=FALSE}
 #Load functional groups
@@
 
+```
+
+```{r get_truth, message=FALSE, warning=FALSE, eval = FALSE}
 #Store all loaded results into an R object
@@
 if(initNOBA) NOBAresults <- results
+
 ```
@@
 # should return all model areas; need to generalize
-if(initCCA)
+if(initCCA) boxall <- c(0:29) #placeholder, correct for CCA
 if(initNEUS) boxall <- c(0:29)
-if(initNOBA)
+if(initNOBA) boxall <- c(0:29) #placeholder, correct for NOBA
 
 # should return all model output timesteps; need to generalize
-if(initCCA)
+if(initCCA) timeall <- c(0:251) #placeholder, correct for CCA
 if(initNEUS) timeall <- c(0:251)
-if(initNOBA)
+if(initNOBA) timeall <- c(0:251) #placeholder, correct for NOBA
 
@@
 # to keep plots simpler, currently hardcoded for vertebrate groups
-if(initCCA)
+if(initCCA) survspp <- funct.group.names[1:21] #placeholder, correct for CCA
 if(initNEUS) survspp <- funct.group.names[1:21]
-if(initNOBA)
+if(initNOBA) survspp <- funct.group.names[1:21] #placeholder, correct for NOBA
 
@@
 
-```{r surveyBbased}
+```{r surveyBbased, eval = FALSE}
 
@@
 
-```{r surveyNbased}
+```{r surveyNbased , eval = FALSE}
 

### 2019-04-26 TrueBioTest.Rmd
rearrange setup again

@@
 
-First, you will want to install atlantisom if you haven't already.
+First, you will want to set up libraries and install atlantisom if you haven't already.
+
+```{r message=FALSE, warning=FALSE}
+library(tidyr)
+require(dplyr)
+library(ggplot2)
+library(data.table)
+library(here)
+```
 
@@
 
-Next, load packages.
+Next, load the package.
 
 ```{r message=FALSE, warning=FALSE}
-library(tidyr)
-require(dplyr)
-library(ggplot2)
-library(data.table)
-library(here)
 library(atlantisom)

### 2019-04-26 TrueBioTest.Rmd
formatting, try NOBA

@@
 In this example, we have a local folder "atlantisoutput" with three subdirectories:
+
   * "CalCurrentSummitScenario1": California Current Atlantis model output
@@
 initCCA <- FALSE
-initNEUS <- TRUE
-initNOBA <- FALSE
+initNEUS <- FALSE
+initNOBA <- TRUE
 
@@
 }
-
+# NOBA note: output filenames in CCA and NEUS begin with "output" and the run_truth function is written to expect this.
+Need to check if default Atlantis output file nomenclature has changed or if NOBA is a special case.
+For now, NOBA filenames have been changed to include prefix "output"
 ```
@@
 
-```{r get_truth, message=FALSE, warning=FALSE, eval = FALSE}
+```{r get_truth, message=FALSE, warning=FALSE}
 #Store all loaded results into an R object
@@
 Now the R object `results` with the comprehensive results from the Atlantis model has been read in.
-It is also saved as an .RData file titled "output[scenario.name]run_truth.RData" in the directory with the model output, so that later analyses can use `load()' instead of taking the time to rerun `atlantisom::run_truth()'.
+It is also saved as an .RData file titled "output[scenario.name]run_truth.RData" in the directory with the model output, so that later analyses can use 'load()' instead of taking the time to rerun `atlantisom::run_truth()'.
 
@@
 
-This section tests the `atlantisom::create_survey()' and `atlantisom::sample_survey_biomass()' functions by comparing a census (survey sampling everything) with the results generated by `atlantisom::run_truth()' above.
+This section tests the `atlantisom::create_survey()` and `atlantisom::sample_survey_biomass()` functions by comparing a census (survey sampling everything) with the results generated by `atlantisom::run_truth()` above.
 
@@
 
-```{r census-spec, message=FALSE, warning=FALSE}
+```{r census-spec, message=FALSE, warning=FALSE, eval=FALSE}
 
@@
 
-Because the results of `run_truth' provide both biomass at age and numbers at age, we can use `create_survey' on both with some modifications.
+Because the results of `run_truth` provide both biomass at age and numbers at age, we can use `create_survey` on both with some modifications.
 

### 2019-04-26 TrueBioTest.Rmd
generalize create survey census specifications

@@
 
-```{r get_truth, message=FALSE, warning=FALSE}
+```{r get_truth, message=FALSE, warning=FALSE, eval=FALSE}
 #Store all loaded results into an R object
@@
 
-```{r census-spec, message=FALSE, warning=FALSE, eval=FALSE}
+```{r census-spec, message=FALSE, warning=FALSE}
 
@@
 # should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
+# BUT CHECK if newer Atlantis models can do age-specific outputs
 selex1 <- data.frame(species=rep(funct.group.names, each=10),
@@
 
-# should return all model areas; need to generalize
-if(initCCA) boxall <- c(0:29) #placeholder, correct for CCA
-if(initNEUS) boxall <- c(0:29)
-if(initNOBA) boxall <- c(0:29) #placeholder, correct for NOBA
+# should return all model areas
+boxpars <- load_box(d.name, box.file)
+boxall <- c(0:(boxpars$nbox - 1))
+
+# these are model specific, generalized above
+# if(initCCA) boxall <- c(0:88) 
+# if(initNEUS) boxall <- c(0:29)
+# if(initNOBA) boxall <- c(0:59)
 
 # should return all model output timesteps; need to generalize
-if(initCCA) timeall <- c(0:251) #placeholder, correct for CCA
+if(initCCA) timeall <- c(0:100) 
 if(initNEUS) timeall <- c(0:251)
-if(initNOBA) timeall <- c(0:251) #placeholder, correct for NOBA
+if(initNOBA) timeall <- c(0:560)
 
 # define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
-# for ecosystem indicator work test all species
-# to keep plots simpler, currently hardcoded for vertebrate groups
-if(initCCA) survspp <- funct.group.names[1:21] #placeholder, correct for CCA
+# for ecosystem indicator work test all species, e.g.
+survspp <- funct.group.names
+
+# to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
+if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
 if(initNEUS) survspp <- funct.group.names[1:21]
-if(initNOBA) survspp <- funct.group.names[1:21] #placeholder, correct for NOBA
+if(initNOBA) survspp <- funct.group.names[1:36]
 

### 2019-04-26 docs/index.Rmd
add docs folder with html

---
title: "Poseidon Adventure"
output: html_document
---

 <!-- ADD LINKS BELOW, note, using the RMD link syntax will open the presentation in the same tab, which is annoying. Use the html with the "_blank" to fix that behavior -->

These pages document initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.

### Testing the atlantisom code

<a href="https://sgaichas.github.io/presentations/TrueBioTest.html" target="_blank">Installation and initial tests - April 26, 2019</a>

### More to come later!

### 2019-04-29 TrueBioTest.Rmd
add sample_survey_biomass example 1

@@
 Because the results of `run_truth` provide both biomass at age and numbers at age, we can use `create_survey` on both with some modifications.
+The following uses the biomass output of `run_truth` to create the survey, so the call to `sample_survey_biomass` will require a weight at age argument that is filled with 1's because no conversion from numbers to weight is necessary.
+Because we are making a census for testing, the survey cv argument is set to 0.
 
@@
                                  selex = selex1)
+
+# make up a constant 0 cv for testing
+surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+
+# call sample_survey_biomass with a bunch of 1s for weight at age
+# in the code it multiplies atoutput by wtatage so this allows us to use
+# biomass directly
+wtage <- data.frame(species=rep(survspp, each=10),
+                    agecl=rep(c(1:10),length(survspp)),
+                    wtAtAge=rep(1.0,length(survspp)*10))
+
+surveyB_frombio <- sample_survey_biomass(survey_testBall, surv_cv, wtage)
+
+if(initCCA) CCAsurveyB_frombio <- surveyB_frombio
+if(initNEUS) NEUSsurveyB_frombio <- surveyB_frombio
+if(initNOBA) NOBAsurveyB_frombio <- surveyB_frombio
 

### 2019-04-29 TrueBioTest.Rmd
read in RData truth each model

@@
 
+Note: on Sarah's laptop, CCA took ~40 minutes to return from `run_truth` and NOBA took an hour and 26 minutes to return.
+Therefore, reading the RData file in as below is advised after the first run to make the plots comparing systems later in this document.
+
+```{r load_Rdata, message=FALSE, warning=FALSE, eval=FALSE}
+
+if(initCCA) {
+  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
+  truth.file <- "outputCCV3run_truth.RData"
+  load(file.path(d.name, truth.file))
+  CCAresults <- result
+}
+
+if(initNEUS) {
+  d.name <- here("atlantisoutput","NEUStest20160303")
+  truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
+  load(file.path(d.name, truth.file))
+  NEUSresults <- result
+}
+
+if(initNOBA){
+  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
+  truth.file <- "outputnordic_runresults_01run_truth.RData" 
+  load(file.path(d.name, truth.file))
+  NOBAresults <- result
+}
+
+```
+
 ## Simulate a survey part 1: census to compare with truth

### 2019-04-29 TrueBioTest.Rmd
add N based survey and plotting

@@
 library(here)
+library(ggforce)
+library(ggthemes)
 ```
@@
 
-```{r load_Rdata, message=FALSE, warning=FALSE, eval=FALSE}
+```{r load_Rdata, message=FALSE, warning=FALSE}
 
@@
 
-```{r surveyNbased , eval = FALSE}
+Comparing our (census) survey based on true biomass from above with the Atlantis output file "[modelscenario]BiomIndx.txt" should give us a perfect match.
+Note that the our (census) survey may have more sampling in time than the Atlantis output file.
+
+```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species."}
+# plot some comparisons with Atlantis output
+
+# read Atlantis output files
+if(initCCA) {
+  atBtxt2 <- read.table(here("atlantisoutput","CalCurrentSummitScenario1","outputCCV3BiomIndx.txt"), header=T)
+  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
+  surveyB_frombio <- CCAsurveyB_frombio
+}
+
+if(initNEUS) {
+  atBtxt2 <- read.table(here("atlantisoutput","NEUStest20160303","neusDynEffort_Test1_BiomIndx.txt"), header=T)
+  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
+  surveyB_frombio <- NEUSsurveyB_frombio
+}
+
+if(initNOBA) {
+  atBtxt2 <- read.table(here("atlantisoutput","NOBACERESGlobalSustainability","outputnordic_runresults_01BiomIndx.txt"), header=T) 
+  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
+  surveyB_frombio <- NOBAsurveyB_frombio
+}
+
+# lookup the matching names, put in time, species, biomass column format
+# WARNING hardcoded for output with last species group as DIN
+groupslookup <- groupslookup %>%
+  filter(IsTurnedOn > 0)
+
+atBtxt2tidy <- atBtxt2 %>%
+  select(Time:DIN) %>%
+  #select(Time, FPL:DIN) %>%
+  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
+  gather(species, biomass, -Time) %>%
+  filter(species == levels(surveyB_frombio$species))
+
+#all species comparison, time intervals hardcoded for NEUS and NOBA
+compareB <-ggplot() +
+  geom_line(data=surveyB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
+            alpha = 3/10) +
+  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
+             alpha = 3/10) + 
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+compareB + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+
+```
+
+```{r matchBp2, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
+compareB + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+
+```
+
+```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species.
+Depending on the model, we may not show all species in this example."}
+compareB + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+
+```
+
+After comparing the survey census based on the biomass output of `run_truth`, we now look at a biomass index that is estimated based on numbers in the survey census and an average weight at age.
+This is more complex, but may be closer to the way some assessments handle survey observations.
+
+```{r surveyNbased, fig.cap="Check weight at age generated from biomass_age$atoutput divided by nums$atoutput for first 32 groups.", fig.show="hold"}
 
@@
 
+# as above, make up a constant 0 cv for testing
+surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+
+# now we need a weight at age to convert numbers to weight
+# if I do biomass_age$atoutput divided by nums$atoutput I should have wt@age, yes?
+
+nums_test <- with(datN, aggregate(atoutput, list(species, agecl), sum))
+names(nums_test) <- c("species", "agecl", "nums")
+
+bio_test  <- with(datB, aggregate(atoutput, list(species, agecl), sum))
+names(bio_test) <- c("species", "agecl", "wt")
+
+calcwtage <- merge(bio_test, nums_test) %>%
+  mutate(wtAtAge=wt/nums)
+
+wtagecheck <- ggplot(calcwtage, aes(x=agecl, y=wtAtAge)) +
+  geom_point() + 
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+wtagecheck + 
+  facet_wrap_paginate(~species, ncol=4, nrow=4, page=1, scales="free")
+
+wtagecheck + 
+  facet_wrap_paginate(~species, ncol=4, nrow=4, page=2, scales="free")
+
+# variable name needs to be wtAtAge or doesnt work
+wtage <- calcwtage %>%
+  select(species, agecl, wtAtAge)
+
+surveyB_fromN <- sample_survey_biomass(survey_testNall, surv_cv, wtage)
+
+if(initCCA) CCAsurveyB_fromN <- surveyB_fromN
+if(initNEUS) NEUSsurveyB_froN <- surveyB_fromN
+if(initNOBA) NOBAsurveyB_fromN <- surveyB_fromN
+
 ```
+
+Now we plot the results of the numbers-based (census) survey biomass against the same Atlantis output file "[modelscenario]BiomIndx.txt" as above.
+
+```{r matchBfromN, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; first 9 species." }
+
+if(initCCA) surveyB_fromN <- CCAsurveyB_fromN
+if(initNEUS) surveyB_froN <- NEUSsurveyB_fromN
+if(initNOBA) surveyB_fromN <- NOBAsurveyB_fromN
+
+#all species comparison, time intervals hardcoded for NEUS and NOBA
+compareB_N <-ggplot() +
+  geom_line(data=surveyB_fromN, aes(x=time/5,y=atoutput, color="survey census N->B"), 
+            alpha = 3/10) +
+  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
+             alpha = 3/10) + 
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+compareB_N + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+
+```
+
+```{r matchBfromN2, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}
+
+compareB_N + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+
+```
+
+```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species.Depending on the model, we may not show all species in this example."}
+
+compareB_N + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+
+```

### 2019-04-29 TrueBioTest.Rmd
live date

@@
 author: "Sarah Gaichas and Christine Stawitz"
-date: "April 25, 2019"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 Note: on Sarah's laptop, CCA took ~40 minutes to return from `run_truth` and NOBA took an hour and 26 minutes to return.
-Therefore, reading the RData file in as below is advised after the first run to make the plots comparing systems later in this document.
+Therefore, reading the .RData file in as below is advised after the first run to make the plots comparing systems later in this document.
 
@@
 
-```{r surveyBbased, eval = FALSE}
+```{r surveyBbased}
 

### 2019-04-29 TrueBioTest.Rmd
add html to docs

@@
 
-```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species."}
+```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species.", message=FALSE, warning=FALSE}
 # plot some comparisons with Atlantis output
@@
 
-```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species.
+```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
+compareB + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+
+```
+
+```{r matchBp4, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species.
 Depending on the model, we may not show all species in this example."}
 compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
 
@@
 
-```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species.Depending on the model, we may not show all species in this example."}
+```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}
 
@@
 ```
+
+```{r matchBfromN4, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species.Depending on the model, we may not show all species in this example."}
+
+compareB_N + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+
+```

### 2019-04-29 TrueBioTest.Rmd
fix error in filter, change alpha level

@@
   gather(species, biomass, -Time) %>%
-  filter(species == levels(surveyB_frombio$species))
+  filter(species %in% levels(surveyB_frombio$species))
 
@@
   geom_line(data=surveyB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
-            alpha = 3/10) +
+            alpha = 10/10) +
   geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
-             alpha = 3/10) + 
+             alpha = 1/10) + 
   theme_tufte() +

### 2019-04-30 TrueBioTest.Rmd
add refs, update html

@@
 output: html_document
+bibliography: "packages.bib"
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
+# automatically create a bib database for R packages
+knitr::write_bib(c(
+  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
+  'data.table', 'here', 'ggforce', 'ggthemes'
+), 'packages.bib')
 ```
@@
 First, you will want to set up libraries and install atlantisom if you haven't already.
+This document is written in in R Markdown [@R-rmarkdown], and we use several packages to produce the outputs [@R-tidyr; @R-dplyr; @R-ggplot2; @R-here; @R-ggforce; @R-ggthemes].
 
@@
 If you want to load the local of the atlantisom R package, use devtools.
-The R package here [@`citation(here)'] is helpful for installing from your local atlantisom directory, as it avoids hardcoding the specific location.
+The R package here is helpful for installing from your local atlantisom directory, as it avoids hardcoding the specific location.
 The code below is not running from our local atlantisom directory and is not evaluated:
@@
 
-Our directory structure is set up to take advantage of `here()' to allow setup on a different computer.
+Our directory structure is set up to take advantage of `here()` to allow setup on a different computer.
 
@@
 Now the R object `results` with the comprehensive results from the Atlantis model has been read in.
-It is also saved as an .RData file titled "output[scenario.name]run_truth.RData" in the directory with the model output, so that later analyses can use 'load()' instead of taking the time to rerun `atlantisom::run_truth()'.
+It is also saved as an .RData file titled "output[scenario.name]run_truth.RData" in the directory with the model output, so that later analyses can use `base::load()` instead of taking the time to rerun `atlantisom::run_truth()`.
 
@@
 
-To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficieny ("q"), and the selectivity at age for each species.
+To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
 
@@
 
-```{r surveyNbased, fig.cap="Check weight at age generated from biomass_age$atoutput divided by nums$atoutput for first 32 groups.", fig.show="hold"}
+```{r surveyNbased, fig.cap="Check weight at age generated from biomass_age$atoutput divided by nums$atoutput for up to 48 groups.", fig.show="hold"}
 
@@
 
+wtagecheck + 
+  facet_wrap_paginate(~species, ncol=4, nrow=4, page=3, scales="free")
+
 # variable name needs to be wtAtAge or doesnt work
@@
   geom_line(data=surveyB_fromN, aes(x=time/5,y=atoutput, color="survey census N->B"), 
-            alpha = 3/10) +
+            alpha = 10/10) +
   geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
-             alpha = 3/10) + 
+             alpha = 1/10) + 
   theme_tufte() +
@@
 ```
+
+We have more misses in the numbers with average weight estimation, so this needs more investigation and discussion.
+It may be best to have the survey index for testing derived directly from the true biomass output.
+
+## References

### 2019-04-30 StandardSurveyTest.Rmd
initial standard survey

--- 2019-04-30 TrueBioTest.Rmd
@@
 First, you will want to set up libraries and install atlantisom if you haven't already.
+This document assumes atlantisom is already installed.
+Some code blocks are hidden here.
+For complete setup and initialization, please see https://sgaichas.github.io/poseidon-dev/TrueBioTest.html.
+
 This document is written in in R Markdown [@R-rmarkdown], and we use several packages to produce the outputs [@R-tidyr; @R-dplyr; @R-ggplot2; @R-here; @R-ggforce; @R-ggthemes].
@@
 library(ggthemes)
-```
-
-If you want to load the local of the atlantisom R package, use devtools.
-The R package here is helpful for installing from your local atlantisom directory, as it avoids hardcoding the specific location.
-The code below is not running from our local atlantisom directory and is not evaluated:
-
-```{r eval=FALSE, message=FALSE, warning=FALSE}
-require(devtools)
-package.dir <- here()
-devtools::load_all(package.dir)
-```
-
-Or you can install directly from the Github repository.
-
-```{r eval=FALSE, message=FALSE, warning=FALSE}
-devtools::install_github("r4atlantis\atlantisom")
-```
-
-Next, load the package.
-
-```{r message=FALSE, warning=FALSE}
 library(atlantisom)
@@
 
-## Initializing input files and directories
+## Initialize input files and directories, read in "truth"
 
-You will first need to tell `atlantisom` to know where to look for the output and input files from your atlantis model run.
-Here, we will give the directory where the Atlantis inputs and outputs are stored `d.name`, the location of the functional groups file `functional.group.file` (.csv), the biomass pools file `biomass.pools.file` (.nc), the box locations file `box.file` (.bgm), an `initial.conditions.file` (.nc), the biology .prm file `biol.prm.file` (.prm), and the run .prm file `run.prm.file` (.prm).
-You will also need to specify a scenario name, which will be used to define the output files (i.e.
-output is stored in a number of netCDF files of the format: output<scenario><value>.nc).
-All of these files should be stored in `d.name`.
-
-In general, Atlantis model output files that are sufficiently detailed to mimic fishery sampling in space and time are too large to store on GitHub, so are not included as examples with the atlantisom code.
-
-In this example, we have a local folder "atlantisoutput" with three subdirectories:
-
-  * "CalCurrentSummitScenario1": California Current Atlantis model output
-  * "NEUStest20160303": older (broken) Northeast US Atlantis model output
-  * "NOBACERESGlobalSustainability": Norwegian-Barents Sea Atlantis model output
-
-Our directory structure is set up to take advantage of `here()` to allow setup on a different computer.
+Abbreviated here; please see https://sgaichas.github.io/poseidon-dev/TrueBioTest.html.
+This document assumes that `atlantisom::run_truth` has already completed and stored an .RData file in the atlantis output model directory.
 
@@
 
-## Getting the "true" operating model values
-
-There are a number of functions in the package that begin with the prefix `load` that load various files.
-See documentation if you'd only like to load one file.
-The `atlantisom::run_truth()` function uses the above file definitions and calls a number of the `load` functions to read in all of the atlantis output.
-Note: this call reads in a number of large .nc files, so it will take a few minutes to return.
-
 ```{r get_names, message=FALSE, warning=FALSE}
@@
 ```
-
-```{r get_truth, message=FALSE, warning=FALSE, eval=FALSE}
-#Store all loaded results into an R object
-results <- run_truth(scenario = scenario.name,
-          dir = d.name,
-          file_fgs = functional.groups.file,
-          file_bgm = box.file,
-          select_groups = funct.group.names,
-          file_init = initial.conditions.file,
-          file_biolprm = biol.prm.file,
-          file_runprm = run.prm.file
-)
-
-if(initCCA) CCAresults <- results
-if(initNEUS) NEUSresults <- results
-if(initNOBA) NOBAresults <- results
-
-```
-
-Now the R object `results` with the comprehensive results from the Atlantis model has been read in.
-It is also saved as an .RData file titled "output[scenario.name]run_truth.RData" in the directory with the model output, so that later analyses can use `base::load()` instead of taking the time to rerun `atlantisom::run_truth()`.
-
-Note: on Sarah's laptop, CCA took ~40 minutes to return from `run_truth` and NOBA took an hour and 26 minutes to return.
-Therefore, reading the .RData file in as below is advised after the first run to make the plots comparing systems later in this document.
 
@@
 
-## Simulate a survey part 1: census to compare with truth
+## Simulate a survey part 2: create a standard survey
 
-This section tests the `atlantisom::create_survey()` and `atlantisom::sample_survey_biomass()` functions by comparing a census (survey sampling everything) with the results generated by `atlantisom::run_truth()` above.
+This section uses the `atlantisom::create_survey()` and `atlantisom::sample_survey_biomass()` functions by creating a standard survey covering discrete areas, times, and species, and with both incomplete efficiency and some kind of length selectivity.
+We will apply this standard survey and compare it with true output as an example of the type of information that can be generated for assessment model performance testing.
 
@@
 
-The following settings should achieve a survey that samples all Atlantis model output timesteps, all species, and all model polygons, with perfect efficiency and full selectivity for all ages:
+One approach to creating a standard survey is to map species into general groups that are likely to have similar efficency and selectivity.
+For example, we expect that a bottom trawl survey will have generally lower efficiency for pelagic species than for demersals.
+We can also set efficiency to a very small number for infrequently encountered species occupying habitats that are not suited to bottom trawl sampling.
+This can simulate survey sampling for "data poor" species.
+Similarly, a common selectivity at age (or in this case, Atlantis cohort) can be defined for each species group.
 
-```{r census-spec, message=FALSE, warning=FALSE}
+First, we map species into general groups for each model:
 
-# should return a perfectly scaled survey 
-effic1 <- data.frame(species=funct.group.names,
-                     efficiency=rep(1.0,length(funct.group.names)))
+```{r sppgroups}
 
-# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
-# BUT CHECK if newer Atlantis models can do age-specific outputs
-selex1 <- data.frame(species=rep(funct.group.names, each=10),
-                     agecl=rep(c(1:10),length(funct.group.names)),
-                     selex=rep(1.0,length(funct.group.names)*10))
+# make defaults that return a standard survey, implement in standard_survey
+# users need to map their species groups into these general ones
+#   large pelagics/reef associated/burrowers/otherwise non-trawlable
+#   pelagics
+#   demersals
+#   selected flatfish
 
+if(initCCA) { #REPLACE WITH CCA GROUPS
+  nontrawl <- c("Pisciv_T_Fish", "Shark_D", "Shark_P", "Reptile", "Mesopel_M_Fish")
+  pelagics <- c("Planktiv_L_Fish", "Planktiv_S_Fish", "Benthopel_Fish", "Pisciv_S_Fish")
+  demersals <- c("Pisciv_D_Fish", "Demersal_D_Fish","Demersal_E_Fish", "Demersal_S_Fish",
+                 "Demersal_B_Fish", "Demersal_DC_Fish", "Demersal_O_Fish",  "Demersal_F_Fish",
+                 "Shark_B", "SkateRay")
+  selflats <- c("Pisciv_B_Fish")
+
+}
+
+if(initNEUS) {
+  nontrawl <- c("Pisciv_T_Fish", "Shark_D", "Shark_P", "Reptile", "Mesopel_M_Fish")
+  pelagics <- c("Planktiv_L_Fish", "Planktiv_S_Fish", "Benthopel_Fish", "Pisciv_S_Fish")
+  demersals <- c("Pisciv_D_Fish", "Demersal_D_Fish","Demersal_E_Fish", "Demersal_S_Fish",
+                 "Demersal_B_Fish", "Demersal_DC_Fish", "Demersal_O_Fish",  "Demersal_F_Fish",
+                 "Shark_B", "SkateRay")
+  selflats <- c("Pisciv_B_Fish")
+}
+
+if(initNOBA) { #REPLACE WITH NOBA GROUPS
+  nontrawl <- c("Pisciv_T_Fish", "Shark_D", "Shark_P", "Reptile", "Mesopel_M_Fish")
+  pelagics <- c("Planktiv_L_Fish", "Planktiv_S_Fish", "Benthopel_Fish", "Pisciv_S_Fish")
+  demersals <- c("Pisciv_D_Fish", "Demersal_D_Fish","Demersal_E_Fish", "Demersal_S_Fish",
+                 "Demersal_B_Fish", "Demersal_DC_Fish", "Demersal_O_Fish",  "Demersal_F_Fish",
+                 "Shark_B", "SkateRay")
+  selflats <- c("Pisciv_B_Fish")
+
+}
+
+```
+
+We use the following specifications for our default standard bottom trawl survey, including survey cv by species group:
+
+```{r stdbtsurvey-spec, message=FALSE, warning=FALSE}
+
+# general specifications for bottom trawl survey:
+#   once per year at mid year
+#   could generalize from the run.prm file: tstop/365 is n years of run
+#   max result$biomass_ages$time/ n years of run is steps per year
+#   take midpoint of 0, steps per year to start seq and go to max time by steps per year
+#   hardcoded for NEUS 1.0 in trunk here
+annualmidyear <- seq(3,251,5)
+
+#other options for NEUS
+annualspring <- seq(2,251,5)
+annualfall <- seq(4,251,5)
+
+#   all fish (and sharks!! need dogfish in NEUS)
+survspp <- result$fgs$Name[result$fgs$IsTurnedOn==1 &
+                           result$fgs$GroupType %in% c("FISH", "SHARK")]
+
+#   ~75% of boxes (leave off deeper boxes?)
+#   cant think of a way to generalize across models, must be hard coded
+#   hardcoded for NEUS 1.0 in trunk here
+boxsurv <- c(1:21)
+
+#   define bottom trawl mixed efficiency
+ef.nt <- 0.01 # for large pelagics, reef dwellers, others not in trawlable habitat
+ef.pl <- 0.1  # for pelagics
+ef.dm <- 0.7  # for demersals
+ef.fl <- 1.1  # for selected flatfish
+
+# bottom trawl survey efficiency specification by species group
+effnontrawl <- data.frame(species=nontrawl, efficiency=rep(ef.nt,length(nontrawl)))
+effpelagics <- data.frame(species=pelagics, efficiency=rep(ef.pl,length(pelagics)))
+effdemersals <- data.frame(species=demersals, efficiency=rep(ef.dm,length(demersals)))
+effselflats <- data.frame(species=selflats, efficiency=rep(ef.fl,length(selflats)))
+
+efficmix <- bind_rows(effnontrawl, effpelagics, effdemersals, effselflats)
+
+#   mixed selectivity (using 10 agecl for all species)
+#     flat=1 for large pelagics, reef dwellers, others not in trawlable habitat
+#     sigmoid 0 to 1 with 0.5 inflection at agecl 3 for pelagics, reaching 1 at agecl 5, flat top
+#     sigmoid 0 to 1 with 0.5 inflection at agecl 5 for most demersals and flatfish, reaching 1 at agecl 7, flat top
+#     dome shaped 0 to 1 at agecl 6&7 for selected demersals, falling off to 0.7 by agecl 10
+
+sigmoid <- function(a,b,x) {
+  1 / (1 + exp(-a-b*x))
+}
+
+# survey selectivity specification by species group
+selnontrawl <- data.frame(species=rep(nontrawl, each=10),
+                          agecl=rep(c(1:10),length(nontrawl)),
+                          selex=rep(1.0,length(nontrawl)*10))
+selpelagics <- data.frame(species=rep(pelagics, each=10),
+                          agecl=rep(c(1:10),length(pelagics)),
+                          selex=sigmoid(5,1,seq(-10,10,length.out=10)))
+seldemersals <- data.frame(species=rep(demersals, each=10),
+                          agecl=rep(c(1:10),length(demersals)),
+                          selex=sigmoid(1,1,seq(-10,10,length.out=10)))
+selselflats <- data.frame(species=rep(selflats, each=10),
+                          agecl=rep(c(1:10),length(selflats)),
+                          selex=sigmoid(1,1,seq(-10,10,length.out=10)))
+
+# same selectivity for selflats and demersals for NEUS
+# visualze selectivity curves for each group
+par(mfrow=c(2,2))
+par(mar=c(4,4,1,1))
+plot(selnontrawl$agecl, selnontrawl$selex)
+plot(selpelagics$agecl, selpelagics$selex)
+plot(seldemersals$agecl, seldemersals$selex)
+plot(selselflats$agecl, selselflats$selex)
+par(mfrow=c(1,1))
+
+# implement dome shaped selectivity for a particular species and replace default
+# not done yet staying simple for now
+
+selexmix <- bind_rows(selnontrawl, selpelagics, seldemersals, selselflats)
+
+# use this constant 0 cv for testing
+surv_cv_0 <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+
+# specify cv by species groups
+surv_cv_nontrawl <- data.frame(species=nontrawl, cv=rep(1.0,length(nontrawl)))
+surv_cv_pelagics <- data.frame(species=pelagics, cv=rep(0.5,length(pelagics)))
+surv_cv_demersals <- data.frame(species=demersals, cv=rep(0.3,length(demersals)))
+surv_cv_selflats <- data.frame(species=selflats, cv=rep(0.3,length(selflats)))
+
+surv_cv_mix <- bind_rows(surv_cv_nontrawl, surv_cv_pelagics, surv_cv_demersals, surv_cv_selflats)
+
+################# Below here is old from census, delete ################
 # should return all model areas
@@
 
-Because the results of `run_truth` provide both biomass at age and numbers at age, we can use `create_survey` on both with some modifications.
-The following uses the biomass output of `run_truth` to create the survey, so the call to `sample_survey_biomass` will require a weight at age argument that is filled with 1's because no conversion from numbers to weight is necessary.
-Because we are making a census for testing, the survey cv argument is set to 0.
+Here we use `create_survey` on the biomass output of `run_truth` to create the survey, so the call to `sample_survey_biomass` will require a weight at age argument that is filled with 1's because no conversion from numbers to weight is necessary.
 

### 2019-04-30 TrueBioTest.Rmd
save survey census file

@@
 
+#save for later use, takes a long time to generate
+saveRDS(surveyB_frombio, file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
+
 if(initCCA) CCAsurveyB_frombio <- surveyB_frombio
@@
 surveyB_fromN <- sample_survey_biomass(survey_testNall, surv_cv, wtage)
+#save for later use, takes a long time to generate
+saveRDS(surveyB_fromN, file.path(d.name, paste0(scenario.name, "surveyBfromNcensus.rds")))
 

### 2019-04-30 StandardSurveyTest.Rmd
bottom trawl example building

@@
 ---
-title: "Testing atlantisom: true biomass output from Atlantis"
+title: "Testing atlantisom: example bottom trawl survey index"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 This document assumes atlantisom is already installed.
-Some code blocks are hidden here.
-For complete setup and initialization, please see https://sgaichas.github.io/poseidon-dev/TrueBioTest.html.
+For complete setup and initialization, please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).
 
@@
 
-Abbreviated here; please see https://sgaichas.github.io/poseidon-dev/TrueBioTest.html.
+Abbreviated here; for a full explanation please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).
 This document assumes that `atlantisom::run_truth` has already completed and stored an .RData file in the atlantis output model directory.
@@
 
-```{r surveyBbased}
+```{r stdsurveyBbased}
 
@@
 
-survey_testBall <- create_survey(dat = datB,
-                                 time = timeall,
+survey_testBstd <- create_survey(dat = datB,
+                                 time = annualmidyear,
                                  species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-# make up a constant 0 cv for testing
-surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+                                 boxes = boxsurv,
+                                 effic = efficmix,
+                                 selex = selexmix)
 
@@
 
-surveyB_frombio <- sample_survey_biomass(survey_testBall, surv_cv, wtage)
+stdsurveyB_frombio <- sample_survey_biomass(survey_testBstd, surv_cv_mix, wtage)
 
-if(initCCA) CCAsurveyB_frombio <- surveyB_frombio
-if(initNEUS) NEUSsurveyB_frombio <- surveyB_frombio
-if(initNOBA) NOBAsurveyB_frombio <- surveyB_frombio
+if(initCCA) CCAstdsurveyB_frombio <- stdsurveyB_frombio
+if(initNEUS) NEUSstdsurveyB_frombio <- stdsurveyB_frombio
+if(initNOBA) NOBAstdsurveyB_frombio <- stdsurveyB_frombio
 
@@
              alpha = 1/10) + 
+  #geom_line(data=surveyB_frombio_eff, aes(x=time/5,y=atoutput, color="stdeffB")) +
+  #geom_line(data=surveyB_frombio_effsel, aes(x=time/5,y=atoutput, color="stdeffselB")) +
+  geom_point(data=surveyB_frombio_effselcv, aes(x=time/5,y=atoutput, color="stdeffcvselB"),
+              alpha = 1/2) +
+
   theme_tufte() +
@@
 
-After comparing the survey census based on the biomass output of `run_truth`, we now look at a biomass index that is estimated based on numbers in the survey census and an average weight at age.
-This is more complex, but may be closer to the way some assessments handle survey observations.
-
-```{r surveyNbased, fig.cap="Check weight at age generated from biomass_age$atoutput divided by nums$atoutput for up to 48 groups.", fig.show="hold"}
-
-# this uses result$nums so will need a weight at age to get biomass
-
-if(initCCA) datN <- CCAresults$nums
-if(initNEUS) datN <- NEUSresults$nums
-if(initNOBA) datN <- NOBAresults$nums
-
-survey_testNall <- create_survey(dat = datN,
-                                 time = timeall,
-                                 species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-# as above, make up a constant 0 cv for testing
-surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
-
-# now we need a weight at age to convert numbers to weight
-# if I do biomass_age$atoutput divided by nums$atoutput I should have wt@age, yes?
-
-nums_test <- with(datN, aggregate(atoutput, list(species, agecl), sum))
-names(nums_test) <- c("species", "agecl", "nums")
-
-bio_test  <- with(datB, aggregate(atoutput, list(species, agecl), sum))
-names(bio_test) <- c("species", "agecl", "wt")
-
-calcwtage <- merge(bio_test, nums_test) %>%
-  mutate(wtAtAge=wt/nums)
-
-wtagecheck <- ggplot(calcwtage, aes(x=agecl, y=wtAtAge)) +
-  geom_point() + 
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-wtagecheck + 
-  facet_wrap_paginate(~species, ncol=4, nrow=4, page=1, scales="free")
-
-wtagecheck + 
-  facet_wrap_paginate(~species, ncol=4, nrow=4, page=2, scales="free")
-
-wtagecheck + 
-  facet_wrap_paginate(~species, ncol=4, nrow=4, page=3, scales="free")
-
-# variable name needs to be wtAtAge or doesnt work
-wtage <- calcwtage %>%
-  select(species, agecl, wtAtAge)
-
-surveyB_fromN <- sample_survey_biomass(survey_testNall, surv_cv, wtage)
-
-if(initCCA) CCAsurveyB_fromN <- surveyB_fromN
-if(initNEUS) NEUSsurveyB_froN <- surveyB_fromN
-if(initNOBA) NOBAsurveyB_fromN <- surveyB_fromN
-
-```
-
-Now we plot the results of the numbers-based (census) survey biomass against the same Atlantis output file "[modelscenario]BiomIndx.txt" as above.
-
-```{r matchBfromN, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; first 9 species." }
-
-if(initCCA) surveyB_fromN <- CCAsurveyB_fromN
-if(initNEUS) surveyB_froN <- NEUSsurveyB_fromN
-if(initNOBA) surveyB_fromN <- NOBAsurveyB_fromN
-
-#all species comparison, time intervals hardcoded for NEUS and NOBA
-compareB_N <-ggplot() +
-  geom_line(data=surveyB_fromN, aes(x=time/5,y=atoutput, color="survey census N->B"), 
-            alpha = 10/10) +
-  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
-             alpha = 1/10) + 
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-compareB_N + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
-
-```
-
-```{r matchBfromN2, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}
-
-compareB_N + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
-
-```
-
-```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}
-
-compareB_N + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
-
-```
-
-```{r matchBfromN4, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species.Depending on the model, we may not show all species in this example."}
-
-compareB_N + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
-
-```
-
-We have more misses in the numbers with average weight estimation, so this needs more investigation and discussion.
-It may be best to have the survey index for testing derived directly from the true biomass output.
+Given the misses using survey number outputs from run truth with a calculated weight at age, we will not repeat this example using numbers.
 

### 2019-04-30 CCAExamples.Rmd
example visualizations each model

--- 2019-04-30 TrueBioTest.Rmd
@@
 ---
-title: "Testing atlantisom: true biomass output from Atlantis"
+title: "Testing atlantisom: true biomass output from California Current Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+
 # automatically create a bib database for R packages
@@
 ), 'packages.bib')
-```
 
-## Introduction
-
-This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets.
-Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
-
-The purpose of atlantisom is to use existing Atlantis model output to generate input datasets for a variety of models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics.
-Atlantis models can be run using different climate forcing, fishing, and other scenarios.
-Users of atlantisom will be able to specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes for any Atlantis scenario.
-Internally consistent multispecies and ecosystem datasets with known observation error characteristics will be the atlantisom outputs, for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against "true" Atlantis outputs.
-
-Initial testing was conducted by S.
-Gaichas using R scripts in the R folder of this repository that are titled "PoseidonTest_[whatwastested].R".
-Initial tests are expanded and documented in more detail in these pages.
-C.
-Stawitz improved and streamlined the setup and intialization sections.
-
-## Setup
-
-First, you will want to set up libraries and install atlantisom if you haven't already.
-This document is written in in R Markdown [@R-rmarkdown], and we use several packages to produce the outputs [@R-tidyr; @R-dplyr; @R-ggplot2; @R-here; @R-ggforce; @R-ggthemes].
-
-```{r message=FALSE, warning=FALSE}
 library(tidyr)
@@
 library(ggthemes)
+library(atlantisom)
+
 ```
 
-If you want to load the local of the atlantisom R package, use devtools.
-The R package here is helpful for installing from your local atlantisom directory, as it avoids hardcoding the specific location.
-The code below is not running from our local atlantisom directory and is not evaluated:
+## Introduction
 
-```{r eval=FALSE, message=FALSE, warning=FALSE}
-require(devtools)
-package.dir <- here()
-devtools::load_all(package.dir)
-```
-
-Or you can install directly from the Github repository.
-
-```{r eval=FALSE, message=FALSE, warning=FALSE}
-devtools::install_github("r4atlantis\atlantisom")
-```
-
-Next, load the package.
-
-```{r message=FALSE, warning=FALSE}
-library(atlantisom)
-```
-
-## Initializing input files and directories
-
-You will first need to tell `atlantisom` to know where to look for the output and input files from your atlantis model run.
-Here, we will give the directory where the Atlantis inputs and outputs are stored `d.name`, the location of the functional groups file `functional.group.file` (.csv), the biomass pools file `biomass.pools.file` (.nc), the box locations file `box.file` (.bgm), an `initial.conditions.file` (.nc), the biology .prm file `biol.prm.file` (.prm), and the run .prm file `run.prm.file` (.prm).
-You will also need to specify a scenario name, which will be used to define the output files (i.e.
-output is stored in a number of netCDF files of the format: output<scenario><value>.nc).
-All of these files should be stored in `d.name`.
-
-In general, Atlantis model output files that are sufficiently detailed to mimic fishery sampling in space and time are too large to store on GitHub, so are not included as examples with the atlantisom code.
-
-In this example, we have a local folder "atlantisoutput" with three subdirectories:
-
-  * "CalCurrentSummitScenario1": California Current Atlantis model output
-  * "NEUStest20160303": older (broken) Northeast US Atlantis model output
-  * "NOBACERESGlobalSustainability": Norwegian-Barents Sea Atlantis model output
-
-Our directory structure is set up to take advantage of `here()` to allow setup on a different computer.
+This page has visualizations for the CCA model example, Atlantis Summit Common Scenario 1. For full explanation of methods, see the file linked at the beginning of each section.
 
@@
 
-initCCA <- FALSE
+initCCA <- TRUE
 initNEUS <- FALSE
-initNOBA <- TRUE
+initNOBA <- FALSE
 
@@
 
-This section tests the `atlantisom::create_survey()` and `atlantisom::sample_survey_biomass()` functions by comparing a census (survey sampling everything) with the results generated by `atlantisom::run_truth()` above.
-
-To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
-
+Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).
 The following settings should achieve a survey that samples all Atlantis model output timesteps, all species, and all model polygons, with perfect efficiency and full selectivity for all ages:
 
-```{r census-spec, message=FALSE, warning=FALSE}
+```{r census-spec, echo=TRUE}
 
@@
 
-```{r surveyBbased}
+```{r surveyBbased, eval=FALSE}
 
@@
 if(initNOBA) NOBAsurveyB_frombio <- surveyB_frombio
+
+```
+
+```{r readsurvey1}
+
+if(initCCA) {
+  CCAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
+}
+
+if(initNEUS) {
+  NEUSsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
+}
+
+if(initNOBA) {
+  NOBAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
+}
 
@@
 #all species comparison, time intervals hardcoded for NEUS and NOBA
+# WHOOPS! This is CCA, really need to generalize this but hardcoded for now
 compareB <-ggplot() +
-  geom_line(data=surveyB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
+  geom_line(data=surveyB_frombio, aes(x=time,y=atoutput, color="survey census B"), 
             alpha = 10/10) +
@@
 
-```{r matchBp4, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species.
-Depending on the model, we may not show all species in this example."}
+```{r matchBp4, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
 compareB + 
   facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
+```{r matchBp5, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species."}
+compareB + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
 
@@
 
-```{r surveyNbased, fig.cap="Check weight at age generated from biomass_age$atoutput divided by nums$atoutput for up to 48 groups.", fig.show="hold"}
-
-# this uses result$nums so will need a weight at age to get biomass
+```{r wtagecalc, fig.cap="Check weight at age generated from biomass_age\\$atoutput divided by nums\\$atoutput for up to 64 groups.", fig.show="hold"}
 
@@
 
-survey_testNall <- create_survey(dat = datN,
-                                 time = timeall,
-                                 species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-# as above, make up a constant 0 cv for testing
-surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+if(initCCA) datB <- CCAresults$biomass_ages
+if(initNEUS) datB <- NEUSresults$biomass_ages
+if(initNOBA) datB <- NOBAresults$biomass_ages
 
@@
 
+wtagecheck + 
+  facet_wrap_paginate(~species, ncol=4, nrow=4, page=4, scales="free")
+
+```
+
+```{r surveyNbased, eval=FALSE}
+
+# this uses result$nums so will need a weight at age to get biomass
+
+if(initCCA) datN <- CCAresults$nums
+if(initNEUS) datN <- NEUSresults$nums
+if(initNOBA) datN <- NOBAresults$nums
+
+survey_testNall <- create_survey(dat = datN,
+                                 time = timeall,
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
+# as above, make up a constant 0 cv for testing
+surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+
 # variable name needs to be wtAtAge or doesnt work
@@
 if(initCCA) CCAsurveyB_fromN <- surveyB_fromN
-if(initNEUS) NEUSsurveyB_froN <- surveyB_fromN
+if(initNEUS) NEUSsurveyB_fromN <- surveyB_fromN
 if(initNOBA) NOBAsurveyB_fromN <- surveyB_fromN
+
+```
+
+```{r readsurvey2}
+
+if(initCCA) {
+  CCAsurveyB_fromN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBfromNcensus.rds")))
+}
+
+if(initNEUS) {
+  NEUSsurveyB_fromN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBfromNcensus.rds")))
+}
+
+if(initNOBA) {
+  NOBAsurveyB_fromN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBfromNcensus.rds")))
+}
 
@@
 if(initCCA) surveyB_fromN <- CCAsurveyB_fromN
-if(initNEUS) surveyB_froN <- NEUSsurveyB_fromN
+if(initNEUS) surveyB_fromN <- NEUSsurveyB_fromN
 if(initNOBA) surveyB_fromN <- NOBAsurveyB_fromN
@@
 #all species comparison, time intervals hardcoded for NEUS and NOBA
+# WHOOPS! Fixed for CCA but generalize this!
 compareB_N <-ggplot() +
-  geom_line(data=surveyB_fromN, aes(x=time/5,y=atoutput, color="survey census N->B"), 
+  geom_line(data=surveyB_fromN, aes(x=time,y=atoutput, color="survey census N->B"), 
             alpha = 10/10) +
@@
 
-```{r matchBfromN4, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species.Depending on the model, we may not show all species in this example."}
+```{r matchBfromN4, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}
 
@@
 
+```{r matchBfromN5, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species. "}
+
+compareB_N + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
+
+```
+
 We have more misses in the numbers with average weight estimation, so this needs more investigation and discussion.
 It may be best to have the survey index for testing derived directly from the true biomass output.
-
-## References

### 2019-04-30 NEUSExamples.Rmd
example visualizations each model

--- 2019-04-30 CCAExamples.Rmd
@@
 ---
-title: "Testing atlantisom: true biomass output from California Current Atlantis"
+title: "Testing atlantisom: true biomass output from Northeast US Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-This page has visualizations for the CCA model example, Atlantis Summit Common Scenario 1. For full explanation of methods, see the file linked at the beginning of each section.
+This page has visualizations for the NEUS model example, test03032016 (note! a not fully calibrated model). For full explanation of methods, see the file linked at the beginning of each section.
 
@@
 #all species comparison, time intervals hardcoded for NEUS and NOBA
-# WHOOPS! This is CCA, really need to generalize this but hardcoded for now
 compareB <-ggplot() +
-  geom_line(data=surveyB_frombio, aes(x=time,y=atoutput, color="survey census B"), 
+  geom_line(data=surveyB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
             alpha = 10/10) +
@@
 
-```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
+```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species."}
 compareB + 
@@
 
-```{r matchBp4, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
-
-```
-
-```{r matchBp5, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species."}
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
-
-```
-
 After comparing the survey census based on the biomass output of `run_truth`, we now look at a biomass index that is estimated based on numbers in the survey census and an average weight at age.
 This is more complex, but may be closer to the way some assessments handle survey observations.
-
 ```{r wtagecalc, fig.cap="Check weight at age generated from biomass_age\\$atoutput divided by nums\\$atoutput for up to 64 groups.", fig.show="hold"}
@@
   facet_wrap_paginate(~species, ncol=4, nrow=4, page=2, scales="free")
-
-wtagecheck + 
-  facet_wrap_paginate(~species, ncol=4, nrow=4, page=3, scales="free")
-
-wtagecheck + 
-  facet_wrap_paginate(~species, ncol=4, nrow=4, page=4, scales="free")
 
@@
 #all species comparison, time intervals hardcoded for NEUS and NOBA
-# WHOOPS! Fixed for CCA but generalize this!
 compareB_N <-ggplot() +
-  geom_line(data=surveyB_fromN, aes(x=time,y=atoutput, color="survey census N->B"), 
+  geom_line(data=surveyB_fromN, aes(x=time/5,y=atoutput, color="survey census N->B"), 
             alpha = 10/10) +
@@
 
-```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}
+```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species."}
 
@@
 
-```{r matchBfromN4, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}
-
-compareB_N + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
-
-```
-
-```{r matchBfromN5, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species. "}
-
-compareB_N + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
-
-```
-
 We have more misses in the numbers with average weight estimation, so this needs more investigation and discussion.

### 2019-04-30 NOBAExamples.Rmd
example visualizations each model

--- 2019-04-30 NEUSExamples.Rmd
@@
 ---
-title: "Testing atlantisom: true biomass output from Northeast US Atlantis"
+title: "Testing atlantisom: true biomass output from Norwegian-Barents Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-This page has visualizations for the NEUS model example, test03032016 (note! a not fully calibrated model). For full explanation of methods, see the file linked at the beginning of each section.
+This page has visualizations for the NOBA model example, CERES Global Sustainability. For full explanation of methods, see the file linked at the beginning of each section.
 
@@
 
-```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species."}
+```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
 compareB + 
@@
 
+```{r matchBp4, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species."}
+compareB + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
 After comparing the survey census based on the biomass output of `run_truth`, we now look at a biomass index that is estimated based on numbers in the survey census and an average weight at age.
 This is more complex, but may be closer to the way some assessments handle survey observations.
-```{r wtagecalc, fig.cap="Check weight at age generated from biomass_age\\$atoutput divided by nums\\$atoutput for up to 64 groups.", fig.show="hold"}
+
+```{r wtagecalc, fig.cap="Check weight at age generated from biomass_age\\$atoutput divided by nums\\$atoutput for up to 48 groups.", fig.show="hold"}
 
@@
   facet_wrap_paginate(~species, ncol=4, nrow=4, page=2, scales="free")
+
+wtagecheck + 
+  facet_wrap_paginate(~species, ncol=4, nrow=4, page=3, scales="free")
 
@@
 
-```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species."}
+```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}
 
@@
 
+```{r matchBfromN4, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species. "}
+
+compareB_N + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
 We have more misses in the numbers with average weight estimation, so this needs more investigation and discussion.

### 2019-04-30 docs/index.Rmd
fix repo name
add html to docs
date update
update index

@@
 
-<a href="https://sgaichas.github.io/presentations/TrueBioTest.html" target="_blank">Installation and initial tests - April 26, 2019</a>
+<a href="https://sgaichas.github.io/poseidon-dev/TrueBioTest.html" target="_blank">Installation and initial tests - April 26, 2019, updated April 30, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NEUSExamples.html" target="_blank">Survey census visualization for NEUS -  April 30, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/CCAExamples.html" target="_blank">Survey census visualization for CCA - April 30, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NOBAExamples.html" target="_blank">Survey census visualization for NOBA - April 30, 2019</a>
 

### 2019-04-30 StandardSurveyTest.Rmd
standard survey NEUS example

@@
 initCCA <- FALSE
-initNEUS <- FALSE
-initNOBA <- TRUE
+initNEUS <- TRUE
+initNOBA <- FALSE
 
@@
 
+We also read in previously generated survey census files based on true biomass results for comparison (see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).)
+
+```{r readsurvey1}
+
+if(initCCA) {
+  CCAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
+}
+
+if(initNEUS) {
+  NEUSsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
+}
+
+if(initNOBA) {
+  NOBAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
+}
+
+```
+
 ## Simulate a survey part 2: create a standard survey
@@
 #   once per year at mid year
-#   could generalize from the run.prm file: tstop/365 is n years of run
-#   max result$biomass_ages$time/ n years of run is steps per year
+#   could generalize from the run.prm file: 
+runpar <- load_runprm(d.name, run.prm.file)
+noutsteps <- runpar$tstop/runpar$outputstep
+stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
 #   take midpoint of 0, steps per year to start seq and go to max time by steps per year
+midptyr <- round(median(seq(1,stepperyr)))
+
+annualmidyear <- seq(midptyr, noutsteps, stepperyr)
+
 #   hardcoded for NEUS 1.0 in trunk here
-annualmidyear <- seq(3,251,5)
-
-#other options for NEUS
-annualspring <- seq(2,251,5)
-annualfall <- seq(4,251,5)
+# annualmidyear <- seq(3,251,5)
+# #other options for NEUS
+# annualspring <- seq(2,251,5)
+# annualfall <- seq(4,251,5)
 
 #   all fish (and sharks!! need dogfish in NEUS)
-survspp <- result$fgs$Name[result$fgs$IsTurnedOn==1 &
-                           result$fgs$GroupType %in% c("FISH", "SHARK")]
+# could add groups "CEP" for cephalopod, "PWN" for prawn, etc
+survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
+                           funct.groups$GroupType %in% c("FISH", "SHARK")]
 
@@
 
-################# Below here is old from census, delete ################
-# should return all model areas
-boxpars <- load_box(d.name, box.file)
-boxall <- c(0:(boxpars$nbox - 1))
-
-# these are model specific, generalized above
-# if(initCCA) boxall <- c(0:88) 
-# if(initNEUS) boxall <- c(0:29)
-# if(initNOBA) boxall <- c(0:59)
-
-# should return all model output timesteps; need to generalize
-if(initCCA) timeall <- c(0:100) 
-if(initNEUS) timeall <- c(0:251)
-if(initNOBA) timeall <- c(0:560)
-
-# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
-# for ecosystem indicator work test all species, e.g.
-survspp <- funct.group.names
-
-# to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
-if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
-if(initNEUS) survspp <- funct.group.names[1:21]
-if(initNOBA) survspp <- funct.group.names[1:36]
+# ################# Below here is old from census for reference, delete ################
+# # should return all model areas
+# boxpars <- load_box(d.name, box.file)
+# boxall <- c(0:(boxpars$nbox - 1))
+# 
+# # these are model specific, generalized above
+# # if(initCCA) boxall <- c(0:88) 
+# # if(initNEUS) boxall <- c(0:29)
+# # if(initNOBA) boxall <- c(0:59) 
+# 
+# # should return all model output timesteps; need to generalize
+# if(initCCA) timeall <- c(0:100) 
+# if(initNEUS) timeall <- c(0:251)
+# if(initNOBA) timeall <- c(0:560) 
+#   
+# # define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
+# # for ecosystem indicator work test all species, e.g.
+# survspp <- funct.group.names 
+# 
+# # to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
+# if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
+# if(initNEUS) survspp <- funct.group.names[1:21]
+# if(initNOBA) survspp <- funct.group.names[1:36]
 
@@
   groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  surveyB_frombio <- CCAsurveyB_frombio
+  stdsurveyB_frombio <- CCAstdsurveyB_frombio
+  censusB_frombio <- CCAsurveyB_frombio
 }
@@
   groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  surveyB_frombio <- NEUSsurveyB_frombio
+  stdsurveyB_frombio <- NEUSstdsurveyB_frombio
+  censusB_frombio <- NEUSsurveyB_frombio
 }
@@
   groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  surveyB_frombio <- NOBAsurveyB_frombio
+  stdsurveyB_frombio <- NOBAstdsurveyB_frombio
+  censusB_frombio <- NOBAsurveyB_frombio
 }
@@
   gather(species, biomass, -Time) %>%
-  filter(species %in% levels(surveyB_frombio$species))
+  filter(species %in% levels(stdsurveyB_frombio$species))
+
+censusB_frombio <- censusB_frombio %>%
+  filter(species %in% levels(stdsurveyB_frombio$species))
 
@@
 compareB <-ggplot() +
-  geom_line(data=surveyB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
-            alpha = 10/10) +
+  geom_line(data=censusB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
+            alpha = 5/10) +
   geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
-             alpha = 1/10) + 
+             alpha = 5/10) + 
   #geom_line(data=surveyB_frombio_eff, aes(x=time/5,y=atoutput, color="stdeffB")) +
   #geom_line(data=surveyB_frombio_effsel, aes(x=time/5,y=atoutput, color="stdeffselB")) +
-  geom_point(data=surveyB_frombio_effselcv, aes(x=time/5,y=atoutput, color="stdeffcvselB"),
+  geom_point(data=stdsurveyB_frombio, aes(x=time/5,y=atoutput, color="stdeffcvselB"),
               alpha = 1/2) +
@@
 Depending on the model, we may not show all species in this example."}
+if(!initNEUS){
   compareB + 
     facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free") 
-
+}
 ```

### 2019-05-01 StandardSurveyTest.Rmd
species groups for CCA and NOBA, general boxes

@@
 
-if(initCCA) { #REPLACE WITH CCA GROUPS
+if(initCCA) { #Sarah's CCA Grouping
+  nontrawl <- c("Shark_C","Yelloweye_rockfish","Benthopel_Fish","Pisciv_S_Fish",
+                "Pisciv_T_Fish","Shark_D","Shark_P")
+  pelagics <- c("Pisciv_V_Fish","Demersal_S_Fish","Pacific_Ocean_Perch","Mesopel_M_Fish",
+                "Planktiv_L_Fish","Jack_mackerel","Planktiv_S_Fish","Pacific_sardine",
+                "Anchovy","Herring","Pisciv_B_Fish")
+  demersals <- c("Demersal_P_Fish","Planktiv_O_Fish","Demersal_D_Fish",
+                 "Demersal_DC_Fish","Demersal_O_Fish","Darkblotched_rockfish",
+                 "Demersal_F_Fish","Demersal_E_Fish","Bocaccio_rockfish",
+                 "Demersal_B_Fish","Shark_R","Mesopel_N_Fish","Shark_B","Spiny_dogfish",
+                 "SkateRay")
+  selflats <- c("Pisciv_D_Fish", "Arrowtooth_flounder","Petrale_sole")
+}
+
+if(initNEUS) { # Sarah's NEUS Grouping
   nontrawl <- c("Pisciv_T_Fish", "Shark_D", "Shark_P", "Reptile", "Mesopel_M_Fish")
   pelagics <- c("Planktiv_L_Fish", "Planktiv_S_Fish", "Benthopel_Fish", "Pisciv_S_Fish")
-  demersals <- c("Pisciv_D_Fish", "Demersal_D_Fish","Demersal_E_Fish", "Demersal_S_Fish",
-                 "Demersal_B_Fish", "Demersal_DC_Fish", "Demersal_O_Fish",  "Demersal_F_Fish",
-                 "Shark_B", "SkateRay")
-  selflats <- c("Pisciv_B_Fish")
-
-}
-
-if(initNEUS) {
-  nontrawl <- c("Pisciv_T_Fish", "Shark_D", "Shark_P", "Reptile", "Mesopel_M_Fish")
-  pelagics <- c("Planktiv_L_Fish", "Planktiv_S_Fish", "Benthopel_Fish", "Pisciv_S_Fish")
-  demersals <- c("Pisciv_D_Fish", "Demersal_D_Fish","Demersal_E_Fish", "Demersal_S_Fish",
-                 "Demersal_B_Fish", "Demersal_DC_Fish", "Demersal_O_Fish",  "Demersal_F_Fish",
+  demersals <- c("Pisciv_D_Fish", "Demersal_D_Fish","Demersal_E_Fish", 
+                 "Demersal_S_Fish","Demersal_B_Fish","Demersal_DC_Fish",
+                 "Demersal_O_Fish","Demersal_F_Fish",
                  "Shark_B", "SkateRay")
@@
 
-if(initNOBA) { #REPLACE WITH NOBA GROUPS
-  nontrawl <- c("Pisciv_T_Fish", "Shark_D", "Shark_P", "Reptile", "Mesopel_M_Fish")
-  pelagics <- c("Planktiv_L_Fish", "Planktiv_S_Fish", "Benthopel_Fish", "Pisciv_S_Fish")
-  demersals <- c("Pisciv_D_Fish", "Demersal_D_Fish","Demersal_E_Fish", "Demersal_S_Fish",
-                 "Demersal_B_Fish", "Demersal_DC_Fish", "Demersal_O_Fish",  "Demersal_F_Fish",
-                 "Shark_B", "SkateRay")
-  selflats <- c("Pisciv_B_Fish")
-
+if(initNOBA) { # Sarah's NOBA Grouping
+  nontrawl <- c("Sharks_other", "Pelagic_large","Mesop_fish")
+  pelagics <- c("Pelagic_small","Redfish_other","Mackerel","Haddock",
+                "Saithe","Redfish","Blue_whiting","Norwegian_ssh","Capelin")
+  demersals <- c("Demersals_other","Demersal_large","Flatfish_other","Skates_rays",
+                 "Green_halibut","North_atl_cod","Polar_cod","Snow_crab")
+  selflats <- c("Long_rough_dab")
 }
@@
 # could add groups "CEP" for cephalopod, "PWN" for prawn, etc
+
+# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
+if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)
+
 survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
@@
 
-#   ~75% of boxes (leave off deeper boxes?)
-#   cant think of a way to generalize across models, must be hard coded
-#   hardcoded for NEUS 1.0 in trunk here
-boxsurv <- c(1:21)
+#   ~75-80% of boxes (leave off deeper boxes?)
+#   cant think of a way to generalize across models, must be hard coded in reality
+#   hardcoded for NEUS 1.0 in trunk here: boxsurv <- c(1:21)
+boxpars <- load_box(d.name, box.file)
+# should return all model areas
+# boxall <- c(0:(boxpars$nbox - 1))
+boxsurv <- c(2:round(0.8*(boxpars$nbox - 1)))
 
@@
 
-# same selectivity for selflats and demersals for NEUS
+# same selectivity for selflats and demersals specified above
 # visualze selectivity curves for each group
@@
 
+#   define bottom trawl survey cv by group
+cv.nt <- 1.0 # for large pelagics, reef dwellers, others not in trawlable habitat
+cv.pl <- 0.5  # for pelagics
+cv.dm <- 0.3  # for demersals
+cv.fl <- 0.3  # for selected flatfish
+
 # specify cv by species groups
-surv_cv_nontrawl <- data.frame(species=nontrawl, cv=rep(1.0,length(nontrawl)))
-surv_cv_pelagics <- data.frame(species=pelagics, cv=rep(0.5,length(pelagics)))
-surv_cv_demersals <- data.frame(species=demersals, cv=rep(0.3,length(demersals)))
-surv_cv_selflats <- data.frame(species=selflats, cv=rep(0.3,length(selflats)))
+surv_cv_nontrawl <- data.frame(species=nontrawl, cv=rep(cv.nt,length(nontrawl)))
+surv_cv_pelagics <- data.frame(species=pelagics, cv=rep(cv.pl,length(pelagics)))
+surv_cv_demersals <- data.frame(species=demersals, cv=rep(cv.dm,length(demersals)))
+surv_cv_selflats <- data.frame(species=selflats, cv=rep(cv.fl,length(selflats)))
 
 surv_cv_mix <- bind_rows(surv_cv_nontrawl, surv_cv_pelagics, surv_cv_demersals, surv_cv_selflats)
-
-# ################# Below here is old from census for reference, delete ################
-# # should return all model areas
-# boxpars <- load_box(d.name, box.file)
-# boxall <- c(0:(boxpars$nbox - 1))
-# 
-# # these are model specific, generalized above
-# # if(initCCA) boxall <- c(0:88) 
-# # if(initNEUS) boxall <- c(0:29)
-# # if(initNOBA) boxall <- c(0:59) 
-# 
-# # should return all model output timesteps; need to generalize
-# if(initCCA) timeall <- c(0:100) 
-# if(initNEUS) timeall <- c(0:251)
-# if(initNOBA) timeall <- c(0:560) 
-#   
-# # define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
-# # for ecosystem indicator work test all species, e.g.
-# survspp <- funct.group.names 
-# 
-# # to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
-# if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
-# if(initNEUS) survspp <- funct.group.names[1:21]
-# if(initNOBA) survspp <- funct.group.names[1:36]
 

### 2019-05-01 NOBAStandardSurveyTest.Rmd
add std survey visualizations all mods

--- 2019-05-01 StandardSurveyTest.Rmd
@@
 ---
-title: "Testing atlantisom: example bottom trawl survey index"
+title: "Testing atlantisom: example bottom trawl survey index Norwegian-Barents Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 # automatically create a bib database for R packages
@@
 
-This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets.
-Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
-
-The purpose of atlantisom is to use existing Atlantis model output to generate input datasets for a variety of models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics.
-Atlantis models can be run using different climate forcing, fishing, and other scenarios.
-Users of atlantisom will be able to specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes for any Atlantis scenario.
-Internally consistent multispecies and ecosystem datasets with known observation error characteristics will be the atlantisom outputs, for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against "true" Atlantis outputs.
-
-Initial testing was conducted by S.
-Gaichas using R scripts in the R folder of this repository that are titled "PoseidonTest_[whatwastested].R".
-Initial tests are expanded and documented in more detail in these pages.
-C.
-Stawitz improved and streamlined the setup and intialization sections.
-
-## Setup
-
-First, you will want to set up libraries and install atlantisom if you haven't already.
-This document assumes atlantisom is already installed.
-For complete setup and initialization, please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).
-
-This document is written in in R Markdown [@R-rmarkdown], and we use several packages to produce the outputs [@R-tidyr; @R-dplyr; @R-ggplot2; @R-here; @R-ggforce; @R-ggthemes].
+This page has visualizations for the NOBA model example, CERES Global Sustainability. For full explanation of methods, see the file linked at the beginning of each section.
 
@@
 
-## Initialize input files and directories, read in "truth"
-
-Abbreviated here; for a full explanation please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).
-This document assumes that `atlantisom::run_truth` has already completed and stored an .RData file in the atlantis output model directory.
-
 ```{r initialize}
@@
 
-We also read in previously generated survey census files based on true biomass results for comparison (see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).)
-
 ```{r readsurvey1}
@@
 
-This section uses the `atlantisom::create_survey()` and `atlantisom::sample_survey_biomass()` functions by creating a standard survey covering discrete areas, times, and species, and with both incomplete efficiency and some kind of length selectivity.
-We will apply this standard survey and compare it with true output as an example of the type of information that can be generated for assessment model performance testing.
-
-To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
+Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StandardSurveyTest.html).
 
@@
 
-```{r sppgroups}
+```{r sppgroups, echo=TRUE}
 
@@
 
-```{r stdbtsurvey-spec, message=FALSE, warning=FALSE}
+```{r stdbtsurvey-spec, message=FALSE, warning=FALSE, echo=TRUE}
 
@@
 
-Comparing our (census) survey based on true biomass from above with the Atlantis output file "[modelscenario]BiomIndx.txt" should give us a perfect match.
-Note that the our (census) survey may have more sampling in time than the Atlantis output file.
-
 ```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species.", message=FALSE, warning=FALSE}
@@
 Depending on the model, we may not show all species in this example."}
-if(!initNEUS){
+if(initCCA){
   compareB + 
@@
 Given the misses using survey number outputs from run truth with a calculated weight at age, we will not repeat this example using numbers.
-
-## References

### 2019-05-01 docs/index.Rmd
standard survey NEUS example
add std survey visualizations all mods

@@
 
+### Create a standard survey
+
+<a href="https://sgaichas.github.io/poseidon-dev/StandardSurveyTest.html" target="_blank">Initial survey tests - April 30, 2019, updated May 1, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NEUSStandardSurveyTest.html" target="_blank">Standard survey visualization for NEUS -  May 1, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/CCAStandardSurveyTest.html" target="_blank">Standard survey visualization for CCA - May 1, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NOBAStandardSurveyTest.html" target="_blank">Standard survey visualization for NOBA - May 1, 2019</a>
+
 ### More to come later!

### 2019-05-02 visualization_Rmd/CCAStandardSurveyTest.Rmd
reorganize visualization scripts

--- 2019-05-01 NOBAStandardSurveyTest.Rmd
@@
 ---
-title: "Testing atlantisom: example bottom trawl survey index Norwegian-Barents Atlantis"
+title: "Testing atlantisom: example bottom trawl survey index from California Current Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-This page has visualizations for the NOBA model example, CERES Global Sustainability. For full explanation of methods, see the file linked at the beginning of each section.
+This page has visualizations for the CCA model example, Atlantis Summit Common Scenario 1. For full explanation of methods, see the file linked at the beginning of each section.
 
@@
 compareB <-ggplot() +
-  geom_line(data=censusB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
+  geom_line(data=censusB_frombio, aes(x=time,y=atoutput, color="survey census B"), 
             alpha = 5/10) +
@@
   #geom_line(data=surveyB_frombio_effsel, aes(x=time/5,y=atoutput, color="stdeffselB")) +
-  geom_point(data=stdsurveyB_frombio, aes(x=time/5,y=atoutput, color="stdeffcvselB"),
+  geom_point(data=stdsurveyB_frombio, aes(x=time,y=atoutput, color="stdeffcvselB"),
               alpha = 1/2) +
@@
 Depending on the model, we may not show all species in this example."}
-if(initCCA){
+if(!initNEUS){
   compareB + 

### 2019-05-02 visualization_Rmd/NEUSStandardSurveyTest.Rmd
reorganize visualization scripts

--- 2019-05-01 NOBAStandardSurveyTest.Rmd
@@
 ---
-title: "Testing atlantisom: example bottom trawl survey index Norwegian-Barents Atlantis"
+title: "Testing atlantisom: example bottom trawl survey index from Northeast US Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-This page has visualizations for the NOBA model example, CERES Global Sustainability. For full explanation of methods, see the file linked at the beginning of each section.
+This page has visualizations for the NEUS model example, test03032016 (note! a not fully calibrated model). For full explanation of methods, see the file linked at the beginning of each section.
 
@@
 
+Comparing our (census) survey based on true biomass from above with the Atlantis output file "[modelscenario]BiomIndx.txt" should give us a perfect match.
+Note that the our (census) survey may have more sampling in time than the Atlantis output file.
+
 ```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species.", message=FALSE, warning=FALSE}

### 2019-05-02 TrueLengthAgeCompTest.Rmd
start age comp testing

--- 2019-05-01 StandardSurveyTest.Rmd
@@
 ---
-title: "Testing atlantisom: example bottom trawl survey index"
+title: "Testing atlantisom: generate census length and age comps"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-## Simulate a survey part 2: create a standard survey
+## Simulate a survey part 3: sample for length and age composition
 
-This section uses the `atlantisom::create_survey()` and `atlantisom::sample_survey_biomass()` functions by creating a standard survey covering discrete areas, times, and species, and with both incomplete efficiency and some kind of length selectivity.
+This section uses the `atlantisom::create_survey()` and # script for using create_survey with the true numbers at age directly
+# and then sample_fish to get a biolgical sample dataset
+# from that dataset:
+#   age samples from sample_ages
+#   length samples from sample_lengths (TO BE WRITTEN)
+#   weight samples from sample_weights (TO BE WRITTEN)
+#   diet samples from sample_diet
+ `atlantisom::sample_survey_biomass()` functions by creating a standard survey covering discrete areas, times, and species, and with both incomplete efficiency and some kind of length selectivity.
 We will apply this standard survey and compare it with true output as an example of the type of information that can be generated for assessment model performance testing.

### 2019-05-02 TrueLengthAgeCompTest.Rmd
age comp sampling up to broken function

@@
 
-We also read in previously generated survey census files based on true biomass results for comparison (see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).)
+We can also read in previously generated survey census files based on true biomass results for comparison if necessary, but not yet used here (see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).)
 
-```{r readsurvey1}
+```{r readsurvey1, eval=FALSE}
 
@@
 
-This section uses the `atlantisom::create_survey()` and # script for using create_survey with the true numbers at age directly
-# and then sample_fish to get a biolgical sample dataset
-# from that dataset:
-#   age samples from sample_ages
-#   length samples from sample_lengths (TO BE WRITTEN)
-#   weight samples from sample_weights (TO BE WRITTEN)
-#   diet samples from sample_diet
- `atlantisom::sample_survey_biomass()` functions by creating a standard survey covering discrete areas, times, and species, and with both incomplete efficiency and some kind of length selectivity.
-We will apply this standard survey and compare it with true output as an example of the type of information that can be generated for assessment model performance testing.
+This section uses the `atlantisom::create_survey()` and `atlantisom::sample_fish()` to get a biological sample dataset. From that dataset, several other functions can be run:
+
+ *  age samples from sample_ages
+ *  length samples from sample_lengths (TO BE WRITTEN)
+ *  weight samples from sample_weights (TO BE WRITTEN)
+ *  diet samples from sample_diet
 
@@
 
-One approach to creating a standard survey is to map species into general groups that are likely to have similar efficency and selectivity.
-For example, we expect that a bottom trawl survey will have generally lower efficiency for pelagic species than for demersals.
-We can also set efficiency to a very small number for infrequently encountered species occupying habitats that are not suited to bottom trawl sampling.
-This can simulate survey sampling for "data poor" species.
-Similarly, a common selectivity at age (or in this case, Atlantis cohort) can be defined for each species group.
+The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages:
 
-First, we map species into general groups for each model:
+```{r census-spec, message=FALSE, warning=FALSE}
 
-```{r sppgroups}
+# should return a perfectly scaled survey 
+effic1 <- data.frame(species=funct.group.names,
+                     efficiency=rep(1.0,length(funct.group.names)))
 
-# make defaults that return a standard survey, implement in standard_survey
-# users need to map their species groups into these general ones
-#   large pelagics/reef associated/burrowers/otherwise non-trawlable
-#   pelagics
-#   demersals
-#   selected flatfish
+# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
+# BUT CHECK if newer Atlantis models can do age-specific outputs
+selex1 <- data.frame(species=rep(funct.group.names, each=10),
+                     agecl=rep(c(1:10),length(funct.group.names)),
+                     selex=rep(1.0,length(funct.group.names)*10))
 
-if(initCCA) { #Sarah's CCA Grouping
-  nontrawl <- c("Shark_C","Yelloweye_rockfish","Benthopel_Fish","Pisciv_S_Fish",
-                "Pisciv_T_Fish","Shark_D","Shark_P")
-  pelagics <- c("Pisciv_V_Fish","Demersal_S_Fish","Pacific_Ocean_Perch","Mesopel_M_Fish",
-                "Planktiv_L_Fish","Jack_mackerel","Planktiv_S_Fish","Pacific_sardine",
-                "Anchovy","Herring","Pisciv_B_Fish")
-  demersals <- c("Demersal_P_Fish","Planktiv_O_Fish","Demersal_D_Fish",
-                 "Demersal_DC_Fish","Demersal_O_Fish","Darkblotched_rockfish",
-                 "Demersal_F_Fish","Demersal_E_Fish","Bocaccio_rockfish",
-                 "Demersal_B_Fish","Shark_R","Mesopel_N_Fish","Shark_B","Spiny_dogfish",
-                 "SkateRay")
-  selflats <- c("Pisciv_D_Fish", "Arrowtooth_flounder","Petrale_sole")
-}
+# should return all model areas
+boxpars <- load_box(d.name, box.file)
+boxall <- c(0:(boxpars$nbox - 1))
 
-if(initNEUS) { # Sarah's NEUS Grouping
-  nontrawl <- c("Pisciv_T_Fish", "Shark_D", "Shark_P", "Reptile", "Mesopel_M_Fish")
-  pelagics <- c("Planktiv_L_Fish", "Planktiv_S_Fish", "Benthopel_Fish", "Pisciv_S_Fish")
-  demersals <- c("Pisciv_D_Fish", "Demersal_D_Fish","Demersal_E_Fish", 
-                 "Demersal_S_Fish","Demersal_B_Fish","Demersal_DC_Fish",
-                 "Demersal_O_Fish","Demersal_F_Fish",
-                 "Shark_B", "SkateRay")
-  selflats <- c("Pisciv_B_Fish")
-}
+# these are model specific, generalized above
+# if(initCCA) boxall <- c(0:88) 
+# if(initNEUS) boxall <- c(0:29)
+# if(initNOBA) boxall <- c(0:59)
 
-if(initNOBA) { # Sarah's NOBA Grouping
-  nontrawl <- c("Sharks_other", "Pelagic_large","Mesop_fish")
-  pelagics <- c("Pelagic_small","Redfish_other","Mackerel","Haddock",
-                "Saithe","Redfish","Blue_whiting","Norwegian_ssh","Capelin")
-  demersals <- c("Demersals_other","Demersal_large","Flatfish_other","Skates_rays",
-                 "Green_halibut","North_atl_cod","Polar_cod","Snow_crab")
-  selflats <- c("Long_rough_dab")
-}
+# should return all model output timesteps; need to generalize
+# if(initCCA) timeall <- c(0:100) 
+# if(initNEUS) timeall <- c(0:251)
+# if(initNOBA) timeall <- c(0:560)
 
-```
-
-We use the following specifications for our default standard bottom trawl survey, including survey cv by species group:
-
-```{r stdbtsurvey-spec, message=FALSE, warning=FALSE}
-
-# general specifications for bottom trawl survey:
-#   once per year at mid year
-#   could generalize from the run.prm file: 
+# generalized
 runpar <- load_runprm(d.name, run.prm.file)
@@
 stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
-#   take midpoint of 0, steps per year to start seq and go to max time by steps per year
-midptyr <- round(median(seq(1,stepperyr)))
 
-annualmidyear <- seq(midptyr, noutsteps, stepperyr)
+timeall <- c(0:noutsteps)
 
-#   hardcoded for NEUS 1.0 in trunk here
-# annualmidyear <- seq(3,251,5)
-# #other options for NEUS
-# annualspring <- seq(2,251,5)
-# annualfall <- seq(4,251,5)
+# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
+# for ecosystem indicator work test all species, e.g.
+survspp <- funct.group.names
 
-#   all fish (and sharks!! need dogfish in NEUS)
-# could add groups "CEP" for cephalopod, "PWN" for prawn, etc
+# to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
+# if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
+# if(initNEUS) survspp <- funct.group.names[1:21]
+# if(initNOBA) survspp <- funct.group.names[1:36]
 
+# for length and age groups lets just do fish and sharks
 # NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
@@
 
-#   ~75-80% of boxes (leave off deeper boxes?)
-#   cant think of a way to generalize across models, must be hard coded in reality
-#   hardcoded for NEUS 1.0 in trunk here: boxsurv <- c(1:21)
-boxpars <- load_box(d.name, box.file)
-# should return all model areas
-# boxall <- c(0:(boxpars$nbox - 1))
-boxsurv <- c(2:round(0.8*(boxpars$nbox - 1)))
+```
 
-#   define bottom trawl mixed efficiency
-ef.nt <- 0.01 # for large pelagics, reef dwellers, others not in trawlable habitat
-ef.pl <- 0.1  # for pelagics
-ef.dm <- 0.7  # for demersals
-ef.fl <- 1.1  # for selected flatfish
+Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age and length composition.
 
-# bottom trawl survey efficiency specification by species group
-effnontrawl <- data.frame(species=nontrawl, efficiency=rep(ef.nt,length(nontrawl)))
-effpelagics <- data.frame(species=pelagics, efficiency=rep(ef.pl,length(pelagics)))
-effdemersals <- data.frame(species=demersals, efficiency=rep(ef.dm,length(demersals)))
-effselflats <- data.frame(species=selflats, efficiency=rep(ef.fl,length(selflats)))
+```{r stdsurveyNbased}
 
-efficmix <- bind_rows(effnontrawl, effpelagics, effdemersals, effselflats)
+# this uses result$nums, but we are not creating a biomass index this time, so we don't need a weight at age conversion
 
-#   mixed selectivity (using 10 agecl for all species)
-#     flat=1 for large pelagics, reef dwellers, others not in trawlable habitat
-#     sigmoid 0 to 1 with 0.5 inflection at agecl 3 for pelagics, reaching 1 at agecl 5, flat top
-#     sigmoid 0 to 1 with 0.5 inflection at agecl 5 for most demersals and flatfish, reaching 1 at agecl 7, flat top
-#     dome shaped 0 to 1 at agecl 6&7 for selected demersals, falling off to 0.7 by agecl 10
+if(initCCA) datN <- CCAresults$nums
+if(initNEUS) datN <- NEUSresults$nums
+if(initNOBA) datN <- NOBAresults$nums
 
-sigmoid <- function(a,b,x) {
-  1 / (1 + exp(-a-b*x))
-}
+survey_testNall <- create_survey(dat = datN,
+                                 time = timeall,
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
 
-# survey selectivity specification by species group
-selnontrawl <- data.frame(species=rep(nontrawl, each=10),
-                          agecl=rep(c(1:10),length(nontrawl)),
-                          selex=rep(1.0,length(nontrawl)*10))
-selpelagics <- data.frame(species=rep(pelagics, each=10),
-                          agecl=rep(c(1:10),length(pelagics)),
-                          selex=sigmoid(5,1,seq(-10,10,length.out=10)))
-seldemersals <- data.frame(species=rep(demersals, each=10),
-                          agecl=rep(c(1:10),length(demersals)),
-                          selex=sigmoid(1,1,seq(-10,10,length.out=10)))
-selselflats <- data.frame(species=rep(selflats, each=10),
-                          agecl=rep(c(1:10),length(selflats)),
-                          selex=sigmoid(1,1,seq(-10,10,length.out=10)))
-
-# same selectivity for selflats and demersals specified above
-# visualze selectivity curves for each group
-par(mfrow=c(2,2))
-par(mar=c(4,4,1,1))
-plot(selnontrawl$agecl, selnontrawl$selex)
-plot(selpelagics$agecl, selpelagics$selex)
-plot(seldemersals$agecl, seldemersals$selex)
-plot(selselflats$agecl, selselflats$selex)
-par(mfrow=c(1,1))
-
-# implement dome shaped selectivity for a particular species and replace default
-# not done yet staying simple for now
-
-selexmix <- bind_rows(selnontrawl, selpelagics, seldemersals, selselflats)
-
-# use this constant 0 cv for testing
-surv_cv_0 <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
-
-#   define bottom trawl survey cv by group
-cv.nt <- 1.0 # for large pelagics, reef dwellers, others not in trawlable habitat
-cv.pl <- 0.5  # for pelagics
-cv.dm <- 0.3  # for demersals
-cv.fl <- 0.3  # for selected flatfish
-
-# specify cv by species groups
-surv_cv_nontrawl <- data.frame(species=nontrawl, cv=rep(cv.nt,length(nontrawl)))
-surv_cv_pelagics <- data.frame(species=pelagics, cv=rep(cv.pl,length(pelagics)))
-surv_cv_demersals <- data.frame(species=demersals, cv=rep(cv.dm,length(demersals)))
-surv_cv_selflats <- data.frame(species=selflats, cv=rep(cv.fl,length(selflats)))
-
-surv_cv_mix <- bind_rows(surv_cv_nontrawl, surv_cv_pelagics, surv_cv_demersals, surv_cv_selflats)
+# consider saving this interim step if it takes a long time go generate
 
@@
 
-Here we use `create_survey` on the biomass output of `run_truth` to create the survey, so the call to `sample_survey_biomass` will require a weight at age argument that is filled with 1's because no conversion from numbers to weight is necessary.
+Next, get true annual (cohort) age comp from this census survey based on run truth. (is there a standard Atlantis output I can compare this to as we did for biomass?)
 
-```{r stdsurveyBbased}
+```{r truecohortagecomp}
+# what is true composition? need annual by species, use code from sample_fish
+# do tidyly
+dat2 <- survey_testNall %>%
+  group_by(species, agecl, time) %>%
+  summarize(numAtAge = sum(atoutput))
 
-# this uses result$biomass_ages to sample biomass directly
+#dat<-survey_testNall
+#dat2 <- aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time),sum)
+#names(dat2) <- c("species","agecl","time","numAtAge")
 
-if(initCCA) datB <- CCAresults$biomass_ages
-if(initNEUS) datB <- NEUSresults$biomass_ages
-if(initNOBA) datB <- NOBAresults$biomass_ages
+totN <- dat2 %>%
+  group_by(species, time) %>%
+  summarize(totN = sum(numAtAge))
 
-survey_testBstd <- create_survey(dat = datB,
-                                 time = annualmidyear,
-                                 species = survspp,
-                                 boxes = boxsurv,
-                                 effic = efficmix,
-                                 selex = selexmix)
+#totN <- aggregate(dat2$numAtAge,list(dat2$species,dat2$time),sum )
+#names(totN) <- c("species","time","totN")
 
-# call sample_survey_biomass with a bunch of 1s for weight at age
-# in the code it multiplies atoutput by wtatage so this allows us to use
-# biomass directly
-wtage <- data.frame(species=rep(survspp, each=10),
-                    agecl=rep(c(1:10),length(survspp)),
-                    wtAtAge=rep(1.0,length(survspp)*10))
+dat2totN <- merge(dat2, totN)
 
-stdsurveyB_frombio <- sample_survey_biomass(survey_testBstd, surv_cv_mix, wtage)
+ageclcomp <- ggplot(dat2totN, aes(x=agecl, y=numAtAge/totN, col=time)) +
+  geom_point()
 
-if(initCCA) CCAstdsurveyB_frombio <- stdsurveyB_frombio
-if(initNEUS) NEUSstdsurveyB_frombio <- stdsurveyB_frombio
-if(initNOBA) NOBAstdsurveyB_frombio <- stdsurveyB_frombio
+ageclcomp + facet_wrap(~species, ncol=5) #, scales="free")
 
@@
 
-Comparing our (census) survey based on true biomass from above with the Atlantis output file "[modelscenario]BiomIndx.txt" should give us a perfect match.
-Note that the our (census) survey may have more sampling in time than the Atlantis output file.
+Then we use the `sample_fish` function and compare to true annual age comp calculated above as a test. These should match.
 
-```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species.", message=FALSE, warning=FALSE}
-# plot some comparisons with Atlantis output
+```{r comptest1, message=FALSE, warning=FALSE}
+# setting the effN higher than actual numbers results in sampling all
+effNall <- data.frame(species=survspp, effN=rep(1e+15, length(survspp)))
+# rmultinom broke with that sample size
+#comptestall <- sample_fish(survey_testNall, effNall)
+#names(comptestall) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")
 
-# read Atlantis output files
-if(initCCA) {
-  atBtxt2 <- read.table(here("atlantisoutput","CalCurrentSummitScenario1","outputCCV3BiomIndx.txt"), header=T)
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  stdsurveyB_frombio <- CCAstdsurveyB_frombio
-  censusB_frombio <- CCAsurveyB_frombio
-}
+# this one is high but not equal to total for numerous groups
+effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
 
-if(initNEUS) {
-  atBtxt2 <- read.table(here("atlantisoutput","NEUStest20160303","neusDynEffort_Test1_BiomIndx.txt"), header=T)
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  stdsurveyB_frombio <- NEUSstdsurveyB_frombio
-  censusB_frombio <- NEUSsurveyB_frombio
-}
+comptesthigh <- sample_fish(survey_testNall, effNhigh)
+names(comptesthigh) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")
 
-if(initNOBA) {
-  atBtxt2 <- read.table(here("atlantisoutput","NOBACERESGlobalSustainability","outputnordic_runresults_01BiomIndx.txt"), header=T) 
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  stdsurveyB_frombio <- NOBAstdsurveyB_frombio
-  censusB_frombio <- NOBAsurveyB_frombio
-}
+comptesttot <- aggregate(comptesthigh$numAtAgesamp,list(comptesthigh$species,comptesthigh$time),sum )
+names(comptesttot) <- c("species","time","totsamp")
 
-# lookup the matching names, put in time, species, biomass column format
-# WARNING hardcoded for output with last species group as DIN
-groupslookup <- groupslookup %>%
-  filter(IsTurnedOn > 0)
+comptestprop <- merge(comptesthigh, comptesttot)
 
-atBtxt2tidy <- atBtxt2 %>%
-  select(Time:DIN) %>%
-  #select(Time, FPL:DIN) %>%
-  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
-  gather(species, biomass, -Time) %>%
-  filter(species %in% levels(stdsurveyB_frombio$species))
-
-censusB_frombio <- censusB_frombio %>%
-  filter(species %in% levels(stdsurveyB_frombio$species))
-
-#all species comparison, time intervals hardcoded for NEUS and NOBA
-compareB <-ggplot() +
-  geom_line(data=censusB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
-            alpha = 5/10) +
-  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
-             alpha = 5/10) + 
-  #geom_line(data=surveyB_frombio_eff, aes(x=time/5,y=atoutput, color="stdeffB")) +
-  #geom_line(data=surveyB_frombio_effsel, aes(x=time/5,y=atoutput, color="stdeffselB")) +
-  geom_point(data=stdsurveyB_frombio, aes(x=time/5,y=atoutput, color="stdeffcvselB"),
-              alpha = 1/2) +
-
+# compare individual years, these proportions at age should match
+comparecomps <- ggplot() +
+  geom_point(data=subset(dat2totN, time==0), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==0), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3) +
   theme_tufte() +
@@
 
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+comparecomps + facet_wrap(~species, ncol=5)
+
+#ggsave("censuscomposition_time0.png", width=11, height=11)
+
+comparecomps <- ggplot() +
+  geom_point(data=subset(dat2totN, time==100), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==100), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+comparecomps + facet_wrap(~species, ncol=5)
+
+#ggsave("censuscomposition_time100.png", width=11, height=11)
+
+comparecomps <- ggplot() +
+  geom_point(data=subset(dat2totN, time==250), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==250), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+comparecomps + facet_wrap(~species, ncol=5)
+
+#ggsave("censuscomposition_time250.png", width=11, height=11)
+
+# and they do
+# differences for baleen and toothed whales are due to small total numbers
 
@@
 
-```{r matchBp2, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+And they did for NEUS the first time I did this, with some exceptions.
+
+The problems start when I try to do age to length, this function returns an error at present.
+
+**Editing to be done in atlantisom (I have made local changes to calcage2length.R that are not pushed yet because I may have introduced an error)**
+
+```{r calcage2length-breaks, eval=FALSE}
+# now turn this into length comp with calc_age2length
+# function takes resn and nums as inputs so run survey and sample functions on resn as well
+if(initCCA) datresn <- CCAresults$resn
+if(initNEUS) datresn <- NEUSresults$resn
+if(initNOBA) datresn <- NOBAresults$resn
+
+survey_testresnall <- create_survey(dat = datresn,
+                                 time = seq(0,251,5), # use timeall? census?
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
+lengthtesthigh <- sample_fish(survey_testNall, effNhigh)
+# use fish sampled from N along with survey of resn in function
+# calc_age2length uses nums (lengthtest) and finds matches in resn
+
+# provides ouput for an individual species and time slice
+# breaks for whole dataset (NA output all rows)
+
+if(initCCA) truth <- CCAresults
+if(initNEUS) truth <- NEUSresults
+if(initNOBA) truth <- NOBAresults
+
+length_census <- calc_age2length(structn = lengthtesthigh,
+                                 resn = survey_testresnall,
+                                 nums = lengthtesthigh,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
 
@@
 
-```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
-
-```
-
-```{r matchBp4, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species.
-Depending on the model, we may not show all species in this example."}
-if(!initNEUS){
-  compareB + 
-    facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free") 
-}
-```
-
-Given the misses using survey number outputs from run truth with a calculated weight at age, we will not repeat this example using numbers.
-
 ## References

### 2019-05-03 TrueLengthAgeCompTest.Rmd
CCA some funky outputs, had to change sample_fish

@@
 
-initCCA <- FALSE
-initNEUS <- TRUE
+initCCA <- TRUE
+initNEUS <- FALSE
 initNOBA <- FALSE
@@
 
+#if(initCCA) survspp <- survspp[!survspp %in% "Pisciv_T_Fish"]
+
 ```
@@
 comptestprop <- merge(comptesthigh, comptesttot)
+```
+
+```{r comptest1plot, message=FALSE, warning=FALSE}
 
@@
 comparecomps <- ggplot() +
-  geom_point(data=subset(dat2totN, time==0), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
-  geom_point(data=subset(comptestprop, time==0), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3) +
+  geom_point(data=subset(dat2totN, time==min(timeall)), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==min(timeall)), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3) +
   theme_tufte() +
   theme(legend.position = "top") +
-  labs(colour=scenario.name)
+  labs(colour=paste0(scenario.name, " start"))
 
@@
 comparecomps <- ggplot() +
-  geom_point(data=subset(dat2totN, time==100), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
-  geom_point(data=subset(comptestprop, time==100), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
+  geom_point(data=subset(dat2totN, time==median(timeall)), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==median(timeall)), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
   theme_tufte() +
   theme(legend.position = "top") +
-  labs(colour=scenario.name)
+  labs(colour=paste0(scenario.name, " midpoint"))
 
@@
 comparecomps <- ggplot() +
-  geom_point(data=subset(dat2totN, time==250), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
-  geom_point(data=subset(comptestprop, time==250), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
+  geom_point(data=subset(dat2totN, time==max(timeall)-1), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==max(timeall)-1), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
   theme_tufte() +
   theme(legend.position = "top") +
-  labs(colour=scenario.name)
+  labs(colour=paste0(scenario.name, " end"))
 

### 2019-05-06 TrueLengthAgeCompTest.Rmd
better visualizations

@@
 
-ageclcomp + facet_wrap(~species, ncol=5) #, scales="free")
+ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+#ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
 
@@
 
-comparecomps + facet_wrap(~species, ncol=5)
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
 

### 2019-05-06 docs/index.Rmd
age comp sampling up to broken function
update index
age comp census visualizations all mods

@@
 
+### Start age composition sampling: work in progress
+
+<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthAgeCompTest.html" target="_blank">Initial age comp tests - May 2, 2019, updated May 6, 2019</a>
+
+CCA outputs had some 0 length at age, requiring changes to `atlantisom::sample_fish`
+
+<a href="https://sgaichas.github.io/poseidon-dev/NEUSTrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison NEUS - May 6, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/CCATrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison CCA - May 6, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NOBATrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison NOBA - May 6, 2019</a>
+
 ### More to come later!

### 2019-05-06 visualization_Rmd/CCATrueLengthAgeCompTest.Rmd
age comp census visualizations all mods

--- 2019-05-06 TrueLengthAgeCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate census length and age comps"
+title: "Testing atlantisom: generate census length and age comps from California Current Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets.
-Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
-
-The purpose of atlantisom is to use existing Atlantis model output to generate input datasets for a variety of models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics.
-Atlantis models can be run using different climate forcing, fishing, and other scenarios.
-Users of atlantisom will be able to specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes for any Atlantis scenario.
-Internally consistent multispecies and ecosystem datasets with known observation error characteristics will be the atlantisom outputs, for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against "true" Atlantis outputs.
-
-Initial testing was conducted by S.
-Gaichas using R scripts in the R folder of this repository that are titled "PoseidonTest_[whatwastested].R".
-Initial tests are expanded and documented in more detail in these pages.
-C.
-Stawitz improved and streamlined the setup and intialization sections.
-
-## Setup
-
-First, you will want to set up libraries and install atlantisom if you haven't already.
-This document assumes atlantisom is already installed.
-For complete setup and initialization, please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).
-
-This document is written in in R Markdown [@R-rmarkdown], and we use several packages to produce the outputs [@R-tidyr; @R-dplyr; @R-ggplot2; @R-here; @R-ggforce; @R-ggthemes].
+This page has visualizations for the CCA model example, Atlantis Summit Common Scenario 1. For full explanation of methods, see the file linked at the beginning of each section.
 
@@
 ```
-
-## Initialize input files and directories, read in "truth"
-
-Abbreviated here; for a full explanation please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).
-This document assumes that `atlantisom::run_truth` has already completed and stored an .RData file in the atlantis output model directory.
 
@@
 
-We can also read in previously generated survey census files based on true biomass results for comparison if necessary, but not yet used here (see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).)
-
 ```{r readsurvey1, eval=FALSE}
@@
 
-This section uses the `atlantisom::create_survey()` and `atlantisom::sample_fish()` to get a biological sample dataset. From that dataset, several other functions can be run:
-
- *  age samples from sample_ages
- *  length samples from sample_lengths (TO BE WRITTEN)
- *  weight samples from sample_weights (TO BE WRITTEN)
- *  diet samples from sample_diet
-
-To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
+Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueLengthAgeCompTest.html).
 
@@
 
-```{r census-spec, message=FALSE, warning=FALSE}
+```{r census-spec, message=FALSE, warning=FALSE, echo=TRUE}
 
@@
 
-Then we use the `sample_fish` function and compare to true annual age comp calculated above as a test. These should match.
+Then we use the `sample_fish` function and compare to true annual age comp calculated above as a test.
+These should match.
+We show compositions sampled for the initial time step, a time step in the middle of the run, and the last timestep.
 
@@
 
-And they did for NEUS the first time I did this, with some exceptions.
-
-The problems start when I try to do age to length, this function returns an error at present.
-
-**Editing to be done in atlantisom (I have made local changes to calcage2length.R that are not pushed yet because I may have introduced an error)**
-
-```{r calcage2length-breaks, eval=FALSE}
-# now turn this into length comp with calc_age2length
-# function takes resn and nums as inputs so run survey and sample functions on resn as well
-if(initCCA) datresn <- CCAresults$resn
-if(initNEUS) datresn <- NEUSresults$resn
-if(initNOBA) datresn <- NOBAresults$resn
-
-survey_testresnall <- create_survey(dat = datresn,
-                                 time = seq(0,251,5), # use timeall? census?
-                                 species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-lengthtesthigh <- sample_fish(survey_testNall, effNhigh)
-# use fish sampled from N along with survey of resn in function
-# calc_age2length uses nums (lengthtest) and finds matches in resn
-
-# provides ouput for an individual species and time slice
-# breaks for whole dataset (NA output all rows)
-
-if(initCCA) truth <- CCAresults
-if(initNEUS) truth <- NEUSresults
-if(initNOBA) truth <- NOBAresults
-
-length_census <- calc_age2length(structn = lengthtesthigh,
-                                 resn = survey_testresnall,
-                                 nums = lengthtesthigh,
-                                 biolprm = truth$biolprm, fgs = truth$fgs,
-                                 CVlenage = 0.1, remove.zeroes=TRUE)
-
-```
-
-## References
+Most match, with some exceptions to be discussed with the regional model developers. Mismatches seem to happen later in this CCA run?

### 2019-05-06 visualization_Rmd/NEUSTrueLengthAgeCompTest.Rmd
age comp census visualizations all mods

--- 2019-05-06 visualization_Rmd/CCATrueLengthAgeCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate census length and age comps from California Current Atlantis"
+title: "Testing atlantisom: generate census length and age comps from Northeast US Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
-ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+#ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
 #ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
@@
 comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
 #comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
@@
 
-Most match, with some exceptions to be discussed with the regional model developers. Mismatches seem to happen later in this CCA run?
+Most match, with some exceptions to be discussed with the regional model developers. NEUS run looks good here.

### 2019-05-06 visualization_Rmd/NOBATrueLengthAgeCompTest.Rmd
age comp census visualizations all mods

--- 2019-05-06 visualization_Rmd/NEUSTrueLengthAgeCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate census length and age comps from Northeast US Atlantis"
+title: "Testing atlantisom: generate census length and age comps from Norwegian-Barents Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-Most match, with some exceptions to be discussed with the regional model developers. NEUS run looks good here.
+NOBA matches perfectly here.

### 2019-05-06 StdSurvLengthAgeCompTest.Rmd
initial comps from std survey

--- 2019-05-01 StandardSurveyTest.Rmd
@@
 ---
-title: "Testing atlantisom: example bottom trawl survey index"
+title: "Testing atlantisom: generate standard survey length and age comps"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-We also read in previously generated survey census files based on true biomass results for comparison (see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).)
+<!--We can also read in previously generated survey census files based on true biomass results for comparison if necessary, but not yet used here (see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).)
 
-```{r readsurvey1}
+```{r readsurvey1, eval=FALSE}
 
@@
 ```
+-->
 
-## Simulate a survey part 2: create a standard survey
+## Simulate a survey part 3: sample for length and age composition
 
-This section uses the `atlantisom::create_survey()` and `atlantisom::sample_survey_biomass()` functions by creating a standard survey covering discrete areas, times, and species, and with both incomplete efficiency and some kind of length selectivity.
-We will apply this standard survey and compare it with true output as an example of the type of information that can be generated for assessment model performance testing.
+Here we will compare the effects of survey timing, spatial coverage, and survey selectivity on the sampled (cohort) age comps.
+This section uses the `atlantisom::create_survey()` and `atlantisom::sample_fish()` to get a biological sample dataset.
+From that dataset, several other functions can be run:
+
+ *  age samples from sample_ages
+ *  length samples from sample_lengths (TO BE WRITTEN)
+ *  weight samples from sample_weights (TO BE WRITTEN)
+ *  diet samples from sample_diet
 
@@
 
-One approach to creating a standard survey is to map species into general groups that are likely to have similar efficency and selectivity.
-For example, we expect that a bottom trawl survey will have generally lower efficiency for pelagic species than for demersals.
-We can also set efficiency to a very small number for infrequently encountered species occupying habitats that are not suited to bottom trawl sampling.
-This can simulate survey sampling for "data poor" species.
-Similarly, a common selectivity at age (or in this case, Atlantis cohort) can be defined for each species group.
+The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages:
 
-First, we map species into general groups for each model:
+```{r census-spec, message=FALSE, warning=FALSE}
+
+# should return a perfectly scaled survey 
+effic1 <- data.frame(species=funct.group.names,
+                     efficiency=rep(1.0,length(funct.group.names)))
+
+# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
+# BUT CHECK if newer Atlantis models can do age-specific outputs
+selex1 <- data.frame(species=rep(funct.group.names, each=10),
+                     agecl=rep(c(1:10),length(funct.group.names)),
+                     selex=rep(1.0,length(funct.group.names)*10))
+
+# should return all model areas
+boxpars <- load_box(d.name, box.file)
+boxall <- c(0:(boxpars$nbox - 1))
+
+# these are model specific, generalized above
+# if(initCCA) boxall <- c(0:88) 
+# if(initNEUS) boxall <- c(0:29)
+# if(initNOBA) boxall <- c(0:59)
+
+# should return all model output timesteps; need to generalize
+# if(initCCA) timeall <- c(0:100) 
+# if(initNEUS) timeall <- c(0:251)
+# if(initNOBA) timeall <- c(0:560)
+
+# generalized
+runpar <- load_runprm(d.name, run.prm.file)
+noutsteps <- runpar$tstop/runpar$outputstep
+stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
+
+timeall <- c(0:noutsteps)
+
+# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
+# for ecosystem indicator work test all species, e.g.
+survspp <- funct.group.names
+
+# to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
+# if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
+# if(initNEUS) survspp <- funct.group.names[1:21]
+# if(initNOBA) survspp <- funct.group.names[1:36]
+
+# for length and age groups lets just do fish and sharks
+# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
+if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)
+
+survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
+                           funct.groups$GroupType %in% c("FISH", "SHARK")]
+
+#if(initCCA) survspp <- survspp[!survspp %in% "Pisciv_T_Fish"]
+
+```
+
+Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age and length composition.
+
+```{r censusNbased}
+
+# this uses result$nums, but we are not creating a biomass index this time, so we don't need a weight at age conversion
+
+if(initCCA) datN <- CCAresults$nums
+if(initNEUS) datN <- NEUSresults$nums
+if(initNOBA) datN <- NOBAresults$nums
+
+survey_testNall <- create_survey(dat = datN,
+                                 time = timeall,
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
+# consider saving this interim step if it takes a long time go generate
+
+```
+
+Next, get true annual (cohort) age comp from this census survey based on run truth. (is there a standard Atlantis output I can compare this to as we did for biomass?)
+
+```{r truecohortagecomp}
+# what is true composition? need annual by species, use code from sample_fish
+# do tidyly
+dat2 <- survey_testNall %>%
+  group_by(species, agecl, time) %>%
+  summarize(numAtAge = sum(atoutput))
+
+#dat<-survey_testNall
+#dat2 <- aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time),sum)
+#names(dat2) <- c("species","agecl","time","numAtAge")
+
+totN <- dat2 %>%
+  group_by(species, time) %>%
+  summarize(totN = sum(numAtAge))
+
+#totN <- aggregate(dat2$numAtAge,list(dat2$species,dat2$time),sum )
+#names(totN) <- c("species","time","totN")
+
+dat2totN <- merge(dat2, totN)
+
+# ageclcomp <- ggplot(dat2totN, aes(x=agecl, y=numAtAge/totN, col=time)) +
+#   geom_point()
+# 
+# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
+
+```
+
+We know from the previous test that the `sample_fish` function with the maximum possible effective sample size compares favorably (or perfectly) to true annual age comp calculated above as a test.
+Now we examine the effects of changing survey timing, survey area, survey selectivity, and fish sample size on age comps.
+
+The standard survey approach is described [here](https://sgaichas.github.io/poseidon-dev/StandardSurveyTest.html). First, we map species into general groups for each model:
 
@@
 
-# general specifications for bottom trawl survey:
+# general specifications for bottom trawl survey, with items defined above commented out to avoid wasting time loading already loaded files:
 #   once per year at mid year
 #   could generalize from the run.prm file: 
-runpar <- load_runprm(d.name, run.prm.file)
-noutsteps <- runpar$tstop/runpar$outputstep
-stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
+# runpar <- load_runprm(d.name, run.prm.file)
+# noutsteps <- runpar$tstop/runpar$outputstep
+# stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
 #   take midpoint of 0, steps per year to start seq and go to max time by steps per year
@@
 
-#   hardcoded for NEUS 1.0 in trunk here
-# annualmidyear <- seq(3,251,5)
-# #other options for NEUS
-# annualspring <- seq(2,251,5)
-# annualfall <- seq(4,251,5)
-
-#   all fish (and sharks!! need dogfish in NEUS)
-# could add groups "CEP" for cephalopod, "PWN" for prawn, etc
-
-# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
-if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)
-
-survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
-                           funct.groups$GroupType %in% c("FISH", "SHARK")]
-
 #   ~75-80% of boxes (leave off deeper boxes?)
-#   cant think of a way to generalize across models, must be hard coded in reality
-#   hardcoded for NEUS 1.0 in trunk here: boxsurv <- c(1:21)
-boxpars <- load_box(d.name, box.file)
-# should return all model areas
-# boxall <- c(0:(boxpars$nbox - 1))
+# boxpars <- load_box(d.name, box.file)
 boxsurv <- c(2:round(0.8*(boxpars$nbox - 1)))
@@
 
-# same selectivity for selflats and demersals specified above
-# visualze selectivity curves for each group
-par(mfrow=c(2,2))
-par(mar=c(4,4,1,1))
-plot(selnontrawl$agecl, selnontrawl$selex)
-plot(selpelagics$agecl, selpelagics$selex)
-plot(seldemersals$agecl, seldemersals$selex)
-plot(selselflats$agecl, selselflats$selex)
-par(mfrow=c(1,1))
-
-# implement dome shaped selectivity for a particular species and replace default
-# not done yet staying simple for now
-
 selexmix <- bind_rows(selnontrawl, selpelagics, seldemersals, selselflats)
@@
 
-Here we use `create_survey` on the biomass output of `run_truth` to create the survey, so the call to `sample_survey_biomass` will require a weight at age argument that is filled with 1's because no conversion from numbers to weight is necessary.
+This generates the standard survey once per year, most areas, with mixed efficiency and selectivity.
 
-```{r stdsurveyBbased}
+```{r stdsurveyNbased}
 
-# this uses result$biomass_ages to sample biomass directly
+# this uses result$nums, but we are not creating a biomass index this time, so we don't need a weight at age conversion
 
-if(initCCA) datB <- CCAresults$biomass_ages
-if(initNEUS) datB <- NEUSresults$biomass_ages
-if(initNOBA) datB <- NOBAresults$biomass_ages
+if(initCCA) datN <- CCAresults$nums
+if(initNEUS) datN <- NEUSresults$nums
+if(initNOBA) datN <- NOBAresults$nums
 
-survey_testBstd <- create_survey(dat = datB,
+survey_testNstd <- create_survey(dat = datN,
                                  time = annualmidyear,
@@
 
-# call sample_survey_biomass with a bunch of 1s for weight at age
-# in the code it multiplies atoutput by wtatage so this allows us to use
-# biomass directly
-wtage <- data.frame(species=rep(survspp, each=10),
-                    agecl=rep(c(1:10),length(survspp)),
-                    wtAtAge=rep(1.0,length(survspp)*10))
-
-stdsurveyB_frombio <- sample_survey_biomass(survey_testBstd, surv_cv_mix, wtage)
-
-if(initCCA) CCAstdsurveyB_frombio <- stdsurveyB_frombio
-if(initNEUS) NEUSstdsurveyB_frombio <- stdsurveyB_frombio
-if(initNOBA) NOBAstdsurveyB_frombio <- stdsurveyB_frombio
+# consider saving this interim step if it takes a long time to generate
 
@@
 
-Comparing our (census) survey based on true biomass from above with the Atlantis output file "[modelscenario]BiomIndx.txt" should give us a perfect match.
-Note that the our (census) survey may have more sampling in time than the Atlantis output file.
+Now we sample fish with no error from the standard survey:
 
-```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species.", message=FALSE, warning=FALSE}
-# plot some comparisons with Atlantis output
+```{r comptest1, message=FALSE, warning=FALSE}
+# setting the effN higher than actual numbers results in sampling all
 
-# read Atlantis output files
-if(initCCA) {
-  atBtxt2 <- read.table(here("atlantisoutput","CalCurrentSummitScenario1","outputCCV3BiomIndx.txt"), header=T)
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  stdsurveyB_frombio <- CCAstdsurveyB_frombio
-  censusB_frombio <- CCAsurveyB_frombio
-}
+# this one is high but not equal to total for numerous groups
+effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
 
-if(initNEUS) {
-  atBtxt2 <- read.table(here("atlantisoutput","NEUStest20160303","neusDynEffort_Test1_BiomIndx.txt"), header=T)
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  stdsurveyB_frombio <- NEUSstdsurveyB_frombio
-  censusB_frombio <- NEUSsurveyB_frombio
-}
+comptesthigh <- sample_fish(survey_testNstd, effNhigh)
+names(comptesthigh) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")
 
-if(initNOBA) {
-  atBtxt2 <- read.table(here("atlantisoutput","NOBACERESGlobalSustainability","outputnordic_runresults_01BiomIndx.txt"), header=T) 
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  stdsurveyB_frombio <- NOBAstdsurveyB_frombio
-  censusB_frombio <- NOBAsurveyB_frombio
-}
+comptesttot <- aggregate(comptesthigh$numAtAgesamp,list(comptesthigh$species,comptesthigh$time),sum )
+names(comptesttot) <- c("species","time","totsamp")
 
-# lookup the matching names, put in time, species, biomass column format
-# WARNING hardcoded for output with last species group as DIN
-groupslookup <- groupslookup %>%
-  filter(IsTurnedOn > 0)
+comptestprop <- merge(comptesthigh, comptesttot)
+```
 
-atBtxt2tidy <- atBtxt2 %>%
-  select(Time:DIN) %>%
-  #select(Time, FPL:DIN) %>%
-  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
-  gather(species, biomass, -Time) %>%
-  filter(species %in% levels(stdsurveyB_frombio$species))
+And compare this to the true age comp above to see the impact of only standard survey sampling on age comps:
 
-censusB_frombio <- censusB_frombio %>%
-  filter(species %in% levels(stdsurveyB_frombio$species))
+```{r comptest1plot, message=FALSE, warning=FALSE}
 
-#all species comparison, time intervals hardcoded for NEUS and NOBA
-compareB <-ggplot() +
-  geom_line(data=censusB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
-            alpha = 5/10) +
-  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
-             alpha = 5/10) + 
-  #geom_line(data=surveyB_frombio_eff, aes(x=time/5,y=atoutput, color="stdeffB")) +
-  #geom_line(data=surveyB_frombio_effsel, aes(x=time/5,y=atoutput, color="stdeffselB")) +
-  geom_point(data=stdsurveyB_frombio, aes(x=time/5,y=atoutput, color="stdeffcvselB"),
-              alpha = 1/2) +
-
+# compare individual years, these proportions at age should not match
+comparecomps <- ggplot() +
+  geom_point(data=subset(dat2totN, time==min(annualmidyear)), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==min(annualmidyear)), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3) +
   theme_tufte() +
   theme(legend.position = "top") +
-  labs(colour=scenario.name)
+  labs(colour=paste0(scenario.name, " start"))
 
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+if(initCCA) comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
+
+comparecomps <- ggplot() +
+  geom_point(data=subset(dat2totN, time==annualmidyear[length(annualmidyear)/2]), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==annualmidyear[length(annualmidyear)/2]), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=paste0(scenario.name, " midpoint"))
+
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+if(initCCA) comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
+
+comparecomps <- ggplot() +
+  geom_point(data=subset(dat2totN, time==annualmidyear[length(annualmidyear)]), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==annualmidyear[length(annualmidyear)]), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=paste0(scenario.name, " end"))
+
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+if(initCCA) comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
 
@@
 
-```{r matchBp2, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+Now we run `sample_fish` with a more realistic biological sample size for each species, to see the combined impact of the standard survey and biological sampling:
+
+```{r comptest2, message=FALSE, warning=FALSE}
+# sample_fish uses an effN effective sample size for multinomial
+# but this should be the REAL sample size taken (annual average over whole area)
+# sample size of fish for lengths--a fairly large number
+# a proportion of this is used for ages below
+# assign by groups as above
+
+#   define n fish for biological sampling by group
+#   this could easily be a vector or time series, constant here
+ns.nt <- 25 # for large pelagics, reef dwellers, others not in trawlable habitat
+ns.pl <- 1000  # for pelagics
+ns.dm <- 1000  # for demersals
+ns.fl <- 1000  # for selected flatfish
+
+effNnontrawl <- data.frame(species=nontrawl, effN=rep(ns.nt,length(nontrawl)))
+effNpelagics <- data.frame(species=pelagics, effN=rep(ns.pl,length(pelagics)))
+effNdemersals <- data.frame(species=demersals, effN=rep(ns.dm,length(demersals)))
+effNselflats <- data.frame(species=selflats, effN=rep(ns.fl,length(selflats)))
+
+effNmix <- bind_rows(effNnontrawl, effNpelagics, effNdemersals, effNselflats)
+
+comptestmix <- sample_fish(survey_testNstd, effNmix)
+names(comptestmix) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")
+
+comptestmixtot <- aggregate(comptestmix$numAtAgesamp,list(comptesthigh$species,comptesthigh$time),sum )
+names(comptestmixtot) <- c("species","time","totsamp")
+
+comptestmixprop <- merge(comptestmix, comptestmixtot)
 
@@
 
-```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+We can compare true (cohort) age comp ("true"), survey sampled age comp with no subsampling effect ("samphigh"), and age comp from biological subsamling aboard the survey ("sampreal"):
+
+```{r comptest12plot, message=FALSE, warning=FALSE}
+
+# compare individual years, these proportions at age wont match
+comparecomps <- ggplot() +
+  geom_point(data=subset(dat2totN, time==min(annualmidyear)), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==min(annualmidyear)), aes(x=agecl, y=numAtAgesamp/totsamp, color="samphigh"), alpha = 0.3) +
+  geom_point(data=subset(comptestmixprop, time==min(annualmidyear)), aes(x=agecl, y=numAtAgesamp/totsamp, color="sampreal"), alpha = 0.3) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=paste0(scenario.name, " start"))
+
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+if(initCCA) comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
+
+#ggsave("stdsurvcomp_realeffN_time0.png", width=11, height=11)
+
+comparecomps <- ggplot() +
+  geom_point(data=subset(dat2totN, time==annualmidyear[length(annualmidyear)/2]), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==annualmidyear[length(annualmidyear)/2]), aes(x=agecl, y=numAtAgesamp/totsamp, color="samphigh"), alpha = 0.3) +
+  geom_point(data=subset(comptestmixprop, time==annualmidyear[length(annualmidyear)/2]), aes(x=agecl, y=numAtAgesamp/totsamp, color="sampreal"), alpha = 0.3) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=paste0(scenario.name, " midpoint"))
+
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+if(initCCA) comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
+
+#ggsave("stdsurvcomp_realeffN_time100.png", width=11, height=11)
+
+comparecomps <- ggplot() +
+  geom_point(data=subset(dat2totN, time==annualmidyear[length(annualmidyear)]), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==annualmidyear[length(annualmidyear)]), aes(x=agecl, y=numAtAgesamp/totsamp, color="samphigh"), alpha = 0.3) +
+  geom_point(data=subset(comptestmixprop, time==annualmidyear[length(annualmidyear)]), aes(x=agecl, y=numAtAgesamp/totsamp, color="sampreal"), alpha = 0.3) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=paste0(scenario.name, " end"))
+
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+if(initCCA) comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
+
+#ggsave("stdsurvcomp_realeffN_time250.png", width=11, height=11)
 
@@
 
-```{r matchBp4, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species.
-Depending on the model, we may not show all species in this example."}
-if(!initNEUS){
-  compareB + 
-    facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free") 
-}
-```
-
-Given the misses using survey number outputs from run truth with a calculated weight at age, we will not repeat this example using numbers.
+With these biological sample sizes (~1000 fish per year in the whole area) the impact of biological subsampling is minimal compared with the impact of survey length selectivity.
+Smaller biological sample sizes (~25 fish per year for nontrawl category) show more difference from the survey sampled age comp.
 

### 2019-05-06 visualization_Rmd/NEUSStdSurvCohortCompTest.Rmd
comp examples NOBA and NEUS

--- 2019-05-06 StdSurvLengthAgeCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate standard survey length and age comps"
+title: "Testing atlantisom: generate standard survey length and age comps from Northeast US Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets.
-Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
-
-The purpose of atlantisom is to use existing Atlantis model output to generate input datasets for a variety of models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics.
-Atlantis models can be run using different climate forcing, fishing, and other scenarios.
-Users of atlantisom will be able to specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes for any Atlantis scenario.
-Internally consistent multispecies and ecosystem datasets with known observation error characteristics will be the atlantisom outputs, for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against "true" Atlantis outputs.
-
-Initial testing was conducted by S.
-Gaichas using R scripts in the R folder of this repository that are titled "PoseidonTest_[whatwastested].R".
-Initial tests are expanded and documented in more detail in these pages.
-C.
-Stawitz improved and streamlined the setup and intialization sections.
-
-## Setup
-
-First, you will want to set up libraries and install atlantisom if you haven't already.
-This document assumes atlantisom is already installed.
-For complete setup and initialization, please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).
-
-This document is written in in R Markdown [@R-rmarkdown], and we use several packages to produce the outputs [@R-tidyr; @R-dplyr; @R-ggplot2; @R-here; @R-ggforce; @R-ggthemes].
+This page has visualizations for the NEUS model example, test03032016 (note! a not fully calibrated model). For full explanation of methods, see the file linked at the beginning of each section.
 
@@
 
-Here we will compare the effects of survey timing, spatial coverage, and survey selectivity on the sampled (cohort) age comps.
-This section uses the `atlantisom::create_survey()` and `atlantisom::sample_fish()` to get a biological sample dataset.
-From that dataset, several other functions can be run:
-
- *  age samples from sample_ages
- *  length samples from sample_lengths (TO BE WRITTEN)
- *  weight samples from sample_weights (TO BE WRITTEN)
- *  diet samples from sample_diet
-
-To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
+Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StdSurv
+LengthAgeCompTest.html).
 
@@
 
-# these are model specific, generalized above
-# if(initCCA) boxall <- c(0:88) 
-# if(initNEUS) boxall <- c(0:29)
-# if(initNOBA) boxall <- c(0:59)
-
-# should return all model output timesteps; need to generalize
-# if(initCCA) timeall <- c(0:100) 
-# if(initNEUS) timeall <- c(0:251)
-# if(initNOBA) timeall <- c(0:560)
-
 # generalized
@@
 
-# to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
-# if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
-# if(initNEUS) survspp <- funct.group.names[1:21]
-# if(initNOBA) survspp <- funct.group.names[1:36]
-
 # for length and age groups lets just do fish and sharks
@@
                            funct.groups$GroupType %in% c("FISH", "SHARK")]
-
-#if(initCCA) survspp <- survspp[!survspp %in% "Pisciv_T_Fish"]
 
@@
 
-```{r comptest2, message=FALSE, warning=FALSE}
+```{r comptest2, message=FALSE, warning=FALSE, echo=TRUE}
 # sample_fish uses an effN effective sample size for multinomial
@@
 Smaller biological sample sizes (~25 fish per year for nontrawl category) show more difference from the survey sampled age comp.
-
-## References

### 2019-05-06 visualization_Rmd/NOBAStdSurvCohortCompTest.Rmd
comp examples NOBA and NEUS

--- 2019-05-06 visualization_Rmd/NEUSStdSurvCohortCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate standard survey length and age comps from Northeast US Atlantis"
+title: "Testing atlantisom: generate standard survey length and age comps from Norwegian-Barents Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"

### 2019-05-06 visualization_Rmd/CCAStdSurvCohortCompTest.Rmd
CCA survey comp sampling

--- 2019-05-06 visualization_Rmd/NOBAStdSurvCohortCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate standard survey length and age comps from Norwegian-Barents Atlantis"
+title: "Testing atlantisom: generate standard survey length and age comps from California Current Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-And compare this to the true age comp above to see the impact of only standard survey sampling on age comps:
+And compare this to the true age comp above to see the impact of only standard survey sampling on age comps (CCA has empty dataset for last year, hardcoding last year-1):
 
@@
 comparecomps <- ggplot() +
-  geom_point(data=subset(dat2totN, time==annualmidyear[length(annualmidyear)]), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
-  geom_point(data=subset(comptestprop, time==annualmidyear[length(annualmidyear)]), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
+  geom_point(data=subset(dat2totN, time==annualmidyear[length(annualmidyear)]-1), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==annualmidyear[length(annualmidyear)]-1), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
   theme_tufte() +
@@
 
-We can compare true (cohort) age comp ("true"), survey sampled age comp with no subsampling effect ("samphigh"), and age comp from biological subsamling aboard the survey ("sampreal"):
+We can compare true (cohort) age comp ("true"), survey sampled age comp with no subsampling effect ("samphigh"), and age comp from biological subsamling aboard the survey ("sampreal") (CCA has empty dataset for last year, hardcoding last year-1):
 
@@
 comparecomps <- ggplot() +
-  geom_point(data=subset(dat2totN, time==annualmidyear[length(annualmidyear)]), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
-  geom_point(data=subset(comptestprop, time==annualmidyear[length(annualmidyear)]), aes(x=agecl, y=numAtAgesamp/totsamp, color="samphigh"), alpha = 0.3) +
-  geom_point(data=subset(comptestmixprop, time==annualmidyear[length(annualmidyear)]), aes(x=agecl, y=numAtAgesamp/totsamp, color="sampreal"), alpha = 0.3) +
+  geom_point(data=subset(dat2totN, time==annualmidyear[length(annualmidyear)]-1), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
+  geom_point(data=subset(comptestprop, time==annualmidyear[length(annualmidyear)]-1), aes(x=agecl, y=numAtAgesamp/totsamp, color="samphigh"), alpha = 0.3) +
+  geom_point(data=subset(comptestmixprop, time==annualmidyear[length(annualmidyear)]-1), aes(x=agecl, y=numAtAgesamp/totsamp, color="sampreal"), alpha = 0.3) +
   theme_tufte() +

### 2019-05-06 docs/index.Rmd
add comp testing to index

@@
 
+Standard survey with composition sampling: impact of selectivity, `sample_fish`
+
+<a href="https://sgaichas.github.io/poseidon-dev/StdSurvLengthAgeCompTest.html" target="_blank">Compare true cohort comp with survey + sampling - May 6, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NEUSStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison NEUS - May 6, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/CCAStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison CCA - May 6, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NOBAStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison NOBA - May 6, 2019</a>
+
 ### More to come later!

### 2019-05-08 TrueLengthAgeCompTest.Rmd
start corrections to length sampling

@@
 # function takes resn and nums as inputs so run survey and sample functions on resn as well
+
 if(initCCA) datresn <- CCAresults$resn
@@
 survey_testresnall <- create_survey(dat = datresn,
-                                 time = seq(0,251,5), # use timeall? census?
+                                 time = timeall, # use timeall? yes to match Nall
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
+#also needs structn, apply survey to this as well
+if(initCCA) datstructn <- CCAresults$structn
+if(initNEUS) datstructn <- NEUSresults$structn
+if(initNOBA) datstructn <- NOBAresults$structn
+
+survey_teststructnall <- create_survey(dat = datstructn,
+                                 time = timeall, # use timeall? yes to match Nall
                                  species = survspp,
@@
 lengthtesthigh <- sample_fish(survey_testNall, effNhigh)
-# use fish sampled from N along with survey of resn in function
-# calc_age2length uses nums (lengthtest) and finds matches in resn
+# use fish sampled from N along with survey of resn and structn in function
+# calc_age2length uses nums (lengthtest) and finds matches in resn and structn
 
@@
 
-length_census <- calc_age2length(structn = lengthtesthigh,
+length_census <- calc_age2length(structn = survey_teststructnall,
                                  resn = survey_testresnall,

### 2019-05-08 TrueLengthCompTest.Rmd
working through length comp sampling

--- 2019-05-08 TrueLengthAgeCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate census length and age comps"
+title: "Testing atlantisom: generate census length comps"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-## Simulate a survey part 3: sample for length and age composition
+## Simulate a survey part 4: sample for length composition
 
@@
 
+Atlantis outputs numbers by cohort (stage-age) and growth informtation, but does not output size of animals directly.
+The function we are testing here, `atlantisom::calc_age2length` converts numbers by cohort to a length composition.
+The [workflow originally envisioned](https://onedrive.live.com/?authkey=%21AFQkOoKRz64TLUw&cid=59547B4CB95EF108&id=59547B4CB95EF108%216291&parId=59547B4CB95EF108%216262&o=OneUp) was to create a survey, sample fish from the survey, then apply this function, but as written it may work better to create a length comp for the whole population, which could then be sampled.
+We'll work out the options here.
+
+### This stuff is biolerplate from other docs, repeated here for completeness: 
 To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
@@
 
-ageclcomp <- ggplot(dat2totN, aes(x=agecl, y=numAtAge/totN, col=time)) +
-  geom_point()
-
-ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
-ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
-ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
-ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
-#ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
+# ageclcomp <- ggplot(dat2totN, aes(x=agecl, y=numAtAge/totN, col=time)) +
+#   geom_point()
+# 
+# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+# #ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
 
@@
 
-```{r comptest1plot, message=FALSE, warning=FALSE}
+```{r comptest1plot, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
 
@@
 
-And they did for NEUS the first time I did this, with some exceptions.
+And they mostly do aside from fish that have 0s in the first year (Planktiv_T_Fish.
 
-The problems start when I try to do age to length, this function returns an error at present.
+The problems start when I try to do age to length, this function returns an error as written when `sample_fish` is run first.
 
-**Editing to be done in atlantisom (I have made local changes to calcage2length.R that are not pushed yet because I may have introduced an error)**
+May 8 update: Found problem, fix may require rethinking sampling flow. Making another rmd (this one!) to deal with this.
+
+So, test 1 is running `calc_age2length` with the full true dataset. This should work because there will be no aggregation of the output:
+
+```{r calcage2length-truth}
+
+if(initCCA) truth <- CCAresults
+if(initNEUS) truth <- NEUSresults
+if(initNOBA) truth <- NOBAresults
+
+length_census <- calc_age2length(structn = truth$structn,
+                                 resn = truth$resn,
+                                 nums = truth$nums,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+```
+
+This produces output. Visualize:
+
+```{r lengthcensus-vis}
+
+# for an example species, by polygon, layer because this is full resolution
+
+```
+
+So the question now is, do we run `create_survey` on the output of this, or do we `create_survey` on all three truth components first, then run `calc_age2length` to get a survey sample?
+This should work and will eliminate the unsampled areas/times/species.
+
+And then `sample_fish` from `calc_age2length` output?
+But also we need to run `sample_fish` to get the cohort (age) output (or do we?
+see below) so the multinomial in there will make different distributions for smaller sample sizes.
+Since we are trying to represent biological sampling as one set of sampled fish which we length, then take a subset of ages and weights, this may not work.
+
+Or, we can sum the length atoutput from calc_age2length over size bins and back in to agecl bins to get age classes, then subsample this for ages?
+We also have mulen and muweight for each size bin (or is it agecl bin?) so can we back the individual weights for the aged fish subsample out of this?
 
@@
 # function takes resn and nums as inputs so run survey and sample functions on resn as well
+# do we need to run the survey if calc_age2length does matching based on sample_fish?
 
@@
 
-lengthtesthigh <- sample_fish(survey_testNall, effNhigh)
+lengthtesthighN <- sample_fish(survey_testNall, effNhigh)
 # use fish sampled from N along with survey of resn and structn in function
 # calc_age2length uses nums (lengthtest) and finds matches in resn and structn
+
+# BUT: expecting raw nums from truth, not sample_fish nums with polygon already aggregated, so this never matches as written.
+
+# solutions, run sample_fish on structn and resn prior to input--*trying*
+# or don't sameple_fish on nums first, but I think we want to sample_fish first in the workflow, then convert that sample to lengths, ages, and weights.
+
+# unfortunately, if we are really sampling with low effN, these will be sampling cohorts with different probabilities across nums, structn, and resn, so do we need a length comp on the whole thing first and then sample it??
+
+lengthtesthighstructn <- sample_fish(survey_teststructnall, effNhigh)
+lengthtesthighresn <- sample_fish(survey_testresnall, effNhigh)
 

### 2019-05-08 TrueLengthCompTest.Rmd
try a survey census

@@
 
-```{r calcage2length-truth}
+```{r calcage2length-truth, warning=FALSE, message=FALSE, eval=FALSE}
 
@@
 
-This produces output. Visualize:
+This produces output, after a looooooong time. Its doing it for all species, which is a problem. Visualize a tiny bit:
 
@@
 # for an example species, by polygon, layer because this is full resolution
+# everyone's favorite predator
+
+atf_length_census <- length_census$natlength %>%
+  filter(species %in% ("Arrowtooth_flounder"))
 
@@
 So the question now is, do we run `create_survey` on the output of this, or do we `create_survey` on all three truth components first, then run `calc_age2length` to get a survey sample?
-This should work and will eliminate the unsampled areas/times/species.
+The latter should work and will eliminate the unsampled areas/times/species.
+
+It should save time to run the census survey to get true length comps, because at least it aggregates over polygon and layer. Test this:
+
+```{r censussurvey-calcage2length, warning=FALSE, message=FALSE}
+
+if(initCCA) truth <- CCAresults
+if(initNEUS) truth <- NEUSresults
+if(initNOBA) truth <- NOBAresults
+
+survey_testresnall <- create_survey(dat = truth$resn,
+                                 time = timeall, # use timeall? yes to match Nall
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
+survey_teststructnall <- create_survey(dat = truth$structn,
+                                 time = timeall, # use timeall? yes to match Nall
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
+# survey_testNall with truth$nums as dat was created above
+
+length_censussurv <- calc_age2length(structn = survey_teststructnall,
+                                 resn = survey_testresnall,
+                                 nums = survey_testNall,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+```
 

### 2019-05-08 TrueLengthCompTest.Rmd
save this, if it ever returns
trying for atf only

@@
 
-So, test 1 is running `calc_age2length` with the full true dataset. This should work because there will be no aggregation of the output:
+So, test 1 is running `calc_age2length` with the full true dataset.
+This should work because there will be no aggregation of the output (no, don't do this with the full dataset!
+example with one species instead):
 
-```{r calcage2length-truth, warning=FALSE, message=FALSE, eval=FALSE}
+```{r calcage2length-truth, warning=FALSE, message=FALSE}
 
@@
 
-length_census <- calc_age2length(structn = truth$structn,
-                                 resn = truth$resn,
-                                 nums = truth$nums,
+# length_census <- calc_age2length(structn = truth$structn,
+#                                  resn = truth$resn,
+#                                  nums = truth$nums,
+#                                  biolprm = truth$biolprm, fgs = truth$fgs,
+#                                  CVlenage = 0.1, remove.zeroes=TRUE)
+# 
+# # never got here with full truth
+# saveRDS(length_census, file.path(d.name, paste0(scenario.name, "length_census.rds")))
+
+# try for a single species
+# everyone's favorite predator
+atf_truestructn <- truth$structn[truth$structn$species == "Arrowtooth_flounder",]
+atf_trueresn <- truth$resn[truth$resn$species == "Arrowtooth_flounder",] 
+atf_truenums <- truth$nums[truth$nums$species == "Arrowtooth_flounder",]
+
+atf_length_census <- calc_age2length(structn = atf_truestructn,
+                                 resn = atf_trueresn,
+                                 nums = atf_truenums,
                                  biolprm = truth$biolprm, fgs = truth$fgs,
@@
 
-This produces output, after a looooooong time. Its doing it for all species, which is a problem. Visualize a tiny bit:
+This produced output for a single species by mistake.
+It ran for hours on the full truth dataset and did not complete (I stopped it).
+Its doing it for all species, which is a problem.
+Rewrote for just one species to see full length comp by area and depth layer.
+
+Visualize a tiny bit:
 
@@
 
-atf_length_census <- length_census$natlength %>%
-  filter(species %in% ("Arrowtooth_flounder"))
-
 ```
@@
 
-It should save time to run the census survey to get true length comps, because at least it aggregates over polygon and layer. Test this:
+Try again with survey sampling for nums, structn and resn, which limits the species list to fish and sharks, and also aggregates over layer and polygon. Test this:
 

### 2019-05-09 TrueLengthCompTest.Rmd
more refinements

@@
 This should work because there will be no aggregation of the output (no, don't do this with the full dataset!
-example with one species instead):
+example with one species instead.
+actually, don't even try this):
 
@@
 
-atf_length_census <- calc_age2length(structn = atf_truestructn,
-                                 resn = atf_trueresn,
-                                 nums = atf_truenums,
-                                 biolprm = truth$biolprm, fgs = truth$fgs,
-                                 CVlenage = 0.1, remove.zeroes=TRUE)
+#atf_length_census <- calc_age2length(structn = atf_truestructn,
+#                                 resn = atf_trueresn,
+#                                 nums = atf_truenums,
+#                                 biolprm = truth$biolprm, fgs = truth$fgs,
+#                                 CVlenage = 0.1, remove.zeroes=TRUE)
+atf_length_census <- lenout
 
@@
 This produced output for a single species by mistake.
-It ran for hours on the full truth dataset and did not complete (I stopped it).
-Its doing it for all species, which is a problem.
-Rewrote for just one species to see full length comp by area and depth layer.
+It ran for hours on the full truth dataset, and on the just arrowtooth full truth dataset and did not complete (I stopped it).
+Testing for just one species and 500 index steps (of 103690) to see full length comp by area and depth layer.
 
@@
 
+lfplot <- ggplot(atf_length_census$natlength, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte()
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1)
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2)
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3)
+
+lfplot + facet_wrap_paginate(~polygon, ncol=4, nrow = 4, page = 1)
+lfplot + facet_wrap_paginate(~polygon, ncol=4, nrow = 4, page = 2)
+lfplot + facet_wrap_paginate(~polygon, ncol=4, nrow = 4, page = 3)
+
 ```
@@
 
-Try again with survey sampling for nums, structn and resn, which limits the species list to fish and sharks, and also aggregates over layer and polygon. Test this:
+Try again with survey sampling for nums, structn and resn, which limits the species list to fish and sharks, and also aggregates over layer and polygon.
+This will still take too long.
+Can we run with the sample_fish in nums?
+Test this:
 
@@
 # survey_testNall with truth$nums as dat was created above
+
+#cut down to single species
+atf_survey_teststructnall <- survey_teststructnall[survey_teststructnall$species == "Arrowtooth_flounder",]
+atf_survey_testresnall <- survey_testresnall[survey_testresnall$species == "Arrowtooth_flounder",]
+atf_survey_testNall <- survey_testNall[survey_testNall$species == "Arrowtooth_flounder",]
 

### 2019-05-10 RethinkSamplingFunctions.Rmd
rethinking sampling

---
title: "Rethinking sampling functions to include weights and numbers"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction: the problem

The original design of [atlantisom](https://github.com/r4atlantis/atlantisom) conceived at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.had the following workflow:

1. apply `run_truth` to a set of Atlantis model outputs, which returns 'atoutput' components in the following units by species, agecl (Atlantis cohort, often in units of maxage/10), polygon (model box), layer (model depth), and time (model toutinc or outputstep):
    + `biomass_eaten$atoutput` is eaten biomass in t for each functional group?
        + `biomass_eaten$vol` is volume aggregated over layers
        + `biomass_eaten$prey` is the prey item eaten?
        + `biomass_eaten$dietcomp` is proportion in diet?
        + `biomass_eaten$bio_eaten` is different from atoutput but how?
    + `biomass_ages$atoutput` is  biomass in t from `calc_biomass_age`
    + `catch$atoutput` is fishery catch in numbers of individuals
    + `nums$atoutput` is numbers of individuals
    + `resn$atoutput` is Nitrogen reserve weight in mg/m^3
    + `structn$atoutput`is Nitrogen reserve weight in mg/m^3
2. additional components of the `run_truth` return (default name `result`) are key factors needed for conversions and interpreting the outputs (described elsewhere):
    + `biolprm`
    + `fgs`
3. apply `create_survey` to the result output of `run_truth`
4. apply `sample_survey_biomass` to the output of `create_survey` to get an index
    + designed to use nums output with an average weight--but estimating average weight proved problematic the way I did it, which was too imprecise
    + can be used directly on biomass_ages output with a vector of average weight=1 (see [True Bio Test](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html))
5. apply `sample_fish` to the output of `create_survey` to generate biological samples
6. apply `calc_age2length` to output of 'sample_fish` to generate length compositions
    + *FIRST BIG PROBLEM!* `calc_age2length` needs weight data at the same level of resolution as the nums data to estimate lengths using the length-weight relationship. 
    + Weight data comes from resn and structn. `calc_age2length` was originally tested with completely disaggregated outputs of `run_truth` and it generates length samples correctly that way
    + Initial `calc_age2length` testing with a real dataset revealed that it is entirely too slow to apply at the level of `run_truth` results for a full Atlantis model run and over many species (TrueLengthCompTest.Rmd)
    + Attempts to aggregate structn and resn results to the `create_survey` and `sample_fish` level were incorrect, because:
    + *we cannot apply functions designed for numbers output that aggregate based on sum to resn and structn outputs which are in units of N per volume*
    + So actually, this problem reaches all the way back to `create_survey` which is designed for nums output and aggregates using a sum:

```{r createsurvey}

knitr::include_graphics("https://dm2301files.storage.live.com/y4mSPipdQ7jyZNNfmpslxt7sCy7DqcQ0wMM40iLU1LLwfprFtKrjNJtHdibguUzNb9qY2vqluaHtpnKsUqkJ_wk3PpGCBtZmLGMdDSF1TVPsjeRWwtGjx8AIdNDiw9xvGmXf4aMnjXJ9DK9QW63JRZusH2uKgO1Mjd-8AEY554sLUG9xC9NFGaktffG2Ep5y8sz09stoHEM2pqG1I_oNtHGbg/CreateSurvey.png?psid=1&width=960&height=720")

```

The next three biological sampling steps after creating the length composition are taking a subset of the `sample_fish` for lengths using `sample_length` (though this could be 100% sample as is often done on a survey), taking a subset of the `sample_fish` for ages and applying ageing error using `sample_ages`, and finally taking a sample of individual fish weights, perhaps the same sample that was aged as is done on some surveys) using `sample_weight`. (So far only `sample_ages` is drafted and the other two functions remain to be written.)

```{r samplingworkflow}

knitr::include_graphics("https://dm2301files.storage.live.com/y4mV1y2g-JoSJFProk32x9SjWofLHwF_WtQc4Ipz8M2ZLc7hypg4yTYAgvB0IjwVLfjVw4fCNFP9l6lbWQIFI8NtzaIHjxgGEmptz9I9cB5jA5VUCJ_gJmSYSfI8_au29U_nuP5-zw-qkBBINifEo8RlJSvOEol5qu5IfI4BLk5YjjxTomMcbDVtOUBIJQHMEZjRUnlPhy7lV4x5ckvz5WfWA/SamplingWorkflow.png?psid=1&width=960&height=720")

```

## Proposed solution

I think we should keep the workflow designed at the Atlantis summit because this is the way surveys generally operate.

We need a way to appropriately aggregate the resn and structn outputs which are densitites. A mean or median across layers and polygons would be appropriate. The `run_truth` function calls `calc_biomass_age` which uses median to aggregate these outputs to generate the biomass_ages result component.

We can either rewrite `aggregateData` to take another argument indicating aggregation by sum or median, or create separate functions `aggregateDataSum` and `aggregateDataMedian` to be used on the numbers/biomass or density outputs, respectively.  Either way we need to ensure that the appropriate aggregation function is used in `create_survey` and `sample_fish`

Current `aggregateData` function, which is strangely not called by `sample_fish`

```{r aggregateData}

knitr::include_graphics("https://dm2301files.storage.live.com/y4mlnKwEPKB-fwO50OZng5CBgAwpz_T3Wc_K8BEdyLLAt-WnSs8Qd4YJY1i4hO0Zja9F6Vx1X3WiIDB1pFlBVW9YQDfKGvczcddyoZbl_bvUmPOsdGN4eHiAgyh_rKZ_Op286QWAPURty0QsJsNOhUNzBAc8326d8cKW9QmTlsncAJDDoU5wMdJgWRoalD939BG4Pg2SbyLxBe9tLfYm1xD_g/aggregateData.png?psid=1&width=960&height=720")

```

While we are at it, catch biomass is apparently calculated by `run_truth` but not exported, so I think we should export that all at once. Right now only catch in numbers is exported, but catch biomass is most often what management is based on and is an input to most assessment models.

For fisheries, the plan would be to run `create_fishery_subset` similar to `create_survey` and then `sample_fish` with the same functions as above to get the catch at length and catch at age samples.

Current `create_fishery_subset` function, as yet untested:

```{r createfishery}

knitr::include_graphics("https://dm2301files.storage.live.com/y4mZYTQWQxl_SZg2SQVNtRiv5OjAyv7ML3FqiVW2dY0SZmUlUO91Mm55pNAamQ-Kz7V_p2dH2UrEIGVttXXmiiN4L6HOBJTexTtqu43r9VH4IOdc-XPmsACvcnmzCrUIjqvr6NPJGwpds4_KxLEvlnfGJ3ZobYsxKwIUhObm6Uyq2deLPjZUuMT06YTELfTWbZBcDIR42waSKhxpUKlQXiaIw/CreateFisherySubset.png?psid=1&width=960&height=720")

```

## To Do List:

1. write an aggregation function for resn and structn that applies the median
2. rewrite `create_survey` to use alternate aggregation functions, or write `create_survey_nums` and `create_survey_wt` functions that use appropriate aggregation
3. rewrite `sample_fish` to use alternate aggregation functions for each input type. Keeping this all in one function makes sense because the functional already takes both numbers and weights as inputs

### 2019-05-10 TrueLengthCompTest.Rmd
rethinking sampling

@@
 
-```{r calcage2length-truth, warning=FALSE, message=FALSE}
+```{r calcage2length-truth, warning=FALSE, message=FALSE, eval=FALSE}
 
@@
 
-```{r lengthcensus-vis}
+```{r lengthcensus-vis, eval=FALSE}
 
@@
 ```
+
+This works, but I can't show you because I ran it line by line from the function.
 
@@
 
-#cut down to single species
+#cut down to single species?
 atf_survey_teststructnall <- survey_teststructnall[survey_teststructnall$species == "Arrowtooth_flounder",]
@@
 
-length_censussurv <- calc_age2length(structn = survey_teststructnall,
-                                 resn = survey_testresnall,
-                                 nums = survey_testNall,
+# still a lot
+# if we apply the highest effN to structn and resn using sample fish, we should have the right atoutput for each agecl to apply to possibly subsampled nums.
+# as above, apply sample_fish to resnall and structnall
+# comptesthigh <- sample_fish(survey_testNall, effNhigh), but I renamed atoutput so do over without that
+
+numsallhigh <- sample_fish(survey_testNall, effNhigh)
+
+#dont sample these, just aggregate them
+structnall <- sample_fish(survey_teststructnall, effNhigh, sample = FALSE)
+
+resnall <-  sample_fish(survey_testresnall, effNhigh, sample = FALSE)
+
+# now cut these down to a single species for testing
+# this should still represent a census but with polygon and layer aggregated
+
+atf_numsallhigh <- numsallhigh[numsallhigh$species == "Arrowtooth_flounder",]
+atf_structnall <- structnall[structnall$species == "Arrowtooth_flounder",]
+atf_resnall <- resnall[resnall$species == "Arrowtooth_flounder",]
+
+atf_length_censussurvsamp <- calc_age2length(structn = atf_structnall,
+                                 resn = atf_resnall,
+                                 nums = atf_numsallhigh,
                                  biolprm = truth$biolprm, fgs = truth$fgs,
@@
 
-And then `sample_fish` from `calc_age2length` output?
-But also we need to run `sample_fish` to get the cohort (age) output (or do we?
-see below) so the multinomial in there will make different distributions for smaller sample sizes.
-Since we are trying to represent biological sampling as one set of sampled fish which we length, then take a subset of ages and weights, this may not work.
+Trying to plot just this sample, ran very quickly. And this didn't work, aggregating structural and residual N is definitely wrong.
 
-Or, we can sum the length atoutput from calc_age2length over size bins and back in to agecl bins to get age classes, then subsample this for ages?
-We also have mulen and muweight for each size bin (or is it agecl bin?) so can we back the individual weights for the aged fish subsample out of this?
+```{r atflengthsamp1-test}
 
-```{r calcage2length-breaks, eval=FALSE}
-# now turn this into length comp with calc_age2length
-# function takes resn and nums as inputs so run survey and sample functions on resn as well
-# do we need to run the survey if calc_age2length does matching based on sample_fish?
+lfplot <- ggplot(atf_length_censussurvsamp$natlength, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte()
 
-if(initCCA) datresn <- CCAresults$resn
-if(initNEUS) datresn <- NEUSresults$resn
-if(initNOBA) datresn <- NOBAresults$resn
-
-survey_testresnall <- create_survey(dat = datresn,
-                                 time = timeall, # use timeall? yes to match Nall
-                                 species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-#also needs structn, apply survey to this as well
-if(initCCA) datstructn <- CCAresults$structn
-if(initNEUS) datstructn <- NEUSresults$structn
-if(initNOBA) datstructn <- NOBAresults$structn
-
-survey_teststructnall <- create_survey(dat = datstructn,
-                                 time = timeall, # use timeall? yes to match Nall
-                                 species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-lengthtesthighN <- sample_fish(survey_testNall, effNhigh)
-# use fish sampled from N along with survey of resn and structn in function
-# calc_age2length uses nums (lengthtest) and finds matches in resn and structn
-
-# BUT: expecting raw nums from truth, not sample_fish nums with polygon already aggregated, so this never matches as written.
-
-# solutions, run sample_fish on structn and resn prior to input--*trying*
-# or don't sameple_fish on nums first, but I think we want to sample_fish first in the workflow, then convert that sample to lengths, ages, and weights.
-
-# unfortunately, if we are really sampling with low effN, these will be sampling cohorts with different probabilities across nums, structn, and resn, so do we need a length comp on the whole thing first and then sample it??
-
-lengthtesthighstructn <- sample_fish(survey_teststructnall, effNhigh)
-lengthtesthighresn <- sample_fish(survey_testresnall, effNhigh)
-
-# provides ouput for an individual species and time slice
-# breaks for whole dataset (NA output all rows)
-
-if(initCCA) truth <- CCAresults
-if(initNEUS) truth <- NEUSresults
-if(initNOBA) truth <- NOBAresults
-
-length_census <- calc_age2length(structn = survey_teststructnall,
-                                 resn = survey_testresnall,
-                                 nums = lengthtesthigh,
-                                 biolprm = truth$biolprm, fgs = truth$fgs,
-                                 CVlenage = 0.1, remove.zeroes=TRUE)
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+# lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+# lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
+# lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
+# lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
+# lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
 
@@
 
+## Back to the drawing board!
+
+Need a couple more functions to do this correctly, see how I am [rethinking](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html).
+
 ## References

### 2019-05-10 docs/index.Rmd
rethinking sampling

@@
 Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
+
+### Read this first on overall workflow
+
+<a href="https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html" target="_blank">Big picture and new functions needed - May 10, 2019</a>
 
@@
 
+### Length compositions
+
+Difficult when you don't have a good idea of the outputs you are using!
+
+<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html" target="_blank">Defeated by length compositions - May 8-9, 2019</a>
+
 ### More to come later!

### 2019-05-10 RethinkSamplingFunctions.Rmd
fix typo

@@
 5. apply `sample_fish` to the output of `create_survey` to generate biological samples
-6. apply `calc_age2length` to output of 'sample_fish` to generate length compositions
+6. apply `calc_age2length` to output of `sample_fish` to generate length compositions
     + *FIRST BIG PROBLEM!* `calc_age2length` needs weight data at the same level of resolution as the nums data to estimate lengths using the length-weight relationship. 

### 2019-05-10 RethinkSamplingFunctions.Rmd
fix another typo

@@
     + `resn$atoutput` is Nitrogen reserve weight in mg/m^3
-    + `structn$atoutput`is Nitrogen reserve weight in mg/m^3
+    + `structn$atoutput`is Nitrogen structural weight in mg/m^3
 2. additional components of the `run_truth` return (default name `result`) are key factors needed for conversions and interpreting the outputs (described elsewhere):

### 2019-05-10 RethinkSamplingFunctions.Rmd
clarify use of create survey

@@
 
-Current `aggregateData` function, which is strangely not called by `sample_fish`
+Current `aggregateData` function, which is strangely not called by `sample_fish`. Or not so strange, since `aggregateData` defaults to aggregate only over layer and maintain polygons in the dataset?
 
@@
 2. rewrite `create_survey` to use alternate aggregation functions, or write `create_survey_nums` and `create_survey_wt` functions that use appropriate aggregation
+    + the latter may be better, because do we want to apply efficiency and selectivity to median weight densities? I don't think so?
+    + the fish still weighs the same, we just don't have the right proportion in the ecosystem (efficiency) or the right proportion at age (selectivty), both of which are taken care of in the numbers.
+    + so maybe `aggregateDensityData` using the standard survey polygons gets us what we need as input to `sample_fish`?
 3.

### 2019-05-10 TrueLengthCompTest.Rmd
restart with new config

@@
 
-if(initCCA){
-  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
-  functional.groups.file <- "CalCurrentV3Groups.csv"
-  biomass.pools.file <- "DIVCalCurrentV3_BIOL.nc"
-  biol.prm.file <- "CalCurrentV3_Biol.prm"
-  box.file <- "CalCurrentV3_utm.bgm"
-  initial.conditions.file <- "DIVCalCurrentV3_BIOL.nc"
-  run.prm.file <- "CalCurrentV3_run.xml"
-  scenario.name <- "CCV3"
-}
+if(initCCA) source(here("config/CCConfig.R"))
 
-if(initNEUS){
-  d.name <- here("atlantisoutput","NEUStest20160303")
-  functional.groups.file <- "NeusGroups.csv" 
-  biomass.pools.file <- ""
-  biol.prm.file <- "at_biol_neus_v15_DE.prm"
-  box.file <- "neus30_2006.bgm"
-  initial.conditions.file <- "inneus_2012.nc"
-  run.prm.file <- "at_run_neus_v15_DE.xml"
-  scenario.name <- "neusDynEffort_Test1_"
-}
+if(initNEUS) source(here("config/NEUSConfig.R"))
 
-if(initNOBA){
-  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
-  functional.groups.file <- "nordic_groups_v04.csv" 
-  biomass.pools.file <- "nordic_biol_v23.nc"
-  biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
-  box.file <- "Nordic02.bgm"
-  initial.conditions.file <- "nordic_biol_v23.nc"
-  run.prm.file <- "nordic_run_v01.xml"
-  scenario.name <- "nordic_runresults_01"
-}
-# NOBA note: output filenames in CCA and NEUS begin with "output" and the run_truth function is written to expect this.
-Need to check if default Atlantis output file nomenclature has changed or if NOBA is a special case.
-For now, NOBA filenames have been changed to include prefix "output"
+if(initNOBA) source(here("config/NOBAConfig.R"))
+
 ```
@@
 if(initCCA) {
-  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
   truth.file <- "outputCCV3run_truth.RData"
   load(file.path(d.name, truth.file))
-  CCAresults <- result
+  truth <- result
 }
@@
 if(initNEUS) {
-  d.name <- here("atlantisoutput","NEUStest20160303")
   truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
   load(file.path(d.name, truth.file))
-  NEUSresults <- result
+  truth <- result
 }
@@
 if(initNOBA){
-  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
   truth.file <- "outputnordic_runresults_01run_truth.RData" 
   load(file.path(d.name, truth.file))
-  NOBAresults <- result
-}
-
-```
-
-We can also read in previously generated survey census files based on true biomass results for comparison if necessary, but not yet used here (see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).)
-
-```{r readsurvey1, eval=FALSE}
-
-if(initCCA) {
-  CCAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
-}
-
-if(initNEUS) {
-  NEUSsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
-}
-
-if(initNOBA) {
-  NOBAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
+  truth <- result
 }
@@
 The function we are testing here, `atlantisom::calc_age2length` converts numbers by cohort to a length composition.
-The [workflow originally envisioned](https://onedrive.live.com/?authkey=%21AFQkOoKRz64TLUw&cid=59547B4CB95EF108&id=59547B4CB95EF108%216291&parId=59547B4CB95EF108%216262&o=OneUp) was to create a survey, sample fish from the survey, then apply this function, but as written it may work better to create a length comp for the whole population, which could then be sampled.
-We'll work out the options here.
+The [workflow originally envisioned](https://onedrive.live.com/?authkey=%21AFQkOoKRz64TLUw&cid=59547B4CB95EF108&id=59547B4CB95EF108%216291&parId=59547B4CB95EF108%216262&o=OneUp) was to create a survey, sample fish from the survey, then apply this function.
+We determined that it will not work better to create a length comp for the whole population, which could then be sampled.
+We rethought what we needed [here](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html) and test the new functions in this document.
 

### 2019-05-10 TrueLengthCompTest.Rmd
clean up and try again

@@
 # should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
-# BUT CHECK if newer Atlantis models can do age-specific outputs
 selex1 <- data.frame(species=rep(funct.group.names, each=10),
@@
 
-# these are model specific, generalized above
-# if(initCCA) boxall <- c(0:88) 
-# if(initNEUS) boxall <- c(0:29)
-# if(initNOBA) boxall <- c(0:59)
-
-# should return all model output timesteps; need to generalize
-# if(initCCA) timeall <- c(0:100) 
-# if(initNEUS) timeall <- c(0:251)
-# if(initNOBA) timeall <- c(0:560)
-
-# generalized
+# generalized timesteps all models
 runpar <- load_runprm(d.name, run.prm.file)
@@
 
-#if(initCCA) survspp <- survspp[!survspp %in% "Pisciv_T_Fish"]
-
 ```
 
-Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age and length composition.
+Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition.
 
@@
 
-# this uses result$nums, but we are not creating a biomass index this time, so we don't need a weight at age conversion
-
-if(initCCA) datN <- CCAresults$nums
-if(initNEUS) datN <- NEUSresults$nums
-if(initNOBA) datN <- NOBAresults$nums
-
-survey_testNall <- create_survey(dat = datN,
+survey_testNall <- create_survey(dat = truth$nums,
                                  time = timeall,
@@
 
-#dat<-survey_testNall
-#dat2 <- aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time),sum)
-#names(dat2) <- c("species","agecl","time","numAtAge")
-
 totN <- dat2 %>%
@@
   summarize(totN = sum(numAtAge))
-
-#totN <- aggregate(dat2$numAtAge,list(dat2$species,dat2$time),sum )
-#names(totN) <- c("species","time","totN")
 
@@
 # ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
-# #ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
 
@@
 
-Then we use the `sample_fish` function and compare to true annual age comp calculated above as a test. These should match.
+Then we use the `sample_fish` function with very high effN and compare to true annual age comp calculated above as a test, which matches as we have shown [here](https://sgaichas.github.io/poseidon-dev/TrueLengthAgeCompTest.html).
 
-```{r comptest1, message=FALSE, warning=FALSE}
-# setting the effN higher than actual numbers results in sampling all
-effNall <- data.frame(species=survspp, effN=rep(1e+15, length(survspp)))
-# rmultinom broke with that sample size
-#comptestall <- sample_fish(survey_testNall, effNall)
-#names(comptestall) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")
+However. We need both numbers and weights to get the length comp, and ultimately the weight at age for use in assessments.
+
+After using functions without thinking first and then [rethinking sampling flow](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html), we will now test new functons to aggregate both densities and numbers appropriately for input into the biological sampling workflow.
+
+What we want is to subset the nums, resn, and structn components of run_truth using the same create_survey design selecting boxes, time, and species.
+We need only apply survey efficiency and selectivity to nums, however.
+So perhaps the new aggregateDensityData function can be applied to resn and structn?
+
+```{r aggdens}
+
+# aggregate true resn per survey design
+survey_aggresnall <- aggregateDensityData(dat = truth$resn,
+                                 time = timeall,
+                                 species = survspp,
+                                 boxes = boxall)
+
+# aggregate true structn per survey design
+survey_aggstructnall <- aggregateDensityData(dat = truth$structn,
+                                 time = timeall,
+                                 species = survspp,
+                                 boxes = boxall)
+
+```
+
+Now we should have inputs to `sample_fish` on the same scale, and they need to be aggregated across boxes into a single biological sample for the whole survey.
+We are not maintaining spatial structure in sampling because it isn't used in most assessments.
+
+To do the proper aggregation and not apply the multinomial sampling to the density data, I rewrite `sample_fish` to apply the median if `sample=FALSE` in the function call. Test this:
+
+```{r censussurvey-calcage2length, warning=FALSE, message=FALSE}
 
@@
 
-comptesthigh <- sample_fish(survey_testNall, effNhigh)
-names(comptesthigh) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")
-
-comptesttot <- aggregate(comptesthigh$numAtAgesamp,list(comptesthigh$species,comptesthigh$time),sum )
-names(comptesttot) <- c("species","time","totsamp")
-
-comptestprop <- merge(comptesthigh, comptesttot)
-```
-
-```{r comptest1plot, message=FALSE, warning=FALSE, echo=FALSE, eval=FALSE}
-
-# compare individual years, these proportions at age should match
-comparecomps <- ggplot() +
-  geom_point(data=subset(dat2totN, time==min(timeall)), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
-  geom_point(data=subset(comptestprop, time==min(timeall)), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3) +
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=paste0(scenario.name, " start"))
-
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
-#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
-
-#ggsave("censuscomposition_time0.png", width=11, height=11)
-
-comparecomps <- ggplot() +
-  geom_point(data=subset(dat2totN, time==median(timeall)), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
-  geom_point(data=subset(comptestprop, time==median(timeall)), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=paste0(scenario.name, " midpoint"))
-
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
-#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
-
-#ggsave("censuscomposition_time100.png", width=11, height=11)
-
-comparecomps <- ggplot() +
-  geom_point(data=subset(dat2totN, time==max(timeall)-1), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
-  geom_point(data=subset(comptestprop, time==max(timeall)-1), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=paste0(scenario.name, " end"))
-
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
-comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
-#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")
-
-#ggsave("censuscomposition_time250.png", width=11, height=11)
-
-# and they do
-# differences for baleen and toothed whales are due to small total numbers
-
-```
-
-And they mostly do aside from fish that have 0s in the first year (Planktiv_T_Fish.
-
-The problems start when I try to do age to length, this function returns an error as written when `sample_fish` is run first.
-
-May 8 update: Found problem, fix may require rethinking sampling flow. Making another rmd (this one!) to deal with this.
-
-So, test 1 is running `calc_age2length` with the full true dataset.
-This should work because there will be no aggregation of the output (no, don't do this with the full dataset!
-example with one species instead.
-actually, don't even try this):
-
-```{r calcage2length-truth, warning=FALSE, message=FALSE, eval=FALSE}
-
-if(initCCA) truth <- CCAresults
-if(initNEUS) truth <- NEUSresults
-if(initNOBA) truth <- NOBAresults
-
-# length_census <- calc_age2length(structn = truth$structn,
-#                                  resn = truth$resn,
-#                                  nums = truth$nums,
-#                                  biolprm = truth$biolprm, fgs = truth$fgs,
-#                                  CVlenage = 0.1, remove.zeroes=TRUE)
-# 
-# # never got here with full truth
-# saveRDS(length_census, file.path(d.name, paste0(scenario.name, "length_census.rds")))
-
-# try for a single species
-# everyone's favorite predator
-atf_truestructn <- truth$structn[truth$structn$species == "Arrowtooth_flounder",]
-atf_trueresn <- truth$resn[truth$resn$species == "Arrowtooth_flounder",] 
-atf_truenums <- truth$nums[truth$nums$species == "Arrowtooth_flounder",]
-
-#atf_length_census <- calc_age2length(structn = atf_truestructn,
-#                                 resn = atf_trueresn,
-#                                 nums = atf_truenums,
-#                                 biolprm = truth$biolprm, fgs = truth$fgs,
-#                                 CVlenage = 0.1, remove.zeroes=TRUE)
-atf_length_census <- lenout
-
-```
-
-This produced output for a single species by mistake.
-It ran for hours on the full truth dataset, and on the just arrowtooth full truth dataset and did not complete (I stopped it).
-Testing for just one species and 500 index steps (of 103690) to see full length comp by area and depth layer.
-
-Visualize a tiny bit:
-
-```{r lengthcensus-vis, eval=FALSE}
-
-# for an example species, by polygon, layer because this is full resolution
-# everyone's favorite predator
-
-lfplot <- ggplot(atf_length_census$natlength, aes(upper.bins)) +
-  geom_bar(aes(weight = atoutput)) +
-  theme_tufte()
-
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1)
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2)
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3)
-
-lfplot + facet_wrap_paginate(~polygon, ncol=4, nrow = 4, page = 1)
-lfplot + facet_wrap_paginate(~polygon, ncol=4, nrow = 4, page = 2)
-lfplot + facet_wrap_paginate(~polygon, ncol=4, nrow = 4, page = 3)
-
-```
-
-This works, but I can't show you because I ran it line by line from the function.
-
-So the question now is, do we run `create_survey` on the output of this, or do we `create_survey` on all three truth components first, then run `calc_age2length` to get a survey sample?
-The latter should work and will eliminate the unsampled areas/times/species.
-
-Try again with survey sampling for nums, structn and resn, which limits the species list to fish and sharks, and also aggregates over layer and polygon.
-This will still take too long.
-Can we run with the sample_fish in nums?
-Test this:
-
-```{r censussurvey-calcage2length, warning=FALSE, message=FALSE}
-
-if(initCCA) truth <- CCAresults
-if(initNEUS) truth <- NEUSresults
-if(initNOBA) truth <- NOBAresults
-
-survey_testresnall <- create_survey(dat = truth$resn,
-                                 time = timeall, # use timeall? yes to match Nall
-                                 species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-survey_teststructnall <- create_survey(dat = truth$structn,
-                                 time = timeall, # use timeall? yes to match Nall
-                                 species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-# survey_testNall with truth$nums as dat was created above
-
-#cut down to single species?
-atf_survey_teststructnall <- survey_teststructnall[survey_teststructnall$species == "Arrowtooth_flounder",]
-atf_survey_testresnall <- survey_testresnall[survey_testresnall$species == "Arrowtooth_flounder",]
-atf_survey_testNall <- survey_testNall[survey_testNall$species == "Arrowtooth_flounder",]
-
-# still a lot
-# if we apply the highest effN to structn and resn using sample fish, we should have the right atoutput for each agecl to apply to possibly subsampled nums.
-# as above, apply sample_fish to resnall and structnall
-# comptesthigh <- sample_fish(survey_testNall, effNhigh), but I renamed atoutput so do over without that
-
+# apply default sample fish as before to get numbers
 numsallhigh <- sample_fish(survey_testNall, effNhigh)
 
-#dont sample these, just aggregate them
-structnall <- sample_fish(survey_teststructnall, effNhigh, sample = FALSE)
+#dont sample these, just aggregate them using median
+structnall <- sample_fish(survey_aggstructnall, effNhigh, sample = FALSE)
 
-resnall <-  sample_fish(survey_testresnall, effNhigh, sample = FALSE)
+resnall <-  sample_fish(survey_aggresnall, effNhigh, sample = FALSE)
 
@@
 
-Trying to plot just this sample, ran very quickly. And this didn't work, aggregating structural and residual N is definitely wrong.
+Maybe this will work, we can hope.
 
@@
 
-## Back to the drawing board!
+## Is this working??
 
-Need a couple more functions to do this correctly, see how I am [rethinking](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html).
+... the suspense is killing me
 

### 2019-05-10 TrueLengthCompTest.Rmd
add full sample, but neglect to save it

@@
 
-Maybe this will work, we can hope.
+Maybe this will work, we can hope. It looks like it did...
 
@@
 lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
-# lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
-# lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
-# lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
-# lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
-# lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
 
@@
 
-... the suspense is killing me
+OK, maybe we have something here. Next I need to see how long it takes to run for all species in the survey.
+
+```{r allspplength}
+
+length_censussurvsamp <- calc_age2length(structn = structnall,
+                                 resn = resnall,
+                                 nums = numsallhigh,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+```
 

### 2019-05-11 TrueLengthCompTest.Rmd
adjustments for the next run
it worked!

@@
 
+# try dir = "v" option for vertical lf comparisons
+
 ```
@@
 
-```{r allspplength}
+```{r allspplength, eval=FALSE}
 
@@
 
+#save for later use, takes a long time to generate
+saveRDS(length_censussurvsamp, file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))
+
 ```
 
+Answer: this ran for 5.5 hours. But it was really only 5 hours 20 minutes that was the last code block! Saving the file is highly recommended.
+
 ## References

### 2019-05-11 TrueLengthCompTest.Rmd
final text adjustments

@@
 
-To do the proper aggregation and not apply the multinomial sampling to the density data, I rewrite `sample_fish` to apply the median if `sample=FALSE` in the function call. Test this:
+To do the proper aggregation and not apply the multinomial sampling to the density data, I rewrote `sample_fish` to apply the median if `sample=FALSE` in the function call. Test this:
 

### 2019-05-11 docs/index.Rmd
Update index.Rmd
add fail filename
add new length comp file

@@
 <a href="https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html" target="_blank">Big picture and new functions needed - May 10, 2019</a>
+
+May 11: Changes to `atlantisom::sample_fish` and new `atlantisom::aggregateDensityData` functions seem to work. See below under [Length compositions].
 
@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html" target="_blank">Defeated by length compositions - May 8-9, 2019</a>
+<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest_fail.html" target="_blank">Defeated by length compositions - May 8-9, 2019</a>
+
+Better when you do:
+
+<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html" target="_blank">Victory: length compositions - May 11, 2019</a>
 

### 2019-05-11 visualization_Rmd/CCATrueLengthCompTest.Rmd
length comp vis CCA

--- 2019-05-11 TrueLengthCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate census length comps"
+title: "Testing atlantisom: generate census length comps for California Current Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-This section uses the `atlantisom::create_survey()` and `atlantisom::sample_fish()` to get a biological sample dataset. From that dataset, several other functions can be run:
-
- *  age samples from sample_ages
- *  length samples from sample_lengths (TO BE WRITTEN)
- *  weight samples from sample_weights (TO BE WRITTEN)
- *  diet samples from sample_diet
-
-Atlantis outputs numbers by cohort (stage-age) and growth informtation, but does not output size of animals directly.
-The function we are testing here, `atlantisom::calc_age2length` converts numbers by cohort to a length composition.
-The [workflow originally envisioned](https://onedrive.live.com/?authkey=%21AFQkOoKRz64TLUw&cid=59547B4CB95EF108&id=59547B4CB95EF108%216291&parId=59547B4CB95EF108%216262&o=OneUp) was to create a survey, sample fish from the survey, then apply this function.
-We determined that it will not work better to create a length comp for the whole population, which could then be sampled.
-We rethought what we needed [here](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html) and test the new functions in this document.
-
-### This stuff is biolerplate from other docs, repeated here for completeness: 
-To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
+Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).
 
@@
 
-```{r stdsurveyNbased}
+```{r stdsurveyNbased, echo=TRUE}
 
@@
 ```
-
-Next, get true annual (cohort) age comp from this census survey based on run truth. (is there a standard Atlantis output I can compare this to as we did for biomass?)
 
@@
 
-However. We need both numbers and weights to get the length comp, and ultimately the weight at age for use in assessments.
+The following gets both numbers and weights to get the length comp, and ultimately the weight at age for use in assessments.
 
-After using functions without thinking first and then [rethinking sampling flow](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html), we will now test new functons to aggregate both densities and numbers appropriately for input into the biological sampling workflow.
+First we subset the nums, resn, and structn components of `run_truth` using the same `create_survey` design selecting boxes, time, and species.
+We need only apply survey efficiency and selectivity to nums, however (done above), so we apply `aggregateDensityData` here.
 
-What we want is to subset the nums, resn, and structn components of run_truth using the same create_survey design selecting boxes, time, and species.
-We need only apply survey efficiency and selectivity to nums, however.
-So perhaps the new aggregateDensityData function can be applied to resn and structn?
-
-```{r aggdens}
+```{r aggdens, echo=TRUE}
 
@@
 
-To do the proper aggregation and not apply the multinomial sampling to the density data, I rewrote `sample_fish` to apply the median if `sample=FALSE` in the function call. Test this:
+To do the proper aggregation and not apply the multinomial sampling to the density data, I rewrote `sample_fish` to apply the median if `sample=FALSE` in the function call.
 
-```{r censussurvey-calcage2length, warning=FALSE, message=FALSE}
+```{r censussurvey-samplefish, warning=FALSE, message=FALSE, echo=TRUE}
 
@@
 
-# now cut these down to a single species for testing
+```
+
+Select a single species for testing. Generating length composition for a single species should take just a few minutes to run.
+
+```{r singlespp-lengthcomp, echo=TRUE}
+
 # this should still represent a census but with polygon and layer aggregated
+# WARNING: selecting a species by name is hardcoded for a particular model
 
-atf_numsallhigh <- numsallhigh[numsallhigh$species == "Arrowtooth_flounder",]
-atf_structnall <- structnall[structnall$species == "Arrowtooth_flounder",]
-atf_resnall <- resnall[resnall$species == "Arrowtooth_flounder",]
+ss_numsallhigh <- numsallhigh[numsallhigh$species == "Yelloweye_rockfish",]
+ss_structnall <- structnall[structnall$species == "Yelloweye_rockfish",]
+ss_resnall <- resnall[resnall$species == "Yelloweye_rockfish",]
 
-atf_length_censussurvsamp <- calc_age2length(structn = atf_structnall,
-                                 resn = atf_resnall,
-                                 nums = atf_numsallhigh,
+ss_length_censussurvsamp <- calc_age2length(structn = ss_structnall,
+                                 resn = ss_resnall,
+                                 nums = ss_numsallhigh,
                                  biolprm = truth$biolprm, fgs = truth$fgs,
@@
 
-Maybe this will work, we can hope. It looks like it did...
+This gives annual length comps:
 
-```{r atflengthsamp1-test}
+```{r ss-lengthsamp1-test}
 
-lfplot <- ggplot(atf_length_censussurvsamp$natlength, aes(upper.bins)) +
+lfplot <- ggplot(ss_length_censussurvsamp$natlength, aes(upper.bins)) +
   geom_bar(aes(weight = atoutput)) +
@@
 
-## Is this working??
+## Run for all species then SAVE
 
-OK, maybe we have something here. Next I need to see how long it takes to run for all species in the survey.
+This code block can take hours (5+ for CCA) to run, so be prepared, but then the saved file can be read back in to have length comps for all surveyed species.
 
-```{r allspplength, eval=FALSE}
+```{r allspplength, eval=FALSE, echo=TRUE}
 
@@
 
-Answer: this ran for 5.5 hours. But it was really only 5 hours 20 minutes that was the last code block! Saving the file is highly recommended.
+Read it back in:
+
+```{r loadlengthcomp, echo=TRUE}
+
+length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))
+
+```
+
+Demo plots for other species:
+
+```{r moreplots, echo=TRUE}
+
+ss_length_censussurvsamp <- length_censussurvsamp[length_censussurvsamp$species == "Herring",]
+
+lfplot <- ggplot(ss_length_censussurvsamp$natlength, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte()
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
+
+```
+
+So this looks like our best approximation of a true length comp for a particular Atlantis run. Comparisons with standard surveys next.
 

### 2019-05-11 visualization_Rmd/NOBATrueLengthCompTest.Rmd
first try NOBA length comps

--- 2019-05-11 visualization_Rmd/CCATrueLengthCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate census length comps for California Current Atlantis"
+title: "Testing atlantisom: generate census length comps for Norwegian Barents Sea Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-This page has visualizations for the CCA model example, Atlantis Summit Common Scenario 1. For full explanation of methods, see the file linked at the beginning of each section.
+This page has visualizations for the NOBA model example, CERES Global Sustainability. For full explanation of methods, see the file linked at the beginning of each section.
 
@@
 
-initCCA <- TRUE
+initCCA <- FALSE
 initNEUS <- FALSE
-initNOBA <- FALSE
+initNOBA <- TRUE
 
@@
 
-ss_numsallhigh <- numsallhigh[numsallhigh$species == "Yelloweye_rockfish",]
-ss_structnall <- structnall[structnall$species == "Yelloweye_rockfish",]
-ss_resnall <- resnall[resnall$species == "Yelloweye_rockfish",]
+ss_numsallhigh <- numsallhigh[numsallhigh$species == "Green_halibut",]
+ss_structnall <- structnall[structnall$species == "Green_halibut",]
+ss_resnall <- resnall[resnall$species == "Green_halibut",]
 
@@
 
-```{r allspplength, eval=FALSE, echo=TRUE}
+```{r allspplength, echo=TRUE}
 
@@
 
-ss_length_censussurvsamp <- length_censussurvsamp[length_censussurvsamp$species == "Herring",]
+ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
+  filter(species == "Norwegian_ssh")
 
-lfplot <- ggplot(ss_length_censussurvsamp$natlength, aes(upper.bins)) +
+lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
   geom_bar(aes(weight = atoutput)) +
@@
 So this looks like our best approximation of a true length comp for a particular Atlantis run. Comparisons with standard surveys next.
-
-## References

### 2019-05-13 visualization_Rmd/NOBATrueLengthCompTest.Rmd
update with last three code blocks not evaluated

@@
 
-Select a single species for testing. Generating length composition for a single species should take just a few minutes to run.
+Select a single species for testing.
+Generating length composition for a single species should take just a few minutes to run.
+Update: for NOBA with many output timesteps, even a single species took 15-20 minutes.
 
@@
 
-This code block can take hours (5+ for CCA) to run, so be prepared, but then the saved file can be read back in to have length comps for all surveyed species.
+This code block can take hours (5+ for CCA and 18+ for NOBA) to run, so be prepared, but then the saved file can be read back in to have length comps for all surveyed species.
+For a normal survey query we would have only one output timestep per year, which should save some time.
 
-```{r allspplength, echo=TRUE}
+```{r allspplength, echo=TRUE, eval=FALSE}
 
@@
 
-```{r loadlengthcomp, echo=TRUE}
+```{r loadlengthcomp, echo=TRUE, eval=FALSE}
 
@@
 
-```{r moreplots, echo=TRUE}
+```{r moreplots, echo=TRUE, eval=FALSE}
 

### 2019-05-13 visualization_Rmd/NOBATrueLengthCompTest.Rmd
change NOBA to annual output

@@
 
+# not enough memory to do all NOBA outputsteps, change to once per year
+midptyr <- round(median(seq(1,stepperyr)))
+
+annualmidyear <- seq(midptyr, noutsteps, stepperyr)
+
 # define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
@@
 
-Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition.
+Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition. *Sampling once per year, rather than output timestep.*
 
@@
 survey_testNall <- create_survey(dat = truth$nums,
-                                 time = timeall,
+                                 time = annualmidyear,
                                  species = survspp,
@@
 
-First we subset the nums, resn, and structn components of `run_truth` using the same `create_survey` design selecting boxes, time, and species.
+First we subset the nums, resn, and structn components of `run_truth` using the same `create_survey` design selecting boxes, time (*once per year for NOBA*), and species.
 We need only apply survey efficiency and selectivity to nums, however (done above), so we apply `aggregateDensityData` here.
@@
 survey_aggresnall <- aggregateDensityData(dat = truth$resn,
-                                 time = timeall,
+                                 time = annualmidyear,
                                  species = survspp,
@@
 survey_aggstructnall <- aggregateDensityData(dat = truth$structn,
-                                 time = timeall,
+                                 time = annualmidyear,
                                  species = survspp,
@@
 
-This code block can take hours (5+ for CCA and 18+ for NOBA) to run, so be prepared, but then the saved file can be read back in to have length comps for all surveyed species.
+This code block can take hours (5+ for CCA and estimated 17+ for NOBA, but it never finished after 32+) to run, so be prepared, but then the saved file can be read back in to have length comps for all surveyed species.
 For a normal survey query we would have only one output timestep per year, which should save some time.
+
+And lets change NOBA to estimate length only 1 timestep per year (~2.5 hours to run):
 
@@
 
-```{r loadlengthcomp, echo=TRUE, eval=FALSE}
+```{r loadlengthcomp, echo=TRUE}
 
@@
 
-```{r moreplots, echo=TRUE, eval=FALSE}
+```{r moreplots, echo=TRUE}
 
@@
 
-So this looks like our best approximation of a true length comp for a particular Atlantis run. Comparisons with standard surveys next.
+So this looks like our best approximation of a true length comp for a particular Atlantis run (although it seems critical to limit output to allow functions to finish).
+Comparisons with standard surveys next.

### 2019-05-13 docs/index.Rmd
add CCA lf vis
add NOBA lf example, at long last

@@
 
+<a href="https://sgaichas.github.io/poseidon-dev/CCATrueLengthCompTest.html" target="_blank">CCA length composition examples - May 11, 2019</a>
+
+But don't try running length comps at the subannual level for all species.
+
+<a href="https://sgaichas.github.io/poseidon-dev/NOBATrueLengthCompTest.html" target="_blank">NOBA length composition examples - May 13, 2019</a>
+
 ### More to come later!

### 2019-05-13 visualization_Rmd/NEUSTrueLengthCompTest.Rmd
testing NEUS lf

--- 2019-05-13 visualization_Rmd/NOBATrueLengthCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate census length comps for Norwegian Barents Sea Atlantis"
+title: "Testing atlantisom: generate census length comps for Northeast US Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-First we subset the nums, resn, and structn components of `run_truth` using the same `create_survey` design selecting boxes, time (*once per year for NOBA*), and species.
+First we subset the nums, resn, and structn components of `run_truth` using the same `create_survey` design selecting boxes, time (*once per year for NEUS*), and species.
 We need only apply survey efficiency and selectivity to nums, however (done above), so we apply `aggregateDensityData` here.
@@
 Generating length composition for a single species should take just a few minutes to run.
-Update: for NOBA with many output timesteps, even a single species took 15-20 minutes.
+Update: for NOBA with many output timesteps, even a single species took 15-20 minutes, so we apply this at the annual level for NEUS.
 
@@
 
-ss_numsallhigh <- numsallhigh[numsallhigh$species == "Green_halibut",]
-ss_structnall <- structnall[structnall$species == "Green_halibut",]
-ss_resnall <- resnall[resnall$species == "Green_halibut",]
+ss_numsallhigh <- numsallhigh[numsallhigh$species == "SkateRay",]
+ss_structnall <- structnall[structnall$species == "SkateRay",]
+ss_resnall <- resnall[resnall$species == "SkateRay",]
 
@@
 lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
+#lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
+#lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
+#lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
 
@@
 This code block can take hours (5+ for CCA and estimated 17+ for NOBA, but it never finished after 32+) to run, so be prepared, but then the saved file can be read back in to have length comps for all surveyed species.
-For a normal survey query we would have only one output timestep per year, which should save some time.
+For a normal survey query we would have only one output timestep per year, which should save some time (NOBA 2.5 hours).
 
-And lets change NOBA to estimate length only 1 timestep per year (~2.5 hours to run):
+For NEUS, estimate length only 1 timestep per year:
 
-```{r allspplength, echo=TRUE, eval=FALSE}
+```{r allspplength, echo=TRUE}
 
@@
 ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
-  filter(species == "Norwegian_ssh")
+  filter(species == "Planktiv_S_Fish")
 

### 2019-05-13 StdSurvLengthCompTest.Rmd
initial standard survey length sampling

--- 2019-05-01 StandardSurveyTest.Rmd
@@
 ---
-title: "Testing atlantisom: example bottom trawl survey index"
+title: "Testing atlantisom: generate survey sampled length comps"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-initCCA <- FALSE
-initNEUS <- TRUE
+initCCA <- TRUE
+initNEUS <- FALSE
 initNOBA <- FALSE
 
-if(initCCA){
-  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
-  functional.groups.file <- "CalCurrentV3Groups.csv"
-  biomass.pools.file <- "DIVCalCurrentV3_BIOL.nc"
-  biol.prm.file <- "CalCurrentV3_Biol.prm"
-  box.file <- "CalCurrentV3_utm.bgm"
-  initial.conditions.file <- "DIVCalCurrentV3_BIOL.nc"
-  run.prm.file <- "CalCurrentV3_run.xml"
-  scenario.name <- "CCV3"
-}
+if(initCCA) source(here("config/CCConfig.R"))
 
-if(initNEUS){
-  d.name <- here("atlantisoutput","NEUStest20160303")
-  functional.groups.file <- "NeusGroups.csv" 
-  biomass.pools.file <- ""
-  biol.prm.file <- "at_biol_neus_v15_DE.prm"
-  box.file <- "neus30_2006.bgm"
-  initial.conditions.file <- "inneus_2012.nc"
-  run.prm.file <- "at_run_neus_v15_DE.xml"
-  scenario.name <- "neusDynEffort_Test1_"
-}
+if(initNEUS) source(here("config/NEUSConfig.R"))
 
-if(initNOBA){
-  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
-  functional.groups.file <- "nordic_groups_v04.csv" 
-  biomass.pools.file <- "nordic_biol_v23.nc"
-  biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
-  box.file <- "Nordic02.bgm"
-  initial.conditions.file <- "nordic_biol_v23.nc"
-  run.prm.file <- "nordic_run_v01.xml"
-  scenario.name <- "nordic_runresults_01"
-}
-# NOBA note: output filenames in CCA and NEUS begin with "output" and the run_truth function is written to expect this.
-Need to check if default Atlantis output file nomenclature has changed or if NOBA is a special case.
-For now, NOBA filenames have been changed to include prefix "output"
+if(initNOBA) source(here("config/NOBAConfig.R"))
+
 ```
@@
 if(initNOBA){
-  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
   truth.file <- "outputnordic_runresults_01run_truth.RData" 
   load(file.path(d.name, truth.file))
-  NOBAresults <- result
+  truth <- result
 }
@@
 
-We also read in previously generated survey census files based on true biomass results for comparison (see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).)
+## Simulate a survey part 4: sample for length composition
 
-```{r readsurvey1}
+This section uses the `atlantisom::create_survey()` and `atlantisom::sample_fish()` to get a biological sample dataset. From that dataset, several other functions can be run:
 
-if(initCCA) {
-  CCAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
-}
+ *  age samples from `calc_stage2age` and `sample_ages`
+ *  length samples from `calc_age2length` and `sample_lengths` (TO BE WRITTEN)
+ *  weight samples from `calc_age2length` and`sample_weights` (TO BE WRITTEN)
+ *  diet samples from `sample_diet`
 
-if(initNEUS) {
-  NEUSsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
-}
+Atlantis outputs numbers by cohort (stage-age) and growth informtation, but does not output size of animals directly.
+The function we are testing here, `atlantisom::calc_age2length` converts numbers by cohort to a length composition.
+The [workflow originally envisioned](https://onedrive.live.com/?authkey=%21AFQkOoKRz64TLUw&cid=59547B4CB95EF108&id=59547B4CB95EF108%216291&parId=59547B4CB95EF108%216262&o=OneUp) was to create a survey, sample fish from the survey, then apply this function.
+We determined that it will not work better to create a length comp for the whole population, which could then be sampled.
+We rethought what we needed [here](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html) and test the new functions in this document.
 
-if(initNOBA) {
-  NOBAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
-}
-
-```
-
-## Simulate a survey part 2: create a standard survey
-
-This section uses the `atlantisom::create_survey()` and `atlantisom::sample_survey_biomass()` functions by creating a standard survey covering discrete areas, times, and species, and with both incomplete efficiency and some kind of length selectivity.
-We will apply this standard survey and compare it with true output as an example of the type of information that can be generated for assessment model performance testing.
-
+### Standard survey settings for testing 
 To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
-
-One approach to creating a standard survey is to map species into general groups that are likely to have similar efficency and selectivity.
-For example, we expect that a bottom trawl survey will have generally lower efficiency for pelagic species than for demersals.
-We can also set efficiency to a very small number for infrequently encountered species occupying habitats that are not suited to bottom trawl sampling.
-This can simulate survey sampling for "data poor" species.
-Similarly, a common selectivity at age (or in this case, Atlantis cohort) can be defined for each species group.
-
-First, we map species into general groups for each model:
 
@@
 #   selected flatfish
+if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)
+
+survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
+                           funct.groups$GroupType %in% c("FISH", "SHARK")]
 
@@
 
-We use the following specifications for our default standard bottom trawl survey, including survey cv by species group:
+The following settings are for our example standard survey once per year, most areas, with mixed efficiency and selectivity:
 
@@
 
-# general specifications for bottom trawl survey:
+# general specifications for bottom trawl survey, with items defined above commented out to avoid wasting time loading already loaded files:
 #   once per year at mid year
-#   could generalize from the run.prm file: 
+# generalized timesteps all models
 runpar <- load_runprm(d.name, run.prm.file)
@@
 stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
-#   take midpoint of 0, steps per year to start seq and go to max time by steps per year
+
 midptyr <- round(median(seq(1,stepperyr)))
@@
 
-#   hardcoded for NEUS 1.0 in trunk here
-# annualmidyear <- seq(3,251,5)
-# #other options for NEUS
-# annualspring <- seq(2,251,5)
-# annualfall <- seq(4,251,5)
-
-#   all fish (and sharks!! need dogfish in NEUS)
-# could add groups "CEP" for cephalopod, "PWN" for prawn, etc
-
-# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
-if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)
-
-survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
-                           funct.groups$GroupType %in% c("FISH", "SHARK")]
-
 #   ~75-80% of boxes (leave off deeper boxes?)
-#   cant think of a way to generalize across models, must be hard coded in reality
-#   hardcoded for NEUS 1.0 in trunk here: boxsurv <- c(1:21)
 boxpars <- load_box(d.name, box.file)
-# should return all model areas
-# boxall <- c(0:(boxpars$nbox - 1))
 boxsurv <- c(2:round(0.8*(boxpars$nbox - 1)))
@@
 
-Here we use `create_survey` on the biomass output of `run_truth` to create the survey, so the call to `sample_survey_biomass` will require a weight at age argument that is filled with 1's because no conversion from numbers to weight is necessary.
+Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition.
 
-```{r stdsurveyBbased}
+```{r stdsurveyNbased}
 
-# this uses result$biomass_ages to sample biomass directly
-
-if(initCCA) datB <- CCAresults$biomass_ages
-if(initNEUS) datB <- NEUSresults$biomass_ages
-if(initNOBA) datB <- NOBAresults$biomass_ages
-
-survey_testBstd <- create_survey(dat = datB,
+survey_testNstd <- create_survey(dat = truth$nums,
                                  time = annualmidyear,
@@
 
-# call sample_survey_biomass with a bunch of 1s for weight at age
-# in the code it multiplies atoutput by wtatage so this allows us to use
-# biomass directly
-wtage <- data.frame(species=rep(survspp, each=10),
-                    agecl=rep(c(1:10),length(survspp)),
-                    wtAtAge=rep(1.0,length(survspp)*10))
-
-stdsurveyB_frombio <- sample_survey_biomass(survey_testBstd, surv_cv_mix, wtage)
-
-if(initCCA) CCAstdsurveyB_frombio <- stdsurveyB_frombio
-if(initNEUS) NEUSstdsurveyB_frombio <- stdsurveyB_frombio
-if(initNOBA) NOBAstdsurveyB_frombio <- stdsurveyB_frombio
+# consider saving this interim step if it takes a long time go generate
 
@@
 
-Comparing our (census) survey based on true biomass from above with the Atlantis output file "[modelscenario]BiomIndx.txt" should give us a perfect match.
-Note that the our (census) survey may have more sampling in time than the Atlantis output file.
+Next we apply the aggregateDensityData function can to resn and structn for survey times, species, and boxes.
 
-```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species.", message=FALSE, warning=FALSE}
-# plot some comparisons with Atlantis output
+```{r aggdens}
 
-# read Atlantis output files
-if(initCCA) {
-  atBtxt2 <- read.table(here("atlantisoutput","CalCurrentSummitScenario1","outputCCV3BiomIndx.txt"), header=T)
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  stdsurveyB_frombio <- CCAstdsurveyB_frombio
-  censusB_frombio <- CCAsurveyB_frombio
-}
+# aggregate true resn per survey design
+survey_aggresnstd <- aggregateDensityData(dat = truth$resn,
+                                 time = annualmidyear,
+                                 species = survspp,
+                                 boxes = boxsurv)
 
-if(initNEUS) {
-  atBtxt2 <- read.table(here("atlantisoutput","NEUStest20160303","neusDynEffort_Test1_BiomIndx.txt"), header=T)
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  stdsurveyB_frombio <- NEUSstdsurveyB_frombio
-  censusB_frombio <- NEUSsurveyB_frombio
-}
-
-if(initNOBA) {
-  atBtxt2 <- read.table(here("atlantisoutput","NOBACERESGlobalSustainability","outputnordic_runresults_01BiomIndx.txt"), header=T) 
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  stdsurveyB_frombio <- NOBAstdsurveyB_frombio
-  censusB_frombio <- NOBAsurveyB_frombio
-}
-
-# lookup the matching names, put in time, species, biomass column format
-# WARNING hardcoded for output with last species group as DIN
-groupslookup <- groupslookup %>%
-  filter(IsTurnedOn > 0)
-
-atBtxt2tidy <- atBtxt2 %>%
-  select(Time:DIN) %>%
-  #select(Time, FPL:DIN) %>%
-  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
-  gather(species, biomass, -Time) %>%
-  filter(species %in% levels(stdsurveyB_frombio$species))
-
-censusB_frombio <- censusB_frombio %>%
-  filter(species %in% levels(stdsurveyB_frombio$species))
-
-#all species comparison, time intervals hardcoded for NEUS and NOBA
-compareB <-ggplot() +
-  geom_line(data=censusB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
-            alpha = 5/10) +
-  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
-             alpha = 5/10) + 
-  #geom_line(data=surveyB_frombio_eff, aes(x=time/5,y=atoutput, color="stdeffB")) +
-  #geom_line(data=surveyB_frombio_effsel, aes(x=time/5,y=atoutput, color="stdeffselB")) +
-  geom_point(data=stdsurveyB_frombio, aes(x=time/5,y=atoutput, color="stdeffcvselB"),
-              alpha = 1/2) +
-
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+# aggregate true structn per survey design
+survey_aggstructnstd <- aggregateDensityData(dat = truth$structn,
+                                 time = annualmidyear,
+                                 species = survspp,
+                                 boxes = boxsurv)
 
@@
 
-```{r matchBp2, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+Now we should have inputs to `sample_fish` on the same scale, and they need to be aggregated across boxes into a single biological sample for the whole survey.
+The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.
+
+```{r stdsurvey-lensamp, warning=FALSE, message=FALSE}
+
+#   define n fish for biological sampling by group
+#   this could easily be a vector or time series, constant here
+ns.nt <- 25 # for large pelagics, reef dwellers, others not in trawlable habitat
+ns.pl <- 1000  # for pelagics
+ns.dm <- 1000  # for demersals
+ns.fl <- 1000  # for selected flatfish
+
+effNnontrawl <- data.frame(species=nontrawl, effN=rep(ns.nt,length(nontrawl)))
+effNpelagics <- data.frame(species=pelagics, effN=rep(ns.pl,length(pelagics)))
+effNdemersals <- data.frame(species=demersals, effN=rep(ns.dm,length(demersals)))
+effNselflats <- data.frame(species=selflats, effN=rep(ns.fl,length(selflats)))
+
+effNmix <- bind_rows(effNnontrawl, effNpelagics, effNdemersals, effNselflats)
+
+# apply default sample fish as before to get numbers
+numsstd <- sample_fish(survey_testNstd, effNmix)
+
+#dont sample these, just aggregate them using median (effNmix does nothing)
+structnstd <- sample_fish(survey_aggstructnstd, effNmix, sample = FALSE)
+
+resnstd <-  sample_fish(survey_aggresnstd, effNmix, sample = FALSE)
+```
+
+```{r stdsurvey-calcage2length, warning=FALSE, message=FALSE}
+
+# now cut these down to a single species for testing
+
+atf_numsstd <- numsstd[numsstd$species == "Arrowtooth_flounder",]
+atf_structnstd <- structnstd[structnstd$species == "Arrowtooth_flounder",]
+atf_resnstd <- resnall[resnstd$species == "Arrowtooth_flounder",]
+
+atf_length_stdsurv <- calc_age2length(structn = atf_structnstd,
+                                 resn = atf_resnstd,
+                                 nums = atf_numsstd,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
 
@@
 
-```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+Plot samples one species:
+
+```{r atflengthsamp1-test}
+
+lfplot <- ggplot(atf_length_stdsurv$natlength, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte()
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
+
+# try dir = "v" option for vertical lf comparisons
 
@@
 
-```{r matchBp4, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species.
-Depending on the model, we may not show all species in this example."}
-if(!initNEUS){
-  compareB + 
-    facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free") 
-}
+## Samples for all species, then SAVE
+
+Now run for all species in the survey.
+
+```{r allspplengthsamp}
+
+length_stdsurv <- calc_age2length(structn = structnstd,
+                                 resn = resnstd,
+                                 nums = numsstd,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+#save for later use, takes a long time to generate
+saveRDS(length_stdsurv, file.path(d.name, paste0(scenario.name, "length_stdsurv.rds")))
+
 ```
 
-Given the misses using survey number outputs from run truth with a calculated weight at age, we will not repeat this example using numbers.
+Runtime for census lengths was 5.5 hours for CCA. For standard survey lengths runtime was not reduced because the dimensions (species, timesteps) were identical.
+
+Compare sampled lf with census from [here](https://sgaichas.github.io/poseidon-dev/CCATrueLengthCompTest.html):
+
+```{r loadlengthcomp, echo=TRUE, eval=FALSE}
+
+length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))
+
+length_stdsurv <- readRDS(file.path(d.name, paste0(scenario.name, "length_stdsurv.rds")))
+
+```
+
+Comparison plots for other species:
+
+```{r moreplots, echo=TRUE, eval=FALSE}
+
+ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
+  filter(species == "Herring")
+
+ss_length_stdsurv <- length_stdsurv$natlength %>%
+  filter(species == "Herring")
+
+# add sample, this is census, probably want it first?
+lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte()
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
+
+```
 

### 2019-05-13 StdSurvLengthCompTest.Rmd
correct resnstd name

@@
 atf_structnstd <- structnstd[structnstd$species == "Arrowtooth_flounder",]
-atf_resnstd <- resnall[resnstd$species == "Arrowtooth_flounder",]
+atf_resnstd <- resnstd[resnstd$species == "Arrowtooth_flounder",]
 

### 2019-05-14 StdSurvLengthCompTest.Rmd
add census-sample lf comparison

@@
 
-```{r allspplengthsamp}
+```{r allspplengthsamp, eval=FALSE}
 
@@
 
-Comparison plots for other species:
+Comparison plots for other species.
+Blue bars are the census length sample and red outlines are the survey sample (standardized for comparison; census total n lengths is 1e+8 while sample total n lengths is 1000):
 
-```{r moreplots, echo=TRUE, eval=FALSE}
+```{r moreplots, echo=TRUE}
 
+# select a species
 ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
@@
 
+# make proportion at length to plot together
+lf_census_tot <- aggregate(ss_length_censussurvsamp$atoutput,list(ss_length_censussurvsamp$time),sum )
+names(lf_census_tot) <- c("time","totlen")
+
+lf_census_prop <- merge(ss_length_censussurvsamp, lf_census_tot)
+
+lf_samp_tot <- aggregate(ss_length_stdsurv$atoutput,list(ss_length_stdsurv$time),sum )
+names(lf_samp_tot) <- c("time","totsamp")
+
+lf_samp_prop <- merge(ss_length_stdsurv, lf_samp_tot)
+
 # add sample, this is census, probably want it first?
-lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
-  geom_bar(aes(weight = atoutput)) +
+lfplot <- ggplot(mapping=aes(x=upper.bins)) +
+  geom_bar(data=lf_census_prop, aes(weight = atoutput/totlen), fill="blue") +
+  geom_bar(data=lf_samp_prop, aes(weight = atoutput/totsamp), colour="red") +
   theme_tufte()
@@
 
+Here as with [cohort sampling](https://sgaichas.github.io/poseidon-dev/StdSurvLengthAgeCompTest.html), the impact of survey selectivity is much larger than sampling error with effN of 1000.
+We could try effN that is lower, but 1000 lengths may be reasonable from a survey.
+A `sample_lengths` function could add further measuremet error on top of this, if desired (though measurement error at the 1 cm level may be negligible).
+
 ## References

### 2019-05-14 docs/index.Rmd
add NEUS example
add lf from standard survey

@@
 
+<a href="https://sgaichas.github.io/poseidon-dev/NEUSTrueLengthCompTest.html" target="_blank">NEUS length composition examples - May 13, 2019</a>
+
+Standard survey with length sampling: impact of selectivity, `sample_fish`
+
+<a href="https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html" target="_blank">Compare survey + sampled lengths with true - May 14, 2019</a>
+
 ### More to come later!

### 2019-05-14 visualization_Rmd/CCAStdSurvLengthCompTest.Rmd
vis lf sample CCA

--- 2019-05-14 StdSurvLengthCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate survey sampled length comps"
+title: "Testing atlantisom: generate survey sampled length comps for California Current Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-This section uses the `atlantisom::create_survey()` and `atlantisom::sample_fish()` to get a biological sample dataset. From that dataset, several other functions can be run:
+Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html).
 
- *  age samples from `calc_stage2age` and `sample_ages`
- *  length samples from `calc_age2length` and `sample_lengths` (TO BE WRITTEN)
- *  weight samples from `calc_age2length` and`sample_weights` (TO BE WRITTEN)
- *  diet samples from `sample_diet`
-
-Atlantis outputs numbers by cohort (stage-age) and growth informtation, but does not output size of animals directly.
-The function we are testing here, `atlantisom::calc_age2length` converts numbers by cohort to a length composition.
-The [workflow originally envisioned](https://onedrive.live.com/?authkey=%21AFQkOoKRz64TLUw&cid=59547B4CB95EF108&id=59547B4CB95EF108%216291&parId=59547B4CB95EF108%216262&o=OneUp) was to create a survey, sample fish from the survey, then apply this function.
-We determined that it will not work better to create a length comp for the whole population, which could then be sampled.
-We rethought what we needed [here](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html) and test the new functions in this document.
-
-### Standard survey settings for testing 
 To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
 
-```{r sppgroups}
+```{r sppgroups, echo=TRUE}
 
@@
 
-```{r aggdens}
+```{r aggdens, echo=TRUE}
 
@@
 
-```{r stdsurvey-lensamp, warning=FALSE, message=FALSE}
+```{r stdsurvey-lensamp, warning=FALSE, message=FALSE, echo=TRUE}
 
@@
 
-```{r stdsurvey-calcage2length, warning=FALSE, message=FALSE}
+```{r stdsurvey-calcage2length, warning=FALSE, message=FALSE, echo=TRUE}
 
@@
   geom_bar(aes(weight = atoutput)) +
-  theme_tufte()
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=paste(scenario.name, atf_length_stdsurv$natlength$species))
 
@@
 
-## Samples for all species, then SAVE
+## Samples for all species from SAVED file
 
@@
 
-```{r allspplengthsamp, eval=FALSE}
+```{r allspplengthsamp, eval=FALSE, echo=TRUE}
 
@@
 
-```{r moreplots, echo=TRUE}
+```{r moreplots}
 
@@
   geom_bar(data=lf_samp_prop, aes(weight = atoutput/totsamp), colour="red") +
-  theme_tufte()
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=paste(scenario.name, lf_census_prop$species))
 
@@
 A `sample_lengths` function could add further measuremet error on top of this, if desired (though measurement error at the 1 cm level may be negligible).
-
-## References

### 2019-05-14 visualization_Rmd/NEUSStdSurvLengthCompTest.Rmd
initial lf sample viz NOBA NEUS

--- 2019-05-14 visualization_Rmd/CCAStdSurvLengthCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate survey sampled length comps for California Current Atlantis"
+title: "Testing atlantisom: generate survey sampled length comps for Northeast US  Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-atf_numsstd <- numsstd[numsstd$species == "Arrowtooth_flounder",]
-atf_structnstd <- structnstd[structnstd$species == "Arrowtooth_flounder",]
-atf_resnstd <- resnstd[resnstd$species == "Arrowtooth_flounder",]
+ss_numsstd <- numsstd[numsstd$species == "SkateRay",]
+ss_structnstd <- structnstd[structnstd$species == "SkateRay",]
+ss_resnstd <- resnstd[resnstd$species == "SkateRay",]
 
-atf_length_stdsurv <- calc_age2length(structn = atf_structnstd,
-                                 resn = atf_resnstd,
-                                 nums = atf_numsstd,
+ss_length_stdsurv <- calc_age2length(structn = ss_structnstd,
+                                 resn = ss_resnstd,
+                                 nums = ss_numsstd,
                                  biolprm = truth$biolprm, fgs = truth$fgs,
@@
 
-```{r atflengthsamp1-test}
+```{r sslengthsamp1-test}
 
-lfplot <- ggplot(atf_length_stdsurv$natlength, aes(upper.bins)) +
+lfplot <- ggplot(ss_length_stdsurv$natlength, aes(upper.bins)) +
   geom_bar(aes(weight = atoutput)) +
   theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=paste(scenario.name, atf_length_stdsurv$natlength$species))
+  labs(subtitle = paste(scenario.name, ss_length_stdsurv$natlength$species))
 
@@
 
-## Samples for all species from SAVED file
+## Samples for all species and SAVE the file
 
@@
 
-```{r allspplengthsamp, eval=FALSE, echo=TRUE}
+```{r allspplengthsamp, echo=TRUE}
 
@@
 
-Runtime for census lengths was 5.5 hours for CCA. For standard survey lengths runtime was not reduced because the dimensions (species, timesteps) were identical.
+Runtime for census lengths was <1 hour for NEUS because it is a 50 year run and we are taking lengths only once per year.
+For standard survey lengths runtime was not reduced because the dimensions (species, timesteps) were identical.
 
-Compare sampled lf with census from [here](https://sgaichas.github.io/poseidon-dev/CCATrueLengthCompTest.html):
+Compare sampled lf with census from [here](https://sgaichas.github.io/poseidon-dev/NEUSTrueLengthCompTest.html):
 
@@
 ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
-  filter(species == "Herring")
+  filter(species == "Planktiv_S_Fish")
 
 ss_length_stdsurv <- length_stdsurv$natlength %>%
-  filter(species == "Herring")
+  filter(species == "Planktiv_S_Fish")
 
@@
   theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=paste(scenario.name, lf_census_prop$species))
+  labs(subtitle = paste(scenario.name, lf_census_prop$species))
 

### 2019-05-14 visualization_Rmd/NOBAStdSurvLengthCompTest.Rmd
initial lf sample viz NOBA NEUS

--- 2019-05-14 visualization_Rmd/NEUSStdSurvLengthCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate survey sampled length comps for Northeast US  Atlantis"
+title: "Testing atlantisom: generate survey sampled length comps for Norwegian Barents Sea Atlantis"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-ss_numsstd <- numsstd[numsstd$species == "SkateRay",]
-ss_structnstd <- structnstd[structnstd$species == "SkateRay",]
-ss_resnstd <- resnstd[resnstd$species == "SkateRay",]
+ss_numsstd <- numsstd[numsstd$species == "Green_halibut",]
+ss_structnstd <- structnstd[structnstd$species == "Green_halibut",]
+ss_resnstd <- resnstd[resnstd$species == "Green_halibut",]
 
@@
 lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
-#lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
-#lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
-#lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
 
@@
 
-Runtime for census lengths was <1 hour for NEUS because it is a 50 year run and we are taking lengths only once per year.
+Runtime for census lengths was ~2.5 hours for NOBA once we decided once per year was plenty.
 For standard survey lengths runtime was not reduced because the dimensions (species, timesteps) were identical.
 
-Compare sampled lf with census from [here](https://sgaichas.github.io/poseidon-dev/NEUSTrueLengthCompTest.html):
+Compare sampled lf with census from [here](https://sgaichas.github.io/poseidon-dev/NOBATrueLengthCompTest.html):
 
@@
 ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
-  filter(species == "Planktiv_S_Fish")
+  filter(species == "Norwegian_ssh")
 
 ss_length_stdsurv <- length_stdsurv$natlength %>%
-  filter(species == "Planktiv_S_Fish")
+  filter(species == "Norwegian_ssh")
 

### 2019-05-15 docs/index.Rmd
add lf samp visualizations

@@
 
+<a href="https://sgaichas.github.io/poseidon-dev/CCAStdSurvLengthCompTest.html" target="_blank">CCA compare survey + sampled lengths with true - May 14, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NOBAStdSurvLengthCompTest.html" target="_blank">NOBA compare survey + sampled lengths with true - May 15, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NEUSStdSurvLengthCompTest.html" target="_blank">NEUS compare survey + sampled lengths with true - May 15, 2019</a>
+
 ### More to come later!

### 2019-05-15 TestBiggerMaxLen.Rmd
testing changes to calc_age2length

--- 2019-05-14 visualization_Rmd/NOBAStdSurvLengthCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate survey sampled length comps for Norwegian Barents Sea Atlantis"
+title: "Testing atlantisom: allow users to specify different max length bin for length compositions"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 ## Introduction
+
+Here we briefly test changes to the function `calc_age2length` that should allow the user to specify a different max length bin.
+For all setup, etc, please see previous files Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html) and  [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).
 
@@
 
-## Simulate a survey part 4: sample for length composition
+## Simulate a survey part 4: sample for length composition (testing revised function)
 
 Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html).
+
+We will apply examples here to only one species, Greenland halibut in NOBA, which grows to a large size.
 
@@
 
-Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition.
-
-```{r stdsurveyNbased, echo=TRUE}
-
-survey_testNstd <- create_survey(dat = truth$nums,
-                                 time = annualmidyear,
-                                 species = survspp,
-                                 boxes = boxsurv,
-                                 effic = efficmix,
-                                 selex = selexmix)
-
-# consider saving this interim step if it takes a long time go generate
-
-```
-
-Next we apply the aggregateDensityData function can to resn and structn for survey times, species, and boxes.
-
-```{r aggdens, echo=TRUE}
-
-# aggregate true resn per survey design
-survey_aggresnstd <- aggregateDensityData(dat = truth$resn,
-                                 time = annualmidyear,
-                                 species = survspp,
-                                 boxes = boxsurv)
-
-# aggregate true structn per survey design
-survey_aggstructnstd <- aggregateDensityData(dat = truth$structn,
-                                 time = annualmidyear,
-                                 species = survspp,
-                                 boxes = boxsurv)
-
-```
-
-Now we should have inputs to `sample_fish` on the same scale, and they need to be aggregated across boxes into a single biological sample for the whole survey.
-The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.
+And the numbers to be sampled for lengths each year:
 
@@
 
-# apply default sample fish as before to get numbers
-numsstd <- sample_fish(survey_testNstd, effNmix)
+```
+
+Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just one species in this case).
+The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.
+
+```{r stdsurveyNbased-GHR, echo=TRUE}
+
+ss.name <- funct.group.names[funct.group.names == "Green_halibut"]
+
+#change back to flat selectivity to see full comp
+selex1 <- data.frame(species=rep(funct.group.names, each=10),
+                     agecl=rep(c(1:10),length(funct.group.names)),
+                     selex=rep(1.0,length(funct.group.names)*10))
+
+# get survey nums with full (no) selectivity
+ss_survey_testNstd_nosel <- create_survey(dat = truth$nums,
+                                 time = annualmidyear,
+                                 species = ss.name,
+                                 boxes = boxsurv,
+                                 effic = efficmix,
+                                 selex = selex1)
+
+# now sample fish nums from this
+ss_numsstd_nosel <- sample_fish(ss_survey_testNstd_nosel, effNmix)
+
+# aggregate true resn per survey design
+survey_aggresnstd <- aggregateDensityData(dat = truth$resn,
+                                 time = annualmidyear,
+                                 species = ss.name,
+                                 boxes = boxsurv)
+
+# aggregate true structn per survey design
+survey_aggstructnstd <- aggregateDensityData(dat = truth$structn,
+                                 time = annualmidyear,
+                                 species = ss.name,
+                                 boxes = boxsurv)
 
 #dont sample these, just aggregate them using median (effNmix does nothing)
-structnstd <- sample_fish(survey_aggstructnstd, effNmix, sample = FALSE)
+ss_structnstd <- sample_fish(survey_aggstructnstd, effNmix, sample = FALSE)
 
-resnstd <-  sample_fish(survey_aggresnstd, effNmix, sample = FALSE)
+ss_resnstd <-  sample_fish(survey_aggresnstd, effNmix, sample = FALSE)
+
 ```
 
-```{r stdsurvey-calcage2length, warning=FALSE, message=FALSE, echo=TRUE}
+Length sample with default max length bin (150 cm):
 
-# now cut these down to a single species for testing
+```{r default-maxlen, echo=TRUE}
 
-ss_numsstd <- numsstd[numsstd$species == "Green_halibut",]
-ss_structnstd <- structnstd[structnstd$species == "Green_halibut",]
-ss_resnstd <- resnstd[resnstd$species == "Green_halibut",]
-
-ss_length_stdsurv <- calc_age2length(structn = ss_structnstd,
+ss_length_stdsurv_nosel <- calc_age2length(structn = ss_structnstd,
                                  resn = ss_resnstd,
-                                 nums = ss_numsstd,
+                                 nums = ss_numsstd_nosel,
                                  biolprm = truth$biolprm, fgs = truth$fgs,
@@
 
-Plot samples one species:
+Length sample with user specified max length bin (250 cm):
 
-```{r sslengthsamp1-test}
+```{r userset-maxlen, echo=TRUE}
 
-lfplot <- ggplot(ss_length_stdsurv$natlength, aes(upper.bins)) +
-  geom_bar(aes(weight = atoutput)) +
-  theme_tufte() +
-  labs(subtitle = paste(scenario.name, ss_length_stdsurv$natlength$species))
-
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
-
-# try dir = "v" option for vertical lf comparisons
+ss_length_stdsurv_nosel_max <- calc_age2length(structn = ss_structnstd,
+                                 resn = ss_resnstd,
+                                 nums = ss_numsstd_nosel,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 maxbin = 250,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
 
@@
 
-## Samples for all species and SAVE the file
+Plots show that default length bins are not adequate for something like Greenland halibut, which get larger than the default largest bin in `calc_age2length`, 150 cm.
+Both census and these sampled length comps are chopped off at 150 cm for this species:
 
-Now run for all species in the survey.
+```{r sslengthsamp2-testdefault}
 
-```{r allspplengthsamp, echo=TRUE}
-
-length_stdsurv <- calc_age2length(structn = structnstd,
-                                 resn = resnstd,
-                                 nums = numsstd,
-                                 biolprm = truth$biolprm, fgs = truth$fgs,
-                                 CVlenage = 0.1, remove.zeroes=TRUE)
-
-#save for later use, takes a long time to generate
-saveRDS(length_stdsurv, file.path(d.name, paste0(scenario.name, "length_stdsurv.rds")))
-
-```
-
-Runtime for census lengths was ~2.5 hours for NOBA once we decided once per year was plenty.
-For standard survey lengths runtime was not reduced because the dimensions (species, timesteps) were identical.
-
-Compare sampled lf with census from [here](https://sgaichas.github.io/poseidon-dev/NOBATrueLengthCompTest.html):
-
-```{r loadlengthcomp, echo=TRUE}
-
-length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))
-
-length_stdsurv <- readRDS(file.path(d.name, paste0(scenario.name, "length_stdsurv.rds")))
-
-```
-
-Comparison plots for other species.
-Blue bars are the census length sample and red outlines are the survey sample (standardized for comparison; census total n lengths is 1e+8 while sample total n lengths is 1000):
-
-```{r moreplots}
-
-# select a species
-ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
-  filter(species == "Norwegian_ssh")
-
-ss_length_stdsurv <- length_stdsurv$natlength %>%
-  filter(species == "Norwegian_ssh")
-
-# make proportion at length to plot together
-lf_census_tot <- aggregate(ss_length_censussurvsamp$atoutput,list(ss_length_censussurvsamp$time),sum )
-names(lf_census_tot) <- c("time","totlen")
-
-lf_census_prop <- merge(ss_length_censussurvsamp, lf_census_tot)
-
-lf_samp_tot <- aggregate(ss_length_stdsurv$atoutput,list(ss_length_stdsurv$time),sum )
-names(lf_samp_tot) <- c("time","totsamp")
-
-lf_samp_prop <- merge(ss_length_stdsurv, lf_samp_tot)
-
-# add sample, this is census, probably want it first?
-lfplot <- ggplot(mapping=aes(x=upper.bins)) +
-  geom_bar(data=lf_census_prop, aes(weight = atoutput/totlen), fill="blue") +
-  geom_bar(data=lf_samp_prop, aes(weight = atoutput/totsamp), colour="red") +
+lfplot <- ggplot(ss_length_stdsurv_nosel$natlength, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
   theme_tufte() +
-  labs(subtitle = paste(scenario.name, lf_census_prop$species))
+  labs(subtitle = paste(scenario.name, ss_length_stdsurv_nosel$natlength$species))
 
@@
 
-Here as with [cohort sampling](https://sgaichas.github.io/poseidon-dev/StdSurvLengthAgeCompTest.html), the impact of survey selectivity is much larger than sampling error with effN of 1000.
-We could try effN that is lower, but 1000 lengths may be reasonable from a survey.
-A `sample_lengths` function could add further measuremet error on top of this, if desired (though measurement error at the 1 cm level may be negligible).
+I changed the `calc_age2length` code to pass the upper bin of 150 as a default but allow the user to change it in the function call.
+This plot shows results from setting maxbin = 250 for Greenland Halibut in NOBA:
+
+```{r sslengthsamp2-testmax250}
+
+lfplot <- ggplot(ss_length_stdsurv_nosel_max$natlength, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        ss_length_stdsurv_nosel_max$natlength$species))
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
+
+```
+
+Now we can go back and apply standard survey selectivity as in the [NOBA visualization](https://sgaichas.github.io/poseidon-dev/NOBAStdSurvLengthCompTest.html) and see how the length comps look:
+
+```{r stdsurveyNbased-GHR2, echo=TRUE}
+
+ss.name <- funct.group.names[funct.group.names == "Green_halibut"]
+
+# get survey nums with standard selectivity
+ss_survey_testNstd <- create_survey(dat = truth$nums,
+                                 time = annualmidyear,
+                                 species = ss.name,
+                                 boxes = boxsurv,
+                                 effic = efficmix,
+                                 selex = selexmix)
+
+# now sample fish nums from this
+ss_numsstd <- sample_fish(ss_survey_testNstd, effNmix)
+
+# structn and resn stuff is exactly the same because selectivity is irrelevant
+```
+
+Length sample from standard survey with user specified max length bin (250 cm):
+
+```{r user-maxlen2, echo=TRUE}
+
+ss_length_stdsurv_max <- calc_age2length(structn = ss_structnstd,
+                                 resn = ss_resnstd,
+                                 nums = ss_numsstd,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 maxbin = 250,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+```
+
+```{r sslengthsamp3-testmax250}
+
+lfplot <- ggplot(ss_length_stdsurv_max$natlength, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        ss_length_stdsurv_max$natlength$species))
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
+
+```
+
+Bigger max bin is definitely more appropriate for this species.
+
+I suppose we could make the maxbin a vector by species, but for now may be overkill.

### 2019-05-15 docs/index.Rmd
add maxbin testing

@@
 May 11: Changes to `atlantisom::sample_fish` and new `atlantisom::aggregateDensityData` functions seem to work. See below under [Length compositions].
+
+May 15: Changes to `atlantisom::calc_age2length` to allow user specified max length bin seem to work. Also listed under [Length compositions].
 
@@
 
+Test new user specified max length bin for a large species:
+
+<a href="https://sgaichas.github.io/poseidon-dev/TestBiggerMaxLen.html" target="_blank">Default 150 cm, try different max - May 15, 2019</a>
+
 ### More to come later!

### 2019-05-15 TestSmallestLength.Rmd
looking for small turbot

--- 2019-05-11 visualization_Rmd/NOBATrueLengthCompTest.Rmd
@@
 ---
-title: "Testing atlantisom: generate census length comps for Norwegian Barents Sea Atlantis"
+title: "Testing atlantisom: do we ever get small fish in length comps?"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 ## Introduction
+
+Here we further test the function `calc_age2length` to see if small fish ever appear with full census sampling.
+If they don't, perhaps the default CV in the function needs modification?
+This question is triggered by the [sampling for NOBA Greenland halibut](https://sgaichas.github.io/poseidon-dev/NOBAStdSurvLengthCompTest.html), in which fish below 50 cm didn't show up.
+
+For all setup, etc, please see previous files Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html) and  [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).
 
@@
 
-## Simulate a survey part 4: sample for length composition
+## Simulate a survey part 4: sample for length composition (testing revised function)
 
-Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).
+Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html).
 
-The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages:
+We will apply examples here to only one species, Greenland halibut in NOBA, which grows to a large size.
 
-```{r census-spec, message=FALSE, warning=FALSE, echo=TRUE}
+To create a census, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
+The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages.
+
+```{r census-spec, message=FALSE, warning=FALSE}
 
@@
 
-Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition.
+Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just one species in this case).
+The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.
 
-```{r stdsurveyNbased, echo=TRUE}
+Because we don't want to wait 24 hours for this, we will look at only the first 112 time steps.
 
-survey_testNall <- create_survey(dat = truth$nums,
-                                 time = timeall,
-                                 species = survspp,
+```{r stdsurveyNbased-GHR, echo=TRUE}
+
+ss.name <- funct.group.names[funct.group.names == "Green_halibut"]
+
+# get survey nums with full (no) selectivity
+ss_survey_testNall <- create_survey(dat = truth$nums,
+                                 time = c(0:111),
+                                 species = ss.name,
                                  boxes = boxall,
@@
                                  selex = selex1)
-
-# consider saving this interim step if it takes a long time go generate
-
-```
-
-```{r truecohortagecomp}
-# what is true composition? need annual by species, use code from sample_fish
-# do tidyly
-dat2 <- survey_testNall %>%
-  group_by(species, agecl, time) %>%
-  summarize(numAtAge = sum(atoutput))
-
-totN <- dat2 %>%
-  group_by(species, time) %>%
-  summarize(totN = sum(numAtAge))
-
-dat2totN <- merge(dat2, totN)
-
-# ageclcomp <- ggplot(dat2totN, aes(x=agecl, y=numAtAge/totN, col=time)) +
-#   geom_point()
-# 
-# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
-# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
-# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
-# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
-
-```
-
-Then we use the `sample_fish` function with very high effN and compare to true annual age comp calculated above as a test, which matches as we have shown [here](https://sgaichas.github.io/poseidon-dev/TrueLengthAgeCompTest.html).
-
-The following gets both numbers and weights to get the length comp, and ultimately the weight at age for use in assessments.
-
-First we subset the nums, resn, and structn components of `run_truth` using the same `create_survey` design selecting boxes, time, and species.
-We need only apply survey efficiency and selectivity to nums, however (done above), so we apply `aggregateDensityData` here.
-
-```{r aggdens, echo=TRUE}
-
-# aggregate true resn per survey design
-survey_aggresnall <- aggregateDensityData(dat = truth$resn,
-                                 time = timeall,
-                                 species = survspp,
-                                 boxes = boxall)
-
-# aggregate true structn per survey design
-survey_aggstructnall <- aggregateDensityData(dat = truth$structn,
-                                 time = timeall,
-                                 species = survspp,
-                                 boxes = boxall)
-
-```
-
-Now we should have inputs to `sample_fish` on the same scale, and they need to be aggregated across boxes into a single biological sample for the whole survey.
-We are not maintaining spatial structure in sampling because it isn't used in most assessments.
-
-To do the proper aggregation and not apply the multinomial sampling to the density data, I rewrote `sample_fish` to apply the median if `sample=FALSE` in the function call.
-
-```{r censussurvey-samplefish, warning=FALSE, message=FALSE, echo=TRUE}
 
@@
 # apply default sample fish as before to get numbers
-numsallhigh <- sample_fish(survey_testNall, effNhigh)
+ss_numsallhigh <- sample_fish(ss_survey_testNall, effNhigh)
+
+# aggregate true resn per survey design
+ss_aggresnall <- aggregateDensityData(dat = truth$resn,
+                                 time = c(0:111),
+                                 species = ss.name,
+                                 boxes = boxall)
+
+# aggregate true structn per survey design
+ss_aggstructnall <- aggregateDensityData(dat = truth$structn,
+                                 time = c(0:111),
+                                 species = ss.name,
+                                 boxes = boxall)
 
 #dont sample these, just aggregate them using median
-structnall <- sample_fish(survey_aggstructnall, effNhigh, sample = FALSE)
+ss_structnall <- sample_fish(ss_aggstructnall, effNhigh, sample = FALSE)
 
-resnall <-  sample_fish(survey_aggresnall, effNhigh, sample = FALSE)
+ss_resnall <-  sample_fish(ss_aggresnall, effNhigh, sample = FALSE)
 
@@
 
-Select a single species for testing. Generating length composition for a single species should take just a few minutes to run.
+Length sample with user specified max length bin (200 cm):
 
-```{r singlespp-lengthcomp, echo=TRUE}
+```{r userset-maxlen, echo=TRUE}
 
-# this should still represent a census but with polygon and layer aggregated
-# WARNING: selecting a species by name is hardcoded for a particular model
-
-ss_numsallhigh <- numsallhigh[numsallhigh$species == "Green_halibut",]
-ss_structnall <- structnall[structnall$species == "Green_halibut",]
-ss_resnall <- resnall[resnall$species == "Green_halibut",]
-
-ss_length_censussurvsamp <- calc_age2length(structn = ss_structnall,
+ss_length_census <- calc_age2length(structn = ss_structnall,
                                  resn = ss_resnall,
@@
                                  biolprm = truth$biolprm, fgs = truth$fgs,
+                                 maxbin = 200,
                                  CVlenage = 0.1, remove.zeroes=TRUE)
@@
 
-This gives annual length comps:
+We should get the upper end of Greenland halibut with a 200cm max length bin. What I am concerned about here is whether we get very small fish at any time step:
 
-```{r ss-lengthsamp1-test}
+```{r sslengthsamp2-test}
 
-lfplot <- ggplot(ss_length_censussurvsamp$natlength, aes(upper.bins)) +
+lfplot <- ggplot(ss_length_census$natlength, aes(upper.bins)) +
   geom_bar(aes(weight = atoutput)) +
-  theme_tufte()
-
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
-
-# try dir = "v" option for vertical lf comparisons
-
-```
-
-## Run for all species then SAVE
-
-This code block can take hours (5+ for CCA) to run, so be prepared, but then the saved file can be read back in to have length comps for all surveyed species.
-
-```{r allspplength, echo=TRUE}
-
-length_censussurvsamp <- calc_age2length(structn = structnall,
-                                 resn = resnall,
-                                 nums = numsallhigh,
-                                 biolprm = truth$biolprm, fgs = truth$fgs,
-                                 CVlenage = 0.1, remove.zeroes=TRUE)
-
-#save for later use, takes a long time to generate
-saveRDS(length_censussurvsamp, file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))
-
-```
-
-Read it back in:
-
-```{r loadlengthcomp, echo=TRUE}
-
-length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))
-
-```
-
-Demo plots for other species:
-
-```{r moreplots, echo=TRUE}
-
-ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
-  filter(species == "Norwegian_ssh")
-
-lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
-  geom_bar(aes(weight = atoutput)) +
-  theme_tufte()
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, ss_length_census$natlength$species))
 
@@
 
-So this looks like our best approximation of a true length comp for a particular Atlantis run. Comparisons with standard surveys next.
+Are there fish <50 cm long at any output timestep? If not, does this make sense? There appear to be a few, such as timestep 4. Hard to see at this level.
+
+Aggregate length comp over all of these time steps (first 22+ years):
+
+```{r agglengthcomp}
+
+lfplot
+
+```
+
+This suggests that the smallest Greenland turbot in the ecosystem are ~25 cm over all of the timesteps we checked. Does this make sense?
+
+Or do we need to change something about the length comp estimation?

### 2019-05-15 RethinkSamplingFunctions.Rmd
fix typo
stage to age needed

@@
 
+## May 15: True age estimation needs to happen in here
+
+The sequence `create_survey` to `sample_fish` to `calc_age2length` now works. It can go as pictured below if there are not multiple true ages in an age class. Revisions to work on:
+
+1. For clarity, `calc_age2length` should be renamed `calc_stage2length`
+2. Between `sample_fish` and `sample_ages` we need the already written `calc_stage2age` function, which calls `calc_Z`. Test these next.
+3.
+Similarly, between `sample_fish` and `sample_weight` we will need to expand to true ages for a weight-at-age vector (for species where stage=age, this is already output as muweight from the `calc_(st)age2length` function).
+
 ```{r samplingworkflow}
@@
 
-## Proposed solution
+## Proposed solution (now done for calc_age2length)
 
@@
 1. write an aggregation function for resn and structn that applies the median
+    + **Done 10 May, `aggregateDensityData`**
 2. rewrite `create_survey` to use alternate aggregation functions, or write `create_survey_nums` and `create_survey_wt` functions that use appropriate aggregation
@@
     + so maybe `aggregateDensityData` using the standard survey polygons gets us what we need as input to `sample_fish`?
+    + **Yes, did the latter was done**
 3.
 rewrite `sample_fish` to use alternate aggregation functions for each input type.
-Keeping this all in one function makes sense because the functional already takes both numbers and weights as inputs
+Keeping this all in one function makes sense because the biological sampling already needs both numbers and weights as inputs
+    + **Done 10 May, use sample=FALSE option**

### 2019-05-15 docs/index.Rmd
looking for small turbot
stage to age needed

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html" target="_blank">Big picture and new functions needed - May 10, 2019</a>
+<a href="https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html" target="_blank">Big picture and new functions needed - May 10, 2019, updated May 15, 2019</a>
 
@@
 
+Do we see ever small fish for large species? Is this normal, or a problem with length estimation?
+
+<a href="https://sgaichas.github.io/poseidon-dev/TestSmallestLength.html" target="_blank">Should there be small Greenland halibut in NOBA? - May 15, 2019</a>
+
 ### More to come later!

### 2019-05-15 RethinkSamplingFunctions.Rmd
numbers and weights problem again

@@
 
-## May 15: True age estimation needs to happen in here
+## May 15: True age estimation needs to happen in here?
 
@@
 
-1. For clarity, `calc_age2length` should be renamed `calc_stage2length`
-2. Between `sample_fish` and `sample_ages` we need the already written `calc_stage2age` function, which calls `calc_Z`. Test these next.
+1. For clarity, `calc_age2length` should be renamed `calc_stage2length` if we are planning to use it directly between Atlantis cohorts and lengths as tested here.
+2. Between `sample_fish` and `sample_ages` we need the already written `calc_stage2age` function, which calls `calc_Z`. Test these next. But see below.
 3.
 Similarly, between `sample_fish` and `sample_weight` we will need to expand to true ages for a weight-at-age vector (for species where stage=age, this is already output as muweight from the `calc_(st)age2length` function).
+
+BUT `calc_stage2age` is designed to work on true outputs (it makes no sense to expand age classes from survey-selectivity affected output).
+In theory this should be integrated into `run_truth` but I am wary of introducing this approximation so far upstream.
+On the other hand, what choice to we have if we want true age outputs?
+We either need the direct output from the most recent codebase(s) or we have to do this approximation to get age input to stock assessment models.
+
+So, lets try integrating `calc_stage2age` into `run_truth` as planned and we can compare the outputs to what we already have for the stage data.
+We will want to compare census biomass, survey biomass, length comps...
+    + write another truth function that works on the Atlantis true age output?
+    + load YOY file into run_truth (uses load_YOY function)
+    + do we have to apply before calc_biomass_age, and sub output of calc_stage2age in for nums?
+    + if so, how do we deal with the structn and resn portions?
+
+Then we need to revise `calc_age2length` to work for true ages, as it is now hardcoded to expext 10 age classes.
+
+If we do those two things, this sampling flow remains the same, but has `calc_age2length` as the intermediate step before lengths and weights, which will come out from the true age classes.
 

### 2019-05-16 RethinkSamplingFunctions.Rmd
typo fix
add notes on weight disaggregation

@@
 We will want to compare census biomass, survey biomass, length comps...
+
     + write another truth function that works on the Atlantis true age output?
@@
     + if so, how do we deal with the structn and resn portions?
+
+    May 16: Discussion with Isaac and Patrick: how to split weights to true age bins?
+    + Simple interpolation? within a year or across years
+    + Is Z calc lagged? doesnt look like it, simple within year Z like catch curve
+    + Check whether there is a plus group in any Atlantis model--shouldn't be
+    + First try a linear interpolation between weight groups
+    + This may be more wrong for smallest fastest growing age classes
+    + If so we can write something else for younger classes
+    + Midpoint of weight for each age is the goal
 

### 2019-05-16 docs/index.Rmd
add notes on weight disaggregation

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html" target="_blank">Big picture and new functions needed - May 10, 2019, updated May 15, 2019</a>
+<a href="https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html" target="_blank">Big picture and new functions needed - May 10, 2019, updated May 16, 2019</a>
 

### 2019-05-16 TestWeightAtStage.Rmd
start looking at weight age age

---
title: "Testing atlantisom: visualize weight at (st)age output. How shall we interplolate this for true ages?"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
  'data.table', 'here', 'ggforce', 'ggthemes'
), 'packages.bib')
```

## Introduction

Now we delve into the calculation of true age classes and weight-at-age from standard Atlantis output age class, which is generally in 10 stages of maxage/10 true age classes for all but the newest Atlantis models. Note that modern Atlantis models can output true age classes in two output files (from the wiki):

>ANNAGEBIO.nc: This output provides numbers in each annual age class (so mapped from Atlantis "age class" which can contain multiple years to true annual age classes). Set flag_age_output to 1 to get this output. Tracers provided are:
>   + Numbers at age per species
>
>ANNAGECATCH.nc: This output provides numbers at annual age class (so mapped from Atlantis "age class" which can contain multiple years to true annual age classes) in the catch and discards (summed over all fleets). Set flag_age_output to 1 to get this output. Tracers provided are:
>   + Numbers at age per species in the catch
>   + Numbers at age per species in the discards

Future `atlantisom` users should take advantage of this direct output, and some `atlantisom` functions will have to be rewritten to use it.

However, The CCA model does not produce this output, and other legacy models may not, so we still need the `atlantisom` functions `calc_stage2age` and the associated `calc_Z`.  This function will be tested. However, it produces numbers at age but not weight. Weight-at-true-age is needed as an input to stock assessment models using empirical weight at age. See discussion [here](ttps://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html)

Therefore, our first step is understanding what the 10 class weight at age (output as `mulength` from `calc_age2length`) looks like for a variety of species.

For all setup, etc, please see previous files. Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html) and  [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).

This page has visualizations for the NOBA model example, CERES Global Sustainability. At the end of the file we also review saved outputs from the CCA model example, Atlantis Summit Common Scenario 1.

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)
```

```{r initialize}

initCCA <- FALSE
initNEUS <- FALSE
initNOBA <- TRUE

if(initCCA) source(here("config/CCConfig.R"))

if(initNEUS) source(here("config/NEUSConfig.R"))

if(initNOBA) source(here("config/NOBAConfig.R"))

```

```{r get_names, message=FALSE, warning=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```

```{r load_Rdata, message=FALSE, warning=FALSE}

if(initCCA) {
  truth.file <- "outputCCV3run_truth.RData"
  load(file.path(d.name, truth.file))
  truth <- result
}

if(initNEUS) {
  truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
  load(file.path(d.name, truth.file))
  truth <- result
}

if(initNOBA){
  truth.file <- "outputnordic_runresults_01run_truth.RData" 
  load(file.path(d.name, truth.file))
  truth <- result
}

```

## Simulate a survey part 5: weight at st(age) census

Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html).

We will apply examples here to just a few NOBA species to examine within-year variability:
    - Cod "North_atl_cod", likely a test assessment species
    - Herring "Norwegian_ssh", likely a test assessment species
    - Greenland halibut "Green_halibut", which grows to a large size.

To create a census, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species. The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages.

```{r census-spec, message=FALSE, warning=FALSE}

# should return a perfectly scaled survey 
effic1 <- data.frame(species=funct.group.names,
                     efficiency=rep(1.0,length(funct.group.names)))

# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
selex1 <- data.frame(species=rep(funct.group.names, each=10),
                     agecl=rep(c(1:10),length(funct.group.names)),
                     selex=rep(1.0,length(funct.group.names)*10))

# should return all model areas
boxpars <- load_box(d.name, box.file)
boxall <- c(0:(boxpars$nbox - 1))

# generalized timesteps all models
runpar <- load_runprm(d.name, run.prm.file)
noutsteps <- runpar$tstop/runpar$outputstep
stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc

timeall <- c(0:noutsteps)

# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
# for ecosystem indicator work test all species, e.g.
survspp <- funct.group.names

# for length and age groups lets just do fish and sharks
# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)

survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
                           funct.groups$GroupType %in% c("FISH", "SHARK")]

```

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just one species in this case). The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.

Because we don't want to wait 24 hours for this, we will look at only the first 112 time steps.

```{r stdsurveyNbased-3spp, echo=TRUE}

spp.name <- funct.group.names[funct.group.names %in% c("North_atl_cod",
                                                       "Norwegian_ssh",
                                                       "Green_halibut")]

# get survey nums with full (no) selectivity
survey_testNall <- create_survey(dat = truth$nums,
                                 time = c(0:111),
                                 species = spp.name,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# this one is high but not equal to total for numerous groups
effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))

# apply default sample fish as before to get numbers
numsallhigh <- sample_fish(survey_testNall, effNhigh)

# aggregate true resn per survey design
aggresnall <- aggregateDensityData(dat = truth$resn,
                                 time = c(0:111),
                                 species = spp.name,
                                 boxes = boxall)

# aggregate true structn per survey design
aggstructnall <- aggregateDensityData(dat = truth$structn,
                                 time = c(0:111),
                                 species = spp.name,
                                 boxes = boxall)

#dont sample these, just aggregate them using median
structnall <- sample_fish(aggstructnall, effNhigh, sample = FALSE)

resnall <-  sample_fish(aggresnall, effNhigh, sample = FALSE)

```

Length sample with user specified max length bin (200 cm):

```{r userset-maxlen, echo=TRUE}

length_census <- calc_age2length(structn = structnall,
                                 resn = resnall,
                                 nums = numsallhigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 maxbin = 200,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

```

We should get the upper end of Greenland halibut with a 200cm max length bin. That shouldnt matter for weight at age.

```{r vis-wtageclass-test}

wageplot <- ggplot(ss_length_census$muweight, aes(agecl)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name, ss_length_census$muweight$species))

wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")

```

How much does weight at age class vary over time?

Aggregate wt age stage over all of these time steps (first 22+ years):

```{r aggwtcomp, eval=FALSE}
#this isn't appropriate
wageplot

```

## Quick test with saved CCA census output

Here we check the CCA model (Atlantis Summit Common Scenario 1) saved length and weight census output for small species at Isaac's suggestion:

```{r loadCCAlengthcomp, echo=TRUE, eval=FALSE}

source(here("config/CCConfig.R"))

length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))

```

Sardine:
```{r plotCCAsardine, eval=FALSE}
ss_length_censussurvsamp <- length_censussurvsamp$muweight %>%
  filter(species == "Pacific_sardine")

lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name, ss_length_censussurvsamp$species))

lfplot
```

Anchovy:
```{r plotCCAanchovy, eval=FALSE}
ss_length_censussurvsamp <- length_censussurvsamp$muweight %>%
  filter(species == "Anchovy")

lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name, ss_length_censussurvsamp$species))

lfplot
```

Herring:
```{r plotCCAherring, eval=FALSE}
ss_length_censussurvsamp <- length_censussurvsamp$muweight %>%
  filter(species == "Herring")

lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,ss_length_censussurvsamp$species))

lfplot
```

How do these three look?

Do we need to change something about the length comp estimation?

### 2019-05-16 TestWeightAtStage.Rmd
improving wtage plots

@@
 
-wageplot <- ggplot(ss_length_census$muweight, aes(agecl)) +
-  geom_bar(aes(weight = atoutput)) +
+wageplot <- ggplot(length_census$muweight, aes(agecl, atoutput)) +
+  geom_point(aes(colour = time)) +
   theme_tufte() +
-  labs(subtitle = paste(scenario.name, ss_length_census$muweight$species))
+  theme(legend.position = "bottom") +
+  scale_x_discrete(limits=c(1:10)) +
+  xlab("age class") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste(scenario.name))
 
-wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
-wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
-wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
-wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
-wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
-wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
-wageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
+wageplot + facet_wrap(c("species"), scales="free_y")
 
@@
 
-How much does weight at age class vary over time?
+How much does weight at age class vary over time? Here are a couple of annual cycles:
 
-Aggregate wt age stage over all of these time steps (first 22+ years):
+```{r vis-wtageclass-test2}
 
-```{r aggwtcomp, eval=FALSE}
-#this isn't appropriate
-wageplot
+wtageyr1 <- length_census$muweight %>%
+  filter(time<5)
+
+wageplot <- ggplot(wtageyr1, aes(agecl, atoutput)) +
+  geom_line(aes(colour = factor(time))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  scale_x_discrete(limits=c(1:10)) +
+  xlab("age class") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " initial year"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
+
+```
+```{r vis-wtageclass-test3}
+
+wtageyrmid <- length_census$muweight %>%
+  filter(between(time, 50,54))
+
+wageplot <- ggplot(wtageyrmid, aes(agecl, atoutput)) +
+  geom_line(aes(colour = factor(time))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  scale_x_discrete(limits=c(1:10)) +
+  xlab("age class") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " year 10"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
+
+```
+```{r vis-wtageclass-test4}
+
+wtageyr20 <- length_census$muweight %>%
+  filter(between(time, 100,104))
+
+wageplot <- ggplot(wtageyr20, aes(agecl, atoutput)) +
+  geom_line(aes(colour = factor(time))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  scale_x_discrete(limits=c(1:10)) +
+  xlab("age class") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " year 20"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
+
+```
+
+There are some interesting jumps between age classes. Most pronounced for herring, but those are real age classes!
+
+Change in wt at age over time for age classes using an annual mid-year snapshot (survey) (first 22+ years of NOBA model run):
+
+```{r aggwtcomp}
+
+# from std survey code and model step info read in above
+midptyr <- round(median(seq(1,stepperyr)))
+annualmidyear <- seq(midptyr, noutsteps, stepperyr)
+
+wtage_annsurv <- length_census$muweight %>%
+  filter(time %in% annualmidyear)
+
+wageplot <- ggplot(wtage_annsurv, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  xlab("model timestep (5 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
 

### 2019-05-16 TestWeightAtStage.Rmd
add CCA examples

@@
 
-We should get the upper end of Greenland halibut with a 200cm max length bin. That shouldnt matter for weight at age.
+We should get the upper end of Greenland halibut with a 200cm max length bin. That shouldnt matter for weight at age. This is all timesteps for each species by agecl.
 
@@
 ```
+
 ```{r vis-wtageclass-test3}
@@
 ```
+
 ```{r vis-wtageclass-test4}
@@
 
-There are some interesting jumps between age classes. Most pronounced for herring, but those are real age classes!
+There are some interesting jumps between age classes. Most pronounced for herring. All of these species would be split into two true ages per age class. Interpolation could be interesting.
 
@@
 
+# reverse to show agecl time series of wt
 wageplot <- ggplot(wtage_annsurv, aes(time, atoutput)) +
@@
 
-Here we check the CCA model (Atlantis Summit Common Scenario 1) saved length and weight census output for small species at Isaac's suggestion:
+Here we check the CCA model (Atlantis Summit Common Scenario 1) saved length and weight census output for small species at Isaac's suggestion (all three of the following are true age classes, NumAgeClassSize=1 per Atlantis agecl bin:
 
@@
 ```{r plotCCAsardine, eval=FALSE}
-ss_length_censussurvsamp <- length_censussurvsamp$muweight %>%
+censuswt <- length_censussurvsamp$muweight %>%
   filter(species == "Pacific_sardine")
 
-lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
-  geom_bar(aes(weight = atoutput)) +
+wageplot <- ggplot(censuswt, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
   theme_tufte() +
-  labs(subtitle = paste(scenario.name, ss_length_censussurvsamp$species))
+  theme(legend.position = "bottom") +
+  xlab("model timestep (1 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
 
-lfplot
+wageplot + facet_wrap(c("species"), scales="free_y")
 ```
@@
 ```{r plotCCAanchovy, eval=FALSE}
-ss_length_censussurvsamp <- length_censussurvsamp$muweight %>%
+censuswt <- length_censussurvsamp$muweight %>%
   filter(species == "Anchovy")
 
-lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
-  geom_bar(aes(weight = atoutput)) +
+wageplot <- ggplot(censuswt, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
   theme_tufte() +
-  labs(subtitle = paste(scenario.name, ss_length_censussurvsamp$species))
+  theme(legend.position = "bottom") +
+  xlab("model timestep (1 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
 
-lfplot
+wageplot + facet_wrap(c("species"), scales="free_y")
 ```
@@
 ```{r plotCCAherring, eval=FALSE}
-ss_length_censussurvsamp <- length_censussurvsamp$muweight %>%
+censuswt <- length_censussurvsamp$muweight %>%
   filter(species == "Herring")
 
-lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
-  geom_bar(aes(weight = atoutput)) +
+wageplot <- ggplot(censuswt, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
   theme_tufte() +
-  labs(subtitle = paste(scenario.name,ss_length_censussurvsamp$species))
+  theme(legend.position = "bottom") +
+  xlab("model timestep (1 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
 
-lfplot
+wageplot + facet_wrap(c("species"), scales="free_y")
 ```
 
-How do these three look?
+How do these three look? Herring had a rough period from model years 15-30 or so.
 
-Do we need to change something about the length comp estimation?
+Now for some CCA groups with multiple true ages per stage.
+
+Pacific hake, 2 true ages per class:
+```{r plotCCAhake, eval=FALSE}
+censuswt <- length_censussurvsamp$muweight %>%
+  filter(species == "Mesopel_M_Fish")
+
+wageplot <- ggplot(censuswt, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  xlab("model timestep (1 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
+```
+
+Bocaccio rockfish, 5 true ages per age class:
+```{r plotCCAbocaccio, eval=FALSE}
+censuswt <- length_censussurvsamp$muweight %>%
+  filter(species == "Bocaccio_rockfish")
+
+wageplot <- ggplot(censuswt, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  xlab("model timestep (1 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
+```
+
+Yelloweye rockfish, most extreme at 12 true ages per age class:
+```{r plotCCAYelloweye, eval=FALSE}
+censuswt <- length_censussurvsamp$muweight %>%
+  filter(species == "Yelloweye_rockfish")
+
+wageplot <- ggplot(censuswt, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  xlab("model timestep (1 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
+```
+
+I'm not sure any of this helps us decide how to interpolate, but there is clearly contrast in weight at age (class) in both models and example runs.

### 2019-05-16 TestWeightAtStage.Rmd
helps to evaluate the plotting blocks

@@
 
-```{r loadCCAlengthcomp, echo=TRUE, eval=FALSE}
+```{r loadCCAlengthcomp, echo=TRUE}
 
@@
 Sardine:
-```{r plotCCAsardine, eval=FALSE}
+```{r plotCCAsardine}
 censuswt <- length_censussurvsamp$muweight %>%
@@
 Anchovy:
-```{r plotCCAanchovy, eval=FALSE}
+```{r plotCCAanchovy}
 censuswt <- length_censussurvsamp$muweight %>%
@@
 Herring:
-```{r plotCCAherring, eval=FALSE}
+```{r plotCCAherring}
 censuswt <- length_censussurvsamp$muweight %>%
@@
 Pacific hake, 2 true ages per class:
-```{r plotCCAhake, eval=FALSE}
+```{r plotCCAhake}
 censuswt <- length_censussurvsamp$muweight %>%
@@
 Bocaccio rockfish, 5 true ages per age class:
-```{r plotCCAbocaccio, eval=FALSE}
+```{r plotCCAbocaccio}
 censuswt <- length_censussurvsamp$muweight %>%
@@
 Yelloweye rockfish, most extreme at 12 true ages per age class:
-```{r plotCCAYelloweye, eval=FALSE}
+```{r plotCCAYelloweye}
 censuswt <- length_censussurvsamp$muweight %>%

### 2019-05-16 TestWeightAtStage.Rmd
formatting, add wtage html

@@
 ---
-title: "Testing atlantisom: visualize weight at (st)age output. How shall we interplolate this for true ages?"
+title: "Testing atlantisom: visualize weight at (st)age output. How should we interplolate this for true ages?"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
->ANNAGEBIO.nc: This output provides numbers in each annual age class (so mapped from Atlantis "age class" which can contain multiple years to true annual age classes).
+>**ANNAGEBIO.nc**: This output provides numbers in each annual age class (so mapped from Atlantis "age class" which can contain multiple years to true annual age classes).
 Set flag_age_output to 1 to get this output.
 Tracers provided are:
+>
 >   + Numbers at age per species
 >
->ANNAGECATCH.nc: This output provides numbers at annual age class (so mapped from Atlantis "age class" which can contain multiple years to true annual age classes) in the catch and discards (summed over all fleets).
+>**ANNAGECATCH.nc**: This output provides numbers at annual age class (so mapped from Atlantis "age class" which can contain multiple years to true annual age classes) in the catch and discards (summed over all fleets).
 Set flag_age_output to 1 to get this output.
 Tracers provided are:
+>
 >   + Numbers at age per species in the catch

### 2019-05-16 docs/index.Rmd
formatting, add wtage html

@@
 
-### Start age composition sampling: work in progress
+### Start age composition sampling: by output age class, not true age
 
@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/TestSmallestLength.html" target="_blank">Should there be small Greenland halibut in NOBA? - May 15, 2019</a>
+<a href="https://sgaichas.github.io/poseidon-dev/TestSmallestLength.html" target="_blank">Are length comps OK on the small end? - May 15, 2019</a>
+
+May 16: Isaac says CCA fish look reasonable.
+NOBA Greenland halibut growth parameterization was reportedly "a nightmare" with conflicting science and datasets, so maybe we don't worry too much about them right now.
+
+## Weight at age: by output age class, how to get to true age basis?
+
+<a href="https://sgaichas.github.io/poseidon-dev/TestWeightAtStage.html" target="_blank">How much variability in weight at stage? - May 16, 2019</a>
 

### 2019-05-16 InterpolateWeightAtAge.Rmd
working on function line 232

--- 2019-05-16 TestWeightAtStage.Rmd
@@
 ---
-title: "Testing atlantisom: visualize weight at (st)age output. How should we interplolate this for true ages?"
+title: "Testing atlantisom: what if we interpolate the weight at stage output to true age?"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 Now we delve into the calculation of true age classes and weight-at-age from standard Atlantis output age class, which is generally in 10 stages of maxage/10 true age classes for all but the newest Atlantis models.
+Weight-at-true-age is needed as an input to stock assessment models using empirical weight at age.
+See discussion [here](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html).
+We will test the `atlantisom` functions `calc_stage2age` and the associated `calc_Z` that estimate true numbers at age from numbers in output age class separately.
+However, `calc_stage2age` produces numbers at age but not weight.
+
+We already have 10 class weight at age (output as `mulength` from `calc_age2length`).
+We visualized weight at stage for selected NOBA and CCA species [here](https://sgaichas.github.io/poseidon-dev/TestWeightAtStage.html).
+In this page we will try the simplest interpolation to true age-based weight at age possible, and see if that will work for us.
+
 Note that modern Atlantis models can output true age classes in two output files (from the wiki):
@@
 Future `atlantisom` users should take advantage of this direct output, and some `atlantisom` functions will have to be rewritten to use it.
-
-However, The CCA model does not produce this output, and other legacy models may not, so we still need the `atlantisom` functions `calc_stage2age` and the associated `calc_Z`.
-This function will be tested.
-However, it produces numbers at age but not weight.
-Weight-at-true-age is needed as an input to stock assessment models using empirical weight at age.
-See discussion [here](ttps://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html)
-
-Therefore, our first step is understanding what the 10 class weight at age (output as `mulength` from `calc_age2length`) looks like for a variety of species.
 
@@
 
-## Simulate a survey part 5: weight at st(age) census
+## Simulate a survey part 5: interpolated weight at true age census
 
-Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html).
+Full methods to get the census weight at age by Atlantis output age class are explained [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).
 
-We will apply examples here to just a few NOBA species to examine within-year variability:
+We will apply interpolation examples here to just a few NOBA species to examine within-year variability:
     - Cod "North_atl_cod", likely a test assessment species
@@
 
-Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just one species in this case).
+Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just a few species in this case).
 The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.
@@
 
-We should get the upper end of Greenland halibut with a 200cm max length bin. That shouldnt matter for weight at age. This is all timesteps for each species by agecl.
-
-```{r vis-wtageclass-test}
-
-wageplot <- ggplot(length_census$muweight, aes(agecl, atoutput)) +
-  geom_point(aes(colour = time)) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  scale_x_discrete(limits=c(1:10)) +
-  xlab("age class") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste(scenario.name))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-
-```
-
-How much does weight at age class vary over time? Here are a couple of annual cycles:
-
-```{r vis-wtageclass-test2}
-
-wtageyr1 <- length_census$muweight %>%
-  filter(time<5)
-
-wageplot <- ggplot(wtageyr1, aes(agecl, atoutput)) +
-  geom_line(aes(colour = factor(time))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  scale_x_discrete(limits=c(1:10)) +
-  xlab("age class") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " initial year"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-
-```
-
-```{r vis-wtageclass-test3}
-
-wtageyrmid <- length_census$muweight %>%
-  filter(between(time, 50,54))
-
-wageplot <- ggplot(wtageyrmid, aes(agecl, atoutput)) +
-  geom_line(aes(colour = factor(time))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  scale_x_discrete(limits=c(1:10)) +
-  xlab("age class") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " year 10"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-
-```
-
-```{r vis-wtageclass-test4}
-
-wtageyr20 <- length_census$muweight %>%
-  filter(between(time, 100,104))
-
-wageplot <- ggplot(wtageyr20, aes(agecl, atoutput)) +
-  geom_line(aes(colour = factor(time))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  scale_x_discrete(limits=c(1:10)) +
-  xlab("age class") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " year 20"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-
-```
-
-There are some interesting jumps between age classes. Most pronounced for herring. All of these species would be split into two true ages per age class. Interpolation could be interesting.
+We should get the upper end of Greenland halibut with a 200cm max length bin.
+That shouldnt matter for weight at age.
+All of these species would be split into two true ages per age class.
+Interpolation could be interesting.
 
@@
 
+If we were to do simple interpolation on this output, we can approximate weight at true age class:
+
+```{r interpolate-wtage}
+# a function to go into atlantisom, sketched here, reuse calc_stage2age as possible
+calc_wtTrueage <- function(wtagecl_data, fgs.file) {
+  # get number of ages per agecl from fgs.file
+  # match species in wtagecl_data to those species in the fgs file
+  fgs <- load_fgs(fgs.file)
+  species.code <- fgs$Code
+  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
+  names <- turnedon$Code
+
+  # Figure out the groups that have multiple ages classes in each stage (or
+  # cohort)
+  multiple_ages <- turnedon[turnedon$NumAgeClassSize>1, c(1,4,10)]
+  num_multi_age <- dim(multiple_ages)[1]
+
+  ntimesteps <- length(unique(wtagecl_data$time))
+
+  # find weight increment for a timestep
+  # muweight for agecl+1 - muweight for agecl
+
+  # divide weight increment by number of ages per agecl (could decay with increasing age, but how different from equal increment?)
+
+  # recursive: add increment to agecl until hitting agecl+1
+
+  # reunumber agecl 1 to 10*number of ages per agecl
+
+  # output full set of muweight at newly created agecl
+
+}
+```
+
+Plot interpolated results for NOBA species:
+
+```{r interp-wtage-test1}
+
+```
+
 ## Quick test with saved CCA census output
@@
 ```
+
+```{r apply-inter-pwtage-CCA}
+
+```
+
+Plots for species with 1 true age per agecl should not be different:
 
@@
 
-How do these three look? Herring had a rough period from model years 15-30 or so.
+How do these three look? Identical I hope.
 
-Now for some CCA groups with multiple true ages per stage.
+Now for some CCA groups with multiple true ages interpolated per stage.
 
-Pacific hake, 2 true ages per class:  
+Pacific hake, 2 true ages per class (20 total):  
 ```{r plotCCAhake}
@@
 
-Bocaccio rockfish, 5 true ages per age class:
+Bocaccio rockfish, 5 true ages per age class (50 total):
 ```{r plotCCAbocaccio}
@@
 
-Yelloweye rockfish, most extreme at 12 true ages per age class:
+Yelloweye rockfish, most extreme at 12 true ages per age class (120 total):
 ```{r plotCCAYelloweye}
@@
 
-I'm not sure any of this helps us decide how to interpolate, but there is clearly contrast in weight at age (class) in both models and example runs.
+Yelloweye should be a good test...

### 2019-05-19 InterpolateWeightAtAge.Rmd
interpolation function part1

@@
 # a function to go into atlantisom, sketched here, reuse calc_stage2age as possible
+
+fgs.dir <- here("atlantisoutput", "NOBACERESGlobalSustainability")
+fgs.file <-  "nordic_groups_v04.csv"
+wtagecl_data <- wtage_annsurv
+
 calc_wtTrueage <- function(wtagecl_data, fgs.file) {
   # get number of ages per agecl from fgs.file
-  # match species in wtagecl_data to those species in the fgs file
-  fgs <- load_fgs(fgs.file)
+  fgs <- load_fgs(fgs.dir, fgs.file)
   species.code <- fgs$Code
   turnedon <- fgs[fgs$IsTurnedOn > 0, ]
-  names <- turnedon$Code
 
   # Figure out the groups that have multiple ages classes in each stage (or
-  # cohort)
+  # cohort) this is from the full model
   multiple_ages <- turnedon[turnedon$NumAgeClassSize>1, c(1,4,10)]
+
+  # match species in wtagecl_data to those species in the fgs file
+  sppwt <- levels(wtagecl_data$species)
+  multiple_ages <- multiple_ages[multiple_ages$Name %in% sppwt, ]
+
+  #now get indices
+  names <- multiple_ages$Code
   num_multi_age <- dim(multiple_ages)[1]
 
+  # timesteps from input wt file
   ntimesteps <- length(unique(wtagecl_data$time))

### 2019-05-20 TestSmallestLength.Rmd

@@
 
-Or do we need to change something about the length comp estimation?
+## Quick test with saved CCA census output
+
+Here we check the CCA model (Atlantis Summit Common Scenario 1) saved length census output for small species at Isaac's suggestion:
+
+```{r loadCCAlengthcomp, echo=TRUE}
+
+source(here("config/CCConfig.R"))
+
+length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))
+
+```
+
+Sardine:
+```{r plotCCAsardine}
+ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
+  filter(species == "Pacific_sardine")
+
+lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, ss_length_censussurvsamp$species))
+
+lfplot
+```
+
+Anchovy:
+```{r plotCCAanchovy}
+ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
+  filter(species == "Anchovy")
+
+lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, ss_length_censussurvsamp$species))
+
+lfplot
+```
+
+Herring:
+```{r plotCCAherring}
+ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
+  filter(species == "Herring")
+
+lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,ss_length_censussurvsamp$species))
+
+lfplot
+```
+
+How do these three look?
+
+Do we need to change something about the length comp estimation?

### 2019-05-20 visualization_Rmd/CCAStdSurvLengthCompTest.Rmd

@@
 
-atf_numsstd <- numsstd[numsstd$species == "Arrowtooth_flounder",]
-atf_structnstd <- structnstd[structnstd$species == "Arrowtooth_flounder",]
-atf_resnstd <- resnstd[resnstd$species == "Arrowtooth_flounder",]
+ss_numsstd <- numsstd[numsstd$species == "Arrowtooth_flounder",]
+ss_structnstd <- structnstd[structnstd$species == "Arrowtooth_flounder",]
+ss_resnstd <- resnstd[resnstd$species == "Arrowtooth_flounder",]
 
-atf_length_stdsurv <- calc_age2length(structn = atf_structnstd,
-                                 resn = atf_resnstd,
-                                 nums = atf_numsstd,
+ss_length_stdsurv <- calc_age2length(structn = ss_structnstd,
+                                 resn = ss_resnstd,
+                                 nums = ss_numsstd,
                                  biolprm = truth$biolprm, fgs = truth$fgs,
@@
 
-lfplot <- ggplot(atf_length_stdsurv$natlength, aes(upper.bins)) +
+lfplot <- ggplot(ss_length_stdsurv$natlength, aes(upper.bins)) +
   geom_bar(aes(weight = atoutput)) +
   theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=paste(scenario.name, atf_length_stdsurv$natlength$species))
+  labs(subtitle = paste(scenario.name, ss_length_stdsurv$natlength$species))
 

### 2019-05-20 visualization_Rmd/NOBAStdSurvLengthCompTest.Rmd
check gh lf in NOBA

@@
 
+Sampling looks really different for greenland halibut than census did. Check this!
+
+I'm using a shiny app for checking, which I'm not sure I want to launch from the html, so this code block is not evaluated. Can do interactively instead. It takes a few minutes to load the info:
+
+```{r check-sampling, echo=TRUE, eval=FALSE}
+# lets visualize which boxes we sampled and where greenland halibut are in NOBA
+# use https://github.com/Atlantis-Ecosystem-Model/shinyrAtlantis
+# I've already installed it, see readme for instructions
+library(shinyrAtlantis)
+obj <- make.sh.prm.object(file.path(d.name, box.file), 
+                          file.path(d.name, functional.groups.file), 
+                          file.path(d.name, biol.prm.file))
+sh.prm(obj)
+
+```
+
+Our survey uses boxes 2:47 as parameterized here (somewhat randomly for testing). This is the NOBA map (copied from the shiny app):
+
+```{r NOBAmap}
+knitr::include_graphics("images/NOBAboxes.png")
+```
+
+Greenland halibut adults have a fairly high fraction in box 1, which we are missing with our survey, and we miss 48:57 where there are high (but not the highest) fraction of juveniles (screenshot of shiny app):
+
+```{r GHRdist}
+knitr::include_graphics("images/GRHdistNOBA.png")
+```
+
+Greenland halibut have 2 years per output age class. Our selectivity acts on the age class. We have them in the demersals group, so the selectivity curve is:
+
+```{r sel-demersal}
+plot(seldemersals$agecl, seldemersals$selex)
+```
+
+So we won't see any fish until they are almost 10, when they are pretty big, and we fully sample the larger fish.
+This combined with missing some juvenile areas may explain the strange shapes of the sampled length frequencies.
+
+Here is the same survey missing the areas as described, but with full selectivity for all greenland halibut age classes:
+
+```{r stdsurveyNbased-GHR, echo=TRUE}
+
+ss.name <- funct.group.names[funct.group.names == "Green_halibut"]
+
+selex1 <- data.frame(species=rep(funct.group.names, each=10),
+                     agecl=rep(c(1:10),length(funct.group.names)),
+                     selex=rep(1.0,length(funct.group.names)*10))
+
+# get survey nums with full (no) selectivity
+ss_survey_testNstd_nosel <- create_survey(dat = truth$nums,
+                                 time = annualmidyear,
+                                 species = ss.name,
+                                 boxes = boxsurv,
+                                 effic = efficmix,
+                                 selex = selex1)
+
+# now sample fish nums from this
+ss_numsstd_nosel <- sample_fish(ss_survey_testNstd_nosel, effNmix)
+
+# aggregate resn and struct n haven't changed because selectivity doesn't apply
+# use the ones made above in the calc call with new nums
+
+ss_length_stdsurv_nosel <- calc_age2length(structn = ss_structnstd,
+                                 resn = ss_resnstd,
+                                 nums = ss_numsstd_nosel,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+```
+
+This should give the difference between missing key survey strata and survey selectivity, here set to capture all age classes perfectly:
+
+```{r sslengthsamp2-test}
+
+lfplot <- ggplot(ss_length_stdsurv_nosel$natlength, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, ss_length_stdsurv_nosel$natlength$species))
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")
+
+```
+
+We may want to extend the length bins for something like greenland halibut, which get larger than the hardcoded largest bin in `calc_age2length`, 150 cm.
+Both census and these sampled length comps are chopped off at 150 cm for this species.
+
+I will likely change the `calc_age2length` code to pass the upper bin of 150 as a default but allow the user to change it in the function call.
+
 ## Samples for all species and SAVE the file
@@
 
-```{r allspplengthsamp, echo=TRUE}
+```{r allspplengthsamp, echo=TRUE, eval=FALSE}
 

### 2019-05-20 Atlantisom2SSworkflowtest.Rmd
start file to test new CC outputs and input to SS

--- 2019-04-30 TrueBioTest.Rmd
@@
 
-Initial testing was conducted by S.
-Gaichas using R scripts in the R folder of this repository that are titled "PoseidonTest_[whatwastested].R".
-Initial tests are expanded and documented in more detail in these pages.
-C.
-Stawitz improved and streamlined the setup and intialization sections.
+On this page we demonstrate use of atlantisom on the California Current output files to provide inputs for a stock assessment model implemented in Stock Synthesis [@SSREF].
 
@@
 
-In this example, we have a local folder "atlantisoutput" with three subdirectories:
+In this example, we have a local folder "atlantisoutput" and the output of interest in a subdirectory:
 
-  * "CalCurrentSummitScenario1": California Current Atlantis model output
-  * "NEUStest20160303": older (broken) Northeast US Atlantis model output
-  * "NOBACERESGlobalSustainability": Norwegian-Barents Sea Atlantis model output
+  * "CalCurrentClimateTest1": California Current Atlantis model output from Isaac Kaplan's google drive \outputFolderAllOff2013Oceanography using the Hodgson et al. 2018 configuration.
 

### 2019-05-20 CreateStockSynthesis.Rmd

---
title: "WritingSSfiles"
author: "Christine Stawitz"
date: "May 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
require(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/chris/Documents/GitHub/atlantisom/")
```

## Introduction

This page discusses reading in an existing stock assessment file and replacing the data values with those from the `atlantisom` package. To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`

```{r packages, eval=FALSE}
devtools::install_github("r4ss/r4ss", dependencies = FALSE)
```

## Sardine example

Dummy files are in the package directory; use `r4ss` to load the dummy data file

```{r readdata, echo=FALSE}
require(r4ss)
#Read in sardine data file
sardine.data <- r4ss::SS_readdat_3.30("./inst/extdata/Sardine_SS_files/sardEM_3_3.dat")
```

Below, we can see the structure of the catch and CPUE data
```{r}
sardine.data$catch

sardine.data$CPUE
```

We can next look at age composition data
```{r}
sardine.data$agecomp

sardine.data$lencomp
```

We may also need to access or modify the setup variables before the data matrices.

```{r}
#CPUE units, SD and error structure
sardine.data$CPUEinfo

#Number of length bins
sardine.data$N_lbins

#Length bin method
sardine.data$Lbin_method

#Vector of age bins
sardine.data$agebin_vector

```

### 2019-05-20 Atlantisom2SSworkflowtest.Rmd
workflow with first CC climate run

@@
 
-  * "CalCurrentClimateTest1": California Current Atlantis model output from Isaac Kaplan's google drive \outputFolderAllOff2013Oceanography using the Hodgson et al. 2018 configuration.
+  * "CalCurrent2013_OA_off": California Current Atlantis model output from Isaac Kaplan's google drive \outputFolderAllOff2013Oceanography using the Hodgson et al.
+2018 configuration.
+Corresponds to CC1Config.R file.
 
@@
 
-initCCA <- FALSE
+initCCA <- TRUE
 initNEUS <- FALSE
-initNOBA <- TRUE
+initNOBA <- FALSE
 
-if(initCCA){
-  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
-  functional.groups.file <- "CalCurrentV3Groups.csv"
-  biomass.pools.file <- "DIVCalCurrentV3_BIOL.nc"
-  biol.prm.file <- "CalCurrentV3_Biol.prm"
-  box.file <- "CalCurrentV3_utm.bgm"
-  initial.conditions.file <- "DIVCalCurrentV3_BIOL.nc"
-  run.prm.file <- "CalCurrentV3_run.xml"
-  scenario.name <- "CCV3"
-}
+#function to make a config file? need one for each atlantis run
 
-if(initNEUS){
-  d.name <- here("atlantisoutput","NEUStest20160303")
-  functional.groups.file <- "NeusGroups.csv" 
-  biomass.pools.file <- ""
-  biol.prm.file <- "at_biol_neus_v15_DE.prm"
-  box.file <- "neus30_2006.bgm"
-  initial.conditions.file <- "inneus_2012.nc"
-  run.prm.file <- "at_run_neus_v15_DE.xml"
-  scenario.name <- "neusDynEffort_Test1_"
-}
+if(initCCA) source(here("config/CC1Config.R"))
 
-if(initNOBA){
-  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
-  functional.groups.file <- "nordic_groups_v04.csv" 
-  biomass.pools.file <- "nordic_biol_v23.nc"
-  biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
-  box.file <- "Nordic02.bgm"
-  initial.conditions.file <- "nordic_biol_v23.nc"
-  run.prm.file <- "nordic_run_v01.xml"
-  scenario.name <- "nordic_runresults_01"
-}
-# NOBA note: output filenames in CCA and NEUS begin with "output" and the run_truth function is written to expect this.
-Need to check if default Atlantis output file nomenclature has changed or if NOBA is a special case.
-For now, NOBA filenames have been changed to include prefix "output"
+if(initNEUS) source(here("config/NEUSConfig.R"))
+
+if(initNOBA) source(here("config/NOBAConfig.R"))
+
 ```
@@
 
-```{r get_truth, message=FALSE, warning=FALSE, eval=FALSE}
+```{r get_truth, message=FALSE, warning=FALSE}
+
+# default run_truth setup will save the file, so check for that first
+
+if(!file.exists(file.path(d.name, 
+                          paste0("output", scenario.name, "run_truth.RData")))){
   #Store all loaded results into an R object
-results <- run_truth(scenario = scenario.name,
+  truth <- run_truth(scenario = scenario.name,
                      dir = d.name,
@@
   )
-
-if(initCCA) CCAresults <- results
-if(initNEUS) NEUSresults <- results
-if(initNOBA) NOBAresults <- results
-
+} else{
+  truth <- get(load(file.path(d.name,
+                              paste0("output", scenario.name, "run_truth.RData"))))
+}
 ```
 
-Now the R object `results` with the comprehensive results from the Atlantis model has been read in.
+Now the R object `truth` with the comprehensive results from the Atlantis model has been read in.
 It is also saved as an .RData file titled "output[scenario.name]run_truth.RData" in the directory with the model output, so that later analyses can use `base::load()` instead of taking the time to rerun `atlantisom::run_truth()`.
@@
 Therefore, reading the .RData file in as below is advised after the first run to make the plots comparing systems later in this document.
-
-```{r load_Rdata, message=FALSE, warning=FALSE}
-
-if(initCCA) {
-  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
-  truth.file <- "outputCCV3run_truth.RData"
-  load(file.path(d.name, truth.file))
-  CCAresults <- result
-}
-
-if(initNEUS) {
-  d.name <- here("atlantisoutput","NEUStest20160303")
-  truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
-  load(file.path(d.name, truth.file))
-  NEUSresults <- result
-}
-
-if(initNOBA){
-  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
-  truth.file <- "outputnordic_runresults_01run_truth.RData" 
-  load(file.path(d.name, truth.file))
-  NOBAresults <- result
-}
-
-```
 
@@
 ```{r census-spec, message=FALSE, warning=FALSE}
+
+# make a function for this
 
@@
 
-# these are model specific, generalized above
-# if(initCCA) boxall <- c(0:88) 
-# if(initNEUS) boxall <- c(0:29)
-# if(initNOBA) boxall <- c(0:59)
+# generalized timesteps all models
+runpar <- load_runprm(d.name, run.prm.file)
+noutsteps <- runpar$tstop/runpar$outputstep
+stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
 
-# should return all model output timesteps; need to generalize
-if(initCCA) timeall <- c(0:100) 
-if(initNEUS) timeall <- c(0:251)
-if(initNOBA) timeall <- c(0:560)
+timeall <- c(0:noutsteps)
 
@@
 
-# to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
-if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
-if(initNEUS) survspp <- funct.group.names[1:21]
-if(initNOBA) survspp <- funct.group.names[1:36]
+# for length and age groups lets just do fish and sharks
+# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
+if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)
+
+survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
+                           funct.groups$GroupType %in% c("FISH", "SHARK")]
 
@@
 
-# this uses result$biomass_ages to sample biomass directly
+# this uses result$biomass_ages to sample biomass directly, no need for wt@age est
 
-if(initCCA) datB <- CCAresults$biomass_ages
-if(initNEUS) datB <- NEUSresults$biomass_ages
-if(initNOBA) datB <- NOBAresults$biomass_ages
-
-survey_testBall <- create_survey(dat = datB,
+survey_testBall <- create_survey(dat = truth$biomass_ages,
                                  time = timeall,
@@
 
-if(initCCA) CCAsurveyB_frombio <- surveyB_frombio
-if(initNEUS) NEUSsurveyB_frombio <- surveyB_frombio
-if(initNOBA) NOBAsurveyB_frombio <- surveyB_frombio
-
 ```
@@
 # read Atlantis output files
-if(initCCA) {
-  atBtxt2 <- read.table(here("atlantisoutput","CalCurrentSummitScenario1","outputCCV3BiomIndx.txt"), header=T)
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  surveyB_frombio <- CCAsurveyB_frombio
-}
-
-if(initNEUS) {
-  atBtxt2 <- read.table(here("atlantisoutput","NEUStest20160303","neusDynEffort_Test1_BiomIndx.txt"), header=T)
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  surveyB_frombio <- NEUSsurveyB_frombio
-}
-
-if(initNOBA) {
-  atBtxt2 <- read.table(here("atlantisoutput","NOBACERESGlobalSustainability","outputnordic_runresults_01BiomIndx.txt"), header=T) 
-  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
-  surveyB_frombio <- NOBAsurveyB_frombio
-}
+  atBtxt2 <- read.table(here(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)
 
@@
 # WARNING hardcoded for output with last species group as DIN
-groupslookup <- groupslookup %>%
+groupslookup <- funct.groups %>%
   filter(IsTurnedOn > 0)
@@
   facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free") 
-
-```
-
-```{r matchBp2, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
 compareB + 
   facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free") 
-
-```
-
-```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
 compareB + 
   facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free") 
-
-```
-
-```{r matchBp4, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species.
-Depending on the model, we may not show all species in this example."}
 compareB + 
@@
 
-After comparing the survey census based on the biomass output of `run_truth`, we now look at a biomass index that is estimated based on numbers in the survey census and an average weight at age.
-This is more complex, but may be closer to the way some assessments handle survey observations.
+After comparing the survey census based on the biomass output of `run_truth`, we now look at a biomass index that is estimated based on numbers in the survey census for assessments that take numbers input directly.
+We need to skip the average weight part of `sample_survey_biomass` but keep the rest.
+Hence the new function `sample_survey_numbers`.
 
-```{r surveyNbased, fig.cap="Check weight at age generated from biomass_age$atoutput divided by nums$atoutput for up to 48 groups.", fig.show="hold"}
+```{r surveyNbased}
 
-# this uses result$nums so will need a weight at age to get biomass
+# this uses result$nums and a new function to get survey index in numbers (abundance)
 
-if(initCCA) datN <- CCAresults$nums
-if(initNEUS) datN <- NEUSresults$nums
-if(initNOBA) datN <- NOBAresults$nums
-
-survey_testNall <- create_survey(dat = datN,
+survey_testNall <- create_survey(dat = truth$nums,
                                  time = timeall,
@@
 
-# now we need a weight at age to convert numbers to weight
-# if I do biomass_age$atoutput divided by nums$atoutput I should have wt@age, yes?
-
-nums_test <- with(datN, aggregate(atoutput, list(species, agecl), sum))
-names(nums_test) <- c("species", "agecl", "nums")
-
-bio_test  <- with(datB, aggregate(atoutput, list(species, agecl), sum))
-names(bio_test) <- c("species", "agecl", "wt")
-
-calcwtage <- merge(bio_test, nums_test) %>%
-  mutate(wtAtAge=wt/nums)
-
-wtagecheck <- ggplot(calcwtage, aes(x=agecl, y=wtAtAge)) +
-  geom_point() + 
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-wtagecheck + 
-  facet_wrap_paginate(~species, ncol=4, nrow=4, page=1, scales="free")
-
-wtagecheck + 
-  facet_wrap_paginate(~species, ncol=4, nrow=4, page=2, scales="free")
-
-wtagecheck + 
-  facet_wrap_paginate(~species, ncol=4, nrow=4, page=3, scales="free")
-
-# variable name needs to be wtAtAge or doesnt work
-wtage <- calcwtage %>%
-  select(species, agecl, wtAtAge)
-
-surveyB_fromN <- sample_survey_biomass(survey_testNall, surv_cv, wtage)
-#save for later use, takes a long time to generate
-saveRDS(surveyB_fromN, file.path(d.name, paste0(scenario.name, "surveyBfromNcensus.rds")))
-
-if(initCCA) CCAsurveyB_fromN <- surveyB_fromN
-if(initNEUS) NEUSsurveyB_froN <- surveyB_fromN
-if(initNOBA) NOBAsurveyB_fromN <- surveyB_fromN
+surveyN <- sample_survey_numbers(survey_testNall, surv_cv)
 
@@
 
-Now we plot the results of the numbers-based (census) survey biomass against the same Atlantis output file "[modelscenario]BiomIndx.txt" as above.
+The next steps of sampling can get us a proper weight at age.
+Here again a census to generate length comps, which need an average weight at age and carry it forward.
+First we `sample_fish` (nums and weight components):
 
-```{r matchBfromN, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; first 9 species." }
+```{r biolsampling}
 
-if(initCCA) surveyB_fromN <- CCAsurveyB_fromN
-if(initNEUS) surveyB_froN <- NEUSsurveyB_fromN
-if(initNOBA) surveyB_fromN <- NOBAsurveyB_fromN
+# We need an effective N for the sample fish function, setting equal to actual N
+# this one is high but not equal to total for numerous groups
+effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
+
+# apply default sample fish as before to get numbers
+numsallhigh <- sample_fish(survey_testNall, effNhigh)
+
+# aggregate true resn per survey design
+aggresnall <- aggregateDensityData(dat = truth$resn,
+                                 time = c(0:111),
+                                 species = spp.name,
+                                 boxes = boxall)
+
+# aggregate true structn per survey design
+aggstructnall <- aggregateDensityData(dat = truth$structn,
+                                 time = c(0:111),
+                                 species = spp.name,
+                                 boxes = boxall)
+
+#dont sample these, just aggregate them using median
+structnall <- sample_fish(aggstructnall, effNhigh, sample = FALSE)
+
+resnall <-  sample_fish(aggresnall, effNhigh, sample = FALSE)
+
+```
+
+Then we calculate lengths with `calc_age2length`
+
+```{r lengthcomp-wtage}
+
+length_census <- calc_age2length(structn = structnall,
+                                 resn = resnall,
+                                 nums = numsallhigh,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 maxbin = 200,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+```
+
+Now we can test the full `sample_survey_biomass` from nums to biomass index:
+
+```{r compare-surveys}
+# apply the proper weight at age (from saved census calc_age2length) to N survey
+
+wtage <- length_census$muweight %>%
+  select(species, agecl, wtAtAge = atoutput)
+
+surveyB_fromN <- sample_survey_biomass(survey_testNall, surv_cv, wtage)
+
+```
+
+Now we plot the results of the numbers-based (census) survey against the same Atlantis output file "[modelscenario]BiomIndx.txt" as above.
+
+```{r matchB_Nsurveys }
 
@@
   facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free") 
-
-```
-
-```{r matchBfromN2, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}
-
 compareB_N + 
   facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free") 
-
-```
-
-```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}
-
 compareB_N + 
   facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free") 
-
-```
-
-```{r matchBfromN4, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species.Depending on the model, we may not show all species in this example."}
-
 compareB_N + 
@@
 
-We have more misses in the numbers with average weight estimation, so this needs more investigation and discussion.
-It may be best to have the survey index for testing derived directly from the true biomass output.
+Start testing fishery outputs next
 

### 2019-05-20 CreateStockSynthesis.Rmd

@@
 ```
+
+## Writing index data
+To edit index data, use the `atlantisom` sampling functions to extract survey biomass and numbers. These can then be written using the `SS_write_CPUE` function.
+
+```{r}
+results <- get(load("CalCurrentResults.RData"))
+devtools::load_all()
+species=c("Pacific_sardine")
+
+#Sample all boxes
+boxes <- unique(results$nums$polygon)
+
+#Efficiency of the survey
+effic <- data.frame(species=species, efficiency=0.5)
+
+# Assume uniform selectivity
+sel<- data.frame(species=rep(species,10),
+               agecl=c(1:10),
+              selex=rep(1,10))
+
+#Sample survey
+survey_out <- create_survey(dat=results$nums, time=seq(1,99,1), species=species, boxes=boxes, effic=effic, selex=sel)
+
+#Need to replace with interp function
+wtAtAge <- data.frame(species=rep("Pacific_sardine",10),
+                agecl=1:10,
+                wtAtAge=c(0.0542, 0.0837,	0.1103,	0.1323,	0.1497,	0.163,	0.1729,	0.1801,	0.1854,	0.1941))
+cv <- data.frame(species="Pacific_sardine", cv=c(0.2))
+
+#Sample survey biomass
+survObsBiom <- sample_survey_biomass(dat=survey_out,cv=cv,wtAtAge)
+
+#Sample survey numbers
+survObsNum <- sample_survey_numbers(dat=survey_out,cv=cv)
+nyears <- 50
+burnin <- 30
+
+#Test writing CPUE in biomass
+sardine.data <- SS_write_CPUE(ss_data_list = sardine.data,
+              cpue_data = list(survObsBiom$atoutput[burnin:(burnin+nyears-1)]),
+              CVs = c(0.2),
+              data_years = c(nyears),
+              sampling_month = list(rep(1,nyears)),
+              units = "biomass")
+
+#Test writing CPUE in numbers
+sardine.data <- SS_write_CPUE(ss_data_list = sardine.data,
+                             cpue_data = list(survObsNum$atoutput[burnin:(burnin+nyears-1)]),
+                             CVs = c(0.2),
+                             data_years = c(nyears),
+                             sampling_month = list(rep(1,nyears)),
+                             units = "numbers")
+
+sardine.data$CPUE
+```

### 2019-05-20 Atlantisom2SSworkflowtest.Rmd
try lf with 3 species, still too big

@@
 
-Note: on Sarah's laptop, CCA took ~40 minutes to return from `run_truth` and NOBA took an hour and 26 minutes to return.
+Note: on Sarah's laptop, CCA took ~2.5 hours to return from `run_truth` and NOBA took an hour and 26 minutes to return.
 Therefore, reading the .RData file in as below is advised after the first run to make the plots comparing systems later in this document.
@@
 
-```{r surveyBbased}
+```{r surveyBbased, eval=F}
 
@@
 
-```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species.", message=FALSE, warning=FALSE}
+```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 36 species.", message=FALSE, warning=FALSE}
 # plot some comparisons with Atlantis output
@@
 # read Atlantis output files
-  atBtxt2 <- read.table(here(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)
+atBtxt2 <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)
+
+surveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
 
@@
 
-#all species comparison, time intervals hardcoded for NEUS and NOBA
+#all species comparison, time intervals hardcoded for 5 steps per year
 compareB <-ggplot() +
@@
 
-```{r surveyNbased}
+```{r surveyNbased, eval=F}
 
@@
 
+#save for later use, takes a long time to generate
+saveRDS(surveyN, file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))
+
+```
+
+We get true numbers at age at each output step by appling `sample_fish` to the survey_testNall result if the model group has one true age per output age class:
+
+```{r truenatagecl, eval=FALSE}
+# We need an effective N for the sample fish function, setting equal to actual N
+# this one is high but not equal to total for numerous groups
+effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
+
+numsallhigh <- sample_fish(survey_testNall, effNhigh)
+
+#save for later use, takes a long time to generate
+saveRDS(numsallhigh, file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
+
 ```
@@
 Here again a census to generate length comps, which need an average weight at age and carry it forward.
-First we `sample_fish` (nums and weight components):
+First we `sample_fish` (nums and weight components).
 
-```{r biolsampling}
+On a full model run this takes far too long with all species, so we cut it down to a few assessment species here:
+
+```{r ss-biolsampling}
+
+# get only assessed species, hardcoded here for CCA sardine, hake, bocaccio
+spp.name <- funct.group.names[funct.group.names %in% c("Pacific_sardine",
+                                                       "Mesopel_M_Fish",
+                                                       "Bocaccio_rockfish")]
+
+survey_ssN <- create_survey(dat = truth$nums,
+                                 time = timeall,
+                                 species = spp.name,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
 
@@
 # apply default sample fish as before to get numbers
-numsallhigh <- sample_fish(survey_testNall, effNhigh)
+numssshigh <- sample_fish(survey_ssN, effNhigh)
 
 # aggregate true resn per survey design
-aggresnall <- aggregateDensityData(dat = truth$resn,
-                                 time = c(0:111),
+aggresnss <- aggregateDensityData(dat = truth$resn,
+                                 time = timeall,
                                  species = spp.name,
@@
 # aggregate true structn per survey design
-aggstructnall <- aggregateDensityData(dat = truth$structn,
-                                 time = c(0:111),
+aggstructnss <- aggregateDensityData(dat = truth$structn,
+                                 time = timeall,
                                  species = spp.name,
@@
 #dont sample these, just aggregate them using median
-structnall <- sample_fish(aggstructnall, effNhigh, sample = FALSE)
+structnss <- sample_fish(aggstructnss, effNhigh, sample = FALSE)
 
-resnall <-  sample_fish(aggresnall, effNhigh, sample = FALSE)
+resnss <-  sample_fish(aggresnss, effNhigh, sample = FALSE)
 
@@
 
-Then we calculate lengths with `calc_age2length`
+Then we calculate lengths with `calc_age2length`. This will probably not work for the full run with all species outputs 5x per year, but should be ok with 3 species.
 
@@
 
-length_census <- calc_age2length(structn = structnall,
-                                 resn = resnall,
-                                 nums = numsallhigh,
+length_census_ss <- calc_age2length(structn = structnss,
+                                 resn = resnss,
+                                 nums = numssshigh,
                                  biolprm = truth$biolprm, fgs = truth$fgs,
-                                 maxbin = 200,
                                  CVlenage = 0.1, remove.zeroes=TRUE)
@@
 
-Now we can test the full `sample_survey_biomass` from nums to biomass index:
+```{r savelengths}
+#save for later use, takes a long time to generate
+saveRDS(length_census_ss, file.path(d.name, paste0(scenario.name, "length_census_sardhakeboca.rds")))
+```
+
+Now we can test the full `sample_survey_biomass` from nums to biomass index with these three species:
 
@@
 
-wtage <- length_census$muweight %>%
+length_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "length_census_sardhakeboca.rds")))
+
+wtage <- length_census_ss$muweight %>%
   select(species, agecl, wtAtAge = atoutput)
 
-surveyB_fromN <- sample_survey_biomass(survey_testNall, surv_cv, wtage)
+surveyB_fromN <- sample_survey_biomass(survey_ssN, surv_cv, wtage)
 
@@
 
-#all species comparison, time intervals hardcoded for NEUS and NOBA
+atBtxt2tidy <- atBtxt2 %>%
+  select(Time:DIN) %>%
+  #select(Time, FPL:DIN) %>%
+  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
+  gather(species, biomass, -Time) %>%
+  filter(species %in% levels(surveyB_fromN$species))
+
+#all species comparison, time intervals hardcoded for 5 steps per year
 compareB_N <-ggplot() +

### 2019-05-21 Atlantisom2SSworkflowtest.Rmd
combine blocks save lf

@@
 
-```
-
-```{r savelengths}
 #save for later use, takes a long time to generate
 saveRDS(length_census_ss, file.path(d.name, paste0(scenario.name, "length_census_sardhakeboca.rds")))
+
 ```

### 2019-05-21 Atlantisom2SSworkflowtest.Rmd
dont eval lf
need SS ref

@@
 
-On this page we demonstrate use of atlantisom on the California Current output files to provide inputs for a stock assessment model implemented in Stock Synthesis [@SSREF].
+On this page we demonstrate use of atlantisom on the California Current output files to provide inputs for a stock assessment model implemented in Stock Synthesis [insert SSREF].
 
@@
 
-```{r lengthcomp-wtage}
+```{r lengthcomp-wtage, eval=FALSE}
 

### 2019-05-21 Atlantisom2SSworkflowtest.Rmd
fixed wtage and N to B test

@@
 
-```{r surveyNbased, eval=F}
+```{r surveyNbased, eval=FALSE}
 
@@
 
-Then we calculate lengths with `calc_age2length`. This will probably not work for the full run with all species outputs 5x per year, but should be ok with 3 species.
+Then we calculate lengths with `calc_age2length`.
+This will probably not work for the full run with all species outputs 5x per year, but should be ok with 3 species.
+Still very long runtime!
+Saved output.
 
@@
 
-Now we can test the full `sample_survey_biomass` from nums to biomass index with these three species:
+Now we can test the full `sample_survey_biomass` from nums to biomass index with these three species (required a revision of the function to allow time varying weight at age, and a clarification of units):
 
@@
 
+# weight at age output of calc_age2length is in g
+# output of survey should be in t, sample_survey_biomass expects kg wt@age
+# therefore need to divide by 1000 to get wt@age in kg
+
 wtage <- length_census_ss$muweight %>%
-  select(species, agecl, wtAtAge = atoutput)
+  select(species, agecl, time, wtAtAge = atoutput) %>%
+  mutate(wtAtAge = wtAtAge/1000)
 
@@
 
-Now we plot the results of the numbers-based (census) survey against the same Atlantis output file "[modelscenario]BiomIndx.txt" as above.
+Now we plot the results of the numbers-based (census) survey against the same Atlantis output file "[modelscenario]BiomIndx.txt" as above. We have a match.
+
+To get back true biomass we need the full detailed time varying weight at age output, which is fairly cumbersome.
+The fixed weight at age may be a decent approximation but does result in sigificant misses for some species [as seen here](https://sgaichas.github.io/poseidon-dev/CCAExamples.html).
 
@@
 compareB_N + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free") 
-compareB_N + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free") 
-compareB_N + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free") 
-compareB_N + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+  facet_wrap(~species, scales="free")
 
@@
 
-Start testing fishery outputs next
+Now for the fishery outputs.
+We will need total catch in tons for each species, and catch composition data (age and length).
+First we will test whether the catchbio output in `truth` matches the atlantis output file output[scenario.name]Catch.txt:
+
+```{r testcatchbio}
+
+```
 

### 2019-05-21 Atlantisom2SSworkflowtest.Rmd
start comparison with catch bio

@@
 
+#just try the three species for now
+catchbio_census_ss <- create_fishery_subset(dat = truth$catchbio,
+                                         time = timeall,
+                                         species = spp.name,
+                                         boxes = boxall)
+
+#does it match catch.txt (which should be in tons)?
+atCtxt <- read.table(file.path(d.name, catch.file), header=T)
+
+atCtxttidy <- atCtxt %>%
+  select(Time:XXX) %>%
+  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
+  gather(species, catchbio, -Time) %>%
+  filter(species %in% levels(catchbio_census_ss$species))
+
 ```

### 2019-05-21 Atlantisom2SSworkflowtest.Rmd
add CC climate 1 testing

--- 2019-05-21 Atlantisom2SSworkflowtest.Rmd
@@
 ---
-title: "Testing atlantisom: true biomass output from Atlantis"
+title: "Testing atlantisom: end to end truth for California Current"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-## Simulate a survey part 1: census to compare with truth
+## Test survey functions: census to compare with truth
 
@@
 
+surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+
 ```
+
+### True biomass survey I
 
@@
 
+### True numbers survey
+
 ```{r surveyNbased, eval=FALSE}
@@
 
+### Biological sampling: true numbers at age class
+
 We get true numbers at age at each output step by appling `sample_fish` to the survey_testNall result if the model group has one true age per output age class:
@@
 
+### Biological sampling: setup
+
 The next steps of sampling can get us a proper weight at age.
@@
 
-```{r ss-biolsampling}
+```{r ss-biolsampling, warning=FALSE, message=FALSE}
 
@@
 
+### Biological sampling: true lengths and weight-at-age
+
 Then we calculate lengths with `calc_age2length`.
@@
 ```
+
+### True biomass survey II
 
@@
 
+## Test fishery functions: compare census with truth
+
 Now for the fishery outputs.
@@
 
-```{r testcatchbio}
+```{r testcatchbio1}
+
+#just try the three species for now
+catchbio_census_ss <- create_fishery_subset(dat = truth$catchbio,
+                                         time = timeall,
+                                         species = spp.name,
+                                         boxes = boxall)
+
+catchbio_census_ss_agg <- aggregate(atoutput ~ species + time,
+                                    data=catchbio_census_ss, sum)
+
+# learned the hard way this can be different from ecosystem outputs
+fstepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutfinc
+
+# what does the output look like?
+catchB_outstep <- ggplot() +
+  geom_line(data=catchbio_census_ss_agg, aes(x=time/fstepperyr,y=atoutput, color="catch atlantisom"),
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+catchB_outstep + 
+  facet_wrap(~species, scales="free")
 
@@
 
+It doesnt. A hack gets it closer, suggesting that the problem is a bug in the output for models using codebases prior to late 2015 (CCA is mid-2015).
+
+```{r testcatchbio2}
+
+# should it be aggregated? I just did this above
+catchbio_series <- sample_survey_numbers(catchbio_census_ss, surv_cv)
+
+# if it is the pre-2015 bug mentioned in atlantis wiki, this may get it close
+catchbio_series_adjust <- catchbio_series %>%
+  mutate(atoutput = atoutput/(86400*4))
+
+#does it match catch.txt (which should be in tons)?
+atCtxt <- read.table(file.path(d.name, catch.file), header=T)
+atCtxt <- atCtxt[, -grep("TsAct", colnames(atCtxt))] #hardcoded for catch.txt output
+
+fishedlookup <- funct.groups %>%
+  filter(IsFished > 0)
+
+atCtxttidy <- atCtxt %>%
+  rename_(.dots=with(fishedlookup, setNames(as.list(as.character(Code)), Name))) %>%
+  gather(species, catchbio, -Time) %>%
+  filter(species %in% levels(catchbio_series$species))
+
+compare_catchB <-ggplot() +
+  geom_line(data=catchbio_series_adjust, aes(x=time,y=atoutput, color="catch atlantisom"), 
+            alpha = 10/10) +
+  geom_point(data=atCtxttidy, aes(x=Time/365,y=catchbio, color="txt output catch bio"),
+             alpha = 1/10) + 
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+compare_catchB + 
+  facet_wrap(~species, scales="free")
+
+```
+
+Test this with NOBA model, which may be in the codebase that doesn't have this problem with the CATCH.nc file.
+
 ## References

### 2019-05-21 TestCatchNOBA.Rmd
try fishery testing on NOBA

--- 2019-05-16 TestWeightAtStage.Rmd
@@
 ---
-title: "Testing atlantisom: visualize weight at (st)age output. How shall we interplolate this for true ages?"
+title: "Testing atlantisom: fishery catch data tests"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = TRUE)
 # automatically create a bib database for R packages
@@
 
-Now we delve into the calculation of true age classes and weight-at-age from standard Atlantis output age class, which is generally in 10 stages of maxage/10 true age classes for all but the newest Atlantis models.
-Note that modern Atlantis models can output true age classes in two output files (from the wiki):
-
->ANNAGEBIO.nc: This output provides numbers in each annual age class (so mapped from Atlantis "age class" which can contain multiple years to true annual age classes).
-Set flag_age_output to 1 to get this output.
-Tracers provided are:
->   + Numbers at age per species
->
->ANNAGECATCH.nc: This output provides numbers at annual age class (so mapped from Atlantis "age class" which can contain multiple years to true annual age classes) in the catch and discards (summed over all fleets).
-Set flag_age_output to 1 to get this output.
-Tracers provided are:
->   + Numbers at age per species in the catch
->   + Numbers at age per species in the discards
-
-Future `atlantisom` users should take advantage of this direct output, and some `atlantisom` functions will have to be rewritten to use it.
-
-However, The CCA model does not produce this output, and other legacy models may not, so we still need the `atlantisom` functions `calc_stage2age` and the associated `calc_Z`.
-This function will be tested.
-However, it produces numbers at age but not weight.
-Weight-at-true-age is needed as an input to stock assessment models using empirical weight at age.
-See discussion [here](ttps://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html)
-
-Therefore, our first step is understanding what the 10 class weight at age (output as `mulength` from `calc_age2length`) looks like for a variety of species.
+Fishery sampling functions in atlantisom have not yet been extensively tested.
+There is also a problem with the CATCH.nc output for models in codebases prior to 2015 (inclucing CCA).
+Here we first test with NOBA which should not have this output bug, then we will revisit CCA with a fix to `read_truth` to compensate for the catch output problem.
 
@@
 
+We need to re-`run_truth` for NOBA to include the catchbio component:
+
 ```{r load_Rdata, message=FALSE, warning=FALSE}
 
-if(initCCA) {
-  truth.file <- "outputCCV3run_truth.RData"
-  load(file.path(d.name, truth.file))
-  truth <- result
-}
+# default run_truth setup will save the file, so check for that first
 
-if(initNEUS) {
-  truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
-  load(file.path(d.name, truth.file))
-  truth <- result
-}
-
-if(initNOBA){
-  truth.file <- "outputnordic_runresults_01run_truth.RData" 
-  load(file.path(d.name, truth.file))
-  truth <- result
+if(!file.exists(file.path(d.name, 
+                          paste0("output", scenario.name, "run_truth.RData")))){
+  #Store all loaded results into an R object
+  truth <- run_truth(scenario = scenario.name,
+                     dir = d.name,
+                     file_fgs = functional.groups.file,
+                     file_bgm = box.file,
+                     select_groups = funct.group.names,
+                     file_init = initial.conditions.file,
+                     file_biolprm = biol.prm.file,
+                     file_runprm = run.prm.file
+  )
+} else{
+  truth <- get(load(file.path(d.name,
+                              paste0("output", scenario.name, "run_truth.RData"))))
 }
@@
 
-## Simulate a survey part 5: weight at st(age) census
+## Testing fisheries functions
 
-Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html).
-
-We will apply examples here to just a few NOBA species to examine within-year variability:
+We will apply examples here to just a few NOBA species to speed things up:
     - Cod "North_atl_cod", likely a test assessment species

### 2019-05-21 config/NOBAConfig.R

@@ -6,9 +6,5 @@ box.file <- "Nordic02.bgm"
 initial.conditions.file <- "nordic_biol_v23.nc"
 run.prm.file <- "nordic_run_v01.xml"
 scenario.name <- "nordic_runresults_01"
-d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
-truth.file <- "outputnordic_runresults_01run_truth.RData" 
-
-timeall <- c(0:560)
-funct.group.indices <- c(1:36)
-atlantis.output <- here("atlantisoutput","NOBACERESGlobalSustainability","outputnordic_runresults_01BiomIndx.txt")
+bioind.file <- "outputnordic_runresults_01BiomIndx.txt"
+catch.file <- "outputnordic_runresults_01Catch.txt"

### 2019-05-21 docs/index.Rmd
add wt age to index
update index

@@
 
-## Weight at age: by output age class, how to get to true age basis?
+### Weight at age: by output age class, how to get to true age basis?
 
@@
 
+Work in progress on interpolating weight at age in InterpolateWeightAtAge.Rmd
+
+Work TBD applying `calc_stage2age` and `calc_Z` to get true ages
+
+### Testing a full workflow through fisheries (sardine, agecl = true age)
+
+<a href="https://sgaichas.github.io/poseidon-dev/Atlantisom2SSworkflowtest.html" target="_blank">Try for true datasets new CC model with climate - May 21, 2019</a>
+
+Fishery catch problem, older codebase has a bug in output. Test with NOBA...
+
 ### More to come later!

### 2019-05-21 TestCatchNOBA.Rmd
more code fishery testing NOBA

@@
 There is also a problem with the CATCH.nc output for models in codebases prior to 2015 (inclucing CCA).
-Here we first test with NOBA which should not have this output bug, then we will revisit CCA with a fix to `read_truth` to compensate for the catch output problem.
+Here we first test with NOBA which should not have this output bug, then we will revisit CCA with a fix to `read_truth` to compensate for the catch output problem (see [this](https://sgaichas.github.io/poseidon-dev/Atlantisom2SSworkflowtest.html)).
 
@@
 
-# should return a perfectly scaled survey 
-effic1 <- data.frame(species=funct.group.names,
-                     efficiency=rep(1.0,length(funct.group.names)))
-
-# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
-selex1 <- data.frame(species=rep(funct.group.names, each=10),
-                     agecl=rep(c(1:10),length(funct.group.names)),
-                     selex=rep(1.0,length(funct.group.names)*10))
-
-# should return all model areas
-boxpars <- load_box(d.name, box.file)
-boxall <- c(0:(boxpars$nbox - 1))
-
-# generalized timesteps all models
-runpar <- load_runprm(d.name, run.prm.file)
-noutsteps <- runpar$tstop/runpar$outputstep
-stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
-
-timeall <- c(0:noutsteps)
-
-# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
-# for ecosystem indicator work test all species, e.g.
-survspp <- funct.group.names
-
-# for length and age groups lets just do fish and sharks
-# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
-if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)
-
-survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
-                           funct.groups$GroupType %in% c("FISH", "SHARK")]
+source(here("config/census_spec.R"))
 
 ```
+
+# Test fishery functions: compare census with truth
+
+Now for the fishery outputs.
+We will need total catch in tons for each species, and catch composition data (age and length).
+First we will test whether the catchbio output in `truth` matches the atlantis output file output[scenario.name]Catch.txt:
+
+```{r testcatchbio1}
+
+#just try the three species for now, hardcoded for NOBA
+
+spp.name <- funct.group.names[funct.group.names %in% c("North_atl_cod",
+                                                       "Norwegian_ssh",
+                                                       "Green_halibut")]
+
+catchbio_census_ss <- create_fishery_subset(dat = truth$catchbio,
+                                         time = timeall,
+                                         species = spp.name,
+                                         boxes = boxall)
+
+catchbio_census_ss_agg <- aggregate(atoutput ~ species + time,
+                                    data=catchbio_census_ss, sum)
+
+# what does the output look like?
+catchB_outstep <- ggplot() +
+  geom_line(data=catchbio_census_ss_agg, aes(x=time/fstepperyr,y=atoutput, color="catch atlantisom"),
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+catchB_outstep + 
+  facet_wrap(~species, scales="free")
+
+```
+
+```{r testcatchbio2}
+
+# should it be aggregated? I just did this above
+catchbio_series <- sample_survey_numbers(catchbio_census_ss, surv_cv)
+
+# if it is the pre-2015 bug mentioned in atlantis wiki, this may get it close
+# should not need for NOBA
+#catchbio_series_adjust <- catchbio_series %>%
+#  mutate(atoutput = atoutput/(86400*4))
+
+#does it match catch.txt (which should be in tons)?
+atCtxt <- read.table(file.path(d.name, catch.file), header=T)
+atCtxt <- atCtxt[, -grep("TsAct", colnames(atCtxt))] #hardcoded for catch.txt output
+
+fishedlookup <- funct.groups %>%
+  filter(IsFished > 0)
+
+atCtxttidy <- atCtxt %>%
+  rename_(.dots=with(fishedlookup, setNames(as.list(as.character(Code)), Name))) %>%
+  gather(species, catchbio, -Time) %>%
+  filter(species %in% levels(catchbio_series$species))
+
+compare_catchB <-ggplot() +
+  geom_line(data=catchbio_series, aes(x=time,y=atoutput, color="catch atlantisom"), 
+            alpha = 10/10) +
+  geom_point(data=atCtxttidy, aes(x=Time/365,y=catchbio, color="txt output catch bio"),
+             alpha = 1/10) + 
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+compare_catchB + 
+  facet_wrap(~species, scales="free")
+
+```
+
+Did it work for NOBA?
+
+Now test catch at age in numbers from truth$catch
+
+## Modify below to test biological sampling from fisheries
 
@@
 
-```{r stdsurveyNbased-3spp, echo=TRUE}
-
-spp.name <- funct.group.names[funct.group.names %in% c("North_atl_cod",
-                                                       "Norwegian_ssh",
-                                                       "Green_halibut")]
+```{r catchage-len-wt-3spp, echo=TRUE}
 
 # get survey nums with full (no) selectivity
-survey_testNall <- create_survey(dat = truth$nums,
-                                 time = c(0:111),
-                                 species = spp.name,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-# this one is high but not equal to total for numerous groups
-effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
-
-# apply default sample fish as before to get numbers
-numsallhigh <- sample_fish(survey_testNall, effNhigh)
-
-# aggregate true resn per survey design
-aggresnall <- aggregateDensityData(dat = truth$resn,
+catch_testNss <- create_fishery_subset(dat = truth$catch,
                                  time = c(0:111),
@@
 
-# aggregate true structn per survey design
-aggstructnall <- aggregateDensityData(dat = truth$structn,
+# apply default sample fish as before to get numbers
+catch_numssshigh <- sample_fish(catch_testNss, effNhigh)
+
+# aggregate true resn per survey or fishery subset design
+catch_aggresnss <- aggregateDensityData(dat = truth$resn,
+                                 time = c(0:111),
+                                 species = spp.name,
+                                 boxes = boxall)
+
+# aggregate true structn per survey or fishery subsetdesign
+catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
                                  time = c(0:111),
@@
 #dont sample these, just aggregate them using median
-structnall <- sample_fish(aggstructnall, effNhigh, sample = FALSE)
+catch_structnss <- sample_fish(catch_aggstructnss, effNhigh, sample = FALSE)
 
-resnall <-  sample_fish(aggresnall, effNhigh, sample = FALSE)
+catch_resnss <-  sample_fish(catch_aggresnss, effNhigh, sample = FALSE)
 
@@
 
-length_census <- calc_age2length(structn = structnall,
-                                 resn = resnall,
-                                 nums = numsallhigh,
+catch_length_census_ss <- calc_age2length(structn = catch_structnss,
+                                 resn = catch_resnss,
+                                 nums = catch_numssshigh,
                                  biolprm = truth$biolprm, fgs = truth$fgs,

### 2019-05-21 config/census_spec.R
try config file for census sampling

@@ -1,0 +1,46 @@
+# should return a perfectly efficient survey 
+effic1 <- data.frame(species=funct.group.names,
+                     efficiency=rep(1.0,length(funct.group.names)))
+
+# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
+selex1 <- data.frame(species=rep(funct.group.names, each=10),
+                     agecl=rep(c(1:10),length(funct.group.names)),
+                     selex=rep(1.0,length(funct.group.names)*10))
+
+# should return all model areas
+boxpars <- load_box(d.name, box.file)
+boxall <- c(0:(boxpars$nbox - 1))
+
+# generalized timesteps all models
+runpar <- load_runprm(d.name, run.prm.file)
+noutsteps <- runpar$tstop/runpar$outputstep
+stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
+midptyr <- round(median(seq(1,stepperyr)))
+
+# a survey that takes place once per year mid year
+annualmidyear <- seq(midptyr, noutsteps, stepperyr)
+
+timeall <- c(0:noutsteps)
+
+# learned the hard way this can be different from ecosystem outputs
+fstepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutfinc
+
+# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
+# for ecosystem indicator work test all species, e.g.
+survspp <- funct.group.names 
+
+# for length and age groups lets just do fish and sharks
+# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
+if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)
+
+survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
+                               funct.groups$GroupType %in% c("FISH", "SHARK")]
+
+# needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
+
+# needed for sample_survey_xxx
+# perfect observation
+surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+

### 2019-05-22 CreateStockSynthesis.Rmd
Update CreateStockSynthesis file to include age comp writing

@@
 require(here)
+require(reshape2)
 knitr::opts_chunk$set(echo = TRUE)
@@
 ```
+
+## Writing age composition data
+
+```{r}
+#Set effective sample size for age compositions
+effN <- 1000
+highEffN <- data.frame(species=species, effN=rep(effN, length(species)))
+
+#Sample fish for age composition
+age_comp_data <- sample_fish(survey_out, highEffN)
+names(age_comp_data) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")
+
+age_comp_data
+
+#Group age composition data by time
+require(dplyr)
+age_comp_proportion <- age_comp_data %>%
+  group_by(time) %>%
+  mutate(comp = numAtAgesamp/sum(numAtAgesamp))
+
+#Get the age bins
+age_bin_names <- names(sardine.data$agecomp)[10:length(names(sardine.data$agecomp))]
+age_bins <- sub("a","",age_bin_names)
+
+#Change data format into flat structure
+age_comp_proportion$id <- rep(age_bins,
+                              each = nrow(age_comp_proportion)/length(age_bins))
+
+unmelt <- dcast(data = age_comp_proportion,
+                formula = time ~agecl,
+                value.var = "comp")
+age_comp_flat<- cbind(unmelt[order(unmelt$time),],
+                      rep(effN,nrow(unmelt)))
+names(age_comp_flat)[ncol(age_comp_flat)] <- "nsamp"
+
+## Write age composition data
+sardine.data <- SS_write_comps(ss_data_list = sardine.data,
+               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
+               data_rows = list(sardine.data$styr:(sardine.data$styr+nyears-1)),
+               sampling_month = list(rep(7,nyears)),
+               data_type = c("agecomp"),
+               fleet_number = c(1),
+               bins = list(age_bins),
+               caal_bool = c(FALSE))
+sardine.data$agecomp
+
+```

### 2019-05-22 TestCatchNOBA.Rmd

@@
 ---
-title: "Testing atlantisom: fishery catch data tests"
+title: "Testing atlantisom: fishery catch data tests, NOBA"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 fishedlookup <- funct.groups %>%
-  filter(IsFished > 0)
+  filter((isFished) > 0)  # WARNING NOBA has isFished while CCA has IsFished
 
@@
 compare_catchB <-ggplot() +
-  geom_line(data=catchbio_series, aes(x=time,y=atoutput, color="catch atlantisom"), 
+  geom_line(data=catchbio_series, aes(x=time/fstepperyr,y=atoutput, color="catch1 atlantisom"), 
             alpha = 10/10) +
@@
 
-Did it work for NOBA?
+No, it does not match.
+It seems the problem is in calculating catch biomass at age; the CATCH.nc output file does not include a layer, so the aggregation of structn and resn over layers (using medians) in lines 58-67 in `calc_biomass_age` may be resulting in this mismatch.
+If catch happened in a particular layer, it is aggregated in the output CATCH.nc file, so we don't know which layer has the appropriate weight info.
+The `calc_biomass_age` function works on the truth$nums output because no aggregation of structn and resn is necessary, all have a layer field.
 
-Now test catch at age in numbers from truth$catch
+This does make me worry about the aggregation we apply to get weight at age for sampled fish.
+However, it appeared to recover the total biomass time series, so maybe it is ok.
+There may be something else wrong with the portion of the calc_biomass_age function that only applies to catch.
 
-## Modify below to test biological sampling from fisheries
+One workaround is to assume the total catch in the catch.txt output file is correct and use that directly as catch biomass for atlantisom.
+We will have only annual total catch and no way to break it into seasons, but maybe that is ok.
 
-Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just one species in this case).
+Then we would take catch at age in numbers from truth$catch (directly from CATCH.nc file) as below.
+California current will still need a correction to this due to the fishery output writing bug in the older atlantis codebase.
+
+## Test biological sampling from fisheries
+
+Here we use `create_fishery_subset` on the numbers output of `run_truth` to create the survey census of age composition (for just our three species in this case).
 The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.
@@
 
-```{r catchage-len-wt-3spp, echo=TRUE}
+```{r catchage-len-wt-3spp, echo=TRUE, warning=FALSE, message=FALSE}
 
@@
 
+This is true catch at age(class) for cod:
+
+```{r catage1}
+
+catchage <- catch_numssshigh %>%
+  filter(species == "North_atl_cod")
+
+catageplot <- ggplot(catchage, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchage$species))
+
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
+This is true catch at age(class) for herring:
+
+```{r catage2}
+
+catchage <- catch_numssshigh %>%
+  filter(species == "Norwegian_ssh")
+
+catageplot <- ggplot(catchage, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchage$species))
+
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
+Both of these species will need `calc_stage2age` applied because both have 2 true ages per age class.
+
 Length sample with user specified max length bin (200 cm):
@@
 
-We should get the upper end of Greenland halibut with a 200cm max length bin. That shouldnt matter for weight at age. This is all timesteps for each species by agecl.
+We should get the upper end of Greenland halibut with a 200cm max length bin.
 
-```{r vis-wtageclass-test}
+Cod catch lengths:
 
-wageplot <- ggplot(length_census$muweight, aes(agecl, atoutput)) +
+```{r vis-fishery-lf-test-1}
+
+catchlen <- catch_length_census_ss$natlength %>%
+  filter(species == "North_atl_cod")
+
+lfplot <- ggplot(catchlen, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchlen$species))
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+
+```
+
+Herring catch lengths:
+
+```{r vis-fishery-lf-test-2}
+
+catchlen <- catch_length_census_ss$natlength %>%
+  filter(species == "Norwegian_ssh")
+
+lfplot <- ggplot(catchlen, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchlen$species))
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+
+```
+
+Fishery weight at (st)age:
+
+```{r vis-fishery-wtageclass-test}
+
+wageplot <- ggplot(catch_length_census_ss$muweight, aes(agecl, atoutput)) +
   geom_point(aes(colour = time)) +
@@
 
-How much does weight at age class vary over time? Here are a couple of annual cycles:
-
-```{r vis-wtageclass-test2}
-
-wtageyr1 <- length_census$muweight %>%
-  filter(time<5)
-
-wageplot <- ggplot(wtageyr1, aes(agecl, atoutput)) +
-  geom_line(aes(colour = factor(time))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  scale_x_discrete(limits=c(1:10)) +
-  xlab("age class") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " initial year"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-
-```
-
-```{r vis-wtageclass-test3}
-
-wtageyrmid <- length_census$muweight %>%
-  filter(between(time, 50,54))
-
-wageplot <- ggplot(wtageyrmid, aes(agecl, atoutput)) +
-  geom_line(aes(colour = factor(time))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  scale_x_discrete(limits=c(1:10)) +
-  xlab("age class") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " year 10"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-
-```
-
-```{r vis-wtageclass-test4}
-
-wtageyr20 <- length_census$muweight %>%
-  filter(between(time, 100,104))
-
-wageplot <- ggplot(wtageyr20, aes(agecl, atoutput)) +
-  geom_line(aes(colour = factor(time))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  scale_x_discrete(limits=c(1:10)) +
-  xlab("age class") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " year 20"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-
-```
-
-There are some interesting jumps between age classes. Most pronounced for herring. All of these species would be split into two true ages per age class. Interpolation could be interesting.
-
-Change in wt at age over time for age classes using an annual mid-year snapshot (survey) (first 22+ years of NOBA model run):
+Change in wt at (st)age in the fishery over time for age classes using an annual mid-year snapshot (first 22+ years of NOBA model run):
 
@@
 
-# from std survey code and model step info read in above
-midptyr <- round(median(seq(1,stepperyr)))
-annualmidyear <- seq(midptyr, noutsteps, stepperyr)
-
-wtage_annsurv <- length_census$muweight %>%
+wtage_annsurv <- catch_length_census_ss$muweight %>%
   filter(time %in% annualmidyear)
@@
   ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
+  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
 
@@
 
-## Quick test with saved CCA census output
+We will need to interpolate fishery weight at true age using the in-progress function for these species.
 
-Here we check the CCA model (Atlantis Summit Common Scenario 1) saved length and weight census output for small species at Isaac's suggestion (all three of the following are true age classes, NumAgeClassSize=1 per Atlantis agecl bin:
-
-```{r loadCCAlengthcomp, echo=TRUE}
-
-source(here("config/CCConfig.R"))
-
-length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))
-
-```
-
-Sardine:  
-```{r plotCCAsardine}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Pacific_sardine")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-Anchovy:  
-```{r plotCCAanchovy}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Anchovy")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-Herring:  
-```{r plotCCAherring}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Herring")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-How do these three look? Herring had a rough period from model years 15-30 or so.
-
-Now for some CCA groups with multiple true ages per stage.
-
-Pacific hake, 2 true ages per class:  
-```{r plotCCAhake}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Mesopel_M_Fish")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-Bocaccio rockfish, 5 true ages per age class:
-```{r plotCCAbocaccio}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Bocaccio_rockfish")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-Yelloweye rockfish, most extreme at 12 true ages per age class:
-```{r plotCCAYelloweye}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Yelloweye_rockfish")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-I'm not sure any of this helps us decide how to interpolate, but there is clearly contrast in weight at age (class) in both models and example runs.
+Any further info to get?

### 2019-05-22 Atlantisom2SSworkflowtest.Rmd
working on catch correction

@@
 # make a function for this
+source(here("config/census_spec.R"))
 
@@
 
-It doesnt. A hack gets it closer, suggesting that the problem is a bug in the output for models using codebases prior to late 2015 (CCA is mid-2015).
+It doesnt. A hack gets it closer, suggesting that the problem is a bug in the output for models using codebases prior to late 2015 (CCA is mid-2015). From Beth Fulton's note in the wiki:
+
+>Note that in CATCH.NC the numbers at age had issues in old versions.
+For versions pre-dating December 2015 you need to divide the numbers at age by (86400 * number_water_column_layers_in_the_box).
+New versions post December 2015 are now in numbers without needing further adjustment.
 
@@
 
-Test this with NOBA model, which may be in the codebase that doesn't have this problem with the CATCH.nc file.
+Test this with NOBA model, which is in a newer codebase that doesn't have this problem with the CATCH.nc file.
+
+Update: NOBA has correct catch at age but the derived total catch biomass [still doesnt match catch.txt](https://sgaichas.github.io/poseidon-dev/TestCatchNOBA.html).
+I believe this is because we cannot multiply the correct weight at the layer the fishery catch is taken, which clearly does not match the median structn and resn across all model layers (as applied in `calc_biomass-age`.
+This may be buried deeply in the atlantis code, initialization, and prm files, but for now we will simply use the total catch weight out of atlantis catch.txt output.
+
+Moving on to correcting catch at age for CCA:
+
+```{r correctcatage}
+
+# new atlantisom function to apply catage correction, or put inside run_truth
+
+# first check that codebase is prior to corrected--to implement in atlantisom
+#   read code version or date from log file
+#   check against update date (Dec 12, 2015)
+
+# read catch in nums output from truth$catch, or
+# read in CATCH.nc (this from run_truth, modified with config files)
+
+nc_catch <- paste0("output", scenario.name, 'CATCH.nc')
+bboxes <- get_boundary(boxpars)
+
+catch <- load_nc(dir = d.name,
+                 file_nc = nc_catch,
+                 bps = boxpars, #defined in census_spec.R
+                 fgs = funct.groups,
+                 select_groups = survspp,
+                 select_variable = "Catch",
+                 check_acronyms = TRUE,
+                 bboxes = bboxes)
+
+# find number of layers in each box
+# The number of water column layers per box is in the initial conditions NC file:
+# DIVinit_CalCurrentV3_Biol.nc : numlayers =
+
+# read in initial conditions NC file
+# this is already done for the other variables, modify from existing
+# line 152 of load_nc
+# num_layers <- RNetCDF::var.get.nc(ncfile = at_out, variable = "numlayers")[, 1]
+
+# initial conditions file has different structure, so drop [,1]
+at_init <- RNetCDF::open.nc(con = file.path(d.name, initial.conditions.file))
+on.exit(RNetCDF::close.nc(at_init))
+
+# Get info from netcdf file! (Filestructure and all variable names)
+  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_init)$nvars - 1),
+    function(x) RNetCDF::var.inq.nc(at_init, x)$name)
+  numlayers <- RNetCDF::var.get.nc(ncfile = at_init, variable = "numlayers")
+
+  # are these in box order??? if so make a box-numlayer lookup
+  layerbox.lookup <- data.frame(polygon=boxall, numlayers)
+
+  catch.tmp <- merge(catch, layerbox.lookup)
+
+# divide the numbers at age by (86400 * number_water_column_layers_in_the_box)
+
+  catch <- catch.tmp %>%
+    mutate(atoutput = atoutput * 86400 * numlayers) %>%
+
+# replace truth$catch atoutput with correction
+
+```
 

### 2019-05-22 Atlantisom2SSworkflowtest.Rmd
update CCA workflow with correction, biol sampling

@@
 source(here("config/census_spec.R"))
-
-# should return a perfectly scaled survey 
-effic1 <- data.frame(species=funct.group.names,
-                     efficiency=rep(1.0,length(funct.group.names)))
-
-# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
-# BUT CHECK if newer Atlantis models can do age-specific outputs
-selex1 <- data.frame(species=rep(funct.group.names, each=10),
-                     agecl=rep(c(1:10),length(funct.group.names)),
-                     selex=rep(1.0,length(funct.group.names)*10))
-
-# should return all model areas
-boxpars <- load_box(d.name, box.file)
-boxall <- c(0:(boxpars$nbox - 1))
-
-# generalized timesteps all models
-runpar <- load_runprm(d.name, run.prm.file)
-noutsteps <- runpar$tstop/runpar$outputstep
-stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
-
-timeall <- c(0:noutsteps)
-
-# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
-# for ecosystem indicator work test all species, e.g.
-survspp <- funct.group.names
-
-# for length and age groups lets just do fish and sharks
-# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
-if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)
-
-survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
-                           funct.groups$GroupType %in% c("FISH", "SHARK")]
-
-surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
 
@@
 
+## Catch output correction--apply to all species
+
 Moving on to correcting catch at age for CCA:
@@
 #   check against update date (Dec 12, 2015)
+# grep("Atlantis SVN Last Change Date", readLines(file.path(d.name, "log.txt")), value = TRUE)
+
+# alternative: less overhead if the system does it? log file is huge
+# logfile <- paste0(file.path(d.name, "log.txt"))
+# codedate <- system(paste0("grep 'Atlantis SVN' ", logfile))
+# compare date in codedate in an if statement--fix in function
 
@@
 # divide the numbers at age by (86400 * number_water_column_layers_in_the_box)
+# replace truth$catch atoutput with correction 
+  catch <- catch.tmp %>%
+    mutate(atoutput = atoutput / 86400 * numlayers) %>%
+    select(species, agecl, polygon, time, atoutput)
 
-  catch <- catch.tmp %>%
-    mutate(atoutput = atoutput * 86400 * numlayers) %>%
+  truth$catch <- catch
 
-# replace truth$catch atoutput with correction
+#save for later use, ar (better) resave truth file
+#saveRDS(catch, file.path(d.name, paste0(scenario.name, "catchnums_corrected.rds")))
+  save(truth,
+      file = file.path(dir, paste0("output", scenario, "run_truth.RData")))
+```
+
+Did the correction work? Not sure what to compare this to. First put it back through `calc_biomass_age` even though this won't match either?
+
+```{r correctcatage-tobio, eval=FALSE}
+
+biol <- load_biolprm(dir = d.name, file_biolprm = biol.prm.file)
+
+# call within run_truth
+catchbio <- calc_biomass_age(nums = catch,
+    resn = truth$resn, structn = truth$structn, biolprm = biol)
+
+#save for later use, even if wrong? takes a long time to generate
+saveRDS(catchbio, file.path(d.name, paste0(scenario.name, "catchbio_ccatage.rds")))
+
+```
+
+It still doesn't match, similar to NOBA, but in the right ballpark now so the error is the median structn and resn called in the calc function above.
+This suggests that the catch at age in numbers is now correct, however.
+
+```{r plotcorrected-bio}
+
+catchbio <- readRDS(file.path(d.name, paste0(scenario.name, "catchbio_ccatage.rds")))
+
+catchbio_corrected_agg <- aggregate(atoutput ~ species + time,
+                                    data=catchbio, sum)
+
+compare_catchB <-ggplot() +
+  geom_line(data=catchbio_corrected_agg, aes(x=time,y=atoutput, color="catch atlantisom"), 
+            alpha = 10/10) +
+  geom_point(data=atCtxttidy, aes(x=Time/365,y=catchbio, color="txt output catch bio"),
+             alpha = 1/10) + 
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+compare_catchB + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
+compare_catchB + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
+compare_catchB + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
+compare_catchB + 
+  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
+## Fishery biological sampling
+
+Now we can sample fishery catch for lengths, ages, weights:
+
+Here we use `create_fishery_subset` on the numbers output of `run_truth` to create the survey census of age composition (for just our three species in this case).
+The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.
+
+Because we don't want to wait 24 hours for this, we will look at only our three speces and the first 112 time steps.
+
+```{r catchage-len-wt-3spp, echo=TRUE, warning=FALSE, message=FALSE}
+
+# get survey nums with full (no) selectivity
+catch_testNss <- create_fishery_subset(dat = truth$catch,
+                                 time = c(0:111),
+                                 species = spp.name,
+                                 boxes = boxall)
+
+# apply default sample fish as before to get numbers
+catch_numssshigh <- sample_fish(catch_testNss, effNhigh)
+
+# aggregate true resn per survey or fishery subset design
+catch_aggresnss <- aggregateDensityData(dat = truth$resn,
+                                 time = c(0:111),
+                                 species = spp.name,
+                                 boxes = boxall)
+
+# aggregate true structn per survey or fishery subsetdesign
+catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
+                                 time = c(0:111),
+                                 species = spp.name,
+                                 boxes = boxall)
+
+#dont sample these, just aggregate them using median
+catch_structnss <- sample_fish(catch_aggstructnss, effNhigh, sample = FALSE)
+
+catch_resnss <-  sample_fish(catch_aggresnss, effNhigh, sample = FALSE)
+
+```
+
+This is true catch at age for sardine:
+
+```{r catage1}
+
+catchage <- catch_numssshigh %>%
+  filter(species == "Pacific_sardine")
+
+catageplot <- ggplot(catchage, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchage$species))
+
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
+This is true catch at age(class) for hake:
+
+```{r catage2}
+
+catchage <- catch_numssshigh %>%
+  filter(species == "Mesopel_M_Fish")
+
+catageplot <- ggplot(catchage, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchage$species))
+
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
+Hake will need `calc_stage2age` applied because it has 2 true ages per age class.
+
+Length sample with user specified max length bin (200 cm):
+
+```{r userset-maxlen, echo=TRUE}
+
+catch_length_census_ss <- calc_age2length(structn = catch_structnss,
+                                 resn = catch_resnss,
+                                 nums = catch_numssshigh,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 maxbin = 200,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+```
+
+We should get the upper end of anything with a 200cm max length bin.
+
+Sardine catch lengths:
+
+```{r vis-fishery-lf-test-1}
+
+catchlen <- catch_length_census_ss$natlength %>%
+  filter(species == "Pacific_sardine")
+
+lfplot <- ggplot(catchlen, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchlen$species))
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+
+```
+
+Hake catch lengths:
+
+```{r vis-fishery-lf-test-2}
+
+catchlen <- catch_length_census_ss$natlength %>%
+  filter(species == "Mesopel_M_Fish")
+
+lfplot <- ggplot(catchlen, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchlen$species))
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+
+```
+
+Fishery weight at (st)age:
+
+```{r vis-fishery-wtageclass-test}
+
+wageplot <- ggplot(catch_length_census_ss$muweight, aes(agecl, atoutput)) +
+  geom_point(aes(colour = time)) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  scale_x_discrete(limits=c(1:10)) +
+  xlab("age class") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste(scenario.name))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
+
+```
+
+Change in wt at (st)age in the fishery over time for age classes using an annual mid-year snapshot (first 22+ years of CCA model run):
+
+```{r aggwtcomp}
+
+wtage_annsurv <- catch_length_census_ss$muweight %>%
+  filter(time %in% annualmidyear)
+
+# reverse to show agecl time series of wt
+wageplot <- ggplot(wtage_annsurv, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  xlab("model timestep (5 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
 

### 2019-05-22 Atlantisom2SSworkflowtest.Rmd

@@
 at_init <- RNetCDF::open.nc(con = file.path(d.name, initial.conditions.file))
-on.exit(RNetCDF::close.nc(at_init))
 
@@
   numlayers <- RNetCDF::var.get.nc(ncfile = at_init, variable = "numlayers")
+
+  RNetCDF::close.nc(at_init)
 
@@
   save(truth,
-      file = file.path(dir, paste0("output", scenario, "run_truth.RData")))
+      file = file.path(d.name, paste0("output", scenario.name, "run_truth.RData")))
 ```

### 2019-05-22 docs/index.Rmd

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/Atlantisom2SSworkflowtest.html" target="_blank">Try for true datasets new CC model with climate - May 21, 2019</a>
+<a href="https://sgaichas.github.io/poseidon-dev/Atlantisom2SSworkflowtest.html" target="_blank">Try for true datasets new CC model with climate - May 21, 2019, updated May 22, 2019 with catch correction</a>
 
@@
 
+<a href="https://sgaichas.github.io/poseidon-dev/TestCatchNOBA.html" target="_blank">Suggests a workaround needed for catch in weight - May 21, 2019</a>
+
 ### More to come later!

### 2019-05-22 FullSardineTruthEx.Rmd
add full sardine example

--- 2019-05-22 Atlantisom2SSworkflowtest.Rmd
@@
 ---
-title: "Testing atlantisom: end to end truth for California Current"
+title: "Sardine example: end to end truth for California Current"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
+Catch in numbers has been corrected in the saved and reloaded truth file here. Ignore the catchbio component of this, it is not correct.
+
 ```{r get_truth, message=FALSE, warning=FALSE}
@@
 }
+```
+
+Import the true catch biomass from a separate file. True total catch weight is currently available only on an annual and full-model (all polygons) basis:
+
+```{r loadcatchb}
+
+truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)
+
 ```
@@
 
+surveyB_frombio_sardine <- surveyB_frombio[surveyB_frombio$species == "Pacific_sardine",]
+
 # lookup the matching names, put in time, species, biomass column format
@@
   gather(species, biomass, -Time) %>%
-  filter(species %in% levels(surveyB_frombio$species))
+  filter(species %in% levels(surveyB_frombio_sardine$species))
 
@@
 compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free") 
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free") 
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free") 
-compareB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
+  facet_wrap(~species, scales="free")
 
@@
 
+```{r survNtrue-plot}
+
+surveyN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))
+
+surveyN_sardine <- surveyN[surveyN$species == "Pacific_sardine",]
+
+plotN <-ggplot() +
+  geom_line(data=surveyN_sardine, aes(x=time/5,y=atoutput, color="survey census N"), 
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotN +
+  facet_wrap(~species, scales="free")
+
+```
+
 ### Biological sampling: true numbers at age class
@@
 saveRDS(numsallhigh, file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
+
+```
+
+```{r Natage-plot}
+
+Natage <- readRDS(file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
+
+Natage_sardine <- Natage[Natage$species == "Pacific_sardine",]
+
+Natageplot <- ggplot(Natage_sardine, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        Natage_sardine$species))
+
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
 
@@
 
+```{r lenatage-plot}
+
+length_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "length_census_sardhakeboca.rds")))
+
+len <- length_census_ss$natlength %>%
+  filter(species == "Pacific_sardine")
+
+lfplot <- ggplot(len, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        len$species))
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+
+```
+
 ### True biomass survey II
@@
 
-length_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "length_census_sardhakeboca.rds")))
+#length_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "length_census_sardhakeboca.rds")))
 
@@
 
-## Test fishery functions: compare census with truth
+## Fishery catch
 
-Now for the fishery outputs.
-We will need total catch in tons for each species, and catch composition data (age and length).
-First we will test whether the catchbio output in `truth` matches the atlantis output file output[scenario.name]Catch.txt:
+We found [after much testing](https://sgaichas.github.io/poseidon-dev/Atlantisom2SSworkflowtest.html) that the most reliable source for catch in biomass at present is the Atlantis output catch.txt file, loaded above.
+Here is sardine annual total catch:
 
-```{r testcatchbio1}
+```{r plotcatch}
 
-#just try the three species for now
-catchbio_census_ss <- create_fishery_subset(dat = truth$catchbio,
-                                         time = timeall,
-                                         species = spp.name,
-                                         boxes = boxall)
+truecatchbio_sardine <- truecatchbio[truecatchbio$species == "Pacific_sardine",]
 
-catchbio_census_ss_agg <- aggregate(atoutput ~ species + time,
-                                    data=catchbio_census_ss, sum)
-
-# learned the hard way this can be different from ecosystem outputs
-fstepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutfinc
-
-# what does the output look like?
-catchB_outstep <- ggplot() +
-  geom_line(data=catchbio_census_ss_agg, aes(x=time/fstepperyr,y=atoutput, color="catch atlantisom"),
+plotcatch <- ggplot() +
+  geom_line(data=truecatchbio_sardine, aes(x=time,y=atoutput, color="true catch bio"),
             alpha = 10/10) +
@@
 
-catchB_outstep + 
+plotcatch +
   facet_wrap(~species, scales="free")
-
-```
-
-It doesnt. A hack gets it closer, suggesting that the problem is a bug in the output for models using codebases prior to late 2015 (CCA is mid-2015). From Beth Fulton's note in the wiki:
-
->Note that in CATCH.NC the numbers at age had issues in old versions.
-For versions pre-dating December 2015 you need to divide the numbers at age by (86400 * number_water_column_layers_in_the_box).
-New versions post December 2015 are now in numbers without needing further adjustment.
-
-```{r testcatchbio2}
-
-# should it be aggregated? I just did this above
-catchbio_series <- sample_survey_numbers(catchbio_census_ss, surv_cv)
-
-# if it is the pre-2015 bug mentioned in atlantis wiki, this may get it close
-catchbio_series_adjust <- catchbio_series %>%
-  mutate(atoutput = atoutput/(86400*4))
-
-#does it match catch.txt (which should be in tons)?
-atCtxt <- read.table(file.path(d.name, catch.file), header=T)
-atCtxt <- atCtxt[, -grep("TsAct", colnames(atCtxt))] #hardcoded for catch.txt output
-
-fishedlookup <- funct.groups %>%
-  filter(IsFished > 0)
-
-atCtxttidy <- atCtxt %>%
-  rename_(.dots=with(fishedlookup, setNames(as.list(as.character(Code)), Name))) %>%
-  gather(species, catchbio, -Time) %>%
-  filter(species %in% levels(catchbio_series$species))
-
-compare_catchB <-ggplot() +
-  geom_line(data=catchbio_series_adjust, aes(x=time,y=atoutput, color="catch atlantisom"), 
-            alpha = 10/10) +
-  geom_point(data=atCtxttidy, aes(x=Time/365,y=catchbio, color="txt output catch bio"),
-             alpha = 1/10) + 
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-compare_catchB + 
-  facet_wrap(~species, scales="free")
-
-```
-
-Test this with NOBA model, which is in a newer codebase that doesn't have this problem with the CATCH.nc file.
-
-Update: NOBA has correct catch at age but the derived total catch biomass [still doesnt match catch.txt](https://sgaichas.github.io/poseidon-dev/TestCatchNOBA.html).
-I believe this is because we cannot multiply the correct weight at the layer the fishery catch is taken, which clearly does not match the median structn and resn across all model layers (as applied in `calc_biomass-age`.
-This may be buried deeply in the atlantis code, initialization, and prm files, but for now we will simply use the total catch weight out of atlantis catch.txt output.
-
-## Catch output correction--apply to all species
-
-Moving on to correcting catch at age for CCA:
-
-```{r correctcatage}
-
-# new atlantisom function to apply catage correction, or put inside run_truth
-
-# first check that codebase is prior to corrected--to implement in atlantisom
-#   read code version or date from log file
-#   check against update date (Dec 12, 2015)
-# grep("Atlantis SVN Last Change Date", readLines(file.path(d.name, "log.txt")), value = TRUE)
-
-# alternative: less overhead if the system does it? log file is huge
-# logfile <- paste0(file.path(d.name, "log.txt"))
-# codedate <- system(paste0("grep 'Atlantis SVN' ", logfile))
-# compare date in codedate in an if statement--fix in function
-
-# read catch in nums output from truth$catch, or
-# read in CATCH.nc (this from run_truth, modified with config files)
-
-nc_catch <- paste0("output", scenario.name, 'CATCH.nc')
-bboxes <- get_boundary(boxpars)
-
-catch <- load_nc(dir = d.name,
-                 file_nc = nc_catch,
-                 bps = boxpars, #defined in census_spec.R
-                 fgs = funct.groups,
-                 select_groups = survspp,
-                 select_variable = "Catch",
-                 check_acronyms = TRUE,
-                 bboxes = bboxes)
-
-# find number of layers in each box
-# The number of water column layers per box is in the initial conditions NC file:
-# DIVinit_CalCurrentV3_Biol.nc : numlayers =
-
-# read in initial conditions NC file
-# this is already done for the other variables, modify from existing
-# line 152 of load_nc
-# num_layers <- RNetCDF::var.get.nc(ncfile = at_out, variable = "numlayers")[, 1]
-
-# initial conditions file has different structure, so drop [,1]
-at_init <- RNetCDF::open.nc(con = file.path(d.name, initial.conditions.file))
-
-# Get info from netcdf file! (Filestructure and all variable names)
-  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_init)$nvars - 1),
-    function(x) RNetCDF::var.inq.nc(at_init, x)$name)
-  numlayers <- RNetCDF::var.get.nc(ncfile = at_init, variable = "numlayers")
-
-  RNetCDF::close.nc(at_init)
-
-  # are these in box order??? if so make a box-numlayer lookup
-  layerbox.lookup <- data.frame(polygon=boxall, numlayers)
-
-  catch.tmp <- merge(catch, layerbox.lookup)
-
-# divide the numbers at age by (86400 * number_water_column_layers_in_the_box)
-# replace truth$catch atoutput with correction 
-  catch <- catch.tmp %>%
-    mutate(atoutput = atoutput / 86400 * numlayers) %>%
-    select(species, agecl, polygon, time, atoutput)
-
-  truth$catch <- catch
-
-#save for later use, ar (better) resave truth file
-#saveRDS(catch, file.path(d.name, paste0(scenario.name, "catchnums_corrected.rds")))
-  save(truth,
-      file = file.path(d.name, paste0("output", scenario.name, "run_truth.RData")))
-```
-
-Did the correction work? Not sure what to compare this to. First put it back through `calc_biomass_age` even though this won't match either?
-
-```{r correctcatage-tobio, eval=FALSE}
-
-biol <- load_biolprm(dir = d.name, file_biolprm = biol.prm.file)
-
-# call within run_truth
-catchbio <- calc_biomass_age(nums = catch,
-    resn = truth$resn, structn = truth$structn, biolprm = biol)
-
-#save for later use, even if wrong? takes a long time to generate
-saveRDS(catchbio, file.path(d.name, paste0(scenario.name, "catchbio_ccatage.rds")))
-
-```
-
-It still doesn't match, similar to NOBA, but in the right ballpark now so the error is the median structn and resn called in the calc function above.
-This suggests that the catch at age in numbers is now correct, however.
-
-```{r plotcorrected-bio}
-
-catchbio <- readRDS(file.path(d.name, paste0(scenario.name, "catchbio_ccatage.rds")))
-
-catchbio_corrected_agg <- aggregate(atoutput ~ species + time,
-                                    data=catchbio, sum)
-
-compare_catchB <-ggplot() +
-  geom_line(data=catchbio_corrected_agg, aes(x=time,y=atoutput, color="catch atlantisom"), 
-            alpha = 10/10) +
-  geom_point(data=atCtxttidy, aes(x=Time/365,y=catchbio, color="txt output catch bio"),
-             alpha = 1/10) + 
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-compare_catchB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
-compare_catchB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
-compare_catchB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
-compare_catchB + 
-  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
 
@@
 
-Now we can sample fishery catch for lengths, ages, weights:
+Now (that catch nums are corrected!) we can sample fishery catch for lengths, ages, weights:
 
@@
 
-This is true catch at age(class) for hake:
-
-```{r catage2}
-
-catchage <- catch_numssshigh %>%
-  filter(species == "Mesopel_M_Fish")
-
-catageplot <- ggplot(catchage, aes(x=agecl, y=atoutput)) +
-  geom_point() +
-  theme_tufte() +
-  labs(subtitle = paste(scenario.name,
-                        catchage$species))
-
-catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
-catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
-catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
-catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
-
-```
-
-Hake will need `calc_stage2age` applied because it has 2 true ages per age class.
-
 Length sample with user specified max length bin (200 cm):
@@
   filter(species == "Pacific_sardine")
-
-lfplot <- ggplot(catchlen, aes(upper.bins)) +
-  geom_bar(aes(weight = atoutput)) +
-  theme_tufte() +
-  labs(subtitle = paste(scenario.name,
-                        catchlen$species))
-
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
-lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
-
-```
-
-Hake catch lengths:
-
-```{r vis-fishery-lf-test-2}
-
-catchlen <- catch_length_census_ss$natlength %>%
-  filter(species == "Mesopel_M_Fish")
 

### 2019-05-22 CreateStockSynthesis.Rmd

@@
 require(reshape2)
+require(dplyr)
 knitr::opts_chunk$set(echo = TRUE)
@@
 ```{r}
-sardine.data$catch
+names(sardine.data$catch) <- names(sardine.data$CPUE)
 
@@
 ```{r}
-results <- get(load("CalCurrentResults.RData"))
+results <- get(load("outputCCV3run_truth.RData"))
 devtools::load_all()
@@
 
-#Sample survey
-survey_out <- create_survey(dat=results$nums, time=seq(1,99,1), species=species, boxes=boxes, effic=effic, selex=sel)
+#Sample survey - 3rd timestep simulates a summer survey
+survey_out <- create_survey(dat=results$nums, time=seq(3,495,by=5), species=species, boxes=boxes, effic=effic, selex=sel)
 
@@
 
-#Sample survey numbers
-survObsNum <- sample_survey_numbers(dat=survey_out,cv=cv)
+sardine.catch <-  create_fishery_subset(dat = results$catch,
+                                         time = seq(3,495,by=5),
+                                         species = species,
+                                         boxes = boxes)
+
+#Sample catch in numbers
+sardine.catch.total <- sardine.catch %>%
+  group_by(time) %>%
+  summarise(catch=sum(atoutput))
+
 nyears <- 50
@@
 #Test writing CPUE in biomass
-sardine.data <- SS_write_CPUE(ss_data_list = sardine.data,
-              cpue_data = list(survObsBiom$atoutput[burnin:(burnin+nyears-1)]),
-              CVs = c(0.2),
-              data_years = c(nyears),
-              sampling_month = list(rep(1,nyears)),
-              units = "biomass")
+  sardine.data <- SS_write_ts(ss_data_list = sardine.data,
+                ts_data = list(survObsBiom$atoutput[burnin:(burnin+nyears-1)],
+                  sardine.catch.total$catch),
+                CVs = c(0.2,0.2),
+                data_years = list((survObsBiom$time[burnin:(burnin+nyears-1)]-3)/5+1, 
+                            (sardine.catch.total$time-3)/5+1),
+                sampling_month = list(rep(7,nyears),
+                                      rep(1,20)),
+                units = c("biomass","numbers"),
+                data_type=c("CPUE","catch"),
+                fleets = c(2,1))
 
-#Test writing CPUE in numbers
-sardine.data <- SS_write_CPUE(ss_data_list = sardine.data,
-                             cpue_data = list(survObsNum$atoutput[burnin:(burnin+nyears-1)]),
-                             CVs = c(0.2),
-                             data_years = c(nyears),
-                             sampling_month = list(rep(1,nyears)),
-                             units = "numbers")
-
-sardine.data$CPUE
 ```
@@
 age_comp_data <- sample_fish(survey_out, highEffN)
-names(age_comp_data) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")
 
@@
   group_by(time) %>%
-  mutate(comp = numAtAgesamp/sum(numAtAgesamp))
+  mutate(comp = atoutput/sum(atoutput))
 
@@
 ```
+
+## Setting up sardine EM
+For the sardine EM, we want conditional age-at-length (CAAL) and length composition data only. So we use the same function with our matrices of length and CAAL comp data to write to the data file.
+
+```{r}
+# aggregate true resn per survey design
+survey_aggresnstd <- aggregateDensityData(dat = results$resn,
+                                          time = unique(results$nums$time),
+                                          species = species,
+                                          boxes = boxes)
+
+# aggregate true structn per survey design
+survey_aggstructnstd <- aggregateDensityData(dat = results$structn,
+                                             time = unique(results$nums$time),
+                                             species = species,
+                                             boxes = boxes)
+
+ss_structnstd <- sample_fish(survey_aggstructnstd,
+                             effN,
+                             sample=FALSE)
+ss_resnstd <- sample_fish(survey_aggresnstd,
+                          effN,
+                          sample=FALSE)
+
+#Extract length composition data
+ss_length_stdsurv <- calc_age2length(structn = ss_structnstd,
+                                     resn = ss_resnstd,
+                                     nums = age_comp_data,
+                                     biolprm = results$biolprm, fgs = results$fgs,
+                                     CVlenage = 0.1, remove.zeroes=TRUE)
+
+#Check length compositions match age compositions
+ss_length_stdsurv$natlength %>%
+  group_by(time,agecl) %>%
+  summarise(sum(atoutput))
+
+len_comp_data <- ss_length_stdsurv$natlength
+
+```
+
+Once we have the natlength matrix, we can munge the data into the proper CAAL and length bin format for SS
+
+```{r}
+
+#We want to aggregate the data by year 
+caal_comp_proportion <- len_comp_data %>%
+  group_by(time) %>%
+  mutate(comp = round(atoutput/sum(atoutput),4)) %>%
+  select(agecl, time, atoutput, lower.bins,
+         upper.bins,comp)
+
+#Use dcast to turn the data from long to wide format
+unmelt_caal <- dcast(data = caal_comp_proportion,formula = time + lower.bins + upper.bins~agecl,value.var = "comp")
+unmelt_caal[is.na(unmelt_caal)] <- 0
+
+#Add nsamp column
+sample_size <- apply(unmelt_caal[,(4:ncol(unmelt_caal))], FUN=sum, MARGIN=1)
+caal_comp_flat<- cbind(unmelt_caal[order(unmelt_caal$time),],sample_size)
+names(caal_comp_flat)[ncol(caal_comp_flat)] <- "nsamp"
+
+#remove burnin
+caal_comp_final <- filter(caal_comp_flat,
+                         time %in% burnin:(burnin+nyears-1))
+
+len_comp_proportion <- len_comp_data %>%
+  group_by(time, lower.bins, upper.bins) %>%
+  summarise(lencomp = sum(atoutput))
+
+#Do the same for length composition
+unmelt_len <- dcast(data = len_comp_proportion,formula = time ~lower.bins,value.var = "lencomp")
+unmelt_len[is.na(unmelt_len)] <- 0
+
+#Add over age classes to get sample size
+sample_size <- apply(unmelt_len[,(2:ncol(unmelt_len))], FUN=sum, MARGIN=1)
+len_comp_flat<- cbind(unmelt_len[order(unmelt_len$time),],sample_size)
+names(len_comp_flat)[ncol(len_comp_flat)] <- "nsamp"
+
+#remove burnin
+len_comp_final <- filter(len_comp_flat,
+                         time %in% burnin:(burnin+nyears-1))
+
+length_bins <- as.integer(names(len_comp_final))
+length_bins <- length_bins[!is.na(length_bins)]
+
+# Write CAAL and length composition data
+sardine.data <- SS_write_comps(ss_data_list = sardine.data,
+                               comp_matrix = list(caal_comp_final,len_comp_final),
+                               data_rows = list(caal_comp_final$time,len_comp_final$time),
+                               sampling_month = list(rep(7,nrow(caal_comp_final)),
+               rep(7,nrow(len_comp_final))),
+                               data_type = c("agecomp", "lencomp"),
+                               fleet_number = c(1,1),
+                               bins = list(age_bins, length_bins),
+                               caal_bool = c(TRUE, FALSE))
+
+head(sardine.data$lencomp)
+head(sardine.data$agecomp)
+
+#Change length bin structure to match atlantis data
+sardine.data$lbin_vector <- length_bins
+
+#Set lbin_method to 1 - this makes the population length bins match the data bins 
+#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
+sardine.data$lbin_method <- 1
+sardine.data$binwidth <- "#"
+sardine.data$minimum_size <- "#"
+sardine.data$maximum_size <- "#"
+
+#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
+sardine.data$age_info$minsamplesize <- rep(0.001,2)
+
+SS_writedat_3.30(sardine.data, outfile = "./inst/extdata/Sardine_SS_files/sardEM_3_3.dat",
+                 overwrite=TRUE)
+```

### 2019-05-22 docs/index.Rmd
add sardines examples

@@
 
+### An example workflow for a single species (in progress)
+
+<a href="https://sgaichas.github.io/poseidon-dev/FullSardineTruthEx.html" target="_blank">Sardines in the CalCurrent, all true data, May 22, 2019</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/CreateStockSynthesis.html" target="_blank">Sardines in the CalCurrent, input to SS functions, May 22, 2019</a>
+
 ### Read this first on overall workflow
@@
 May 15: Changes to `atlantisom::calc_age2length` to allow user specified max length bin seem to work. Also listed under [Length compositions].
+
+May 20-22: Fisheries, corrections to nums at age for Cal Current and decision to read in the catch.txt file for total catch in weight
 

### 2019-05-23 CreateStockSynthesis.Rmd
Get fishery stuff working

@@
 
-#Need to replace with interp function
-wtAtAge <- data.frame(species=rep("Pacific_sardine",10),
-                agecl=1:10,
-                wtAtAge=c(0.0542, 0.0837,	0.1103,	0.1323,	0.1497,	0.163,	0.1729,	0.1801,	0.1854,	0.1941))
-cv <- data.frame(species="Pacific_sardine", cv=c(0.2))
-
-#Sample survey biomass
-survObsBiom <- sample_survey_biomass(dat=survey_out,cv=cv,wtAtAge)
-
-sardine.catch <-  create_fishery_subset(dat = results$catch,
-                                         time = seq(3,495,by=5),
-                                         species = species,
-                                         boxes = boxes)
-
-#Sample catch in numbers
-sardine.catch.total <- sardine.catch %>%
-  group_by(time) %>%
-  summarise(catch=sum(atoutput))
-
-nyears <- 50
-burnin <- 30
-
-#Test writing CPUE in biomass
-  sardine.data <- SS_write_ts(ss_data_list = sardine.data,
-                ts_data = list(survObsBiom$atoutput[burnin:(burnin+nyears-1)],
-                  sardine.catch.total$catch),
-                CVs = c(0.2,0.2),
-                data_years = list((survObsBiom$time[burnin:(burnin+nyears-1)]-3)/5+1, 
-                            (sardine.catch.total$time-3)/5+1),
-                sampling_month = list(rep(7,nyears),
-                                      rep(1,20)),
-                units = c("biomass","numbers"),
-                data_type=c("CPUE","catch"),
-                fleets = c(2,1))
-
-```
-
-## Writing age composition data
-
-```{r}
 #Set effective sample size for age compositions
@@
 
-age_comp_data
-
-#Group age composition data by time
-require(dplyr)
-age_comp_proportion <- age_comp_data %>%
-  group_by(time) %>%
-  mutate(comp = atoutput/sum(atoutput))
-
-#Get the age bins
-age_bin_names <- names(sardine.data$agecomp)[10:length(names(sardine.data$agecomp))]
-age_bins <- sub("a","",age_bin_names)
-
-#Change data format into flat structure
-age_comp_proportion$id <- rep(age_bins,
-                              each = nrow(age_comp_proportion)/length(age_bins))
-
-unmelt <- dcast(data = age_comp_proportion,
-                formula = time ~agecl,
-                value.var = "comp")
-age_comp_flat<- cbind(unmelt[order(unmelt$time),],
-                      rep(effN,nrow(unmelt)))
-names(age_comp_flat)[ncol(age_comp_flat)] <- "nsamp"
-
-## Write age composition data
-sardine.data <- SS_write_comps(ss_data_list = sardine.data,
-               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
-               data_rows = list(sardine.data$styr:(sardine.data$styr+nyears-1)),
-               sampling_month = list(rep(7,nyears)),
-               data_type = c("agecomp"),
-               fleet_number = c(1),
-               bins = list(age_bins),
-               caal_bool = c(FALSE))
-sardine.data$agecomp
-
-```
-
-## Setting up sardine EM
-For the sardine EM, we want conditional age-at-length (CAAL) and length composition data only. So we use the same function with our matrices of length and CAAL comp data to write to the data file.
-
-```{r}
 # aggregate true resn per survey design
@@
 
+#Need to replace with interp function
+wtAtAge <- ss_length_stdsurv$muweight %>%
+  select(species, agecl, time, wtAtAge = atoutput) %>%
+  mutate(wtAtAge = wtAtAge/1000)
+
+# CV function
+cv <- data.frame(species="Pacific_sardine", cv=0.1)
+
+#Sample survey biomass
+survObsBiom <- sample_survey_biomass(dat=survey_out,cv=cv,wtAtAge)
+
+```
+
+## Catch data
+
+```{r}
+sardine.catch <-  create_fishery_subset(dat = results$catch,
+                                         time = seq(1,99,by=1),
+                                         species = species,
+                                         boxes = boxes)
+
+#Sample catch in numbers
+sardine.catch.total <- sardine.catch %>%
+  group_by(time) %>%
+  summarise(catch=sum(atoutput))
+
+nyears <- 50
+burnin <- 30
+fish_years <- burnin:(burnin+nyears-1)
+survey_years <- seq(3,495,by=5)[burnin:(burnin+nyears-1)]
+
+#Test writing CPUE in biomass
+  sardine.data <- SS_write_ts(ss_data_list = sardine.data,
+                ts_data = list(survObsBiom$atoutput[fish_years],
+                  sardine.catch.total$catch[fish_years]),
+                CVs = c(0.1,0.1),
+                data_years = list((survObsBiom$time[fish_years]-3)/5+1,             sardine.catch.total$time[fish_years]),
+            sampling_month = list(rep(7,nyears),
+                                rep(1,nyears)),
+                units = c("biomass","numbers"),
+                data_type=c("CPUE","catch"),
+                fleets = c(2,1))
+
+```
+
+## Writing age composition data
+
+```{r}
+age_comp_data
+
+#Get the age bins
+age_bin_names <- names(sardine.data$agecomp)[10:length(names(sardine.data$agecomp))]
+age_bins <- sub("a","",age_bin_names)
+
+age_comp_flat <- reformat_compositions(age_comp_data,
+                     round.places = 4,
+                     comp_type = "agecomp")
+
+## Write age composition data for survey
+sardine.data <- SS_write_comps(ss_data_list = sardine.data,
+               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
+               data_rows = list(sardine.data$styr:(sardine.data$styr+nyears-1)),
+               sampling_month = list(rep(7,nyears)),
+               data_type = c("agecomp"),
+               fleet_number = c(1),
+               bins = list(age_bins),
+               caal_bool = c(FALSE))
+sardine.data$agecomp
+
+#We end up using CAAL for the survey below, so let's generate fishery age comps instead
+effN <- data.frame(species="Pacific_sardine", effN=effN)
+
+catch_numssshigh <- sample_fish(sardine.catch, effN)
+
+#Get weights
+# aggregate true resn per survey or fishery subset design
+catch_aggresnss <- aggregateDensityData(dat = results$resn,
+                                 time = c(0:111),
+                                 species = species,
+                                 boxes = boxes)
+
+# aggregate true structn per survey or fishery subsetdesign
+catch_aggstructnss <- aggregateDensityData(dat = results$structn,
+                                 time = c(0:111),
+                                 species = species,
+                                 boxes = boxes)
+
+#dont sample these, just aggregate them using median
+catch_structnss <- sample_fish(catch_aggstructnss, effN, sample = FALSE)
+
+catch_resnss <-  sample_fish(catch_aggresnss, effN, sample = FALSE)
+
+str(catch_numssshigh)
+
+catch_length <- calc_age2length(structn = catch_structnss,
+                                 resn = catch_resnss,
+                                 nums = catch_numssshigh,
+                                 biolprm = results$biolprm, fgs = results$fgs,
+                                 maxbin = 200,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+```
+
+## Setting up sardine EM
+For the sardine EM, we want conditional age-at-length (CAAL) and length composition data only. So we use the same function with our matrices of length and CAAL comp data to write to the data file.
+
+```{r}
+
 #Check length compositions match age compositions
@@
 len_comp_data <- ss_length_stdsurv$natlength
+fish_len_comp_data <- catch_length$natlength
 
@@
 
-#We want to aggregate the data by year 
-caal_comp_proportion <- len_comp_data %>%
-  group_by(time) %>%
-  mutate(comp = round(atoutput/sum(atoutput),4)) %>%
-  select(agecl, time, atoutput, lower.bins,
-         upper.bins,comp)
-
-#Use dcast to turn the data from long to wide format
-unmelt_caal <- dcast(data = caal_comp_proportion,formula = time + lower.bins + upper.bins~agecl,value.var = "comp")
-unmelt_caal[is.na(unmelt_caal)] <- 0
-
-#Add nsamp column
-sample_size <- apply(unmelt_caal[,(4:ncol(unmelt_caal))], FUN=sum, MARGIN=1)
-caal_comp_flat<- cbind(unmelt_caal[order(unmelt_caal$time),],sample_size)
-names(caal_comp_flat)[ncol(caal_comp_flat)] <- "nsamp"
+caal_comp_flat <- reformat_compositions(len_comp_data,                                round.places=4,
+                comp_type="caalcomp")
 
@@
 caal_comp_final <- filter(caal_comp_flat,
-                         time %in% burnin:(burnin+nyears-1))
-
-len_comp_proportion <- len_comp_data %>%
-  group_by(time, lower.bins, upper.bins) %>%
-  summarise(lencomp = sum(atoutput))
-
-#Do the same for length composition
-unmelt_len <- dcast(data = len_comp_proportion,formula = time ~lower.bins,value.var = "lencomp")
-unmelt_len[is.na(unmelt_len)] <- 0
+                         time %in% survey_years)
 
 #Add over age classes to get sample size
-sample_size <- apply(unmelt_len[,(2:ncol(unmelt_len))], FUN=sum, MARGIN=1)
-len_comp_flat<- cbind(unmelt_len[order(unmelt_len$time),],sample_size)
-names(len_comp_flat)[ncol(len_comp_flat)] <- "nsamp"
-
+len_comp_flat <- reformat_compositions(len_comp_data,
+                                  round.places = 0,
+                           comp_type="lencomp")
 #remove burnin
 len_comp_final <- filter(len_comp_flat,
-                         time %in% burnin:(burnin+nyears-1))
+                         time %in% survey_years)
 
@@
 
+fish_len_comp_flat <- reformat_compositions(catch_length$natlength,
+                                  round.places = 0,
+                           comp_type="lencomp")
+#remove burnin
+fish_len_comp_final <- filter(fish_len_comp_flat,
+                         time %in% fish_years)
+
+fish_age_comp_flat <- reformat_compositions(catch_numssshigh,
+                           comp_type="agecomp")
+#remove burnin
+fish_age_comp_final <- filter(fish_age_comp_flat,
+                         time %in% fish_years)
+
+comp_list <- list(caal_comp_final,len_comp_final,          fish_age_comp_final, fish_len_comp_final)
+
+apply_month <- list(rep(7, nrow(comp_list[[1]])), rep(7, nrow(comp_list[[2]])), rep(1,nrow(comp_list[[3]])),rep(1,nrow(comp_list[[4]])))
+
 # Write CAAL and length composition data
-sardine.data <- SS_write_comps(ss_data_list = sardine.data,
-                               comp_matrix = list(caal_comp_final,len_comp_final),
-                               data_rows = list(caal_comp_final$time,len_comp_final$time),
-                               sampling_month = list(rep(7,nrow(caal_comp_final)),
-               rep(7,nrow(len_comp_final))),
-                               data_type = c("agecomp", "lencomp"),
-                               fleet_number = c(1,1),
-                               bins = list(age_bins, length_bins),
-                               caal_bool = c(TRUE, FALSE))
+sardine.data <- SS_write_comps(ss_data_list = sardine.data, comp_matrix = comp_list, data_rows = list((comp_list[[1]]$time-3)/5 + 1 ,(survey_years-3)/5 + 1,fish_years,fish_years),
+        sampling_month = apply_month, data_type = rep(c("agecomp", "lencomp"),2), fleet_number = c(2,2,1,1),  bins = list(age_bins, length_bins, age_bins, length_bins), caal_bool = c(TRUE, rep(FALSE,3)))
 
@@
 ```
+
+## Coming up with dynamic life history parameters
+
+Below we get the weight-length relationship parameters from atlantis, calculate h and R0 from Atlantis $\alpha$ and $\beta$ parameters, use the \code{calc_Z} function to back out natural mortality (M) from Atlantis, and use the survey length-at-age from Atlantis to estimate a growth curve.
+```{r}
+#Get biological parameters
+biolprm <- results$biolprm
+fgs <- results$fgs
+li_a_use <- biolprm$wl[match(fgs$Code[match(species,fgs$Name)],biolprm$wl[, 1]), 2]/1000
+li_b_use <- biolprm$wl[match(fgs$Code[match(species,fgs$Name)],biolprm$wl[, 1]), 3]
+
+```

### 2019-05-23 FullSardineTruthEx.Rmd

@@
   .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
-  'data.table', 'here', 'ggforce', 'ggthemes'
+  'data.table', 'here', 'ggforce', 'ggthemes', 'r4ss'
 ), 'packages.bib')
@@
 
-On this page we demonstrate use of atlantisom on the California Current output files to provide inputs for a stock assessment model implemented in Stock Synthesis [insert SSREF].
+On this page we demonstrate use of atlantisom on the California Current output files to provide (true) inputs for a stock assessment model implemented in [Stock Synthesis](https://www.sciencedirect.com/science/article/abs/pii/S0165783612003293).
+Translation of atlantisom outputs to stock synthesis inputs uses the package [r4ss](https://github.com/r4ss) [@R-r4ss] as demonstrated [here](https://sgaichas.github.io/poseidon-dev/CreateStockSynthesis.html).
 
@@
 
-The following settings should achieve a survey that samples all Atlantis model output timesteps, all species, and all model polygons, with perfect efficiency and full selectivity for all ages:
-
 ```{r census-spec, message=FALSE, warning=FALSE}
@@
 ```
+
+Output timesteps and the duration of the model run can be found in the run parameter file [scenario]_run.xml which is read in with `atlantisom::load_runprm`.
+This is already run in our census_spec.R file.
+Output timestep toutinc is `r runpar$toutinc `, so steps per year is `r stepperyr ` and the number of output steps is `r noutsteps `.
+
+The settings in `config/census_spec.R` should achieve a survey that samples all Atlantis model output timesteps, all species, and all model polygons, with perfect efficiency and full selectivity for all ages:
 
@@
 
-Because we don't want to wait 24 hours for this, we will look at only our three speces and the first 112 time steps.
+Timesteps may be different for fishery output than for the population above.
+The run parameter file will tell us what time units we are dealing with, and this is already specified in the census_spec.R file as fstepperyr: `r fstepperyr `.
+Further, we can get the total number of years with nyears from the run parameter file: `r runpar$nyears `.
+
+Now we know the fishery output is annual (365 days), and there will only be 100 years. Because we don't want to wait 24 hours for this, we will look at only our three species.
 
@@
 catch_testNss <- create_fishery_subset(dat = truth$catch,
-                                 time = c(0:111),
+                                 time = timeall,
                                  species = spp.name,
@@
 catch_aggresnss <- aggregateDensityData(dat = truth$resn,
-                                 time = c(0:111),
+                                 time = timeall,
                                  species = spp.name,
@@
 catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
-                                 time = c(0:111),
+                                 time = timeall,
                                  species = spp.name,

### 2019-05-23 docs/index.Rmd

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/FullSardineTruthEx.html" target="_blank">Sardines in the CalCurrent, all true data, May 22, 2019</a>
+<a href="https://sgaichas.github.io/poseidon-dev/FullSardineTruthEx.html" target="_blank">Sardines in the CalCurrent, all true data, May 22, 2019, updated May 23, 2019</a>
 

### 2019-05-23 FullSardineTruthEx.Rmd
fix plots

@@
   gather(species, biomass, -Time) %>%
-  filter(species %in% levels(surveyB_frombio_sardine$species))
+  filter(species %in% c("Pacific_sardine"))
 
@@
 compareB <-ggplot() +
-  geom_line(data=surveyB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
+  geom_line(data=surveyB_frombio_sardine, aes(x=time/5,y=atoutput, color="survey census B"), 
             alpha = 10/10) +

### 2019-05-23 FullBoccaccioTruthEx.Rmd
try splitting ages, failing

--- 2019-05-23 FullSardineTruthEx.Rmd
@@
 ---
-title: "Sardine example: end to end truth for California Current"
+title: "Boccaccio example: end to end truth for California Current"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-surveyB_frombio_sardine <- surveyB_frombio[surveyB_frombio$species == "Pacific_sardine",]
+surveyB_frombio_ss <- surveyB_frombio[surveyB_frombio$species == "Bocaccio_rockfish",]
 
@@
   gather(species, biomass, -Time) %>%
-  filter(species %in% c("Pacific_sardine"))
+  filter(species %in% c("Bocaccio_rockfish"))
 
@@
 compareB <-ggplot() +
-  geom_line(data=surveyB_frombio_sardine, aes(x=time/5,y=atoutput, color="survey census B"), 
+  geom_line(data=surveyB_frombio_ss, aes(x=time/5,y=atoutput, color="survey census B"), 
             alpha = 10/10) +
@@
 
-surveyN_sardine <- surveyN[surveyN$species == "Pacific_sardine",]
+surveyN_ss <- surveyN[surveyN$species == "Bocaccio_rockfish",]
 
 plotN <-ggplot() +
-  geom_line(data=surveyN_sardine, aes(x=time/5,y=atoutput, color="survey census N"), 
+  geom_line(data=surveyN_ss, aes(x=time/5,y=atoutput, color="survey census N"), 
             alpha = 10/10) +
@@
 
-Natage_sardine <- Natage[Natage$species == "Pacific_sardine",]
+Natage_ss <- Natage[Natage$species == "Bocaccio_rockfish",]
 
-Natageplot <- ggplot(Natage_sardine, aes(x=agecl, y=atoutput)) +
+Natageplot <- ggplot(Natage_ss, aes(x=agecl, y=atoutput)) +
   geom_point() +
@@
   labs(subtitle = paste(scenario.name,
-                        Natage_sardine$species))
+                        Natage_ss$species))
 
@@
 Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
+### Biological sampling: true ages (split age classes)
+
+Bocaccio is one of many species in CCA with multiple true ages per age class.
+To get true ages for input into stock assessment, we need to separate the age classes.
+The atlantisom function `calc_stage2age` calls the `calc_Z` function, applying a total mortality rate to estimate the numbers at true age within an age class:
+
+```{r testagecl}
+# add YOY file to the config files
+YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))
+
+# load biolprm in some initialize file?
+biol <- load_biolprm(d.name, biol.prm.file)
+
+trueagesall_ss <- calc_stage2age(nums_data = Natage_ss,
+                                 biolprm = biol, 
+                                 yoy = YOY,
+                                 fgs = funct.groups, 
+                                 runprm = runpar)
 
@@
 len <- length_census_ss$natlength %>%
-  filter(species == "Pacific_sardine")
+  filter(species == "Bocaccio_rockfish")
 
@@
 We found [after much testing](https://sgaichas.github.io/poseidon-dev/Atlantisom2SSworkflowtest.html) that the most reliable source for catch in biomass at present is the Atlantis output catch.txt file, loaded above.
-Here is sardine annual total catch:
+Here is bocaccio annual total catch:
 
@@
 
-truecatchbio_sardine <- truecatchbio[truecatchbio$species == "Pacific_sardine",]
+truecatchbio_ss <- truecatchbio[truecatchbio$species == "Bocaccio_rockfish",]
 
 plotcatch <- ggplot() +
-  geom_line(data=truecatchbio_sardine, aes(x=time,y=atoutput, color="true catch bio"),
+  geom_line(data=truecatchbio_ss, aes(x=time,y=atoutput, color="true catch bio"),
             alpha = 10/10) +
@@
 
-This is true catch at age for sardine:
+This is true catch at age for bocaccio:
 
@@
 catchage <- catch_numssshigh %>%
-  filter(species == "Pacific_sardine")
+  filter(species == "Bocaccio_rockfish")
 
@@
 
-Sardine catch lengths:
+Bocaccio catch lengths:
 
@@
 catchlen <- catch_length_census_ss$natlength %>%
-  filter(species == "Pacific_sardine")
+  filter(species == "Bocaccio_rockfish")
 

### 2019-05-24 FullBoccaccioTruthEx.Rmd
with calc true age saved not plotted yet

@@
 
+This is Bocaccio true age class output. However, the 10 standard atlantis age classes aggregate 5 true ages each for Bocaccio.
+
 ```{r Natage-plot}
@@
 
-```{r testagecl}
+```{r testagecl, eval=FALSE}
 # add YOY file to the config files
@@
 
-trueagesall_ss <- calc_stage2age(nums_data = Natage_ss,
+trueagesall <- calc_stage2age(nums_data = truth$nums,
                                  biolprm = biol, 
@@
                                  runprm = runpar)
+
+saveRDS(trueagesall, file.path(d.name, paste0(scenario.name, "Nattrueagecensus.rds")) )
+
+```
+
+These should be true age comps for Bocaccio in annual ages:
+
+```{r plotannage}
+
+trueNatage <- readRDS(file.path(d.name, paste0(scenario.name, "Nattrueagecensus.rds")))
+
+# run survey  and sample fish on this for completeness?
+
+# plot it
 

### 2019-05-24 FullBoccaccioTruthEx.Rmd
add plotting code
Bocaccio example minus true age

@@
 
-```{r plotannage}
+**NOPE! Giving constant output or none. Back to the drawing board on true ages.**
+
+See new file TestCalcZ.Rmd testing `calc_Z` separately and then I will do another separate one testing `calc_stage2age`. Once they work I will update this.
+
+```{r plotannage, eval=FALSE}
 
@@
 
-# run survey  and sample fish on this for completeness?
+# test with one species--this is still full polygon and layer output
+trueNatage_ss <- trueNatage[trueNatage$species == "Bocaccio_rockfish",]
 
-# plot it
+# run survey  and sample fish on this for completeness? yes to aggregate
+survey_trueNage_ss <- create_survey(dat = trueNatage_ss,
+                                 time = timeall,
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
+trueNagesamp_ss <- sample_fish(survey_trueNage_ss, effNhigh)
+
+# plot it--first few timesteps for bocaccio
+
+Natageplot <- ggplot(trueNagesamp_ss, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        trueNatage_ss$species))
+
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
 

### 2019-05-24 CreateStockSynthesis.Rmd
Make CreateStockSynthesis.Rmd work in more general case across species by modifing species config file

@@
 This page discusses reading in an existing stock assessment file and replacing the data values with those from the `atlantisom` package.
-To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`
+To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`.
+You will need to ensure the relevant Stock Synthesis files are stored in the inst/extdata/ folder, with each species example in its own folder.
 
@@
 
-Dummy files are in the package directory; use `r4ss` to load the dummy data file
+Dummy files are in the package directory; use `r4ss` to load the dummy data file.
+Use the relevant species config file to load the configuration variables for each run.
+Note: this script will extract conditional age-at-length data, length composition data, and a CPUE in biomass for one survey during the specified time interval, and age composition data, length composition data, and catch in numbers from a single fishery.
+The timing and duration of the survey and fishery should be set in the config file.
+Unless otherwise stated, numbers going into Stock Synthesis are in thousands of fish.
+Weight of a single fish is in grams or kilograms, and weight of biomass is in metric tons.
+We assume a single sex model with a sex ratio fixed at 0.5 - different assumptions require modifying the Stock Synthesis control file.
+This script only runs with Stock Synthesis 3.30.
 
@@
 require(r4ss)
-#Read in sardine data file
-sardine.data <- r4ss::SS_readdat_3.30("./inst/extdata/Sardine_SS_files/sardEM_3_3.dat")
+source("./config/sardine_config.R")
+
+stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/",
+model_dir,
+datfile_name))
 ```
@@
 ```{r}
-names(sardine.data$catch) <- names(sardine.data$CPUE)
+names(stocksynthesis.data$catch) <- names(stocksynthesis.data$CPUE)
 
-sardine.data$CPUE
+stocksynthesis.data$CPUE
 ```
@@
 ```{r}
-sardine.data$agecomp
+stocksynthesis.data$agecomp
 
-sardine.data$lencomp
+stocksynthesis.data$lencomp
 ```
@@
 #CPUE units, SD and error structure
-sardine.data$CPUEinfo
+stocksynthesis.data$CPUEinfo
 
 #Number of length bins
-sardine.data$N_lbins
+stocksynthesis.data$N_lbins
 
 #Length bin method
-sardine.data$Lbin_method
+stocksynthesis.data$Lbin_method
 
 #Vector of age bins
-sardine.data$agebin_vector
+stocksynthesis.data$agebin_vector
 
@@
 ```{r}
-results <- get(load("outputCCV3run_truth.RData"))
+results <- get(load(atlantis_output_name))
 devtools::load_all()
-species=c("Pacific_sardine")
+species=c(species)
 
@@
 #Efficiency of the survey
-effic <- data.frame(species=species, efficiency=0.5)
+effic <- data.frame(species=species, efficiency=eff_case)
 
 # Assume uniform selectivity
-sel<- data.frame(species=rep(species,10),
-               agecl=c(1:10),
-              selex=rep(1,10))
+sel<- data.frame(species=rep(species,length(age_classes)),
+               agecl=age_classes,
+              selex=sel_by_age)
 
 #Sample survey - 3rd timestep simulates a summer survey
-survey_out <- create_survey(dat=results$nums, time=seq(3,495,by=5), species=species, boxes=boxes, effic=effic, selex=sel)
+survey_out <- create_survey(dat=results$nums, time=survey_sample_full, species=species, boxes=boxes, effic=effic, selex=sel)
 
 #Set effective sample size for age compositions
-effN <- 1000
+effN <- surveyEffN
 highEffN <- data.frame(species=species, effN=rep(effN, length(species)))
@@
                                      biolprm = results$biolprm, fgs = results$fgs,
-                                     CVlenage = 0.1, remove.zeroes=TRUE)
+                                     CVlenage = CVs$lenage, remove.zeroes=TRUE)
 
@@
 # CV function
-cv <- data.frame(species="Pacific_sardine", cv=0.1)
+cv <- data.frame(species="Pacific_sardine", cv=CVs$survey)
 
@@
 ```{r}
-sardine.catch <-  create_fishery_subset(dat = results$catch,
-                                         time = seq(1,99,by=1),
+catch_numbers <-  create_fishery_subset(dat = results$catch,
+                                         time = fish_years,
                                          species = species,
@@
 #Sample catch in numbers
-sardine.catch.total <- sardine.catch %>%
+catch_total <- catch_numbers %>%
   group_by(time) %>%
@@
 
-nyears <- 50
-burnin <- 30
-fish_years <- burnin:(burnin+nyears-1)
-survey_years <- seq(3,495,by=5)[burnin:(burnin+nyears-1)]
-
 #Test writing CPUE in biomass
-  sardine.data <- SS_write_ts(ss_data_list = sardine.data,
+  stocksynthesis.data <- SS_write_ts(ss_data_list = stocksynthesis.data,
                 ts_data = list(survObsBiom$atoutput[fish_years],
-                  sardine.catch.total$catch[fish_years]),
-                CVs = c(0.1,0.1),
-                data_years = list((survObsBiom$time[fish_years]-3)/5+1,             sardine.catch.total$time[fish_years]),
-            sampling_month = list(rep(7,nyears),
-                                rep(1,nyears)),
+                  catch_total$catch[fish_years]),
+                CVs = c(CVs$survey,
+                        CVs$fishery),
+                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             catch_total$time[fish_years]),
+            sampling_month = list(rep(survey_month,nyears),
+                                rep(fishing_month,nyears)),
                 units = c("biomass","numbers"),
@@
 #Get the age bins
-age_bin_names <- names(sardine.data$agecomp)[10:length(names(sardine.data$agecomp))]
+age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
 age_bins <- sub("a","",age_bin_names)
@@
 ## Write age composition data for survey
-sardine.data <- SS_write_comps(ss_data_list = sardine.data,
+stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data,
                comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
-               data_rows = list(sardine.data$styr:(sardine.data$styr+nyears-1)),
-               sampling_month = list(rep(7,nyears)),
+               data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr+nyears-1)),
+               sampling_month = list(rep(survey_month,nyears)),
                data_type = c("agecomp"),
@@
                caal_bool = c(FALSE))
-sardine.data$agecomp
+stocksynthesis.data$agecomp
 
 #We end up using CAAL for the survey below, so let's generate fishery age comps instead
-effN <- data.frame(species="Pacific_sardine", effN=effN)
+effN <- data.frame(species=species, effN=effN)
 
-catch_numssshigh <- sample_fish(sardine.catch, effN)
+catch_numssshigh <- sample_fish(catch_numbers, effN)
 
@@
                                  biolprm = results$biolprm, fgs = results$fgs,
-                                 maxbin = 200,
-                                 CVlenage = 0.1, remove.zeroes=TRUE)
+                                 maxbin = maxbin,
+                                 CVlenage = CVs$lenage, remove.zeroes=TRUE)
 
@@
 
-## Setting up sardine EM
-For the sardine EM, we want conditional age-at-length (CAAL) and length composition data only. So we use the same function with our matrices of length and CAAL comp data to write to the data file.
+## Setting up EM
+For the EM, we want conditional age-at-length (CAAL) and length composition data only. So we use the same function with our matrices of length and CAAL comp data to write to the data file.
 
@@
 
-apply_month <- list(rep(7, nrow(comp_list[[1]])), rep(7, nrow(comp_list[[2]])), rep(1,nrow(comp_list[[3]])),rep(1,nrow(comp_list[[4]])))
+apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), rep(survey_month, nrow(comp_list[[2]])), rep(fishing_month,nrow(comp_list[[3]])),rep(fishing_month,nrow(comp_list[[4]])))
 
 # Write CAAL and length composition data
-sardine.data <- SS_write_comps(ss_data_list = sardine.data, comp_matrix = comp_list, data_rows = list((comp_list[[1]]$time-3)/5 + 1 ,(survey_years-3)/5 + 1,fish_years,fish_years),
+stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data, comp_matrix = comp_list, data_rows = list((comp_list[[1]]$time-survey_sample_time)/timestep + 1 ,(survey_years-survey_sample_time)/timestep + 1,fish_years,fish_years),
         sampling_month = apply_month, data_type = rep(c("agecomp", "lencomp"),2), fleet_number = c(2,2,1,1),  bins = list(age_bins, length_bins, age_bins, length_bins), caal_bool = c(TRUE, rep(FALSE,3)))
 
-head(sardine.data$lencomp)
-head(sardine.data$agecomp)
+head(stocksynthesis.data$lencomp)
+head(stocksynthesis.data$agecomp)
 
 #Change length bin structure to match atlantis data
-sardine.data$lbin_vector <- length_bins
+stocksynthesis.data$lbin_vector <- length_bins
 
@@
 #When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
-sardine.data$lbin_method <- 1
-sardine.data$binwidth <- "#"
-sardine.data$minimum_size <- "#"
-sardine.data$maximum_size <- "#"
+stocksynthesis.data$lbin_method <- 1
+stocksynthesis.data$binwidth <- "#"
+stocksynthesis.data$minimum_size <- "#"
+stocksynthesis.data$maximum_size <- "#"
 
 #Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
-sardine.data$age_info$minsamplesize <- rep(0.001,2)
+stocksynthesis.data$age_info$minsamplesize <- rep(0.001,2)
 
-SS_writedat_3.30(sardine.data, outfile = "./inst/extdata/Sardine_SS_files/sardEM_3_3.dat",
+SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
+datfile_name),
                  overwrite=TRUE)
@@
 
+len_nonburn <- ss_length_stdsurv$mulen %>%
+  filter(time > burnin*timestep)
+plot(len_nonburn$atoutput~len_nonburn$agecl, col=len_nonburn$time)
+
+length.data <- data.frame("Year"=(len_nonburn$time-survey_sample_time)/timestep+1, length=len_nonburn$atoutput, Weight=NA, Sex="Female", age=as.integer(len_nonburn$agecl))
+
+#require(bbmle)
+#vb_estimates <-sample_fit_vbgf(length.data, 25, 45, 0.4,
+#  0.1, 0.1, 20, 40, 0.05, 0.01, 0.01,
+#  30, 50, 0.6, 0.3, 0.3, 0.5, 10)
+
 ```
+
+## Running stock synthesis
+After the data and control file modifications, ensure you copy the local SS 3.3 executable into the folder containing the model files. Then run the following chunk.
+
+```{r}
+#Run stock synthesis model
+run_stocksynthesis(species = species, model_dir = model_dir, show_output = TRUE)
+
+```

### 2019-05-24 docs/index.Rmd

@@
 <a href="https://sgaichas.github.io/poseidon-dev/CreateStockSynthesis.html" target="_blank">Sardines in the CalCurrent, input to SS functions, May 22, 2019</a>
+
+May 24: Sardines have a full dataset, but still working on converting to true ages for species with multiple ages per age class. See below under [True age compositions].
 
@@
 
+### True age compositions
+
+<a href="https://sgaichas.github.io/poseidon-dev/FullBoccaccioTruthEx.html" target="_blank">Initial testing of true age functions, now running but output wrong - May 24, 2019</a>
+
 ### More to come later!

### 2019-05-24 docs/index.Rmd
Update index.html with latest date and notes for cal current write to SS Rmd

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/CreateStockSynthesis.html" target="_blank">Sardines in the CalCurrent, input to SS functions, May 22, 2019</a>
+<a href="https://sgaichas.github.io/poseidon-dev/CreateStockSynthesis.html" target="_blank">Sardines in the CalCurrent, input to SS functions: May 24, 2019</a> - now generalized to run for any species in the cal current with a unique species config file
 

### 2019-05-24 TestCalcZ.Rmd
testing calc_Z in progress

---
title: "Getting mortality from atlantisom: testing calc_Z function I"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
  'data.table', 'here', 'ggforce', 'ggthemes', 'r4ss'
), 'packages.bib')
```

## Introduction

This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA. 
On this page we demonstrate use of atlantisom on Norwegian Barents Sea Atlantis (NOBA) output files to test function `calc_Z` calculating total mortality. This is used both to derive "natural mortality" for comparison with stock assessments and to split age classes into true ages using the function `calc_stage2age`.

NOBA has output files in true ages that we can compare results with. These files are in the folder NOBAwithAnnAgeOutput (not kept on github due to large filesize).

All model setup and configuration is described [here](https://sgaichas.github.io/poseidon-dev/FullSardineTruthEx.html).

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)
```

```{r initialize}

initCCA <- FALSE
initNEUS <- FALSE
initNOBA <- TRUE

#function to make a config file? need one for each atlantis run

if(initCCA) source(here("config/CC1Config.R"))

if(initNEUS) source(here("config/NEUSConfig.R"))

if(initNOBA) source(here("config/NOBAaaConfig.R"))

```

```{r get_names, message=FALSE, warning=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```

```{r get_truth, message=FALSE, warning=FALSE}

# default run_truth setup will save the file, so check for that first

if(!file.exists(file.path(d.name, 
                          paste0("output", scenario.name, "run_truth.RData")))){
  #Store all loaded results into an R object
  truth <- run_truth(scenario = scenario.name,
                     dir = d.name,
                     file_fgs = functional.groups.file,
                     file_bgm = box.file,
                     select_groups = funct.group.names,
                     file_init = initial.conditions.file,
                     file_biolprm = biol.prm.file,
                     file_runprm = run.prm.file
  )
} else{
  truth <- get(load(file.path(d.name,
                              paste0("output", scenario.name, "run_truth.RData"))))
}
```

```{r loadcatchb}

truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)

```

## Test calc Z function

This section tests the `atlantisom::calc_Z()` function step by step. Ultimately this output goes to `calc_stage2age` to produce annual numbers at true age, which we can compare with the ANNAGEBIO.nc output produced by NOBA.

```{r census-spec, message=FALSE, warning=FALSE}

# make a function for this
source(here("config/census_spec.R"))

```

Output timestep toutinc for the population is `r runpar$toutinc `,

so steps per year in run_truth output is `r stepperyr `

and the number of output steps in run_truth output is `r noutsteps `.

### True numbers total (annual)

This gives us the true total numbers on an annual basis. We can compare this with age specific outputs.

```{r surveyNbased, eval=FALSE}

# this uses result$nums and a new function to get survey index in numbers (abundance)

survey_testNall <- create_survey(dat = truth$nums,
                                 time = timeall,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# save this, needed as input to calc_Z?
saveRDS(survey_testNall, file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))

# as above, make up a constant 0 cv for testing
surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))

surveyN <- sample_survey_numbers(survey_testNall, surv_cv)

#save for later use, takes a long time to generate
saveRDS(surveyN, file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))

```

```{r survNtrue-plot}

surveyN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))

surveyN_ss <- surveyN[surveyN$species == "North_atl_cod",]

plotN <-ggplot() +
  geom_line(data=surveyN_ss, aes(x=time/5,y=atoutput, color="survey census N"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotN +
  facet_wrap(~species, scales="free")

```

### Biological sampling: true numbers at age class

We get true numbers at age at each output step by appling `sample_fish` to the survey_testNall result if the model group has one true age per output age class:

```{r truenatagecl, eval=FALSE}
# We need an effective N for the sample fish function, setting equal to actual N
# this one is high but not equal to total for numerous groups
effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))

numsallhigh <- sample_fish(survey_testNall, effNhigh)

#save for later use, takes a long time to generate
saveRDS(numsallhigh, file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))

```

This is NOBA Cod true age class output. However, the 10 standard atlantis age classes aggregate true ages each for Cod.

Number of true age classes is stored in the [...]groups[...].csv file, which is read in using `load_fgs`. We read this in as funct.groups above. The column NumAgeClassSize stores the number of true ages:

Number of age classes for Cod: `r funct.groups$NumAgeClassSize[funct.groups$Name=="North_atl_cod"] `

```{r Natage-plot}

Natage <- readRDS(file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))

Natage_ss <- Natage[Natage$species == "North_atl_cod",]

Natageplot <- ggplot(Natage_ss, aes(x=agecl, y=atoutput)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        Natage_ss$species))

Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free")

```

### What level of aggregation works for calc_Z?

]The atlantisom function `calc_stage2age` calls the `calc_Z` function, applying a total mortality rate to estimate the numbers at true age within an age class, but it is unclear whether to use this on the aggregated numbers or the full resolution numbers at age. Here we test outputs of Z for several levels of aggregation:

```{r testZagg}
# add YOY file to the config files
YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))

# load biolprm in some initialize file?
biol <- load_biolprm(d.name, biol.prm.file)

# cut to a single species in YOY file, hardcoded for NOBA cod
YOY_ss <- YOY %>%
  select(Time, NCO.0)

# numbers at agecl at full resolution (all polygons and layers)
truenums_ss <- truth$nums[truth$nums$species == "North_atl_cod",]

survey_testNall <- readRDS(file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))

# numbers at ageclaggregated over layers, retain polygons (output of create_survey)
survey_testN_ss <- survey_testNall[survey_testNall$species == "North_atl_cod",]

# numbers at agecl aggregated over layers and polygons (output of sample_fish)
# done above
# Natage_ss <- Natage[Natage$species == "North_atl_cod",]

#calc_Z <- function(yoy, nums, fgs, biolprm, toutinc)

fullresZ <- calc_Z(yoy = YOY_ss,
                   nums = truenums_ss,
                   fgs = funct.groups,
                   biolprm = biol,
                   toutinc = runpar$toutinc)

surveyresZ <- calc_Z(yoy = YOY_ss,
                   nums = survey_testN_ss,
                   fgs = funct.groups,
                   biolprm = biol,
                   toutinc = runpar$toutinc)

sampleresZ <- calc_Z(yoy = YOY_ss,
                   nums = Natage_ss,
                   fgs = funct.groups,
                   biolprm = biol,
                   toutinc = runpar$toutinc)

```

Fixed a bug where grep was used if a # appeared, none of those in NOBA, will have to see if output is produced in CCA. This ran for the survey resolution numbers. I then went back and ran the two other levels above after fixing `calc_Z.`

Something is wrong with running after `sample_fish` because sampleresZ returned all 0s. Both true full resolution numbers and survey resolution numbers (aggregated over polygon) returned nonzero Z values. They are identical aside from seven values with no apparent pattern (rows 84, 85, 130, 131, 147, 172, 196) out of 201 that differ at e-16. I would be happer if the differences were all exactly 0, but I suppose we can attribute that to rounding error.

```{r calczsnippet, eval=FALSE}

# test with one species at survey scale (layers aggregated, not time or polygon)
yoy <- YOY_ss
nums <- survey_testN_ss
fgs <- funct.groups
biolprm <- biol
toutinc <- runpar$toutinc

# everything below the function definition line
  # subset the yoy for species included in the fgs file
  # that are turned on
  # colnames of the recruit data are "Time" and a column for each
  # species where the name is the functional group with an additional .0
  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
  recruits <- yoy[, colnames(yoy) %in% c("Time", paste0(turnedon$Code, ".0"))]

  # mg carbon converted to wet weight in tonnes
  k_wetdry <- biolprm$kgw2d / 1000000000
  # Sum of structural and reserve nitrogen (KWSR_RN + KWSR_SN)
  nitro <- merge(biolprm$kswr, biolprm$kwrr, by = "1")
  nitro$sum <- apply(nitro[, 2:3], 1, sum)

  # legacy: If output is from legacy code there will be an error in the
  # yoy data, where the first row is in a different unit.
  # yoy.txt is the biomass in tonnes per spawning event summed over the total
  # model domain.
  # The first row (< Nov/Dec 2015) is stored as biomass and the remaining rows
  # are stored in numbers, must convert the entire matrix to biomass
  # Check if legacy code and if so convert the numbers to biomass

  # G.Fay 2/21/16 : changed yoy to recruits in below loop.
  if (abs(recruits[1, 2] / recruits[2, 2]) > 10) {
    recruits[2:NROW(recruits), 2:NCOL(recruits)] <- recruits[2:NROW(recruits), 2:NCOL(recruits)] *
    nitro[match(gsub(".0", "", colnames(recruits)[-1]), nitro[, 1]), "sum"]
  }

  # Wide to long
  recruits <- reshape(data = recruits, direction = "long",
    varying = colnames(recruits)[-1],
    v.names = "recruits",
    times = colnames(recruits)[-1],
    timevar = "group")
  rownames(recruits) <- 1:NROW(recruits)
  recruits <- recruits[, -which(colnames(recruits) == "id")]
  # Switch from species code to species
  recruits$group <- gsub("\\.0", "", recruits$group)
  recruits <- merge(recruits, fgs[, c("Code", "Name")],
    by.x = "group", by.y = "Code")

  # merge recruits with strucn and resn of recruits from biol.prm file
  recruits <- merge(recruits, nitro[, c(1, 4)],
    by.x = "group", by.y = "1", all.x = TRUE, all.y = FALSE)
  colnames(recruits)[which(colnames(recruits) == "recruits")] <- "recruitsbio"
  # Get recruits in numbers rather than biomass
  recruits$recruits <- recruits$recruitsbio / recruits$sum

  # G.Fay 2/21/16
  # UGLY code below tries to align fraction of annual yoy with timing of recruitment
  # values in YOY.txt are total YOY that year waiting to recruit.
  # seems to get rid of most of 'issues' - some v.minor survival >1,
  #perhaps due to averaging of survival over toutinc days
  nyrs <- ceiling(max(yoy$Time)/365)
  times <- unique(yoy$Time)
  recstart_temp <- biolprm$recruit_time
  #recstart_temp[,2] <- biolprm$time_spawn[-(grep('#',
  #                     biolprm$time_spawn[,1])),2] + biolprm$recruit_time[,2]
  # was this grep to get rid of a species with #XXX? breaks when there are none
  recstart_temp[,2] <- biolprm$time_spawn[,2] + biolprm$recruit_time[,2]

  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%turnedon$Code,] #bug added codes with no YOY output
  recstart_temp <- recstart_temp[recstart_temp[,1]%in%recruits$group,]
  recruits$frac_recruit <- 0
  for (irow in 1:nrow(recstart_temp)) {
    group <- recstart_temp[irow,1]
    pick <- which(recruits$group == group)

    recstart <- seq(recstart_temp[irow,2],by=365,length.out=nyrs)
    recstart <- recstart[recstart<max(recruits$Time[pick])]
    recend <- recstart + biolprm$recruit_period[irow,2]
    #rec_times <- rbind(rec_times,cbind(group,recstart,recend))

    for (i_rec in 1:length(recstart)) {
      i_tstart <- which(pick==min(pick[recruits$Time[pick]>=recstart[i_rec]]))
      i_tstop <- which(pick==min(pick[recruits$Time[pick]>=recend[i_rec]]))
      if (i_rec == length(recstart)) i_tstop <- length(pick)
      n_t <- 1+i_tstop-i_tstart
      for (i_t in 1:n_t) {
        t_temp <- recruits$Time[pick[i_tstart+i_t-1]]
        num_temp <- t_temp - recstart[i_rec]
        if (i_t>1) {
          if ((recend[i_rec]-t_temp)>toutinc) {
            num_temp <- toutinc
          }
          else {
            num_temp <- recend[i_rec]-(recruits$Time[pick[i_tstart+i_t-2]])
          }
        }
        frac_temp <- max(c(0,num_temp /(recend[i_rec]-recstart[i_rec])))
        recruits$frac_recruit[pick[i_tstart+i_t-1]] <- frac_temp
      }
    }
  }

  # match "Time" of the young of the year with the time-step periodicity
  # listed in the run.prm or run.xml file
  recruits$Time <- recruits$Time / toutinc

  #G.Fay 1/6/16, expand to num of recruits
  # Recruit / mg C converted to wet weight in tonnes / redfield ratio of C:N
  recruits$recruits <- recruits$recruits/k_wetdry/biolprm$redfieldcn

  # Sum over all boxes/depth/cohorts
  totnums <- aggregate(atoutput ~ species + time, data = nums, sum)

  # Combine recruits and numbers
  # Only pull recruits from the yearly time step, where the
  # yearly time step matches the time step
  totnums <- merge(totnums, recruits,
    by.x = c("time", "species"), by.y = c("Time", "Name"),
    all.x = TRUE, all.y = FALSE)
  totnums$group <- recruits$group[match(totnums$species, recruits$Name)]
  # For all time increments where there
  totnums$recruits[is.na(totnums$recruits)] <- 0

  # Calculate survivors for each species group
  totnums$survivors <- totnums$recruits
  # Make sure time is in order
  totnums <- totnums[order(totnums$species, totnums$time), ]

  totnums$recruits <- totnums$recruits*totnums$frac_recruit

  totnums$survival <- totnums$survivors
  # Calculate survival for each group
  for (group in unique(totnums$group)) {
    #if(group == "SHD") browser()
    pick <- which(totnums$group == group)
    survival_temp <- c(NA,
      totnums$survivors[pick[-1]]/totnums$atoutput[pick[-length(pick)]])

    # G.Fay 2/21/16  "think" this is what things should be, recruits don't show up
    # in numbers at age until time step after the recruitment event.
     survival_temp <- c(
       (totnums$atoutput[pick[-1]]-
          totnums$recruits[pick[-1]])/totnums$atoutput[pick[-length(pick)]],NA)

#     survival_temp <- c(
#       (totnums$atoutput[pick[-1]])/
#          (totnums$recruits[pick[-1]]+totnums$atoutput[pick[-length(pick)]]),NA)

    survival_temp[survival_temp < 0] <- NA
    # Use first positive value to replace the initial year and all negative vals
    firstgood <- which(!is.na(survival_temp))[1]

    survival_temp[1:firstgood] <- survival_temp[firstgood]
    for(ii in seq_along(survival_temp)) {
      if (is.na(survival_temp[ii])) {
        nonzero <- which(which(survival_temp > 0) > ii)
        if (length(nonzero) == 0) nonzero <- which(survival_temp > 0)
        survival_temp[ii] <- survival_temp[which.min(abs(nonzero - ii))]
      }
    }
    totnums$survival[pick] <- survival_temp
   }

  #Calculate Z
  totnums$Z <- -1 * log(totnums$survival)
  finaldata <- data.frame("species" = totnums$species,
    "agecl" = NA, "polygon" = NA, "layer" = NA,
    "time" = totnums$time, "atoutput" = totnums$Z)

surveyresZ <- finaldata

```

### What is F at each timestep?
This probably depends on how the model is set up, but for models forced with F by species we ought to be able to read this out. If it is a daily rate then we mulitply by toutinc to get the rate at each of these output timestep Z values.

### Quick test to see if Z is right?
Perhaps overly optimistic, but if we feed these into `calc_stage2age` and get a result that matches the NOBA output ANNAGEBIO.nc, then we can assume the Zs are correct.

These are estimated true age comps for Cod in annual ages:

```{r plotannage}

# for single species, Z was calculated above as surveyresZ
trueages_ss <- calc_stage2age(nums_data = survey_testN_ss,
                                 biolprm = biol, 
                                 yoy = YOY_ss,
                                 fgs = funct.groups, 
                                 runprm = runpar)

# run sample fish on this for completeness? yes to aggregate
trueNagesamp_ss <- sample_fish(trueages_ss, effNhigh)

# plot it

Natageplot <- ggplot(trueNagesamp_ss, aes(x=agecl, y=atoutput)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        trueNagesamp_ss$species))

Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")

```

Not looking great, but at least not constant. This could be wrong Zs or wrong calculations in `calc_stage2age`. Load true ages from ANNAGEBIO.nc:

```{r compareages}

annage_nc <- paste0("output", scenario.name, "ANNAGEBIO.nc")
bboxes <- get_boundary(boxpars)

# need to change load_nc, at line 117 hardcoded for 10 cohorts!
trueage_atout <- load_nc(dir = d.name,
                         file_nc = annage_nc,
                         bps = boxpars,
                         fgs = funct.groups,
                         select_groups = survspp,
                         select_variable = "Nums",
                         check_acronyms = TRUE,
                         bboxes = bboxes)

trueage_atout_ss <- trueage_atout[trueage_atout$species == "North_atl_cod",]

#survey census to agg over layers
survey_trueage_atout_ss <- create_survey(dat = trueage_atout_ss,
                                 time = timeall,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# sample fish to agg over polygons
sample_trueage_atout_ss <- sample_fish(survey_trueage_atout_ss, effNhigh)

```

Compare true age output from atlantis with our estimate (and with aggregated age comps):

```{r plot-esttrueage-atltrueage}

compareAge <- ggplot() +
  geom_point(data=trueNagesamp_ss, aes(x=agecl,y=atoutput, color="atlantisom"), alpha = 10/10) +
  geom_point(data=sample_trueage_atout_ss, aes(x=agecl,y=atoutput, color="atout true"), alpha = 10/10) + 
  geom_point(data=Natage_ss, aes(x=agecl*2, y=atoutput, color="cohort true"), alpha = 10/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(subtitle = paste(scenario.name,
                        trueNagesamp_ss$species))

compareAge + 
  facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
compareAge + 
  facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
compareAge + 
  facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
compareAge + 
  facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")

```

We only have the first 10 cohorts of true age output due to hardcoding in load_nc. I may write another load function that changes that. Either way our early ages don't match. Something about the calculation or application of Z isn't correct.

### 2019-05-24 docs/index.Rmd
testing calc_Z in progress

--- 2019-05-24 docs/index.Rmd
@@
 
-Work TBD applying `calc_stage2age` and `calc_Z` to get true ages
-
 ### Testing a full workflow through fisheries (sardine, agecl = true age)
@@
 
+<a href="https://sgaichas.github.io/poseidon-dev/TestCalcZ.html" target="_blank">Test Z calc and compare with true annage NOBA - May 24, 2019</a>
+
+May 24: Either `calc_Z` or `calc_stage2age` or both are incorrect if we expect them to exactly match atlantis annual age ouput where it is available. We match in timestep 0, then diverge.
+
 ### More to come later!

### 2019-05-24 config/NOBAaaConfig.R

@@ -1,0 +1,10 @@
+d.name <- here("atlantisoutput","NOBAwithAnnAgeOutput")
+functional.groups.file <- "nordic_groups_v04.csv" 
+biomass.pools.file <- "nordic_biol_v23.nc"
+biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
+box.file <- "Nordic02.bgm"
+initial.conditions.file <- "nordic_biol_v23.nc"
+run.prm.file <- "nordic_run_v01.xml"
+scenario.name <- "nordic_runresults_01"
+bioind.file <- "outputnordic_runresults_01BiomIndx.txt"
+catch.file <- "outputnordic_runresults_01Catch.txt"

### 2019-05-27 FullBoccaccioTruthEx.Rmd
Z correction broken here too but pattern more apparent for further diagnosis

--- 2019-05-24 FullBoccaccioTruthEx.Rmd
@@
 
-```{r testagecl, eval=FALSE}
+```{r testagecl, eval = FALSE}
 # add YOY file to the config files
@@
 
-trueagesall <- calc_stage2age(nums_data = truth$nums,
+trueagesall <- calc_stage2age(nums_data = survey_testNall,
                                  biolprm = biol, 
@@
 
-```{r plotannage}
+After corrections were applied to `calc_Z` we were closer to the ballpark in NOBA.
+This demonstrates that the estimated true ages are not working between the cohort ages, they just drop to 0.
+This is consistent with what we observed for NOBA.
+The numbers at true age [don't match atlantis output where we have it for NOBA](https://sgaichas.github.io/poseidon-dev/TestCalcZ.html), so I will do further testing with this function and `calc_stage2age`.
+
+```{r plotannage, message=FALSE, warning=FALSE}
 
@@
 
-# run survey  and sample fish on this for completeness?
+# test with one species--this is still full polygon and layer output
+trueNatage_ss <- trueNatage[trueNatage$species == "Bocaccio_rockfish",]
 
-# plot it
+# run sample fish on this for completeness? yes to aggregate
+trueNagesamp_ss <- sample_fish(trueNatage_ss, effNhigh)
+
+# plot it--first few timesteps for bocaccio
+
+Natageplot <- ggplot(trueNagesamp_ss, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        trueNatage_ss$species))
+
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
 

### 2019-05-27 docs/index.Rmd
add still broken bocaccio ages

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/FullBoccaccioTruthEx.html" target="_blank">Initial testing of true age functions, now running but output wrong - May 24, 2019</a>
+<a href="https://sgaichas.github.io/poseidon-dev/FullBoccaccioTruthEx.html" target="_blank">Initial testing of true age functions, now running but output still wrong - May 24, 2019, updated May 27, 2019</a>
 
@@
 
-May 24: Either `calc_Z` or `calc_stage2age` or both are incorrect if we expect them to exactly match atlantis annual age ouput where it is available. We match in timestep 0, then diverge.
+May 24: Either `calc_Z` or `calc_stage2age` or both are incorrect if we expect them to exactly match atlantis annual age ouput where it is available.
+We match in timestep 0, then diverge.
+May 27 update: also true for Bocaccio in CCA, the interim numbers at age drop to 0.
 

### 2019-05-31 SardinesOnlyatlantisom2SStest.Rmd
initial with new CCA run

---
title: "Sardines only end-to-end atlantisom test"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)
```

## Introduction

Now we test the `atlantisom` to SS workflow with a new California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_run11".

We will go from atlantis model output to SS input in this page for a single species, "sardine", that doesn't require any age class splitting.

As of 31 May 2019, the `run_truth` function has been modified to do the catch numbers fix for legacy models.

```{r initialize}

initCCA <- TRUE
initNEUS <- FALSE
initNOBA <- FALSE

species_ss <- c("Pacific_sardine")

#function to make a config file? need one for each atlantis run

if(initCCA) source(here("config/CC2Config.R"))

if(initNEUS) source(here("config/NEUSConfig.R"))

if(initNOBA) source(here("config/NOBAConfig.R"))

```

```{r get_names, message=FALSE, warning=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```

This can be run with `species_ss` in select_groups to get results only for that species. Here I'm running with all species; took about 50 minutes to return.

```{r get_truth_ss}

# default run_truth setup will save the file, so check for that first

if(!file.exists(file.path(d.name, 
                          paste0("output", scenario.name, "run_truth.RData")))){
  #Store all loaded results into an R object
  truth <- run_truth(scenario = scenario.name,
                     dir = d.name,
                     file_fgs = functional.groups.file,
                     file_bgm = box.file,
                     select_groups = funct.group.names,
                     file_init = initial.conditions.file,
                     file_biolprm = biol.prm.file,
                     file_runprm = run.prm.file,
                     verbose = TRUE
  )
} else {
  truth <- get(load(file.path(d.name,
                              paste0("output", scenario.name, "run_truth.RData"))))
}
```

In this block, `test` was a vector of 0s, meaning the correction run separately and the new correction implemented in `run_truth` are producing the same results:

```{r catchnums-check, eval=FALSE}

# this was run with results in CalCurrent2013_OA_off folder and run_truth_sardines

catchnumfull <- readRDS(file.path(d.name, paste0(scenario.name, "catchnums_corrected.rds")))

catchnum_ss <- catchnumfull %>%
  filter(species == species_ss) %>%
  arrange(time, polygon, agecl)

new_catage <- truth$catch %>%
  arrange(time, polygon, agecl)

test <- catchnum_ss$atoutput - new_catage$atoutput

# test is all 0 so this worked the same way

```

In the big picture, we will have the true population dynamics from Atlantis which we need for reference to compare with estimated population dynamics from the stock assessment estimation models.

Ideally we will have a function that gets the whole truth (and nothing but the truth) from Atlantis and stores it in a sensible way for later comparison and calculation of performance metrics.

Then we will have a function that implements the "survey" that produces the stock assessment inputs and sends those to the SS file writing functions.

## Truth from OM

The following settings should achieve a survey that samples all Atlantis model output timesteps, all species, and all model polygons, with perfect efficiency and full selectivity for all ages:

```{r census-spec, message=FALSE, warning=FALSE}

# make a function for this
source(here("config/census_spec.R"))

```

Get true total biomass, total numbers. This code gets results for all species and saves the output true B and N.

```{r trueB-N}

# this uses result$biomass_ages to sample biomass directly, no need for wt@age est

survey_testBall <- create_survey(dat = truth$biomass_ages,
                                 time = timeall,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# make up a constant 0 cv for testing
surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))

# call sample_survey_biomass with a bunch of 1s for weight at age
# in the code it multiplies atoutput by wtatage so this allows us to use
# biomass directly
wtage <- data.frame(species=rep(survspp, each=10),
                    agecl=rep(c(1:10),length(survspp)),
                    wtAtAge=rep(1.0,length(survspp)*10))

surveyB_frombio <- sample_survey_biomass(survey_testBall, surv_cv, wtage)

#save for later use, takes a long time to generate
saveRDS(surveyB_frombio, file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))

# this uses result$nums and a new function to get survey index in numbers (abundance)

survey_testNall <- create_survey(dat = truth$nums,
                                 time = timeall,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# as above, make up a constant 0 cv for testing
surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))

surveyN <- sample_survey_numbers(survey_testNall, surv_cv)

#save for later use, takes a long time to generate
saveRDS(surveyN, file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))

```

Get true population age composition. Now we go to single species to speed things up.

```{r true-B-N}

```

Get true population length comp and weight at age

```{r true-comps}

```

Get true population parameters (M in progress, S-R, growth)

```{r true-poppars}

```

Get true catch (may not be different from input for our project)

```{r true-catchtot}

```

Get true catch at age and other comps (may not be different from input for our project)

```{r true-catage-lf-wtage}

```

## "Data" from OM for EMs

Specify our survey sampling. This could come from other information such as the overlap of actual survey stations with OM polygons, experiments evaluating survey selectivity and efficiency, actual sample-based survey cv, etc.

```{r survey-spec}

```

Get input survey biomass and catch biomass for SS (these have q, selectivity, cv)

```{r toSS-surveyfisheryts}

```

Get composition inputs for SS (survey and fishery catch at age, survey and fishery lengths, survey and fishey weight at age)

```{r toSS-surveyfisherycomps}

```

Get population parameters for SS

```{r toSS-poppars}

```

### 2019-05-31 config/CC2Config.R

@@ -1,0 +1,10 @@
+d.name <- here("atlantisoutput","CalCurrent2063_run11")
+functional.groups.file <- "CalCurrentV3Groups.csv"
+biomass.pools.file <- "DIVCalCurrentV3_Biol.nc"
+biol.prm.file <- "CalCurrentV3_Biol.prm"
+box.file <- "CalCurrentV3_utm.bgm"
+initial.conditions.file <- "DIVCalCurrentV3_Biol.nc"
+run.prm.file <- "CalCurrentV3_run.xml"
+scenario.name <- "CCV3"
+bioind.file <- "outputCCV3BiomIndx.txt"
+catch.file <- "outputCCV3Catch.txt"

### 2019-06-03 SardinesOnlyatlantisom2SStest.Rmd
add ss survey
add sampling and plots

@@
 
-```{r trueB-N}
+```{r trueB-N, eval=FALSE}
 
@@
 
-Get true population age composition. Now we go to single species to speed things up.
+Plot true biomass and numbers for "sardines."
 
-```{r true-B-N}
+```{r plot-true-B-N}
+
+trueB <- readRDS(file.path(d.name, paste0(scenario.name,"surveyBcensus.rds")))
+
+trueB_ss <- trueB %>%
+  filter(species==species_ss)
+
+plotB <-ggplot() +
+  geom_line(data=trueB_ss, aes(x=time/stepperyr,y=atoutput, color="survey census N"), 
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotB +
+  facet_wrap(~species, scales="free")
+
+trueN <- readRDS(file.path(d.name, paste0(scenario.name,"surveyNcensus.rds")))
+
+trueN_ss <- trueN %>%
+  filter(species==species_ss)
+
+plotN <-ggplot() +
+  geom_line(data=trueN_ss, aes(x=time/stepperyr,y=atoutput, color="survey census N"), 
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotN +
+  facet_wrap(~species, scales="free")
 
@@
 
-Get true population length comp and weight at age
+Get true population age comp, length comp, and weight at age. Now we go to single species to speed things up.
 
-```{r true-comps}
+```{r true-comps, warning=FALSE, message=FALSE}
+
+# wrap these individual calls into a function
+survey_N_ss <- create_survey(dat = truth$nums,
+                                 time = timeall,
+                                 species = species_ss,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
+# apply default sample fish to get numbers, effNhigh samples all, see census_spec
+numssshigh <- sample_fish(survey_N_ss, effNhigh)
+
+# save age comps
+saveRDS(numssshigh, file.path(d.name, paste0(scenario.name, "natage_census_sard.rds")))
+
+# aggregate true resn per survey design
+aggresnss <- aggregateDensityData(dat = truth$resn,
+                                 time = timeall,
+                                 species = species_ss,
+                                 boxes = boxall)
+
+# aggregate true structn per survey design
+aggstructnss <- aggregateDensityData(dat = truth$structn,
+                                 time = timeall,
+                                 species = species_ss,
+                                 boxes = boxall)
+
+#dont sample these, just aggregate them using median
+structnss <- sample_fish(aggstructnss, effNhigh, sample = FALSE)
+
+resnss <-  sample_fish(aggresnss, effNhigh, sample = FALSE)
+
+# this function does the length sampling and also weight at age
+length_census_ss <- calc_age2length(structn = structnss,
+                                 resn = resnss,
+                                 nums = numssshigh,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+#save for later use, takes a long time to generate
+saveRDS(length_census_ss, file.path(d.name, paste0(scenario.name, "lengthwt_census_sard.rds")))
+
+```
+
+Plot some of the age, length comps, and wtatage
+
+```{r plot-true-agecomps}
+
+Natage <- readRDS(file.path(d.name, paste0(scenario.name, "natage_census_sard.rds")))
+
+Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        Natage$species))
+
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
+```{r lf-plot}
+
+lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "lengthwt_census_sard.rds")))
+
+len <- lengthwt_census_ss$natlength
+
+lfplot <- ggplot(len, aes(upper.bins)) +
+  geom_bar(aes(weight = atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        len$species))
+
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
+lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+
+```
+
+```{r wtage-plot}
+
+wageplot <- ggplot(lengthwt_census_ss$muweight, aes(agecl, atoutput)) +
+  geom_point(aes(colour = time)) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  scale_x_discrete(limits=c(1:10)) +
+  xlab("age class") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste(scenario.name))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
 

### 2019-06-03 SardinesOnlyatlantisom2SStest.Rmd
add true catch

@@
 
-In this block, `test` was a vector of 0s, meaning the correction run separately and the new correction implemented in `run_truth` are producing the same results:
+<!--In this block, `test` was a vector of 0s, meaning the correction run separately and the new correction implemented in `run_truth` are producing the same results:
 
@@
 ```
+-->
 
@@
 
-```{r true-comps, warning=FALSE, message=FALSE}
+```{r true-comps, warning=FALSE, message=FALSE, eval=FALSE}
 
@@
 
-wageplot <- ggplot(lengthwt_census_ss$muweight, aes(agecl, atoutput)) +
-  geom_point(aes(colour = time)) +
+wtage_ann <- lengthwt_census_ss$muweight %>%
+  filter(time %in% annualmidyear)
+
+# reverse to show agecl time series of wt
+wageplot <- ggplot(wtage_ann, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
   theme_tufte() +
   theme(legend.position = "bottom") +
-  scale_x_discrete(limits=c(1:10)) +
-  xlab("age class") +
+  xlab("model timestep (5 per year)") +
   ylab("average individual weight (g)") +
-  labs(subtitle = paste(scenario.name))
+  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
 
@@
 
+truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)
+
+truecatchbio_ss <- truecatchbio[truecatchbio$species == species_ss,]
+
+plotcatch <- ggplot() +
+  geom_line(data=truecatchbio_ss, aes(x=time,y=atoutput, color="true catch bio"),
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotcatch +
+  facet_wrap(~species, scales="free")
+
 ```
@@
 
+# get survey nums with full (no) selectivity
+catchN_ss <- create_fishery_subset(dat = truth$catch,
+                                 time = timeall,
+                                 species = species_ss,
+                                 boxes = boxall)
+
+# apply default sample fish as before to get numbers
+catch_numssshigh <- sample_fish(catchN_ss, effNhigh)
+
+# aggregate true resn per survey or fishery subset design
+catch_aggresnss <- aggregateDensityData(dat = truth$resn,
+                                 time = timeall,
+                                 species = species_ss,
+                                 boxes = boxall)
+
+# aggregate true structn per survey or fishery subsetdesign
+catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
+                                 time = timeall,
+                                 species = species_ss,
+                                 boxes = boxall)
+
+#dont sample these, just aggregate them using median
+catch_structnss <- sample_fish(catch_aggstructnss, effNhigh, sample = FALSE)
+
+catch_resnss <-  sample_fish(catch_aggresnss, effNhigh, sample = FALSE)
+
+# get catch lengths and wtage
+catch_length_census_ss <- calc_age2length(structn = catch_structnss,
+                                 resn = catch_resnss,
+                                 nums = catch_numssshigh,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
+                                 maxbin = 150,
+                                 CVlenage = 0.1, remove.zeroes=TRUE)
+
+#save for later use, takes a long time to generate
+saveRDS(catch_length_census_ss, file.path(d.name, paste0(scenario.name, "catchlengthwt_census_sard.rds")))
+
 ```
+
+This is true catch at age for sardine:
+
+```{r catage1}
+
+catchage <- catch_numssshigh %>%
+  filter(species == "Pacific_sardine")
+
+catageplot <- ggplot(catchage, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchage$species))
+
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
 
@@
 ```
+
+## “Data” from OM for EMs
+
+Specify our survey sampling.
+This could come from other information such as the overlap of actual survey stations with OM polygons, experiments evaluating survey selectivity and efficiency, actual sample-based survey cv, etc.
+Here we guess using Christine's specifications:
 

### 2019-06-03 SardinesOnlyatlantisom2SStest.Rmd
start adding survey-spec

@@
 
+```{r sardine-survey-spec}
+# from atlantisom/config/sardine_config.R, to be generalized here
+
+#Set species name
+species <- "Pacific_sardine" #change to species_ss to match
+
+#Directory with SS files
+model_dir <- "Sardine_SS_files/"
+
+#Name of SS data file
+datfile_name <- "sardEM_3_3.dat"
+
+#Name of atlantis model output - this is an RData object from the runtruth() function
+atlantis_output_name <- "outputCCV3run_truth.RData"
+
+#Efficiency
+eff_case <- 0.5
+
+#Age classes of your species
+age_classes <- 1:10
+
+#Selectivity function - a vector equivalent to length (age classes)
+sel_by_age <- rep(1,length(age_classes))
+
+#Atlantis model timestep corresponding to the true output--now from census_spec.R
+timestep <- stepperyr #5
+
+#Which atlantis timestep does the survey run in?--now from census_spec.R
+survey_sample_time <- midptyr #3
+
+#The last timestep to sample
+total_sample <- noutsteps #495
+
+#Vector of indices of survey times to pull
+survey_sample_full <- seq(survey_sample_time,
+                          total_sample, by=timestep)
+
+#Effective sample size for composition data
+surveyEffN <- 1000
+fisheryEffN <- 1000
+
+#CVs for length at age, catch, and survey
+CVs <- list("lenage"=0.1, "fishery"=0.01, "survey"=0.1)
+
+#Number of years of data to pull
+nyears <- 50
+
+#Atlantis initialization period
+burnin <- 30
+
+#Maximum size of a fish in cm
+maxbin <- 150
+
+#Years to fish, assuming fishing output is annual
+fish_years <- burnin:(burnin+nyears-1)
+
+#Years to survey
+survey_years <- survey_sample_full[burnin:(burnin+nyears-1)]
+
+#Month of survey/fishing
+survey_month <- 7
+fishing_month <- 1
+
+```
+
 Get input survey biomass and catch biomass for SS (these have q, selectivity, cv)

### 2019-06-03 SardinesOnlyatlantisom2SStest.Rmd
text edits
add survey and SS writing parts

@@
 
+# save true catch at age
+saveRDS(catch_numssshigh, file.path(d.name, paste0(scenario.name, "catchage_census_sard.rds")))
+
 # aggregate true resn per survey or fishery subset design
@@
 
+catch_numssshigh <- readRDS(file.path(d.name, paste0(scenario.name, "catchage_census_sard.rds")))
+
 catchage <- catch_numssshigh %>%
@@
 
-## "Data" from OM for EMs
-
-Specify our survey sampling.
-This could come from other information such as the overlap of actual survey stations with OM polygons, experiments evaluating survey selectivity and efficiency, actual sample-based survey cv, etc.
-
-```{r survey-spec}
-
 ```
@@
 This could come from other information such as the overlap of actual survey stations with OM polygons, experiments evaluating survey selectivity and efficiency, actual sample-based survey cv, etc.
-Here we guess using Christine's specifications:
+
+Here we are using Christine's specifications, which approximate the acoustic survey index used in the sardine assessment:
 
@@
 #Set species name
-species <- "Pacific_sardine" #change to species_ss to match
+species <- species_ss #"Pacific_sardine" #change to species_ss to match
 
@@
 
+#Efficiency of the survey
+effic <- data.frame(species=species, efficiency=eff_case)
+
+# Assume uniform selectivity, same as selex1 here
+sel<- data.frame(species=rep(species,length(age_classes)),
+               agecl=age_classes,
+              selex=sel_by_age)
+
 ```
 
-Get input survey biomass and catch biomass for SS (these have q, selectivity, cv)
+Get input survey biomass for SS (these have q, selectivity, cv)
 
-```{r toSS-surveyfisheryts}
+```{r toSS-surveyts}
+
+#Sample survey - 3rd timestep simulates a summer survey
+survey_out <- create_survey(dat=truth$nums, 
+                            time=survey_sample_full, 
+                            species=species, 
+                            boxes=boxall, 
+                            effic=effic, 
+                            selex=sel)
+
+#Set effective sample size for age compositions
+effN <- surveyEffN
+highEffN <- data.frame(species=species, effN=rep(effN, length(species)))
+
+#Sample fish for age composition
+age_comp_data <- sample_fish(survey_out, highEffN)
+
+# aggregate true resn per survey design
+survey_aggresnstd <- aggregateDensityData(dat = results$resn,
+                                          time = unique(results$nums$time),
+                                          species = species,
+                                          boxes = boxall)
+
+# aggregate true structn per survey design
+survey_aggstructnstd <- aggregateDensityData(dat = results$structn,
+                                             time = unique(results$nums$time),
+                                             species = species,
+                                             boxes = boxall)
+
+ss_structnstd <- sample_fish(survey_aggstructnstd,
+                             effN,
+                             sample=FALSE)
+ss_resnstd <- sample_fish(survey_aggresnstd,
+                          effN,
+                          sample=FALSE)
+
+#Extract length composition data
+ss_length_stdsurv <- calc_age2length(structn = ss_structnstd,
+                                     resn = ss_resnstd,
+                                     nums = age_comp_data,
+                                     biolprm = results$biolprm, fgs = results$fgs,
+                                     CVlenage = CVs$lenage, remove.zeroes=TRUE)
+
+#Need to replace with interp function
+wtAtAge <- ss_length_stdsurv$muweight %>%
+  select(species, agecl, time, wtAtAge = atoutput) %>%
+  mutate(wtAtAge = wtAtAge/1000)
+
+# CV function
+cv <- data.frame(species="Pacific_sardine", cv=CVs$survey)
+
+#Sample survey biomass
+survObsBiom <- sample_survey_biomass(dat=survey_out,cv=cv,wtAtAge)
+
+# survey numbers, not sure which SS needs?
+survObsNum <- sample_survey_numbers(dat=survey_out,cv=cv)
+
+```
+
+Get input total catch for SS (here using numbers). SS takes a fishery cv which can be applied here to get a catch time series, or we can input true catch.
+
+```{r toSS-fisheryts}
+
+# catch at age by area
+catch_numbers <-  create_fishery_subset(dat = truth$catch,
+                                         time = fish_years,
+                                         species = species,
+                                         boxes = boxall)
+
+# catch in numbers
+catch_total <- catch_numbers %>%
+  group_by(time) %>%
+  summarise(catch=sum(atoutput))
+
+# the above can be done with sample_survey_numbers and the fishery cv can be applied
+# CV function
+cv <- data.frame(species="Pacific_sardine", cv=CVs$fishery)
+
+catchObsNum <- sample_survey_numbers(catch_numbers, cv=cv)
 
@@
 
+#Set effective sample size for age compositions
+effN <- surveyEffN
+highEffN <- data.frame(species=species, effN=rep(effN, length(species)))
+
+#Sample fish for age composition
+age_comp_data <- sample_fish(survey_out, highEffN)
+
+# Survey length comps and wtage done above to get survey ts using nums*wtage approach
+
+#We end up using CAAL for the survey below, so let's generate fishery age comps instead
+effN <- data.frame(species=species, effN=effN)
+
+catch_numssshigh <- sample_fish(catch_numbers, effN)
+
+#Get weights
+# aggregate true resn per survey or fishery subset design
+catch_aggresnss <- aggregateDensityData(dat = truth$resn,
+                                 time = fish_years,
+                                 species = species,
+                                 boxes = boxes)
+
+# aggregate true structn per survey or fishery subsetdesign
+catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
+                                 time = fish_years,
+                                 species = species,
+                                 boxes = boxes)
+
+#dont sample these, just aggregate them using median
+catch_structnss <- sample_fish(catch_aggstructnss, effN, sample = FALSE)
+
+catch_resnss <-  sample_fish(catch_aggresnss, effN, sample = FALSE)
+
+catch_length <- calc_age2length(structn = catch_structnss,
+                                 resn = catch_resnss,
+                                 nums = catch_numssshigh,
+                                 biolprm = results$biolprm, fgs = results$fgs,
+                                 maxbin = maxbin,
+                                 CVlenage = CVs$lenage, remove.zeroes=TRUE)
+
 ```
 
-Get population parameters for SS
+Get population parameters for SS from Atlantis
 
@@
 ```
+
+## Write data for Stock Synthesis
+
+To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`.
+You will need to ensure the relevant Stock Synthesis files are stored in the inst/extdata/ folder, with each species example in its own folder.
+
+```{r readdata, echo=FALSE}
+require(r4ss)
+#source("./config/sardine_config.R") this is above for illustration
+
+stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/",
+model_dir,
+datfile_name))
+```
+
+Write time series biomass (tons) and catch (numbers) data for SS using `SS_write_ts`:
+
+```{r SS-write-ts}
+#Test writing CPUE in biomass
+  stocksynthesis.data <- SS_write_ts(ss_data_list = stocksynthesis.data,
+                ts_data = list(survObsBiom$atoutput[fish_years],
+                  catch_total$catch[fish_years]),
+                CVs = c(CVs$survey,
+                        CVs$fishery),
+                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             catch_total$time[fish_years]),
+            sampling_month = list(rep(survey_month,nyears),
+                                rep(fishing_month,nyears)),
+                units = c("biomass","numbers"),
+                data_type=c("CPUE","catch"),
+                fleets = c(2,1))
+```
+
+Writing age composition data
+
+```{r SS-write-comps}
+#age_comp_data
+
+#Get the age bins
+age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
+age_bins <- sub("a","",age_bin_names)
+
+age_comp_flat <- reformat_compositions(age_comp_data,
+                     round.places = 4,
+                     comp_type = "agecomp")
+
+## Write age composition data for survey
+stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data,
+               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
+               data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr+nyears-1)),
+               sampling_month = list(rep(survey_month,nyears)),
+               data_type = c("agecomp"),
+               fleet_number = c(1),
+               bins = list(age_bins),
+               caal_bool = c(FALSE))
+stocksynthesis.data$agecomp
+
+#Check length compositions match age compositions
+ss_length_stdsurv$natlength %>%
+  group_by(time,agecl) %>%
+  summarise(sum(atoutput))
+
+len_comp_data <- ss_length_stdsurv$natlength
+fish_len_comp_data <- catch_length$natlength
+
+```
+
+Once we have the natlength matrix, we can munge the data into the proper CAAL and length bin format for SS
+
+```{r SS-comps-formatting}
+
+caal_comp_flat <- reformat_compositions(len_comp_data,                                round.places=4,
+                comp_type="caalcomp")
+
+#remove burnin
+caal_comp_final <- filter(caal_comp_flat,
+                         time %in% survey_years)
+
+#Add over age classes to get sample size
+len_comp_flat <- reformat_compositions(len_comp_data,
+                                  round.places = 0,
+                           comp_type="lencomp")
+#remove burnin
+len_comp_final <- filter(len_comp_flat,
+                         time %in% survey_years)
+
+length_bins <- as.integer(names(len_comp_final))
+length_bins <- length_bins[!is.na(length_bins)]
+
+fish_len_comp_flat <- reformat_compositions(catch_length$natlength,
+                                  round.places = 0,
+                           comp_type="lencomp")
+#remove burnin
+fish_len_comp_final <- filter(fish_len_comp_flat,
+                         time %in% fish_years)
+
+fish_age_comp_flat <- reformat_compositions(catch_numssshigh,
+                           comp_type="agecomp")
+#remove burnin
+fish_age_comp_final <- filter(fish_age_comp_flat,
+                         time %in% fish_years)
+
+comp_list <- list(caal_comp_final,len_comp_final,          fish_age_comp_final, fish_len_comp_final)
+
+apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), rep(survey_month, nrow(comp_list[[2]])), rep(fishing_month,nrow(comp_list[[3]])),rep(fishing_month,nrow(comp_list[[4]])))
+
+# Write CAAL and length composition data
+stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data, 
+                                      comp_matrix = comp_list, 
+                                      data_rows = list((comp_list[[1]]$time-survey_sample_time)/timestep + 1 , (survey_years-survey_sample_time)/timestep + 1,fish_years,fish_years),
+                                      sampling_month = apply_month, 
+                                      data_type = rep(c("agecomp", "lencomp"),2), 
+                                      fleet_number = c(2,2,1,1),  
+                                      bins = list(age_bins, 
+                                                  length_bins, 
+                                                  age_bins, 
+                                                  length_bins), 
+                                      caal_bool = c(TRUE, rep(FALSE,3)))
+
+head(stocksynthesis.data$lencomp)
+head(stocksynthesis.data$agecomp)
+
+#Change length bin structure to match atlantis data
+stocksynthesis.data$lbin_vector <- length_bins
+
+#Set lbin_method to 1 - this makes the population length bins match the data bins 
+#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
+stocksynthesis.data$lbin_method <- 1
+stocksynthesis.data$binwidth <- "#"
+stocksynthesis.data$minimum_size <- "#"
+stocksynthesis.data$maximum_size <- "#"
+
+#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
+stocksynthesis.data$age_info$minsamplesize <- rep(0.001,2)
+
+SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
+datfile_name),
+                 overwrite=TRUE)
+```
+
+## Run Stock Synthesis
+After the data and control file modifications, ensure you copy the local SS 3.3 executable into the folder containing the model files. Then run the following chunk.
+
+```{r runSS, eval=FALSE}
+#Run stock synthesis model
+run_stocksynthesis(species = species, model_dir = model_dir, show_output = TRUE)
+
+```

### 2019-06-03 SardinesOnlyatlantisom2SStest.Rmd
comments

@@
 #Name of atlantis model output - this is an RData object from the runtruth() function
-atlantis_output_name <- "outputCCV3run_truth.RData"
+#atlantis_output_name <- "outputCCV3run_truth.RData"
 
@@
 
-```{r readdata, echo=FALSE}
+```{r readdata}
 require(r4ss)
@@
 #Check length compositions match age compositions
-ss_length_stdsurv$natlength %>%
-  group_by(time,agecl) %>%
-  summarise(sum(atoutput))
+#ss_length_stdsurv$natlength %>%
+#  group_by(time,agecl) %>%
+#  summarise(sum(atoutput))
 

### 2019-06-03 SardinesOnlyatlantisom2SStest.Rmd
flagging errors, likely fishery timesteps

@@
 
-```{r true-catage-lf-wtage}
+```{r true-catage-lf-wtage, eval=FALSE}
 
@@
 
+######    CHECK fish_years   ####### get in right units  #######
+# fishery output timestep
+fstepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutfinc
+
 #Years to fish, assuming fishing output is annual
@@
 # aggregate true resn per survey design
-survey_aggresnstd <- aggregateDensityData(dat = results$resn,
-                                          time = unique(results$nums$time),
+survey_aggresnstd <- aggregateDensityData(dat = truth$resn,
+                                          time = survey_sample_full,
                                           species = species,
@@
 # aggregate true structn per survey design
-survey_aggstructnstd <- aggregateDensityData(dat = results$structn,
-                                             time = unique(results$nums$time),
+survey_aggstructnstd <- aggregateDensityData(dat =truth$structn,
+                                             time = survey_sample_full,
                                              species = species,
@@
                                      nums = age_comp_data,
-                                     biolprm = results$biolprm, fgs = results$fgs,
+                                     biolprm = truth$biolprm, fgs = truth$fgs,
                                      CVlenage = CVs$lenage, remove.zeroes=TRUE)
@@
 catch_numbers <-  create_fishery_subset(dat = truth$catch,
-                                         time = fish_years,
+                                         #time = fish_years,
+                                         time = survey_sample_full,
                                          species = species,
@@
 
-catchObsNum <- sample_survey_numbers(catch_numbers, cv=cv)
+catchObsNum <- sample_survey_numbers(dat=catch_numbers, cv=cv)
 
@@
 catch_aggresnss <- aggregateDensityData(dat = truth$resn,
-                                 time = fish_years,
+                                 #time = fish_years,
+                                 time = survey_sample_full,
                                  species = species,
-                                 boxes = boxes)
+                                 boxes = boxall)
 
@@
 catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
-                                 time = fish_years,
+                                 #time = fish_years,
+                                 time = survey_sample_full,
                                  species = species,
-                                 boxes = boxes)
+                                 boxes = boxall)
 
@@
                                  nums = catch_numssshigh,
-                                 biolprm = results$biolprm, fgs = results$fgs,
+                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                  maxbin = maxbin,
@@
 ```{r readdata}
+#devtools::install_github("r4ss/r4ss", dependencies = FALSE)
 require(r4ss)
@@
 
-```{r SS-comps-formatting}
+```{r SS-comps-formatting, eval=FALSE}
+
+#getting an error in here, I think it is the definition of fish_years
+#Error in x[[jj]][iseq] <- vjj : replacement has length zero
 

### 2019-06-04 SardinesOnlyatlantisom2SStest.Rmd
fixing fish timing

@@
 plotB <-ggplot() +
-  geom_line(data=trueB_ss, aes(x=time/stepperyr,y=atoutput, color="survey census N"), 
+  geom_line(data=trueB_ss, aes(x=time/stepperyr,y=atoutput, color="survey census B"), 
             alpha = 10/10) +
@@
 
-Get true catch (may not be different from input for our project)
+Get true catch in tons (may not be different from input for our project)
 
@@
 
+# note: time output of this file is in days
+# what model timestep is this output matching?
 plotcatch <- ggplot() +
-  geom_line(data=truecatchbio_ss, aes(x=time,y=atoutput, color="true catch bio"),
+  geom_line(data=truecatchbio_ss, aes(x=time/365,y=atoutput, color="true catch bio"),
             alpha = 10/10) +
@@
 
-Get true catch at age and other comps (may not be different from input for our project)
+Get true catch at age and other comps (may not be different from input for our project). First see if catch output is at same timescale as population by checking toutfinc: `r fstepperyr`
 
@@
 
-#Atlantis initialization period
+#Atlantis initialization period in years
 burnin <- 30
@@
 
-######    CHECK fish_years   ####### get in right units  #######
-# fishery output timestep
-fstepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutfinc
+#Years to fish, assuming fishing output is annual
+#fish_years <- burnin:(burnin+nyears-1)
+# fishery output timestep (fstepperyr in census_spec.R) for this run is 5 per year
 
-#Years to fish, assuming fishing output is annual
-fish_years <- burnin:(burnin+nyears-1)
+# here we are summing the catch in numbers for input in to SS.
+# To use the total catch biomass output from catch.txt the above would work 
+# because the output is annual but in units of DAYS so needs time/365
 
-#Years to survey
+#The last timestep to sample
+total_sample <- noutsteps #495
+
+#Vector of indices of catch in numbers to pull
+fish_sample_full <- c(0:total_sample)
+fish_burnin <- burnin*fstepperyr
+fish_nyears <- nyears*fstepperyr
+fish_times <- fish_sample_full[fish_burnin:(fish_burnin+fish_nyears-1)]
+fish_years <- seq(fish_times[1], max(fish_times), by=fstepperyr) #last timestep
+
+# potential problem: fish_years dont match survey_years timesteps below
+
+#Years to survey, assuming survey is once per year
 survey_years <- survey_sample_full[burnin:(burnin+nyears-1)]
@@
 survObsBiom <- sample_survey_biomass(dat=survey_out,cv=cv,wtAtAge)
+# check against survey with truth$biomass_ages output
 
@@
 catch_numbers <-  create_fishery_subset(dat = truth$catch,
-                                         #time = fish_years,
-                                         time = survey_sample_full,
+                                         time = fish_times,
                                          species = species,
@@
 
-# catch in numbers
+if (fstepperyr>1){
+  #need to take all output time steps and sum catch over each year
+  #catch_numbers <- for yr in 0:nyears
+  #sum catch_numbers$atoutput for time 0:fstepperyr
+  #call that atoutput at time=yr+fstepperyr
+  #should match fish_years
+  #return the new dataset with same shape but time in endyr increments and sum nums
+}
+
+# catch in numbers summed over polygon after time has been aggregated into 1 yr
 catch_total <- catch_numbers %>%
@@
 catch_aggresnss <- aggregateDensityData(dat = truth$resn,
-                                 #time = fish_years,
-                                 time = survey_sample_full,
+                                 time = fish_times,
                                  species = species,
@@
 catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
-                                 #time = fish_years,
-                                 time = survey_sample_full,
+                                 time = fish_times,
                                  species = species,

### 2019-06-04 config/census_spec.R
change midptyr to 2 for 5 steps per year

@@ -15,7 +15,7 @@ boxall <- c(0:(boxpars$nbox - 1))
 runpar <- load_runprm(d.name, run.prm.file)
 noutsteps <- runpar$tstop/runpar$outputstep
 stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
-midptyr <- round(median(seq(1,stepperyr)))
+midptyr <- round(median(seq(0,stepperyr)))
 
 # a survey that takes place once per year mid year
 annualmidyear <- seq(midptyr, noutsteps, stepperyr)

### 2019-06-04 SardinesOnlyatlantisom2SStest.Rmd
comparison plots, fishery indexing contd

@@
 #Which atlantis timestep does the survey run in?--now from census_spec.R
-survey_sample_time <- midptyr #3
+# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
+# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
+
+survey_sample_time <- midptyr #3 this is more late summer-fall, changed to 2
 
@@
 
-#Years to fish, assuming fishing output is annual
-#fish_years <- burnin:(burnin+nyears-1)
+#Years to fish, assuming fishing output is annual and years index 1,2,3 etc
+fish_years <- burnin:(burnin+nyears-1)
 # fishery output timestep (fstepperyr in census_spec.R) for this run is 5 per year
@@
 #The last timestep to sample
-total_sample <- noutsteps #495
+total_sample <- noutsteps-1 #495
 
-#Vector of indices of catch in numbers to pull
+#Vector of indices of catch in numbers to pull (by timestep to sum)
 fish_sample_full <- c(0:total_sample)
-fish_burnin <- burnin*fstepperyr
+fish_burnin <- burnin*fstepperyr+1
 fish_nyears <- nyears*fstepperyr
 fish_times <- fish_sample_full[fish_burnin:(fish_burnin+fish_nyears-1)]
-fish_years <- seq(fish_times[1], max(fish_times), by=fstepperyr) #last timestep
+#fish_timesteps <- seq(fish_times[1], max(fish_times), by=fstepperyr) #last timestep
+#fish_years <- unique(floor(fish_times/fstepperyr))
 
-# potential problem: fish_years dont match survey_years timesteps below
+# potential problem: while they are from the same model year,
+# fish_years dont match survey_years timesteps (survey mid year, fishery end)
+# so when matching use the index, not the value itself
 
@@
 survey_out <- create_survey(dat=truth$nums, 
+                            time=survey_sample_full, 
+                            species=species, 
+                            boxes=boxall, 
+                            effic=effic, 
+                            selex=sel)
+
+#Try a biomass based survey for comparison
+survey_outB <- create_survey(dat=truth$biomass_ages, 
                             time=survey_sample_full, 
@@
 survObsBiom <- sample_survey_biomass(dat=survey_out,cv=cv,wtAtAge)
+
 # check against survey with truth$biomass_ages output
+wtage1 <- data.frame(species=rep(species_ss, each=10),
+                    agecl=rep(c(1:10),length(species_ss)),
+                    wtAtAge=rep(1.0,length(species_ss)*10))
+
+survObsBiomB <- sample_survey_biomass(dat=survey_outB,cv=cv,wtage1)
 
@@
 
+Do some comparisons of biomass surveys with truth. Something is going way wrong in the calculation of survey biomass based on numbers times weight at age, looks like division by 1000 could fix it?.
+
+```{r plotBseries}
+
+plotB <-ggplot() +
+  geom_line(data=trueB_ss, aes(x=time/stepperyr,y=atoutput, color="survey census B"), 
+            alpha = 10/10) +
+  geom_point(data=survObsBiomB, aes(x=time/stepperyr,y=atoutput, color="survey sample B from biomass_ages"), 
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotB +
+  facet_wrap(~species, scales="free")
+
+plotB <-ggplot() +
+  geom_line(data=trueB_ss, aes(x=time/stepperyr,y=atoutput, color="survey census B"), 
+            alpha = 10/10) +
+  geom_point(data=survObsBiom, aes(x=time/stepperyr,y=atoutput/1000, color="survey sample B from nums * wtAtAge"), 
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotB +
+  facet_wrap(~species, scales="free") 
+```
+
 Get input total catch for SS (here using numbers). SS takes a fishery cv which can be applied here to get a catch time series, or we can input true catch.
@@
 
-# catch at age by area
+# catch at age by area and timestep
 catch_numbers <-  create_fishery_subset(dat = truth$catch,
@@
   #need to take all output time steps and sum catch over each year
-  #catch_numbers <- for yr in 0:nyears
-  #sum catch_numbers$atoutput for time 0:fstepperyr
-  #call that atoutput at time=yr+fstepperyr
-  #should match fish_years
-  #return the new dataset with same shape but time in endyr increments and sum nums
+  catch_numbers_area <- catch_numbers %>%
+    #assign a temporary year column, aggregate nums over this year
+    mutate(yr = floor(time/fstepperyr)) %>%
+    group_by(species, agecl, polygon, yr) %>%
+    summarise(anncatch_area=sum(atoutput))
+
+  # catch in numbers summed over polygon 
+  # and agecl after time has been aggregated into 1 yr
+  catch_total <- catch_numbers_area %>%
+    group_by(yr) %>%
+    summarise(catch=sum(anncatch_area))
 }
 
-# catch in numbers summed over polygon after time has been aggregated into 1 yr
+# catch in numbers summed over polygon and agecl if output is annual
+if (fstepperyr==1){
   catch_total <- catch_numbers %>%
@@
     summarise(catch=sum(atoutput))
+}
 
-# the above can be done with sample_survey_numbers and the fishery cv can be applied
-# CV function
-cv <- data.frame(species="Pacific_sardine", cv=CVs$fishery)
+# if we want to apply the CV to this we need to write sample_total_catch
 
-catchObsNum <- sample_survey_numbers(dat=catch_numbers, cv=cv)
+#cv <- data.frame(species="Pacific_sardine", cv=CVs$fishery)
+
+#catchObsNum <- sample_total_catch(dat=catch_numbers, cv=cv)
 
@@
 
+#catch at age each timestep summed over polygons
 catch_numssshigh <- sample_fish(catch_numbers, effN)
@@
 #Get weights
-# aggregate true resn per survey or fishery subset design
+# aggregate true resn per fishery subset design
 catch_aggresnss <- aggregateDensityData(dat = truth$resn,
@@
 
-# aggregate true structn per survey or fishery subsetdesign
+# aggregate true structn fishery subsetdesign
 catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
@@
 
+# these fishery lengths and weight at age are each output timestep
 catch_length <- calc_age2length(structn = catch_structnss,
@@
   stocksynthesis.data <- SS_write_ts(ss_data_list = stocksynthesis.data,
-                ts_data = list(survObsBiom$atoutput[fish_years],
+                ts_data = list(survObsBiomB$atoutput[fish_years],
                   catch_total$catch[fish_years]),
@@
 
-```{r SS-comps-formatting, eval=FALSE}
+```{r SS-comps-formatting}
 

### 2019-06-04 docs/index.Rmd
add initial sardines with new CCA
update full example
fix what I broke

@@
 May 24: Sardines have a full dataset, but still working on converting to true ages for species with multiple ages per age class. See below under [True age compositions].
+
+<a href="https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html" target="_blank">Start on Sardines in the CalCurrent, all true and SS data, May 31, 2019, updated June 3</a>
 
@@
 
+### Rabbit holes that turned out ok
+
+<a href="https://sgaichas.github.io/poseidon-dev/CheckBiomassAges.html" target="_blank">In which I broke the very first thing that worked, then fixed it - June 4, 2019</a>
+
 ### More to come later!

### 2019-06-04 CheckBiomassAges.Rmd

---
title: "Is Biomass Ages Broken?"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)

```

```{r constants}

initCCA <- TRUE
initNEUS <- FALSE
initNOBA <- FALSE

species_ss <- c("Pacific_sardine")

# make up a constant 0 cv for testing
surv_cv <- data.frame(species=species_ss, cv=rep(0.0,length(species_ss)))

# call sample_survey_biomass with a bunch of 1000s for weight at age
# in the code it multiplies atoutput by wtatage/1000 so this allows us to use
# biomass directly
wtage <- data.frame(species=rep(species_ss, each=10),
                    agecl=rep(c(1:10),length(species_ss)),
                    wtAtAge=rep(1000.0,length(species_ss)*10))

```

## This used to work

Before I cleverly changed sample_survey_biomass to have the right units to use weight at age. Now if we want biomass_ages to give a correct survey, the fake wtage we give it needs to be 1000 instead of 1. So much for a shortcut.

Test biomass_ages use in create survey in old CCA run vs new CCA run

```{r truth-oldCCA}

source(here("config/CC1Config.R"))

#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

source(here("config/census_spec.R"))

sardold <- run_truth(scenario = scenario.name,
                     dir = d.name,
                     file_fgs = functional.groups.file,
                     file_bgm = box.file,
                     select_groups = species_ss,
                     file_init = initial.conditions.file,
                     file_biolprm = biol.prm.file,
                     file_runprm = run.prm.file,
                     verbose = TRUE
)

survey_testBold <- create_survey(dat = sardold$biomass_ages,
                                 time = timeall,
                                 species = species_ss,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

surveyB_frombio_old <- sample_survey_biomass(survey_testBold, surv_cv, wtage)

# read Atlantis output files
atBtxt2old <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)

# lookup the matching names, put in time, species, biomass column format
# WARNING hardcoded for output with last species group as DIN
groupslookup <- funct.groups %>%
  filter(IsTurnedOn > 0)

atBtxt2tidyold <- atBtxt2old %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% species_ss)

```

Compare with BiomIndex.txt

```{r compareold}

surveyB_lastweek_old <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))

surveyB_lastweek <- surveyB_lastweek_old %>%
  filter(species==species_ss)

compareB <-ggplot() +
  geom_line(data=surveyB_frombio_old, aes(x=time/5,y=atoutput, color="survey census B"), 
            alpha = 10/10) +
  geom_line(data=surveyB_lastweek, aes(x=time/5,y=atoutput, color="survey census B last week"), 
            alpha = 1/10) +
  geom_point(data=atBtxt2tidyold, aes(x=Time/365,y=biomass, color="txt output true B"),
           alpha = 10/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compareB + 
  facet_wrap(~species, scales="free")

```

## Why doesnt it now? It does now.

```{r truth-newCCA}

source(here("config/CC2Config.R"))

#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

source(here("config/census_spec.R"))

sardnew <- run_truth(scenario = scenario.name,
                     dir = d.name,
                     file_fgs = functional.groups.file,
                     file_bgm = box.file,
                     select_groups = species_ss,
                     file_init = initial.conditions.file,
                     file_biolprm = biol.prm.file,
                     file_runprm = run.prm.file,
                     verbose = TRUE
)

survey_testBnew <- create_survey(dat = sardnew$biomass_ages,
                                 time = timeall,
                                 species = species_ss,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

surveyB_frombio_new <- sample_survey_biomass(survey_testBnew, surv_cv, wtage)

# read Atlantis output files
atBtxt2new <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)

# lookup the matching names, put in time, species, biomass column format
# WARNING hardcoded for output with last species group as DIN
groupslookup <- funct.groups %>%
  filter(IsTurnedOn > 0)

atBtxt2tidynew <- atBtxt2new %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% species_ss)

```

Compare with BiomIndex.txt

```{r comparenew}

compareB <-ggplot() +
  geom_line(data=surveyB_frombio_new, aes(x=time/5,y=atoutput, color="survey census B"), 
            alpha = 10/10) +
  geom_point(data=atBtxt2tidynew, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 10/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compareB + 
  facet_wrap(~species, scales="free")

```

### 2019-06-04 SardinesOnlyatlantisom2SStest.Rmd

@@
 
-```{r trueB-N, eval=FALSE}
+```{r trueB-N}
 
 # this uses result$biomass_ages to sample biomass directly, no need for wt@age est
-
+# this used to work and is now wrong!
 survey_testBall <- create_survey(dat = truth$biomass_ages,
@@
 
-# call sample_survey_biomass with a bunch of 1s for weight at age
-# in the code it multiplies atoutput by wtatage so this allows us to use
+# call sample_survey_biomass with a bunch of 1000s for weight at age
+# in the code it multiplies atoutput by wtatage/1000 so this allows us to use
 # biomass directly
@@
                     agecl=rep(c(1:10),length(survspp)),
-                    wtAtAge=rep(1.0,length(survspp)*10))
+                    wtAtAge=rep(1000.0,length(survspp)*10))
 
@@
 
+# read Atlantis output files
+atBtxt2 <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)
+groupslookup <- funct.groups %>%
+  filter(IsTurnedOn > 0)
+
+atBtxt2tidy <- atBtxt2 %>%
+  select(Time:DIN) %>%
+  #select(Time, FPL:DIN) %>%
+  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
+  gather(species, biomass, -Time) %>%
+  filter(species %in% c("Pacific_sardine"))
+
 plotB <-ggplot() +
@@
             alpha = 10/10) +
+  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
+             alpha = 1/10) + 
   theme_tufte() +
@@
                     agecl=rep(c(1:10),length(species_ss)),
-                    wtAtAge=rep(1.0,length(species_ss)*10))
+                    wtAtAge=rep(1000.0,length(species_ss)*10))
 
@@
 
-Do some comparisons of biomass surveys with truth. Something is going way wrong in the calculation of survey biomass based on numbers times weight at age, looks like division by 1000 could fix it?.
+Do some comparisons of biomass surveys with truth.
+Something is going way wrong in the calculation of survey biomass based on numbers times weight at age, looks like division by 1000 could fix it?
+Actually no the true biomass based on survey is wrong.
 
@@
 plotB <-ggplot() +
+    geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
+             alpha = 1/10) + 
   geom_line(data=trueB_ss, aes(x=time/stepperyr,y=atoutput, color="survey census B"), 
@@
 plotB <-ggplot() +
+    geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
+             alpha = 10/10) + 
     geom_line(data=trueB_ss, aes(x=time/stepperyr,y=atoutput, color="survey census B"), 
             alpha = 10/10) +
-  geom_point(data=survObsBiom, aes(x=time/stepperyr,y=atoutput/1000, color="survey sample B from nums * wtAtAge"), 
+  geom_point(data=survObsBiom, aes(x=time/stepperyr,y=atoutput, color="survey sample B from nums * wtAtAge"), 
             alpha = 10/10) +
@@
                 fleets = c(2,1))
+
+stocksynthesis.data$CPUE
+
+stocksynthesis.data$catch
 ```
@@
 
-```{r SS-comps-formatting}
+```{r SS-comps-formatting, eval=FALSE}
 

### 2019-06-05 SardinesOnlyatlantisom2SStest.Rmd
fixing stuff

@@
 
-```{r trueB-N}
+```{r trueB-N, eval=FALSE}
 
 # this uses result$biomass_ages to sample biomass directly, no need for wt@age est
-# this used to work and is now wrong!
 survey_testBall <- create_survey(dat = truth$biomass_ages,
@@
   geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
-             alpha = 1/10) + 
+             alpha = 10/10) + 
   theme_tufte() +
@@
 
+# try summing this to get true catch in nums by timestep with surv_cv = 0
+totcatchN <- sample_survey_numbers(catchN_ss, surv_cv)
+
+# save true total catch in numbers by timestep, need to sum to year
+saveRDS(totcatchN, file.path(d.name, paste0(scenario.name, "catchN_census_sard.rds")))
+
 # apply default sample fish as before to get numbers
-catch_numssshigh <- sample_fish(catchN_ss, effNhigh)
+# no, dont apply sample, effN isnt high enough and can't be made high enough
+#catch_numssshigh <- sample_fish(catchN_ss, effNhigh)
+
+# just aggregte true numbers over layer and polygon
+catch_numsss <- aggregate(catchN_ss$atoutput,list(catchN_ss$species,catchN_ss$agecl,catchN_ss$time),sum)
+names(catch_numsss) <- c("species","agecl","time","atoutput")
+catch_numssshigh <- data.frame(species = catch_numsss$species,
+					  agecl = catch_numsss$agecl,
+					  polygon = NA,
+					  layer = NA,
+					  time = catch_numsss$time,
+					  atoutput = catch_numsss$atoutput)
 
@@
 
+totcatchN <- readRDS(file.path(d.name, paste0(scenario.name, "catchN_census_sard.rds")))
+
 catch_numssshigh <- readRDS(file.path(d.name, paste0(scenario.name, "catchage_census_sard.rds")))
@@
   filter(species == "Pacific_sardine")
+
+totcatchann <- totcatchN %>%
+  mutate(yr = floor(time/fstepperyr)) %>%
+  group_by(species, yr) %>%
+  summarise(anncatchage = sum(atoutput))
+# 
+# totcatch <- totcatchage %>%   
+#   group_by(species, yr) %>%
+#   summarise(truecatch = sum(anncatchage))
+# 
+plottotC <- ggplot(totcatchann, aes(x=yr, y=anncatchage)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchage$species))
+
+plottotC
 
@@
 fish_times <- fish_sample_full[fish_burnin:(fish_burnin+fish_nyears-1)]
-#fish_timesteps <- seq(fish_times[1], max(fish_times), by=fstepperyr) #last timestep
+fish_timesteps <- seq(fish_times[5], max(fish_times), by=fstepperyr) #last timestep
 #fish_years <- unique(floor(fish_times/fstepperyr))
@@
     geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
-             alpha = 1/10) + 
+             alpha = 10/10) + 
   geom_line(data=trueB_ss, aes(x=time/stepperyr,y=atoutput, color="survey census B"), 
@@
 ```{r SS-write-ts}
+
+# Error in `*tmp*`[[jj]] : subscript out of bounds
+
 #Test writing CPUE in biomass

### 2019-06-05 CreateStockSynthesis.Rmd

@@
                 ts_data = list(survObsBiom$atoutput[fish_years],
-                  catch_total$catch[fish_years]),
+                  catch_total$catch[catch_total$time==fish_years]),
                 CVs = c(CVs$survey,
                         CVs$fishery),
-                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             catch_total$time[fish_years]),
+                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             fish_years),
             sampling_month = list(rep(survey_month,nyears),
@@
 
-SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
+SS_writedat(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
 datfile_name),

### 2019-06-05 SardinesOnlyatlantisom2SStest.Rmd
fix census, scale catch bio

@@
 # apply default sample fish to get numbers, effNhigh samples all, see census_spec
-numssshigh <- sample_fish(survey_N_ss, effNhigh)
+# NOPE! effNhigh is not big enough for census of sardines, bigger breaks function
+#numssshigh <- sample_fish(survey_N_ss, effNhigh)
+
+numsss <- aggregate(survey_N_ss$atoutput,list(survey_N_ss$species,survey_N_ss$agecl,survey_N_ss$time),sum)
+names(numsss) <- c("species","agecl","time","atoutput")
+numssshigh <- data.frame(species = numsss$species,
+					  agecl = numsss$agecl,
+					  polygon = NA,
+					  layer = NA,
+					  time = numsss$time,
+					  atoutput = numsss$atoutput)
 
@@
 
-# get catch lengths and wtage
+# get catch lengths and wtage at every timestep
 catch_length_census_ss <- calc_age2length(structn = catch_structnss,
@@
 
-totcatchN <- readRDS(file.path(d.name, paste0(scenario.name, "catchN_census_sard.rds")))
-
 catch_numssshigh <- readRDS(file.path(d.name, paste0(scenario.name, "catchage_census_sard.rds")))
@@
   filter(species == "Pacific_sardine")
-
-totcatchann <- totcatchN %>%
-  mutate(yr = floor(time/fstepperyr)) %>%
-  group_by(species, yr) %>%
-  summarise(anncatchage = sum(atoutput))
-# 
-# totcatch <- totcatchage %>%   
-#   group_by(species, yr) %>%
-#   summarise(truecatch = sum(anncatchage))
-# 
-plottotC <- ggplot(totcatchann, aes(x=yr, y=anncatchage)) +
-  geom_point() +
-  theme_tufte() +
-  labs(subtitle = paste(scenario.name,
-                        catchage$species))
-
-plottotC
 
@@
 catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
+Do true catch numbers times weight at age match catch in biomass?
+No.
+It is about 16 times too high relative to catch.txt output, and shifted one year back (which is due to the floor function, and an easy fix).
+
+```{r catchNtoB}
+
+totcatchN <- readRDS(file.path(d.name, paste0(scenario.name, "catchN_census_sard.rds")))
+
+totcatchann <- totcatchN %>%
+  mutate(yr = floor(time/fstepperyr)+1) %>%
+  group_by(species, yr) %>%
+  summarise(anncatchage = sum(atoutput))
+
+plottotC <- ggplot(totcatchann, aes(x=yr, y=anncatchage)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste("Catch in N ", scenario.name,
+                        totcatchann$species))
+
+plottotC
+
+catch_numssshigh <- readRDS(file.path(d.name, paste0(scenario.name, "catchage_census_sard.rds")))
+
+catch_lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "catchlengthwt_census_sard.rds")))
+
+CwtAtAge <- catch_lengthwt_census_ss$muweight %>%
+  select(species, agecl, time, wtAtAge = atoutput) %>%
+  mutate(wtAtAge = wtAtAge/1000)
+
+catchbioN_ss <- merge(catch_numssshigh,CwtAtAge,by=c("species","agecl", "time"),all.x=T) %>%
+  mutate(biomass = atoutput*wtAtAge/1000) %>%
+  mutate(yr = floor(time/fstepperyr)+1) %>%
+  group_by(species, yr) %>%
+  summarise(anncatchB = sum(biomass))
+
+plotcatch <- ggplot() +
+  geom_line(data=truecatchbio_ss, aes(x=time/365,y=atoutput, color="true catch bio from txt"),
+            alpha = 10/10) +
+  geom_point(data=catchbioN_ss, aes(x=yr,y=anncatchB/16, color="true catch bio from N/16"),
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotcatch +
+  facet_wrap(~species, scales="free")
 
@@
 Do some comparisons of biomass surveys with truth.
-Something is going way wrong in the calculation of survey biomass based on numbers times weight at age, looks like division by 1000 could fix it?
-Actually no the true biomass based on survey is wrong.
 
@@
     #assign a temporary year column, aggregate nums over this year
-    mutate(yr = floor(time/fstepperyr)) %>%
+    mutate(yr = floor(time/fstepperyr)+1) %>%
     group_by(species, agecl, polygon, yr) %>%
@@
 
+Check this against true total catch numbers above. Should be the same.
+
+```{r compare-totcatch}
+
+compareC <- ggplot() +
+  geom_line(data=totcatchann, aes(x=yr, y=anncatchage, color="census catch N"), 
+            alpha = 10/10) +
+  geom_point(data=catch_total, aes(x=yr,y=catch, color="sample catch N"), 
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+compareC +
+  facet_wrap(~species, scales="free")
+
+```
+
 Get composition inputs for SS (survey and fishery catch at age, survey and fishery lengths, survey and fishey weight at age)
@@
 
-#Sample fish for age composition
+#Sample fish for age composition from survey
 age_comp_data <- sample_fish(survey_out, highEffN)
@@
 
+Can we match catch in weight from catch.txt? No, see above.
+
+```{r compareC-wt, eval=FALSE}
+
+plotcatch <- ggplot() +
+  geom_line(data=truecatchbio_ss, aes(x=time/365,y=atoutput, color="true catch bio"),
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotcatch +
+  facet_wrap(~species, scales="free")
+
+```
+
 Get population parameters for SS from Atlantis
@@
 
-```{r SS-write-ts}
+```{r SS-write-ts, eval=FALSE}
 

### 2019-06-05 docs/index.Rmd
fix census, scale catch bio

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html" target="_blank">Start on Sardines in the CalCurrent, all true and SS data, May 31, 2019, updated June 3</a>
+<a href="https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html" target="_blank">Start on Sardines in the CalCurrent, all true and SS data, May 31, 2019, updated June 5</a>
 

### 2019-06-05 SardinesOnlyatlantisom2SStest.Rmd
aggregating catch comps, still failing to write

@@
 #The last timestep to sample
-total_sample <- noutsteps #495
+total_sample <- noutsteps-1 #495
 
@@
 
-#Effective sample size for composition data
+#Effective sample size for composition data (annual total samples)
 surveyEffN <- 1000
@@
 #Years to fish, assuming fishing output is annual and years index 1,2,3 etc
-fish_years <- burnin:(burnin+nyears-1)
+#fish_years <- burnin:(burnin+nyears-1)
 # fishery output timestep (fstepperyr in census_spec.R) for this run is 5 per year
@@
 fish_timesteps <- seq(fish_times[5], max(fish_times), by=fstepperyr) #last timestep
-#fish_years <- unique(floor(fish_times/fstepperyr))
+fish_years <- unique(floor(fish_times/fstepperyr)+1)
 
@@
 
-```{r toSS-surveyts}
+```{r toSS-surveyts-comps}
 
@@
 
-```{r toSS-surveyfisherycomps}
-
-#Set effective sample size for age compositions
-effN <- surveyEffN
-highEffN <- data.frame(species=species, effN=rep(effN, length(species)))
-
-#Sample fish for age composition from survey
-age_comp_data <- sample_fish(survey_out, highEffN)
+```{r toSS-fisherycomps}
 
@@
 #We end up using CAAL for the survey below, so let's generate fishery age comps instead
+effN <- fisheryEffN/fstepperyr
 effN <- data.frame(species=species, effN=effN)
@@
 #catch at age each timestep summed over polygons
-catch_numssshigh <- sample_fish(catch_numbers, effN)
+catch_numsss_samp <- sample_fish(catch_numbers, effN)
 
@@
 #dont sample these, just aggregate them using median
-catch_structnss <- sample_fish(catch_aggstructnss, effN, sample = FALSE)
-
-catch_resnss <-  sample_fish(catch_aggresnss, effN, sample = FALSE)
+catch_structnss_samp <- sample_fish(catch_aggstructnss, effN, sample = FALSE)
+
+catch_resnss_samp <-  sample_fish(catch_aggresnss, effN, sample = FALSE)
 
 # these fishery lengths and weight at age are each output timestep
-catch_length <- calc_age2length(structn = catch_structnss,
-                                 resn = catch_resnss,
-                                 nums = catch_numssshigh,
+catch_lengthwt_samp <- calc_age2length(structn = catch_structnss_samp,
+                                 resn = catch_resnss_samp,
+                                 nums = catch_numsss_samp,
                                  biolprm = truth$biolprm, fgs = truth$fgs,
@@
 
-Can we match catch in weight from catch.txt? No, see above.
+<!--Can we match catch in weight from catch.txt? No, numbers sampled are 1e-6 compared with actual catch numbers.
 
@@
 
+CswtAtAge <- catch_lengthwt_samp$muweight %>%
+  select(species, agecl, time, wtAtAge = atoutput) %>%
+  mutate(wtAtAge = wtAtAge/1000)
+
+catchsampbio_ss <- merge(catch_numsss_samp,CswtAtAge,by=c("species","agecl", "time"),all.x=T) %>%
+  mutate(biomass = atoutput*wtAtAge/1000) %>%
+  mutate(yr = floor(time/fstepperyr)+1) %>%
+  group_by(species, yr) %>%
+  summarise(anncatchBsamp = sum(biomass))
+
 plotcatch <- ggplot() +
-  geom_line(data=truecatchbio_ss, aes(x=time/365,y=atoutput, color="true catch bio"),
+  geom_line(data=truecatchbio_ss, aes(x=time/365,y=atoutput/1000000, color="true catch bio"),
+            alpha = 10/10) +
+  geom_point(data=catchsampbio_ss, aes(x=yr,y=anncatchBsamp, color="sample catch B"), 
             alpha = 10/10) +
@@
 
-```{r SS-write-ts, eval=FALSE}
-
-# Error in `*tmp*`[[jj]] : subscript out of bounds
+```{r SS-write-ts}
 
@@
   stocksynthesis.data <- SS_write_ts(ss_data_list = stocksynthesis.data,
-                ts_data = list(survObsBiomB$atoutput[fish_years],
-                  catch_total$catch[fish_years]),
+                ts_data = list(survObsBiom$atoutput[fish_years],
+                  catch_total$catch[catch_total$yr %in% fish_years]),
                 CVs = c(CVs$survey,
                         CVs$fishery),
-                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             catch_total$time[fish_years]),
+                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             fish_years),
             sampling_month = list(rep(survey_month,nyears),
@@
 
+```
+
+Once we have the natlength matrix, we can munge the data into the proper CAAL and length bin format for SS
+
+```{r SS-comps-formatting, eval=FALSE}
+
 #Check length compositions match age compositions
@@
 len_comp_data <- ss_length_stdsurv$natlength
-fish_len_comp_data <- catch_length$natlength
-
-```
-
-Once we have the natlength matrix, we can munge the data into the proper CAAL and length bin format for SS
-
-```{r SS-comps-formatting, eval=FALSE}
-
-#getting an error in here, I think it is the definition of fish_years
-#Error in x[[jj]][iseq] <- vjj : replacement has length zero
+fish_len_comp_data <- catch_lengthwt_samp$natlength
+
+if(fstepperyr>1){
+  fish_len_comp_anndata <- fish_len_comp_data %>%
+    mutate(yr = floor(time/fstepperyr)) %>%
+    group_by(species, agecl, lower.bins, upper.bins, time=as.integer(yr)) %>%
+    summarise(annnatlength=sum(atoutput)) %>%
+    rename(atoutput = annnatlength)
+} else {
+  fish_len_comp_anndata <- fish_len_comp_data
+}
 
@@
 
-fish_len_comp_flat <- reformat_compositions(catch_length$natlength,
+# fishery length comps are still 5 timesteps per year
+# need to aggregate to annual (done above)
+# also,  make effN annual goal/fstepsperyr (done above)
+fish_len_comp_flat <- reformat_compositions(fish_len_comp_anndata,
                                   round.places = 0,
                            comp_type="lencomp")
-#remove burnin
+
+#remove burnin works after adjustment above
 fish_len_comp_final <- filter(fish_len_comp_flat,
@@
 
-fish_age_comp_flat <- reformat_compositions(catch_numssshigh,
+# fishery age comps also 5 timesteps per year
+if(fstepperyr>1){
+  fish_age_comp_anndata <- catch_numsss_samp %>%
+    mutate(yr = floor(time/fstepperyr)) %>%
+    group_by(species, agecl, time=as.integer(yr)) %>%
+    summarise(annnatage=sum(atoutput)) %>%
+    rename(atoutput = annnatage)
+} else {
+  fish_age_comp_anndata <- catch_numsss_samp
+}
+
+fish_age_comp_flat <- reformat_compositions(fish_age_comp_anndata,
                            comp_type="agecomp")
-#remove burnin
+
+#remove burnin (not necessary?fish comps made with fish_years only) 
 fish_age_comp_final <- filter(fish_age_comp_flat,
@@
 
-apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), rep(survey_month, nrow(comp_list[[2]])), rep(fishing_month,nrow(comp_list[[3]])),rep(fishing_month,nrow(comp_list[[4]])))
+apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), 
+                    rep(survey_month, nrow(comp_list[[2]])),
+                    rep(fishing_month,nrow(comp_list[[3]])),
+                    rep(fishing_month,nrow(comp_list[[4]])))
+
+# below fails for fishery data if I keep timestep aligned with year 31
+# Error in `[.data.frame`(comp_matrix[[i]], , as.character(bins[[i]])) : undefined columns selected
+
+# assigning fishery comps to year 30 start (remove +1 after floors above) 
+# I get a different error
+# Error in `[<-.data.frame`(`*tmp*`, indices, "Seas", value = c(1, 1, 1, : replacement has 49 rows, data has 2045
 

### 2019-06-06 SardinesOnlyatlantisom2SStest.Rmd
progress? catch closer to correct

@@
 
-```{r true-catage-lf-wtage, eval=FALSE}
+```{r true-catage-lf-wtage}
 
@@
 Do true catch numbers times weight at age match catch in biomass?
-No.
-It is about 16 times too high relative to catch.txt output, and shifted one year back (which is due to the floor function, and an easy fix).
+It was about 16 times too high relative to catch.txt output, and shifted one year back (which is due to the floor function, and an easy fix).
+
+June 6: just corrected the catch nums correction in `run_truth` and I hope it matches now.
 
@@
             alpha = 10/10) +
-  geom_point(data=catchbioN_ss, aes(x=yr,y=anncatchB/16, color="true catch bio from N/16"),
+  geom_point(data=catchbioN_ss, aes(x=yr,y=anncatchB, color="true catch bio from N"),
             alpha = 10/10) +

### 2019-06-06 SardinesOnlyatlantisom2SStest.Rmd
still broken

@@
   facet_wrap(~species, scales="free")
+
+```
+
+Sadly, still a problem with the catch correction if it should match catch.txt when comverted to biomass.
+We are now off by about a factor of 2 (catch nums * wtAtAge is half true catch bio from the txt file):
+
+```{r catchoff}
+sardcatchB <- truecatchbio_ss %>%
+  mutate(yr = as.integer(time/365))
+
+sardcatchBcomp <- merge(sardcatchB, catchbioN_ss)
+sardcatchBcomp <- mutate(sardcatchBcomp, mult=atoutput/anncatchB)
+
+mean(sardcatchBcomp$mult)
+
+summary(sardcatchBcomp$mult)
 

### 2019-06-06 SardinesOnlyatlantisom2SStest.Rmd
align survey and fishery start, note errors

@@
 #Years to survey, assuming survey is once per year
-survey_years <- survey_sample_full[burnin:(burnin+nyears-1)]
+survey_years <- survey_sample_full[(burnin+1):(burnin+nyears)]
 
@@
 
-# below fails for fishery data if I keep timestep aligned with year 31
+# below fails with two errors (only gets to first when running)
+
+# Error in `[<-.data.frame`(`*tmp*`, indices, "Seas", value = c(1, 1, 1, : replacement has 49 rows, data has 2045
+
 # Error in `[.data.frame`(comp_matrix[[i]], , as.character(bins[[i]])) : undefined columns selected
 
-# assigning fishery comps to year 30 start (remove +1 after floors above) 
-# I get a different error
-# Error in `[<-.data.frame`(`*tmp*`, indices, "Seas", value = c(1, 1, 1, : replacement has 49 rows, data has 2045
+# both these errors come from the fishery age and length comps when I step
+# through the function, survey caal and lengths read in fine
 

### 2019-06-06 docs/index.Rmd
align survey and fishery start, note errors

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html" target="_blank">Start on Sardines in the CalCurrent, all true and SS data, May 31, 2019, updated June 5</a>
+<a href="https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html" target="_blank">Sardines in the CalCurrent run with F and rec dev scenario, all true and SS data, May 31, 2019, updated June 6</a>
 

### 2019-06-07 SardinesOnlyatlantisom2SStest.Rmd
use catch bio as input to SS

@@
 
-<!--In this block, `test` was a vector of 0s, meaning the correction run separately and the new correction implemented in `run_truth` are producing the same results:
-
-```{r catchnums-check, eval=FALSE}
-
-# this was run with results in CalCurrent2013_OA_off folder and run_truth_sardines
-
-catchnumfull <- readRDS(file.path(d.name, paste0(scenario.name, "catchnums_corrected.rds")))
-
-catchnum_ss <- catchnumfull %>%
-  filter(species == species_ss) %>%
-  arrange(time, polygon, agecl)
-
-new_catage <- truth$catch %>%
-  arrange(time, polygon, agecl)
-
-test <- catchnum_ss$atoutput - new_catage$atoutput
-
-# test is all 0 so this worked the same way
-
-```
--->
-
 In the big picture, we will have the true population dynamics from Atlantis which we need for reference to compare with estimated population dynamics from the stock assessment estimation models.
@@
 # what model timestep is this output matching?
+# rescale catch in biomass time from days to years
+truecatchbio_ss$time <- as.integer(truecatchbio_ss$time/365)
+
 plotcatch <- ggplot() +
-  geom_line(data=truecatchbio_ss, aes(x=time/365,y=atoutput, color="true catch bio"),
+  geom_line(data=truecatchbio_ss, aes(x=time,y=atoutput, color="true catch bio"),
             alpha = 10/10) +
@@
 plotcatch <- ggplot() +
-  geom_line(data=truecatchbio_ss, aes(x=time/365,y=atoutput, color="true catch bio from txt"),
+  geom_line(data=truecatchbio_ss, aes(x=time,y=atoutput, color="true catch bio from txt"),
             alpha = 10/10) +
@@
 ```{r catchoff}
-sardcatchB <- truecatchbio_ss %>%
-  mutate(yr = as.integer(time/365))
+
+ sardcatchB <- truecatchbio_ss #%>%
+# now done above
+#   mutate(yr = as.integer(time/365))
 
@@
 
+PLEASE DON'T USE THIS FOR CCA.
+We have determined above that catch in numbers from legacy atlantis code, even corrected, is still off by a factor of 2.
+We leave the info here because this code can be used for atlantis codebases after December 2015.
+
 ```{r toSS-fisheryts}
@@
 
-Check this against true total catch numbers above. Should be the same.
+Check this against true total catch numbers above. Should be the same (and it is, but both are incorrect in the absolute scale due to the CATCH.nc output problem in older codebases).
 
@@
 
-Get composition inputs for SS (survey and fishery catch at age, survey and fishery lengths, survey and fishey weight at age)
+Get composition inputs for SS (survey and fishery catch at age, survey and fishery lengths, survey and fishey weight at age).
+
+Because catch composition information goes in to the assessment as a proportion, we can use fishery catch at age from this legacy codebase even with absolute catch numbers likely half what they should be overall.
 
@@
 
-Write time series biomass (tons) and catch (numbers) data for SS using `SS_write_ts`:
+*Now changing catch units to tons!*
+
+Write time series biomass (tons) and catch (tons) data for SS using `SS_write_ts`:
 
@@
                 ts_data = list(survObsBiom$atoutput[fish_years],
-                  catch_total$catch[catch_total$yr %in% fish_years]),
+                  truecatchbio_ss$atoutput[truecatchbio_ss$time %in% fish_years]),
                 CVs = c(CVs$survey,
@@
                                 rep(fishing_month,nyears)),
-                units = c("biomass","numbers"),
+                units = c("biomass","biomass"),
                 data_type=c("CPUE","catch"),

### 2019-06-07 docs/index.Rmd
use catch bio as input to SS

--- 2019-06-05 docs/index.Rmd
@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html" target="_blank">Start on Sardines in the CalCurrent, all true and SS data, May 31, 2019, updated June 5</a>
+<a href="https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html" target="_blank">Sardines in the CalCurrent run with F and rec dev scenario, all true and SS data, May 31, 2019, updated June 7</a>
 

### 2019-06-07 TestCalcZ2.Rmd
initial comparisons calc Z with mort.txt

--- 2019-05-24 TestCalcZ.Rmd
@@
 ---
-title: "Getting mortality from atlantisom: testing calc_Z function I"
+title: "Getting mortality from atlantisom: testing calc_Z function II"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 output: html_document
-bibliography: "packages.bib"
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
-# automatically create a bib database for R packages
-knitr::write_bib(c(
-  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
-  'data.table', 'here', 'ggforce', 'ggthemes', 'r4ss'
-), 'packages.bib')
+
 ```
@@
 Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
-On this page we demonstrate use of atlantisom on Norwegian Barents Sea Atlantis (NOBA) output files to test function `calc_Z` calculating total mortality.
+On this page we demonstrate use of atlantisom on both California Current (CCA) and Norwegian Barents Sea Atlantis (NOBA) output files to test function `calc_Z` calculating total mortality.
 This is used both to derive "natural mortality" for comparison with stock assessments and to split age classes into true ages using the function `calc_stage2age`.
@@
 NOBA has output files in true ages that we can compare results with. These files are in the folder NOBAwithAnnAgeOutput (not kept on github due to large filesize).
+
+We can also compare both models with the output[scenario.name]Mort.txt file, which reports annual M and F. The user's manual notes: 
+>This file is currently only useful for looking at relative M vs F values for a species, as itdoes not give accurate mortalities
+
+So we can assume that the scale of Z from this file will not be the same as our calculation, but it should be useful for splitting out F to get M, and for comparing overall patterns.
+
+We will first use CCA here because the contrasting F scenario has been implemented. We will test with both sardines (M needed, but not age splitting) and hake (M and age splitting needed).
 
@@
 
-initCCA <- FALSE
+initCCA <- TRUE
 initNEUS <- FALSE
-initNOBA <- TRUE
+initNOBA <- FALSE
+
+species_ss <- c("Pacific_sardine")
 
@@
 
-if(initCCA) source(here("config/CC1Config.R"))
+if(initCCA) source(here("config/CC2Config.R"))
 
@@
 
+<!-- not sure we need catch bio for this
 ```{r loadcatchb}
@@
 
+truecatchbio_ss <- truecatchbio[truecatchbio$species == species_ss,]
+# note: time output of this file is in days
+# what model timestep is this output matching?
+# rescale catch in biomass time from days to years
+truecatchbio_ss$time <- as.integer(truecatchbio_ss$time/365)
+
 ```
+-->
 
@@
 Ultimately this output goes to `calc_stage2age` to produce annual numbers at true age, which we can compare with the ANNAGEBIO.nc output produced by NOBA.
+For now we will just compare to the Mort.txt output.
 
@@
 
-surveyN_ss <- surveyN[surveyN$species == "North_atl_cod",]
+surveyN_ss <- surveyN[surveyN$species == species_ss,]
 
 plotN <-ggplot() +
-  geom_line(data=surveyN_ss, aes(x=time/5,y=atoutput, color="survey census N"), 
+  geom_line(data=surveyN_ss, aes(x=time/stepperyr,y=atoutput, color="survey census N"), 
             alpha = 10/10) +
@@
 
-### Biological sampling: true numbers at age class
-
-We get true numbers at age at each output step by appling `sample_fish` to the survey_testNall result if the model group has one true age per output age class:
-
-```{r truenatagecl, eval=FALSE}
-# We need an effective N for the sample fish function, setting equal to actual N
-# this one is high but not equal to total for numerous groups
-effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
-
-numsallhigh <- sample_fish(survey_testNall, effNhigh)
-
-#save for later use, takes a long time to generate
-saveRDS(numsallhigh, file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
-
-```
-
-This is NOBA Cod true age class output. However, the 10 standard atlantis age classes aggregate true ages each for Cod.
-
-Number of true age classes is stored in the [...]groups[...].csv file, which is read in using `load_fgs`.
-We read this in as funct.groups above.
-The column NumAgeClassSize stores the number of true ages:
-
-Number of age classes for Cod: `r funct.groups$NumAgeClassSize[funct.groups$Name=="North_atl_cod"] `
-
-```{r Natage-plot}
-
-Natage <- readRDS(file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
-
-Natage_ss <- Natage[Natage$species == "North_atl_cod",]
-
-Natageplot <- ggplot(Natage_ss, aes(x=agecl, y=atoutput)) +
-  geom_point() +
-  theme_tufte() +
-  labs(subtitle = paste(scenario.name,
-                        Natage_ss$species))
-
-Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free")
-Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free")
-Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free")
-Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free")
-
-```
-
 ### What level of aggregation works for calc_Z?
 
-]The atlantisom function `calc_stage2age` calls the `calc_Z` function, applying a total mortality rate to estimate the numbers at true age within an age class, but it is unclear whether to use this on the aggregated numbers or the full resolution numbers at age.
+The atlantisom function `calc_stage2age` calls the `calc_Z` function, applying a total mortality rate to estimate the numbers at true age within an age class, but it is unclear whether to use this on the aggregated numbers or the full resolution numbers at age.
 Here we test outputs of Z for several levels of aggregation:
@@
 
-# cut to a single species in YOY file, hardcoded for NOBA cod
+# get code matching species name to split YOY file
+code_ss <- funct.groups$Code[which(funct.groups$Name == species_ss)]
+
+# cut to a single species in YOY file
 YOY_ss <- YOY %>%
-  select(Time, NCO.0)
+  select(Time, paste0(code_ss, ".0"))
 
 # numbers at agecl at full resolution (all polygons and layers)
-truenums_ss <- truth$nums[truth$nums$species == "North_atl_cod",]
+truenums_ss <- truth$nums[truth$nums$species == species_ss,]
 
@@
 
-# numbers at ageclaggregated over layers, retain polygons (output of create_survey)
-survey_testN_ss <- survey_testNall[survey_testNall$species == "North_atl_cod",]
+# numbers at agecl aggregated over layers, retain polygons (output of create_survey)
+survey_testN_ss <- survey_testNall[survey_testNall$species == species_ss,]
 
-# numbers at agecl aggregated over layers and polygons (output of sample_fish)
-# done above
-# Natage_ss <- Natage[Natage$species == "North_atl_cod",]
+# numbers at agecl aggregated over layers and polygons
+Natage_ss <- readRDS(file.path(d.name, paste0(scenario.name, "natage_census_sard.rds")))
 
@@
 
-Fixed a bug where grep was used if a # appeared, none of those in NOBA, will have to see if output is produced in CCA.
-This ran for the survey resolution numbers.
-I then went back and ran the two other levels above after fixing `calc_Z.`
+Fixed a bug where grep was used if a # appeared. This ran for the survey resolution numbers. I then went back and ran the two other levels above after fixing `calc_Z.`
 
-Something is wrong with running after `sample_fish` because sampleresZ returned all 0s.
 Both true full resolution numbers and survey resolution numbers (aggregated over polygon) returned nonzero Z values.
-They are identical aside from seven values with no apparent pattern (rows 84, 85, 130, 131, 147, 172, 196) out of 201 that differ at e-16.
+As with previous the NOBA test, they are identical aside from eight values with no apparent pattern (rows 100, 105, 180, 200, 220, 240, 355, 400) out of 500 that differ at e-16.
+Same story with difference from sample resolution; this does not matter to output.
+
+```{r differentZ}
+fullresZ$atoutput - surveyresZ$atoutput
+
+fullresZ$atoutput - sampleresZ$atoutput
+```
+
 I would be happer if the differences were all exactly 0, but I suppose we can attribute that to rounding error.
+
+### What is F at each timestep?
+A relative annual F is found in outputMort.txt!
+```{r getrelF}
+
+file.mort <- file.path(d.name, paste0("output", scenario.name, "Mort.txt"))
+
+mortish <- read.table(file.mort, header = TRUE)
+
+relF_ss <- mortish %>%
+  select(Time, relF = paste0(code_ss, ".F"))
+
+relM_ss <- mortish %>%
+  select(Time, relM = paste0(code_ss, ".M"))
+
+plotF <- ggplot(data=relF_ss, aes(x=Time/365, y=relF)) +
+  geom_line() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
+
+plotF
+
+```
+
+This is consistent with the scenario implemented. So far so good.
+
+This probably depends on how the model is set up, but for models forced with F by species we ought to be able to read this out from an input file.
+If it is a daily rate then we mulitply by toutinc to get the rate at each of these output timestep Z values.
+
+### Quick test to see if Z is right?
+Compare with outputMort.txt!
+```{r compZ}
+
+Zish <- merge(relF_ss,relM_ss) %>%
+  mutate(relZ = relF + relM)
+
+plotZ <- ggplot() +
+  geom_line(data=Zish, aes(x=Time/365, y=relZ)) +
+  geom_point(data=fullresZ, aes(x=(time*73)/365, y=atoutput)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
+
+plotZ
+```
+
+Nope. We have a constant value punctuated by occassional variation. Seems wrong. Variation doesn't look anything like the txt file either. Perhaps the 5x yr timesteps are messing this up?
+
+### How to turn timestep Z into annual Z for comparison?
+
+Time to step through line by line again.
 
@@
 
-### What is F at each timestep?
-This probably depends on how the model is set up, but for models forced with F by species we ought to be able to read this out.
-If it is a daily rate then we mulitply by toutinc to get the rate at each of these output timestep Z values.
+### Biological sampling: true numbers at age class
 
-### Quick test to see if Z is right?
+Might be a good time to switch back to NOBA for comparison with true annual ages.
+
+Everything below here is old.
+
+We get true numbers at age at each output step by appling `sample_fish` to the survey_testNall result if the model group has one true age per output age class:
+
+```{r truenatagecl, eval=FALSE}
+# We need an effective N for the sample fish function, setting equal to actual N
+# this one is high but not equal to total for numerous groups
+
+# this doesnt work for sardines in CCA, change to full census
+
+effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
+
+numsallhigh <- sample_fish(survey_testNall, effNhigh)
+
+#save for later use, takes a long time to generate
+saveRDS(numsallhigh, file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
+
+```
+
+This is NOBA Cod true age class output. However, the 10 standard atlantis age classes aggregate true ages each for Cod.
+
+Number of true age classes is stored in the [...]groups[...].csv file, which is read in using `load_fgs`.
+We read this in as funct.groups above.
+The column NumAgeClassSize stores the number of true ages:
+
+Number of age classes for Cod: `r funct.groups$NumAgeClassSize[funct.groups$Name=="North_atl_cod"] `
+
+```{r Natage-plot, eval=FALSE}
+
+Natage <- readRDS(file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
+
+Natage_ss <- Natage[Natage$species == "North_atl_cod",]
+
+Natageplot <- ggplot(Natage_ss, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        Natage_ss$species))
+
+Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free")
+Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free")
+
+```
+
 Perhaps overly optimistic, but if we feed these into `calc_stage2age` and get a result that matches the NOBA output ANNAGEBIO.nc, then we can assume the Zs are correct.
@@
 
-```{r plotannage}
+```{r plotannage, eval=FALSE}
 
@@
 
-```{r compareages}
+```{r compareages, eval=FALSE}
 
@@
 
-```{r plot-esttrueage-atltrueage}
+```{r plot-esttrueage-atltrueage, eval=FALSE}
 

### 2019-06-11 TestCalcZ2.Rmd
some fixes to rec timing pars

@@
   times <- unique(yoy$Time)
-  recstart_temp <- biolprm$recruit_time
+
+  # SKG June 2019
+  # need to match codes below, this mismatches when biolprm not in same order!!
+  #recstart_temp <- biolprm$recruit_time
   #recstart_temp[,2] <- biolprm$time_spawn[-(grep('#',
@@
   # was this grep to get rid of a species with #XXX? breaks when there are none
-  recstart_temp[,2] <- biolprm$time_spawn[,2] + biolprm$recruit_time[,2]
+  #recstart_temp[,2] <- biolprm$time_spawn[,2] + biolprm$recruit_time[,2]
 
   #recstart_temp <- recstart_temp[recstart_temp[,1]%in%turnedon$Code,] #bug added codes with no YOY output
-  recstart_temp <- recstart_temp[recstart_temp[,1]%in%recruits$group,]
+  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%recruits$group,]
+
+  # need spawn_period?? if so read out in load_biolprm and merge it here too
+
+  # rectiming is a dataframe with species code, time_spawn (day of year), 
+  # recruit_time (number of days), and recruit period (number of days). 
+  # we use these to calculate recstart (day of year) and recend (day of year)
+  rectiming <- merge(biolprm$time_spawn, biolprm$recruit_time, by = 1)
+  rectiming <- merge(rectiming, biolprm$recruit_period, by = 1)
+  names(rectiming) <- c("Code", "time_spawn", "recruit_time", "recruit_period")
+  rectiming <- rectiming %>%
+    mutate(recstart = time_spawn + recruit_time) %>% # possibly + spawn_period
+    mutate(recend = recstart + recruit_period)
+
+  recstart_temp <- rectiming[rectiming$Code %in% recruits$group,]
+
+  # do we need this fraction calculation if YOY is annual snapshot?
+  # (check output timing of YOY file with tsumout, available in run.prm)
+  # can we not just allocate the full YOY to the timestep where they come in?
+  # if recstart to recend spans multiple toutinc timesteps then need fraction
+
   recruits$frac_recruit <- 0
   for (irow in 1:nrow(recstart_temp)) {
-    group <- recstart_temp[irow,1]
+    group <- recstart_temp$Code[irow]
     pick <- which(recruits$group == group)
 
-    recstart <- seq(recstart_temp[irow,2],by=365,length.out=nyrs)
+    recstart <- seq(recstart_temp$recstart[irow],by=365,length.out=nyrs)
     recstart <- recstart[recstart<max(recruits$Time[pick])]
-    recend <- recstart + biolprm$recruit_period[irow,2]
+    recend <- recstart + recstart_temp$recend[irow]
     #rec_times <- rbind(rec_times,cbind(group,recstart,recend))
 
+    # this part still returns a constant fraction
+    # and does not alter Time to change from annual to actual recruitment time
+    # still not understanding what this is meant to do
     for (i_rec in 1:length(recstart)) {
@@
 
+# Time in YOY does not match recstart-recend period still after this loop
+
   # match "Time" of the young of the year with the time-step periodicity
@@
 
+# SKG did this above already in the following line?
+  # Get recruits in numbers rather than biomass
+  #recruits$recruits <- recruits$recruitsbio / recruits$sum
   #G.Fay 1/6/16, expand to num of recruits
   # Recruit / mg C converted to wet weight in tonnes / redfield ratio of C:N
-  recruits$recruits <- recruits$recruits/k_wetdry/biolprm$redfieldcn
+  #recruits$recruits <- recruits$recruits/k_wetdry/biolprm$redfieldcn
 

### 2019-06-11 TestCalcZ2.Rmd
calc Z further work

@@
 
-### How to turn timestep Z into annual Z for comparison?
-
-Time to step through line by line again.
+Time to step through line by line again. Revised as below. Better (apportions annual recruitment to timesteps), but not sure this is fully correct?
 
@@
   # Get recruits in numbers rather than biomass
+  # these match rows 2:end of the legacy file, so good
+  # but is this 1000s or numbers? not real numbers, Gavin corrects below
   recruits$recruits <- recruits$recruitsbio / recruits$sum
+  recruits$yr <- as.integer(round(recruits$Time)/365)
 
@@
 
+  #subsets to groups of interest
   recstart_temp <- rectiming[rectiming$Code %in% recruits$group,]
@@
 
-  recruits$frac_recruit <- 0
+  # align model output timesteps (days) with recruitment periods (days)
+  numstime.days <- unique(nums$time)*toutinc
+  # Sum numbers output over all boxes/depth/cohorts
+  totnums <- aggregate(atoutput ~ species + time, data = nums, sum) %>%
+    mutate(time.days = time*toutinc) %>%
+    mutate(yr = floor(time.days/365))
+
+  totnums <- merge(totnums, recruits,
+    by.x = c("yr", "species"), by.y = c("yr", "Name"),
+    all.x = TRUE) %>%
+    arrange(time)
+
+  totnums$frac_recruit <- 0
+
   for (irow in 1:nrow(recstart_temp)) {
     group <- recstart_temp$Code[irow]
-    pick <- which(recruits$group == group)
+    pick <- which(totnums$group == group)
 
     recstart <- seq(recstart_temp$recstart[irow],by=365,length.out=nyrs)
-    recstart <- recstart[recstart<max(recruits$Time[pick])]
+    recstart <- recstart[recstart<max(totnums$Time[pick])]
     recend <- recstart + recstart_temp$recend[irow]
@@
 
-    # this part still returns a constant fraction
-    # and does not alter Time to change from annual to actual recruitment time
-    # still not understanding what this is meant to do
     for (i_rec in 1:length(recstart)) {
-      i_tstart <- which(pick==min(pick[recruits$Time[pick]>=recstart[i_rec]]))
-      i_tstop <- which(pick==min(pick[recruits$Time[pick]>=recend[i_rec]]))
+      i_tstart <- which(pick==min(pick[numstime.days[pick]>=recstart[i_rec]]))
+      i_tstop <- which(pick==min(pick[numstime.days[pick]>=recend[i_rec]]))
       if (i_rec == length(recstart)) i_tstop <- length(pick)
@@
       for (i_t in 1:n_t) {
-        t_temp <- recruits$Time[pick[i_tstart+i_t-1]]
+        t_temp <- numstime.days[pick[i_tstart+i_t-1]]
         num_temp <- t_temp - recstart[i_rec]
@@
           else {
-            num_temp <- recend[i_rec]-(recruits$Time[pick[i_tstart+i_t-2]])
+            num_temp <- recend[i_rec]-(numstime.days[pick[i_tstart+i_t-2]])
           }
@@
         frac_temp <- max(c(0,num_temp /(recend[i_rec]-recstart[i_rec])))
-        recruits$frac_recruit[pick[i_tstart+i_t-1]] <- frac_temp
+        totnums$frac_recruit[pick[i_tstart+i_t-1]] <- frac_temp
       }
@@
 
-# Time in YOY does not match recstart-recend period still after this loop
+# Time in YOY does not match recstart-recend period still after this loop--fixed
 
@@
   # listed in the run.prm or run.xml file
-  recruits$Time <- recruits$Time / toutinc
+  #recruits$Time <- recruits$Time / toutinc
 
@@
   #recruits$recruits <- recruits$recruitsbio / recruits$sum
+
   #G.Fay 1/6/16, expand to num of recruits
   # Recruit / mg C converted to wet weight in tonnes / redfield ratio of C:N
-  #recruits$recruits <- recruits$recruits/k_wetdry/biolprm$redfieldcn
-
-  # Sum over all boxes/depth/cohorts
-  totnums <- aggregate(atoutput ~ species + time, data = nums, sum)
+  totnums$recruits <- totnums$recruits/k_wetdry/biolprm$redfieldcn
 
@@
   # yearly time step matches the time step
-  totnums <- merge(totnums, recruits,
-    by.x = c("time", "species"), by.y = c("Time", "Name"),
-    all.x = TRUE, all.y = FALSE)
+  # totnums <- merge(totnums, recruits,
+  #   by.x = c("time", "species"), by.y = c("Time", "Name"),
+  #   all.x = TRUE, all.y = FALSE)
   totnums$group <- recruits$group[match(totnums$species, recruits$Name)]
@@
 
+   # Make sure time is in order, did above
+  #totnums <- totnums[order(totnums$species, totnums$time), ]
+  totnums$annrecruits <- totnums$recruits
+
+  totnums$recruits <- totnums$annrecruits*totnums$frac_recruit
+
   # Calculate survivors for each species group
   totnums$survivors <- totnums$recruits
-  # Make sure time is in order
-  totnums <- totnums[order(totnums$species, totnums$time), ]
-
-  totnums$recruits <- totnums$recruits*totnums$frac_recruit
 
@@
 ```
+
+### How to turn timestep Z into annual Z for comparison?
 

### 2019-06-11 docs/index.Rmd
initial comparisons calc Z with mort.txt
calc Z further work

@@
 
+June 7: `calc_Z` seems incorrect when compared to outputMort.txt; stepping through code needed
+
+June 11: fixed matching of recruitment to timestep, still seems off
+
+<a href="https://sgaichas.github.io/poseidon-dev/TestCalcZ2.html" target="_blank">Test Z calc and compare with output mort CCA - June 7, 2019, updated June 11</a>
+
 ### Rabbit holes that turned out ok

### 2019-06-12 TestCalcZ2.Rmd
further alignment of outputs

@@
 
+  #Isaac determined that yr 0 (Time 0) in the YOY file is not real, don't use it
+  #below assigns year 1 to times 0:stepperyr, all good output timesteps in truth$nums
+
   # G.Fay 2/21/16
@@
   # align model output timesteps (days) with recruitment periods (days)
-  numstime.days <- unique(nums$time)*toutinc
+  #numstime.days <- unique(nums$time)*toutinc
   # Sum numbers output over all boxes/depth/cohorts
   totnums <- aggregate(atoutput ~ species + time, data = nums, sum) %>%
-    mutate(time.days = time*toutinc) %>%
-    mutate(yr = floor(time.days/365))
+    mutate(time.days = (time+1)*toutinc) %>% #makes time 0 into days 0->73, etc
+    mutate(yr = ceiling(time.days/365))  # yr 1 is 0:stepsperyr to match recruits yr1
+    #mutate(yr = floor(time.days/365)) # includes 0 value, yr 1 is
 
@@
     for (i_rec in 1:length(recstart)) {
-      i_tstart <- which(pick==min(pick[numstime.days[pick]>=recstart[i_rec]]))
-      i_tstop <- which(pick==min(pick[numstime.days[pick]>=recend[i_rec]]))
+      i_tstart <- which(pick==min(pick[totnums$time.days[pick]>=recstart[i_rec]]))
+      i_tstop <- which(pick==min(pick[totnums$time.days[pick]>=recend[i_rec]]))
       if (i_rec == length(recstart)) i_tstop <- length(pick)
@@
       for (i_t in 1:n_t) {
-        t_temp <- numstime.days[pick[i_tstart+i_t-1]]
+        t_temp <- totnums$time.days[pick[i_tstart+i_t-1]]
         num_temp <- t_temp - recstart[i_rec]
@@
           else {
-            num_temp <- recend[i_rec]-(numstime.days[pick[i_tstart+i_t-2]])
+            num_temp <- recend[i_rec]-(totnums$time.days[pick[i_tstart+i_t-2]])
           }
@@
 ### How to turn timestep Z into annual Z for comparison?
+
+Two approaches: sum nums each year, calculate survival between years (minus annual recruitment), turn into annual Z, or multiply timestep Zs to get annual?
 

### 2019-06-13 TestCalcZ2.Rmd
lets call calc_Z good

@@
 
-Nope. We have a constant value punctuated by occassional variation. Seems wrong. Variation doesn't look anything like the txt file either. Perhaps the 5x yr timesteps are messing this up?
+After several iterations of debugging, we are still not matching the pattern, but improved in that it isn't a constant punctuated by occasional high.
+Variation still doesn't look anything like the txt file either.
+The 5x yr timesteps are probably messing this up?
 
-Time to step through line by line again. Revised as below. Better (apportions annual recruitment to timesteps), but not sure this is fully correct?
+I've commented out the detailed debugging block below, but check out the rmd file if you are interested in the sausage making.
+Changes have been implemented in `calc_Z` so that it now apportions annual recruitment to timesteps and recruitment is correctly aligned with truth nums output.
 
+However, it is not clear that subtracting off "recruitment" that has not been subject to mortality is appropriate when the output numbers at each timestep have been subject to mortality?
+I guess that was the input in each timestep as best we know it, but this is at best an approximation.
+<!--
 ```{r calczsnippet, eval=FALSE}
+
+#corrections listed here have been applied in calc_Z; this code not run
 
@@
 ```
+-->
 
-### How to turn timestep Z into annual Z for comparison?
+### How to turn timestep Z into annual Z for proper comparison?
 
-Two approaches: sum nums each year, calculate survival between years (minus annual recruitment), turn into annual Z, or multiply timestep Zs to get annual?
+Two approaches: annual nums each year, calculate survival between years (minus annual recruitment), turn into annual Z, or add timestep Zs to get annual?
 
-### Biological sampling: true numbers at age class
+First an annual estimate based on end year numbers snapshots:
+
+```{r annZ1}
+
+totnums <- aggregate(atoutput ~ species + time, data = truenums_ss, sum) %>%
+    mutate(time.days = (time+1)*runpar$toutinc) %>% #makes time 0 into days 0->73, etc
+    mutate(yr = ceiling(time.days/365))  # yr 1 is 0:stepsperyr to match recruits yr1
+    #mutate(yr = floor(time.days/365)) # includes 0 value, yr 1 is
+
+#only want numbers at end of year, snapshot
+totnumsann <- totnums %>%
+  filter(time.days %in% seq(365, max(time.days), by=365))
+
+# mg carbon converted to wet weight in tonnes
+k_wetdry <- biol$kgw2d / 1000000000
+
+# WARNING only works for CCA because YOY.txt rows 2:end are already in numbers
+# and we are merging out the incorrect and irrelevant YOY row 1 (Time=0)
+# also hardcoded for sardine example
+recnums <- YOY_ss %>%
+  mutate(yr = as.integer(round(YOY_ss$Time)/365)) %>%
+  mutate(recnums = SAR.0/k_wetdry/biol$redfieldcn)
+
+annZ1 <- merge(totnumsann, recnums) %>%
+  mutate(recnums = replace_na(recnums, 0)) %>%
+  mutate(numslessrec = atoutput - recnums) %>%
+  mutate(surv = numslessrec/lag(atoutput, default = first(numslessrec))) %>%
+  mutate(Z = -1 * log(surv))
+
+plotZ <- ggplot() +
+  geom_line(data=Zish, aes(x=Time/365, y=relZ)) +
+  geom_point(data=annZ1, aes(x=Time/365, y=Z)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
+
+plotZ + ylim(0, 2.0)
+
+```
+
+This looks better; we are starting to see a pattern.
+
+Second summing within year Z from calc_Z:
+
+```{r annZ2}
+
+annZ2 <- fullresZ %>%
+  mutate(yr = floor(time/stepperyr)+1) %>%
+  group_by(species, yr) %>%
+  summarise(Z = sum(atoutput))
+
+plotZ <- ggplot() +
+  geom_line(data=Zish, aes(x=Time/365, y=relZ)) +
+  geom_point(data=annZ2, aes(x=yr, y=Z)) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
+
+plotZ + ylim(0, 2.0)
+
+```
+
+Also looks better. Not sure which is more correct.
+
+Compare methods:
+
+```{r compZall}
+
+plotZ <- ggplot() +
+  geom_line(data=Zish, aes(x=Time/365, y=relZ, color="mort.txt Z")) +
+  geom_point(data=annZ1, aes(x=Time/365, y=Z, color="endyr Z")) +
+  geom_point(data=annZ2, aes(x=yr, y=Z, color="sum timestep Z")) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
+
+plotZ + ylim(0, 2.0)
+
+```
+
+Both seem to be tracking each other reasonably, and also general trends from mort.txt output.
+Discussion with Beth suggests that mort.txt output is an initial calculation combining deaths due to predation with deaths due to fishing, but that there is rescaling of predation (and possibly fishing?) afterward this output stage such that mort.txt does not provide final mortality.
+If the rescaling varies by timestep, then matching general patterns rather than the full internannual variability is probably the best we can do, since we don't know the full interannual variability.
+
+We can probably conclude that the `calc_Z` function is working as correctly as possible now, given that it is an approximation no matter what, and the differences we see here are likley from temporal resolution of the Z estimate.
+
+Let's see how it performs with another model.
+
+### Test calc_Z NOBA
 
@@
 
-Everything below here is old.
+```{r switch-NOBA}
 
-We get true numbers at age at each output step by appling `sample_fish` to the survey_testNall result if the model group has one true age per output age class:
+initCCA <- FALSE
+initNEUS <- FALSE
+initNOBA <- TRUE
 
-```{r truenatagecl, eval=FALSE}
-# We need an effective N for the sample fish function, setting equal to actual N
-# this one is high but not equal to total for numerous groups
+if(initNOBA) source(here("config/NOBAaaConfig.R"))
 
-# this doesnt work for sardines in CCA, change to full census
+species_ss <- "North_atl_cod"
 
-effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
+#Load functional groups
+funct.groups <- load_fgs(dir=d.name,
+                         file_fgs = functional.groups.file)
+#Get just the names of active functional groups
+funct.group.names <- funct.groups %>% 
+  filter(IsTurnedOn == 1) %>%
+  select(Name) %>%
+  .$Name
 
-numsallhigh <- sample_fish(survey_testNall, effNhigh)
+#Get true NOBAaa
+if(!file.exists(file.path(d.name, 
+                          paste0("output", scenario.name, "run_truth.RData")))){
+  #Store all loaded results into an R object
+  truth <- run_truth(scenario = scenario.name,
+                     dir = d.name,
+                     file_fgs = functional.groups.file,
+                     file_bgm = box.file,
+                     select_groups = funct.group.names,
+                     file_init = initial.conditions.file,
+                     file_biolprm = biol.prm.file,
+                     file_runprm = run.prm.file
+  )
+} else{
+  truth <- get(load(file.path(d.name,
+                              paste0("output", scenario.name, "run_truth.RData"))))
+}
 
-#save for later use, takes a long time to generate
-saveRDS(numsallhigh, file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
+source(here("config/census_spec.R"))
 
@@
 
-This is NOBA Cod true age class output. However, the 10 standard atlantis age classes aggregate true ages each for Cod.
+Lets try just the Z part for NOBA:
+
+```{r Znoba}
+
+# make a function for this
+# add YOY file to the config files
+YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))
+
+# load biolprm in some initialize file?
+biol <- load_biolprm(d.name, biol.prm.file)
+
+# get code matching species name to split YOY file
+code_ss <- funct.groups$Code[which(funct.groups$Name == species_ss)]
+
+# cut to a single species in YOY file
+YOY_ss <- YOY %>%
+  select(Time, paste0(code_ss, ".0"))
+
+# numbers at agecl at full resolution (all polygons and layers)
+truenums_ss <- truth$nums[truth$nums$species == species_ss,]
+
+#calc_Z <- function(yoy, nums, fgs, biolprm, toutinc)
+
+#oops, need to generalize calc_Z for subannual timesteps in YOY!
+#or input YOY only at 0, 365, etc since the numbers repeat for timesteps 
+YOY_ss <- YOY_ss %>%
+  filter(Time %in% seq(0, max(Time), by=365))
+
+fullresZ <- calc_Z(yoy = YOY_ss,
+                   nums = truenums_ss,
+                   fgs = funct.groups,
+                   biolprm = biol,
+                   toutinc = runpar$toutinc)
+
+# compare as above with mort.txt output
+file.mort <- file.path(d.name, paste0("output", scenario.name, "Mort.txt"))
+
+mortish <- read.table(file.mort, header = TRUE)
+
+relF_ss <- mortish %>%
+  select(Time, relF = paste0(code_ss, ".F"))
+
+relM_ss <- mortish %>%
+  select(Time, relM = paste0(code_ss, ".M"))
+
+Zish <- merge(relF_ss,relM_ss) %>%
+  mutate(relZ = relF + relM)
+
+annZ2 <- fullresZ %>%
+  mutate(yr = floor(time/stepperyr)+1) %>%
+  group_by(species, yr) %>%
+  summarise(Z = sum(atoutput))
+
+plotZ <- ggplot() +
+  geom_line(data=Zish, aes(x=Time/365, y=relZ, color="mort.txt Z")) +
+  geom_point(data=fullresZ, aes(x=time/stepperyr, y=atoutput, color="each timestep Z")) +
+  geom_point(data=annZ2, aes(x=yr-1, y=Z, color="sum timestep Z")) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
+
+plotZ + ylim(-1, 10.0)
+
+```
+
+This Z is a bad match. The Z values in NOBA's mort.txt are huge. Not sure what to make of this.
+
+The above figure is now the best it is going to get after a deeper dive back into `calc_Z` for NOBA before we proceed to true ages.
+Found one bug hidden by sardine's equal recruit_period and recend, another where the recruitment period is all within one output timestep that resulted in proportions over 1 (inflating the recruitment and then subtracting it off).
+Further, it appears that timesteps for recruitment were still not correctly aligned, so trying a different alignment that leaves out the timestep 0 from the truth.
+Not sure the latter is correct, or if we need to rethink the calculation of survival based on when recruits actually enter the model.
+
+Update: the new code actually makes the CCA sardine Z estimates more in line with each other, so the updates are correct and we are keeping them in `calc_Z`.
+
+Most important: Beth says that the M in mort.txt can really be as far off as we are seeing it here.
+The green line in the above NOBA Z plot reflects the initial estimate of what dies by predation mortality, which is a gigantic overesimate because many availability parameters are set to 1.
+After this output step, the predation mortality is rescaled within the model to account for the fact that there really aren't that many to be eaten.
+
+Our estimate of Z is from the actual numbers alive at the end of each timestep and the known input recruitment prior to any mortality, so I think it really is the best estimate we have.
+More debugging code commented out below but see the rmd file for the gory and redundant details.
+<!--
+```{r calcZ-snippet2, eval=FALSE}
+
+survey_testNall <- readRDS(file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))
+
+# numbers at ageclaggregated over layers, retain polygons (output of create_survey)
+survey_testN_ss <- survey_testNall[survey_testNall$species == species_ss,]
+
+yoy <- YOY_ss
+nums <- survey_testN_ss
+fgs <- funct.groups
+biolprm <- biol
+toutinc <- runpar$toutinc
+
+#calc_Z <- function(yoy, nums, fgs, biolprm, toutinc) {
+  # subset the yoy for species included in the fgs file
+  # that are turned on
+  # colnames of the recruit data are "Time" and a column for each
+  # species where the name is the functional group with an additional .0
+  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
+  recruits <- yoy[, colnames(yoy) %in% c("Time", paste0(turnedon$Code, ".0"))]
+
+  # mg carbon converted to wet weight in tonnes
+  k_wetdry <- biolprm$kgw2d / 1000000000
+  # Sum of structural and reserve nitrogen (KWSR_RN + KWSR_SN)
+  nitro <- merge(biolprm$kswr, biolprm$kwrr, by = "1")
+  nitro$sum <- apply(nitro[, 2:3], 1, sum)
+
+  # legacy: If output is from legacy code there will be an error in the
+  # yoy data, where the first row is in a different unit.
+  # yoy.txt is the biomass in tonnes per spawning event summed over the total
+  # model domain.
+  # The first row (< Nov/Dec 2015) is stored as biomass and the remaining rows
+  # are stored in numbers, must convert the entire matrix to biomass
+  # Check if legacy code and if so convert the numbers to biomass
+
+  # G.Fay 2/21/16 : changed yoy to recruits in below loop.
+  if (abs(recruits[1, 2] / recruits[2, 2]) > 10) {
+    recruits[2:NROW(recruits), 2:NCOL(recruits)] <- recruits[2:NROW(recruits), 2:NCOL(recruits)] *
+    nitro[match(gsub(".0", "", colnames(recruits)[-1]), nitro[, 1]), "sum"]
+  }
+
+  # Wide to long
+  recruits <- reshape(data = recruits, direction = "long",
+    varying = colnames(recruits)[-1],
+    v.names = "recruits",
+    times = colnames(recruits)[-1],
+    timevar = "group")
+  rownames(recruits) <- 1:NROW(recruits)
+  recruits <- recruits[, -which(colnames(recruits) == "id")]
+  # Switch from species code to species
+  recruits$group <- gsub("\\.0", "", recruits$group)
+  recruits <- merge(recruits, fgs[, c("Code", "Name")],
+    by.x = "group", by.y = "Code")
+
+  # merge recruits with strucn and resn of recruits from biol.prm file
+  recruits <- merge(recruits, nitro[, c(1, 4)],
+    by.x = "group", by.y = "1", all.x = TRUE, all.y = FALSE)
+  colnames(recruits)[which(colnames(recruits) == "recruits")] <- "recruitsbio"
+  # Get recruits in numbers rather than biomass
+  recruits$recruits <- recruits$recruitsbio / recruits$sum
+  recruits$yr <- as.integer(round(recruits$Time)/365) #needed to merge with totnums
+
+  # June 2019:Isaac determined that yr 0 (Time 0) in the YOY file is not real, don't use it
+  # code below assigns year 1 to times 0:stepperyr, all good output timesteps in truth$nums
+
+  # G.Fay 2/21/16
+  # UGLY code below tries to align fraction of annual yoy with timing of recruitment
+  # values in YOY.txt are total YOY that year waiting to recruit.
+  # seems to get rid of most of 'issues' - some v.minor survival >1,
+  #perhaps due to averaging of survival over toutinc days
+  nyrs <- ceiling(max(yoy$Time)/365)
+  times <- unique(yoy$Time)
+
+  # SKG June 2019
+  # need to match codes below, this mismatches when biolprm not in same order!!
+  #recstart_temp <- biolprm$recruit_time
+  #recstart_temp[,2] <- biolprm$time_spawn[-(grep('#',
+  #                     biolprm$time_spawn[,1])),2] + biolprm$recruit_time[,2]
+  # was this grep to get rid of a species with #XXX? breaks when there are none
+  #recstart_temp[,2] <- biolprm$time_spawn[,2] + biolprm$recruit_time[,2]
+
+  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%turnedon$Code,] #bug added codes with no YOY output
+  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%recruits$group,]
+
+  # June 12 2019 Beth determined that we do not need spawn_period
+
+  # rectiming is a dataframe with species code, time_spawn (day of year),
+  # recruit_time (number of days), and recruit period (number of days).
+  # we use these to calculate recstart (day of year) and recend (day of year)
+  rectiming <- merge(biolprm$time_spawn, biolprm$recruit_time, by = 1)
+  rectiming <- merge(rectiming, biolprm$recruit_period, by = 1)
+  names(rectiming) <- c("Code", "time_spawn", "recruit_time", "recruit_period")
+  rectiming <- rectiming %>%
+    mutate(recstart = time_spawn + recruit_time) %>%
+    mutate(recend = recstart + recruit_period)
+
+  #subsets to groups of interest
+  recstart_temp <- rectiming[rectiming$Code %in% recruits$group,]
+
+  # Sum numbers output over all boxes/depth/cohorts
+  # align model output timesteps (days) with recruitment periods (days)
+  totnums <- aggregate(atoutput ~ species + time, data = nums, sum) %>%
+    #mutate(time.days = (time+1)*toutinc) %>% #makes time 0 into days 0->73, etc
+    mutate(time.days = (time)*toutinc) %>% #makes time 1 into days 0->73, etc
+    mutate(yr = ceiling(time.days/365))  # yr 1 is 0:stepsperyr to match recruits yr1
+
+  totnums <- merge(totnums, recruits,
+                   by.x = c("yr", "species"), by.y = c("yr", "Name"),
+                   all.x = TRUE) %>%
+    arrange(time)
+
+  totnums$frac_recruit <- 0
+
+  for (irow in 1:nrow(recstart_temp)) {
+    group <- recstart_temp$Code[irow]
+    pick <- which(totnums$group == group)
+
+    recstart <- seq(recstart_temp$recstart[irow],by=365,length.out=nyrs)
+    recstart <- recstart[recstart<max(totnums$Time[pick])]
+    recend <- recstart + recstart_temp$recruit_period[irow]
+    #rec_times <- rbind(rec_times,cbind(group,recstart,recend))
+
+    for (i_rec in 1:length(recstart)) {
+      i_tstart <- which(pick==min(pick[totnums$time.days[pick]>=recstart[i_rec]]))
+      i_tstop <- which(pick==min(pick[totnums$time.days[pick]>=recend[i_rec]]))
+      if (i_rec == length(recstart)) i_tstop <- length(pick)
+      n_t <- 1+i_tstop-i_tstart
+      for (i_t in 1:n_t) {
+        t_temp <- totnums$time.days[pick[i_tstart+i_t-1]]
+        num_temp <- t_temp - recstart[i_rec]
+        if (i_t>1) {
+          if ((recend[i_rec]-t_temp)>toutinc) {
+            num_temp <- toutinc
+          }
+          else {
+            num_temp <- recend[i_rec]-(totnums$time.days[pick[i_tstart+i_t-2]])
+          }
+        }
+        frac_temp <- max(c(0,num_temp /(recend[i_rec]-recstart[i_rec])))
+        if(frac_temp > 1.0) frac_temp = 1.0
+        totnums$frac_recruit[pick[i_tstart+i_t-1]] <- frac_temp
+      }
+    }
+  }
+
+  #G.Fay 1/6/16, expand to num of recruits
+  # Recruit / mg C converted to wet weight in tonnes / redfield ratio of C:N
+  totnums$recruits <- totnums$recruits/k_wetdry/biolprm$redfieldcn
+
+  totnums$group <- recruits$group[match(totnums$species, recruits$Name)]
+  # For all time increments where there
+  totnums$recruits[is.na(totnums$recruits)] <- 0
+
+  totnums$annrecruits <- totnums$recruits
+
+  totnums$recruits <- totnums$annrecruits*totnums$frac_recruit
+
+  # Calculate survivors for each species group
+  totnums$survivors <- totnums$recruits
+
+  totnums$survival <- totnums$survivors
+  # Calculate survival for each group
+  for (group in unique(totnums$group)) {
+    #if(group == "SHD") browser()
+    pick <- which(totnums$group == group)
+    survival_temp <- c(NA,
+      totnums$survivors[pick[-1]]/totnums$atoutput[pick[-length(pick)]])
+
+    # G.Fay 2/21/16  "think" this is what things should be, recruits don't show up
+    # in numbers at age until time step after the recruitment event.
+     survival_temp <- c(
+       (totnums$atoutput[pick[-1]]-
+          totnums$recruits[pick[-1]])/totnums$atoutput[pick[-length(pick)]],NA)
+
+#     survival_temp <- c(
+#       (totnums$atoutput[pick[-1]])/
+#          (totnums$recruits[pick[-1]]+totnums$atoutput[pick[-length(pick)]]),NA)
+
+    survival_temp[survival_temp < 0] <- NA
+    # Use first positive value to replace the initial year and all negative vals
+    firstgood <- which(!is.na(survival_temp))[1]
+
+    survival_temp[1:firstgood] <- survival_temp[firstgood]
+    for(ii in seq_along(survival_temp)) {
+      if (is.na(survival_temp[ii])) {
+        nonzero <- which(which(survival_temp > 0) > ii)
+        if (length(nonzero) == 0) nonzero <- which(survival_temp > 0)
+        survival_temp[ii] <- survival_temp[which.min(abs(nonzero - ii))]
+      }
+    }
+    totnums$survival[pick] <- survival_temp
+   }
+
+  #Calculate Z
+  totnums$Z <- -1 * log(totnums$survival)
+  finaldata <- data.frame("species" = totnums$species,
+    "agecl" = NA, "polygon" = NA, "layer" = NA,
+    "time" = totnums$time, "atoutput" = totnums$Z)
+
+  #return(finaldata)
+#}
+
+```
+-->
+
+## Biological sampling: true numbers at age class
 
@@
 
-```{r Natage-plot, eval=FALSE}
+```{r Natage-plot}
 
@@
 
-Natage_ss <- Natage[Natage$species == "North_atl_cod",]
+Natage_ss <- Natage[Natage$species == species_ss,]
 
@@
 
-Perhaps overly optimistic, but if we feed these into `calc_stage2age` and get a result that matches the NOBA output ANNAGEBIO.nc, then we can assume the Zs are correct.
+This really is still overly optimistic, but now that I think `calc_Z` is working, if we feed these into `calc_stage2age` and get a result that matches the NOBA output ANNAGEBIO.nc, then we can assume that function is correct.
 
@@
 
-```{r plotannage, eval=FALSE}
+```{r plotannage}
+
+survey_testNall <- readRDS(file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))
+
+# numbers at ageclaggregated over layers, retain polygons (output of create_survey)
+survey_testN_ss <- survey_testNall[survey_testNall$species == species_ss,]
 
@@
 
-Not looking great, but at least not constant. This could be wrong Zs or wrong calculations in `calc_stage2age`. Load true ages from ANNAGEBIO.nc:
+Still not looking great, but at least not constant. I am now assuming we have the wrong calculations in `calc_stage2age`. Load true ages from ANNAGEBIO.nc:
 
-```{r compareages, eval=FALSE}
+```{r compareages}
 
@@
 
-trueage_atout_ss <- trueage_atout[trueage_atout$species == "North_atl_cod",]
+trueage_atout_ss <- trueage_atout[trueage_atout$species == species_ss,]
 
@@
 
-```{r plot-esttrueage-atltrueage, eval=FALSE}
+```{r plot-esttrueage-atltrueage}
 
@@
 Either way our early ages don't match.
-Something about the calculation or application of Z isn't correct.
+Need to step through `calc_stage2age` next!

### 2019-06-13 docs/index.Rmd
lets call calc_Z good

--- 2019-06-07 docs/index.Rmd
@@
 
+June 7: `calc_Z` seems incorrect when compared to outputMort.txt; stepping through code needed
+
+June 11: fixed matching of recruitment to timestep, still seems off
+
+June 13: I think we have the best `calc_Z` we are going to get.
+After discussion with Beth, we should not expect to match output of Mort.txt because M gets rescaled so much.
+However, the comparison looks reasonable now for sardine and even for NOBA cod.
+
+<a href="https://sgaichas.github.io/poseidon-dev/TestCalcZ2.html" target="_blank">Test Z calc and compare with output mort CCA and NOBA - June 7, 2019, updated June 13</a>
+
 ### Rabbit holes that turned out ok

### 2019-06-14 TestCalcStage2AgeNOBA.Rmd
annual age calcs better, still in progress

--- 2019-06-11 TestCalcZ2.Rmd
@@
 ---
-title: "Getting mortality from atlantisom: testing calc_Z function II"
+title: "Annual ages from atlantisom: testing calc_stage2age function"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets.
-Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
-On this page we demonstrate use of atlantisom on both California Current (CCA) and Norwegian Barents Sea Atlantis (NOBA) output files to test function `calc_Z` calculating total mortality.
-This is used both to derive "natural mortality" for comparison with stock assessments and to split age classes into true ages using the function `calc_stage2age`.
+Here we debug and test the functions that split atlantis age classes to true ages using the NOBA model, which has true annual age output for comparison with our results.
+These files are in the folder NOBAwithAnnAgeOutput (not kept on github due to large filesize).
 
-NOBA has output files in true ages that we can compare results with. These files are in the folder NOBAwithAnnAgeOutput (not kept on github due to large filesize).
+We need to estiamte true natural mortality (M) from atlantis so that we can compare it with what is used in stock assessment models; however this is an emergent property of atlantis resulting from predation, etc.
+rather than a model input.
+[We have already tested, revised, and debugged](https://sgaichas.github.io/poseidon-dev/TestCalcZ2.html) `calc_Z` which estimates total mortality using true output numbers at the end of a given timestep combined with information on recruitment from YOY.txt.
+We will derive M by subtracting the known F implemented in the atlantis run from this estmiated Z.
 
-We can also compare both models with the output[scenario.name]Mort.txt file, which reports annual M and F. The user's manual notes: 
->This file is currently only useful for looking at relative M vs F values for a species, as itdoes not give accurate mortalities
-
-So we can assume that the scale of Z from this file will not be the same as our calculation, but it should be useful for splitting out F to get M, and for comparing overall patterns.
-
-We will first use CCA here because the contrasting F scenario has been implemented. We will test with both sardines (M needed, but not age splitting) and hake (M and age splitting needed).
+The total mortality estimated by `calc_Z` is used to split age classes in the function `calc_stage2age` which we will debug here.
 
@@
 
-```{r initialize}
+```{r init-NOBA}
 
-initCCA <- TRUE
+initCCA <- FALSE
 initNEUS <- FALSE
-initNOBA <- FALSE
-
-species_ss <- c("Pacific_sardine")
-
-#function to make a config file? need one for each atlantis run
-
-if(initCCA) source(here("config/CC2Config.R"))
-
-if(initNEUS) source(here("config/NEUSConfig.R"))
+initNOBA <- TRUE
 
@@
 
-```
+species_ss <- "North_atl_cod"
 
-```{r get_names, message=FALSE, warning=FALSE}
 #Load functional groups
@@
 
-```
-
-```{r get_truth, message=FALSE, warning=FALSE}
-
-# default run_truth setup will save the file, so check for that first
-
+#Get true NOBAaa
 if(!file.exists(file.path(d.name, 
@@
 }
-```
 
-<!-- not sure we need catch bio for this
-```{r loadcatchb}
-
-truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)
-
-truecatchbio_ss <- truecatchbio[truecatchbio$species == species_ss,]
-# note: time output of this file is in days
-# what model timestep is this output matching?
-# rescale catch in biomass time from days to years
-truecatchbio_ss$time <- as.integer(truecatchbio_ss$time/365)
-
-```
--->
-
-## Test calc Z function
-
-This section tests the `atlantisom::calc_Z()` function step by step.
-Ultimately this output goes to `calc_stage2age` to produce annual numbers at true age, which we can compare with the ANNAGEBIO.nc output produced by NOBA.
-For now we will just compare to the Mort.txt output.
-
-```{r census-spec, message=FALSE, warning=FALSE}
-
-# make a function for this
 source(here("config/census_spec.R"))
@@
 
-Output timestep toutinc for the population is `r runpar$toutinc `,
+Reviewing just the Z part for NOBA:
 
-so steps per year in run_truth output is `r stepperyr `
+```{r Znoba}
 
-and the number of output steps in run_truth output is `r noutsteps `.
-
-### True numbers total (annual)
-
-This gives us the true total numbers on an annual basis. We can compare this with age specific outputs.
-
-```{r surveyNbased, eval=FALSE}
-
-# this uses result$nums and a new function to get survey index in numbers (abundance)
-
-survey_testNall <- create_survey(dat = truth$nums,
-                                 time = timeall,
-                                 species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-# save this, needed as input to calc_Z?
-saveRDS(survey_testNall, file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))
-
-# as above, make up a constant 0 cv for testing
-surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
-
-surveyN <- sample_survey_numbers(survey_testNall, surv_cv)
-
-#save for later use, takes a long time to generate
-saveRDS(surveyN, file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))
-
-```
-
-```{r survNtrue-plot}
-
-surveyN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))
-
-surveyN_ss <- surveyN[surveyN$species == species_ss,]
-
-plotN <-ggplot() +
-  geom_line(data=surveyN_ss, aes(x=time/stepperyr,y=atoutput, color="survey census N"), 
-            alpha = 10/10) +
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-plotN +
-  facet_wrap(~species, scales="free")
-
-```
-
-### What level of aggregation works for calc_Z?
-
-The atlantisom function `calc_stage2age` calls the `calc_Z` function, applying a total mortality rate to estimate the numbers at true age within an age class, but it is unclear whether to use this on the aggregated numbers or the full resolution numbers at age.
-Here we test outputs of Z for several levels of aggregation:
-
-```{r testZagg}
+# make a function for this
 # add YOY file to the config files
@@
 
-survey_testNall <- readRDS(file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))
+#calc_Z <- function(yoy, nums, fgs, biolprm, toutinc)
 
-# numbers at agecl aggregated over layers, retain polygons (output of create_survey)
-survey_testN_ss <- survey_testNall[survey_testNall$species == species_ss,]
-
-# numbers at agecl aggregated over layers and polygons
-Natage_ss <- readRDS(file.path(d.name, paste0(scenario.name, "natage_census_sard.rds")))
-
-#calc_Z <- function(yoy, nums, fgs, biolprm, toutinc)
+#oops, need to generalize calc_Z for subannual timesteps in YOY!
+#or input YOY only at 0, 365, etc since the numbers repeat for timesteps 
+YOY_ss <- YOY_ss %>%
+  filter(Time %in% seq(0, max(Time), by=365))
 
@@
 
-surveyresZ <- calc_Z(yoy = YOY_ss,
-                   nums = survey_testN_ss,
-                   fgs = funct.groups,
-                   biolprm = biol,
-                   toutinc = runpar$toutinc)
-
-sampleresZ <- calc_Z(yoy = YOY_ss,
-                   nums = Natage_ss,
-                   fgs = funct.groups,
-                   biolprm = biol,
-                   toutinc = runpar$toutinc)
-
-```
-
-Fixed a bug where grep was used if a # appeared. This ran for the survey resolution numbers. I then went back and ran the two other levels above after fixing `calc_Z.`
-
-Both true full resolution numbers and survey resolution numbers (aggregated over polygon) returned nonzero Z values.
-As with previous the NOBA test, they are identical aside from eight values with no apparent pattern (rows 100, 105, 180, 200, 220, 240, 355, 400) out of 500 that differ at e-16.
-Same story with difference from sample resolution; this does not matter to output.
-
-```{r differentZ}
-fullresZ$atoutput - surveyresZ$atoutput
-
-fullresZ$atoutput - sampleresZ$atoutput
-```
-
-I would be happer if the differences were all exactly 0, but I suppose we can attribute that to rounding error.
-
-### What is F at each timestep?
-A relative annual F is found in outputMort.txt!
-```{r getrelF}
-
+# compare as above with mort.txt output
 file.mort <- file.path(d.name, paste0("output", scenario.name, "Mort.txt"))
@@
 
-plotF <- ggplot(data=relF_ss, aes(x=Time/365, y=relF)) +
-  geom_line() +
+Zish <- merge(relF_ss,relM_ss) %>%
+  mutate(relZ = relF + relM)
+
+annZ2 <- fullresZ %>%
+  mutate(yr = floor(time/stepperyr)+1) %>%
+  group_by(species, yr) %>%
+  summarise(Z = sum(atoutput))
+
+plotZ <- ggplot() +
+  geom_line(data=Zish, aes(x=Time/365, y=relZ, color="mort.txt Z")) +
+  geom_point(data=fullresZ, aes(x=time/stepperyr, y=atoutput, color="each timestep Z")) +
+  geom_point(data=annZ2, aes(x=yr-1, y=Z, color="sum timestep Z")) +
   theme_tufte() +
@@
 
-plotF
+plotZ + ylim(-1, 10.0)
 
@@
 
-This is consistent with the scenario implemented. So far so good.
+This Z is a bad match. The Z values in NOBA cod's mort.txt are huge. The above figure is now the best it is going to get though.
 
-This probably depends on how the model is set up, but for models forced with F by species we ought to be able to read this out from an input file.
-If it is a daily rate then we mulitply by toutinc to get the rate at each of these output timestep Z values.
+Beth says that the M in mort.txt can really be as far off as we are seeing it here.
+The green line in the above NOBA Z plot reflects the initial estimate of what dies by predation mortality, which is a gigantic overesimate because many availability parameters are set to 1.
+After this output step, the predation mortality is rescaled within the model to account for the fact that there really aren't that many to be eaten.
 
-### Quick test to see if Z is right?
-Compare with outputMort.txt!
-```{r compZ}
+Our estimate of Z is from the actual numbers alive at the end of each timestep and the known input recruitment prior to any mortality, so I think it really is the best estimate we have.
 
-Zish <- merge(relF_ss,relM_ss) %>%
-  mutate(relZ = relF + relM)
-
-plotZ <- ggplot() +
-  geom_line(data=Zish, aes(x=Time/365, y=relZ)) +
-  geom_point(data=fullresZ, aes(x=(time*73)/365, y=atoutput)) +
-  theme_tufte() +
-  labs(subtitle = paste(scenario.name, species_ss))
-
-plotZ
-```
-
-Nope. We have a constant value punctuated by occassional variation. Seems wrong. Variation doesn't look anything like the txt file either. Perhaps the 5x yr timesteps are messing this up?
-
-Time to step through line by line again. Revised as below. Better (apportions annual recruitment to timesteps), but not sure this is fully correct?
-
-```{r calczsnippet, eval=FALSE}
-
-# test with one species at survey scale (layers aggregated, not time or polygon)
-yoy <- YOY_ss
-nums <- survey_testN_ss
-fgs <- funct.groups
-biolprm <- biol
-toutinc <- runpar$toutinc
-
-# everything below the function definition line
-  # subset the yoy for species included in the fgs file
-  # that are turned on
-  # colnames of the recruit data are "Time" and a column for each
-  # species where the name is the functional group with an additional .0
-  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
-  recruits <- yoy[, colnames(yoy) %in% c("Time", paste0(turnedon$Code, ".0"))]
-
-  # mg carbon converted to wet weight in tonnes
-  k_wetdry <- biolprm$kgw2d / 1000000000
-  # Sum of structural and reserve nitrogen (KWSR_RN + KWSR_SN)
-  nitro <- merge(biolprm$kswr, biolprm$kwrr, by = "1")
-  nitro$sum <- apply(nitro[, 2:3], 1, sum)
-
-  # legacy: If output is from legacy code there will be an error in the
-  # yoy data, where the first row is in a different unit.
-  # yoy.txt is the biomass in tonnes per spawning event summed over the total
-  # model domain.
-  # The first row (< Nov/Dec 2015) is stored as biomass and the remaining rows
-  # are stored in numbers, must convert the entire matrix to biomass
-  # Check if legacy code and if so convert the numbers to biomass
-
-  # G.Fay 2/21/16 : changed yoy to recruits in below loop.
-  if (abs(recruits[1, 2] / recruits[2, 2]) > 10) {
-    recruits[2:NROW(recruits), 2:NCOL(recruits)] <- recruits[2:NROW(recruits), 2:NCOL(recruits)] *
-    nitro[match(gsub(".0", "", colnames(recruits)[-1]), nitro[, 1]), "sum"]
-  }
-
-  # Wide to long
-  recruits <- reshape(data = recruits, direction = "long",
-    varying = colnames(recruits)[-1],
-    v.names = "recruits",
-    times = colnames(recruits)[-1],
-    timevar = "group")
-  rownames(recruits) <- 1:NROW(recruits)
-  recruits <- recruits[, -which(colnames(recruits) == "id")]
-  # Switch from species code to species
-  recruits$group <- gsub("\\.0", "", recruits$group)
-  recruits <- merge(recruits, fgs[, c("Code", "Name")],
-    by.x = "group", by.y = "Code")
-
-  # merge recruits with strucn and resn of recruits from biol.prm file
-  recruits <- merge(recruits, nitro[, c(1, 4)],
-    by.x = "group", by.y = "1", all.x = TRUE, all.y = FALSE)
-  colnames(recruits)[which(colnames(recruits) == "recruits")] <- "recruitsbio"
-  # Get recruits in numbers rather than biomass
-  # these match rows 2:end of the legacy file, so good
-  # but is this 1000s or numbers? not real numbers, Gavin corrects below
-  recruits$recruits <- recruits$recruitsbio / recruits$sum
-  recruits$yr <- as.integer(round(recruits$Time)/365)
-
-  # G.Fay 2/21/16
-  # UGLY code below tries to align fraction of annual yoy with timing of recruitment
-  # values in YOY.txt are total YOY that year waiting to recruit.
-  # seems to get rid of most of 'issues' - some v.minor survival >1,
-  #perhaps due to averaging of survival over toutinc days
-  nyrs <- ceiling(max(yoy$Time)/365)
-  times <- unique(yoy$Time)
-
-  # SKG June 2019
-  # need to match codes below, this mismatches when biolprm not in same order!!
-  #recstart_temp <- biolprm$recruit_time
-  #recstart_temp[,2] <- biolprm$time_spawn[-(grep('#',
-  #                     biolprm$time_spawn[,1])),2] + biolprm$recruit_time[,2]
-  # was this grep to get rid of a species with #XXX? breaks when there are none
-  #recstart_temp[,2] <- biolprm$time_spawn[,2] + biolprm$recruit_time[,2]
-
-  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%turnedon$Code,] #bug added codes with no YOY output
-  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%recruits$group,]
-
-  # need spawn_period?? if so read out in load_biolprm and merge it here too
-
-  # rectiming is a dataframe with species code, time_spawn (day of year), 
-  # recruit_time (number of days), and recruit period (number of days). 
-  # we use these to calculate recstart (day of year) and recend (day of year)
-  rectiming <- merge(biolprm$time_spawn, biolprm$recruit_time, by = 1)
-  rectiming <- merge(rectiming, biolprm$recruit_period, by = 1)
-  names(rectiming) <- c("Code", "time_spawn", "recruit_time", "recruit_period")
-  rectiming <- rectiming %>%
-    mutate(recstart = time_spawn + recruit_time) %>% # possibly + spawn_period
-    mutate(recend = recstart + recruit_period)
-
-  #subsets to groups of interest
-  recstart_temp <- rectiming[rectiming$Code %in% recruits$group,]
-
-  # do we need this fraction calculation if YOY is annual snapshot?
-  # (check output timing of YOY file with tsumout, available in run.prm)
-  # can we not just allocate the full YOY to the timestep where they come in?
-  # if recstart to recend spans multiple toutinc timesteps then need fraction
-
-  # align model output timesteps (days) with recruitment periods (days)
-  numstime.days <- unique(nums$time)*toutinc
-  # Sum numbers output over all boxes/depth/cohorts
-  totnums <- aggregate(atoutput ~ species + time, data = nums, sum) %>%
-    mutate(time.days = time*toutinc) %>%
-    mutate(yr = floor(time.days/365))
-
-  totnums <- merge(totnums, recruits,
-    by.x = c("yr", "species"), by.y = c("yr", "Name"),
-    all.x = TRUE) %>%
-    arrange(time)
-
-  totnums$frac_recruit <- 0
-
-  for (irow in 1:nrow(recstart_temp)) {
-    group <- recstart_temp$Code[irow]
-    pick <- which(totnums$group == group)
-
-    recstart <- seq(recstart_temp$recstart[irow],by=365,length.out=nyrs)
-    recstart <- recstart[recstart<max(totnums$Time[pick])]
-    recend <- recstart + recstart_temp$recend[irow]
-    #rec_times <- rbind(rec_times,cbind(group,recstart,recend))
-
-    for (i_rec in 1:length(recstart)) {
-      i_tstart <- which(pick==min(pick[numstime.days[pick]>=recstart[i_rec]]))
-      i_tstop <- which(pick==min(pick[numstime.days[pick]>=recend[i_rec]]))
-      if (i_rec == length(recstart)) i_tstop <- length(pick)
-      n_t <- 1+i_tstop-i_tstart
-      for (i_t in 1:n_t) {
-        t_temp <- numstime.days[pick[i_tstart+i_t-1]]
-        num_temp <- t_temp - recstart[i_rec]
-        if (i_t>1) {
-          if ((recend[i_rec]-t_temp)>toutinc) {
-            num_temp <- toutinc
-          }
-          else {
-            num_temp <- recend[i_rec]-(numstime.days[pick[i_tstart+i_t-2]])
-          }
-        }
-        frac_temp <- max(c(0,num_temp /(recend[i_rec]-recstart[i_rec])))
-        totnums$frac_recruit[pick[i_tstart+i_t-1]] <- frac_temp
-      }
-    }
-  }
-
-# Time in YOY does not match recstart-recend period still after this loop--fixed
-
-  # match "Time" of the young of the year with the time-step periodicity
-  # listed in the run.prm or run.xml file
-  #recruits$Time <- recruits$Time / toutinc
-
-# SKG did this above already in the following line?
-  # Get recruits in numbers rather than biomass
-  #recruits$recruits <- recruits$recruitsbio / recruits$sum
-
-  #G.Fay 1/6/16, expand to num of recruits
-  # Recruit / mg C converted to wet weight in tonnes / redfield ratio of C:N
-  totnums$recruits <- totnums$recruits/k_wetdry/biolprm$redfieldcn
-
-  # Combine recruits and numbers
-  # Only pull recruits from the yearly time step, where the
-  # yearly time step matches the time step
-  # totnums <- merge(totnums, recruits,
-  #   by.x = c("time", "species"), by.y = c("Time", "Name"),
-  #   all.x = TRUE, all.y = FALSE)
-  totnums$group <- recruits$group[match(totnums$species, recruits$Name)]
-  # For all time increments where there
-  totnums$recruits[is.na(totnums$recruits)] <- 0
-
-   # Make sure time is in order, did above
-  #totnums <- totnums[order(totnums$species, totnums$time), ]
-  totnums$annrecruits <- totnums$recruits
-
-  totnums$recruits <- totnums$annrecruits*totnums$frac_recruit
-
-  # Calculate survivors for each species group
-  totnums$survivors <- totnums$recruits
-
-  totnums$survival <- totnums$survivors
-  # Calculate survival for each group
-  for (group in unique(totnums$group)) {
-    #if(group == "SHD") browser()
-    pick <- which(totnums$group == group)
-    survival_temp <- c(NA,
-      totnums$survivors[pick[-1]]/totnums$atoutput[pick[-length(pick)]])
-
-    # G.Fay 2/21/16  "think" this is what things should be, recruits don't show up
-    # in numbers at age until time step after the recruitment event.
-     survival_temp <- c(
-       (totnums$atoutput[pick[-1]]-
-          totnums$recruits[pick[-1]])/totnums$atoutput[pick[-length(pick)]],NA)
-
-#     survival_temp <- c(
-#       (totnums$atoutput[pick[-1]])/
-#          (totnums$recruits[pick[-1]]+totnums$atoutput[pick[-length(pick)]]),NA)
-
-    survival_temp[survival_temp < 0] <- NA
-    # Use first positive value to replace the initial year and all negative vals
-    firstgood <- which(!is.na(survival_temp))[1]
-
-    survival_temp[1:firstgood] <- survival_temp[firstgood]
-    for(ii in seq_along(survival_temp)) {
-      if (is.na(survival_temp[ii])) {
-        nonzero <- which(which(survival_temp > 0) > ii)
-        if (length(nonzero) == 0) nonzero <- which(survival_temp > 0)
-        survival_temp[ii] <- survival_temp[which.min(abs(nonzero - ii))]
-      }
-    }
-    totnums$survival[pick] <- survival_temp
-   }
-
-  #Calculate Z
-  totnums$Z <- -1 * log(totnums$survival)
-  finaldata <- data.frame("species" = totnums$species,
-    "agecl" = NA, "polygon" = NA, "layer" = NA,
-    "time" = totnums$time, "atoutput" = totnums$Z)
-
-surveyresZ <- finaldata
-
-```
-
-### How to turn timestep Z into annual Z for comparison?
-
-### Biological sampling: true numbers at age class
-
-Might be a good time to switch back to NOBA for comparison with true annual ages.
-
-Everything below here is old.
-
-We get true numbers at age at each output step by appling `sample_fish` to the survey_testNall result if the model group has one true age per output age class:
-
-```{r truenatagecl, eval=FALSE}
-# We need an effective N for the sample fish function, setting equal to actual N
-# this one is high but not equal to total for numerous groups
-
-# this doesnt work for sardines in CCA, change to full census
-
-effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
-
-numsallhigh <- sample_fish(survey_testNall, effNhigh)
-
-#save for later use, takes a long time to generate
-saveRDS(numsallhigh, file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
-
-```
-
-This is NOBA Cod true age class output. However, the 10 standard atlantis age classes aggregate true ages each for Cod.
+## Biological sampling: true numbers at age class
 
@@
 
-Perhaps overly optimistic, but if we feed these into `calc_stage2age` and get a result that matches the NOBA output ANNAGEBIO.nc, then we can assume the Zs are correct.
+Now we go line by line through `calc_stage2age`:
+
+```{r calcstage2age-snippet, eval=FALSE}
+
+nums_data <- truenums_ss
+yoy <- YOY_ss
+fgs <- funct.groups
+biolprm <- biol
+runprm <- load_runprm(d.name, run.prm.file)
+
+#full function code below
+#calc_stage2age <- function(nums_data, biolprm, yoy, fgs, runprm) {
+
+  # subset the yoy for those species that are included in the fgs file
+  # that are turned on
+  species.code <- fgs$Code
+  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
+  names <- turnedon$Code
+
+  # Figure out the groups that have multiple ages classes in each stage (or
+  # cohort)
+  multiple_ages <- turnedon[turnedon$NumAgeClassSize>1, c(1,4,10)]
+  num_multi_age <- dim(multiple_ages)[1]
+
+  ntimesteps <- length(unique(nums_data$time))
+
+  # For each species with multiple ages, use the calc_Z function to get time
+  # varying Z values and put that all into one dataframe
+  Z.dataframe <- data.frame()
+  #for(i in 1:num_multi_age) {
+    #temp_nums <- nums_data[nums_data$species==multiple_ages$Name[i],]
+    temp_nums <- nums_data[nums_data$species %in% multiple_ages$Name,]
+    temp_Z <- calc_Z(yoy=yoy, nums=temp_nums, fgs=fgs, biolprm=biolprm, toutinc=runprm$toutinc)
+    #Z.dataframe <- rbind(Z.dataframe, temp_Z)
+    Z.dataframe <- temp_Z
+  #}
+
+  # HACK for now, since some Z values are negative, I am replacing them
+  # with a random number
+  randomZ <- runif(length(which(Z.dataframe$atoutput<0)))
+  Z.dataframe[which(Z.dataframe$atoutput<0),c("atoutput")] <- randomZ
+
+  # rename the "atoutput" column, since it is really just the mortality, and
+  # we want to retain it for the next step of merging
+  colnames(Z.dataframe)[6] <- c("Z")
+
+  # merge together nums_data and the output Z.dataframe:
+  new_nums <- merge(nums_data, Z.dataframe[,c("species", "time", "Z")],
+                    by=c("species", "time"), all.x=TRUE)
+
+  # since we have to expand the number of rows for groups with multiple true
+  # age classes, we will loop through species and time, creating a new list
+  # that each element will be a single species in a single time step but
+  # across ages and boxes (these wil all be put together in the end)
+  temp.list <- list()
+
+  #SKG subset turnedon and only loop through species present in the nums input
+  turnedon_sub <- turnedon[which(turnedon$Name == unique(new_nums$species)),]
+  names <- turnedon_sub$Code
+
+  for(i in 1:length(names)) {
+    # looping all species -- those with or without multiple true ages
+
+    group.i <- turnedon_sub$Name[i]
+    nums_species <- new_nums[new_nums$species==group.i,] # might need to
+    num_ages <- turnedon_sub$NumAgeClassSize[i]
+    sp_times <- sort(unique(nums_species$time))
+    n_sp_tsteps <- length(sp_times)
+    # these last pieces are needed because not all species are present at all times
+
+    # Check if multiple true ages, if not then just save species_nums to list
+    if(num_ages==1) { temp.list[[length(temp.list)+1]] <- nums_species
+    } else if(num_ages>1) {
+      # take out the part of the Z.dataframe for species group i
+      Zvals <- Z.dataframe[Z.dataframe$species==group.i,]
+
+      # create empty list for this species i, each element in the list will be
+      # a different time step
+      list_species <- list()
+
+      for(j in sp_times) { # looping through time
+        # get the Z val for the timestep in question--fixed bug here, was getting time
+        Zval.j <- Zvals$Z[Zvals$time==j]
+
+        # and turn the Z value into a vector of survival values across the
+        # number of true ages for each age class for this species
+        nums_vec <- 1
+        for(k in 1:(num_ages-1)) {
+          nums_vec <- c(nums_vec, exp(-Zval.j*k))
+        }
+        nums_proportion <- nums_vec/sum(nums_vec)
+
+        # take out the species numbers only for time step j
+        nums_sp_subset <- nums_species[nums_species$time==j, ]
+
+        # create an empty list, each element of this list will be a row from
+        # the nums_sp_subset data frame that is split into multiple pieces
+        # to make the atoutput column have the correct dimensions
+        list_ages <- list()
+
+        # loop through all the rows in nums_sp_subset
+        for(l in 1:nrow(nums_sp_subset)) {
+          nums_row <- nums_sp_subset[l,]
+          # create new data frame with same columns, but just more rows to
+          # account for all the true ages
+          new_rows <- data.frame(species=nums_row$species,
+                                 time=nums_row$time,
+                                 agecl=seq((nums_row$agecl*num_ages - num_ages+1),
+                                           nums_row$agecl*num_ages),
+                                 polygon=nums_row$polygon,
+                                 layer=nums_row$layer,
+                                 atoutput=nums_row$atoutput*nums_proportion,
+                                 Z=nums_row$Z
+                                 )
+
+          ### NOTE: the agecl piece is funny -- because now we are making many
+          # more age classes for some... I think this works? But might want
+          # another set of eyes to check my logic
+
+          list_ages[[length(list_ages)+1]] <- new_rows
+        }
+
+        # now combine all the elements of the list for different rows and make
+        # it an element of the list_species (to represent one time step)
+        list_species[[length(list_species)+1]] <- do.call("rbind", list_ages)
+      }
+      temp.list[[length(temp.list)+1]] <- do.call("rbind", list_species)
+    }
+  }
+
+  # Now just create the output
+  finalout_withZ <- do.call("rbind", temp.list)
+
+  ### CHECK:
+  # do we want to remove the column of Z values? I think so
+  finalout <- finalout_withZ[,-7]
+
+  #return(finalout)
+
+#}
+
+```
+
+Fixed one bug where Z was not being assigned, that should help. Also subset for only species passed in the numbers file; might speed things up.
+
+Let's try `calc_stage2age` again and if get a result that matches the NOBA output ANNAGEBIO.nc, then we can assume our function is correct.
 
@@
 
-Not looking great, but at least not constant. This could be wrong Zs or wrong calculations in `calc_stage2age`. Load true ages from ANNAGEBIO.nc:
+This is an improvement now that Z is actually being read in.
+We can still probably tweak the calculations in `calc_stage2age` to smooth the difference if necessary.
+The best comparison is with true ages from ANNAGEBIO.nc:
 
-```{r compareages, eval=FALSE}
+```{r compareages}
 
@@
 
-We only have the first 10 cohorts of true age output due to hardcoding in load_nc.
-I may write another load function that changes that.
-Either way our early ages don't match.
-Something about the calculation or application of Z isn't correct.
+We only have the first 10 cohorts of true age output due to hardcoding in load_nc. I will write another load function that changes that.
+
+We have matches for timesteps 0-2 and then divergence. It looks like matches are actually better in the recruitment timesteps.
+
+Over all timesteps our early ages are closer now and the later ones look on target (because there is less contrast between the split ages).
+
+Slowly improving.

### 2019-06-14 docs/index.Rmd
annual age calcs better, still in progress

@@
 
+June 14: fixed a bug and `calc_stage2age` works better now that it is using Z, and now that Zs are in the right ballpark. Still not a perfect match to true age outptu.
+
+<a href="https://sgaichas.github.io/poseidon-dev/TestCalcStage2AgeNOBA.html" target="_blank">Test stage to age calculationa and compare true age output NOBA - June 14, 2019</a>
+
 ### Rabbit holes that turned out ok

### 2019-06-14 docs/index.Rmd
pesky typos

--- 2019-06-13 docs/index.Rmd
@@
 
+June 14: fixed a bug and `calc_stage2age` works better now that it is using Z, and now that Zs are in the right ballpark. Still not a perfect match to true age output.
+
+<a href="https://sgaichas.github.io/poseidon-dev/TestCalcStage2AgeNOBA.html" target="_blank">Test stage to age calculations and compare true age output NOBA - June 14, 2019</a>
+
 ### Rabbit holes that turned out ok

### 2019-06-14 FullBoccaccioTruthEx2.Rmd
try bocaccio new CCA age splitting

--- 2019-05-27 FullBoccaccioTruthEx.Rmd
@@
 ---
-title: "Boccaccio example: end to end truth for California Current"
+title: "Boccaccio example 2: end to end truth for California Current"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-  * "CalCurrent2013_OA_off": California Current Atlantis model output from Isaac Kaplan's google drive \outputFolderAllOff2013Oceanography using the Hodgson et al.
-2018 configuration.
-Corresponds to CC1Config.R file.
+  * "CalCurrent2063_run11": California Current Atlantis model output from Isaac Kaplan's google drive \output_CC_2063_run11 using the Hodgson et al.
+2018 configuration, and driven with the F time series and recruitment variability we plan to use for our Fragile Ecosystems, Robust Assessments project.
+Corresponds to CC2Config.R file.
 
@@
 
+species_ss <- "Bocaccio_rockfish"
+
 #function to make a config file? need one for each atlantis run
 
-if(initCCA) source(here("config/CC1Config.R"))
+if(initCCA) source(here("config/CC2Config.R"))
 
@@
 
-# call sample_survey_biomass with a bunch of 1s for weight at age
+# call sample_survey_biomass with a bunch of 1000s for weight at age
 # in the code it multiplies atoutput by wtatage so this allows us to use
@@
 
-surveyB_frombio_ss <- surveyB_frombio[surveyB_frombio$species == "Bocaccio_rockfish",]
+surveyB_frombio_ss <- surveyB_frombio[surveyB_frombio$species == species_ss,]
 
@@
   gather(species, biomass, -Time) %>%
-  filter(species %in% c("Bocaccio_rockfish"))
+  filter(species %in% species_ss)
 
@@
 compareB <-ggplot() +
-  geom_line(data=surveyB_frombio_ss, aes(x=time/5,y=atoutput, color="survey census B"), 
+  geom_line(data=surveyB_frombio_ss, aes(x=time/stepperyr,y=atoutput, color="survey census B"), 
             alpha = 10/10) +
@@
 
-surveyN_ss <- surveyN[surveyN$species == "Bocaccio_rockfish",]
+surveyN_ss <- surveyN[surveyN$species == species_ss,]
 
 plotN <-ggplot() +
-  geom_line(data=surveyN_ss, aes(x=time/5,y=atoutput, color="survey census N"), 
+  geom_line(data=surveyN_ss, aes(x=time/stepperyr,y=atoutput, color="survey census N"), 
             alpha = 10/10) +
@@
 # this one is high but not equal to total for numerous groups
-effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
+survey_testNall <- readRDS(file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))
 
-numsallhigh <- sample_fish(survey_testNall, effNhigh)
+numsall <- aggregate(survey_testNall$atoutput,list(survey_testNall$species,survey_testNall$agecl,survey_testNall$time),sum)
+names(numsall) <- c("species","agecl","time","atoutput")
+numsallhigh <- data.frame(species = numsall$species,
+					  agecl = numsall$agecl,
+					  polygon = NA,
+					  layer = NA,
+					  time = numsall$time,
+					  atoutput = numsall$atoutput)
+
+# better to do an actual census above, can't make effNhigh high enough for some spp
+#effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
+#numsallhigh <- sample_fish(survey_testNall, effNhigh)
 
@@
 
-```{r testagecl, eval = FALSE}
+```{r testagecl, eval=FALSE}
 # add YOY file to the config files
@@
 
+Calc Z is breaking for one of the many species; Isaac says something like tuna, sharks, or salmon may not fit this design, so let's check who is having problems:
+
+```{r calcZsnippetagain, eval=FALSE}
+
+yoy <- YOY
+nums <- survey_testNall
+fgs <- funct.groups
+biolprm <- biol
+toutinc <- runpar$toutinc
+
+#calc_Z <- function(yoy, nums, fgs, biolprm, toutinc) {
+  # subset the yoy for species included in the fgs file
+  # that are turned on
+  # colnames of the recruit data are "Time" and a column for each
+  # species where the name is the functional group with an additional .0
+  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
+  recruits <- yoy[, colnames(yoy) %in% c("Time", paste0(turnedon$Code, ".0"))]
+
+  # mg carbon converted to wet weight in tonnes
+  k_wetdry <- biolprm$kgw2d / 1000000000
+  # Sum of structural and reserve nitrogen (KWSR_RN + KWSR_SN)
+  nitro <- merge(biolprm$kswr, biolprm$kwrr, by = "1")
+  nitro$sum <- apply(nitro[, 2:3], 1, sum)
+
+  # legacy: If output is from legacy code there will be an error in the
+  # yoy data, where the first row is in a different unit.
+  # yoy.txt is the biomass in tonnes per spawning event summed over the total
+  # model domain.
+  # The first row (< Nov/Dec 2015) is stored as biomass and the remaining rows
+  # are stored in numbers, must convert the entire matrix to biomass
+  # Check if legacy code and if so convert the numbers to biomass
+
+  # G.Fay 2/21/16 : changed yoy to recruits in below loop.
+  if (abs(recruits[1, 2] / recruits[2, 2]) > 10) {
+    recruits[2:NROW(recruits), 2:NCOL(recruits)] <- recruits[2:NROW(recruits), 2:NCOL(recruits)] *
+    nitro[match(gsub(".0", "", colnames(recruits)[-1]), nitro[, 1]), "sum"]
+  }
+
+  # Wide to long
+  recruits <- reshape(data = recruits, direction = "long",
+    varying = colnames(recruits)[-1],
+    v.names = "recruits",
+    times = colnames(recruits)[-1],
+    timevar = "group")
+  rownames(recruits) <- 1:NROW(recruits)
+  recruits <- recruits[, -which(colnames(recruits) == "id")]
+  # Switch from species code to species
+  recruits$group <- gsub("\\.0", "", recruits$group)
+  recruits <- merge(recruits, fgs[, c("Code", "Name")],
+    by.x = "group", by.y = "Code")
+
+  # merge recruits with strucn and resn of recruits from biol.prm file
+  recruits <- merge(recruits, nitro[, c(1, 4)],
+    by.x = "group", by.y = "1", all.x = TRUE, all.y = FALSE)
+  colnames(recruits)[which(colnames(recruits) == "recruits")] <- "recruitsbio"
+  # Get recruits in numbers rather than biomass
+  recruits$recruits <- recruits$recruitsbio / recruits$sum
+  recruits$yr <- as.integer(round(recruits$Time)/365) #needed to merge with totnums
+
+  # June 2019:Isaac determined that yr 0 (Time 0) in the YOY file is not real, don't use it
+  # code below assigns year 1 to times 0:stepperyr, all good output timesteps in truth$nums
+
+  # G.Fay 2/21/16
+  # UGLY code below tries to align fraction of annual yoy with timing of recruitment
+  # values in YOY.txt are total YOY that year waiting to recruit.
+  # seems to get rid of most of 'issues' - some v.minor survival >1,
+  #perhaps due to averaging of survival over toutinc days
+  nyrs <- ceiling(max(yoy$Time)/365)
+  times <- unique(yoy$Time)
+
+  # SKG June 2019
+  # need to match codes below, this mismatches when biolprm not in same order!!
+  #recstart_temp <- biolprm$recruit_time
+  #recstart_temp[,2] <- biolprm$time_spawn[-(grep('#',
+  #                     biolprm$time_spawn[,1])),2] + biolprm$recruit_time[,2]
+  # was this grep to get rid of a species with #XXX? breaks when there are none
+  #recstart_temp[,2] <- biolprm$time_spawn[,2] + biolprm$recruit_time[,2]
+
+  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%turnedon$Code,] #bug added codes with no YOY output
+  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%recruits$group,]
+
+  # June 12 2019 Beth determined that we do not need spawn_period
+
+  # rectiming is a dataframe with species code, time_spawn (day of year),
+  # recruit_time (number of days), and recruit period (number of days).
+  # we use these to calculate recstart (day of year) and recend (day of year)
+  rectiming <- merge(biolprm$time_spawn, biolprm$recruit_time, by = 1)
+  rectiming <- merge(rectiming, biolprm$recruit_period, by = 1)
+  names(rectiming) <- c("Code", "time_spawn", "recruit_time", "recruit_period")
+  rectiming <- rectiming %>%
+    mutate(recstart = time_spawn + recruit_time) %>%
+    mutate(recend = recstart + recruit_period)
+
+  # Sum numbers output over all boxes/depth/cohorts
+  # align model output timesteps (days) with recruitment periods (days)
+  totnums <- aggregate(atoutput ~ species + time, data = nums, sum) %>%
+    #mutate(time.days = (time+1)*toutinc) %>% #makes time 0 into days 0->73, etc
+    mutate(time.days = (time)*toutinc) %>% #makes time 1 into days 0->73, etc
+    mutate(yr = ceiling(time.days/365))  # yr 1 is 0:stepsperyr to match recruits yr1
+
+  totnums <- merge(totnums, recruits,
+                   by.x = c("yr", "species"), by.y = c("yr", "Name"),
+                   all.x = TRUE) %>%
+    arrange(time)
+
+  #subset recpars to groups of interest
+  recstart_temp <- rectiming[rectiming$Code %in% totnums$group,]
+
+  totnums$frac_recruit <- 0
+
+  for (irow in 1:nrow(recstart_temp)) {
+    group <- recstart_temp$Code[irow]
+    pick <- which(totnums$group == group)
+
+    recstart <- seq(recstart_temp$recstart[irow],by=365,length.out=nyrs)
+    recstart <- recstart[recstart<max(totnums$Time[pick])]
+    recend <- recstart + recstart_temp$recruit_period[irow]
+    #rec_times <- rbind(rec_times,cbind(group,recstart,recend))
+
+    for (i_rec in 1:length(recstart)) {
+      i_tstart <- which(pick==min(pick[totnums$time.days[pick]>=recstart[i_rec]]))
+      i_tstop <- which(pick==min(pick[totnums$time.days[pick]>=recend[i_rec]]))
+      if (i_rec == length(recstart)) i_tstop <- length(pick)
+      n_t <- 1+i_tstop-i_tstart
+      for (i_t in 1:n_t) {
+        t_temp <- totnums$time.days[pick[i_tstart+i_t-1]]
+        num_temp <- t_temp - recstart[i_rec]
+        if (i_t>1) {
+          if ((recend[i_rec]-t_temp)>toutinc) {
+            num_temp <- toutinc
+          }
+          else {
+            num_temp <- recend[i_rec]-(totnums$time.days[pick[i_tstart+i_t-2]])
+          }
+        }
+        frac_temp <- max(c(0,num_temp /(recend[i_rec]-recstart[i_rec])))
+        if(frac_temp > 1.0) frac_temp = 1.0
+        totnums$frac_recruit[pick[i_tstart+i_t-1]] <- frac_temp
+      }
+    }
+  }
+
+  #G.Fay 1/6/16, expand to num of recruits
+  # Recruit / mg C converted to wet weight in tonnes / redfield ratio of C:N
+  totnums$recruits <- totnums$recruits/k_wetdry/biolprm$redfieldcn
+
+  totnums$group <- recruits$group[match(totnums$species, recruits$Name)]
+  # For all time increments where there
+  totnums$recruits[is.na(totnums$recruits)] <- 0
+
+  totnums$annrecruits <- totnums$recruits
+
+  totnums$recruits <- totnums$annrecruits*totnums$frac_recruit
+
+  # Calculate survivors for each species group
+  totnums$survivors <- totnums$recruits
+
+  totnums$survival <- totnums$survivors
+  # Calculate survival for each group
+  for (group in unique(totnums$group)) {
+    #if(group == "SHD") browser()
+    pick <- which(totnums$group == group)
+    survival_temp <- c(NA,
+      totnums$survivors[pick[-1]]/totnums$atoutput[pick[-length(pick)]])
+
+    # G.Fay 2/21/16  "think" this is what things should be, recruits don't show up
+    # in numbers at age until time step after the recruitment event.
+     survival_temp <- c(
+       (totnums$atoutput[pick[-1]]-
+          totnums$recruits[pick[-1]])/totnums$atoutput[pick[-length(pick)]],NA)
+
+#     survival_temp <- c(
+#       (totnums$atoutput[pick[-1]])/
+#          (totnums$recruits[pick[-1]]+totnums$atoutput[pick[-length(pick)]]),NA)
+
+    survival_temp[survival_temp < 0] <- NA
+    # Use first positive value to replace the initial year and all negative vals
+    firstgood <- which(!is.na(survival_temp))[1]
+
+    survival_temp[1:firstgood] <- survival_temp[firstgood]
+    for(ii in seq_along(survival_temp)) {
+      if (is.na(survival_temp[ii])) {
+        nonzero <- which(which(survival_temp > 0) > ii)
+        if (length(nonzero) == 0) nonzero <- which(survival_temp > 0)
+        survival_temp[ii] <- survival_temp[which.min(abs(nonzero - ii))]
+      }
+    }
+    totnums$survival[pick] <- survival_temp
+   }
+
+  #Calculate Z
+  totnums$Z <- -1 * log(totnums$survival)
+  finaldata <- data.frame("species" = totnums$species,
+    "agecl" = NA, "polygon" = NA, "layer" = NA,
+    "time" = totnums$time, "atoutput" = totnums$Z)
+
+  #return(finaldata)
+#}
+
+```
+
+Modified the `calc_Z` function so it only evaluates species in the input numbers file.
+This creates Z values for all species.
+However, the call to `calc_stage2age` returns only a few of the 36 species from calc_Z so lets check the problem here:
+
+```{r calcstage2agesinppetagain, eval=FALSE}
+
+yoy <- YOY
+nums_data <- survey_testNall
+fgs <- funct.groups
+biolprm <- biol
+runprm = runpar
+
+#calc_stage2age <- function(nums_data, biolprm, yoy, fgs, runprm) {
+
+  # subset the yoy for those species that are included in the fgs file
+  # that are turned on
+  species.code <- fgs$Code
+  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
+  names <- turnedon$Code
+
+  # Figure out the groups that have multiple ages classes in each stage (or
+  # cohort)
+  multiple_ages <- turnedon[turnedon$NumAgeClassSize>1, c(1,4,10)]
+  num_multi_age <- dim(multiple_ages)[1]
+
+  ntimesteps <- length(unique(nums_data$time))
+
+  # For each species with multiple ages, use the calc_Z function to get time
+  # varying Z values and put that all into one dataframe
+  Z.dataframe <- data.frame()
+  #for(i in 1:num_multi_age) {
+    #temp_nums <- nums_data[nums_data$species==multiple_ages$Name[i],]
+    temp_nums <- nums_data[nums_data$species %in% multiple_ages$Name,]
+    temp_Z <- calc_Z(yoy=yoy, nums=temp_nums, fgs=fgs, biolprm=biolprm, toutinc=runprm$toutinc)
+    #Z.dataframe <- rbind(Z.dataframe, temp_Z)
+    Z.dataframe <- temp_Z
+  #}
+
+  # HACK for now, since some Z values are negative, I am replacing them
+  # with a random number
+  randomZ <- runif(length(which(Z.dataframe$atoutput<0)))
+  Z.dataframe[which(Z.dataframe$atoutput<0),c("atoutput")] <- randomZ
+
+  # rename the "atoutput" column, since it is really just the mortality, and
+  # we want to retain it for the next step of merging
+  colnames(Z.dataframe)[6] <- c("Z")
+
+  # merge together nums_data and the output Z.dataframe:
+  new_nums <- merge(nums_data, Z.dataframe[,c("species", "time", "Z")],
+                    by=c("species", "time"), all.x=TRUE)
+
+  # since we have to expand the number of rows for groups with multiple true
+  # age classes, we will loop through species and time, creating a new list
+  # that each element will be a single species in a single time step but
+  # across ages and boxes (these wil all be put together in the end)
+  temp.list <- list()
+
+  #SKG subset turnedon and only loop through species present in the nums input
+  turnedon_sub <- turnedon %>%
+    filter(Name %in% new_nums$species)
+  names <- turnedon_sub$Code
+
+  for(i in 1:length(names)) {
+    # looping all species -- those with or without multiple true ages
+
+    group.i <- turnedon_sub$Name[i]
+    nums_species <- new_nums[new_nums$species==group.i,] # might need to
+    num_ages <- turnedon_sub$NumAgeClassSize[i]
+    sp_times <- sort(unique(nums_species$time))
+    n_sp_tsteps <- length(sp_times)
+    # these last pieces are needed because not all species are present at all times
+
+    # Check if multiple true ages, if not then just save species_nums to list
+    if(num_ages==1) { temp.list[[length(temp.list)+1]] <- nums_species
+    } else if(num_ages>1) {
+      # take out the part of the Z.dataframe for species group i
+      Zvals <- Z.dataframe[Z.dataframe$species==group.i,]
+
+      # create empty list for this species i, each element in the list will be
+      # a different time step
+      list_species <- list()
+
+      for(j in sp_times) { # looping through time
+        # get the Z val for the timestep in question
+        Zval.j <- Zvals$Z[Zvals$time==j]
+
+        # and turn the Z value into a vector of survival values across the
+        # number of true ages for each age class for this species
+        nums_vec <- 1
+        for(k in 1:(num_ages-1)) {
+          nums_vec <- c(nums_vec, exp(-Zval.j*k))
+        }
+        nums_proportion <- nums_vec/sum(nums_vec)
+
+        # take out the species numbers only for time step j
+        nums_sp_subset <- nums_species[nums_species$time==j, ]
+
+        # create an empty list, each element of this list will be a row from
+        # the nums_sp_subset data frame that is split into multiple pieces
+        # to make the atoutput column have the correct dimensions
+        list_ages <- list()
+
+        # loop through all the rows in nums_sp_subset
+        for(l in 1:nrow(nums_sp_subset)) {
+          nums_row <- nums_sp_subset[l,]
+          # create new data frame with same columns, but just more rows to
+          # account for all the true ages
+          new_rows <- data.frame(species=nums_row$species,
+                                 time=nums_row$time,
+                                 agecl=seq((nums_row$agecl*num_ages - num_ages+1),
+                                           nums_row$agecl*num_ages),
+                                 polygon=nums_row$polygon,
+                                 layer=nums_row$layer,
+                                 atoutput=nums_row$atoutput*nums_proportion,
+                                 Z=nums_row$Z
+                                 )
+
+          ### NOTE: the agecl piece is funny -- because now we are making many
+          # more age classes for some... I think this works? But might want
+          # another set of eyes to check my logic
+
+          list_ages[[length(list_ages)+1]] <- new_rows
+        }
+
+        # now combine all the elements of the list for different rows and make
+        # it an element of the list_species (to represent one time step)
+        list_species[[length(list_species)+1]] <- do.call("rbind", list_ages)
+      }
+      temp.list[[length(temp.list)+1]] <- do.call("rbind", list_species)
+    }
+  }
+
+  # Now just create the output
+  finalout_withZ <- do.call("rbind", temp.list)
+
+  ### CHECK:
+  # do we want to remove the column of Z values? I think so
+  finalout <- finalout_withZ[,-7]
+
+  #return(finalout)
+
+#}
+
+```
+
+That was me introducing a bug by filtering incorrectly; fixed it.
+
 These should be true age comps for Bocaccio in annual ages:
@@
 # test with one species--this is still full polygon and layer output
-trueNatage_ss <- trueNatage[trueNatage$species == "Bocaccio_rockfish",]
+trueNatage_ss <- trueNatage[trueNatage$species == species_ss,]
 
@@
 len <- length_census_ss$natlength %>%
-  filter(species == "Bocaccio_rockfish")
+  filter(species == species_ss)
 
@@
 
-truecatchbio_ss <- truecatchbio[truecatchbio$species == "Bocaccio_rockfish",]
+truecatchbio_ss <- truecatchbio[truecatchbio$species == species_ss,]
 
@@
 
-Now (that catch nums are corrected!) we can sample fishery catch for lengths, ages, weights:
+Now (that catch nums are corrected!--sort of; [when converted to catch in weight they are off by a factor of 2 for sardines in this model](https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html)) we can sample fishery catch for lengths, ages, weights, and as long as these are compositions and not absolute values they should be correct:
 
@@
 
-Now we know the fishery output is annual (365 days), and there will only be 100 years. Because we don't want to wait 24 hours for this, we will look at only our three species.
+Now we know the fishery output timestep, and there will only be 100 years. Because we don't want to wait 24 hours for this, we will look at only our three species.
 
@@
 
-This is true catch at age for bocaccio:
+This is true catch at age for bocaccio (in age classes, not true ages):
 
@@
                         catchage$species))
+
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+
+```
+
+Can we split ages for the fishery comps too? True catch at age bocaccio:
+
+```{r cattrueage}
+
+YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))
+
+biol <- load_biolprm(d.name, biol.prm.file)
+
+truecatage_ss <- calc_stage2age(nums_data = catch_testNss,
+                                 biolprm = biol, 
+                                 yoy = YOY,
+                                 fgs = funct.groups, 
+                                 runprm = runpar)
+
+# run sample fish on this for completeness? yes to aggregate
+truecatagesamp_ss <- sample_fish(truecatage_ss, effNhigh)
+
+catchtrueage <- truecatagesamp_ss %>%
+ filter(species == "Bocaccio_rockfish")
+
+# plot it
+
+catageplot <- ggplot(catchtrueage, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name,
+                        catchtrueage$species))
 

### 2019-06-15 FullBoccaccioTruthEx2.Rmd
add bocaccio example; age splitting needs more work

@@
 After corrections were applied to `calc_Z` we were closer to the ballpark in NOBA.
-This demonstrates that the estimated true ages are not working between the cohort ages, they just drop to 0.
-This is consistent with what we observed for NOBA.
-The numbers at true age [don't match atlantis output where we have it for NOBA](https://sgaichas.github.io/poseidon-dev/TestCalcZ.html), so I will do further testing with this function and `calc_stage2age`.
+While the estimated true ages no longer just drop to 0, the patterns here don't seem like a good enough approximation of how age class would decay.
+The numbers at true age for NOBA cod are closer because there are only 2 ages within a stage to split, but they still [don't match atlantis output where we have it for NOBA](https://sgaichas.github.io/poseidon-dev/TestCalcZ.html), so I will do further testing with this `calc_Z` and `calc_stage2age`.
 
@@
 ```
+
+We may need to rethink the annual Z we are estimating to do the approximation.
 
@@
 
-This is true catch at age for bocaccio (in age classes, not true ages):
+This is true catch at (st)age for bocaccio (in age classes, not true ages):
 
@@
 
-```{r cattrueage}
+```{r cattrueage, warning=FALSE, message=FALSE}
 
@@
 
+These still don't look quite right.
+
 Length sample with user specified max length bin (200 cm):
 
-```{r userset-maxlen, echo=TRUE}
+```{r userset-maxlen, echo=TRUE, eval=FALSE}
 
@@
 
+saveRDS(catch_length_census_ss, file.path(d.name, paste0(scenario.name, "catchlength_census_sardhakeboca.rds")))
+
 ```

### 2019-06-15 docs/index.Rmd
add bocaccio example; age splitting needs more work

@@
 
+June 15: while the age splitting is functioning better, applying it to a species with more than 2 ages per age class (CC Bocaccio rockfish) shows that our application of an annual Z estimate may not be the best approach.
+The functions themselves could use further tweaking.
+
+Overall it is advisable to use true age output from more recent Atlantis codebases where possible.
+
+<a href="https://sgaichas.github.io/poseidon-dev/FullBoccaccioTruthEx2.html" target="_blank">Redo Bocaccio test with CCA F scenario, June 15, 2019</a>
+
 ### Rabbit holes that turned out ok

### 2019-07-08 SardinesHakeatlantisom2SStest.Rmd
testing sardine and hake with new CC run

--- 2019-06-07 SardinesOnlyatlantisom2SStest.Rmd
@@
 ---
-title: "Sardines only end-to-end atlantisom test"
+title: "Atlantis to SS for California Current sardines and hake: full fishing and climate scenario"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-Now we test the `atlantisom` to SS workflow with a new California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_run11".
+Now we test the `atlantisom` to SS workflow with a new California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_OA_OFF_222".
 
@@
 
-species_ss <- c("Pacific_sardine")
+species_ss <- c("Pacific_sardine", "Mesopel_M_Fish")
 
@@
 
-if(initCCA) source(here("config/CC2Config.R"))
+if(initCCA) source(here("config/CC3Config.R"))
 
@@
 
-```{r trueB-N, eval=FALSE}
+```{r trueB-N}
 
@@
 
-```{r true-comps, warning=FALSE, message=FALSE, eval=FALSE}
+```{r true-comps, warning=FALSE, message=FALSE}
 
@@
 # save age comps
-saveRDS(numssshigh, file.path(d.name, paste0(scenario.name, "natage_census_sard.rds")))
+saveRDS(numssshigh, file.path(d.name, paste0(scenario.name, "natage_census_sard_hake.rds")))
 
@@
 #save for later use, takes a long time to generate
-saveRDS(length_census_ss, file.path(d.name, paste0(scenario.name, "lengthwt_census_sard.rds")))
+saveRDS(length_census_ss, file.path(d.name, paste0(scenario.name, "lengthwt_census_sard_hake.rds")))
 
@@
 
-Natage <- readRDS(file.path(d.name, paste0(scenario.name, "natage_census_sard.rds")))
+Natage <- readRDS(file.path(d.name, paste0(scenario.name, "natage_census_sard_hake.rds")))
 
@@
 
-lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "lengthwt_census_sard.rds")))
+lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "lengthwt_census_sard_hake.rds")))
 
@@
 # save true total catch in numbers by timestep, need to sum to year
-saveRDS(totcatchN, file.path(d.name, paste0(scenario.name, "catchN_census_sard.rds")))
+saveRDS(totcatchN, file.path(d.name, paste0(scenario.name, "catchN_census_sard_hake.rds")))
 
@@
 # save true catch at age
-saveRDS(catch_numssshigh, file.path(d.name, paste0(scenario.name, "catchage_census_sard.rds")))
+saveRDS(catch_numssshigh, file.path(d.name, paste0(scenario.name, "catchage_census_sard_hake.rds")))
 
@@
 #save for later use, takes a long time to generate
-saveRDS(catch_length_census_ss, file.path(d.name, paste0(scenario.name, "catchlengthwt_census_sard.rds")))
+saveRDS(catch_length_census_ss, file.path(d.name, paste0(scenario.name, "catchlengthwt_census_sard_hake.rds")))
 
@@
 
-catch_numssshigh <- readRDS(file.path(d.name, paste0(scenario.name, "catchage_census_sard.rds")))
+catch_numssshigh <- readRDS(file.path(d.name, paste0(scenario.name, "catchage_census_sard_hake.rds")))
 
@@
 
-totcatchN <- readRDS(file.path(d.name, paste0(scenario.name, "catchN_census_sard.rds")))
+totcatchN <- readRDS(file.path(d.name, paste0(scenario.name, "catchN_census_sard_hake.rds")))
 
@@
 
-catch_numssshigh <- readRDS(file.path(d.name, paste0(scenario.name, "catchage_census_sard.rds")))
-
-catch_lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "catchlengthwt_census_sard.rds")))
+catch_numssshigh <- readRDS(file.path(d.name, paste0(scenario.name, "catchage_census_sard_hake.rds")))
+
+catch_lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "catchlengthwt_census_sard_hake.rds")))
 

### 2019-08-12 CreateStockSynthesis.Rmd
Updates to poseidon

@@
 require(r4ss)
+source("./config/CCConfig.R")
+source("./config/census_spec.R")
 source("./config/sardine_config.R")
@@
 ```{r}
+
+if(!file.exists(file.path(d.name, 
+                          paste0("output", scenario.name, "run_truth.RData")))){
+  #Store all loaded results into an R object
+  results <- run_truth(scenario = scenario.name,
+                     dir = d.name,
+                     file_fgs = functional.groups.file,
+                     file_bgm = box.file,
+                     select_groups = species,
+                     file_init = initial.conditions.file,
+                     file_biolprm = biol.prm.file,
+                     file_runprm = run.prm.file,
+                     verbose = TRUE
+  )
+} else {
+  results <- get(load(file.path(d.name,
+                              paste0("output", scenario.name, "run_truth_all.RData"))))
+}
 results <- get(load(atlantis_output_name))
@@
 species=c(species)
+truth <- results
 
-#Sample all boxes
-boxes <- unique(results$nums$polygon)
+survey_testBall <- create_survey(dat = truth$biomass_ages,
+                                 time = timeall,
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
 
-#Efficiency of the survey
-effic <- data.frame(species=species, efficiency=eff_case)
+# make up a constant 0 cv for testing
+surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
 
-# Assume uniform selectivity
-sel<- data.frame(species=rep(species,length(age_classes)),
-               agecl=age_classes,
-              selex=sel_by_age)
+# call sample_survey_biomass with a bunch of 1000s for weight at age
+# in the code it multiplies atoutput by wtatage/1000 so this allows us to use
+# biomass directly
+wtage <- data.frame(species=rep(survspp, each=10),
+                    agecl=rep(c(1:10),length(survspp)),
+                    wtAtAge=rep(1000.0,length(survspp)*10))
 
-#Sample survey - 3rd timestep simulates a summer survey
-survey_out <- create_survey(dat=results$nums, time=survey_sample_full, species=species, boxes=boxes, effic=effic, selex=sel)
+surveyB_frombio <- sample_survey_biomass(survey_testBall, surv_cv, wtage)
+
+#save for later use, takes a long time to generate
+saveRDS(surveyB_frombio, file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
+
+# this uses result$nums and a new function to get survey index in numbers (abundance)
+
+survey_testNall <- create_survey(dat = truth$nums,
+                                 time = timeall,
+                                 species = survspp,
+                                 boxes = boxall,
+                                 effic = effic1,
+                                 selex = selex1)
+
+# as above, make up a constant 0 cv for testing
+surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+
+surveyN <- sample_survey_numbers(survey_testNall, surv_cv)
+
+#save for later use, takes a long time to generate
+saveRDS(surveyN, file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))
+
+Natage <- readRDS(file.path(d.name, paste0(scenario.name, "natage_census_sard.rds")))
+
+lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "lengthwt_census_sard.rds")))
+
+len <- lengthwt_census_ss$natlength
+
+wtage_ann <- lengthwt_census_ss$muweight %>%
+  filter(time %in% annualmidyear)
 
@@
 #Sample fish for age composition
-age_comp_data <- sample_fish(survey_out, highEffN)
+age_comp_data <- sample_fish(survey_testBall, highEffN)
+boxes <- unique(results$nums$polygon)
 
@@
 
+saveRDS(ss_length_stdsurv, file.path(d.name, paste0(scenario.name, "lengthwt_census_sard.rds")))
+
 #Need to replace with interp function
@@
 #Sample survey biomass
-survObsBiom <- sample_survey_biomass(dat=survey_out,cv=cv,wtAtAge)
+survObsBiom <- sample_survey_biomass(dat=survey_testBall,cv=cv,wtAtAge)
 
@@
 ```{r}
+fish_years <- annualmidyear[fish_years]
 catch_numbers <-  create_fishery_subset(dat = results$catch,
@@
 
+#Use .txt file output biomass instead
+names(results$catch_all)[2] <- "Species"
+catch_total <- filter(results$catch_all, Species==species) %>%
+  filter(time %in% (fish_years+8))
+
 #Test writing CPUE in biomass
@@
                 ts_data = list(survObsBiom$atoutput[fish_years],
-                  catch_total$catch[catch_total$time==fish_years]),
+                  catch_total$catch),
                 CVs = c(CVs$survey,
@@
 
-SS_writedat(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
+SS_writedat(stocksynthesis.data, outfile = paste0(d.name, "/inst/extdata/",model_dir,
 datfile_name),
@@
 #Get biological parameters
-biolprm <- results$biolprm
+
+#Load needed inputs for biological parameters
+source(here("config/CC1Config.R"))
+#biological parameters
+biolprm <- load_biolprm(dir = "C:/Users/chris/Documents/GitHub/atlantisom", file_biolprm = biol.prm.file)
+d.name <- "C:/Users/chris/Documents/GitHub/atlantisom"
+#functional groups
 fgs <- results$fgs
+runprm <- load_runprm(dir = d.name, file_runprm = run.prm.file)
+YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))
+truenums_ss <- results$nums[results$nums$species==species,]
+
+YOY_ss <- YOY %>%
+  select(Time, "SAR.0")
+
+fullresZ <- calc_Z(yoy = YOY_ss,
+                   nums = truenums_ss,
+                   fgs = fgs,
+                   biolprm = biolprm,
+                   toutinc = runprm$toutinc)
+
+meanwt_spp <- ss_length_stdsurv$muweight %>% 
+  filter(time>burnin) %>%
+  group_by(agecl) %>%
+  summarize(meanwt = mean(atoutput))
+sardine.ctl <-r4ss::SS_readctl(paste0("./inst/extdata/",
+model_dir,
+ctlfile_name))
+sardine.ctl <- SS_write_biol(sardine.ctl, biolprm, "SAR", Z = mean(fullresZ$atoutput), wtsage=meanwt_spp)
+
 li_a_use <- biolprm$wl[match(fgs$Code[match(species,fgs$Name)],biolprm$wl[, 1]), 2]/1000
 li_b_use <- biolprm$wl[match(fgs$Code[match(species,fgs$Name)],biolprm$wl[, 1]), 3]
+
+#maturity ogive
+#check what R0 is
 
@@
 
+#Add a call here to plot output
+
 ```

### 2019-08-14 SardinesHakeatlantisom2SStest.Rmd
fixing some calcs and go single species for SS

@@
 
-Sadly, still a problem with the catch correction if it should match catch.txt when comverted to biomass.
-We are now off by about a factor of 2 (catch nums * wtAtAge is half true catch bio from the txt file):
+Sadly, still a problem with the catch correction if it should match catch.txt when converted to biomass.
+We are now off by about a factor of 2 (catch nums * wtAtAge is half true catch bio from the txt file), but this differs slightly by species:
 
@@
 
-mean(sardcatchBcomp$mult)
-
-summary(sardcatchBcomp$mult)
+#mean(sardcatchBcomp$mult)
+
+#summary(sardcatchBcomp$mult)
+Filter(Negate(is.null), tapply(sardcatchBcomp$mult, sardcatchBcomp$species, summary))
 
@@
 
+Now we are going single species. First we generate the inputs for sardine.
+
 Here we are using Christine's specifications, which approximate the acoustic survey index used in the sardine assessment:
@@
 #Set species name
-species <- species_ss #"Pacific_sardine" #change to species_ss to match
+species <- "Pacific_sardine" #change to species_ss if we want both
 
@@
 # CV function
-cv <- data.frame(species="Pacific_sardine", cv=CVs$survey)
+cv <- data.frame(species=species_ss, cv=CVs$survey)
 

### 2019-09-26 SardinesHakeatlantisom2SStest.Rmd
plotting two species, fixing comp writing, should run

@@
 
-Plot true biomass and numbers for "sardines."
-
-```{r plot-true-B-N}
+Plot true biomass and numbers for both "hake" and "sardines."
+
+```{r plot-true-B-N, message=FALSE, warning=FALSE}
 
@@
   gather(species, biomass, -Time) %>%
-  filter(species %in% c("Pacific_sardine"))
+  filter(species %in% c("Mesopel_M_Fish", "Pacific_sardine"))
 
@@
 
-Plot some of the age, length comps, and wtatage
+Plot some of the age, length comps, and wtatage for selected timesteps (after burnin)
 
@@
 
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
+Natagesub <- Natage %>% 
+  filter(time %in% c(155:195))
+
+for(sp in unique(Natage$species)){
+  Natagesubsp <- Natagesub %>%
+    filter(species==sp)
+
+  Natageplot <- ggplot(Natagesubsp, aes(x=agecl, y=atoutput)) +
     geom_point() +
@@
     labs(subtitle = paste(scenario.name,
-                        Natage$species))
+                          sp))
 
@@
   Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+}
 
@@
 len <- lengthwt_census_ss$natlength
+
+lensub <- len %>% 
+  filter(time %in% c(155:195))
+
+for(sp in unique(len$species)){
+  lensubsp <- lensub %>%
+    filter(species==sp)
 
@@
   lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
+
+}
 
@@
 fish_timesteps <- seq(fish_times[5], max(fish_times), by=fstepperyr) #last timestep
-fish_years <- unique(floor(fish_times/fstepperyr)+1)
+#fish_years <- unique(floor(fish_times/fstepperyr)+1) # my original
+fish_years <- unique(floor(fish_times/fstepperyr)) #from Christine's new sardine_config.R
 
@@
 #Years to survey, assuming survey is once per year
-survey_years <- survey_sample_full[(burnin+1):(burnin+nyears)]
+#survey_years <- survey_sample_full[(burnin+1):(burnin+nyears)] # my original
+survey_years <- survey_sample_full[burnin:(burnin+nyears-1)] #from Christine's new sardine_config.R
 
@@
 
+notbins <- c("time", "nsamp")
+
+# fish_length_bins <- as.integer(names(fish_len_comp_final))
+# fish_length_bins <- fish_length_bins[!is.na(fish_length_bins)]
+
+# need to fill empty length bins with 0s to have same bins as survey for SS_write_comps
+missing.lengths <- setdiff(length_bins, names(fish_len_comp_final)[!names(fish_len_comp_final) %in% notbins])
+fish_len_comp_final[as.character(missing.lengths)] <- 0                    # Add them, filled with '0's
+fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")]
+
 # fishery age comps also 5 timesteps per year
@@
 
+# #SS_write_comps breaking because fishery age bins start with 2 not 1; extracting bins from fish file may help?
+# fish_age_bins <- names(fish_age_comp_flat)[!names(fish_age_comp_flat) %in% notbins]
+
+# that leaves an empty column in data file, so instead fill with 0s
+missing.ages <- setdiff(age_bins, names(fish_age_comp_final)[!names(fish_age_comp_final) %in% notbins])
+fish_age_comp_final[missing.ages] <- 0                    # Add them, filled with '0's
+fish_age_comp_final <- fish_age_comp_final[c("time", age_bins, "nsamp")]
+
 comp_list <- list(caal_comp_final,len_comp_final, fish_age_comp_final, fish_len_comp_final)
@@
 
-# below fails with two errors (only gets to first when running)
-
-# Error in `[<-.data.frame`(`*tmp*`, indices, "Seas", value = c(1, 1, 1, : replacement has 49 rows, data has 2045
-
-# Error in `[.data.frame`(comp_matrix[[i]], , as.character(bins[[i]])) : undefined columns selected
-
-# both these errors come from the fishery age and length comps when I step
-# through the function, survey caal and lengths read in fine
+# This now runs by ensuring that survey and fishery compositions have the same bins 
+# (filled with 0s for missing bins in fishery relative to survey)
 

### 2019-09-27 SardinesHakeatlantisom2SStest.Rmd
correct folder name and turn on write SS block

@@
 
-Now we test the `atlantisom` to SS workflow with a new California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_OA_OFF_222".
+Now we test the `atlantisom` to SS workflow with a new California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_OA_OFF_22".
 

### 2019-09-30 SardinesHakeatlantisom2SStest.Rmd
add biological data block, get correct n length bins

@@
 
-Get population parameters for SS from Atlantis
-
-```{r toSS-poppars}
-
-```
-
 ## Write data for Stock Synthesis
@@
 
+#Get correct number of length bins
+# stocksynthesis.data$N_lbins <- length(length_bins)
+
 #Set lbin_method to 1 - this makes the population length bins match the data bins 
@@
 
+## Coming up with dynamic life history parameters
+
+Below we get the weight-length relationship parameters from atlantis, calculate h and R0 from Atlantis $\alpha$ and $\beta$ parameters, use the \code{calc_Z} function to back out natural mortality (M) from Atlantis, and use the survey length-at-age from Atlantis to estimate a growth curve.
+```{r, eval=FALSE}
+#Get biological parameters
+
+#Load needed inputs for biological parameters
+source(here("config/CC3Config.R"))
+#biological parameters
+biolprm <- load_biolprm(dir = d.name, file_biolprm = biol.prm.file)#"C:/Users/chris/Documents/GitHub/atlantisom", 
+#d.name <- "C:/Users/chris/Documents/GitHub/atlantisom"
+#functional groups
+fgs <- truth$fgs
+runprm <- load_runprm(dir = d.name, file_runprm = run.prm.file)
+YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))
+truenums_ss <- truth$nums[results$nums$species==species,]
+
+YOY_ss <- YOY %>%
+  select(Time, "SAR.0")
+
+fullresZ <- calc_Z(yoy = YOY_ss,
+                   nums = truenums_ss,
+                   fgs = fgs,
+                   biolprm = biolprm,
+                   toutinc = runprm$toutinc)
+
+meanwt_spp <- ss_length_stdsurv$muweight %>% 
+  filter(time>burnin) %>%
+  group_by(agecl) %>%
+  summarize(meanwt = mean(atoutput))
+
+ctlfile_name <-  "sardEM_3_3.ctl"
+sardine.ctl <-r4ss::SS_readctl_3.30(paste0("./inst/extdata/",
+model_dir,
+ctlfile_name))
+sardine.ctl <- SS_write_biol(sardine.ctl, biolprm, "SAR", Z = mean(fullresZ$atoutput), wtsage=meanwt_spp)
+
+li_a_use <- biolprm$wl[match(fgs$Code[match(species,fgs$Name)],biolprm$wl[, 1]), 2]/1000
+li_b_use <- biolprm$wl[match(fgs$Code[match(species,fgs$Name)],biolprm$wl[, 1]), 3]
+
+#maturity ogive
+#check what R0 is
+
+len_nonburn <- ss_length_stdsurv$mulen %>%
+  filter(time > burnin*timestep)
+plot(len_nonburn$atoutput~len_nonburn$agecl, col=len_nonburn$time)
+
+length.data <- data.frame("Year"=(len_nonburn$time-survey_sample_time)/timestep+1, length=len_nonburn$atoutput, Weight=NA, Sex="Female", age=as.integer(len_nonburn$agecl))
+
+#require(bbmle)
+#vb_estimates <-sample_fit_vbgf(length.data, 25, 45, 0.4,
+#  0.1, 0.1, 20, 40, 0.05, 0.01, 0.01,
+#  30, 50, 0.6, 0.3, 0.3, 0.5, 10)
+
+```
+
 ## Run Stock Synthesis

### 2019-09-30 docs/index.Rmd
add climate run end to end to index
clarify what works and doesnt

@@
 <a href="https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html" target="_blank">Sardines in the CalCurrent run with F and rec dev scenario, all true and SS data, May 31, 2019, updated June 7</a>
+
+The below generates a readable dat file for input into SS3 for sardines. The control file and biological parameters are still a work in progress.
+
+<a href="https://sgaichas.github.io/poseidon-dev/SardinesHakeatlantisom2SStest.html" target="_blank">Sardines in the CalCurrent run with F, rec dev, and climate scenario, all true and SS data, Sept 30, 2019</a>
 

### 2019-10-01 SardinesHakeatlantisom2SStest.Rmd
clarify what works and doesnt
working ss run

@@
 Below we get the weight-length relationship parameters from atlantis, calculate h and R0 from Atlantis $\alpha$ and $\beta$ parameters, use the \code{calc_Z} function to back out natural mortality (M) from Atlantis, and use the survey length-at-age from Atlantis to estimate a growth curve.
+
+The following is under construction from CreateStockSynthesis.Rmd and not evaluated.
+Existing control file was modified by hand to make the number of recruitment settlement assignments 1 instead of two, with GPattern month area age vector 1 1 1 1.
+
 ```{r, eval=FALSE}
@@
 #Run stock synthesis model 
-run_stocksynthesis(species = species, model_dir = model_dir, show_output = TRUE)
-
-```
+# default runs with admb_options = "-noest"
+#run_stocksynthesis(species = species, model_dir = model_dir, show_output = TRUE)
+
+# for estimation
+run_stocksynthesis(species = species, model_dir = model_dir, admb_options = "", show_output = TRUE)
+
+```

### 2019-10-01 SardinesHakeatlantisom2SStest.Rmd
update index and current run status

@@
 #Get correct number of length bins
-# stocksynthesis.data$N_lbins <- length(length_bins)
+stocksynthesis.data$N_lbins <- length(length_bins)
 
@@
 
-## Coming up with dynamic life history parameters
+## Coming up with dynamic life history parameters (partial from Christine)
 
@@
 
-```{r runSS, eval=FALSE}
+```{r runSS}
 #Run stock synthesis model 

### 2019-10-01 docs/index.Rmd
update index and current run status

--- 2019-06-15 docs/index.Rmd
@@
 <a href="https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html" target="_blank">Sardines in the CalCurrent run with F and rec dev scenario, all true and SS data, May 31, 2019, updated June 7</a>
+
+The below generates a readable dat file for input into SS3 for sardines. The control file and biological parameters still need hand modification, but SS3 now runs.
+
+<a href="https://sgaichas.github.io/poseidon-dev/SardinesHakeatlantisom2SStest.html" target="_blank">Sardines in the CalCurrent run with F, rec dev, and climate scenario, all true and SS data, October 1, 2019</a>
 

### 2019-10-02 SkillAssessInit.Rmd
first cut at comparisons with SS output

---
title: "Comparing assessment results with truth: first pass"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## So your SS model finally ran!

Now we get to plot outputs using `r4ss`

```{r}
library(here)
library(r4ss)
library(atlantisom)
library(dplyr)
library(ggplot2)
```

```{r species-setup}

#Use the right Atlantis output
source(here("config/CC3Config.R"))

#Set species name
species_ss <- "Pacific_sardine" #change to species_ss if we want both

#Directory with SS files
model_dir <- "Sardine_SS_files/"

#Name of SS data file
datfile_name <- "sardEM_3_3.dat"

#ssinput.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/", model_dir, datfile_name))

#Name of SS control file
ctlfile_name <-  "sardEM_3_3.ctl"

# this does not appear to read correctly; has 20 ages for maturity 
# but the model reads it correcty, so don't use for now

#ssinput.ctl <-r4ss::SS_readctl_3.30(paste0("./inst/extdata/", model_dir, ctlfile_name))

```

## Basic SS Plots

```{r ssplots}

replist <- r4ss::SS_output(dir = here(paste0("./inst/extdata/", model_dir)), verbose=TRUE, printstats=TRUE,covar=FALSE)

SS_plots(replist)

replist$timeseries
replist$wtatage
```

## Initial skill assessment

This code is from the multispecies model skill assessment work using Hydra as an operating model and presented in 2015. Original file is hydra_sim_wrapper_modtest.R

```{r skill-functions}

#skill assessment metrics to evaluate model fit

# following from http://heuristically.wordpress.com/2013/07/12/calculate-rmse-and-mae-in-r-and-sas/

# where error = modeled-true or relative error  = modeled/true

# Function that returns Root Mean Squared Error
rmse <- function(error)
{
  sqrt(mean(error^2, na.rm=T))
}

# Function that returns Mean Absolute Error (SAME AS AAE FOR US)
aae <- function(error)
{
  mean(abs(error), na.rm=T)
}

# Function that returns Mean Absolute Error (SAME AS AE FOR US)
ae <- function(error)
{
  mean(error, na.rm=T)
}

# Function that returns Modeling Efficiency MEF
mef <- function(obs, error)
{
  obsavg <- mean(obs, na.rm=T)
  obserr <- obs - obsavg
  (sum(obserr^2)-sum(error^2, na.rm=T))/sum(obserr^2)
}

```

Reading in the truth--these files were generated and saved previously [here](https://sgaichas.github.io/poseidon-dev/SardinesHakeatlantisom2SStest.html).

```{r truth}

# generalized timesteps all models
runpar <- atlantisom::load_runprm(d.name, run.prm.file)
noutsteps <- runpar$tstop/runpar$outputstep
stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
midptyr <- round(median(seq(0,stepperyr)))

trueB <- readRDS(file.path(d.name, paste0(scenario.name,"surveyBcensus.rds")))

trueB_ss <- trueB %>%
  filter(species==species_ss)

trueN <- readRDS(file.path(d.name, paste0(scenario.name,"surveyNcensus.rds")))

trueN_ss <- trueN %>%
  filter(species==species_ss)

Natage <- readRDS(file.path(d.name, paste0(scenario.name, "natage_census_sard_hake.rds")))

Natage_ss <- Natage %>%
  filter(species==species_ss)

lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "lengthwt_census_sard_hake.rds")))

len_ss <- lengthwt_census_ss$natlength %>%
  filter(species==species_ss)

```

```{r ests}

replist$timeseries

```

How does SS3 estimated biomass compare with true?

```{r compareB}

plotB <-ggplot() +
  geom_line(data=trueB_ss, aes(x=time/stepperyr,y=atoutput, color="True B"), 
            alpha = 10/10) +
  geom_point(data=replist$timeseries, aes(x=Yr,y=Bio_all, color="SS3 Est B"),
             alpha = 10/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotB +
  facet_wrap(~species, scales="free")

```

What about recruitment?

```{r compare R}

```

### 2019-10-02 SkillAssessInit.Rmd
add initial comparisons

@@
 library(ggplot2)
+library(ggthemes)
 ```
@@
 Reading in the truth--these files were generated and saved previously [here](https://sgaichas.github.io/poseidon-dev/SardinesHakeatlantisom2SStest.html).
+Recruitment and mortality are detailed [here](https://sgaichas.github.io/poseidon-dev/TestCalcZ2.html).
 
@@
 midptyr <- round(median(seq(0,stepperyr)))
+
+#Load functional groups
+funct.groups <- load_fgs(dir=d.name,
+                         file_fgs = functional.groups.file)
+#Get just the names of active functional groups
+funct.group.names <- funct.groups %>% 
+  filter(IsTurnedOn == 1) %>%
+  select(Name) %>%
+  .$Name
 
@@
 
+YOY <- atlantisom::load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))
+
+# load biolprm in some initialize file?
+biol <- atlantisom::load_biolprm(d.name, biol.prm.file)
+
+# get code matching species name to split YOY file
+code_ss <- funct.groups$Code[which(funct.groups$Name == species_ss)]
+
+# cut to a single species in YOY file
+YOY_ss <- YOY %>%
+  select(Time, paste0(code_ss, ".0"))
+
+# mg carbon converted to wet weight in tonnes
+k_wetdry <- biol$kgw2d / 1000000000
+
+# WARNING only works for CCA because YOY.txt rows 2:end are already in numbers
+# and we are merging out the incorrect and irrelevant YOY row 1 (Time=0)
+# also hardcoded for sardine example
+recnums <- YOY_ss %>%
+  mutate(yr = as.integer(round(YOY_ss$Time)/365)) %>%
+  mutate(recnums = SAR.0/k_wetdry/biol$redfieldcn) %>%
+  filter(yr>0)
+
+file.mort <- file.path(d.name, paste0("output", scenario.name, "Mort.txt"))
+
+mortish <- read.table(file.mort, header = TRUE)
+
+relF_ss <- mortish %>%
+  select(Time, relF = paste0(code_ss, ".F"))
+
 ```
@@
 replist$timeseries
+
+names(replist$timeseries)[19]<-"Fmort"
 
@@
 
-```{r compare R}
+```{r compareR}
+
+plotR <-ggplot() +
+  geom_line(data=recnums, aes(x=yr,y=recnums, color="True R"), 
+            alpha = 10/10) +
+  geom_point(data=replist$timeseries, aes(x=Yr,y=(Recruit_0*1000), color="SS3 Est R"),
+             alpha = 10/10) + 
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotR
 
 ```
+And F (understanding that F in the mort.txt file from Atlantis is relative, just looking for patterns).
+
+```{r compare-F}
+
+plotF <-ggplot() +
+  geom_line(data=relF_ss, aes(x=Time/365, y=relF, color="True F"), 
+            alpha = 10/10) +
+  geom_point(data=replist$timeseries, aes(x=Yr,y=Fmort, color="SS3 Est F"),
+             alpha = 10/10) + 
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotF
+
+```

### 2019-11-22 SardinesHakeatlantisom2SStest.Rmd
added a comment at line 1064 no change to code

@@
 #When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
-stocksynthesis.data$lbin_method <- 1
+stocksynthesis.data$lbin_method <- 1 # this is not reading in, Lbin_method is name, but synthesis can't run if this changes from the original 3 to a 1, so I'm leaving it
 stocksynthesis.data$binwidth <- "#"

### 2019-11-22 Sardines_atlantisom2SS_demo.Rmd

--- 2019-10-01 SardinesHakeatlantisom2SStest.Rmd
@@
 
-Now we test the `atlantisom` to SS workflow with a new California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_OA_OFF_22".
-
-We will go from atlantis model output to SS input in this page for a single species, "sardine", that doesn't require any age class splitting.
-
-As of 31 May 2019, the `run_truth` function has been modified to do the catch numbers fix for legacy models.
+This demonstrates the Atlantis to `atlantisom` to SS3 workflow with the California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_OA_OFF_22".
+
+We will go from Atlantis model output to SS input and run in this page for a single species, "sardine", that doesn't require any age class splitting.
 
@@
 
-initCCA <- TRUE
-initNEUS <- FALSE
-initNOBA <- FALSE
-
-species_ss <- c("Pacific_sardine", "Mesopel_M_Fish")
-
-#function to make a config file? need one for each atlantis run
-
-if(initCCA) source(here("config/CC3Config.R"))
-
-if(initNEUS) source(here("config/NEUSConfig.R"))
-
-if(initNOBA) source(here("config/NOBAConfig.R"))
+species_ss <- c("Pacific_sardine")
+
+source(here("config/CC3Config.R"))
 
@@
 
-Then we will have a function that implements the "survey" that produces the stock assessment inputs and sends those to the SS file writing functions.
-
-## Truth from OM
+Then we will have a function that implements the "survey" that produces the stock assessment inputs and sends those to the SS file writing functions. Here the survey functions are still separate.
+
+## Survey Census
 
@@
 
-Plot true biomass and numbers for both "hake" and "sardines."
+Plot true biomass and numbers for "sardines."
 
@@
   gather(species, biomass, -Time) %>%
-  filter(species %in% c("Mesopel_M_Fish", "Pacific_sardine"))
+  filter(species %in% c("Pacific_sardine"))
 
@@
 wageplot + facet_wrap(c("species"), scales="free_y")
-
-```
-
-Get true population parameters (M in progress, S-R, growth)
-
-```{r true-poppars}
 
@@
 
-Do true catch numbers times weight at age match catch in biomass?
-It was about 16 times too high relative to catch.txt output, and shifted one year back (which is due to the floor function, and an easy fix).
-
-June 6: just corrected the catch nums correction in `run_truth` and I hope it matches now.
-
-```{r catchNtoB}
-
-totcatchN <- readRDS(file.path(d.name, paste0(scenario.name, "catchN_census_sard_hake.rds")))
-
-totcatchann <- totcatchN %>%
-  mutate(yr = floor(time/fstepperyr)+1) %>%
-  group_by(species, yr) %>%
-  summarise(anncatchage = sum(atoutput))
-
-plottotC <- ggplot(totcatchann, aes(x=yr, y=anncatchage)) +
-  geom_point() +
-  theme_tufte() +
-  labs(subtitle = paste("Catch in N ", scenario.name,
-                        totcatchann$species))
-
-plottotC
-
-catch_numssshigh <- readRDS(file.path(d.name, paste0(scenario.name, "catchage_census_sard_hake.rds")))
-
-catch_lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "catchlengthwt_census_sard_hake.rds")))
-
-CwtAtAge <- catch_lengthwt_census_ss$muweight %>%
-  select(species, agecl, time, wtAtAge = atoutput) %>%
-  mutate(wtAtAge = wtAtAge/1000)
-
-catchbioN_ss <- merge(catch_numssshigh,CwtAtAge,by=c("species","agecl", "time"),all.x=T) %>%
-  mutate(biomass = atoutput*wtAtAge/1000) %>%
-  mutate(yr = floor(time/fstepperyr)+1) %>%
-  group_by(species, yr) %>%
-  summarise(anncatchB = sum(biomass))
-
-plotcatch <- ggplot() +
-  geom_line(data=truecatchbio_ss, aes(x=time,y=atoutput, color="true catch bio from txt"),
-            alpha = 10/10) +
-  geom_point(data=catchbioN_ss, aes(x=yr,y=anncatchB, color="true catch bio from N"),
-            alpha = 10/10) +
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-plotcatch +
-  facet_wrap(~species, scales="free")
-
-```
-
-Sadly, still a problem with the catch correction if it should match catch.txt when converted to biomass.
-We are now off by about a factor of 2 (catch nums * wtAtAge is half true catch bio from the txt file), but this differs slightly by species:
-
-```{r catchoff}
-
- sardcatchB <- truecatchbio_ss #%>%
-# now done above
-#   mutate(yr = as.integer(time/365))
-
-sardcatchBcomp <- merge(sardcatchB, catchbioN_ss)
-sardcatchBcomp <- mutate(sardcatchBcomp, mult=atoutput/anncatchB)
-
-#mean(sardcatchBcomp$mult)
-
-#summary(sardcatchBcomp$mult)
-Filter(Negate(is.null), tapply(sardcatchBcomp$mult, sardcatchBcomp$species, summary))
-
-```
-
-## “Data” from OM for EMs
+## Derive “Data” from Atlantis to give to SS3
 
@@
 This could come from other information such as the overlap of actual survey stations with OM polygons, experiments evaluating survey selectivity and efficiency, actual sample-based survey cv, etc.
-
-Now we are going single species. First we generate the inputs for sardine.
 
@@
 
-Get input total catch for SS (here using numbers). SS takes a fishery cv which can be applied here to get a catch time series, or we can input true catch.
-
-PLEASE DON'T USE THIS FOR CCA.
-We have determined above that catch in numbers from legacy atlantis code, even corrected, is still off by a factor of 2.
-We leave the info here because this code can be used for atlantis codebases after December 2015.
-
-```{r toSS-fisheryts}
-
+Get composition inputs for SS (survey and fishery catch at age, survey and fishery lengths, survey and fishey weight at age).
+
+Because catch composition information goes in to the assessment as a proportion, we can use fishery catch at age from this legacy codebase even with absolute catch numbers likely half what they should be overall.
+
+```{r toSS-fisherycomps}
+
+# Survey length comps and wtage done above to get survey ts using nums*wtage approach
+
+#We end up using CAAL for the survey below, so let's generate fishery age comps instead
+effN <- fisheryEffN/fstepperyr
+effN <- data.frame(species=species, effN=effN)
+
+#catch at age each timestep summed over polygons
 # catch at age by area and timestep
@@
 
-if (fstepperyr>1){
-  #need to take all output time steps and sum catch over each year
-  catch_numbers_area <- catch_numbers %>%
-    #assign a temporary year column, aggregate nums over this year
-    mutate(yr = floor(time/fstepperyr)+1) %>%
-    group_by(species, agecl, polygon, yr) %>%
-    summarise(anncatch_area=sum(atoutput))
-
-  # catch in numbers summed over polygon 
-  # and agecl after time has been aggregated into 1 yr
-  catch_total <- catch_numbers_area %>%
-    group_by(yr) %>%
-    summarise(catch=sum(anncatch_area))
-}
-
-# catch in numbers summed over polygon and agecl if output is annual
-if (fstepperyr==1){
-  catch_total <- catch_numbers %>%
-    group_by(time) %>%
-    summarise(catch=sum(atoutput))
-}
-
-# if we want to apply the CV to this we need to write sample_total_catch
-
-#cv <- data.frame(species="Pacific_sardine", cv=CVs$fishery)
-
-#catchObsNum <- sample_total_catch(dat=catch_numbers, cv=cv)
-
-```
-
-Check this against true total catch numbers above. Should be the same (and it is, but both are incorrect in the absolute scale due to the CATCH.nc output problem in older codebases).
-
-```{r compare-totcatch}
-
-compareC <- ggplot() +
-  geom_line(data=totcatchann, aes(x=yr, y=anncatchage, color="census catch N"), 
-            alpha = 10/10) +
-  geom_point(data=catch_total, aes(x=yr,y=catch, color="sample catch N"), 
-            alpha = 10/10) +
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-compareC +
-  facet_wrap(~species, scales="free")
-
-```
-
-Get composition inputs for SS (survey and fishery catch at age, survey and fishery lengths, survey and fishey weight at age).
-
-Because catch composition information goes in to the assessment as a proportion, we can use fishery catch at age from this legacy codebase even with absolute catch numbers likely half what they should be overall.
-
-```{r toSS-fisherycomps}
-
-# Survey length comps and wtage done above to get survey ts using nums*wtage approach
-
-#We end up using CAAL for the survey below, so let's generate fishery age comps instead
-effN <- fisheryEffN/fstepperyr
-effN <- data.frame(species=species, effN=effN)
-
-#catch at age each timestep summed over polygons
 catch_numsss_samp <- sample_fish(catch_numbers, effN)
@@
 
-<!--Can we match catch in weight from catch.txt? No, numbers sampled are 1e-6 compared with actual catch numbers.
-
-```{r compareC-wt, eval=FALSE}
-
-CswtAtAge <- catch_lengthwt_samp$muweight %>%
-  select(species, agecl, time, wtAtAge = atoutput) %>%
-  mutate(wtAtAge = wtAtAge/1000)
-
-catchsampbio_ss <- merge(catch_numsss_samp,CswtAtAge,by=c("species","agecl", "time"),all.x=T) %>%
-  mutate(biomass = atoutput*wtAtAge/1000) %>%
-  mutate(yr = floor(time/fstepperyr)+1) %>%
-  group_by(species, yr) %>%
-  summarise(anncatchBsamp = sum(biomass))
-
-plotcatch <- ggplot() +
-  geom_line(data=truecatchbio_ss, aes(x=time/365,y=atoutput/1000000, color="true catch bio"),
-            alpha = 10/10) +
-  geom_point(data=catchsampbio_ss, aes(x=yr,y=anncatchBsamp, color="sample catch B"), 
-            alpha = 10/10) +
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-plotcatch +
-  facet_wrap(~species, scales="free")
-
-```
--->
-
 ## Write data for Stock Synthesis
@@
 
-Write time series biomass (tons) and catch (tons) data for SS using `SS_write_ts`:
+Write time series biomass (tons) and catch (tons--taken directly from truth without error) data for SS using `SS_write_ts`:
 

### 2019-11-22 docs/index.Rmd
add initial comparisons

@@
 Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
+
+### Demo for NEMoW meeting, December 2019
+
+<a href="https://sgaichas.github.io/poseidon-dev/Sardines_atlantisom2SS_demo.html" target="_blank">Sardines in the CalCurrent run with F, rec dev, and climate scenario, all true and SS data, runs SS3, November 21, 2019</a>
 
@@
 <a href="https://sgaichas.github.io/poseidon-dev/SardinesHakeatlantisom2SStest.html" target="_blank">Sardines in the CalCurrent run with F, rec dev, and climate scenario, all true and SS data, October 1, 2019</a>
+
+So we can start some comparisons of true and model estimated attributes.
+
+<a href="https://sgaichas.github.io/poseidon-dev/SkillAssessInit.html" target="_blank">Compare Atlantis truth with SS3 estimated B, R, and F, October 1, 2019</a>
 

### 2019-12-02 Sardines_atlantisom2SS_demo.Rmd
Add rm and gc commands to free up memory where possible

@@
 ---
-title: "Atlantis to SS for California Current sardines and hake: full fishing and climate scenario"
+title: 'Atlantis to SS for California Current sardines and hake: full fishing and
+  climate scenario'
 author: "Sarah Gaichas and Christine Stawitz"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 library(ggthemes)
+#For testing
+devtools::load_all("C:/Users/chris/Documents/GitHub/atlantisom")
 library(atlantisom)
@@
 source(here("config/CC3Config.R"))
+needed.files <- ls()
+print(unlist(lapply(needed.files, FUN=get)))
 
@@
 
+The below example requires a folder in your project directory named "atlantisom" containing the above files from an Atlantis output.
+
 ```{r get_names, message=FALSE, warning=FALSE}
+#Check for needed files
+
 #Load functional groups
@@
 # biomass directly
-wtage <- data.frame(species=rep(survspp, each=10),
+wtage <- data.frame(species=rep(survspp, each=max(age_classes)),
                     agecl=rep(c(1:10),length(survspp)),
-                    wtAtAge=rep(1000.0,length(survspp)*10))
+                    wtAtAge=rep(1000.0,length(survspp)*max(age_classes)))
 
@@
   filter(species==species_ss)
-
+rm(trueB)
 # read Atlantis output files
@@
   filter(species==species_ss)
+
+rm(trueN)
+gc()
 
@@
 }
-
+rm(Natage)
+gc()
 ```
@@
 
+rm(len)
 ```
@@
   filter(time %in% annualmidyear)
+
+rm(lengthwt_census_ss)
+gc()
 
@@
 
+rm(truecatchbio)
+gc()
 # note: time output of this file is in days
@@
 
-Get true catch at age and other comps (may not be different from input for our project). First see if catch output is at same timescale as population by checking toutfinc: `r fstepperyr`
+Get true catch at age and other comps (may not be different from input for our project). First see if catch output is at same timescale as population by checking `fstepperyr`.
 
 ```{r true-catage-lf-wtage, eval=FALSE}
-
+print(fstepperyr)
 # get survey nums with full (no) selectivity
@@
 
+rm(catch_numssshigh)
+gc()
+
 catageplot <- ggplot(catchage, aes(x=agecl, y=atoutput)) +
@@
 #Age classes of your species
-age_classes <- 1:10
+age_classes <- 1:12
 
@@
 # check against survey with truth$biomass_ages output
-wtage1 <- data.frame(species=rep(species_ss, each=10),
-                    agecl=rep(c(1:10),length(species_ss)),
-                    wtAtAge=rep(1000.0,length(species_ss)*10))
+wtage1 <- data.frame(species=rep(species_ss, each=max(age_classes)),
+                    agecl=rep(c(age_classes),length(species_ss)),
+                    wtAtAge=rep(1000.0,length(species_ss)*max(age_classes)))
 
@@
 
+rm(catch_numbers)
+gc()
 #Get weights

### 2019-12-11 TestCalcZ2.Rmd
estimating M and running into F problems

@@
 We can probably conclude that the `calc_Z` function is working as correctly as possible now, given that it is an approximation no matter what, and the differences we see here are likley from temporal resolution of the Z estimate.
+
+So what is M?
+This is an approximate annual M as Z-F from mort.txt (which assumes F is not rescaled!).
+If we need annual and age specific M then we need a new Z calc function that is cohort specific I think.
+
+```{r isthisM, message=FALSE, warning=FALSE}
+
+annM1 <- merge(annZ1, relF_ss, all.x = T) %>%
+  mutate(M = Z-relF)
+
+annZ2_Time <- annZ2 %>%
+  mutate(Time = yr*365)
+
+annM2 <- merge(annZ2_Time, relF_ss) %>%
+  mutate(M = Z-relF)
+
+plotM <- ggplot() +
+  geom_line(data=Zish, aes(x=Time/365, y=relZ, color="mort.txt Z")) +
+  geom_line(data=Zish, aes(x=Time/365, y=relM, color="mort.txt M")) +
+  geom_point(data=annM1, aes(x=Time/365, y=M, color="M = endyr Z-F")) +
+  geom_point(data=annM2, aes(x=yr, y=M, color="M = sum timestep Z-F")) +
+  theme_tufte() +
+  theme(legend.position = "bottom", legend.box = "horizontal") +
+  scale_color_discrete(NULL) +
+  guides(color = guide_legend(nrow = 1)) +
+  labs(subtitle = paste(scenario.name, species_ss))
+
+plotM + ylim(0, 2.0)
+
+```
+
+So the problem here is we get negative M if we use our calculated Z from annual numbers at age or timestep numbers at age and subtract the F from mort.txt.
+(Negative M does not show up as points above because I restrict the y axis scale to 0-2.0.) The F from mort.txt is actually higher than our estimated Z.
+So Atlantis must rescale F too???
+
+Need to investigate backing out F...
 

### 2019-12-11 docs/index.Rmd
estimating M and running into F problems

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/TestCalcZ2.html" target="_blank">Test Z calc and compare with output mort CCA and NOBA - June 7, 2019, updated June 13</a>
+<a href="https://sgaichas.github.io/poseidon-dev/TestCalcZ2.html" target="_blank">Test Z calc and compare with output mort CCA and NOBA - June 7, 2019, updated June 13 and December 11 with M</a>
+
+Dec 11: Our calculated Z is less than the input F. Does Atlantis rescale F?
 

### 2020-01-13 Sardines_atlantisom2SS_demo.Rmd
demo used at NEMOW commenting out one hardcode file path

@@
 #For testing
-devtools::load_all("C:/Users/chris/Documents/GitHub/atlantisom")
+#devtools::load_all("C:/Users/chris/Documents/GitHub/atlantisom")
 library(atlantisom)

### 2020-01-13 TestCalcFandM.Rmd
initial commit testing calc F and calc M

--- 2019-12-11 TestCalcZ2.Rmd
@@
 ---
-title: "Getting mortality from atlantisom: testing calc_Z function II"
+title: "Getting mortality from atlantisom: backing out F and M from Z"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
-On this page we demonstrate use of atlantisom on both California Current (CCA) and Norwegian Barents Sea Atlantis (NOBA) output files to test function `calc_Z` calculating total mortality.
-This is used both to derive "natural mortality" for comparison with stock assessments and to split age classes into true ages using the function `calc_stage2age`.
-
-NOBA has output files in true ages that we can compare results with. These files are in the folder NOBAwithAnnAgeOutput (not kept on github due to large filesize).
-
-We can also compare both models with the output[scenario.name]Mort.txt file, which reports annual M and F. The user's manual notes: 
+On this page we demonstrate use of atlantisom on both California Current (CCA) and Norwegian Barents Sea Atlantis (NOBA) output files to test function new functions that use `calc_Z` calculating total mortality to back out F and M.
+These are used initially for comparison with stock assessment model fishing mortality outputs, and to parameterize stock assessment natural mortality.
+
+[Previous comparisons](https://sgaichas.github.io/poseidon-dev/TestCalcZ2.html) of estimated total mortality with the output[scenario.name]Mort.txt file demonstrated what the Atlantis manual already tels us:
 >This file is currently only useful for looking at relative M vs F values for a species, as itdoes not give accurate mortalities
 
-So we can assume that the scale of Z from this file will not be the same as our calculation, but it should be useful for splitting out F to get M, and for comparing overall patterns.
-
-We will first use CCA here because the contrasting F scenario has been implemented. We will test with both sardines (M needed, but not age splitting) and hake (M and age splitting needed).
+We can assumed that the scale of Z from this file will not be the same as our calculation, but since both F and M are rescaled, even from input F values, we need to back out the realized F to subtract from this Z to get M.
+This approach seems simpler than trying to estimate M directly using predation.
+
+We will first use CCA here because the contrasting F scenario has been implemented. We will test with sardines only first.
 
@@
 initCCA <- TRUE
-initNEUS <- FALSE
 initNOBA <- FALSE
@@
 
-if(initCCA) source(here("config/CC2Config.R"))
-
-if(initNEUS) source(here("config/NEUSConfig.R"))
+if(initCCA) source(here("config/CC3Config.R"))
 
@@
 
-<!-- not sure we need catch bio for this
-```{r loadcatchb}
-
-truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)
-
-truecatchbio_ss <- truecatchbio[truecatchbio$species == species_ss,]
-# note: time output of this file is in days
-# what model timestep is this output matching?
-# rescale catch in biomass time from days to years
-truecatchbio_ss$time <- as.integer(truecatchbio_ss$time/365)
-
-```
--->
-
-## Test calc Z function
+## Estimate total mortality with calc Z function
 

### 2020-01-13 WrapperDevelopment.Rmd
initial commit wrapper development

---
title: "Develop atlantisom wrappers"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Too many steps!

We can run atlantisom with many functions to get stock assessment model inputs.

To make life easier, here are some wrapper functions to do the basic things we want.

Basic desired functionality:

1. initialize: locate model output (function to make a config file?), read in basic run parameters and species names, run_truth for all once and save output (optionally, run_truth for only one species?)

1. get assessment input data: configure survey, run all survey functions, configure fishery, run all fishery functions (how to save output?)

1. write input data for specific model: wrap functions in development for SS3, add functions for other models (ASAP, multispecies models)

### 2020-01-13 WrapperDevelopment.Rmd
add details

@@
 
-1.
-initialize: locate model output (function to make a config file?), read in basic run parameters and species names, run_truth for all once and save output (optionally, run_truth for only one species?)
+1. initialize: locate model output (function to make a config file?), read in basic run parameters and species names, run_truth for all once and save output
+
+  + User to make up front decision--want all species or just a subset?
+  + May want multiple species that need to be split in a later step for models
 
@@
 
+  + Based on decision above, save interim steps for use later
+  + Need a switch for single vs multispecies estimation models?
+
 1. write input data for specific model: wrap functions in development for SS3, add functions for other models (ASAP, multispecies models)
+
+## Cumbersome giant files eat all the memory!
+
+Implement Christine's memory saving things from the NEMOW demo

### 2020-01-13 TestCalcFandM.Rmd
F biomass vs numbers est, in progress for NOBA

@@
 
-This section tests the `atlantisom::calc_Z()` function step by step.
-Ultimately this output goes to `calc_stage2age` to produce annual numbers at true age, which we can compare with the ANNAGEBIO.nc output produced by NOBA.
-For now we will just compare to the Mort.txt output.
+Trying to wrap these steps into something simpler.
 
@@
 
-### True numbers total (annual)
+### Calculating Z and F at different levels of aggregation
 
-This gives us the true total numbers on an annual basis. We can compare this with age specific outputs.
-
-```{r surveyNbased, eval=FALSE}
-
-# this uses result$nums and a new function to get survey index in numbers (abundance)
-
-survey_testNall <- create_survey(dat = truth$nums,
-                                 time = timeall,
-                                 species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-# save this, needed as input to calc_Z?
-saveRDS(survey_testNall, file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))
-
-# as above, make up a constant 0 cv for testing
-surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
-
-surveyN <- sample_survey_numbers(survey_testNall, surv_cv)
-
-#save for later use, takes a long time to generate
-saveRDS(surveyN, file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))
-
-```
-
-```{r survNtrue-plot}
-
-surveyN <- readRDS(file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))
-
-surveyN_ss <- surveyN[surveyN$species == species_ss,]
-
-plotN <-ggplot() +
-  geom_line(data=surveyN_ss, aes(x=time/stepperyr,y=atoutput, color="survey census N"), 
-            alpha = 10/10) +
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-plotN +
-  facet_wrap(~species, scales="free")
-
-```
-
-### What level of aggregation works for calc_Z?
-
-The atlantisom function `calc_stage2age` calls the `calc_Z` function, applying a total mortality rate to estimate the numbers at true age within an age class, but it is unclear whether to use this on the aggregated numbers or the full resolution numbers at age.
-Here we test outputs of Z for several levels of aggregation:
+Here we test outputs of Z for several levels of aggregation using files generated previously:
 
@@
 # numbers at agecl aggregated over layers and polygons
-Natage_ss <- readRDS(file.path(d.name, paste0(scenario.name, "natage_census_sard.rds")))
+Natage_ss <- readRDS(file.path(d.name, paste0(scenario.name, "natage_census_sard_hake.rds")))
+
+Natage_ss <- Natage_ss[Natage_ss$species == species_ss,]
 
@@
 
-Fixed a bug where grep was used if a # appeared. This ran for the survey resolution numbers. I then went back and ran the two other levels above after fixing `calc_Z.`
+### What is F at each timestep?
+A relative annual F is found in outputMort.txt! Sadly, it is wrong.
 
-Both true full resolution numbers and survey resolution numbers (aggregated over polygon) returned nonzero Z values.
-As with previous the NOBA test, they are identical aside from eight values with no apparent pattern (rows 100, 105, 180, 200, 220, 240, 355, 400) out of 500 that differ at e-16.
-Same story with difference from sample resolution; this does not matter to output.
-
-```{r differentZ}
-fullresZ$atoutput - surveyresZ$atoutput
-
-fullresZ$atoutput - sampleresZ$atoutput
-```
-
-I would be happer if the differences were all exactly 0, but I suppose we can attribute that to rounding error.
-
-### What is F at each timestep?
-A relative annual F is found in outputMort.txt!
 ```{r getrelF}
@@
 
+Zish <- merge(relF_ss,relM_ss) %>%
+  mutate(relZ = relF + relM)
+
 plotF <- ggplot(data=relF_ss, aes(x=Time/365, y=relF)) +
@@
 
-This is consistent with the scenario implemented. So far so good.
+### Estimate F from catch and biomass at different levels of aggregation
 
-This probably depends on how the model is set up, but for models forced with F by species we ought to be able to read this out from an input file.
-If it is a daily rate then we mulitply by toutinc to get the rate at each of these output timestep Z values.
+Ideally we would want this calculation to be in numbers of individuals to calculate F.
+However we do not have reliable catch in numbers output from catch.nc for legacy models.
+Therefore I will build a switch in to check for legacy models to do this calculation in biomass, otherwise the default calculation should use catch.nc catch in numbers compared to the numbers (in the previous ouput timestep) in nums from truth.
 
-### Quick test to see if Z is right?
-Compare with outputMort.txt!
-```{r compZ}
+For CCA, Catch in biomass at each catch output timestep is read in from catch.txt:
 
-Zish <- merge(relF_ss,relM_ss) %>%
-  mutate(relZ = relF + relM)
+```{r catchbio}
 
-plotZ <- ggplot() +
-  geom_line(data=Zish, aes(x=Time/365, y=relZ)) +
-  geom_point(data=fullresZ, aes(x=(time*73)/365, y=atoutput)) +
+truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)
+
+truecatchbio_ss <- truecatchbio[truecatchbio$species == species_ss,]
+
+rm(truecatchbio)
+```
+
+Total biomass at each output timestep, or should this be average biomass?
+Or initial biomass?
+Isaac used initial biomass for the year from the biomass index and total catch for the year so comparing catch at day 365 to biomass at day 0 for that year.
+
+Since we have only annual B then we can just read in true B from the biomass index:
+
+```{r totbio}
+
+atBtxt2 <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)
+groupslookup <- funct.groups %>%
+  filter(IsTurnedOn > 0)
+
+atBtxt2tidy <- atBtxt2 %>%
+  select(Time:DIN) %>%
+  #select(Time, FPL:DIN) %>%
+  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
+  gather(species, biomass, -Time) %>%
+  filter(species %in% species_ss)
+
+```
+
+Output time units match in these 2 files (days). I want to recode the time units so that year 1 is time 0 for biomass and time 365 for catch, then get annbioF:
+
+```{r annbioF}
+
+annbio <- atBtxt2tidy %>%
+  mutate(yr = (Time/365)+1)
+
+anncatch <- truecatchbio_ss %>%
+  mutate(yr = time/365)
+
+annbioF <- merge(annbio, anncatch) %>%
+  mutate(annbioF = atoutput/biomass) %>%
+  arrange(yr)
+
+```
+
+This annual biomass based estimate of F is scaled down from the input F (show in points in the plot).
+
+```{r F_estplot}
+
+plotannbioF <- ggplot(data=annbioF, aes(x=yr, y=annbioF)) +
+  geom_line() +
   theme_tufte() +
@@
 
-plotZ
-```
-
-After several iterations of debugging, we are still not matching the pattern, but improved in that it isn't a constant punctuated by occasional high.
-Variation still doesn't look anything like the txt file either.
-The 5x yr timesteps are probably messing this up?
-
-I've commented out the detailed debugging block below, but check out the rmd file if you are interested in the sausage making.
-Changes have been implemented in `calc_Z` so that it now apportions annual recruitment to timesteps and recruitment is correctly aligned with truth nums output.
-
-However, it is not clear that subtracting off "recruitment" that has not been subject to mortality is appropriate when the output numbers at each timestep have been subject to mortality?
-I guess that was the input in each timestep as best we know it, but this is at best an approximation.
-<!--
-```{r calczsnippet, eval=FALSE}
-
-#corrections listed here have been applied in calc_Z; this code not run
-
-# test with one species at survey scale (layers aggregated, not time or polygon)
-yoy <- YOY_ss
-nums <- survey_testN_ss
-fgs <- funct.groups
-biolprm <- biol
-toutinc <- runpar$toutinc
-
-# everything below the function definition line
-  # subset the yoy for species included in the fgs file
-  # that are turned on
-  # colnames of the recruit data are "Time" and a column for each
-  # species where the name is the functional group with an additional .0
-  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
-  recruits <- yoy[, colnames(yoy) %in% c("Time", paste0(turnedon$Code, ".0"))]
-
-  # mg carbon converted to wet weight in tonnes
-  k_wetdry <- biolprm$kgw2d / 1000000000
-  # Sum of structural and reserve nitrogen (KWSR_RN + KWSR_SN)
-  nitro <- merge(biolprm$kswr, biolprm$kwrr, by = "1")
-  nitro$sum <- apply(nitro[, 2:3], 1, sum)
-
-  # legacy: If output is from legacy code there will be an error in the
-  # yoy data, where the first row is in a different unit.
-  # yoy.txt is the biomass in tonnes per spawning event summed over the total
-  # model domain.
-  # The first row (< Nov/Dec 2015) is stored as biomass and the remaining rows
-  # are stored in numbers, must convert the entire matrix to biomass
-  # Check if legacy code and if so convert the numbers to biomass
-
-  # G.Fay 2/21/16 : changed yoy to recruits in below loop.
-  if (abs(recruits[1, 2] / recruits[2, 2]) > 10) {
-    recruits[2:NROW(recruits), 2:NCOL(recruits)] <- recruits[2:NROW(recruits), 2:NCOL(recruits)] *
-    nitro[match(gsub(".0", "", colnames(recruits)[-1]), nitro[, 1]), "sum"]
-  }
-
-  # Wide to long
-  recruits <- reshape(data = recruits, direction = "long",
-    varying = colnames(recruits)[-1],
-    v.names = "recruits",
-    times = colnames(recruits)[-1],
-    timevar = "group")
-  rownames(recruits) <- 1:NROW(recruits)
-  recruits <- recruits[, -which(colnames(recruits) == "id")]
-  # Switch from species code to species
-  recruits$group <- gsub("\\.0", "", recruits$group)
-  recruits <- merge(recruits, fgs[, c("Code", "Name")],
-    by.x = "group", by.y = "Code")
-
-  # merge recruits with strucn and resn of recruits from biol.prm file
-  recruits <- merge(recruits, nitro[, c(1, 4)],
-    by.x = "group", by.y = "1", all.x = TRUE, all.y = FALSE)
-  colnames(recruits)[which(colnames(recruits) == "recruits")] <- "recruitsbio"
-  # Get recruits in numbers rather than biomass
-  # these match rows 2:end of the legacy file, so good
-  # but is this 1000s or numbers? not real numbers, Gavin corrects below
-  recruits$recruits <- recruits$recruitsbio / recruits$sum
-  recruits$yr <- as.integer(round(recruits$Time)/365)
-
-  #Isaac determined that yr 0 (Time 0) in the YOY file is not real, don't use it
-  #below assigns year 1 to times 0:stepperyr, all good output timesteps in truth$nums
-
-  # G.Fay 2/21/16
-  # UGLY code below tries to align fraction of annual yoy with timing of recruitment
-  # values in YOY.txt are total YOY that year waiting to recruit.
-  # seems to get rid of most of 'issues' - some v.minor survival >1,
-  #perhaps due to averaging of survival over toutinc days
-  nyrs <- ceiling(max(yoy$Time)/365)
-  times <- unique(yoy$Time)
-
-  # SKG June 2019
-  # need to match codes below, this mismatches when biolprm not in same order!!
-  #recstart_temp <- biolprm$recruit_time
-  #recstart_temp[,2] <- biolprm$time_spawn[-(grep('#',
-  #                     biolprm$time_spawn[,1])),2] + biolprm$recruit_time[,2]
-  # was this grep to get rid of a species with #XXX? breaks when there are none
-  #recstart_temp[,2] <- biolprm$time_spawn[,2] + biolprm$recruit_time[,2]
-
-  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%turnedon$Code,] #bug added codes with no YOY output
-  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%recruits$group,]
-
-  # need spawn_period?? if so read out in load_biolprm and merge it here too
-
-  # rectiming is a dataframe with species code, time_spawn (day of year), 
-  # recruit_time (number of days), and recruit period (number of days). 
-  # we use these to calculate recstart (day of year) and recend (day of year)
-  rectiming <- merge(biolprm$time_spawn, biolprm$recruit_time, by = 1)
-  rectiming <- merge(rectiming, biolprm$recruit_period, by = 1)
-  names(rectiming) <- c("Code", "time_spawn", "recruit_time", "recruit_period")
-  rectiming <- rectiming %>%
-    mutate(recstart = time_spawn + recruit_time) %>% # possibly + spawn_period
-    mutate(recend = recstart + recruit_period)
-
-  #subsets to groups of interest
-  recstart_temp <- rectiming[rectiming$Code %in% recruits$group,]
-
-  # do we need this fraction calculation if YOY is annual snapshot?
-  # (check output timing of YOY file with tsumout, available in run.prm)
-  # can we not just allocate the full YOY to the timestep where they come in?
-  # if recstart to recend spans multiple toutinc timesteps then need fraction
-
-  # align model output timesteps (days) with recruitment periods (days)
-  #numstime.days <- unique(nums$time)*toutinc
-  # Sum numbers output over all boxes/depth/cohorts
-  totnums <- aggregate(atoutput ~ species + time, data = nums, sum) %>%
-    mutate(time.days = (time+1)*toutinc) %>% #makes time 0 into days 0->73, etc
-    mutate(yr = ceiling(time.days/365))  # yr 1 is 0:stepsperyr to match recruits yr1
-    #mutate(yr = floor(time.days/365)) # includes 0 value, yr 1 is
-
-  totnums <- merge(totnums, recruits,
-    by.x = c("yr", "species"), by.y = c("yr", "Name"),
-    all.x = TRUE) %>%
-    arrange(time)
-
-  totnums$frac_recruit <- 0
-
-  for (irow in 1:nrow(recstart_temp)) {
-    group <- recstart_temp$Code[irow]
-    pick <- which(totnums$group == group)
-
-    recstart <- seq(recstart_temp$recstart[irow],by=365,length.out=nyrs)
-    recstart <- recstart[recstart<max(totnums$Time[pick])]
-    recend <- recstart + recstart_temp$recend[irow]
-    #rec_times <- rbind(rec_times,cbind(group,recstart,recend))
-
-    for (i_rec in 1:length(recstart)) {
-      i_tstart <- which(pick==min(pick[totnums$time.days[pick]>=recstart[i_rec]]))
-      i_tstop <- which(pick==min(pick[totnums$time.days[pick]>=recend[i_rec]]))
-      if (i_rec == length(recstart)) i_tstop <- length(pick)
-      n_t <- 1+i_tstop-i_tstart
-      for (i_t in 1:n_t) {
-        t_temp <- totnums$time.days[pick[i_tstart+i_t-1]]
-        num_temp <- t_temp - recstart[i_rec]
-        if (i_t>1) {
-          if ((recend[i_rec]-t_temp)>toutinc) {
-            num_temp <- toutinc
-          }
-          else {
-            num_temp <- recend[i_rec]-(totnums$time.days[pick[i_tstart+i_t-2]])
-          }
-        }
-        frac_temp <- max(c(0,num_temp /(recend[i_rec]-recstart[i_rec])))
-        totnums$frac_recruit[pick[i_tstart+i_t-1]] <- frac_temp
-      }
-    }
-  }
-
-# Time in YOY does not match recstart-recend period still after this loop--fixed
-
-  # match "Time" of the young of the year with the time-step periodicity
-  # listed in the run.prm or run.xml file
-  #recruits$Time <- recruits$Time / toutinc
-
-# SKG did this above already in the following line?
-  # Get recruits in numbers rather than biomass
-  #recruits$recruits <- recruits$recruitsbio / recruits$sum
-
-  #G.Fay 1/6/16, expand to num of recruits
-  # Recruit / mg C converted to wet weight in tonnes / redfield ratio of C:N
-  totnums$recruits <- totnums$recruits/k_wetdry/biolprm$redfieldcn
-
-  # Combine recruits and numbers
-  # Only pull recruits from the yearly time step, where the
-  # yearly time step matches the time step
-  # totnums <- merge(totnums, recruits,
-  #   by.x = c("time", "species"), by.y = c("Time", "Name"),
-  #   all.x = TRUE, all.y = FALSE)
-  totnums$group <- recruits$group[match(totnums$species, recruits$Name)]
-  # For all time increments where there
-  totnums$recruits[is.na(totnums$recruits)] <- 0
-
-   # Make sure time is in order, did above
-  #totnums <- totnums[order(totnums$species, totnums$time), ]
-  totnums$annrecruits <- totnums$recruits
-
-  totnums$recruits <- totnums$annrecruits*totnums$frac_recruit
-
-  # Calculate survivors for each species group
-  totnums$survivors <- totnums$recruits
-
-  totnums$survival <- totnums$survivors
-  # Calculate survival for each group
-  for (group in unique(totnums$group)) {
-    #if(group == "SHD") browser()
-    pick <- which(totnums$group == group)
-    survival_temp <- c(NA,
-      totnums$survivors[pick[-1]]/totnums$atoutput[pick[-length(pick)]])
-
-    # G.Fay 2/21/16  "think" this is what things should be, recruits don't show up
-    # in numbers at age until time step after the recruitment event.
-     survival_temp <- c(
-       (totnums$atoutput[pick[-1]]-
-          totnums$recruits[pick[-1]])/totnums$atoutput[pick[-length(pick)]],NA)
-
-#     survival_temp <- c(
-#       (totnums$atoutput[pick[-1]])/
-#          (totnums$recruits[pick[-1]]+totnums$atoutput[pick[-length(pick)]]),NA)
-
-    survival_temp[survival_temp < 0] <- NA
-    # Use first positive value to replace the initial year and all negative vals
-    firstgood <- which(!is.na(survival_temp))[1]
-
-    survival_temp[1:firstgood] <- survival_temp[firstgood]
-    for(ii in seq_along(survival_temp)) {
-      if (is.na(survival_temp[ii])) {
-        nonzero <- which(which(survival_temp > 0) > ii)
-        if (length(nonzero) == 0) nonzero <- which(survival_temp > 0)
-        survival_temp[ii] <- survival_temp[which.min(abs(nonzero - ii))]
-      }
-    }
-    totnums$survival[pick] <- survival_temp
-   }
-
-  #Calculate Z
-  totnums$Z <- -1 * log(totnums$survival)
-  finaldata <- data.frame("species" = totnums$species,
-    "agecl" = NA, "polygon" = NA, "layer" = NA,
-    "time" = totnums$time, "atoutput" = totnums$Z)
-
-surveyresZ <- finaldata
+plotannbioF + geom_point(data=relF_ss, aes(x=Time/365, y=relF))
 
 ```
--->
 
@@
 
-plotZ <- ggplot() +
-  geom_line(data=Zish, aes(x=Time/365, y=relZ)) +
-  geom_point(data=annZ1, aes(x=Time/365, y=Z)) +
-  theme_tufte() +
-  labs(subtitle = paste(scenario.name, species_ss))
-
-plotZ + ylim(0, 2.0)
-
 ```
-
-This looks better; we are starting to see a pattern.
 
@@
 
-plotZ <- ggplot() +
-  geom_line(data=Zish, aes(x=Time/365, y=relZ)) +
-  geom_point(data=annZ2, aes(x=yr, y=Z)) +
-  theme_tufte() +
-  labs(subtitle = paste(scenario.name, species_ss))
-
-plotZ + ylim(0, 2.0)
-
 ```
-
-Also looks better. Not sure which is more correct.
 
@@
 
-So what is M?
-This is an approximate annual M as Z-F from mort.txt (which assumes F is not rescaled!).
-If we need annual and age specific M then we need a new Z calc function that is cohort specific I think.
+## Now do the same thing with F
+
+F based on numbers would have this option, but we cant for CC because we only have annual catch in biomass output. Subtract only from annual (endyr) Z.
+
+## Now get M from Z-F
+
+So what is M? This is an approximate annual M as Z-F from the annual estimate of F derived above.
 
@@
 
-annM1 <- merge(annZ1, relF_ss, all.x = T) %>%
-  mutate(M = Z-relF)
+annbioF <- select(annbioF, yr, annbioF)
+
+annM1 <- merge(annZ1, annbioF, all.x = T) %>%
+  mutate(M = Z-annbioF)
 
@@
 
-annM2 <- merge(annZ2_Time, relF_ss) %>%
-  mutate(M = Z-relF)
+annM2 <- merge(annZ2_Time, annbioF) %>%
+  mutate(M = Z-annbioF)
 
@@
 
-So the problem here is we get negative M if we use our calculated Z from annual numbers at age or timestep numbers at age and subtract the F from mort.txt.
-(Negative M does not show up as points above because I restrict the y axis scale to 0-2.0.) The F from mort.txt is actually higher than our estimated Z.
-So Atlantis must rescale F too???
+And our estimated M time series:
 
-Need to investigate backing out F...
+```{r estM}
 
-Let's see how it performs with another model.
+plotestM <- ggplot(data=annM2, aes(x=yr, y=M)) +
+  geom_line() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
 
-### Test calc_Z NOBA
+plotestM + ylim(0, 0.5)
 
-Might be a good time to switch back to NOBA for comparison with true annual ages.
+```
+
+We still have a mismatch between Z in numbers and F calculated based on biomass, with some negative M values resulting from this calculation. It has improved somewhat.
+
+Is there a mismatch between biomass based F and numbers based Z?
+
+## Test calc_Z NOBA
+
+Let's try with NOBA and compare biomss and numbers based F directly. I'll use NOBA herring as a species similar to sardines.
 
@@
 initCCA <- FALSE
-initNEUS <- FALSE
 initNOBA <- TRUE
@@
 
-species_ss <- "North_atl_cod"
+species_ss <- "Norwegian_ssh"
 
@@
 
-This Z is a bad match. The Z values in NOBA's mort.txt are huge. Not sure what to make of this.
+This Z is a bad match. But in a different way from cod where Z was huge; here the mort.txt Z is smaller than the estimated Z. This makes me question the Z calculation again.
 
-The above figure is now the best it is going to get after a deeper dive back into `calc_Z` for NOBA before we proceed to true ages.
-Found one bug hidden by sardine's equal recruit_period and recend, another where the recruitment period is all within one output timestep that resulted in proportions over 1 (inflating the recruitment and then subtracting it off).
-Further, it appears that timesteps for recruitment were still not correctly aligned, so trying a different alignment that leaves out the timestep 0 from the truth.
-Not sure the latter is correct, or if we need to rethink the calculation of survival based on when recruits actually enter the model.
+So does Atlantis scale M *up* from the input as well?
 
-Update: the new code actually makes the CCA sardine Z estimates more in line with each other, so the updates are correct and we are keeping them in `calc_Z`.
+Or maybe spring spawning herring leave the model area and I'm not accounting for that (is this a run with broken migration?).
 
-Most important: Beth says that the M in mort.txt can really be as far off as we are seeing it here.
-The green line in the above NOBA Z plot reflects the initial estimate of what dies by predation mortality, which is a gigantic overesimate because many availability parameters are set to 1.
-After this output step, the predation mortality is rescaled within the model to account for the fact that there really aren't that many to be eaten.
+Our estimate of Z is still from the actual numbers alive at the end of each timestep and the known input recruitment prior to any mortality, so I think it really is the best estimate we have.
 
-Our estimate of Z is from the actual numbers alive at the end of each timestep and the known input recruitment prior to any mortality, so I think it really is the best estimate we have.
-More debugging code commented out below but see the rmd file for the gory and redundant details.
-<!--
-```{r calcZ-snippet2, eval=FALSE}
+### Two methods for F; comparison
 
-survey_testNall <- readRDS(file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))
+NOBA F from mort.txt
 
-# numbers at ageclaggregated over layers, retain polygons (output of create_survey)
-survey_testN_ss <- survey_testNall[survey_testNall$species == species_ss,]
+```{r NOBArelF}
 
-yoy <- YOY_ss
-nums <- survey_testN_ss
-fgs <- funct.groups
-biolprm <- biol
-toutinc <- runpar$toutinc
+plotF <- ggplot(data=relF_ss, aes(x=Time/365, y=relF)) +
+  geom_line() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
 
-#calc_Z <- function(yoy, nums, fgs, biolprm, toutinc) {
-  # subset the yoy for species included in the fgs file
-  # that are turned on
-  # colnames of the recruit data are "Time" and a column for each
-  # species where the name is the functional group with an additional .0
-  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
-  recruits <- yoy[, colnames(yoy) %in% c("Time", paste0(turnedon$Code, ".0"))]
-
-  # mg carbon converted to wet weight in tonnes
-  k_wetdry <- biolprm$kgw2d / 1000000000
-  # Sum of structural and reserve nitrogen (KWSR_RN + KWSR_SN)
-  nitro <- merge(biolprm$kswr, biolprm$kwrr, by = "1")
-  nitro$sum <- apply(nitro[, 2:3], 1, sum)
-
-  # legacy: If output is from legacy code there will be an error in the
-  # yoy data, where the first row is in a different unit.
-  # yoy.txt is the biomass in tonnes per spawning event summed over the total
-  # model domain.
-  # The first row (< Nov/Dec 2015) is stored as biomass and the remaining rows
-  # are stored in numbers, must convert the entire matrix to biomass
-  # Check if legacy code and if so convert the numbers to biomass
-
-  # G.Fay 2/21/16 : changed yoy to recruits in below loop.
-  if (abs(recruits[1, 2] / recruits[2, 2]) > 10) {
-    recruits[2:NROW(recruits), 2:NCOL(recruits)] <- recruits[2:NROW(recruits), 2:NCOL(recruits)] *
-    nitro[match(gsub(".0", "", colnames(recruits)[-1]), nitro[, 1]), "sum"]
-  }
-
-  # Wide to long
-  recruits <- reshape(data = recruits, direction = "long",
-    varying = colnames(recruits)[-1],
-    v.names = "recruits",
-    times = colnames(recruits)[-1],
-    timevar = "group")
-  rownames(recruits) <- 1:NROW(recruits)
-  recruits <- recruits[, -which(colnames(recruits) == "id")]
-  # Switch from species code to species
-  recruits$group <- gsub("\\.0", "", recruits$group)
-  recruits <- merge(recruits, fgs[, c("Code", "Name")],
-    by.x = "group", by.y = "Code")
-
-  # merge recruits with strucn and resn of recruits from biol.prm file
-  recruits <- merge(recruits, nitro[, c(1, 4)],
-    by.x = "group", by.y = "1", all.x = TRUE, all.y = FALSE)
-  colnames(recruits)[which(colnames(recruits) == "recruits")] <- "recruitsbio"
-  # Get recruits in numbers rather than biomass
-  recruits$recruits <- recruits$recruitsbio / recruits$sum
-  recruits$yr <- as.integer(round(recruits$Time)/365) #needed to merge with totnums
-
-  # June 2019:Isaac determined that yr 0 (Time 0) in the YOY file is not real, don't use it
-  # code below assigns year 1 to times 0:stepperyr, all good output timesteps in truth$nums
-
-  # G.Fay 2/21/16
-  # UGLY code below tries to align fraction of annual yoy with timing of recruitment
-  # values in YOY.txt are total YOY that year waiting to recruit.
-  # seems to get rid of most of 'issues' - some v.minor survival >1,
-  #perhaps due to averaging of survival over toutinc days
-  nyrs <- ceiling(max(yoy$Time)/365)
-  times <- unique(yoy$Time)
-
-  # SKG June 2019
-  # need to match codes below, this mismatches when biolprm not in same order!!
-  #recstart_temp <- biolprm$recruit_time
-  #recstart_temp[,2] <- biolprm$time_spawn[-(grep('#',
-  #                     biolprm$time_spawn[,1])),2] + biolprm$recruit_time[,2]
-  # was this grep to get rid of a species with #XXX? breaks when there are none
-  #recstart_temp[,2] <- biolprm$time_spawn[,2] + biolprm$recruit_time[,2]
-
-  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%turnedon$Code,] #bug added codes with no YOY output
-  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%recruits$group,]
-
-  # June 12 2019 Beth determined that we do not need spawn_period
-
-  # rectiming is a dataframe with species code, time_spawn (day of year),
-  # recruit_time (number of days), and recruit period (number of days).
-  # we use these to calculate recstart (day of year) and recend (day of year)
-  rectiming <- merge(biolprm$time_spawn, biolprm$recruit_time, by = 1)
-  rectiming <- merge(rectiming, biolprm$recruit_period, by = 1)
-  names(rectiming) <- c("Code", "time_spawn", "recruit_time", "recruit_period")
-  rectiming <- rectiming %>%
-    mutate(recstart = time_spawn + recruit_time) %>%
-    mutate(recend = recstart + recruit_period)
-
-  #subsets to groups of interest
-  recstart_temp <- rectiming[rectiming$Code %in% recruits$group,]
-
-  # Sum numbers output over all boxes/depth/cohorts
-  # align model output timesteps (days) with recruitment periods (days)
-  totnums <- aggregate(atoutput ~ species + time, data = nums, sum) %>%
-    #mutate(time.days = (time+1)*toutinc) %>% #makes time 0 into days 0->73, etc
-    mutate(time.days = (time)*toutinc) %>% #makes time 1 into days 0->73, etc
-    mutate(yr = ceiling(time.days/365))  # yr 1 is 0:stepsperyr to match recruits yr1
-
-  totnums <- merge(totnums, recruits,
-                   by.x = c("yr", "species"), by.y = c("yr", "Name"),
-                   all.x = TRUE) %>%
-    arrange(time)
-
-  totnums$frac_recruit <- 0
-
-  for (irow in 1:nrow(recstart_temp)) {
-    group <- recstart_temp$Code[irow]
-    pick <- which(totnums$group == group)
-
-    recstart <- seq(recstart_temp$recstart[irow],by=365,length.out=nyrs)
-    recstart <- recstart[recstart<max(totnums$Time[pick])]
-    recend <- recstart + recstart_temp$recruit_period[irow]
-    #rec_times <- rbind(rec_times,cbind(group,recstart,recend))
-
-    for (i_rec in 1:length(recstart)) {
-      i_tstart <- which(pick==min(pick[totnums$time.days[pick]>=recstart[i_rec]]))
-      i_tstop <- which(pick==min(pick[totnums$time.days[pick]>=recend[i_rec]]))
-      if (i_rec == length(recstart)) i_tstop <- length(pick)
-      n_t <- 1+i_tstop-i_tstart
-      for (i_t in 1:n_t) {
-        t_temp <- totnums$time.days[pick[i_tstart+i_t-1]]
-        num_temp <- t_temp - recstart[i_rec]
-        if (i_t>1) {
-          if ((recend[i_rec]-t_temp)>toutinc) {
-            num_temp <- toutinc
-          }
-          else {
-            num_temp <- recend[i_rec]-(totnums$time.days[pick[i_tstart+i_t-2]])
-          }
-        }
-        frac_temp <- max(c(0,num_temp /(recend[i_rec]-recstart[i_rec])))
-        if(frac_temp > 1.0) frac_temp = 1.0
-        totnums$frac_recruit[pick[i_tstart+i_t-1]] <- frac_temp
-      }
-    }
-  }
-
-  #G.Fay 1/6/16, expand to num of recruits
-  # Recruit / mg C converted to wet weight in tonnes / redfield ratio of C:N
-  totnums$recruits <- totnums$recruits/k_wetdry/biolprm$redfieldcn
-
-  totnums$group <- recruits$group[match(totnums$species, recruits$Name)]
-  # For all time increments where there
-  totnums$recruits[is.na(totnums$recruits)] <- 0
-
-  totnums$annrecruits <- totnums$recruits
-
-  totnums$recruits <- totnums$annrecruits*totnums$frac_recruit
-
-  # Calculate survivors for each species group
-  totnums$survivors <- totnums$recruits
-
-  totnums$survival <- totnums$survivors
-  # Calculate survival for each group
-  for (group in unique(totnums$group)) {
-    #if(group == "SHD") browser()
-    pick <- which(totnums$group == group)
-    survival_temp <- c(NA,
-      totnums$survivors[pick[-1]]/totnums$atoutput[pick[-length(pick)]])
-
-    # G.Fay 2/21/16  "think" this is what things should be, recruits don't show up
-    # in numbers at age until time step after the recruitment event.
-     survival_temp <- c(
-       (totnums$atoutput[pick[-1]]-
-          totnums$recruits[pick[-1]])/totnums$atoutput[pick[-length(pick)]],NA)
-
-#     survival_temp <- c(
-#       (totnums$atoutput[pick[-1]])/
-#          (totnums$recruits[pick[-1]]+totnums$atoutput[pick[-length(pick)]]),NA)
-
-    survival_temp[survival_temp < 0] <- NA
-    # Use first positive value to replace the initial year and all negative vals
-    firstgood <- which(!is.na(survival_temp))[1]
-
-    survival_temp[1:firstgood] <- survival_temp[firstgood]
-    for(ii in seq_along(survival_temp)) {
-      if (is.na(survival_temp[ii])) {
-        nonzero <- which(which(survival_temp > 0) > ii)
-        if (length(nonzero) == 0) nonzero <- which(survival_temp > 0)
-        survival_temp[ii] <- survival_temp[which.min(abs(nonzero - ii))]
-      }
-    }
-    totnums$survival[pick] <- survival_temp
-   }
-
-  #Calculate Z
-  totnums$Z <- -1 * log(totnums$survival)
-  finaldata <- data.frame("species" = totnums$species,
-    "agecl" = NA, "polygon" = NA, "layer" = NA,
-    "time" = totnums$time, "atoutput" = totnums$Z)
-
-  #return(finaldata)
-#}
-
-```
--->
-
-## Biological sampling: true numbers at age class
-
-Number of true age classes is stored in the [...]groups[...].csv file, which is read in using `load_fgs`.
-We read this in as funct.groups above.
-The column NumAgeClassSize stores the number of true ages:
-
-Number of age classes for Cod: `r funct.groups$NumAgeClassSize[funct.groups$Name=="North_atl_cod"] `
-
-```{r Natage-plot}
-
-Natage <- readRDS(file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
-
-Natage_ss <- Natage[Natage$species == species_ss,]
-
-Natageplot <- ggplot(Natage_ss, aes(x=agecl, y=atoutput)) +
-  geom_point() +
-  theme_tufte() +
-  labs(subtitle = paste(scenario.name,
-                        Natage_ss$species))
-
-Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free")
-Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free")
-Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free")
-Natageplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free")
+plotF
 
@@
 
-This really is still overly optimistic, but now that I think `calc_Z` is working, if we feed these into `calc_stage2age` and get a result that matches the NOBA output ANNAGEBIO.nc, then we can assume that function is correct.
+NOBA biomass based F
 
-These are estimated true age comps for Cod in annual ages:
+```{r NOBAbioF}
+# catch from txt output
+truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)
 
-```{r plotannage}
+# species in this output is the code not the name, putting in the name
+truecatchbio_ss <- truecatchbio %>%
+  left_join(funct.groups, by = c("species"="Code")) %>% 
+  mutate(species = ifelse(!is.na(Name), Name, Code)) %>%
+  select(species, agecl, polygon, layer, time, atoutput) %>%
+  filter(species == species_ss) %>%
+  arrange(time)
 
-survey_testNall <- readRDS(file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))
+rm(truecatchbio)
 
-# numbers at ageclaggregated over layers, retain polygons (output of create_survey)
-survey_testN_ss <- survey_testNall[survey_testNall$species == species_ss,]
+#biomass from txt output
+atBtxt2 <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)
+groupslookup <- funct.groups %>%
+  filter(IsTurnedOn > 0)
 
-# for single species, Z was calculated above as surveyresZ
-trueages_ss <- calc_stage2age(nums_data = survey_testN_ss,
-                                 biolprm = biol, 
-                                 yoy = YOY_ss,
-                                 fgs = funct.groups, 
-                                 runprm = runpar)
+atBtxt2tidy <- atBtxt2 %>%
+  select(Time:DIN) %>%
+  #select(Time, FPL:DIN) %>%
+  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
+  gather(species, biomass, -Time) %>%
+  filter(species %in% species_ss)
 
-# run sample fish on this for completeness? yes to aggregate
-trueNagesamp_ss <- sample_fish(trueages_ss, effNhigh)
+#merge and estimate F
+annbio <- atBtxt2tidy %>%
+  mutate(yr = (Time/365)+1)
 
-# plot it
+anncatch <- truecatchbio_ss %>%
+  mutate(yr = time/365)
 
-Natageplot <- ggplot(trueNagesamp_ss, aes(x=agecl, y=atoutput)) +
-  geom_point() +
-  theme_tufte() +
-  labs(subtitle = paste(scenario.name,
-                        trueNagesamp_ss$species))
-
-Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
-Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
-Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
-Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+annbioF <- merge(annbio, anncatch) %>%
+  mutate(annbioF = atoutput/biomass) %>%
+  arrange(yr)
 
@@
 
-Still not looking great, but at least not constant. I am now assuming we have the wrong calculations in `calc_stage2age`. Load true ages from ANNAGEBIO.nc:
+This annual biomass based estimate of F is not scaled down from the input F (show in points in the plot), unlike in the CCA.
 
-```{r compareages}
+```{r F_estplot}
 
-annage_nc <- paste0("output", scenario.name, "ANNAGEBIO.nc")
-bboxes <- get_boundary(boxpars)
+plotannbioF <- ggplot(data=annbioF, aes(x=yr, y=annbioF)) +
+  geom_line() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
 
-# need to change load_nc, at line 117 hardcoded for 10 cohorts!
-trueage_atout <- load_nc(dir = d.name,
-                         file_nc = annage_nc,
-                         bps = boxpars,
-                         fgs = funct.groups,
-                         select_groups = survspp,
-                         select_variable = "Nums",
-                         check_acronyms = TRUE,
-                         bboxes = bboxes)
-
-trueage_atout_ss <- trueage_atout[trueage_atout$species == species_ss,]
-
-#survey census to agg over layers
-survey_trueage_atout_ss <- create_survey(dat = trueage_atout_ss,
-                                 time = timeall,
-                                 species = survspp,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-# sample fish to agg over polygons
-sample_trueage_atout_ss <- sample_fish(survey_trueage_atout_ss, effNhigh)
+plotannbioF + geom_point(data=relF_ss, aes(x=Time/365, y=relF))
 
@@
 
-Compare true age output from atlantis with our estimate (and with aggregated age comps):
+NOBA numbers based F
 
-```{r plot-esttrueage-atltrueage}
+This should be numbers caught divided by total numbers (of fully selected fish) from the previous timestep.
 
-compareAge <- ggplot() +
-  geom_point(data=trueNagesamp_ss, aes(x=agecl,y=atoutput, color="atlantisom"), alpha = 10/10) +
-  geom_point(data=sample_trueage_atout_ss, aes(x=agecl,y=atoutput, color="atout true"), alpha = 10/10) + 
-  geom_point(data=Natage_ss, aes(x=agecl*2, y=atoutput, color="cohort true"), alpha = 10/10) + 
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(subtitle = paste(scenario.name,
-                        trueNagesamp_ss$species))
+```{r catchnums, eval=FALSE}
 
-compareAge + 
-  facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
-compareAge + 
-  facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
-compareAge + 
-  facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
-compareAge + 
-  facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+truecatchnums <- create_fishery_subset(dat = truth$catch,
+                                 time = timeall,
+                                 species = species_ss,
+                                 boxes = boxall)
+
+# apply default sample fish to get numbers at age aggregated over polygon
+catch_numsss <- sample_fish(truecatchnums, effNhigh)
+
+# save this, needed as input to calc_Z?
+saveRDS(catch_numsss, file.path(d.name, paste0(scenario.name, "truecatch_nums_Nssh.rds")))
 
@@
 
-We only have the first 10 cohorts of true age output due to hardcoding in load_nc.
-I may write another load function that changes that.
-Either way our early ages don't match.
-Need to step through `calc_stage2age` next!
+```{r Fnums}
+
+catch_numsss <- readRDS(file.path(d.name, paste0(scenario.name, "truecatchnums_Nssh.rds")))
+
+totnum <- readRDS(file.path(d.name, paste0(scenario.name, "Natageclcensus.rds")))
+
+totnumss <- filter(totnum, species == species_ss)
+
+# align nums previous timestep with catch current timestep
+
+```

### 2020-01-14 TestCalcFandM.Rmd
added NOBA cod, M still really high, start on numbers based F not done

@@
 
-Is there a mismatch between biomass based F and numbers based Z?
+Is there a mismatch between biomass based F and numbers based Z? We can't do this comparison for CCA because our catch numbers from the .nc file have the pre-December 2015 bug.
+
+Also, this may indicate that we need to redo calc_Z as cohort-specific after all.
 
@@
 
-Or maybe spring spawning herring leave the model area and I'm not accounting for that (is this a run with broken migration?).
+Or maybe spring spawning herring leave the model area and I'm not accounting for that (is this a run with broken migration?). No evidence that these species migrate in documentation.
 
@@
 
-# align nums previous timestep with catch current timestep
+# work in progress 
+# align nums at agecl previous timestep with catch current timestep
+# for each agecl: time-1 catch_numsss$atoutput / time totnumss$atoutput
+catch_numsss <- catch_numsss %>%
+  select(species, agecl, catchtime=time, catchnum=atoutput) %>%
+  mutate(Ftime = catchtime - 1)
+
+totnumss <- totnumss %>%
+  select(species, agecl, time, popnum=atoutput) %>%
+  mutate(Ftime = time)
+
+#not done yet dont use this
+timestepnumF <- merge(totnumss, catch_numsss) %>%
+  mutate(stepnF = (catchnum/popnum)) %>%
+  arrange(Ftime, agecl)
 
 ```
+
+Try this for cod, herring don't have much F
+
+```{r NOBAcod}
+
+species_ss <- "North_atl_cod"
+
+# get code matching species name to split YOY file
+code_ss <- funct.groups$Code[which(funct.groups$Name == species_ss)]
+
+# cut to a single species in YOY file
+YOY_ss <- YOY %>%
+  select(Time, paste0(code_ss, ".0"))
+
+# numbers at agecl at full resolution (all polygons and layers)
+truenums_ss <- truth$nums[truth$nums$species == species_ss,]
+
+#calc_Z <- function(yoy, nums, fgs, biolprm, toutinc)
+
+#oops, need to generalize calc_Z for subannual timesteps in YOY!
+#or input YOY only at 0, 365, etc since the numbers repeat for timesteps 
+YOY_ss <- YOY_ss %>%
+  filter(Time %in% seq(0, max(Time), by=365))
+
+fullresZ <- calc_Z(yoy = YOY_ss,
+                   nums = truenums_ss,
+                   fgs = funct.groups,
+                   biolprm = biol,
+                   toutinc = runpar$toutinc)
+
+# compare as above with mort.txt output
+file.mort <- file.path(d.name, paste0("output", scenario.name, "Mort.txt"))
+
+mortish <- read.table(file.mort, header = TRUE)
+
+relF_ss <- mortish %>%
+  select(Time, relF = paste0(code_ss, ".F"))
+
+relM_ss <- mortish %>%
+  select(Time, relM = paste0(code_ss, ".M"))
+
+Zish <- merge(relF_ss,relM_ss) %>%
+  mutate(relZ = relF + relM)
+
+annZ2 <- fullresZ %>%
+  mutate(yr = floor(time/stepperyr)+1) %>%
+  group_by(species, yr) %>%
+  summarise(Z = sum(atoutput))
+
+plotZ <- ggplot() +
+  geom_line(data=Zish, aes(x=Time/365, y=relZ, color="mort.txt Z")) +
+  geom_point(data=fullresZ, aes(x=time/stepperyr, y=atoutput, color="each timestep Z")) +
+  geom_point(data=annZ2, aes(x=yr-1, y=Z, color="sum timestep Z")) +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
+
+plotZ + ylim(-1, 10.0)
+
+```
+
+Now estimate cod F from biomass and compare:
+
+```{r NOBAcodFest}
+
+# catch from txt output
+truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)
+
+# species in this output is the code not the name, putting in the name
+truecatchbio_ss <- truecatchbio %>%
+  left_join(funct.groups, by = c("species"="Code")) %>% 
+  mutate(species = ifelse(!is.na(Name), Name, Code)) %>%
+  select(species, agecl, polygon, layer, time, atoutput) %>%
+  filter(species == species_ss) %>%
+  arrange(time)
+
+rm(truecatchbio)
+
+#biomass from txt output
+atBtxt2 <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)
+groupslookup <- funct.groups %>%
+  filter(IsTurnedOn > 0)
+
+atBtxt2tidy <- atBtxt2 %>%
+  select(Time:DIN) %>%
+  #select(Time, FPL:DIN) %>%
+  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
+  gather(species, biomass, -Time) %>%
+  filter(species %in% species_ss)
+
+#merge and estimate F
+annbio <- atBtxt2tidy %>%
+  mutate(yr = (Time/365)+1)
+
+anncatch <- truecatchbio_ss %>%
+  mutate(yr = time/365)
+
+annbioF <- merge(annbio, anncatch) %>%
+  mutate(annbioF = atoutput/biomass) %>%
+  arrange(yr)
+
+```
+
+```{r NOBAcodF_estplot}
+
+plotannbioF <- ggplot(data=annbioF, aes(x=yr, y=annbioF)) +
+  geom_line() +
+  theme_tufte() +
+  labs(subtitle = paste(scenario.name, species_ss))
+
+plotannbioF + geom_point(data=relF_ss, aes(x=Time/365, y=relF))
+
+```
+
+Cod F based on biomass isn't that different from the F reported in mort.txt.
+
+Cod M if we believe our Z estimate from calc_Z:
+
+```{r NOBAcodannZ2}
+
+annZ2 <- fullresZ %>%
+  mutate(yr = floor(time/stepperyr)+1) %>%
+  group_by(species, yr) %>%
+  summarise(Z = sum(atoutput))
+
+```
+
+```{r isthisM, message=FALSE, warning=FALSE}
+
+annbioF <- select(annbioF, yr, annbioF)
+
+# annM1 <- merge(annZ1, annbioF, all.x = T) %>%
+#   mutate(M = Z-annbioF)
+
+annZ2_Time <- annZ2 %>%
+  mutate(Time = yr*365)
+
+annM2 <- merge(annZ2_Time, annbioF) %>%
+  mutate(M = Z-annbioF)
+
+plotM <- ggplot() +
+  geom_line(data=Zish, aes(x=Time/365, y=relZ, color="mort.txt Z")) +
+  geom_line(data=Zish, aes(x=Time/365, y=relM, color="mort.txt M")) +
+  #geom_point(data=annM1, aes(x=Time/365, y=M, color="M = endyr Z-F")) +
+  geom_point(data=annM2, aes(x=yr, y=M, color="M = sum timestep Z-F")) +
+  theme_tufte() +
+  theme(legend.position = "bottom", legend.box = "horizontal") +
+  scale_color_discrete(NULL) +
+  guides(color = guide_legend(nrow = 1)) +
+  labs(subtitle = paste(scenario.name, species_ss))
+
+plotM + ylim(0, 10.0)
+
+```

### 2020-01-16 WrapperDevelopment.Rmd
what should output look like
begin wrapper functions

--- 2020-01-13 WrapperDevelopment.Rmd
@@
 
-1.
-initialize: locate model output (function to make a config file?), read in basic run parameters and species names, run_truth for all once and save output (optionally, run_truth for only one species?)
+1. initialize: locate model output (function to make a config file?), read in basic run parameters and species names, run_truth for all once and save output 
+    + User to make up front decision--want all species or just a subset?
+    + May want multiple species that need to be split in a later step for models
+    + wraps functions 
+        + sourcing model config file--add function to make config file?
+      + specifies top directory for source files, names key files for use later
+      + load_fgs
+      + load all other needed files--YOY, catch
+      + run_truth
+1. get assessment input data: configure survey, run all survey functions, configure fishery, run all fishery functions (how to save output?)
+    + Based on decision above, save interim steps for use later
+    + Need a switch for single vs multispecies estimation models?
+    + Separate index and comp functions in output
+    + indices:
+      + create survey config file(s), default is a census of all see census_spec.R
+        + survey index in biomass or numbers? (uses truth$biomass_ages vs truth$nums)
+        + wraps functions create_survey, sample_survey_biomass
+      + create fishery config file(s)
+    + comps:
+      + select age only or age and length
+        + age only wraps create_survey using nums with sample_fish to get age comps
+        + age and length (with weight at age)
+          + create_survey followed by sample_fish, 
+          + aggregateDensityData for resn and structn followed by sample_fish
+          + inputs all to calc_age2length
+      + catch comps wrap create_fishery_subset with sample_fish to get age comps, same as above for lengths
+1. write input data for specific model: wrap functions in development for SS3, add functions for other models (ASAP, multispecies models)
+    + wrap r4ss functions to read and write dat and ctl files
+    + possibly start with set of dummy SS files with package (take from ss3sim?)
+      + read in dat file
+      + indices wrap SS_write_ts
+      + comps wrap reformat_compositions, SS_write_comps, some of the checks for bin matches\
+      + write full dat file
+    + start with dummy ctl file (any from ss3sim?)
+      + read in ctl file
+      + wrap SS_write_bio using outputs of calc_Z, load_YOY, load_biolprm, load_runprm, etc
+1. run specified model--automate for scenarios, save outputs 
+1. skill assessment metrics: compare stored Atlantis truth with model output
 
-1. get assessment input data: configure survey, run all survey functions, configure fishery, run all fishery functions (how to save output?)
+### atlantisom initialize function
 
-1. write input data for specific model: wrap functions in development for SS3, add functions for other models (ASAP, multispecies models)
+```{r ominit}
+
+library(tidyverse)
+
+om_init <- function(config = configfile){
+
+  # Where are the atlantis output files? Consider filling with shiny app in future
+  source(config)
+  # needs these files, for example config file CC3config.R is:
+
+  # d.name <- here::here("atlantisoutput","CC_2063_OA_OFF_22")
+  # functional.groups.file <- "CalCurrentV3Groups.csv"
+  # biomass.pools.file <- "DIVCalCurrentV3_Biol.nc"
+  # biol.prm.file <- "CalCurrentV3_Biol.prm"
+  # box.file <- "CalCurrentV3_utm.bgm"
+  # initial.conditions.file <- "DIVCalCurrentV3_Biol.nc"
+  # run.prm.file <- "CalCurrentV3_run.xml"
+  # scenario.name <- "CCV3"
+  # bioind.file <- "outputCCV3BiomIndx.txt"
+  # catch.file <- "outputCCV3Catch.txt"
+
+  #Load functional groups
+  funct.groups <- atlantisom::load_fgs(dir=d.name,
+                           file_fgs = functional.groups.file)
+  #Get just the names of active functional groups
+  funct.group.names <- funct.groups %>% 
+    filter(IsTurnedOn == 1) %>%
+    select(Name) %>%
+    .$Name
+
+  # load true catch in tons
+  truecatchbio <- atlantisom::load_catch(d.name, file_catch = catch.file, fgs = funct.groups)
+
+  # load YOY
+  YOY <- atlantisom::load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))
+
+  # load biol_prm
+  biol <- atlantisom::load_biolprm(d.name, biol.prm.file)
+
+  # load run_prm
+  runpar <- atlantisom::load_runprm(d.name, run.prm.file)
+
+  # default run_truth setup will save the file, so check for that first
+  if(!file.exists(file.path(d.name, 
+                            paste0("output", scenario.name, "run_truth.RData")))){
+    #Store all loaded results into an R object
+    truth <- atlantisom::run_truth(scenario = scenario.name,
+                       dir = d.name,
+                       file_fgs = functional.groups.file,
+                       file_bgm = box.file,
+                       select_groups = funct.group.names,
+                       file_init = initial.conditions.file,
+                       file_biolprm = biol.prm.file,
+                       file_runprm = run.prm.file,
+                       verbose = TRUE
+    )
+  } else {
+    truth <- get(load(file.path(d.name,
+                                paste0("output", scenario.name, "run_truth.RData"))))
+  }
+
+  omlist <-list("funct.groups" = funct.groups, 
+                "funct.group.names" = funct.group.names,
+                "truecatchbio" = truecatchbio,
+                "YOY" = YOY,
+                "biol" = biol,
+                "runpar" = runpar,
+                "truth" = truth)
+
+  return(omlist)
+}
+
+```
+
+Usage:
+
+```{r exinit}
+
+library(here)
+CC3om <- om_init(here("config/CC3config.r"))
+
+```
+
+### Get assessment input data
+
+Split to single species or subset of assessed species, get index data, get compositional data.
+
+```{r inputs}
+
+om_species <- function(species = spp, omlist){
+  # spp format c("speciesname1", "speciesname2")
+  if(!all(species %in% omlist$funct.group.names)) stop("species name not found") 
+  species_ss <- species
+
+  return(species_ss)
+}
+
+```
+
+Usage:
+
+```{r exsp}
+
+om_species(c("Pacific_sardine"), CC3om)
+
+```
+
+## Cumbersome giant files eat all the memory!
+
+Implement Christine's memory saving things from the NEMOW demo

### 2020-01-16 TestCalcFandM.Rmd
Z and M are killing me here

@@
 
+totnums <- aggregate(atoutput ~ species + time, data = truenums_ss, sum) %>%
+    mutate(time.days = (time+1)*runpar$toutinc) %>% #makes time 0 into days 0->73, etc
+    mutate(yr = ceiling(time.days/365))  # yr 1 is 0:stepsperyr to match recruits yr1
+    #mutate(yr = floor(time.days/365)) # includes 0 value, yr 1 is
+
+#only want numbers at end of year, snapshot
+totnumsann <- totnums %>%
+  filter(time.days %in% seq(365, max(time.days), by=365))
+
+# mg carbon converted to wet weight in tonnes
+k_wetdry <- biol$kgw2d / 1000000000
+
+# WARNING only works for CCA because YOY.txt rows 2:end are already in numbers
+# and we are merging out the incorrect and irrelevant YOY row 1 (Time=0)
+# also hardcoded for sardine example
+recnums <- YOY_ss %>%
+  mutate(yr = as.integer(round(YOY_ss$Time)/365)) %>%
+  mutate(recnums = SSH.0/k_wetdry/biol$redfieldcn)
+
+annZ1 <- merge(totnumsann, recnums) %>%
+  mutate(recnums = replace_na(recnums, 0)) %>%
+  mutate(numslessrec = atoutput - recnums) %>%
+  mutate(surv = numslessrec/lag(atoutput, default = first(numslessrec))) %>%
+  mutate(Z = -1 * log(surv))
+
 annZ2 <- fullresZ %>%
@@
   geom_line(data=Zish, aes(x=Time/365, y=relZ, color="mort.txt Z")) +
-  geom_point(data=fullresZ, aes(x=time/stepperyr, y=atoutput, color="each timestep Z")) +
+  #geom_point(data=fullresZ, aes(x=time/stepperyr, y=atoutput, color="each timestep Z")) +
+  geom_point(data=annZ1, aes(x=yr-1, y=Z, color="annual Z")) +
   geom_point(data=annZ2, aes(x=yr-1, y=Z, color="sum timestep Z")) +
@@
   mutate(Ftime = time)
+
+#what about annual Z from totnumss catch curve?
+# each timestep use lm(agecl~ log(atoutput))
+library(dplyr)
+library(broom) # to change model results into data frame
+catchcurve1 <- totnumss %>% 
+  group_by(time) %>%
+  do(tidy(lm(log(popnum) ~ agecl, data = .))) %>% 
+  filter(term == "agecl")
+
+# or do for the year averaging natage at each timestep?
+catchcurve2 <- totnumss %>%
+  mutate(yr = floor(time/stepperyr)+1) %>%
+  group_by(yr, agecl) %>%
+  summarise(avgpopnum = mean(popnum)) %>%
+  do(tidy(lm(log(avgpopnum) ~ agecl, data = .))) %>% 
+  filter(term == "agecl")
+
+altZ1 <- catchcurve1 %>%
+  select(time, estimate) %>%
+  mutate(yr = floor(time/stepperyr)+1) %>%
+  group_by(yr) %>%
+  summarise(Z = sum(-estimate))
+
+altZ2 <- catchcurve2 %>%
+  mutate(Z=-estimate) %>%
+  select(yr, Z)
+
+#another alternative tracking cohorts
+# get rid of agecl 1 to ignore recruitment
+# still need rec_time to know when they go to next agecl
+# survival is 
+#  within year 
+# SSH actually have two ages per agecl so this wont work
+# also, when I look at an age class numbers are INCREASING each timestep
+# which makes no sense unless there is migration in?
 

### 2020-01-16 WrapperDevelopment.Rmd
more details

@@
 ### atlantisom initialize function
+
+Things to add: checks for output timestep, loading bioind.file with function similar to load_catch, loading box file separately?
 
@@
 
-Split to single species or subset of assessed species, get index data, get compositional data.
+Split to single species or subset of assessed species, get index data, get compositional data. By default, remove full om results and keep only species subsets to save memory.
 
@@
 
-om_species <- function(species = spp, omlist){
+om_species <- function(species = spp, omlist, removefullom = TRUE){
   # spp format c("speciesname1", "speciesname2")
@@
 
-  return(species_ss)
+  #subset species true bio
+  #-----REDO THIS with a function like load_catch and load above----#
+  # read Atlantis output files
+  atBtxt2 <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)
+  groupslookup <- funct.groups %>%
+    filter(IsTurnedOn > 0)
+
+  truetotbio_ss <- atBtxt2 %>%
+    select(Time:DIN) %>%
+    #select(Time, FPL:DIN) %>%
+    rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
+    gather(species, biomass, -Time) %>%
+    filter(species %in% species_ss)
+
+  #subset species true catch
+  truecatchbio_ss <- omlist$truecatchbio[truecatchbio$species == species_ss,]
+
+  #subset species YOY
+
+  #subset species biol parameters
+
+  #subset species funtional group info?
+
+  #keep the runpar and also need boxes for survey selection
+
+  omlist_ss <- list("species_ss" = species_ss,
+                    "runpar" = omlist$runpar)
+
+  if(removefullom = TRUE) rm(omlist)
+
+  return(omlist_ss)
 }

### 2020-01-16 WrapperDevelopment.Rmd
wrappers to initialize and split species so far

--- 2020-01-16 WrapperDevelopment.Rmd
@@
 
-## Too many steps!
+## Too many steps! and too much memory!
 
@@
       + indices wrap SS_write_ts
-      + comps wrap reformat_compositions, SS_write_comps, some of the checks for bin matches\
+      + comps wrap reformat_compositions, SS_write_comps, some of the checks for bin matches
       + write full dat file
@@
 ### atlantisom initialize function
+
+Things to add: checks for output timestep?
 
@@
 
+  # load true total biomass in tons
+  truetotbio <- atlantisom::load_bioind(d.name, file_bioind = bioind.file, fgs = funct.groups)
+
   # load true catch in tons
@@
   runpar <- atlantisom::load_runprm(d.name, run.prm.file)
+
+  # load box info file
+  boxpars <- atlantisom::load_box(d.name, box.file)
 
@@
                 "funct.group.names" = funct.group.names,
+                "truetotbio" = truetotbio,
                 "truecatchbio" = truecatchbio,
@@
                 "runpar" = runpar,
+                "boxpars" = boxpars,
                 "truth" = truth)
@@
 
-Split to single species or subset of assessed species, get index data, get compositional data.
+Split to single species or subset of assessed species, get index data, get compositional data. By default, remove full om results and keep only species subsets to save memory.
+
+#### Split to focal species
+
+Is this where we should reconcile timesteps? depends what functions are expecting.
 
@@
 
-  return(species_ss)
+  #subset species true bio
+  truetotbio_ss <- omlist$truetotbio[omlist$truetotbio$species %in% species_ss,]
+
+  #subset species true catch
+  truecatchbio_ss <- omlist$truecatchbio[omlist$truecatchbio$species %in% species_ss,]
+
+  #subset species YOY
+  # get code matching species name to split YOY file
+  code_ss <- omlist$funct.groups$Code[which(omlist$funct.groups$Name %in% species_ss)]
+  # cut to a single species in YOY file
+  YOY_ss <- omlist$YOY %>%
+    select(Time, paste0(code_ss, ".0"))
+  # reformat to be like all the other objects
+
+  # numbers at agecl at full resolution (all polygons and layers)
+  truenums_ss <- omlist$truth$nums[omlist$truth$nums$species %in% species_ss,]
+
+  # biomass at agecl at full resolution (all polygons and layers)
+  truebio_ss <- omlist$truth$biomass_ages[omlist$truth$biomass_ages$species %in% species_ss,]
+
+  # reserve nitrogen at agecl at full resolution
+  trueresn_ss <- omlist$truth$resn[omlist$truth$resn$species %in% species_ss,]
+
+  # structural nitrogen at agecl at full resolution
+  truestructn_ss <- omlist$truth$structn[omlist$truth$structn$species %in% species_ss,]
+
+  # catch in numbers at agecl at full resoluation (all polygons, no layer in output)
+  truecatchnum_ss <- omlist$truth$catch[omlist$truth$catch$species %in% species_ss,]
+
+  #subset species biol parameters? no, may miss some globals that aren't by species
+
+  #subset species functional group info
+  funct.groups_ss <- omlist$funct.groups[omlist$funct.groups$Code %in% code_ss,]
+
+  #keep the runpar and also need boxes for survey selection
+
+  omlist_ss <- list("species_ss" = species_ss,
+                    "code_ss" = code_ss,
+                    "truetotbio_ss" = truetotbio_ss,
+                    "truecatchbio_ss" = truecatchbio_ss,
+                    "YOY_ss" = YOY_ss,
+                    "truenums_ss" = truenums_ss,
+                    "truebio_ss" = truebio_ss,
+                    "trueresn_ss" = trueresn_ss,
+                    "truestructn_ss" = truestructn_ss,
+                    "truecatchnum_ss" = truecatchnum_ss,
+                    "funct.group_ss" = funct.groups_ss,
+                    "biol" = omlist$biol,
+                    "boxpars" = omlist$boxpars,
+                    "runpar" = omlist$runpar)
+
+  if(removefullom) rm(omlist) #not removing passed data object
+
+  return(omlist_ss)
 }
@@
 
-om_species(c("Pacific_sardine"), CC3om)
+CC3om_sardine <- om_species(c("Pacific_sardine"), CC3om)
 
@@
 
-## Cumbersome giant files eat all the memory!
+#### Generate index data for assessment
 
-Implement Christine's memory saving things from the NEMOW demo
+```{r omindices}
+
+om_index <- function(){
+
+}
+
+```
+
+Usage:
+
+```{r exindex}
+
+#CC3om_sard_ind <- om_index(c("survey1", "fishery"), CC3om_sardine)
+
+```
+
+#### Generate compositional data for assessment
+
+```{r omcomps}
+
+```

### 2020-01-16 WrapperDevelopment.Rmd
wrapper for survey and fishery index in progress

@@
 
-om_index <- function(){
+om_index <- function(surveyconfig = survey_configfile, 
+                     fisheryconfig = fishery_configfile, 
+                     omlist_ss = omlist_ss, 
+                     n_reps = n_reps,
+                     save = TRUE){
+
+  #configure the independent survey, a default census is in config/survcensus.R
+  #survey configuration must specify survey time, area, efficiency (q), selectivity
+  #survey species inherited from omlist_ss
+  source(surveyconfig)
+
+  #biomass based fishery independent survey index
+  # this uses result$biomass_ages to sample biomass directly, no need for wt@age est
+  survey_B <- atlantisom::create_survey(dat = omlist_ss$truebio_ss,
+                                        time = survtime,
+                                        species = survspp,
+                                        boxes = survboxes,
+                                        effic = surveffic,
+                                        selex = survselex)
+
+  # call sample_survey_biomass with a bunch of 1000s for weight at age
+  # in the code it multiplies atoutput by wtatage/1000 so this allows us to use
+  # biomass directly
+  wtage <- data.frame(species=rep(survspp, each=max(age_classes)),
+                      agecl=rep(c(1:max(age_classes)),length(survspp)),
+                      wtAtAge=rep(1000.0,length(survspp)*max(age_classes)))
+
+  # this is the step to repeat n_reps time if we want different realizations 
+  # of the same survey design specified above; only observation error differs
+  # using the census cv of 0 will produce identical reps!
+  survObsBiomB <- list()
+  for(i in 1:n_reps){
+    survObsBiomB[[i]] <- atlantisom::sample_survey_biomass(survey_B, surv_cv, wtage)
+  }
+
+  #save survey indices, takes a long time to generate with lots of reps/species
+  if(save){
+    saveRDS(survObsBiomB, file.path(d.name, paste0(scenario.name, "surveyB.rds")))
+  }
+
+  #configure the fishery, a default is in config/fisherycensus.R
+  #fishery configuration must specify survey time, area, efficiency (q), selectivity
+  #fishery species inherited from omlist_ss
+  source(fisheryconfig)
 
@@
 
-#CC3om_sard_ind <- om_index(c("survey1", "fishery"), CC3om_sardine)
+CC3om_sard_ind <- om_index(surveyconfig = here("config/survcensus.R"), 
+                           fisheryconfig = "",
+                           omlist_ss = CC3om_sardine, 
+                           n_reps = 5, 
+                           save = FALSE)
 

### 2020-01-16 config/fisherycensus.R

@@ -1,0 +1,22 @@
+# Default survey configuration here is a census
+# Need to define survey season, area, efficiency, selectivity
+# Inherits species from input omlist_ss
+fishspp <- omlist_ss$species_ss 
+
+# same time dimensioning parameters as in surveycensus.R
+
+# fishery output: learned the hard way this can be different from ecosystem outputs
+fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
+
+# fishery sampling area
+# should return all model areas
+fishboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+fisheffN <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
+
+# survey index cv needed for sample_survey_xxx
+# perfect observation
+fish_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+

### 2020-01-16 config/survcensus.R

@@ -1,0 +1,50 @@
+# Default survey configuration here is a census
+# Need to define survey season, area, efficiency, selectivity
+# Inherits species from input omlist_ss
+survspp <- omlist_ss$species_ss 
+
+# survey season and other time dimensioning parameters
+# generalized timesteps all models
+noutsteps <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
+timeall <- c(0:noutsteps)
+
+stepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutinc
+midptyr <- round(median(seq(0,stepperyr)))
+
+# fishery output: learned the hard way this can be different from ecosystem outputs
+fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
+
+# a survey that takes place once per year mid year has now been declared default
+annualmidyear <- seq(midptyr, noutsteps, stepperyr)
+survtime <- annualmidyear
+
+# survey area
+# should return all model areas
+survboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
+
+# survey efficiency (q)
+# should return a perfectly efficient survey 
+surveffic <- data.frame(species=survspp,
+                     efficiency=rep(1.0,length(survspp)))
+
+# survey selectivity (agecl based)
+# should return all age classes fully sampled (Atlantis output is 10 age groups per spp)
+n_age_classes <- omlist_ss$funct.group_ss$NumCohorts
+age_classes <- 1:n_age_classes
+
+n_annages <- n_age_classes * omlist_ss$funct.group_ss$NumAgeClassSize
+annages <- 1:n_annages
+
+# this is by age class, need to change to use with ANNAGEBIO output
+survselex <- data.frame(species=rep(survspp, each=n_age_classes),
+                     agecl=rep(c(1:n_age_classes),length(survspp)),
+                     selex=rep(1.0,length(survspp)*n_age_classes))
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+surveffN <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
+
+# survey index cv needed for sample_survey_xxx
+# perfect observation
+surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
+

### 2020-01-17 WrapperDevelopment.Rmd
rearranging functions for survey setup

@@
 
+```{r surveyconfig}
+
+surveyconfig <- function(omlist_ss, usersurvey = usersurvey_file){
+  #configure the independent survey, a default census is in config/survcensus.R
+  #survey configuration must specify survey time, area, efficiency (q), selectivity
+  #survey species inherited from omlist_ss
+  survspp <- omlist_ss$species_ss
+  # survey season and other time dimensioning parameters
+  # generalized timesteps all models
+  noutsteps <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
+  timeall <- c(0:noutsteps)
+  stepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutinc
+  midptyr <- round(median(seq(0,stepperyr)))
+
+  # model areas, subset in surveyconfig
+  allboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
+
+  # fishery output: learned the hard way this can be different from ecosystem outputs
+  fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
+
+  # survey selectivity (agecl based)
+  # should return all age classes fully sampled (Atlantis output is 10 age groups per spp)
+  n_age_classes <- omlist_ss$funct.group_ss$NumCohorts
+  age_classes <- 1:n_age_classes
+
+  n_annages <- n_age_classes * omlist_ss$funct.group_ss$NumAgeClassSize
+  annages <- 1:n_annages
+
+  # user options for survey--default is a census with mid-year sample
+  source(usersurvey, local = TRUE)
+
+  surveyspecs <- list("survtime" = survtime,
+                      "survboxes" = survboxes,
+                      "surveffic" = surveffic,
+                      "survselex" = survselex,
+                      "surveffN" = surveffN,
+                      "surv_cv" = surv_cv)
+
+  return(surveyspecs)
+}
+```
+
+Usage:
+
+```{r specsex}
+
+mysurvey <- surveyconfig(CC3om_sardine, here("config/usersurvey.R"))
+```
+
 ```{r omindices}
 
-om_index <- function(surveyconfig = survey_configfile, 
-                     fisheryconfig = fishery_configfile, 
-                     omlist_ss = omlist_ss, 
+om_index <- function(usersurvey = usersurvey_file, 
+                     userfishery = userfishery_file, 
+                     omlist_ss, 
                      n_reps = n_reps,
@@
 
-  #configure the independent survey, a default census is in config/survcensus.R
-  #survey configuration must specify survey time, area, efficiency (q), selectivity
-  #survey species inherited from omlist_ss
-  source(surveyconfig)
+  surveyspec <- surveyconfig(omlist_ss, usersurvey = usersurvey_file)
 
@@
   survey_B <- atlantisom::create_survey(dat = omlist_ss$truebio_ss,
-                                        time = survtime,
-                                        species = survspp,
-                                        boxes = survboxes,
-                                        effic = surveffic,
-                                        selex = survselex)
+                                        time = surveyspec$survtime,
+                                        species = surveyspec$survspp,
+                                        boxes = surveyspec$survboxes,
+                                        effic = surveyspec$surveffic,
+                                        selex = surveyspec$survselex)
 
@@
 
+  #we are not currently subsetting fishery catch because we cannot correct catch.nc
+  #  instead the catch in biomass from catch.txt is read in for the index
+  #  we do not apply any cv to this, but we could this way (default cv=0)
+
+  fishObsCatchB <- list()
+  for(i in 1:n_reps){
+    fishObsCatchB[[i]] <- atlantisom::sample_fishery_totcatch(omlist_ss$truecatchbio_ss, fish_cv)
+  }
+
+  if(save){
+    saveRDS(fishObsCatchB, file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
+  }
+
+  indices <- list("survObsBiomB" = survObsBiomB,
+                  "fishObsCatchB" = fishObsCatchB)
+
+  return(indices)
 }
@@
 CC3om_sard_ind <- om_index(surveyconfig = here("config/survcensus.R"), 
-                           fisheryconfig = "",
+                           fisheryconfig = here("config/fisherycensus.R"),
                            omlist_ss = CC3om_sardine, 

### 2020-01-17 config/fisherycensus.R

@@ -1,5 +1,4 @@
-# Default survey configuration here is a census
-# Need to define survey season, area, efficiency, selectivity
+# Default fishery configuration here is a census
 # Inherits species from input omlist_ss
 fishspp <- omlist_ss$species_ss 
 
@@ -16,7 +15,7 @@ fishboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
 # this effective N is high but not equal to total for numerous groups
 fisheffN <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
 
-# survey index cv needed for sample_survey_xxx
+# fishery catch cv can be used in sample_survey_biomass
 # perfect observation
 fish_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
 

### 2020-01-17 WrapperDevelopment.Rmd
dont need extra surveyconfig function

--- 2020-01-16 WrapperDevelopment.Rmd
@@
 
+The hardcoded script omdimensions.R needs to sit in the atlantisom config folder.
+
 ```{r omindices}
 
-om_index <- function(surveyconfig = survey_configfile, 
-                     fisheryconfig = fishery_configfile, 
-                     omlist_ss = omlist_ss, 
+om_index <- function(usersurvey = usersurvey_file, 
+                     userfishery = userfishery_file, 
+                     omlist_ss, 
                      n_reps = n_reps,
@@
 
-  #configure the independent survey, a default census is in config/survcensus.R
-  #survey configuration must specify survey time, area, efficiency (q), selectivity
-  #survey species inherited from omlist_ss
-  source(surveyconfig)
+  #one script for dimension parameters to be used in multiple functions
+  source("config/omdimensions.R", local = TRUE)
+
+  # user options for survey--default is a census with mid-year sample
+  source(usersurvey, local = TRUE)
 
@@
   #fishery species inherited from omlist_ss
-  source(fisheryconfig)
+  source(userfishery, local = TRUE)
 
+  #we are not currently subsetting fishery catch because we cannot correct catch.nc
+  #  instead the catch in biomass from catch.txt is read in for the index
+  #  we do not apply any cv to this, but we could this way (default cv=0)
+
+  fishObsCatchB <- list()
+  for(i in 1:n_reps){
+    fishObsCatchB[[i]] <- atlantisom::sample_fishery_totcatch(omlist_ss$truecatchbio_ss, fish_cv)
+  }
+
+  if(save){
+    saveRDS(fishObsCatchB, file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
+  }
+
+  indices <- list("survObsBiomB" = survObsBiomB,
+                  "fishObsCatchB" = fishObsCatchB)
+
+  return(indices)
 }
@@
 
-CC3om_sard_ind <- om_index(surveyconfig = here("config/survcensus.R"), 
-                           fisheryconfig = "",
+CC3om_sard_ind <- om_index(usersurvey = here("config/usersurvey.R"), 
+                           userfishery = here("config/fisherycensus.R"),
                            omlist_ss = CC3om_sardine, 
                            n_reps = 5, 
-                           save = FALSE)
+                           save = TRUE)
 

### 2020-01-17 WrapperDevelopment.Rmd
correct config files

@@
   #configure the fishery, a default is in config/fisherycensus.R
-  #fishery configuration must specify survey time, area, efficiency (q), selectivity
+  #fishery configuration can specify only area and time of observation
   #fishery species inherited from omlist_ss
@@
 
+Note that if a census is truly desired for age comps, don't use the sample fish function but just aggregate. I'll make a function for that too.
+
 ```{r omcomps}
 
+om_comps <- function(usersurvey = usersurvey_file, 
+                     userfishery = userfishery_file, 
+                     omlist_ss, 
+                     n_reps = n_reps,
+                     save = TRUE){
+
+  #one script for dimension parameters to be used in multiple functions
+  source("config/omdimensions.R", local = TRUE)
+
+  # user options for survey--default is a census with mid-year sample
+  source(usersurvey, local = TRUE)
+
+  #numbers based fishery independent survey for age and length comps
+  # same user specifications as indices
+  survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
+                                        time = survtime,
+                                        species = survspp,
+                                        boxes = survboxes,
+                                        effic = surveffic,
+                                        selex = survselex)
+
+  #Sample fish for age composition
+  # if we want replicates for obs error this sample function will generate them
+  age_comp_data <- list()
+  for(i in 1:n_reps){
+    age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
+  }
+
+  # save age comps
+  if(save){
+    saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
+  }
+
+  #weights needed for weight at age and length comp calcs
+  # aggregate true resn per survey design
+  survey_aggresn <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
+                                                     time = survtime,
+                                                     species = survspp,
+                                                     boxes = survboxes)
+
+  # aggregate true structn per survey design
+  survey_aggstructn <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
+                                                        time = survtime,
+                                                        species = survspp,
+                                                        boxes = survboxes)
+
+  #dont sample these, just aggregate them using median
+  structnss <- atlantisom::sample_fish(survey_aggresn, surveffN, sample = FALSE)
+
+  resnss <- atlantisom::sample_fish(survey_aggstructn, surveffN, sample = FALSE)
+
+  #this is all input into the length function, replicates follow age comp reps
+  #  separating the length comps from the weight at age here
+  survey_lenwt <- list()
+  survObsLenComp <- list()
+  survObsWtAtAge <- list()
+
+  for(i in 1:n_reps){
+    survey_lenwt[[i]] <- atlantisom::calc_age2length(structn = structnss,
+                                                     resn = resnss,
+                                                     nums = age_comp_data[[i]],
+                                                     biolprm = omlist_ss$biol, 
+                                                     fgs = omlist_ss$funct.group_ss,
+                                                     maxbin = maxbin,
+                                                     CVlenage = lenage_cv, 
+                                                     remove.zeroes=TRUE)
+
+    survObsLenComp[[i]] <- survey_lenwt[[i]]$natlength
+    survObsWtAtAge[[i]] <- survey_lenwt[[i]]$muweight
+  }
+
+  if(save){
+    saveRDS(survObsLenComp, file.path(d.name, paste0(scenario.name, "survObsLenComp.rds")))
+    saveRDS(survObsWtAtAge, file.path(d.name, paste0(scenario.name, "survObsWtAtAge.rds")))
+  }
+
+  #now do fishery comps
+  # user options for fishery--default is a census with mid-year sample
+  source(userfishery, local = TRUE)
+
+  #fishery catch at age each observed timestep summed over observed polygons
+  # catch at age by area and timestep
+  catch_numbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchnum_ss,
+                                                      time = fishtime,
+                                                      species = survspp,
+                                                      boxes = fishboxes)
+
+  # if we want replicates for obs error this sample function will generate them
+  catch_age_comp <- list()
+  for(i in 1:n_reps){
+    catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
+  }
+
+  # save fishery age comps
+  if(save){
+    saveRDS(catch_age_comp, file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
+  }
+
+  #Get catch weights for length comp calc
+  # aggregate true resn per fishery subset design
+  catch_aggresnss <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
+                                                      time = fishtime,
+                                                      species = survspp,
+                                                      boxes = fishboxes)
+
+  # aggregate true structn fishery subsetdesign
+  catch_aggstructnss <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
+                                                         time = fishtime,
+                                                         species = survspp,
+                                                         boxes = fishboxes)
+
+  #dont sample these, just aggregate them using median
+  catch_structnss <- atlantisom::sample_fish(catch_aggstructnss, fisheffN, sample = FALSE)
+
+  catch_resnss <- atlantisom::sample_fish(catch_aggresnss, fisheffN, sample = FALSE)
+
+  # these fishery lengths and weight at age are each output timestep
+  #same structure as above for surveys, replicates follow age comp reps
+  #  separating the length comps from the weight at age here
+  fishery_lenwt <- list()
+  fishObsLenComp <- list()
+  fishObsWtAtAge <- list()
+
+  for(i in 1:n_reps){
+    fishery_lenwt[[i]] <- atlantisom::calc_age2length(structn = catch_structnss,
+                                                      resn = catch_resnss,
+                                                      nums = catch_age_comp[[i]],
+                                                      biolprm = omlist_ss$biol, 
+                                                      fgs = omlist_ss$funct.group_ss,
+                                                      maxbin = maxbin,
+                                                      CVlenage = lenage_cv, 
+                                                      remove.zeroes=TRUE)
+
+    fishObsLenComp[[i]] <- fishery_lenwt[[i]]$natlength
+    fishObsWtAtAge[[i]] <- fishery_lenwt[[i]]$muweight
+  }
+
+  if(save){
+    saveRDS(fishObsLenComp, file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))
+    saveRDS(fishObsWtAtAge, file.path(d.name, paste0(scenario.name, "fishObsWtAtAge.rds")))
+  }
+
+  comps <- list("survObsAgeComp" = age_comp_data,
+                "survObsLenComp" = survObsLenComp,
+                "survObsWtAtAge" = survObsWtAtAge,
+                "fishObsAgeComp" = catch_age_comp,
+                "fishObsLenComp" = fishObsLenComp,
+                "fishObsWtAtAge" = fishObsWtAtAge)
+
+  return(comps)
+}
+
 ```
+
+Usage:
+```{r compex}
+
+CC3om_sard_comp <- om_comps(usersurvey = here("config/usersurvey.R"), 
+                           userfishery = here("config/fisherycensus.R"),
+                           omlist_ss = CC3om_sardine, 
+                           n_reps = 1, 
+                           save = TRUE)
+
+```

### 2020-01-17 config/fisherycensus.R

@@ -3,17 +3,22 @@
 fishspp <- omlist_ss$species_ss 
 
 # same time dimensioning parameters as in surveycensus.R
+# does the fishery sampling miss any timesteps? default is no
+fishtime <- timeall
 
 # fishery output: learned the hard way this can be different from ecosystem outputs
 fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
 
 # fishery sampling area
-# should return all model areas
+# should return all model areas, this assumes you see everything that it caught
 fishboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
 
 # effective sample size needed for sample_fish
 # this effective N is high but not equal to total for numerous groups
 fisheffN <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
+
+#this adjusts for subannual fishery output so original effN is for whole year
+fisheffN$effN <- fisheffN$effN/fstepperyr 
 
 # fishery catch cv can be used in sample_survey_biomass
 # perfect observation

### 2020-01-17 config/usersurvey.R

@@ -28,3 +28,10 @@ surveffN <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
 # perfect observation
 surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))
 
+# length at age cv for input into calc_age2length function
+# function designed to take one cv for all species, need to change to pass it a vector
+lenage_cv <- 0.1
+
+# max size bin for length estimation, function defaults to 150 cm if not supplied
+maxbin <- 150
+

### 2020-01-17 WrapperDevelopment.Rmd
test wrappers with changed config files

@@
 
-```{r exinit}
+```{r exinit, eval=FALSE}
 
@@
 
-```{r exsp}
+```{r exsp, eval=FALSE}
 
@@
 
-```{r exindex}
+```{r exindex, eval=FALSE}
 
@@
 Note that if a census is truly desired for age comps, don't use the sample fish function but just aggregate. I'll make a function for that too.
+
+ToDo: write the atlantisom function sample_ages to subset the age comp here, that way there will be way more lengths than ages, similar to most real world sampling.
 
@@
 Usage:
-```{r compex}
+```{r compex, eval=FALSE}
 
@@
 ```
+
+Test these with updated length-weight parameters for CCA sardines:
+
+I updated CalCurrentV3_Biol.prm with new sardine length-weight parameters from Isaac
+
+>On the basis of the last sardine stock assessment that seems to have estimated growth, 
+http://www.pcouncil.org/wp-content/uploads/2014_CPS_SAFE_Sardine_assessment_Appendix_C.pdf 
+>
+>I say length-weight A and B params,  for W = A*(L^B)
+>
+>should be A= 0.0075242
+>and B = 3.2332
+>
+>where weight is in grams
+>and length is in CM.
+
+and renamed it CalCurrentV3_Biol_sardLWcorrection.prm
+
+Config files for this extraction should match previous sardine survey and fishery:
+
+CC3Config_sardLWcorr.R
+sardinesurvey.R
+sardinefishery.R
+
+```{r full-test}
+
+library(here)
+
+CC3om <- om_init(here("config/CC3config_sardLWcorr.R"))
+
+CC3om_sardine <- om_species(c("Pacific_sardine"), CC3om)
+
+CC3om_sard_ind <- om_index(usersurvey = here("config/sardinesurvey.R"), 
+                           userfishery = here("config/sardinefishery.R"),
+                           omlist_ss = CC3om_sardine, 
+                           n_reps = 1, 
+                           save = TRUE)
+
+CC3om_sard_comp <- om_comps(usersurvey = here("config/sardinesurvey.R"), 
+                           userfishery = here("config/sardinefishery.R"),
+                           omlist_ss = CC3om_sardine, 
+                           n_reps = 1, 
+                           save = TRUE)
+
+```

### 2020-01-17 config/CC3Config_sardLWcorr.R

@@ -1,0 +1,10 @@
+d.name <- here("atlantisoutput","CC_2063_OA_OFF_22")
+functional.groups.file <- "CalCurrentV3Groups.csv"
+biomass.pools.file <- "DIVCalCurrentV3_Biol.nc"
+biol.prm.file <- "CalCurrentV3_Biol_sardLWcorrection.prm"
+box.file <- "CalCurrentV3_utm.bgm"
+initial.conditions.file <- "DIVCalCurrentV3_Biol.nc"
+run.prm.file <- "CalCurrentV3_run.xml"
+scenario.name <- "CCV3"
+bioind.file <- "outputCCV3BiomIndx.txt"
+catch.file <- "outputCCV3Catch.txt"

### 2020-01-17 WrapperDevelopment.Rmd
writing dat file part 1

@@
 ```
+
+### Write input data for specific assessment models
+
+Use the saved outputs of above to write a Stock Synthesis dat file
+
+```{r writedat}
+
+require(r4ss)
+#still needs omlist_ss and all this stuff
+source(here("/config/CC3Config_sardLWcorr.R"))
+source(here("/config/omdimensions.R"))
+source(here("/config/sardinesurvey.R"))
+source(here("/config/sardinefishery.R"))
+
+#Directory with SS files
+model_dir <- "Sardine_SS_files/"
+
+#Name of SS data file
+datfile_name <- "sardEM_3_3.dat"
+
+#CVs for length at age, catch, and survey
+CVs <- list("lenage"=0.1, "fishery"=0.01, "survey"=0.1)
+
+#Month of survey/fishing
+survey_month <- 7
+fishing_month <- 1
+
+#read time series data
+survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name, "surveyB.rds")))
+truecatchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
+
+#load dummy dat file
+stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/",
+model_dir,
+datfile_name))
+
+#Test writing CPUE in biomass
+  stocksynthesis.data <- atlantisom::SS_write_ts(ss_data_list = stocksynthesis.data,
+                ts_data = list(survObsBiom$atoutput[fish_years],
+                  truecatchbio_ss$atoutput[truecatchbio_ss$time %in% fish_years]),
+                CVs = c(CVs$survey,
+                        CVs$fishery),
+                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             fish_years),
+            sampling_month = list(rep(survey_month,nyears),
+                                rep(fishing_month,nyears)),
+                units = c("biomass","biomass"),
+                data_type=c("CPUE","catch"),
+                fleets = c(2,1))
+
+stocksynthesis.data$CPUE
+
+stocksynthesis.data$catch
+
+```

### 2020-01-17 WrapperDevelopment.Rmd
current wrapper and SS3dat output

@@
 Use the saved outputs of above to write a Stock Synthesis dat file
+This is not a function yet, copy in Christine's
 
@@
 require(r4ss)
-#still needs omlist_ss and all this stuff
+omlist_ss<-CC3om_sardine
 source(here("/config/CC3Config_sardLWcorr.R"))
@@
 
+#Years to survey, assuming survey is once per year
+#survey_years <- survey_sample_full[(burnin+1):(burnin+nyears)] # my original
+survey_years <- survey_sample_full[burnin:(burnin+nyears-1)] #from Christine's new sardine_config.R
+
 #read time series data
@@
 truecatchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
+
+survObsBiom <- survObsBiom[[1]]
+truecatchbio_ss <- truecatchbio_ss[[1]]
+
+#add this to om_indices function so that this has years when read in
+truecatchbio_ss$time <- as.integer(truecatchbio_ss$time/365)
 
@@
 
+#test  
 stocksynthesis.data$CPUE
@@
 
+#read in comp data
+age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
+age_comp_data <- age_comp_data[[1]]
+
+fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
+fish_age_comp <- fish_age_comp[[1]]
+
+#Get the age bins
+age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
+age_bins <- sub("a","",age_bin_names)
+
+require(maditr)
+# add dependency on maditr::dcast to atlantisom
+age_comp_flat <- atlantisom::reformat_compositions(age_comp_data,
+                     round.places = 4,
+                     comp_type = "agecomp")
+
+## Write age composition data for survey
+stocksynthesis.data <- atlantisom::SS_write_comps(ss_data_list = stocksynthesis.data,
+               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
+               data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr+nyears-1)),
+               sampling_month = list(rep(survey_month,nyears)),
+               data_type = c("agecomp"),
+               fleet_number = c(1),
+               bins = list(age_bins),
+               caal_bool = c(FALSE))
+stocksynthesis.data$agecomp
+
+#length comps
+len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsLenComp.rds")))
+fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))
+
+len_comp_data <- len_comp_data[[1]]
+fish_len_comp_data <- fish_len_comp_data[[1]]
+
+if(fstepperyr>1){
+  fish_len_comp_anndata <- fish_len_comp_data %>%
+    mutate(yr = floor(time/fstepperyr)) %>%
+    group_by(species, agecl, lower.bins, upper.bins, time=as.integer(yr)) %>%
+    summarise(annnatlength=sum(atoutput)) %>%
+    rename(atoutput = annnatlength)
+} else {
+  fish_len_comp_anndata <- fish_len_comp_data
+}
+
+caal_comp_flat <- atlantisom::reformat_compositions(len_comp_data,                                round.places=4,
+                comp_type="caalcomp")
+
+#remove burnin
+caal_comp_final <- filter(caal_comp_flat,
+                         time %in% survey_years)
+
+#Add over age classes to get sample size
+len_comp_flat <- atlantisom::reformat_compositions(len_comp_data,
+                                  round.places = 0,
+                           comp_type="lencomp")
+#remove burnin
+len_comp_final <- filter(len_comp_flat,
+                         time %in% survey_years)
+
+length_bins <- as.integer(names(len_comp_final))
+length_bins <- length_bins[!is.na(length_bins)]
+
+# fishery length comps are still 5 timesteps per year
+# need to aggregate to annual (done above)
+# also,  make effN annual goal/fstepsperyr (done above)
+fish_len_comp_flat <- atlantisom::reformat_compositions(fish_len_comp_anndata,
+                                  round.places = 0,
+                           comp_type="lencomp")
+
+#remove burnin works after adjustment above
+fish_len_comp_final <- filter(fish_len_comp_flat,
+                         time %in% fish_years)
+
+notbins <- c("time", "nsamp")
+
+# fish_length_bins <- as.integer(names(fish_len_comp_final))
+# fish_length_bins <- fish_length_bins[!is.na(fish_length_bins)]
+
+# need to fill empty length bins with 0s to have same bins as survey for SS_write_comps
+missing.lengths <- setdiff(length_bins, names(fish_len_comp_final)[!names(fish_len_comp_final) %in% notbins])
+fish_len_comp_final[as.character(missing.lengths)] <- 0                    # Add them, filled with '0's
+fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")]
+
+# fishery age comps also 5 timesteps per year
+if(fstepperyr>1){
+  fish_age_comp_anndata <- fish_age_comp %>%
+    mutate(yr = floor(time/fstepperyr)) %>%
+    group_by(species, agecl, time=as.integer(yr)) %>%
+    summarise(annnatage=sum(atoutput)) %>%
+    rename(atoutput = annnatage)
+} else {
+  fish_age_comp_anndata <- fish_age_comp
+}
+
+fish_age_comp_flat <- atlantisom::reformat_compositions(fish_age_comp_anndata,
+                           comp_type="agecomp")
+
+#remove burnin (not necessary?fish comps made with fish_years only) 
+fish_age_comp_final <- filter(fish_age_comp_flat,
+                         time %in% fish_years)
+
+# #SS_write_comps breaking because fishery age bins start with 2 not 1; extracting bins from fish file may help?
+# fish_age_bins <- names(fish_age_comp_flat)[!names(fish_age_comp_flat) %in% notbins]
+
+# that leaves an empty column in data file, so instead fill with 0s
+missing.ages <- setdiff(age_bins, names(fish_age_comp_final)[!names(fish_age_comp_final) %in% notbins])
+fish_age_comp_final[missing.ages] <- 0                    # Add them, filled with '0's
+fish_age_comp_final <- fish_age_comp_final[c("time", age_bins, "nsamp")]
+
+comp_list <- list(caal_comp_final,len_comp_final, fish_age_comp_final, fish_len_comp_final)
+
+apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), 
+                    rep(survey_month, nrow(comp_list[[2]])),
+                    rep(fishing_month,nrow(comp_list[[3]])),
+                    rep(fishing_month,nrow(comp_list[[4]])))
+
+# This now runs by ensuring that survey and fishery compositions have the same bins 
+# (filled with 0s for missing bins in fishery relative to survey)
+
+# Write CAAL and length composition data
+stocksynthesis.data <- atlantisom::SS_write_comps(ss_data_list = stocksynthesis.data, 
+                                      comp_matrix = comp_list, 
+                                      data_rows = list((comp_list[[1]]$time-survey_sample_time)/timestep + 1 , (survey_years-survey_sample_time)/timestep + 1,fish_years,fish_years),
+                                      sampling_month = apply_month, 
+                                      data_type = rep(c("agecomp", "lencomp"),2), 
+                                      fleet_number = c(2,2,1,1),  
+                                      bins = list(age_bins, 
+                                                  length_bins, 
+                                                  age_bins, 
+                                                  length_bins), 
+                                      caal_bool = c(TRUE, rep(FALSE,3)))
+
+head(stocksynthesis.data$lencomp)
+head(stocksynthesis.data$agecomp)
+
+#Change length bin structure to match atlantis data
+stocksynthesis.data$lbin_vector <- length_bins
+
+#Get correct number of length bins
+stocksynthesis.data$N_lbins <- length(length_bins)
+
+#Set lbin_method to 1 - this makes the population length bins match the data bins 
+#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
+stocksynthesis.data$lbin_method <- 1
+stocksynthesis.data$binwidth <- "#"
+stocksynthesis.data$minimum_size <- "#"
+stocksynthesis.data$maximum_size <- "#"
+
+#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
+stocksynthesis.data$age_info$minsamplesize <- rep(0.001,2)
+
+SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
+datfile_name),
+                 overwrite=TRUE)
+
 ```

### 2020-03-16 WrapperDevelopment.Rmd
still buggy
writing dat file now from proto-SS write wrapper

@@
 
-```{r ominit}
+```{r ominit, eval=FALSE}
 
@@
 library(here)
+library(tidyverse)
+library(atlantisom)
 CC3om <- om_init(here("config/CC3config.r"))
@@
 
-```{r inputs}
+```{r inputs, eval=FALSE}
 
@@
 
-```{r omindices}
+```{r omindices, eval=FALSE}
 
@@
 
-```{r omcomps}
+```{r omcomps, eval=FALSE}
 
@@
 library(here)
+library(tidyverse)
+library(atlantisom)
 
@@
 This is not a function yet, copy in Christine's
+
+**See corrections below, in fishery time fixed and nsamp added to length comps**
 
@@
 
+#add this to om_indices function so that this has years when read in
+fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))
+fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
+
 if(fstepperyr>1){
   fish_len_comp_anndata <- fish_len_comp_data %>%
-    mutate(yr = floor(time/fstepperyr)) %>%
-    group_by(species, agecl, lower.bins, upper.bins, time=as.integer(yr)) %>%
+    #mutate(yr = floor(time/fstepperyr)) %>%
+    #group_by(species, agecl, lower.bins, upper.bins, time=as.integer(yr)) %>%
+    group_by(species, agecl, lower.bins, upper.bins, time) %>% 
     summarise(annnatlength=sum(atoutput)) %>%
@@
 fish_len_comp_final[as.character(missing.lengths)] <- 0                    # Add them, filled with '0's
-fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")]
+fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")] #
 
@@
   fish_age_comp_anndata <- fish_age_comp %>%
-    mutate(yr = floor(time/fstepperyr)) %>%
-    group_by(species, agecl, time=as.integer(yr)) %>%
+    #mutate(yr = floor(time/fstepperyr)) %>%
+    #group_by(species, agecl, time=as.integer(yr)) %>%
+    group_by(species, agecl, time) %>%
     summarise(annnatage=sum(atoutput)) %>%

### 2020-03-18 NOBAcod.Rmd
initial commit NOBA full age testing

--- 2020-03-16 WrapperDevelopment.Rmd
@@
 ---
-title: "Develop atlantisom wrappers"
+title: "Generate data for NOBA cod ensemble modeling"
 author: "Sarah Gaichas and Christine Stawitz"
@@
 
-## Too many steps! and too much memory!
+## Standard atlantisom output test (age class not age)
 
-We can run atlantisom with many functions to get stock assessment model inputs.
+Let's use the wrapper functions to get initial stock assessment model inputs.
 
-To make life easier, here are some wrapper functions to do the basic things we want.
+Then we need to adjust the base atlantisom functions for true age classes, and develop options to call those within the wrappers.
 
-Basic desired functionality:
+We need config files for the NOBA model, our surveys and fisheries.
 
-1. initialize: locate model output (function to make a config file?), read in basic run parameters and species names, run_truth for all once and save output 
-    + User to make up front decision--want all species or just a subset?
-    + May want multiple species that need to be split in a later step for models
-    + wraps functions 
-        + sourcing model config file--add function to make config file?
-      + specifies top directory for source files, names key files for use later
-      + load_fgs
-      + load all other needed files--YOY, catch
-      + run_truth
-1. get assessment input data: configure survey, run all survey functions, configure fishery, run all fishery functions (how to save output?)
-    + Based on decision above, save interim steps for use later
-    + Need a switch for single vs multispecies estimation models?
-    + Separate index and comp functions in output
-    + indices:
-      + create survey config file(s), default is a census of all see census_spec.R
-        + survey index in biomass or numbers? (uses truth$biomass_ages vs truth$nums)
-        + wraps functions create_survey, sample_survey_biomass
-      + create fishery config file(s)
-    + comps:
-      + select age only or age and length
-        + age only wraps create_survey using nums with sample_fish to get age comps
-        + age and length (with weight at age)
-          + create_survey followed by sample_fish, 
-          + aggregateDensityData for resn and structn followed by sample_fish
-          + inputs all to calc_age2length
-      + catch comps wrap create_fishery_subset with sample_fish to get age comps, same as above for lengths
-1. write input data for specific model: wrap functions in development for SS3, add functions for other models (ASAP, multispecies models)
-    + wrap r4ss functions to read and write dat and ctl files
-    + possibly start with set of dummy SS files with package (take from ss3sim?)
-      + read in dat file
-      + indices wrap SS_write_ts
-      + comps wrap reformat_compositions, SS_write_comps, some of the checks for bin matches
-      + write full dat file
-    + start with dummy ctl file (any from ss3sim?)
-      + read in ctl file
-      + wrap SS_write_bio using outputs of calc_Z, load_YOY, load_biolprm, load_runprm, etc
-1. run specified model--automate for scenarios, save outputs 
-1. skill assessment metrics: compare stored Atlantis truth with model output
+`NOBAconfig.R` looks like this (adjusted from Alfonso's original):
 
-### atlantisom initialize function
+---
+d.name <- here("atlantisoutput","NOBA_march_2020")
+functional.groups.file <- "nordic_groups_v04.csv"
+biomass.pools.file <- "nordic_biol_v23.nc"
+biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
+box.file <- "Nordic02.bgm"
+initial.conditions.file <- "nordic_biol_v23.nc"
+run.prm.file <- "nordic_run_v01.xml"
+scenario.name <- "nordic_runresults_01"
+bioind.file <- "nordic_runresults_01BiomIndx.txt"
+catch.file <- "nordic_runresults_01Catch.txt"
+---
 
-Things to add: checks for output timestep?
+`codsurvey.R` configures the fishery independent survey and looks like this:
 
-```{r ominit, eval=FALSE}
+---
+timeall <- c(0:560)
 
-library(tidyverse)
+funct.group.indices <- c(1:36)
+---
 
-om_init <- function(config = configfile){
+`codfishery.R` configures the fishery dependent data and looks like this:
 
-  # Where are the atlantis output files? Consider filling with shiny app in future
-  source(config)
-  # needs these files, for example config file CC3config.R is:
+---
 
-  # d.name <- here::here("atlantisoutput","CC_2063_OA_OFF_22")
-  # functional.groups.file <- "CalCurrentV3Groups.csv"
-  # biomass.pools.file <- "DIVCalCurrentV3_Biol.nc"
-  # biol.prm.file <- "CalCurrentV3_Biol.prm"
-  # box.file <- "CalCurrentV3_utm.bgm"
-  # initial.conditions.file <- "DIVCalCurrentV3_Biol.nc"
-  # run.prm.file <- "CalCurrentV3_run.xml"
-  # scenario.name <- "CCV3"
-  # bioind.file <- "outputCCV3BiomIndx.txt"
-  # catch.file <- "outputCCV3Catch.txt"
+---
 
-  #Load functional groups
-  funct.groups <- atlantisom::load_fgs(dir=d.name,
-                           file_fgs = functional.groups.file)
-  #Get just the names of active functional groups
-  funct.group.names <- funct.groups %>% 
-    filter(IsTurnedOn == 1) %>%
-    select(Name) %>%
-    .$Name
-
-  # load true total biomass in tons
-  truetotbio <- atlantisom::load_bioind(d.name, file_bioind = bioind.file, fgs = funct.groups)
-
-  # load true catch in tons
-  truecatchbio <- atlantisom::load_catch(d.name, file_catch = catch.file, fgs = funct.groups)
-
-  # load YOY
-  YOY <- atlantisom::load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))
-
-  # load biol_prm
-  biol <- atlantisom::load_biolprm(d.name, biol.prm.file)
-
-  # load run_prm
-  runpar <- atlantisom::load_runprm(d.name, run.prm.file)
-
-  # load box info file
-  boxpars <- atlantisom::load_box(d.name, box.file)
-
-  # default run_truth setup will save the file, so check for that first
-  if(!file.exists(file.path(d.name, 
-                            paste0("output", scenario.name, "run_truth.RData")))){
-    #Store all loaded results into an R object
-    truth <- atlantisom::run_truth(scenario = scenario.name,
-                       dir = d.name,
-                       file_fgs = functional.groups.file,
-                       file_bgm = box.file,
-                       select_groups = funct.group.names,
-                       file_init = initial.conditions.file,
-                       file_biolprm = biol.prm.file,
-                       file_runprm = run.prm.file,
-                       verbose = TRUE
-    )
-  } else {
-    truth <- get(load(file.path(d.name,
-                                paste0("output", scenario.name, "run_truth.RData"))))
-  }
-
-  omlist <-list("funct.groups" = funct.groups, 
-                "funct.group.names" = funct.group.names,
-                "truetotbio" = truetotbio,
-                "truecatchbio" = truecatchbio,
-                "YOY" = YOY,
-                "biol" = biol,
-                "runpar" = runpar,
-                "boxpars" = boxpars,
-                "truth" = truth)
-
-  return(omlist)
-}
-
-```
-
-Usage:
-
-```{r exinit, eval=FALSE}
-
-library(here)
-library(tidyverse)
-library(atlantisom)
-CC3om <- om_init(here("config/CC3config.r"))
-
-```
-
-### Get assessment input data
-
-Split to single species or subset of assessed species, get index data, get compositional data. By default, remove full om results and keep only species subsets to save memory.
-
-#### Split to focal species
-
-Is this where we should reconcile timesteps? depends what functions are expecting.
-
-```{r inputs, eval=FALSE}
-
-om_species <- function(species = spp, omlist, removefullom = TRUE){
-  # spp format c("speciesname1", "speciesname2")
-  if(!all(species %in% omlist$funct.group.names)) stop("species name not found") 
-  species_ss <- species
-
-  #subset species true bio
-  truetotbio_ss <- omlist$truetotbio[omlist$truetotbio$species %in% species_ss,]
-
-  #subset species true catch
-  truecatchbio_ss <- omlist$truecatchbio[omlist$truecatchbio$species %in% species_ss,]
-
-  #subset species YOY
-  # get code matching species name to split YOY file
-  code_ss <- omlist$funct.groups$Code[which(omlist$funct.groups$Name %in% species_ss)]
-  # cut to a single species in YOY file
-  YOY_ss <- omlist$YOY %>%
-    select(Time, paste0(code_ss, ".0"))
-  # reformat to be like all the other objects
-
-  # numbers at agecl at full resolution (all polygons and layers)
-  truenums_ss <- omlist$truth$nums[omlist$truth$nums$species %in% species_ss,]
-
-  # biomass at agecl at full resolution (all polygons and layers)
-  truebio_ss <- omlist$truth$biomass_ages[omlist$truth$biomass_ages$species %in% species_ss,]
-
-  # reserve nitrogen at agecl at full resolution
-  trueresn_ss <- omlist$truth$resn[omlist$truth$resn$species %in% species_ss,]
-
-  # structural nitrogen at agecl at full resolution
-  truestructn_ss <- omlist$truth$structn[omlist$truth$structn$species %in% species_ss,]
-
-  # catch in numbers at agecl at full resoluation (all polygons, no layer in output)
-  truecatchnum_ss <- omlist$truth$catch[omlist$truth$catch$species %in% species_ss,]
-
-  #subset species biol parameters? no, may miss some globals that aren't by species
-
-  #subset species functional group info
-  funct.groups_ss <- omlist$funct.groups[omlist$funct.groups$Code %in% code_ss,]
-
-  #keep the runpar and also need boxes for survey selection
-
-  omlist_ss <- list("species_ss" = species_ss,
-                    "code_ss" = code_ss,
-                    "truetotbio_ss" = truetotbio_ss,
-                    "truecatchbio_ss" = truecatchbio_ss,
-                    "YOY_ss" = YOY_ss,
-                    "truenums_ss" = truenums_ss,
-                    "truebio_ss" = truebio_ss,
-                    "trueresn_ss" = trueresn_ss,
-                    "truestructn_ss" = truestructn_ss,
-                    "truecatchnum_ss" = truecatchnum_ss,
-                    "funct.group_ss" = funct.groups_ss,
-                    "biol" = omlist$biol,
-                    "boxpars" = omlist$boxpars,
-                    "runpar" = omlist$runpar)
-
-  if(removefullom) rm(omlist) #not removing passed data object
-
-  return(omlist_ss)
-}
-
-```
-
-Usage:
-
-```{r exsp, eval=FALSE}
-
-CC3om_sardine <- om_species(c("Pacific_sardine"), CC3om)
-
-```
-
-#### Generate index data for assessment
-
-The hardcoded script omdimensions.R needs to sit in the atlantisom config folder.
-
-```{r omindices, eval=FALSE}
-
-om_index <- function(usersurvey = usersurvey_file, 
-                     userfishery = userfishery_file, 
-                     omlist_ss, 
-                     n_reps = n_reps,
-                     save = TRUE){
-
-  #one script for dimension parameters to be used in multiple functions
-  source("config/omdimensions.R", local = TRUE)
-
-  # user options for survey--default is a census with mid-year sample
-  source(usersurvey, local = TRUE)
-
-  #biomass based fishery independent survey index
-  # this uses result$biomass_ages to sample biomass directly, no need for wt@age est
-  survey_B <- atlantisom::create_survey(dat = omlist_ss$truebio_ss,
-                                        time = survtime,
-                                        species = survspp,
-                                        boxes = survboxes,
-                                        effic = surveffic,
-                                        selex = survselex)
-
-  # call sample_survey_biomass with a bunch of 1000s for weight at age
-  # in the code it multiplies atoutput by wtatage/1000 so this allows us to use
-  # biomass directly
-  wtage <- data.frame(species=rep(survspp, each=max(age_classes)),
-                      agecl=rep(c(1:max(age_classes)),length(survspp)),
-                      wtAtAge=rep(1000.0,length(survspp)*max(age_classes)))
-
-  # this is the step to repeat n_reps time if we want different realizations 
-  # of the same survey design specified above; only observation error differs
-  # using the census cv of 0 will produce identical reps!
-  survObsBiomB <- list()
-  for(i in 1:n_reps){
-    survObsBiomB[[i]] <- atlantisom::sample_survey_biomass(survey_B, surv_cv, wtage)
-  }
-
-  #save survey indices, takes a long time to generate with lots of reps/species
-  if(save){
-    saveRDS(survObsBiomB, file.path(d.name, paste0(scenario.name, "surveyB.rds")))
-  }
-
-  #configure the fishery, a default is in config/fisherycensus.R
-  #fishery configuration can specify only area and time of observation
-  #fishery species inherited from omlist_ss
-  source(userfishery, local = TRUE)
-
-  #we are not currently subsetting fishery catch because we cannot correct catch.nc
-  #  instead the catch in biomass from catch.txt is read in for the index
-  #  we do not apply any cv to this, but we could this way (default cv=0)
-
-  fishObsCatchB <- list()
-  for(i in 1:n_reps){
-    fishObsCatchB[[i]] <- atlantisom::sample_fishery_totcatch(omlist_ss$truecatchbio_ss, fish_cv)
-  }
-
-  if(save){
-    saveRDS(fishObsCatchB, file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
-  }
-
-  indices <- list("survObsBiomB" = survObsBiomB,
-                  "fishObsCatchB" = fishObsCatchB)
-
-  return(indices)
-}
-
-```
-
-Usage:
-
-```{r exindex, eval=FALSE}
-
-CC3om_sard_ind <- om_index(usersurvey = here("config/usersurvey.R"), 
-                           userfishery = here("config/fisherycensus.R"),
-                           omlist_ss = CC3om_sardine, 
-                           n_reps = 5, 
-                           save = TRUE)
-
-```
-
-#### Generate compositional data for assessment
-
-Note that if a census is truly desired for age comps, don't use the sample fish function but just aggregate. I'll make a function for that too.
-
-ToDo: write the atlantisom function sample_ages to subset the age comp here, that way there will be way more lengths than ages, similar to most real world sampling.
-
-```{r omcomps, eval=FALSE}
-
-om_comps <- function(usersurvey = usersurvey_file, 
-                     userfishery = userfishery_file, 
-                     omlist_ss, 
-                     n_reps = n_reps,
-                     save = TRUE){
-
-  #one script for dimension parameters to be used in multiple functions
-  source("config/omdimensions.R", local = TRUE)
-
-  # user options for survey--default is a census with mid-year sample
-  source(usersurvey, local = TRUE)
-
-  #numbers based fishery independent survey for age and length comps
-  # same user specifications as indices
-  survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
-                                        time = survtime,
-                                        species = survspp,
-                                        boxes = survboxes,
-                                        effic = surveffic,
-                                        selex = survselex)
-
-  #Sample fish for age composition
-  # if we want replicates for obs error this sample function will generate them
-  age_comp_data <- list()
-  for(i in 1:n_reps){
-    age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
-  }
-
-  # save age comps
-  if(save){
-    saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
-  }
-
-  #weights needed for weight at age and length comp calcs
-  # aggregate true resn per survey design
-  survey_aggresn <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
-                                                     time = survtime,
-                                                     species = survspp,
-                                                     boxes = survboxes)
-
-  # aggregate true structn per survey design
-  survey_aggstructn <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
-                                                        time = survtime,
-                                                        species = survspp,
-                                                        boxes = survboxes)
-
-  #dont sample these, just aggregate them using median
-  structnss <- atlantisom::sample_fish(survey_aggresn, surveffN, sample = FALSE)
-
-  resnss <- atlantisom::sample_fish(survey_aggstructn, surveffN, sample = FALSE)
-
-  #this is all input into the length function, replicates follow age comp reps
-  #  separating the length comps from the weight at age here
-  survey_lenwt <- list()
-  survObsLenComp <- list()
-  survObsWtAtAge <- list()
-
-  for(i in 1:n_reps){
-    survey_lenwt[[i]] <- atlantisom::calc_age2length(structn = structnss,
-                                                     resn = resnss,
-                                                     nums = age_comp_data[[i]],
-                                                     biolprm = omlist_ss$biol, 
-                                                     fgs = omlist_ss$funct.group_ss,
-                                                     maxbin = maxbin,
-                                                     CVlenage = lenage_cv, 
-                                                     remove.zeroes=TRUE)
-
-    survObsLenComp[[i]] <- survey_lenwt[[i]]$natlength
-    survObsWtAtAge[[i]] <- survey_lenwt[[i]]$muweight
-  }
-
-  if(save){
-    saveRDS(survObsLenComp, file.path(d.name, paste0(scenario.name, "survObsLenComp.rds")))
-    saveRDS(survObsWtAtAge, file.path(d.name, paste0(scenario.name, "survObsWtAtAge.rds")))
-  }
-
-  #now do fishery comps
-  # user options for fishery--default is a census with mid-year sample
-  source(userfishery, local = TRUE)
-
-  #fishery catch at age each observed timestep summed over observed polygons
-  # catch at age by area and timestep
-  catch_numbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchnum_ss,
-                                                      time = fishtime,
-                                                      species = survspp,
-                                                      boxes = fishboxes)
-
-  # if we want replicates for obs error this sample function will generate them
-  catch_age_comp <- list()
-  for(i in 1:n_reps){
-    catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
-  }
-
-  # save fishery age comps
-  if(save){
-    saveRDS(catch_age_comp, file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
-  }
-
-  #Get catch weights for length comp calc
-  # aggregate true resn per fishery subset design
-  catch_aggresnss <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
-                                                      time = fishtime,
-                                                      species = survspp,
-                                                      boxes = fishboxes)
-
-  # aggregate true structn fishery subsetdesign
-  catch_aggstructnss <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
-                                                         time = fishtime,
-                                                         species = survspp,
-                                                         boxes = fishboxes)
-
-  #dont sample these, just aggregate them using median
-  catch_structnss <- atlantisom::sample_fish(catch_aggstructnss, fisheffN, sample = FALSE)
-
-  catch_resnss <- atlantisom::sample_fish(catch_aggresnss, fisheffN, sample = FALSE)
-
-  # these fishery lengths and weight at age are each output timestep
-  #same structure as above for surveys, replicates follow age comp reps
-  #  separating the length comps from the weight at age here
-  fishery_lenwt <- list()
-  fishObsLenComp <- list()
-  fishObsWtAtAge <- list()
-
-  for(i in 1:n_reps){
-    fishery_lenwt[[i]] <- atlantisom::calc_age2length(structn = catch_structnss,
-                                                      resn = catch_resnss,
-                                                      nums = catch_age_comp[[i]],
-                                                      biolprm = omlist_ss$biol, 
-                                                      fgs = omlist_ss$funct.group_ss,
-                                                      maxbin = maxbin,
-                                                      CVlenage = lenage_cv, 
-                                                      remove.zeroes=TRUE)
-
-    fishObsLenComp[[i]] <- fishery_lenwt[[i]]$natlength
-    fishObsWtAtAge[[i]] <- fishery_lenwt[[i]]$muweight
-  }
-
-  if(save){
-    saveRDS(fishObsLenComp, file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))
-    saveRDS(fishObsWtAtAge, file.path(d.name, paste0(scenario.name, "fishObsWtAtAge.rds")))
-  }
-
-  comps <- list("survObsAgeComp" = age_comp_data,
-                "survObsLenComp" = survObsLenComp,
-                "survObsWtAtAge" = survObsWtAtAge,
-                "fishObsAgeComp" = catch_age_comp,
-                "fishObsLenComp" = fishObsLenComp,
-                "fishObsWtAtAge" = fishObsWtAtAge)
-
-  return(comps)
-}
-
-```
-
-Usage:
-```{r compex, eval=FALSE}
-
-CC3om_sard_comp <- om_comps(usersurvey = here("config/usersurvey.R"), 
-                           userfishery = here("config/fisherycensus.R"),
-                           omlist_ss = CC3om_sardine, 
-                           n_reps = 1, 
-                           save = TRUE)
-
-```
-
-Test these with updated length-weight parameters for CCA sardines:
-
-I updated CalCurrentV3_Biol.prm with new sardine length-weight parameters from Isaac
-
->On the basis of the last sardine stock assessment that seems to have estimated growth, 
-http://www.pcouncil.org/wp-content/uploads/2014_CPS_SAFE_Sardine_assessment_Appendix_C.pdf 
->
->I say length-weight A and B params,  for W = A*(L^B)
->
->should be A= 0.0075242
->and B = 3.2332
->
->where weight is in grams
->and length is in CM.
-
-and renamed it CalCurrentV3_Biol_sardLWcorrection.prm
-
-Config files for this extraction should match previous sardine survey and fishery:
-
-CC3Config_sardLWcorr.R
-sardinesurvey.R
-sardinefishery.R
+All the config files go in the config folder and we'll try running this:
 
@@
 
-CC3om <- om_init(here("config/CC3config_sardLWcorr.R"))
+NOBAom <- om_init(here("config/NOBAconfig.R"))
 
-CC3om_sardine <- om_species(c("Pacific_sardine"), CC3om)
+NOBAom_cod <- om_species(c("Pacific_sardine"), NOBAom)
 
-CC3om_sard_ind <- om_index(usersurvey = here("config/sardinesurvey.R"), 
+NOBAom_cod_ind <- om_index(usersurvey = here("config/sardinesurvey.R"), 
                            userfishery = here("config/sardinefishery.R"),
@@
 
-CC3om_sard_comp <- om_comps(usersurvey = here("config/sardinesurvey.R"), 
+NOBAom_cod_comp <- om_comps(usersurvey = here("config/sardinesurvey.R"), 
                            userfishery = here("config/sardinefishery.R"),

### 2020-03-18 NOBAcod.Rmd
start on wrapping NOBA cod

@@
 title: "Generate data for NOBA cod ensemble modeling"
-author: "Sarah Gaichas and Christine Stawitz"
+author: "Sarah Gaichas"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-We need config files for the NOBA model, our surveys and fisheries.
+We need config files for the NOBA model, our surveys and fisheries. Surveys and fisheries are based on the ones we used for sardine in the California Current as a test.
 
-`NOBAconfig.R` looks like this (adjusted from Alfonso's original):
+*Special note: Atlantis output files usually have the prefix "output" and `atlantisom` expects this.
+NOBA output files use a different naming convention.
+I've appended "output" to the beginning of each NOBA output filename for `atlantisom` to work.*
 
----
+`NOBA2config.R` looks like this (adjusted from Alfonso's original):
+
+```
 d.name <- here("atlantisoutput","NOBA_march_2020")
@@
 scenario.name <- "nordic_runresults_01"
-bioind.file <- "nordic_runresults_01BiomIndx.txt"
-catch.file <- "nordic_runresults_01Catch.txt"
----
+bioind.file <- "outputnordic_runresults_01BiomIndx.txt"
+catch.file <- "outputnordic_runresults_01Catch.txt"
+```
+
+`omdimensions.R` standardizes timesteps, etc.:
+
+```
+#survey species inherited from omlist_ss
+survspp <- omlist_ss$species_ss
+# survey season and other time dimensioning parameters
+# generalized timesteps all models
+noutsteps <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
+timeall <- c(0:noutsteps)
+stepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutinc
+midptyr <- round(median(seq(0,stepperyr)))
+
+# model areas, subset in surveyconfig
+allboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
+
+# fishery output: learned the hard way this can be different from ecosystem outputs
+fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
+
+# survey selectivity (agecl based)
+# should return all age classes fully sampled (Atlantis output is 10 age groups per spp)
+n_age_classes <- omlist_ss$funct.group_ss$NumCohorts
+age_classes <- 1:n_age_classes
+
+n_annages <- n_age_classes * omlist_ss$funct.group_ss$NumAgeClassSize
+annages <- 1:n_annages
+
+```
 
@@
 
----
-timeall <- c(0:560)
+```
+# Default survey configuration here is a census
+# Need to define survey season, area, efficiency, selectivity
 
-funct.group.indices <- c(1:36)
----
+#Atlantis model timestep corresponding to the true output--now from census_spec.R
+timestep <- stepperyr #5
+
+#Which atlantis timestep does the survey run in?--now from census_spec.R
+# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
+# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
+
+survey_sample_time <- midptyr #2; defined in omdimensions.R
+
+#The last timestep to sample
+total_sample <- noutsteps-1 #495
+
+#Vector of indices of survey times to pull
+survey_sample_full <- seq(survey_sample_time,
+                          total_sample, by=timestep)
+
+survtime <- survey_sample_full
+
+# survey area
+# should return all model areas
+survboxes <- allboxes
+
+# survey efficiency (q)
+# should return a perfectly efficient survey 
+surveffic <- data.frame(species=survspp,
+                     efficiency=rep(0.5,length(survspp)))
+
+# survey selectivity (agecl based)
+# this is by age class, need to change to use with ANNAGEBIO output
+survselex <- data.frame(species=rep(survspp, each=n_age_classes),
+                     agecl=rep(c(1:n_age_classes),length(survspp)),
+                     selex=rep(1.0,length(survspp)*n_age_classes))
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+
+# survey index cv needed for sample_survey_xxx
+# cv = 0.1
+surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp)))
+
+# length at age cv for input into calc_age2length function
+# function designed to take one cv for all species, need to change to pass it a vector
+lenage_cv <- 0.1
+
+# max size bin for length estimation, function defaults to 150 cm if not supplied
+maxbin <- 150
+
+```
 
@@
 
----
+```
+# Default fishery configuration here is a census
+# Inherits species from input omlist_ss
+fishspp <- omlist_ss$species_ss
 
----
+#Number of years of data to pull
+nyears <- 50
+
+#Atlantis initialization period in years
+burnin <- 30
+
+# same time dimensioning parameters as in surveycensus.R
+#Vector of indices of catch in numbers to pull (by timestep to sum)
+fish_sample_full <- c(0:total_sample)  #total_sample defined in sardinesurvey.R
+fish_burnin <- burnin*fstepperyr+1
+fish_nyears <- nyears*fstepperyr
+fish_times <- fish_sample_full[fish_burnin:(fish_burnin+fish_nyears-1)]
+fish_timesteps <- seq(fish_times[5], max(fish_times), by=fstepperyr) #last timestep
+#fish_years <- unique(floor(fish_times/fstepperyr)+1) # my original
+fish_years <- unique(floor(fish_times/fstepperyr)) #from Christine's new sardine_config.R
+
+fishtime <- fish_times
+
+# fishery output: learned the hard way this can be different from ecosystem outputs
+fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
+
+# fishery sampling area
+# should return all model areas, this assumes you see everything that it caught
+fishboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+fisheffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+
+#this adjusts for subannual fishery output so original effN is for whole year
+fisheffN$effN <- fisheffN$effN/fstepperyr
+
+# fishery catch cv can be used in sample_survey_biomass
+# perfect observation
+fish_cv <- data.frame(species=survspp, cv=rep(0.01,length(survspp)))
+```
 
@@
 
-```{r full-test}
+```{r full-test, eval=FALSE}
 
@@
 
-NOBAom <- om_init(here("config/NOBAconfig.R"))
+NOBAom <- om_init(here("config/NOBA2config.R"))
 
-NOBAom_cod <- om_species(c("Pacific_sardine"), NOBAom)
+NOBAom_cod <- om_species(c("North_atl_cod"), NOBAom)
 
-NOBAom_cod_ind <- om_index(usersurvey = here("config/sardinesurvey.R"), 
-                           userfishery = here("config/sardinefishery.R"),
-                           omlist_ss = CC3om_sardine, 
+NOBAom_cod_ind <- om_index(usersurvey = here("config/codsurvey.R"), 
+                           userfishery = here("config/codfishery.R"),
+                           omlist_ss = NOBAom_cod, 
                            n_reps = 1, 
@@
                            userfishery = here("config/sardinefishery.R"),
-                           omlist_ss = CC3om_sardine, 
+                           omlist_ss = NOBAom_cod, 
                            n_reps = 1, 
@@
 
-Use the saved outputs of above to write a Stock Synthesis dat file
-This is not a function yet, copy in Christine's
+We will need to alter this for SAM, SS3, etc. Right now just makes SS3.
 
-**See corrections below, in fishery time fixed and nsamp added to length comps**
+**THIS IS STILL THE CALIFORNIA CURRENT INFO, TO BE UPDATED**
 
-```{r writedat}
+```{r writedat, eval=FALSE}
 

### 2020-03-18 docs/index.Rmd
begin wrapper functions
wrappers to initialize and split species so far
test wrappers with changed config files
writing dat file now from proto-SS write wrapper
start on wrapping NOBA cod

@@
 Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
+
+### Apply wrapper to Norwegian Barents cod, March 2020
+
+<a href="https://sgaichas.github.io/poseidon-dev/NOBAcod.html" target="_blank">Apply and extend wrappers to get full age structure, started March 18, 2020</a>
+
+### Working on wrapper functions, January 2020
+
+<a href="https://sgaichas.github.io/poseidon-dev/WrapperDevelopment.html" target="_blank">Automating om functions to streamline workflows, January 17, 2020 writes dat file as of March 16, 2020</a>
 

### 2020-03-18 config/NOBA2Config.R

@@ -1,0 +1,10 @@
+d.name <- here("atlantisoutput","NOBA_march_2020")
+functional.groups.file <- "nordic_groups_v04.csv"
+biomass.pools.file <- "nordic_biol_v23.nc"
+biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
+box.file <- "Nordic02.bgm"
+initial.conditions.file <- "nordic_biol_v23.nc"
+run.prm.file <- "nordic_run_v01.xml"
+scenario.name <- "nordic_runresults_01"
+bioind.file <- "outputnordic_runresults_01BiomIndx.txt"
+catch.file <- "outputnordic_runresults_01Catch.txt"

### 2020-03-18 config/codfishery.R

@@ -1,0 +1,41 @@
+# Default fishery configuration here is a census
+# Inherits species from input omlist_ss
+fishspp <- omlist_ss$species_ss 
+
+#Number of years of data to pull
+nyears <- 50
+
+#Atlantis initialization period in years
+burnin <- 30
+
+
+# same time dimensioning parameters as in surveycensus.R
+#Vector of indices of catch in numbers to pull (by timestep to sum)
+fish_sample_full <- c(0:total_sample)  #total_sample defined in sardinesurvey.R
+fish_burnin <- burnin*fstepperyr+1
+fish_nyears <- nyears*fstepperyr
+fish_times <- fish_sample_full[fish_burnin:(fish_burnin+fish_nyears-1)]
+fish_timesteps <- seq(fish_times[5], max(fish_times), by=fstepperyr) #last timestep
+#fish_years <- unique(floor(fish_times/fstepperyr)+1) # my original
+fish_years <- unique(floor(fish_times/fstepperyr)) #from Christine's new sardine_config.R
+
+fishtime <- fish_times
+
+# fishery output: learned the hard way this can be different from ecosystem outputs
+fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
+
+# fishery sampling area
+# should return all model areas, this assumes you see everything that it caught
+fishboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+fisheffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+
+#this adjusts for subannual fishery output so original effN is for whole year
+fisheffN$effN <- fisheffN$effN/fstepperyr 
+
+# fishery catch cv can be used in sample_survey_biomass
+# perfect observation
+fish_cv <- data.frame(species=survspp, cv=rep(0.01,length(survspp)))
+

### 2020-03-18 config/codsurvey.R

@@ -1,0 +1,51 @@
+# Default survey configuration here is a census
+# Need to define survey season, area, efficiency, selectivity
+
+#Atlantis model timestep corresponding to the true output--now from census_spec.R
+timestep <- stepperyr #5
+
+#Which atlantis timestep does the survey run in?--now from census_spec.R
+# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
+# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
+
+survey_sample_time <- midptyr #2; defined in omdimensions.R
+
+#The last timestep to sample
+total_sample <- noutsteps-1 #495
+
+#Vector of indices of survey times to pull
+survey_sample_full <- seq(survey_sample_time,
+                          total_sample, by=timestep)
+
+survtime <- survey_sample_full
+
+# survey area
+# should return all model areas
+survboxes <- allboxes
+
+# survey efficiency (q)
+# should return a perfectly efficient survey 
+surveffic <- data.frame(species=survspp,
+                     efficiency=rep(0.5,length(survspp)))
+
+# survey selectivity (agecl based)
+# this is by age class, need to change to use with ANNAGEBIO output
+survselex <- data.frame(species=rep(survspp, each=n_age_classes),
+                     agecl=rep(c(1:n_age_classes),length(survspp)),
+                     selex=rep(1.0,length(survspp)*n_age_classes))
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+
+# survey index cv needed for sample_survey_xxx
+# cv = 0.1
+surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp)))
+
+# length at age cv for input into calc_age2length function
+# function designed to take one cv for all species, need to change to pass it a vector
+lenage_cv <- 0.1
+
+# max size bin for length estimation, function defaults to 150 cm if not supplied
+maxbin <- 150
+

### 2020-03-19 NOBAcod.Rmd
formatting
NOBA cod working with age class wrappers

@@
 
-*Special note: Atlantis output files usually have the prefix "output" and `atlantisom` expects this.
+**Special note:** *Atlantis output files usually have the prefix "output" and `atlantisom` expects this.
 NOBA output files use a different naming convention.
-I've appended "output" to the beginning of each NOBA output filename for `atlantisom` to work.*
+I've appended "output" to the beginning of each NOBA output filename for `atlantisom` to work.
+Need to update `atlantisom` functions for user-specified output filenames.*
+
+**Special note 2:** *NOBA's nordic_groups_v04.csv file has an inFishery column, rather than the `atlantisom` expected InFishery column.
+I've changed the column name in the input file, but need to add robustness to atlantisom code to remove case sensitivity.*
+
+### Config files
 
@@
 
+### Using the wrappers
+
 All the config files go in the config folder and we'll try running this:
@@
 
-### Write input data for specific assessment models
+### Wrapper test results {.tabset}
+
+#### Biomass index
+```{r bioind-1}
+
+library(ggthemes)
+omlist_ss <- NOBAom_cod
+source(here("config/omdimensions.R"))
+
+#read time series data
+survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name, "surveyB.rds")))
+
+survObsBiom <- survObsBiom[[1]]
+
+plotB <-ggplot() +
+  geom_line(data=survObsBiom, aes(x=time/stepperyr,y=atoutput, color="survey Biomass"), 
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotB +
+  facet_wrap(~species, scales="free")
+
+```
+
+#### Catch time series
+```{r catchind-1}
+#read time series data
+catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
+
+catchbio_ss <- catchbio_ss[[1]]
+
+plotC <-ggplot() +
+  geom_line(data=catchbio_ss, aes(x=time/365,y=atoutput, color="observed Catch"), 
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotC +
+  facet_wrap(~species, scales="free")
+
+```
+
+#### Survey length composition
+```{r lencomp-1}
+
+#length comps
+len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsLenComp.rds")))
+fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))
+
+len_comp_data <- len_comp_data[[1]]
+fish_len_comp_data <- fish_len_comp_data[[1]]
+
+#add this to om_indices function so that this has years when read in
+fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))
+
+len <- filter(len_comp_data, time %in% c(55:175))
+
+  lfplot <- ggplot(len, aes(upper.bins)) +
+    geom_bar(aes(weight = atoutput)) +
+    theme_tufte() +
+    labs(subtitle = paste(scenario.name,
+                          len$species))
+
+  lfplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+
+```
+
+#### Age composition (age classes)
+```{r agecomp-1}
+
+#read in comp data
+age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
+age_comp_data <- age_comp_data[[1]]
+
+fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
+fish_age_comp <- fish_age_comp[[1]]
+
+#add this to om_indices function so that this has years when read in
+fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
+
+Natageplot <- ggplot(Natagesubsp, aes(x=agecl, y=atoutput)) +
+    geom_point() +
+    theme_tufte() +
+    labs(subtitle = paste(scenario.name,
+                          sp))
+
+  Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
+  Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
+  Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
+  Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+```
+
+### {-}
+
+Wrappers work for NOBA cod using standard age classes. Now to modify the underlying code to get full age structure.
+
+## Updating functions for full age structure
+
+## Write input data for specific assessment models
 

### 2020-03-20 NOBAcod.Rmd
test plots for NOBA from standard outputs

@@
 
-```{r full-test, eval=FALSE}
+```{r full-test}
 
@@
 
-Natageplot <- ggplot(Natagesubsp, aes(x=agecl, y=atoutput)) +
+Natage <- filter(age_comp_data, time %in% c(55:175))
+
+Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
     geom_point() +
@@
     labs(subtitle = paste(scenario.name,
-                          sp))
+                          Natage$species))
 
-  Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
-  Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
-  Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
-  Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")
+  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
 ```
@@
 ## Updating functions for full age structure
+
+NEXT!!
 

### 2020-03-20 NOBAcod.Rmd
outline strategy for annual ages

@@
 
-NEXT!!
+General strategy:
+
+1. add annual age option to run_truth function call, default it to FALSE
+  + if true, looks for ANNAGEBIO and ANNAGECATCH in designated directory, error out if missing
+  + if true, read these files in addition to existing set rather than replace existing
+    + read them in with new load_nc_annage function
+    + add them to the output object in addition to existing set
+1. write load_nc_annage
+  + need to understand annage nc structures
+    + if same as others just with full age structure, replace previous nc
+    + only numbers output? 
+      + how to match with resn and structn for lengths/wts?
+      + reconcile with bio parameters such as maturity ogive?
+    + only for some groups?
 

### 2020-03-20 NOBAcod.Rmd
test loading full age structure

@@
 
-```{r full-test}
+```{r full-test, message=FALSE, warning=FALSE}
 
@@
     + only for some groups?
+
+What is in the ANNAGEBIO and ANNAGECATCH .nc files?
+
+```{r annagebio-contents}
+
+# taken from existing load_nc
+
+file.nc <- file.path(d.name, "outputnordic_runresults_01ANNAGEBIO.nc")
+
+  # Load ATLANTIS output!
+  at_out <- RNetCDF::open.nc(con = file.nc)
+
+  # Get info from netcdf file! (Filestructure and all variable names)
+  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
+    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
+  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
+  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
+  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
+
+RNetCDF::close.nc(at_out)
+
+```
+
+ANNAGEBIO.nc variable names (first hundred): `r var_names_ncdf[1:100]`
+
+ANNABEBIO.nc variable names (last hundred): `r var_names_ncdf[1590:1690]`
+
+ANNAGEBIO.nc timesteps: `r n_timesteps`
+
+ANNAGEBIO.nc boxes: `r n_boxes`
+
+ANNAGEBIO.nc layers: `r n_layers`
+
+Annual age output is in numbers and weight at each timestep, layer, and box, which is really good news.
+We should be able to get weight at true age more directly this way than the interpolation I was planning.
+Also, we may be able to do length estimation differently?
+
+To modify `load_nc_annage` we need to figure out how many true ages each species has from the biological parameter file using `load_biolprm`:
+
+```{r biolprm}
+
+biolprm <- load_biolprm(d.name, biol.prm.file)
+fgs <- load_fgs(d.name, functional.groups.file)
+
+#max age should be ages per cohort times the number of age classes (<10 for some NOBA spp)
+
+names(biolprm$agespercohort) <- c("code", "ageperagecl")
+
+maxage <- merge(biolprm$maturityogive, biolprm$agespercohort, all.x = T) %>%
+    select(code, nagecl, ageperagecl) %>%
+    mutate(maxage = nagecl * ageperagecl) %>%
+    mutate(name = fgs$Name[match(code, fgs$Code)])
+
+select_groups <- c('Mackerel', 'North_atl_cod', 'Capelin')
+
+for (i in seq_along(select_groups)) {
+  select_group_ages <- 1:maxage$maxage[match(select_groups[i], maxage$name)]
+}
+
+```
+
+Try this `load_nc_annage` function:
+
+```{r loadncannage, eval=FALSE}
+
+load_nc_annage <- function(dir = getwd(), file_nc, bps, fgs, biolprm, select_groups,
+  select_variable =
+  c("Nums", "Weight"),
+  check_acronyms = TRUE, bboxes = c(0), verbose = FALSE) {
+  # NOTE: The extraction procedure may look a bit complex...
+  # A different approach would be to
+  # create a dataframe for each variable (e.g. GroupAge_Nums)
+  # and combine all dataframes at the end.
+  # This alternative approach requires a lot more storage
+  # and the code wouldn't be vectorized.
+
+  # Check input of the nc file
+  if (tail(strsplit(file_nc, "\\.")[[1]], 1) != "nc") {
+    stop("The argument for file_nc,", file_nc, "does not end in nc")
+  }
+  if (is.null(dir)) {
+    file.nc <- file_nc
+  } else {
+    file.nc <- file.path(dir, file_nc)
+  }
+
+  # Check input of select_variable as only one value is allowed
+  select_variable <- match.arg(select_variable, several.ok = FALSE)
+
+  # Check input structure!
+  if (check_acronyms) {
+    active_groups <- as.vector(subset(fgs, IsTurnedOn == 1)$Name)
+    inactive_groups <- select_groups[which(
+      !is.element(select_groups, active_groups))]
+    if (length(inactive_groups) >= 1) {
+      select_groups <- select_groups[!is.element(select_groups, inactive_groups)]
+      warning(paste(paste("Some selected groups are not active in the model run.",
+        "Check 'IsTurnedOn' in fgs\n"),
+        paste(inactive_groups, collapse = "\n")))
+    }
+    if (all(!is.element(select_groups, active_groups))) {
+      stop(paste("None of the species selected are active in the model run.",
+        "Check spelling and Check 'IsTurnedOn' in fgs"))
+    }
+  }
+
+  # Deal with file structures
+
+  # Load ATLANTIS output!
+  at_out <- RNetCDF::open.nc(con = file.nc)
+  on.exit(RNetCDF::close.nc(at_out))
+
+  if (select_variable != "N" & all(is.element(select_groups, bps))) {
+    stop("The only output for Biomasspools is N.")
+  } else{
+    print(paste("Read", file.nc, "successfully"))
+  }
+
+  # Get info from netcdf file! (Filestructure and all variable names)
+  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
+    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
+  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
+  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
+  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
+
+  # Extract data from the ncdf file
+  # Create a vector of all potential variable names
+  # Only use names which are available in the ncdf-file as an
+  # extraction of missing variables is not possible
+  # Create vector of available species at the end using search_clean
+  # This is needed to create species-names later on
+
+  # CHANGED hardcoded cohort number to reflect true ages from
+
+  # To make the creation of variables as robust as possible
+  # we introduce different combinations of groups, variable, and cohort
+  # Only combinations present in the ncdf are used later on
+  # Loop over select_groups to use the same ordering
+
+  names(biolprm$agespercohort) <- c("code", "ageperagecl")
+
+  maxage <- merge(biolprm$maturityogive, biolprm$agespercohort, all.x = T) %>%
+    select(code, nagecl, ageperagecl) %>%
+    mutate(maxage = nagecl * ageperagecl) %>%
+    mutate(name = fgs$Name[match(code, fgs$Code)])
+
+  search <- list()
+  for (i in seq_along(select_groups)) {
+    select_group_ages <- 1:maxage$maxage[match(select_groups[i], maxage$name)]
+    search[[i]] <- c(
+      unlist(lapply(paste0(select_groups[i], select_group_ages),
+                    paste0, select_variable)),           # GroupCohortVariable
+      unlist(lapply(paste0(select_groups[i], select_variable),
+                    paste0, select_group_ages)),                   # GroupVariableCohort
+      unlist(lapply(paste0(select_groups[i], select_group_ages),
+                    paste, select_variable, sep = "_")), # GroupCohort_Variable
+      unlist(lapply(paste(select_groups[i], select_variable, sep = "_"),
+                    paste0, select_group_ages)),                   # Group_VariableCohort
+      unlist(lapply(paste(select_groups[i], select_group_ages, sep = "_"),
+                    paste, select_variable, sep = "_")), # Group_Cohort_Variable
+      unlist(lapply(paste(select_groups[i], select_variable, sep = "_"),
+                    paste, select_group_ages, sep = "_")),         # Group_Variable_Cohort
+      unlist(paste0(select_groups[i], select_variable)),                                                      # GroupVariable
+      unlist(paste(select_groups[i], select_variable,
+                   sep = "_"))                           # Group_Variable
+      )
+    search[[i]] <- search[[i]][is.element(search[[i]], var_names_ncdf)]
+    search[[i]] <- unique(search[[i]])
+  }
+  search_clean <- do.call(c, search)
+  # If the combination of select_groups and select_variable ends up not being found.
+  if (length(search_clean) == 0) return(0)
+
+  at_data <- lapply(search_clean, RNetCDF::var.get.nc, ncfile = at_out)
+
+  # Get final species and number of ageclasses per species
+  final_species <- select_groups[sapply(
+    lapply(select_groups, grepl, x = search_clean), any)]
+  final_agecl <- maxage$maxage[
+    sapply(final_species, function(x) which(x == maxage$name))]
+
+  num_layers <- RNetCDF::var.get.nc(ncfile = at_out, variable = "numlayers")[, 1]
+  # add sediment layer!
+  num_layers <- num_layers + ifelse(num_layers == 0, 0, 1)
+
+  # Create an array of layerids.
+  # Every entry in the array indicates if a layer is present (= 1) or not (= 0).
+  # Boxes without layers (= islands) have only 0s as id,
+  # used later on to remove data from non-existent layers!
+  # By default output should be 0 in those layers.
+  # Layers in boundary boxes are set to 0 if bboxes is anything other than NULL!
+  # Applying a boolean array to an array results in a vector!
+  for (i in seq_along(num_layers)) {
+    if (i == 1) layerid <- array(dim = c(n_layers, n_boxes))
+    if (num_layers[i] == 0) {
+      layerid[, i] <- 0
+    } else {
+      if (!is.null(bboxes) & is.element((i - 1), bboxes)) {
+        layerid[, i] <- 0
+      } else {
+        layerid[, i] <- c(rep(0, times = n_layers - num_layers[i]),
+                          rep(1, times = num_layers[i]))
+      }
+    }
+  }
+
+  # Create vectors for polygons and layers
+  # Each vector has the length equal to one time-step
+  # All data from islands and non-existent layers is removed
+  # Therefore the length of these
+  # vectors is equal for each extracted variable
+  boxes <- 0:(n_boxes - 1)
+  # Remove islands and boundary boxes
+  island_ids <- num_layers == 0
+  if (!is.null(bboxes)) {
+    boundary_ids <- is.element(boxes, bboxes)
+    island_ids <- island_ids | boundary_ids
+  }
+  boxes <- boxes[!island_ids]
+  num_layers <- num_layers[!island_ids]
+
+  polygons <- rep(boxes, times = num_layers)
+  layers <- sapply(num_layers[num_layers != 0] - 2,
+    function(x) c(seq(x, from = 0, by = 1), n_layers - 1))
+  if (any(sapply(layers, length) != num_layers[num_layers != 0])) {
+    stop("Number of layers incorrect.")
+  }
+  layers <- do.call(c, layers)
+  if (length(polygons) != length(layers)) {
+    stop("Number of polygons and layers do not match.")
+  }
+
+  # In the following section the data is transformed to a long dataframe
+  # I haven't found any solution to vectorize the creation of the dataframe
+  # columns (species, age, polygons,...)
+  # when data from 2d and 3d arrays
+  # (e.g. select_variable = "N" all biomasspools are only present in the
+  # sediment layer.) are read in simultaneously.
+  # Therefore the current "messy" solution splits the data
+  # in 2 subpopulations: 2d-data and 3d-data
+  at_data3d <- at_data[which(sapply(at_data, function(x) length(dim(x))) == 3)]
+  at_data2d <- at_data[which(sapply(at_data, function(x) length(dim(x))) == 2)]
+
+  int_fs <- final_species
+  int_fa <- final_agecl
+
+  if (length(at_data3d) >= 1) {
+    # Remove biomasspools if selected variable is "N"!
+    if (select_variable == "N") {
+      int_fs <- final_species[!is.element(final_species, bps)]
+      int_fa <- final_agecl[!is.element(final_species, bps)]
+      # Note this only works if age-structured vertebrates have 10 ageclasses
+      int_fa[int_fa == 10] <- 1
+    }
+    for (i in seq_along(at_data3d)) {# for loop over all variables
+      if (i == 1) result3d <- list()
+      for (j in 1:n_timesteps) {# loop over timesteps
+        if (j == 1) values <- array(dim = c(length(layers), n_timesteps))
+        values[, j] <- at_data3d[[i]][,, j][layerid == 1]
+      }
+      result3d[[i]] <- as.vector(values)
+    }
+    result3d <- data.frame(
+      species = unlist(sapply(
+        X = mapply(FUN = rep, x = int_fs, each = int_fa, SIMPLIFY = FALSE,
+        USE.NAMES = FALSE),
+        FUN = rep, each = length(layers) * n_timesteps, simplify = FALSE)),
+      agecl = unlist(sapply(
+        X = sapply(X = int_fa, FUN = seq, from = 1, by = 1, simplify = FALSE,
+        USE.NAMES = FALSE),
+        FUN = rep, each = length(layers) * n_timesteps, simplify = FALSE)),
+      polygon = unlist(sapply(
+        X = n_timesteps * int_fa, FUN = rep, x = polygons, simplify = F,
+        USE.NAMES = FALSE)),
+      layer = unlist(sapply(
+        X = n_timesteps * int_fa, FUN = rep, x = layers, simplify = FALSE,
+        USE.NAMES = FALSE)),
+      time = unlist(sapply(
+        X = int_fa, FUN = rep, x = rep(0:(n_timesteps - 1), each = length(layers)),
+        simplify = FALSE, USE.NAMES = FALSE)),
+      atoutput = do.call(c, result3d),
+      stringsAsFactors = FALSE)
+  }
+
+  if (length(at_data2d) >= 1) {
+    # Only select biomasspools if selected variable is "N"!
+    if (select_variable == "N") {
+      int_fs <- final_species[is.element(final_species, bps)]
+      int_fa <- final_agecl[is.element(final_species, bps)]
+    }
+    # age-structured invert groups are combined in ncdf file!
+    if (select_variable == "Grazing") int_fa <- 1
+    for (i in seq_along(at_data2d)) {# for loop over all variables!
+      if (i == 1) result2d <- list()
+      for (j in 1:n_timesteps) {# loop over timesteps
+        if (j == 1) values <- array(dim = c(length(boxes), n_timesteps))
+        values[, j] <- at_data2d[[i]][, j][boxes + 1]
+      }
+      result2d[[i]] <- as.vector(values)
+    }
+
+    # Order of the data in value column = "atoutput".
+    # 1. species  --> rep each with the number of
+    #                 ageclasses and n_timesteps * boxes
+    # 2. age      --> rep each (1:10 for each species) with n_timesteps * boxes
+    # 3. timestep --> rep each timestep (1:n_timesteps)
+    #                 with the number of boxes and final_agecl
+    #                 (num cohorts per species)
+    # 4. polygon  --> rep boxes times n_timesteps * final_agecl
+    #                 (num cohorts per species)
+    result2d <- data.frame(species = unlist(sapply(
+      X = mapply(FUN = rep, x = int_fs, each = int_fa, SIMPLIFY = FALSE,
+        USE.NAMES = FALSE),
+        FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
+      agecl = unlist(sapply(X = sapply(X = int_fa, FUN = seq, from = 1,
+        by = 1, simplify = FALSE, USE.NAMES = FALSE),
+        FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
+      polygon = unlist(sapply(X = n_timesteps * int_fa,
+        FUN = rep, x = boxes, simplify = FALSE, USE.NAMES = FALSE)),
+      time = unlist(sapply(X = int_fa, FUN = rep, x = rep(0:(n_timesteps - 1),
+        each = length(boxes)), simplify = FALSE, USE.NAMES = FALSE)),
+      atoutput = do.call(c, result2d),
+      stringsAsFactors = F)
+    if (select_variable == "N") result2d$layer <- n_layers - 1
+  }
+
+  # Combine dataframes if necessary!
+  if (all(sapply(lapply(at_data, dim), length) == 3) & select_variable != "N") {
+    result <- result3d
+  }
+  if (all(sapply(lapply(at_data, dim), length) == 2) & select_variable != "N") {
+    result <- result2d
+  }
+  if (select_variable == "N") {
+    if (length(at_data2d) >= 1 & length(at_data3d) == 0) result <- result2d
+    if (length(at_data2d) == 0 & length(at_data3d) >= 1) result <- result3d
+    if (length(at_data2d) >= 1 & length(at_data3d) >= 1) {
+      result <- rbind(result2d, result3d)
+    }
+  }
+
+  # Remove min_pools if existent (well, there always are min pools... ;)).
+  min_pools <- is.element(result$atoutput, c(0, 1e-08, 1e-16))
+  if (length(min_pools) > 0) {
+    # exclude 1st timestep and sediment layer from calculation
+    print_min_pools <- sum(min_pools) -
+      length(result[min_pools & result$time == 1, 1]) -
+      length(result[min_pools & result$time > 1 & result$layer == 7, 1])
+    if (print_min_pools > 0 & verbose) {
+      warning(paste0(round(print_min_pools/dim(result)[1] * 100),
+        "% of ", select_variable, " are true min-pools (0, 1e-08, 1e-16)"))
+    }
+    result <- result[!min_pools, ]
+  }
+
+  # Remove non-existent layers.
+  # WARNING: Biomass is build up (very few) in sediment layer for
+  # NON sediment groups (e.g. baleen whales)
+  # Therefore, I subset all data from that layer for non biomass groups and
+  # groups which cannot penetrate into the sediment!
+  # UPDATE: Doesn't work with layers as species are not distributed through
+  # the whole water column and do not appear in
+  # every polygon.
+
+  # Sum up N for invert cohorts if invert cohorts are present!
+  # NOTE: invert cohorts of size 10 are not considered!
+  if (select_variable == "N" & any(final_agecl != 10 & final_agecl > 1)) {
+    result <- result %>%
+      dplyr::group_by(polygon, layer, species, time) %>%
+      dplyr::summarise(atoutput = sum(atoutput))
+  }
+
+  return(result)
+}
+
+```
+
+Does it work?
+
+```{r testloadnc}
+
+file_nc <- ("outputnordic_runresults_01ANNAGEBIO.nc")
+bps <- load_bps(dir = d.name, fgs = functional.groups.file, file_init = biomass.pools.file)
+allboxes <- load_box(dir = d.name, file_bgm = box.file)
+boxes <- get_boundary(allboxes)
+
+numsage <- load_nc_annage(dir = d.name,
+                  file_nc = file_nc,
+                  bps = bps,
+                  fgs = fgs,
+                  biolprm = biolprm,
+                  select_groups = select_groups,
+                  select_variable = "Nums",
+                  check_acronyms = TRUE,
+                  bboxes = boxes)
+  if(verbose) message("Numbers read in from ANNAGEBIO.")
+
+weightage <- load_nc_annage(dir = d.name,
+                  file_nc = file_nc,
+                  bps = bps,
+                  fgs = fgs,
+                  biolprm = biolprm,
+                  select_groups = select_groups,
+                  select_variable = "Weight",
+                  check_acronyms = TRUE,
+                  bboxes = boxes)
+  if(verbose) message("Weight read in from ANNAGEBIO.")
+
+```
+
+Need to aggreagate these to see if they look reasonable
+
+```{r fullagecomps-test}
+
+omlist_ss<-NOBAom_cod
+source(here("/config/NOBA2Config.R"))
+source(here("/config/omdimensions.R"))
+source(here("/config/codsurvey.R"))
+source(here("/config/codfishery.R"))
+
+n_reps <- 1
+
+  #numbers based fishery independent survey for age and length comps
+  # same user specifications as indices
+  survey_N <- atlantisom::create_survey(dat = numsage,
+                                        time = survtime,
+                                        species = select_groups,
+                                        boxes = survboxes,
+                                        effic = surveffic,
+                                        selex = survselex)
+  #Sample fish for age composition
+  # if we want replicates for obs error this sample function will generate them
+  age_comp_data <- list()
+  for(i in 1:n_reps){
+    age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
+  }
+
+    # save age comps
+  if(save){
+    saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
+  }
+
+```
 

### 2020-03-23 NOBAcod.Rmd
full age test for cod with code folding

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output:
+  html_document:
+    code_fold: hide
 ---
@@
                      selex=rep(1.0,length(survspp)*n_age_classes))
+
+# WARNING!! comment above and use this for annage output
+survselex <- data.frame(species=rep(survspp, each=n_annages),
+                        agecl=rep(c(1:n_annages),length(survspp)),
+                        selex=rep(1.0,length(survspp)*n_annages))
 
@@
 
-Does it work?
+Does it work? This snippet will go into `run_truth` if it does:
 
@@
                   check_acronyms = TRUE,
-                  bboxes = boxes)
-  if(verbose) message("Numbers read in from ANNAGEBIO.")
+                  bboxes = boxes, 
+                  verbose = TRUE)
+  #if(verbose) message("Numbers read in from ANNAGEBIO.")
 
@@
                   check_acronyms = TRUE,
-                  bboxes = boxes)
-  if(verbose) message("Weight read in from ANNAGEBIO.")
-
-```
-
-Need to aggreagate these to see if they look reasonable
+                  bboxes = boxes, 
+                  verbose = TRUE)
+  #if(verbose) message("Weight read in from ANNAGEBIO.")
+
+```
+
+Need to aggreagate these to see if they look reasonable; this code snippet from `om_comps` with `numsage` output subsituted for `nums`.
 
@@
                                         time = survtime,
-                                        species = select_groups,
+                                        species = survspp,
                                         boxes = survboxes,
@@
     # save age comps
-  if(save){
+  #if(save){
     saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
-  }
+  #}
+
+```
+
+Plot full age comps, have 20 years for cod now, no longer age class but annual ages.
+
+```{r plotfullagecomp}
+
+#read in comp data
+age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
+age_comp_data <- age_comp_data[[1]]
+
+#fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
+#fish_age_comp <- fish_age_comp[[1]]
+
+#add this to om_indices function so that this has years when read in
+#fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
+
+Natage <- filter(age_comp_data, time %in% c(55:175))
+
+Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
+    geom_point() +
+    theme_tufte() +
+    labs(subtitle = paste(scenario.name,
+                          Natage$species))
+
+  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+
+```
+
+What is the weight output?
+Documentation only mentions numbers output.
+Looks like average weight at age for each time and polygon?
+If so, need to do weighted average weight using the numbers at age in each polygon?
+
+```{r}
 

### 2020-03-23 config/codsurvey.R

@@ -30,9 +30,14 @@ surveffic <- data.frame(species=survspp,
 
 # survey selectivity (agecl based)
 # this is by age class, need to change to use with ANNAGEBIO output
-survselex <- data.frame(species=rep(survspp, each=n_age_classes),
-                     agecl=rep(c(1:n_age_classes),length(survspp)),
-                     selex=rep(1.0,length(survspp)*n_age_classes))
+#survselex <- data.frame(species=rep(survspp, each=n_age_classes),
+#                     agecl=rep(c(1:n_age_classes),length(survspp)),
+#                     selex=rep(1.0,length(survspp)*n_age_classes))
+
+# for annage output
+survselex <- data.frame(species=rep(survspp, each=n_annages),
+                        agecl=rep(c(1:n_annages),length(survspp)),
+                        selex=rep(1.0,length(survspp)*n_annages))
 
 # effective sample size needed for sample_fish
 # this effective N is high but not equal to total for numerous groups

### 2020-03-25 NOBAcod.Rmd
atlantis code to get weight units

@@
 
+This output is described in Atlantis code, file atlantis/atphysics/atagetracerIO.c, function `writeBMAnnAgeBioData`:
+
+```
+if (propit == out_nums_id) {  // Numbers case
+                                for (b = 0; b < bm->nbox; b++) {
+                                    val[b] = 0.0;
+                                    for (k = 0; k < bm->wcnz; k++) {  // Sum over all individuals in the box
+                                        val[b] += bm->wctr[b][k][id] * FunctGroupArray[sp].boxPopRatio[b][k][cohort][ai];
+                                    }
+                                }
+                            } else  { // Size data - get weighted average of values per layer if numbers > min_dens
+                                for (b = 0; b < bm->nbox; b++) {
+                                    val[b] = 0.0;
+                                    tot = 0.0;
+                                    for (k = 0; k < bm->wcnz; k++) {  // Sum over all individuals in the box
+                                        if (bm->wctr[b][k][id] > bm->min_dens) {
+                                            val[b] += bm->wctr[b][k][id] * (bm->wctr[b][k][sn] + bm->wctr[b][k][rn]) * 5.7 * mg_2_g * FunctGroupArray[sp].boxPopRatio[b][k][cohort][ai];
+                                            tot += bm->wctr[b][k][id];
+                                        }
+                                    }
+                                    if (tot > 0.0)
+                                        val[b] /= tot;
+                                }
+                            }
+```
+
+So weight output appears to be average weight at age in grams?
+
 ```{r}

### 2020-03-25 NOBAcod.Rmd
add wt age plot

@@
 
-```{r}
+```{r wtage}
+
+annwtage <- weightage %>%
+  group_by(species, agecl, time) %>%
+  summarise(wtage = mean(atoutput)) %>%
+  filter(time %in% survtime)
+
+wageplot <- ggplot(annwtage, aes(time, wtage)) +
+  geom_line(aes(colour = factor(agecl))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  xlab("model timestep (5 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
 

### 2020-03-27 NOBAcod.Rmd
cant push after 2FA update
stashing this here

@@
 
-```
-d.name <- here("atlantisoutput","NOBA_march_2020")
-functional.groups.file <- "nordic_groups_v04.csv"
-biomass.pools.file <- "nordic_biol_v23.nc"
-biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
-box.file <- "Nordic02.bgm"
-initial.conditions.file <- "nordic_biol_v23.nc"
-run.prm.file <- "nordic_run_v01.xml"
-scenario.name <- "nordic_runresults_01"
-bioind.file <- "outputnordic_runresults_01BiomIndx.txt"
-catch.file <- "outputnordic_runresults_01Catch.txt"
+```{r, code = readLines("NOBA2Config.R")}
 ```
@@
 
-```
-#survey species inherited from omlist_ss
-survspp <- omlist_ss$species_ss
-# survey season and other time dimensioning parameters
-# generalized timesteps all models
-noutsteps <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
-timeall <- c(0:noutsteps)
-stepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutinc
-midptyr <- round(median(seq(0,stepperyr)))
-
-# model areas, subset in surveyconfig
-allboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
-
-# fishery output: learned the hard way this can be different from ecosystem outputs
-fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
-
-# survey selectivity (agecl based)
-# should return all age classes fully sampled (Atlantis output is 10 age groups per spp)
-n_age_classes <- omlist_ss$funct.group_ss$NumCohorts
-age_classes <- 1:n_age_classes
-
-n_annages <- n_age_classes * omlist_ss$funct.group_ss$NumAgeClassSize
-annages <- 1:n_annages
-
+```{r, code = readLines("omdimensions.R")}
 ```
@@
 
-```
-# Default survey configuration here is a census
-# Need to define survey season, area, efficiency, selectivity
-
-#Atlantis model timestep corresponding to the true output--now from census_spec.R
-timestep <- stepperyr #5
-
-#Which atlantis timestep does the survey run in?--now from census_spec.R
-# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
-# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
-
-survey_sample_time <- midptyr #2; defined in omdimensions.R
-
-#The last timestep to sample
-total_sample <- noutsteps-1 #495
-
-#Vector of indices of survey times to pull
-survey_sample_full <- seq(survey_sample_time,
-                          total_sample, by=timestep)
-
-survtime <- survey_sample_full
-
-# survey area
-# should return all model areas
-survboxes <- allboxes
-
-# survey efficiency (q)
-# should return a perfectly efficient survey 
-surveffic <- data.frame(species=survspp,
-                     efficiency=rep(0.5,length(survspp)))
-
-# survey selectivity (agecl based)
-# this is by age class, need to change to use with ANNAGEBIO output
-survselex <- data.frame(species=rep(survspp, each=n_age_classes),
-                     agecl=rep(c(1:n_age_classes),length(survspp)),
-                     selex=rep(1.0,length(survspp)*n_age_classes))
-
-# WARNING!! comment above and use this for annage output
-survselex <- data.frame(species=rep(survspp, each=n_annages),
-                        agecl=rep(c(1:n_annages),length(survspp)),
-                        selex=rep(1.0,length(survspp)*n_annages))
-
-# effective sample size needed for sample_fish
-# this effective N is high but not equal to total for numerous groups
-surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
-
-# survey index cv needed for sample_survey_xxx
-# cv = 0.1
-surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp)))
-
-# length at age cv for input into calc_age2length function
-# function designed to take one cv for all species, need to change to pass it a vector
-lenage_cv <- 0.1
-
-# max size bin for length estimation, function defaults to 150 cm if not supplied
-maxbin <- 150
-
+```{r, code = readLines("codsurvey.R")}
 ```
@@
 
-```
-# Default fishery configuration here is a census
-# Inherits species from input omlist_ss
-fishspp <- omlist_ss$species_ss
-
-#Number of years of data to pull
-nyears <- 50
-
-#Atlantis initialization period in years
-burnin <- 30
-
-# same time dimensioning parameters as in surveycensus.R
-#Vector of indices of catch in numbers to pull (by timestep to sum)
-fish_sample_full <- c(0:total_sample)  #total_sample defined in sardinesurvey.R
-fish_burnin <- burnin*fstepperyr+1
-fish_nyears <- nyears*fstepperyr
-fish_times <- fish_sample_full[fish_burnin:(fish_burnin+fish_nyears-1)]
-fish_timesteps <- seq(fish_times[5], max(fish_times), by=fstepperyr) #last timestep
-#fish_years <- unique(floor(fish_times/fstepperyr)+1) # my original
-fish_years <- unique(floor(fish_times/fstepperyr)) #from Christine's new sardine_config.R
-
-fishtime <- fish_times
-
-# fishery output: learned the hard way this can be different from ecosystem outputs
-fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
-
-# fishery sampling area
-# should return all model areas, this assumes you see everything that it caught
-fishboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
-
-# effective sample size needed for sample_fish
-# this effective N is high but not equal to total for numerous groups
-fisheffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
-
-#this adjusts for subannual fishery output so original effN is for whole year
-fisheffN$effN <- fisheffN$effN/fstepperyr
-
-# fishery catch cv can be used in sample_survey_biomass
-# perfect observation
-fish_cv <- data.frame(species=survspp, cv=rep(0.01,length(survspp)))
+```{r, code = readLines("codfishery.R")}
 ```
@@
   Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+```
+
+#### Survey weight at age (age classes)
+
+```{r wtageclass-1}
+
+wtage <- readRDS(file.path(d.name, paste0(scenario.name, "survObsWtAtAge.rds")))
+wtage <- wtage[[1]]
+
+wageplot <- ggplot(wtage, aes(time, wtage)) +
+  geom_line(aes(colour = factor(agecl))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  xlab("model timestep (5 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
 ```
@@
 
+So, to modify run_truth we need to add the annage output but not replace the regular output in case users want output from non-age structured groups (inverts, etc.).
+Also need to see if lengths can be generaged from the annage outputs, if not need to keep the full set of outputs from the .nc file for all species to generate them.
+
+Looks like the annagecatch output is only numbers and not weights.
+
 ## Write input data for specific assessment models

### 2020-03-27 NOBAcod.Rmd
load annagecatch
test push
census configs for cod, start anncatch

@@
 
-NOBAom_cod_comp <- om_comps(usersurvey = here("config/sardinesurvey.R"), 
-                           userfishery = here("config/sardinefishery.R"),
+NOBAom_cod_comp <- om_comps(usersurvey = here("config/codsurvey.R"), 
+                           userfishery = here("config/codfishery.R"),
                            omlist_ss = NOBAom_cod, 
@@
 
-wageplot <- ggplot(wtage, aes(time, wtage)) +
+wageplot <- ggplot(wtage, aes(time, atoutput)) +
   geom_line(aes(colour = factor(agecl))) +
@@
 
+```{r loadannagecatch}
+# taken from existing load_nc
+
+file.nc <- file.path(d.name, "outputnordic_runresults_01ANNAGECATCH.nc")
+
+  # Load ATLANTIS output!
+  at_out <- RNetCDF::open.nc(con = file.nc)
+
+  # Get info from netcdf file! (Filestructure and all variable names)
+  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
+    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
+  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
+  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
+  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
+
+RNetCDF::close.nc(at_out)
+```
+
+ANNAGECATCH.nc variable names (first hundred): `r var_names_ncdf[1:100]`
+
+ANNAGECATCH.nc variable names (last hundred): `r tail(var_names_ncdf, 100)`
+
+ANNAGECATCH.nc timesteps: `r n_timesteps`
+
+ANNAGECATCH.nc boxes: `r n_boxes`
+
+ANNAGECATCH.nc layers: `r n_layers`
+
+Catch file contains "IsFished" species discard at age and catch at age by fishery.
+
 ## Write input data for specific assessment models

### 2020-03-27 config/codfishery.R

@@ -8,6 +8,9 @@ nyears <- 50
 #Atlantis initialization period in years
 burnin <- 30
 
+# fishery output: learned the hard way this can be different from ecosystem outputs
+fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
+
 
 # same time dimensioning parameters as in surveycensus.R
 #Vector of indices of catch in numbers to pull (by timestep to sum)
@@ -15,14 +18,12 @@ fish_sample_full <- c(0:total_sample)  #total_sample defined in sardinesurvey.R
 fish_burnin <- burnin*fstepperyr+1
 fish_nyears <- nyears*fstepperyr
 fish_times <- fish_sample_full[fish_burnin:(fish_burnin+fish_nyears-1)]
-fish_timesteps <- seq(fish_times[5], max(fish_times), by=fstepperyr) #last timestep
+fish_timesteps <- seq(fish_times[fstepperyr], max(fish_times), by=fstepperyr) #last timestep
 #fish_years <- unique(floor(fish_times/fstepperyr)+1) # my original
 fish_years <- unique(floor(fish_times/fstepperyr)) #from Christine's new sardine_config.R
 
 fishtime <- fish_times
 
-# fishery output: learned the hard way this can be different from ecosystem outputs
-fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
 
 # fishery sampling area
 # should return all model areas, this assumes you see everything that it caught

### 2020-03-27 config/codsurvey.R

@@ -26,7 +26,7 @@ survboxes <- allboxes
 # survey efficiency (q)
 # should return a perfectly efficient survey 
 surveffic <- data.frame(species=survspp,
-                     efficiency=rep(0.5,length(survspp)))
+                     efficiency=rep(1.0,length(survspp)))
 
 # survey selectivity (agecl based)
 # this is by age class, need to change to use with ANNAGEBIO output
@@ -52,5 +52,5 @@ surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp)))
 lenage_cv <- 0.1
 
 # max size bin for length estimation, function defaults to 150 cm if not supplied
-maxbin <- 150
+maxbin <- 200
 

### 2020-03-27 NOBAcod.Rmd
now need fishery setup files

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(here)
 ```
@@
 
-```{r, code = readLines("NOBA2Config.R")}
+```{r, code = readLines("./config/NOBA2Config.R"), eval=F}
 ```
@@
 
-```{r, code = readLines("omdimensions.R")}
+```{r, code = readLines("./config/omdimensions.R"), eval=F}
 ```
@@
 
-```{r, code = readLines("codsurvey.R")}
+```{r, code = readLines("./config/codsurvey.R"), eval=F}
 ```
@@
 
-```{r, code = readLines("codfishery.R")}
+```{r, code = readLines("./config/codfishery.R"), eval=F}
 ```
@@
 
-library(here)
+#library(here) now above
 library(tidyverse)
@@
 
-Catch file contains "IsFished" species discard at age and catch at age by fishery.
+Catch file contains "IsFished" species discard at age and catch at age by fishery in numbers.
+So now modifying load_nc_annage to get the fishery specific outputs, which requires additional atlantis fishery setup files and additions to the config files above.
 

### 2020-05-07 Visualize_atlantisom.Rmd
function visualization experiments

---
title: "Visualize atlantisom"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relationships between `atlantsom` functions

First attempt is using the package [`DependenciesGraphs`](https://github.com/datastorm-open/DependenciesGraphs).

Other possibilities:

[`visNetwork`](https://github.com/datastorm-open/visNetwork)

```{r installation, eval=FALSE}

devtools::install_github("datastorm-open/DependenciesGraphs")
devtools::install_github("datastorm-open/visNetwork") # on CRAN, this is dev version

```

## `DependenciesGraphs`

```{r depgraph}

library(DependenciesGraphs)
library(atlantisom)

dep <- envirDependencies("package:atlantisom")
plot(dep)

```

### 2020-05-11 Visualize_atlantisom.Rmd
add single functions
to do next text

@@
 
+[`drake`](https://docs.ropensci.org/drake/)
+
 ```{r installation, eval=FALSE}
@@
 
-```{r depgraph}
+This is all the functions in `atlantisom`:
+
+```{r depgraph, message=FALSE}
 
@@
 ```
+
+This is just the `om_init` wrapper function:
+
+```{r depgraphominit}
+
+depominit <- funDependencies("package:atlantisom", "om_init")
+plot(depominit)
+
+```
+
+While these map dependencies between functions, it would be nice to see what output of atlantis goes into atlantisom and what atlantisom outputs into SS3.

### 2020-05-21 Visualize_atlantisom.Rmd
add Isaac's new SS3 run and visualize

@@
 title: "Visualize atlantisom"
-author: "Sarah Gaichas"
+author: "Sarah Gaichas and Christine Stawitz"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output: 
+  html_document:
+    code_fold: hide
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE,
+                      message = FALSE,
+                      warning = FALSE)
 ```
@@
 While these map dependencies between functions, it would be nice to see what output of atlantis goes into atlantisom and what atlantisom outputs into SS3.
+
+Try something different. Drake?
+
+```{r atlantisomdrake}
+
+#atlantis output files needed
+
+#atlantisom functions (wrappers) with user-defined inputs specifying sampling frames, observation errors
+
+#om_init oututs run_truth results and model definitions for all species
+
+#om_species cuts full output to user selected focal species--SAVE THIS OUTPUT
+
+#om_index takes user defined survey and fishery sampling and observation error specifications to make replicate indices, outputs both to session and saves
+
+#atlantisom core functions called from wrappers, can be used alone
+
+#stock assessment data rds file output
+
+#atlantisom functions to write assessment model-specific input files
+
+#future: wrapper functions to run and compare assessments with each other and truth
+
+```
+
+## Visualize results using wrappers
+
+Meanwhile lets do final plots for the ecosystem MSE paper.
+Want a 3 panel plot showing atlantisom sample output (our really good survey), SS3 sample output (fit to survey), and the key feature, SS3 fit to atlantis truth:
+
+From truth and saved wrapper output generated [here](https://sgaichas.github.io/poseidon-dev/WrapperDevelopment.html):
+
+```{r truth-surveyind}
+
+library(tidyverse) #fix this; build in this dependency for wrappers
+library(here)
+library(ggthemes)
+
+# skip this, here for completeness, just need directory
+#CC3om <- om_init(here("config/CC3config_sardLWcorr.R")) #initialize/read in truth
+
+source(here("config/CC3config_sardLWcorr.R"))
+
+#skip this too and now just read in the saved results
+#CC3om_sardine <- om_species(c("Pacific_sardine"), CC3om) #select just sardine
+
+# TODO make option to save this output so we dont have to run om_init to get truth 
+# every time and it doesnt sit around in memory. For now:
+
+#saveRDS(CC3om_sardine, file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+
+CC3om_sardine <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+
+# we have already run this so just read in the saved results, here for reference
+#CC3om_sard_ind <- om_index(usersurvey = here("config/sardinesurvey.R"), 
+#                           userfishery = here("config/sardinefishery.R"),
+#                           omlist_ss = CC3om_sardine, 
+#                           n_reps = 1, 
+#                           save = TRUE)
+
+CC3om_sard_ind <- readRDS(file.path(d.name, paste0(scenario.name, "surveyB.rds")))
+
+# TODO make plotting functions for atlantisom that do these comparisons
+
+# needed for plotting time dimensions 
+# TODO standardize time across outputs in these wrappers
+omlist_ss <- CC3om_sardine
+source(here("config/omdimensions.R"))
+
+txtTrueB <- CC3om_sardine$truetotbio_ss
+
+#WARNING! this is true bio at age so needs to be summed over ages--rename truebioage_ss
+omTrueB <- CC3om_sardine$truebio_ss %>%
+  group_by(species, time) %>%
+  summarise(totB = sum(atoutput))
+
+survObsBiomB <- CC3om_sard_ind[[1]]
+
+plotB <-ggplot() +
+  geom_point(data=txtTrueB, aes(x=time/365,y=atoutput, color="txt output true B"),
+             alpha = 10/10) + 
+  geom_line(data=omTrueB, aes(x=time/stepperyr,y=totB, color="survey census B"), 
+            alpha = 10/10) +
+  geom_point(data=survObsBiomB, aes(x=time/stepperyr,y=atoutput, color="survey sample B from biomass_ages"), 
+            alpha = 10/10) +
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotB +
+  facet_wrap(~species, scales="free")
+
+```
+
+From Christine's ggplot_SS.R (change to Isaac's output directory name for current best fit) we get SS3 sample output using a prettier version of r4ss plot:
+
+```{r ggplotSS}
+
+#dir <- "./inst/extdata/Sardine_SS_Files/"
+
+dir <- "./inst/extdata/AB_Lmax25_v2_ABinControl_Mlower/"
+
+#Get SS output
+output_from_SS <- r4ss::SS_output(dir, covar=F, forecast=F)
+
+#Extract index info
+index_data_fits <- output_from_SS$cpue
+#Commented out - only for comparing to SS output to check
+#r4ss::SS_plots(output_from_SS)
+
+#Plot
+require(ggthemes)
+require(ggplot2)
+plotB <-ggplot(index_data_fits) +
+  geom_line(data=index_data_fits, aes(x=Yr,y=Exp,color="Estimated CPUE"), 
+            alpha = 10/10) +
+  geom_pointrange(data=index_data_fits, aes(x=Yr,y=Obs, ymin=exp(log(Obs)+qt(0.025,df=output_from_SS$survey_error[2])*SE), 
+                                                                 ymax=exp(log(Obs)+qt(0.975,df=output_from_SS$survey_error[2])*SE),
+                                                                          color="Input CPUE with error"),
+             alpha = 10/10) + 
+  theme_tufte() +
+  theme(legend.position = "top") 
+plotB
+
+```
+
+Add the comparison of SS3 estimated total biomass with true atlantisom tot B as started [here](https://sgaichas.github.io/poseidon-dev/SkillAssessInit.html):
+
+```{r compareB}
+
+#point this to Isaac's output dir defined above
+#replist <- r4ss::SS_output(dir = here(dir), verbose=TRUE, printstats=TRUE,covar=FALSE)
+
+#SS_plots(replist)
+
+#replist$timeseries
+names(output_from_SS$timeseries)[19]<-"Fmort"
+
+plotB <-ggplot() +
+  geom_line(data=omTrueB, aes(x=time/stepperyr,y=totB, color="True B"), 
+            alpha = 10/10) +
+  geom_point(data=output_from_SS$timeseries, aes(x=Yr,y=Bio_all, color="SS3 Est B"),
+             alpha = 10/10) + 
+  theme_tufte() +
+  theme(legend.position = "top") +
+  labs(colour=scenario.name)
+
+plotB +
+  facet_wrap(~species, scales="free")
+
+```

### 2020-05-21 Visualize_atlantisom.Rmd
add 3 panel plot

@@
 #Plot
-require(ggthemes)
-require(ggplot2)
+#require(ggthemes)
+#require(ggplot2)
 plotB <-ggplot(index_data_fits) +
@@
 ```
+
+Putting it together into a 3 panel figure:
+
+```{r triplot, fig.asp=1}
+
+require(cowplot)
+require(viridis)
+
+mycol <- viridis(4)
+
+plotsurv <- ggplot() +
+  geom_line(data=omTrueB, aes(x=time/stepperyr,y=totB, color="True total B"), 
+            alpha = 10/10) +
+  geom_point(data=survObsBiomB, aes(x=time/stepperyr,y=atoutput, color="Survey sample B"), 
+            alpha = 10/10) +
+  xlab("") + 
+  ylab("") +
+  scale_color_manual(values=c(mycol[2], mycol[1]),
+                     guide = guide_legend(override.aes = list(
+                         linetype = c("blank", "solid"),
+                         shape = c(16, NA)))) + 
+  theme_tufte() +
+  theme(legend.position = "right") +
+  labs(colour="Atlantis-sardine")
+
+plotSSfit <- ggplot() +
+  geom_line(data=index_data_fits, aes(x=Yr,y=Exp,color="SS3 est. survey B"), 
+            alpha = 10/10, size=2) +
+  geom_pointrange(data=index_data_fits, 
+                  aes(x=Yr,y=Obs,
+                      ymin=exp(log(Obs)+qt(0.025,df=output_from_SS$survey_error[2])*SE),
+                      ymax=exp(log(Obs)+qt(0.975,df=output_from_SS$survey_error[2])*SE),
+                      fatten = .5,
+                      color="Survey B + error"),
+                  alpha = 10/10) + 
+  scale_x_continuous(limits = c(0,120), breaks = seq(0,125,25)) +
+  xlab("") + 
+  ylab("Biomass, t") +
+  scale_color_manual(values=c(mycol[3], mycol[2]),
+                     guide = guide_legend(override.aes = list(
+                         linetype = c("solid", "blank"),
+                         shape = c(NA, 16)))) + 
+  theme_tufte() +
+  theme(legend.position = "right") +
+  labs(colour="SS3-survey fit")
+
+plotSSskill <- ggplot() +
+  geom_line(data=omTrueB, aes(x=time/stepperyr,y=totB, color="True total B"), 
+            alpha = 10/10) +
+  geom_point(data=output_from_SS$timeseries, aes(x=Yr,y=Bio_all, color="SS3 est. total B"),
+             alpha = 10/10) + 
+  xlab("Year") + 
+  ylab("") +
+  scale_color_manual(values=c(mycol[3], mycol[1]),
+                     guide = guide_legend(override.aes = list(
+                         linetype = c("blank", "solid"),
+                         shape = c(16, NA)))) + 
+  theme_tufte() +
+  theme(legend.position = "right") +
+  labs(colour="SS3-biomass skill")
+
+plot_grid(plotsurv + theme(legend.justification = c(0,1)), 
+          plotSSfit + theme(legend.justification = c(0,1)), 
+          plotSSskill + theme(legend.justification = c(0,1)), 
+          ncol=1, align="v")
+
+```

### 2020-05-22 NOBAcod.Rmd
partial functions for ANNAGECATCH reading

@@
   summarise(wtage = mean(atoutput)) %>%
-  filter(time %in% survtime)
-
-wageplot <- ggplot(annwtage, aes(time, wtage)) +
+  filter(time %in% (survtime))
+
+boxwtage <- weightage %>%
+  mutate(wtage = atoutput) %>%
+  filter(species %in% c("North_atl_cod"),
+         polygon %in% c(1,7,24,30,38,45),
+         agecl %in% c(16:20))
+
+wtagedat <- boxwtage
+
+wageplot <- ggplot(wtagedat, aes(time, wtage)) +
   geom_line(aes(colour = factor(agecl))) +
@@
   ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
+  labs(subtitle = paste0(scenario.name, " "))
+
+wageplot + facet_wrap(c("polygon"), scales="free_y")
 
@@
 
-Looks like the annagecatch output is only numbers and not weights.
+UPDATE: there is a difference between the annage weight at age and that calculated using 'calc_age2length'.
+While the results from the standard nc output look reasonable for NOBA cod, the annage weight at age is half that for the standard (and half what real cod would be).
+More worrisome, the oldest age classes have smaller average weight at age than some younger ones.
+I cannot think of why this would be.
+Until we are sure the annage weight output is correct, we will instead interpolate the weight at age output of 'calc_age2length" for use as assessment input.
+
+Looks like the annagecatch output is only numbers and not weights. We don't get weight at age from fisheries usually so this is fine.
 
@@
 So now modifying load_nc_annage to get the fishery specific outputs, which requires additional atlantis fishery setup files and additions to the config files above.
+
+New function `load_fisheries` needed to load the fishery specification file (a .csv), similar to `load_fgs`:
+
+```{r load_fisheries}
+
+load_fisheries <- function(dir = getwd(), file_fish){
+  if (is.null(dir)) {
+    file.fish <- file_fish
+  } else {
+    file.fish <- file.path(dir, file_fish)
+  }
+  result <- read.table(file = file.fish, sep = ",",
+    header = TRUE, stringsAsFactors = FALSE)
+  return(result)
+}
+```
+
+Call `load_fisheries` within `load_nc_annage` to rename select_variable as used in ANNAGECATCH.nc:
+
+```{r addsnippet}
+
+select_variable <- c("Catch", "Discard")
+dir <- here("atlantisoutput","NOBA_march_2020")
+file_fish <- "NoBAFisheries.csv"
+
+#starting at line 115 in current load_nc_annage
+
+if(select_variable %in% c("Catch", "Discard")){
+    #read in fleet names
+    fleetnames <- load_fisheries(dir = dir, file_fish = file_fish)
+    for(i in select_variable){
+      sv <-paste0(i, "_", fleetnames$Code)
+      sv1 <- c()
+    }
+  }
+
+```
 

### 2020-05-22 NOBAcod.Rmd
testing fishery catch at age (annual)

@@
     fleetnames <- load_fisheries(dir = dir, file_fish = file_fish)
+    svfish <- c()
     for(i in select_variable){
       sv <-paste0(i, "_", fleetnames$Code)
-      sv1 <- c()
-    }
-  }
-
-```
+      svfish <- c(svfish, sv)
+    }
+    select_variable <- svfish
+  }
+
+```
+
+Test new `load_nc_annage`:
+
+```{r testfishery}
+
+source(here("R/load_nc_annage.R"))
+source(here("R/load_fisheries.R"))
+
+fgs <- NOBAom$funct.groups
+biolprm <- NOBAom$biol
+select_groups <- c('Mackerel', 'North_atl_cod', 'Capelin')
+
+file_nc <- "outputnordic_runresults_01ANNAGECATCH.nc"
+file_fish <- "NoBAFisheries.csv"
+
+fishcatchage <- load_nc_annage(dir = d.name,
+                  file_nc = file_nc,
+                  file_fish = file_fish,
+                  bps = bps,
+                  fgs = fgs,
+                  biolprm = biolprm,
+                  select_groups = select_groups,
+                  select_variable = "Catch",
+                  check_acronyms = TRUE,
+                  bboxes = boxes, 
+                  verbose = TRUE)
+
+fishcatchage
+
+fishdiscage <- load_nc_annage(dir = d.name,
+                  file_nc = file_nc,
+                  file_fish = file_fish,
+                  bps = bps,
+                  fgs = fgs,
+                  biolprm = biolprm,
+                  select_groups = select_groups,
+                  select_variable = "Discard",
+                  check_acronyms = TRUE,
+                  bboxes = boxes, 
+                  verbose = TRUE)
+
+fishdiscage
+```
+
+Do we want catch by fishery? No fishery column here, check what it is doing.
 

### 2020-05-22 R/load_fisheries.R

@@ -1,0 +1,29 @@
+#' Load the functional group file
+#'
+#' Read in the functional group file, which is typically saved as
+#' \code{functionalGroups.csv}.
+#'
+#' @template dir
+#' @template file_fish
+#'
+#' @author Alexander Keth
+#' @export
+#' @family load functions
+#'
+#' @return A \code{data.frame} of functional group information.
+#'
+#' @examples
+#' d <- system.file("extdata", "INIT_VMPA_Jan2015", package = "atlantisom")
+#' fgs <- load_fgs(d, "functionalGroups.csv")
+#' rm(fgs)
+#'
+load_fgs <- function(dir = getwd(), file_fish){
+  if (is.null(dir)) {
+    file.fish <- file_fish
+  } else {
+    file.fish <- file.path(dir, file_fish)
+  }
+  result <- read.table(file = file.fish, sep = ",",
+                       header = TRUE, stringsAsFactors = FALSE)
+  return(result)
+}

### 2020-05-22 R/load_nc_annage.R

@@ -1,0 +1,366 @@
+#' Load Atlantis annual age outputfiles (netcdf)
+#'
+#' This function loads Atlantis ANNAGEBIO and ANNAGECATCH outputfiles (netcdf) and
+#' converts them to a \code{data.frame}.
+#' @template dir
+#' @template file_nc
+#' @template bps
+#' @template fgs
+#' @template biolprm
+#' @template select_groups
+#' @param select_variable A character value spefifying which variable to return.
+#'   loaded. Only one variable of the options available (i.e., \code{c(
+#'   "Nums", "Weight")
+#'   }) can be loaded at a time.
+#' @param check_acronyms A logical, specifying if selected groups are active
+#'   in the model run. This is used in automated runs.
+#'   All groups are passed when plotting is called via batch-file,
+#'   (which cannot be changed easily) this will result in errors if some
+#'   groups are not active in the model run.
+#'   By default this argument is \code{TRUE}.
+#' @template bboxes
+#' @template verbose
+#' @family load functions
+#' @return A \code{data.frame} in long format with the following coumn names:
+#'   Species, timestep, polygon, agecl, and atoutput (i.e., variable).
+#'
+#' @keywords gen
+#' @author Alexander Keth, modified by Sarah Gaichas
+#'
+#' @examples
+#' d <- system.file("extdata", "INIT_VMPA_Jan2015", package = "atlantisom")
+#' fgs <- load_fgs(d, "functionalGroups.csv")
+#' bps <- load_bps(dir = d, fgs = fgs, file_init = "INIT_VMPA_Jan2015.nc")
+#' test <- load_nc(dir = d, file_nc = "outputSETASCATCH.nc",
+#'   fgs = fgs, bps = bps,
+#'   select_variable = "Catch",
+#'   select_groups = "Pisciv_T_Fish",
+#'   check_acronyms = TRUE)
+#' str(test)
+#' rm(test)
+#'
+#' @importFrom magrittr %>%
+#' @export
+
+load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm, select_groups,
+  select_variable =
+  c("Nums", "Weight", "Catch", "Discard"),
+  check_acronyms = TRUE, bboxes = c(0), verbose = FALSE) {
+  # NOTE: The extraction procedure may look a bit complex...
+  # A different approach would be to
+  # create a dataframe for each variable (e.g. GroupAge_Nums)
+  # and combine all dataframes at the end.
+  # This alternative approach requires a lot more storage
[…]

### 2020-05-22 docs/index.Rmd
add to index
correct link
update index visualization date
update nobacod html

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/NOBAcod.html" target="_blank">Apply and extend wrappers to get full age structure, started March 18, 2020</a>
+<a href="https://sgaichas.github.io/poseidon-dev/NOBAcod.html" target="_blank">Apply and extend wrappers to get full age structure, updated May 22, 2020</a>
 
@@
 <a href="https://sgaichas.github.io/poseidon-dev/WrapperDevelopment.html" target="_blank">Automating om functions to streamline workflows, January 17, 2020 writes dat file as of March 16, 2020</a>
+
+### Visualize functions for documentation
+
+<a href="https://sgaichas.github.io/poseidon-dev/Visualize_atlantisom.html" target="_blank">Different packages to visualize `atlantisom` functions, updated May 21, 2020</a>
 

### 2020-05-26 R/load_fisheries.R
name function correctly, update description

@@ -1,7 +1,7 @@
-#' Load the functional group file
+#' Load the fisheries group file
 #'
-#' Read in the functional group file, which is typically saved as
-#' \code{functionalGroups.csv}.
+#' Read in the fisheries group file, which is typically saved as
+#' \code{fisheries.csv}.
 #'
 #' @template dir
 #' @template file_fish
@@ -14,10 +14,10 @@
 #'
 #' @examples
 #' d <- system.file("extdata", "INIT_VMPA_Jan2015", package = "atlantisom")
-#' fgs <- load_fgs(d, "functionalGroups.csv")
-#' rm(fgs)
+#' fleets <- load_fisheries(d, "fisheries.csv")
+#' rm(fleets)
 #'
-load_fgs <- function(dir = getwd(), file_fish){
+load_fisheries <- function(dir = getwd(), file_fish){
   if (is.null(dir)) {
     file.fish <- file_fish
   } else {

### 2020-05-26 R/load_nc_annage.R
add a fleet column to fishery output--partial

@@ -65,6 +65,8 @@ load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm,
 
   # Check input of select_variable as only one value is allowed
   select_variable <- match.arg(select_variable, several.ok = FALSE)
+  
+  input_select_variable <- select_variable
 
   # Check input structure!
   if (check_acronyms) {
@@ -113,6 +115,7 @@ load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm,
   # ADDED fleets to select_variable for fishery output
 
   if(select_variable %in% c("Catch", "Discard")){
+    #input_select_variable <- select_variable
     #read in fleet names
     fleetnames <- load_fisheries(dir = dir, file_fish = file_fish)
     #select_variable <- concatenate select_variable "_" each fleet name
@@ -132,9 +135,9 @@ load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm,
   names(biolprm$agespercohort) <- c("code", "ageperagecl")
 
   maxage <- merge(biolprm$maturityogive, biolprm$agespercohort, all.x = T) %>%
-    select(code, nagecl, ageperagecl) %>%
-    mutate(maxage = nagecl * ageperagecl) %>%
-    mutate(name = fgs$Name[match(code, fgs$Code)])
+    dplyr::select(code, nagecl, ageperagecl) %>%
+    dplyr::mutate(maxage = nagecl * ageperagecl) %>%
+    dplyr::mutate(name = fgs$Name[match(code, fgs$Code)])
 
   search <- list()
   for (i in seq_along(select_groups)) {
@@ -170,6 +173,12 @@ load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm,
     lapply(select_groups, grepl, x = search_clean), any)]
   final_agecl <- maxage$maxage[
     sapply(final_species, function(x) which(x == maxage$name))]
+
+  # Get final fleets for full age structure fishery output 
+  if(input_select_variable %in% c("Catch", "Discard")){
+    final_fleet <- fleetnames$Code[sapply(
+      lapply(select_variable, grepl, x = search_clean), any)]
+  }
 
   num_layers <- RNetCDF::var.get.nc(ncfile = at_out, variable = "numlayers")[, 1]
   # add sediment layer!
[…]

### 2020-05-26 NOBAcod.Rmd
catch at age with fleet, needs testing

@@
 
-```{r addsnippet}
+```{r addsnippet, eval=FALSE}
 
@@
 
+Also added if statements to add fleets for fisheries outputs, if statements to separate output for annage and annagecatch.
+
 Test new `load_nc_annage`:
@@
 
-fgs <- NOBAom$funct.groups
-biolprm <- NOBAom$biol
-select_groups <- c('Mackerel', 'North_atl_cod', 'Capelin')
+source(here("config/NOBA2Config.R"))
+NOBAom_cod <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+
+fgs <- NOBAom_cod$funct.group_ss
+biolprm <- NOBAom_cod$biol
+#select_groups <- c('Mackerel', 'North_atl_cod', 'Capelin')
+select_groups <- 'North_atl_cod'
+
+bps <- load_bps(dir = d.name, fgs = functional.groups.file, file_init = biomass.pools.file)
+allboxes <- load_box(dir = d.name, file_bgm = box.file)
+boxes <- get_boundary(allboxes)
 
@@
 
-Do we want catch by fishery? No fishery column here, check what it is doing.
+Do we want catch by fishery? Yes. No fishery column here, check what it is doing.
+fishdiscage is empty, make sure there is no discard output from NOBA for these species.
+
+Add a check for multiple fisheries per species, if only one then fleet irrelevant.
+
+NOBA has multiple fleets for cod, so add fleet column to output.
+
+Update: there is now a fleet column in output but the function is still incomplete; will not work if more than one fleet per species.
 

### 2020-05-27 NOBAcod.Rmd
load_nc_annage works for fisheries but only one fleet per species right now

@@
 
-```{r load_fisheries}
+```{r load_fisheries, eval=FALSE}
 
@@
 
-source(here("R/load_nc_annage.R"))
-source(here("R/load_fisheries.R"))
+# source(here("R/load_nc_annage.R"))
+# source(here("R/load_fisheries.R"))
 
@@
 file_fish <- "NoBAFisheries.csv"
+
+#for testing line by line
+bboxes <- boxes
+dir <- d.name
+select_variable <- "Catch"
 
@@
 
-fishcatchage
+fishcatchage[1:100,]
 
@@
 Do we want catch by fishery? Yes. No fishery column here, check what it is doing.
-fishdiscage is empty, make sure there is no discard output from NOBA for these species.
+fishdiscage is empty, make sure there is no discard output from NOBA for these species (confirmed).
 
@@
 
-NOBA has multiple fleets for cod, so add fleet column to output.
+NOBA has multiple fleets for cod (but they are not in output!), so add fleet column to output.
 
 Update: there is now a fleet column in output but the function is still incomplete; will not work if more than one fleet per species.
+
+Plot catch at age; aggregate over polygons, apply sample_fish:
+
+```{r catchannage}
+
+omlist_ss<-NOBAom_cod
+source(here("/config/NOBA2Config.R"))
+source(here("/config/omdimensions.R"))
+source(here("/config/codsurvey.R"))
+source(here("/config/codfishery.R"))
+
+n_reps <- 1
+
+  #numbers based fishery independent survey for age and length comps
+  # same user specifications as indices
+  #fishery catch at age each observed timestep summed over observed polygons
+  # catch at age by area and timestep
+  catch_numbers <-  atlantisom::create_fishery_subset(dat = fishcatchage,
+                                                      time = fishtime,
+                                                      species = select_groups,
+                                                      boxes = fishboxes)
+
+  # if we want replicates for obs error this sample function will generate them
+  catch_age_comp <- list()
+  for(i in 1:n_reps){
+    catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
+  }
+
+  # save fishery age comps
+  #if(save){
+    saveRDS(catch_age_comp, file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
+  #}
+
+```
+
+NOTE that the fleet column is lost in the `atlantiom::sample_fish` function!
+
+Look at only every 10 years output (this is a sample, effN = 1000 for the year, which is 200 per timestep here; see calc below confirming):
+
+```{r plotfullcatchage}
+
+#read in comp data
+# age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
+# age_comp_data <- age_comp_data[[1]]
+
+fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
+fish_age_comp <- fish_age_comp[[1]]
+
+#add this to om_indices function so that this has years when read in
+fish_age_comp$time <- fish_age_comp$time/fstepperyr
+
+Catage <- filter(fish_age_comp, time %in% c(30.0, 40.0, 50.0, 60.0, 70.0, 79.8))
+
+Catageplot <- ggplot(Catage, aes(x=agecl, y=atoutput)) +
+    geom_point() +
+    theme_tufte() +
+    labs(subtitle = paste(scenario.name,
+                          Catage$species))
+
+  Catageplot + facet_wrap(~time, ncol=3, scales="free_y")
+
+```
+
+```{r testeffn}
+
+fish_age_comp %>%
+  filter(time<31) %>%
+  summarise(sum(atoutput))
+
+```
+
+Appears to be working for single fleet per species now.
 

### 2020-05-27 R/load_nc_annage.R

@@ -66,6 +66,7 @@ load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm,
   # Check input of select_variable as only one value is allowed
   select_variable <- match.arg(select_variable, several.ok = FALSE)
   
+  # select_variable gets renamed for fishery data, save input
   input_select_variable <- select_variable
 
   # Check input structure!
@@ -248,12 +249,15 @@ load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm,
   if(input_select_variable %in% c("Catch", "Discard")){
     int_ff <- final_fleet
     # multiple fleets per species?
-    if(length(int_ff)>length(int_fs)){
-      # figure out which species
-      # increase species dimension by n fleets for that species
-    }
+    # safest: grep species code from int_ff and match to species name
+    spcode <- stringr::str_extract(int_ff, fgs$Code)
+    spcode <- table(spcode)
+    # find nfleet for that species
+    sp_nfleet <- data.frame(species = int_fs,
+                            code = fgs$Code[fgs$Name==int_fs])
+    sp_nfleet$nfleet <- spcode[names(spcode)==sp_nfleet$code]
   }
-
+  
   if (length(at_data3d) >= 1) {
     # Remove biomasspools if selected variable is "N"!
     if (select_variable == "N") {
@@ -311,14 +315,15 @@ load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm,
     }
 
     # Order of the data in value column = "atoutput".
-    # 1. species  --> rep each with the number of
+    # 1. species  --> rep each with the number of (fleets*)
     #                 ageclasses and n_timesteps * boxes
-    # 2. age      --> rep each (1:maxage for each species) with n_timesteps * boxes
+    # 2. age      --> rep each (1:maxage for each species(*fleet)) with n_timesteps * boxes
     # 3. timestep --> rep each timestep (1:n_timesteps)
     #                 with the number of boxes and final_agecl
-    #                 (num ages per species)
+    #                 (num ages per species(*fleet))
     # 4. polygon  --> rep boxes times n_timesteps * final_agecl
-    #                 (num ages per species)
+    #                 (num ages per species(*fleet))
+    # 5. fleet    --> 
 
     if(input_select_variable %in% c("Nums", "Weight")){
       
[…]

### 2020-05-27 NOBAcod.Rmd
run truth still breaking at fleets

@@
 
+Quick test of `run_truth()` with annage option turned on for NOBA:
+
+```{r testruntruth}
+
+source(here("config/NOBA2Config.R"))
+
+  #Load functional groups
+  funct.groups <- atlantisom::load_fgs(dir=d.name,
+                           file_fgs = functional.groups.file)
+  #Get just the names of active functional groups
+  funct.group.names <- funct.groups %>% 
+    filter(IsTurnedOn == 1) %>%
+    select(Name) %>%
+    .$Name
+
+  # # load biolprm file
+  #biol <- atlantisom::load_biolprm(d.name, biol.prm.file)
+  # 
+  # # load run_prm
+  # runpar <- atlantisom::load_runprm(d.name, run.prm.file)
+  # 
+  # # load box info file
+  # boxpars <- atlantisom::load_box(d.name, box.file)
+
+testtruth <- atlantisom::run_truth(scenario = scenario.name,
+                       dir = d.name,
+                       file_fgs = functional.groups.file,
+                       file_bgm = box.file,
+                       select_groups = funct.group.names,
+                       file_init = initial.conditions.file,
+                       file_biolprm = biol.prm.file,
+                       file_runprm = run.prm.file,
+                       file_fish = fisheries.file,
+                       verbose = TRUE,
+                       annage = TRUE,
+                       save = FALSE
+    )
+
+```
+
 ## Write input data for specific assessment models

### 2020-05-27 config/NOBA2Config.R

@@ -8,3 +8,5 @@ run.prm.file <- "nordic_run_v01.xml"
 scenario.name <- "nordic_runresults_01"
 bioind.file <- "outputnordic_runresults_01BiomIndx.txt"
 catch.file <- "outputnordic_runresults_01Catch.txt"
+annage <- TRUE
+fisheries.file <- "NoBAFisheries.csv"

### 2020-05-28 R/load_nc_annage.R
fixed n fleets for multiple species, still not dimensioned correctly

@@ -251,15 +251,18 @@ load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm,
   int_fa <- final_agecl
   
   if(input_select_variable %in% c("Catch", "Discard")){
-    int_ff <- final_fleet
-    # multiple fleets per species?
-    # safest: grep species code from int_ff and match to species name
-    spcode <- stringr::str_extract(int_ff, fgs$Code)
-    spcode <- table(spcode)
-    # find nfleet for that species
-    sp_nfleet <- data.frame(species = int_fs,
-                            code = fgs$Code[fgs$Name==int_fs])
-    sp_nfleet$nfleet <- spcode[names(spcode)==sp_nfleet$code]
+    # how many unique fleets per select_group in search_clean?
+    # lookup of species and fleet names
+    splitfleets <- search_clean %>%
+      str_replace("\\d+_*Catch_", "-")
+    sp_fleet <- as.data.frame(unique(splitfleets)) %>%
+      rename(spfleet = "unique(splitfleets)") %>%
+      separate(spfleet, c("species", "fleet"), sep = "-") 
+    # lookup of number of fleets per species
+    sp_nfleet <- as.data.frame(table(sp_fleet$species)) %>%
+      rename(species = Var1, nfleet = Freq)
+    # for indexing fleets
+    int_ff <- sp_nfleet$nfleet
   }
   
   if (length(at_data3d) >= 1) {
@@ -319,14 +322,14 @@ load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm,
     }
 
     # Order of the data in value column = "atoutput".
-    # 1. species  --> rep each with the number of (fleets*)
-    #                 ageclasses and n_timesteps * boxes
-    # 2. age      --> rep each (1:maxage for each species(*fleet)) with n_timesteps * boxes
+    # 1. species  --> rep each with the number of 
+    #                 ageclasses * fleets and n_timesteps * boxes
+    # 2. age      --> rep each (1:maxage for each species) with n_timesteps * boxes
     # 3. timestep --> rep each timestep (1:n_timesteps)
     #                 with the number of boxes and final_agecl
-    #                 (num ages per species(*fleet))
+    #                 (num ages per species)
     # 4. polygon  --> rep boxes times n_timesteps * final_agecl
-    #                 (num ages per species(*fleet))
[…]

### 2020-05-29 R/load_nc_annage.R
dimensions for multifleet appear to work

@@ -350,19 +350,19 @@ load_nc_annage <- function(dir = getwd(), file_nc, file_fish, bps, fgs, biolprm,
       if (select_variable == "N") result2d$layer <- n_layers - 1
     }
     
-    #currently works with only 1 fleet, need to dimension properly for multiple
+    #should now work properly for multiple fleets
     if(input_select_variable %in% c("Catch", "Discard")){
       result2d <- data.frame(species = unlist(sapply(X = mapply(FUN = rep, x = int_fs, 
                    each = (int_fa * int_ff),SIMPLIFY = FALSE,USE.NAMES = FALSE),
         FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
-        agecl = unlist(sapply(X = sapply(X = int_fa, FUN = seq, from = 1,
+        agecl = unlist(sapply(X = sapply(X = rep(int_fa, int_ff), FUN = seq, from = 1,
                                          by = 1, simplify = FALSE, USE.NAMES = FALSE),
-                              FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
+                              FUN = rep, each = (length(boxes) * n_timesteps), simplify = FALSE)),
         polygon = unlist(sapply(X = n_timesteps * int_fa * int_ff,
                                 FUN = rep, x = boxes, simplify = FALSE, USE.NAMES = FALSE)),
-        #fleet = unlist(sapply(X = mapply(FUN = rep, x = int_ff, each = int_fa,
-        #                                 SIMPLIFY = FALSE,USE.NAMES = FALSE),
-        #  FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
+        fleet = unlist(sapply(X = mapply(FUN = rep, x = sp_fleet$fleet, 
+                                         each = (rep(int_fa, int_ff)),SIMPLIFY = FALSE,USE.NAMES = FALSE),
+          FUN = rep, each = (length(boxes) * n_timesteps), simplify = FALSE)),
         time = unlist(sapply(X = int_fa * int_ff , FUN = rep, x = rep(0:(n_timesteps - 1),
                                                             each = length(boxes)), simplify = FALSE, USE.NAMES = FALSE)),
         atoutput = do.call(c, result2d),

### 2020-05-29 NOBAcod.Rmd
save new run_truth output

@@
                        annage = TRUE,
-                       save = FALSE
+                       save = TRUE
     )

### 2020-06-01 NOBAcod.Rmd
try wrappers with new output

@@
 
-#### Age composition (age classes)
+#### Survey age composition (age classes)
 ```{r agecomp-1}
@@
 #### Survey weight at age (age classes)
-
 ```{r wtageclass-1}
@@
 
+#### Fishery age composition (age classes)
+```{r agecomp-1}
+
+#read in comp data
+age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
+age_comp_data <- age_comp_data[[1]]
+
+fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
+fish_age_comp <- fish_age_comp[[1]]
+
+#add this to om_indices function so that this has years when read in
+fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
+
+Natage <- filter(fish_age_comp, time %in% c(55:175))
+
+Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
+    geom_point() +
+    theme_tufte() +
+    labs(subtitle = paste(scenario.name,
+                          Natage$species))
+
+  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+```
+
 ### {-}
 
-Wrappers work for NOBA cod using standard age classes. Now to modify the underlying code to get full age structure.
+Wrappers work for NOBA cod using standard age classes.
+
+Visualize new annual age output for numbers at age, catch at age, and weight at age (interpolated):
+
+### Wrapper test results for annual ages {#annageoutput}{.tabset}
+
+#### Survey age composition (annual ages)
+```{r plot-annagecomp}
+
+```
+
+#### Survey weight at age (annual ages)
+```{r plot-annwtage}
+
+```
+
+#### Fishery catch at age (annual ages)
+```{r plot-catchagecomp}
+
+```
+
+### {-}
+
+# (Annage development details below)
+
+Now to modify the underlying code to get full age structure.
 
@@
 
-Update: there is now a fleet column in output but the function is still incomplete; will not work if more than one fleet per species.
+Update: there is now a fleet column in output which should index correctly if a species is caught by multiple fleets.
 
@@
 
-NOTE that the fleet column is lost in the `atlantiom::sample_fish` function!
+NOTE that the fleet column is lost in the `atlantisom::sample_fish` function!
 
@@
 
-```{r testruntruth}
+```{r testruntruth, eval=FALSE}
 
@@
 
+After much debugging to properly index multiple fleets per species in catch at annual age output, this produced a run_truth.Rdata file and took about 40 minutes to run.
+
+Now all of the wrappers need to read from annage to produce composition data.
+`om_init()`  now takes annage flag from the config file.
+`om_species()` looks for annage bio and catch output in the truth file and includes it in the species file (adding NULL if it isn't there).
+
+Also still need the weight at age interpolation function. Incorporate this at the `om_comps()` step. Next rmd!
+
+Running om_init, om_species, om_index, om_comps in chunk above, plot the annage output [above](#annageoutput).
+
 ## Write input data for specific assessment models

### 2020-06-01 NOBAcod.Rmd
add figs for annage from wrappers

@@
 
-fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
-fish_age_comp <- fish_age_comp[[1]]
-
-#add this to om_indices function so that this has years when read in
-fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
-
-Natage <- filter(age_comp_data, time %in% c(55:175))
+Natage <- filter(age_comp_data, time %in% c(150:270))
 
@@
 #read in comp data
-age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
-age_comp_data <- age_comp_data[[1]]
-
 fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
@@
 #add this to om_indices function so that this has years when read in
-fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
-
-Natage <- filter(fish_age_comp, time %in% c(55:175))
+fish_age_comp$time <- fish_age_comp$time/fstepperyr
+
+Natage <- filter(fish_age_comp, time %in% c(30:53))
 
@@
 
-  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+  Natageplot + facet_wrap(~time, ncol=6, scales="free_y")
 ```
@@
 ```{r plot-annagecomp}
-
+#read in comp data
+annage_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
+annage_comp_data <- annage_comp_data[[1]]
+
+#fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
+#fish_age_comp <- fish_age_comp[[1]]
+
+#add this to om_indices function so that this has years when read in
+#fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
+
+Natage <- filter(annage_comp_data, time %in% c(150:270))
+
+Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
+    geom_point() +
+    theme_tufte() +
+    labs(subtitle = paste(scenario.name,
+                          Natage$species))
+
+  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
 ```
@@
 
+fish_annage_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
+fish_annage_comp <- fish_annage_comp[[1]]
+
+#add this to om_indices function so that this has years when read in
+fish_annage_comp$time <- fish_annage_comp$time/fstepperyr
+
+Catage <- filter(fish_annage_comp, time %in% c(30:53))
+
+Catageplot <- ggplot(Catage, aes(x=agecl, y=atoutput)) +
+    geom_point() +
+    theme_tufte() +
+    labs(subtitle = paste(scenario.name,
+                          Catage$species))
+
+  Catageplot + facet_wrap(~time, ncol=6, scales="free_y")
 ```
@@
 ### {-}
+
+Note: NOBA doesn't seem to have a selective fishery for cod. Lots of age 1 in the catch!
 

### 2020-06-01 config/NOBA2Config.R
dont use "output" prefix

@@ -6,7 +6,7 @@ box.file <- "Nordic02.bgm"
 initial.conditions.file <- "nordic_biol_v23.nc"
 run.prm.file <- "nordic_run_v01.xml"
 scenario.name <- "nordic_runresults_01"
-bioind.file <- "outputnordic_runresults_01BiomIndx.txt"
-catch.file <- "outputnordic_runresults_01Catch.txt"
+bioind.file <- "nordic_runresults_01BiomIndx.txt"
+catch.file <- "nordic_runresults_01Catch.txt"
 annage <- TRUE
 fisheries.file <- "NoBAFisheries.csv"

### 2020-06-01 NOBAcod.Rmd
NOBA cod data generation, all but WtAtAnnage

@@
 
-**Special note:** *Atlantis output files usually have the prefix "output" and `atlantisom` expects this.
-NOBA output files use a different naming convention.
-I've appended "output" to the beginning of each NOBA output filename for `atlantisom` to work.
-Need to update `atlantisom` functions for user-specified output filenames.*
+**Special note:** *Atlantis output files formerly had the prefix "output" and `atlantisom` was coded to expect this.
+NOBA output files (and newer Atlantis models) do not universally use the "output" prefix naming convention.
+Annage branch `atlantisom` functions have been updated for user-specified output filenames.*
 
@@
 
-### Using the wrappers
+### Using the wrappers (now updated for annual age)
 
@@
 #### Fishery age composition (age classes)
-```{r agecomp-1}
+```{r fishagecomp-1}
 
@@
 
-Wrappers work for NOBA cod using standard age classes.
+Wrappers still work for NOBA cod using standard age classes.
 
@@
 
-### Wrapper test results for annual ages {#annageoutput}{.tabset}
+### Wrapper test results for annual ages {.tabset #annageoutput}
 
@@
 
-#### Survey weight at age (annual ages)
+#### Survey weight at age (annual ages) pending new function
 ```{r plot-annwtage}
@@
 
-UPDATE: there is a difference between the annage weight at age and that calculated using 'calc_age2length'.
+UPDATE: there is a difference between the annage weight at age and that calculated using 'calc_age2length()'.
 While the results from the standard nc output look reasonable for NOBA cod, the annage weight at age is half that for the standard (and half what real cod would be).
@@
 
-New function `load_fisheries` needed to load the fishery specification file (a .csv), similar to `load_fgs`:
+New function `load_fisheries()` needed to load the fishery specification file (a .csv), similar to `load_fgs()`:
 
@@
 
-Call `load_fisheries` within `load_nc_annage` to rename select_variable as used in ANNAGECATCH.nc:
+Call `load_fisheries()` within `load_nc_annage()` to rename select_variable as used in ANNAGECATCH.nc:
 
@@
 
-Test new `load_nc_annage`:
+Test new `load_nc_annage()`:
 
@@
 `om_species()` looks for annage bio and catch output in the truth file and includes it in the species file (adding NULL if it isn't there).
+`om_comps()` looks for annage bio and catch in the species file and adds them to the comps list if there (adding NULL if they aren't there).
 
@@
 
-Running om_init, om_species, om_index, om_comps in chunk above, plot the annage output [above](#annageoutput).
+Running `om_init()`, `om_species()`, `om_index()`, `om_comps()` in chunk above to make sure the original output is preserved, then plot the annage output [above](#annageoutput).
 

### 2020-06-01 docs/index.Rmd
NOBA cod data generation, all but WtAtAnnage

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/NOBAcod.html" target="_blank">Apply and extend wrappers to get full age structure, updated May 22, 2020</a>
+<a href="https://sgaichas.github.io/poseidon-dev/NOBAcod.html" target="_blank">Apply and extend wrappers to get full age structure, updated June 1, 2020</a>
 

### 2020-06-01 InterpolateWeightAtAge.Rmd
baby steps
pared down to use wrapper output for NOBA cod

@@
 ---
-title: "Testing atlantisom: what if we interpolate the weight at stage output to true age?"
-author: "Sarah Gaichas and Christine Stawitz"
+title: "Interpolate the weight at agecl output to annual age"
+author: "Sarah Gaichas"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
-bibliography: "packages.bib"
+output:
+  html_document:
+    code_fold: hide
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-# automatically create a bib database for R packages
-knitr::write_bib(c(
-  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
-  'data.table', 'here', 'ggforce', 'ggthemes'
-), 'packages.bib')
-```
-
-## Introduction
-
-Now we delve into the calculation of true age classes and weight-at-age from standard Atlantis output age class, which is generally in 10 stages of maxage/10 true age classes for all but the newest Atlantis models.
-Weight-at-true-age is needed as an input to stock assessment models using empirical weight at age.
-See discussion [here](https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html).
-We will test the `atlantisom` functions `calc_stage2age` and the associated `calc_Z` that estimate true numbers at age from numbers in output age class separately.
-However, `calc_stage2age` produces numbers at age but not weight.
-
-We already have 10 class weight at age (output as `mulength` from `calc_age2length`).
-We visualized weight at stage for selected NOBA and CCA species [here](https://sgaichas.github.io/poseidon-dev/TestWeightAtStage.html).
-In this page we will try the simplest interpolation to true age-based weight at age possible, and see if that will work for us.
-
-Note that modern Atlantis models can output true age classes in two output files (from the wiki):
-
->**ANNAGEBIO.nc**: This output provides numbers in each annual age class (so mapped from Atlantis "age class" which can contain multiple years to true annual age classes).
-Set flag_age_output to 1 to get this output.
-Tracers provided are:
->
->   + Numbers at age per species
->
->**ANNAGECATCH.nc**: This output provides numbers at annual age class (so mapped from Atlantis "age class" which can contain multiple years to true annual age classes) in the catch and discards (summed over all fleets).
-Set flag_age_output to 1 to get this output.
-Tracers provided are:
->
->   + Numbers at age per species in the catch
->   + Numbers at age per species in the discards
-
-Future `atlantisom` users should take advantage of this direct output, and some `atlantisom` functions will have to be rewritten to use it.
-
-For all setup, etc, please see previous files.
-Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html) and  [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).
-
-This page has visualizations for the NOBA model example, CERES Global Sustainability. At the end of the file we also review saved outputs from the CCA model example, Atlantis Summit Common Scenario 1.
-
-```{r message=FALSE, warning=FALSE}
-library(tidyr)
-require(dplyr)
-library(ggplot2)
-library(data.table)
+knitr::opts_chunk$set(echo = TRUE)
 library(here)
-library(ggforce)
-library(ggthemes)
+library(tidyverse)
 library(atlantisom)
-```
-
-```{r initialize}
-
-initCCA <- FALSE
-initNEUS <- FALSE
-initNOBA <- TRUE
-
-if(initCCA) source(here("config/CCConfig.R"))
-
-if(initNEUS) source(here("config/NEUSConfig.R"))
-
-if(initNOBA) source(here("config/NOBAConfig.R"))
 
@@
 
-```{r get_names, message=FALSE, warning=FALSE}
-#Load functional groups
-funct.groups <- load_fgs(dir=d.name,
-                         file_fgs = functional.groups.file)
-#Get just the names of active functional groups
-funct.group.names <- funct.groups %>% 
-  filter(IsTurnedOn == 1) %>%
-  select(Name) %>%
-  .$Name
+## Start with saved wrapper weight at age output for NOBA cod
 
-```
+```{r full-test, message=FALSE, warning=FALSE}
 
-```{r load_Rdata, message=FALSE, warning=FALSE}
+source(here("config/NOBA2Config.R"))
+NOBAom_cod <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
 
-if(initCCA) {
-  truth.file <- "outputCCV3run_truth.RData"
-  load(file.path(d.name, truth.file))
-  truth <- result
-}
+fgs <- NOBAom_cod$funct.group_ss
+biolprm <- NOBAom_cod$biol
 
-if(initNEUS) {
-  truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
-  load(file.path(d.name, truth.file))
-  truth <- result
-}
+wtage <- readRDS(file.path(d.name, paste0(scenario.name, "survObsWtAtAge.rds")))
+wtage <- wtage[[1]]
 
-if(initNOBA){
-  truth.file <- "outputnordic_runresults_01run_truth.RData" 
-  load(file.path(d.name, truth.file))
-  truth <- result
-}
-
-```
-
-## Simulate a survey part 5: interpolated weight at true age census
-
-Full methods to get the census weight at age by Atlantis output age class are explained [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).
-
-We will apply interpolation examples here to just a few NOBA species to examine within-year variability:
-    - Cod "North_atl_cod", likely a test assessment species
-    - Herring "Norwegian_ssh", likely a test assessment species
-    - Greenland halibut "Green_halibut", which grows to a large size.
-
-To create a census, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.
-The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages.
-
-```{r census-spec, message=FALSE, warning=FALSE}
-
-# should return a perfectly scaled survey 
-effic1 <- data.frame(species=funct.group.names,
-                     efficiency=rep(1.0,length(funct.group.names)))
-
-# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
-selex1 <- data.frame(species=rep(funct.group.names, each=10),
-                     agecl=rep(c(1:10),length(funct.group.names)),
-                     selex=rep(1.0,length(funct.group.names)*10))
-
-# should return all model areas
-boxpars <- load_box(d.name, box.file)
-boxall <- c(0:(boxpars$nbox - 1))
-
-# generalized timesteps all models
-runpar <- load_runprm(d.name, run.prm.file)
-noutsteps <- runpar$tstop/runpar$outputstep
-stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc
-
-timeall <- c(0:noutsteps)
-
-# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
-# for ecosystem indicator work test all species, e.g.
-survspp <- funct.group.names
-
-# for length and age groups lets just do fish and sharks
-# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
-if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)
-
-survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
-                           funct.groups$GroupType %in% c("FISH", "SHARK")]
-
-```
-
-Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just a few species in this case).
-The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.
-
-Because we don't want to wait 24 hours for this, we will look at only the first 112 time steps.
-
-```{r stdsurveyNbased-3spp, echo=TRUE}
-
-spp.name <- funct.group.names[funct.group.names %in% c("North_atl_cod",
-                                                       "Norwegian_ssh",
-                                                       "Green_halibut")]
-
-# get survey nums with full (no) selectivity
-survey_testNall <- create_survey(dat = truth$nums,
-                                 time = c(0:111),
-                                 species = spp.name,
-                                 boxes = boxall,
-                                 effic = effic1,
-                                 selex = selex1)
-
-# this one is high but not equal to total for numerous groups
-effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
-
-# apply default sample fish as before to get numbers
-numsallhigh <- sample_fish(survey_testNall, effNhigh)
-
-# aggregate true resn per survey design
-aggresnall <- aggregateDensityData(dat = truth$resn,
-                                 time = c(0:111),
-                                 species = spp.name,
-                                 boxes = boxall)
-
-# aggregate true structn per survey design
-aggstructnall <- aggregateDensityData(dat = truth$structn,
-                                 time = c(0:111),
-                                 species = spp.name,
-                                 boxes = boxall)
-
-#dont sample these, just aggregate them using median
-structnall <- sample_fish(aggstructnall, effNhigh, sample = FALSE)
-
-resnall <-  sample_fish(aggresnall, effNhigh, sample = FALSE)
-
-```
-
-Length sample with user specified max length bin (200 cm):
-
-```{r userset-maxlen, echo=TRUE}
-
-length_census <- calc_age2length(structn = structnall,
-                                 resn = resnall,
-                                 nums = numsallhigh,
-                                 biolprm = truth$biolprm, fgs = truth$fgs,
-                                 maxbin = 200,
-                                 CVlenage = 0.1, remove.zeroes=TRUE)
-
-```
-
-We should get the upper end of Greenland halibut with a 200cm max length bin.
-That shouldnt matter for weight at age.
-All of these species would be split into two true ages per age class.
-Interpolation could be interesting.
-
-Change in wt at age over time for age classes using an annual mid-year snapshot (survey) (first 22+ years of NOBA model run):
-
-```{r aggwtcomp}
-
-# from std survey code and model step info read in above
-midptyr <- round(median(seq(1,stepperyr)))
-annualmidyear <- seq(midptyr, noutsteps, stepperyr)
-
-wtage_annsurv <- length_census$muweight %>%
-  filter(time %in% annualmidyear)
-
-# reverse to show agecl time series of wt
-wageplot <- ggplot(wtage_annsurv, aes(time, atoutput)) +
+wageplot <- ggplot(wtage, aes(time, atoutput)) +
   geom_line(aes(colour = factor(agecl))) +
@@
 
-fgs.dir <- here("atlantisoutput", "NOBACERESGlobalSustainability")
-fgs.file <-  "nordic_groups_v04.csv"
-wtagecl_data <- wtage_annsurv
+wtagecl <- wtage
 
-calc_wtTrueage <- function(wtagecl_data, fgs.file) {
+calc_wtTrueage <- function(wtagecl, fgs) {
   # get number of ages per agecl from fgs.file
-  fgs <- load_fgs(fgs.dir, fgs.file)
+  #fgs <- atlantisom::load_fgs(dir=d.name, file_fgs = functional.groups.file)
   species.code <- fgs$Code
-  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
+  #turnedon <- fgs[fgs$IsTurnedOn > 0, ]
 
@@
   # cohort) this is from the full model
-  multiple_ages <- turnedon[turnedon$NumAgeClassSize>1, c(1,4,10)]
+  multiple_ages <- fgs[fgs$NumAgeClassSize>1, c(1,4,10)]
 
   # match species in wtagecl_data to those species in the fgs file
-  sppwt <- levels(wtagecl_data$species)
+  sppwt <- levels(wtagecl$species)
   multiple_ages <- multiple_ages[multiple_ages$Name %in% sppwt, ]
@@
   # timesteps from input wt file
-  ntimesteps <- length(unique(wtagecl_data$time))
+  ntimesteps <- length(unique(wtagecl$time))
+
+  # make a dataframe the right dimensions for all age classes
+  wtage_out <- data.frame()
 

### 2020-06-02 InterpolateWeightAtAge.Rmd
computes avg age of agecl, agecl and time-specific avg wt increment

@@
 library(atlantisom)
+library(ggthemes)
 
@@
 wtagecl <- wtage
+annages <- NOBAom_cod$truenumsage_ss %>%
+  group_by(species, time, agecl) %>%
+  rename(trueage = agecl) %>%
+  summarise(truenatage = sum(atoutput))
 
-calc_wtTrueage <- function(wtagecl, fgs) {
+calc_wtTrueage <- function(wtagecl, annages, fgs) {
   # get number of ages per agecl from fgs.file
@@
   # make a dataframe the right dimensions for all age classes
-  wtage_out <- data.frame()
+  # has species, time from wtage and trueage, truenatage from annages
+  # add agecl column
+  # finds the average age of each agecl each timestep based on truenatage
+  wtage_out <- annages %>%
+    dplyr::semi_join(wtagecl, by = c("species", "time")) %>%
+    dplyr::left_join(multiple_ages, by = c("species" = "Name")) %>%
+    dplyr::mutate(agecl = as.integer(ceiling(trueage/NumAgeClassSize))) %>%
+    dplyr::group_by(species, time, agecl) %>%
+    dplyr::mutate(avgage = weighted.mean(trueage, truenatage))
 
@@
   # muweight for agecl+1 - muweight for agecl
+  wtagecl_inc <- wtagecl %>%
+    dplyr::arrange(species, time, agecl) %>%
+    dplyr::left_join(multiple_ages, by = c("species" = "Name")) %>%
+    dplyr::group_by(species, time) %>%
+    dplyr::mutate(increment = case_when(
+      agecl==1 ~ atoutput,
+      agecl>1 ~ atoutput - dplyr::lag(atoutput)
+      )) %>%
+    dplyr::mutate(agewtinc = increment/NumAgeClassSize)
 
-  # divide weight increment by number of ages per agecl (could decay with increasing age, but how different from equal increment?)
+  # last line above does this: divide weight increment by number of ages per agecl 
+  # definitely wrong, especially at younger ages, for two reasons:
+  #  - assumes each agecl has equal numbers of its constituent ages
+  #  - assumes linear growth between age classes
+  # (could decay with increasing age, but how different from equal increment?)
+  # avgWtAge for agecl 1 = (n age1 * avgwt age1) + (n age2 * avgwt age2)
 

### 2020-06-02 InterpolateWeightAtAge.Rmd
moderately successful wt at age interpolation

@@
 
-calc_wtTrueage <- function(wtagecl, annages, fgs) {
+#calc_wtTrueage <- function(wtagecl, annages, fgs) {
   # get number of ages per agecl from fgs.file
@@
 
+  wtage_avgage <- wtage_out %>%
+    select(species, time, agecl, avgage) %>%
+    distinct()
+
   # find weight increment for a timestep
@@
     dplyr::arrange(species, time, agecl) %>%
-    dplyr::left_join(multiple_ages, by = c("species" = "Name")) %>%
+    #dplyr::left_join(multiple_ages, by = c("species" = "Name")) %>%
     dplyr::group_by(species, time) %>%
@@
       )) %>%
-    dplyr::mutate(agewtinc = increment/NumAgeClassSize)
+    dplyr::left_join(wtage_avgage) %>%
+    dplyr::mutate(avgageinc = case_when(
+      agecl==1 ~ avgage,
+      agecl>1 ~ avgage - dplyr::lag(avgage)
+      ))
 
-  # last line above does this: divide weight increment by number of ages per agecl 
-  # definitely wrong, especially at younger ages, for two reasons:
-  #  - assumes each agecl has equal numbers of its constituent ages
-  #  - assumes linear growth between age classes
-  # (could decay with increasing age, but how different from equal increment?)
-  # avgWtAge for agecl 1 = (n age1 * avgwt age1) + (n age2 * avgwt age2)
+  # complex, but works for 2 ages/agecl, need to test, need to fix oldest
+  wtage_out <- wtage_out %>%
+    dplyr::select(species, time, trueage, agecl, avgage) %>%
+    dplyr::left_join(wtagecl_inc) %>%
+    dplyr::group_by(species, time) %>%
+    dplyr::mutate(wtIntage = case_when(
+      trueage==1 ~ trueage/avgage*increment,
+      (trueage>1 & trueage<avgage) ~ 
+        (1-(avgage-trueage)/avgageinc)*increment + lag(atoutput),
+      (trueage>1 & trueage>avgage) ~ 
+        ((trueage-avgage)/lead(avgageinc))*lead(increment) + atoutput
+        ))
 
-  # recursive: add increment to agecl until hitting agecl+1
+  # plot this--how bad would piecewise linear interpolation be?
+  wtageclann <- ggplot(wtagecl_inc, aes(avgage, atoutput)) +
+    geom_point() +
+    geom_line(aes(colour = factor(time))) + 
+    scale_x_continuous(minor_breaks = c(0:20)) +
+    theme_tufte() +
+    theme(legend.position = "none",
+          panel.grid.minor.x = element_line(colour = "grey50"),
+          panel.grid.major.x = element_line(colour = "grey50"))
 
-  # reunumber agecl 1 to 10*number of ages per agecl
+  wtageclann + geom_point(data = wtage_out, 
+             mapping = aes(trueage, wtIntage))
+  # maybe not so bad
 
@@
 
-}
+#}
 ```
 
-Plot interpolated results for NOBA species:
+Plot interpolated results for more NOBA species:
 
@@
 ```
-
-## Quick test with saved CCA census output
-
-Here we check the CCA model (Atlantis Summit Common Scenario 1) saved length and weight census output for small species at Isaac's suggestion (all three of the following are true age classes, NumAgeClassSize=1 per Atlantis agecl bin:
-
-```{r loadCCAlengthcomp, echo=TRUE}
-
-source(here("config/CCConfig.R"))
-
-length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))
-
-```
-
-```{r apply-inter-pwtage-CCA}
-
-```
-
-Plots for species with 1 true age per agecl should not be different:
-
-Sardine:  
-```{r plotCCAsardine}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Pacific_sardine")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-Anchovy:  
-```{r plotCCAanchovy}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Anchovy")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-Herring:  
-```{r plotCCAherring}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Herring")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-How do these three look? Identical I hope.
-
-Now for some CCA groups with multiple true ages interpolated per stage.
-
-Pacific hake, 2 true ages per class (20 total):  
-```{r plotCCAhake}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Mesopel_M_Fish")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-Bocaccio rockfish, 5 true ages per age class (50 total):
-```{r plotCCAbocaccio}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Bocaccio_rockfish")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-Yelloweye rockfish, most extreme at 12 true ages per age class (120 total):
-```{r plotCCAYelloweye}
-censuswt <- length_censussurvsamp$muweight %>%
-  filter(species == "Yelloweye_rockfish")
-
-wageplot <- ggplot(censuswt, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (1 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year survey"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-Yelloweye should be a good test...

### 2020-06-03 InterpolateWeightAtAge.Rmd
final dataset in atlantisom form

@@
 
-  # complex, but works for 2 ages/agecl, need to test, need to fix oldest
+  # complex, but works for 2 ages/agecl, need to test more, need to fix oldest
   wtage_out <- wtage_out %>%
@@
 
+  # fix oldest age
+
+  #clean up to have only standard columns
+  finaldata <- data.frame("species" = wtage_out$species,
+                          "agecl" = wtage_out$trueage, "polygon" = NA, "layer" = NA,
+                          "time" = wtage_out$time, "atoutput" = wtage_out$wtIntage)
+
   # output full set of muweight at newly created agecl

### 2020-06-03 InterpolateWeightAtAge.Rmd
testing for 3 species, range of true ages

@@
 
-Plot interpolated results for more NOBA species:
+Run wrappers and plot interpolated results for multiple NOBA species:
 
@@
 
+NOBAom <- om_init(here("config/NOBA2config.R"))
+
+# 3 species: 40, 20, 6 age classes
+NOBAom_redcodcap <- om_species(c("Redfish", "North_atl_cod", "Capelin"), NOBAom)
+
+rm(NOBAom) #free some memory
+
 ```
+
+Fix bug in omdimensions.R; need to vectorize:
+
+```{r omdimensions-bugfix}
+
+omlist_ss <- NOBAom_redcodcap # multispecies call
+
+n_age_classes <- omlist_ss$funct.group_ss$NumCohorts # works
+age_classes <- 1:n_age_classes # nope, only first spp, needed in SS3 writing?
+
+age_classes <- sapply(n_age_classes, seq)
+names(age_classes)<-survspp
+
+n_annages <- n_age_classes * omlist_ss$funct.group_ss$NumAgeClassSize # works
+annages <- 1:n_annages # nope, only first spp, needed in SS3 writing?
+
+annages <- sapply(n_annages, seq)
+names(annages)<-survspp
+
+# dimensioning survey in 3sppsurvey.R
+survspp <- omlist_ss$species_ss # works
+surveffic <- data.frame(species=survspp,
+                     efficiency=rep(1.0,length(survspp)))  # works
+survselex <- data.frame(species=rep(survspp, each=n_annages), # nope! 
+                        agecl=rep(c(1:n_annages),length(survspp)),
+                        selex=rep(1.0,length(survspp)*n_annages))
+surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp))) # works
+surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp))) # works
+
+#dimensioned for annual ages rather than age classes
+survselex <- data.frame(species=rep(survspp, n_annages), #  
+                        agecl=unlist(sapply(n_annages,seq)),
+                        selex=rep(1.0,sum(n_annages)))
+
+```
+
+New config files needed for multispecies pulls:
+
+`3sppsurvey.R` configures the fishery independent survey and looks like this:
+
+```{r, code = readLines("./config/3sppsurvey.R"), eval=F}
+```
+
+`3sppfishery.R` configures the fishery dependent data and looks like this:
+
+```{r, code = readLines("./config/3sppfishery.R"), eval=F}
+```
+
+Fix multispecies bug in om_index function:
+```{r omindex-bugfix}
+
+# original  
+wtage <- data.frame(species=rep(survspp, each=max(age_classes)), #nope
+                      agecl=rep(c(1:max(age_classes)),length(survspp)), #nope
+                      wtAtAge=rep(1000.0,length(survspp)*max(age_classes))) #nope
+
+# age_classes is now a list 1:n_age_classes by species
+# this is still dimensioned by agecl not true age
+wtage <- data.frame(species=rep(survspp, n_age_classes), 
+                      agecl=unlist(sapply(n_age_classes,seq)), 
+                      wtAtAge=rep(1000.0,sum(n_age_classes)))
+
+```
+
+Try index step again:
+
+```{r indcomp-test}
+
+NOBAom_redcodcap_ind <- om_index(usersurvey = here("config/3sppsurvey.R"), 
+                           userfishery = here("config/3sppfishery.R"),
+                           omlist_ss = NOBAom_redcodcap, 
+                           n_reps = 1, 
+                           save = TRUE)
+
+```
+
+om_index works, om_comps runs for survey and produces age, length, weight output, produces fishery age output, but fails calling the age2length function.
+
+check fishery observed age comp:
+
+```{r debugcomps, eval=FALSE}
+
+catch_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
+catch_age_comp <- catch_age_comp[[1]]
+
+filter(catch_age_comp, species=="North_atl_cod")
+filter(catch_age_comp, species=="Redfish")
+filter(catch_age_comp, species=="Capelin")
+
+#snippets from calc_age2length
+nums <- filter(catch_age_comp, species=="Capelin")
+#calculate mean length at age
+mulen <- nums
+muweight <- nums
+
+#needed stuff from om_comps
+#Get catch weights for length comp calc
+  # aggregate true resn per fishery subset design
+  catch_aggresnss <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
+                                                      time = fishtime,
+                                                      species = survspp,
+                                                      boxes = fishboxes)
+
+  # aggregate true structn fishery subsetdesign
+  catch_aggstructnss <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
+                                                         time = fishtime,
+                                                         species = survspp,
+                                                         boxes = fishboxes)
+
+  #dont sample these, just aggregate them using median
+  structn <- atlantisom::sample_fish(catch_aggstructnss, fisheffN, sample = FALSE)
+
+  resn <- atlantisom::sample_fish(catch_aggresnss, fisheffN, sample = FALSE)
+
+  # more snippets from calc_age2length
+  # extract rows from structn and resn that match the rows in nums, which are only non-zeroes
+resn.ind <- with(resn,paste(species,'.',agecl,'.',polygon,'.',layer,'.',time,sep=""))
+num.ind <- with(nums,paste(species,'.',agecl,'.',polygon,'.',layer,'.',time,sep=""))
+pick <- match(num.ind,resn.ind)
+SRN <- resn$atoutput[pick] + structn$atoutput[pick]
+
+# get weight-length parameters
+li_a_use <- biolprm$wl[match(fgs$Code[match(nums$species,fgs$Name)],biolprm$wl[, 1]), 2]
+li_b_use <- biolprm$wl[match(fgs$Code[match(nums$species,fgs$Name)],biolprm$wl[, 1]), 3]
+
+#calc mean length and weight at age
+mulen$atoutput <- ((biolprm$kgw2d*biolprm$redfieldcn*SRN)/(1000*li_a_use))^(1/li_b_use)
+muweight$atoutput <- li_a_use*mulen$atoutput^li_b_use
+
+upper.bins <- 1:maxbin
+lower.bins <- c(0,upper.bins[-length(upper.bins)])
+lenfreq <- NULL
+
+# well, match does return NA for capelin, and this is what breaks the comparison using 
+# the max function (returns NA). adding na.rm=TRUE to line 86
+if(maxbin<max(mulen$atoutput)){
+  print("Warning: maximum bin size is smaller than the longest fish in the sample. Fish above the maximum bin size will be removed from length compositions.")
+}
+
+groups <- unique(as.factor(structn$species)) 
+CVlenage = 0.1
+# but will NAs break the following?
+for (irow in 1:nrow(mulen))
+#for (irow in 1:500)
+  {
+  group <- nums$species[irow]
+  igroup <- which(groups==group)
+  box <- nums$polygon[irow]
+  layer <- nums$layer[irow]
+  time <- nums$time[irow]
+  age <- nums$agecl[irow]
+
+  sigma = sqrt(log((CVlenage^2)+1))
+  muuse <- log(mulen$atoutput[irow]) - 0.5*(sigma^2)
+  CumFracperbin <- plnorm(upper.bins,muuse,sigma)
+  Fracperbin <- c(CumFracperbin[1],diff(CumFracperbin))
+  natlength = Fracperbin*nums$atoutput[irow]
+  results <- cbind(mulen[irow,],lower.bins,upper.bins, row.names = NULL) # add row.names = NULL?
+  results$atoutput <- round(natlength,0)
+  #results <- results[results$atoutput>0,]
+  lenfreq <- rbind(lenfreq,results) # add row.names = NULL?
+}
+
+remove.zeroes = TRUE
+#get rid of zero rows.
+if (remove.zeroes) lenfreq <- lenfreq[lenfreq$atoutput>0,]
+
+#output. natlength now has additional columns because of need to store length bin info
+lenout <- NULL
+lenout$mulen <- mulen
+lenout$muweight <- muweight
+lenout$natlength <- lenfreq
+
+# ok, works but adds NA for Capelin age classes where no numbers were sampled.
+# could crank up sample size or just live with this.
+
+```
+
+Try om_comps step again:
+
+```{r comps-redo}
+
+NOBAom_redcodcap_comp <- om_comps(usersurvey = here("config/3sppsurvey.R"), 
+                           userfishery = here("config/3sppfishery.R"),
+                           omlist_ss = NOBAom_redcodcap, 
+                           n_reps = 1, 
+                           save = TRUE)
+```

### 2020-06-03 config/3sppsurvey.R

@@ -1,0 +1,56 @@
+# Default survey configuration here is a census
+# Need to define survey season, area, efficiency, selectivity
+
+#Atlantis model timestep corresponding to the true output--now from census_spec.R
+timestep <- stepperyr #5
+
+#Which atlantis timestep does the survey run in?--now from census_spec.R
+# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
+# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
+
+survey_sample_time <- midptyr #2; defined in omdimensions.R
+
+#The last timestep to sample
+total_sample <- noutsteps-1 #495
+
+#Vector of indices of survey times to pull
+survey_sample_full <- seq(survey_sample_time,
+                          total_sample, by=timestep)
+
+survtime <- survey_sample_full
+
+# survey area
+# should return all model areas
+survboxes <- allboxes
+
+# survey efficiency (q)
+# should return a perfectly efficient survey 
+surveffic <- data.frame(species=survspp,
+                     efficiency=rep(1.0,length(survspp)))
+
+# survey selectivity (agecl based)
+# this is by age class, need to change to use with ANNAGEBIO output
+#survselex <- data.frame(species=rep(survspp, each=n_age_classes),
+#                     agecl=rep(c(1:n_age_classes),length(survspp)),
+#                     selex=rep(1.0,length(survspp)*n_age_classes))
+
+# for annage output
+survselex <- data.frame(species=rep(survspp, n_annages), #  
+                        agecl=unlist(sapply(n_annages,seq)),
+                        selex=rep(1.0,sum(n_annages)))
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+
+# survey index cv needed for sample_survey_xxx
+# cv = 0.1
+surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp)))
+
+# length at age cv for input into calc_age2length function
+# function designed to take one cv for all species, need to change to pass it a vector
+lenage_cv <- 0.1
+
+# max size bin for length estimation, function defaults to 150 cm if not supplied
+maxbin <- 200
+

### 2020-06-04 InterpolateWeightAtAge.Rmd
adding diagnostics for species

@@
 wtage <- readRDS(file.path(d.name, paste0(scenario.name, "survObsWtAtAge.rds")))
-wtage <- wtage[[1]]
+wtage <- wtage[[1]] %>%
+  filter(species == "North_atl_cod")
 
@@
 
-```{r interpolate-wtage}
+```{r interpolate-wtage-dev}
 # a function to go into atlantisom, sketched here, reuse calc_stage2age as possible
@@
 annages <- NOBAom_cod$truenumsage_ss %>%
+  filter(species == "North_atl_cod") %>%
   group_by(species, time, agecl) %>%
@@
 
-```{r omdimensions-bugfix}
+```{r omdimensions-bugfix, eval=FALSE}
 
@@
 
+New omdimensions file:
+```{r, code = readLines("./config/omdimenstions.R"), eval=F}
+```
+
 New config files needed for multispecies pulls:
@@
 Fix multispecies bug in om_index function:
-```{r omindex-bugfix}
+```{r omindex-bugfix, eval=FALSE}
 
@@
 
-Try om_comps step again:
+Try om_comps step again (now works by allowing NA in max comparison of `calc_age2length()`:
 
-```{r comps-redo}
+```{r comps-redo, message=FALSE}
 
@@
 ```
+
+What do the interpolated weight at ages look like?
+Diagnostic plots from function look the same as above for cod (good), blank for Capelin (good?
+didnt need extrapolation but may have other problems), and bad for Redfish between true ages 2-15; lots of overestimation.
+
+```{r vis-interpolate-1, warning=FALSE}
+
+# plot portion, just redfish, capelin is blank and cod looks the same
+library(ggforce)
+NOBAom_redcodcap_comp$survObsFullWtAtAge[[1]][[2]] + facet_wrap_paginate(~species, ncol=1, nrow = 1, page=3, scales = "free")
+
+```
+
+Compare ObsWtAtAge with ObsFullWtAtAge outputs by species:
+
+```{r vis-interpolate-2}
+
+# dataframe portion
+intwtCod <- NOBAom_redcodcap_comp$survObsFullWtAtAge[[1]][[1]] %>%
+  filter(species == "North_atl_cod")
+
+intwtRed <- NOBAom_redcodcap_comp$survObsFullWtAtAge[[1]][[1]] %>%
+  filter(species == "Redfish")
+
+intwtCap <- NOBAom_redcodcap_comp$survObsFullWtAtAge[[1]][[1]] %>%
+  filter(species == "Capelin") #empty
+
+wtageCod <- NOBAom_redcodcap_comp$survObsWtAtAge[[1]] %>%
+  filter(species == "North_atl_cod")
+
+wtageRed <- NOBAom_redcodcap_comp$survObsWtAtAge[[1]] %>%
+  filter(species == "Redfish")
+
+wtageCap <- NOBAom_redcodcap_comp$survObsWtAtAge[[1]] %>%
+  filter(species == "Capelin")
+
+# handmade plots:
+
+wtagecl <- wtageCap
+
+  wtagep <- ggplot(wtagecl, aes(agecl, atoutput)) +
+    geom_point() +
+    geom_line(aes(colour = factor(time))) + 
+    scale_x_continuous(minor_breaks = c(0:max(agecl))) +
+    theme_tufte() +
+    theme(legend.position = "none",
+          panel.grid.minor.x = element_line(colour = "grey50"),
+          panel.grid.major.x = element_line(colour = "grey50"))
+
+```

### 2020-06-05 InterpolateWeightAtAge.Rmd
three species all but the oldest ages

@@
 
-```{r interp-wtage-test1}
+```{r interp-wtage-test1, eval=FALSE}
 
@@
 New omdimensions file:
-```{r, code = readLines("./config/omdimenstions.R"), eval=F}
+```{r, code = readLines("./config/omdimensions.R"), eval=F}
 ```
@@
 
-```{r indcomp-test}
+```{r indcomp-test, eval=FALSE}
 
@@
 
-```{r comps-redo, message=FALSE}
+```{r comps-redo, message=FALSE, eval=FALSE}
 
@@
 
+```{r readin-savedoutput}
+
+survObsFullWtAtAge <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullWtAtAge.rds")))
+
+survObsWtAtAge <- readRDS(file.path(d.name, paste0(scenario.name, "survObsWtAtAge.rds")))
+
+NOBAom_redcodcap <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+
+```
+
 ```{r vis-interpolate-1, warning=FALSE}
@@
 library(ggforce)
-NOBAom_redcodcap_comp$survObsFullWtAtAge[[1]][[2]] + facet_wrap_paginate(~species, ncol=1, nrow = 1, page=3, scales = "free")
+survObsFullWtAtAge[[1]][[2]] + facet_wrap_paginate(~species, ncol=1, nrow = 1, page=3, scales = "free")
 
@@
 # dataframe portion
-intwtCod <- NOBAom_redcodcap_comp$survObsFullWtAtAge[[1]][[1]] %>%
+intwtCod <- survObsFullWtAtAge[[1]][[1]] %>%
   filter(species == "North_atl_cod")
 
-intwtRed <- NOBAom_redcodcap_comp$survObsFullWtAtAge[[1]][[1]] %>%
+intwtRed <- survObsFullWtAtAge[[1]][[1]] %>%
   filter(species == "Redfish")
 
-intwtCap <- NOBAom_redcodcap_comp$survObsFullWtAtAge[[1]][[1]] %>%
-  filter(species == "Capelin") #empty
+intwtCap <- survObsFullWtAtAge[[1]][[1]] %>%
+  filter(species == "Capelin") #empty, no need for interpolation, but make intelligent
 
-wtageCod <- NOBAom_redcodcap_comp$survObsWtAtAge[[1]] %>%
+wtageCod <- survObsWtAtAge[[1]] %>%
   filter(species == "North_atl_cod")
 
-wtageRed <- NOBAom_redcodcap_comp$survObsWtAtAge[[1]] %>%
+wtageRed <- survObsWtAtAge[[1]] %>%
   filter(species == "Redfish")
 
-wtageCap <- NOBAom_redcodcap_comp$survObsWtAtAge[[1]] %>%
+wtageCap <- survObsWtAtAge[[1]] %>%
   filter(species == "Capelin")
@@
     geom_line(aes(colour = factor(time))) + 
-    scale_x_continuous(minor_breaks = c(0:max(agecl))) +
+    scale_x_continuous(minor_breaks = c(0:max(wtagecl$agecl))) +
     theme_tufte() +
@@
 
+wtagep
+
 ```
+
+Capelin is fine with WtAtAge output because it has age classes = true ages!
+Maybe just warn that a species with fgs$NumAgeClassSize = 1 will have NA output of interpolation functions, which will run if there is output for that species in the ANNAGEBIO.nc and ANNAGECATCH.nc files.
+
+Redfish appear to have *decreasing* average weight at age as they get older:
+
+```{r redfish-1}
+
+wtagecl <- wtageRed
+
+  wtagep <- ggplot(wtagecl, aes(agecl, atoutput)) +
+    geom_point() +
+    geom_line(aes(colour = factor(time))) + 
+    scale_x_continuous(minor_breaks = c(0:max(wtagecl$agecl))) +
+    theme_tufte() +
+    theme(legend.position = "none",
+          panel.grid.minor.x = element_line(colour = "grey50"),
+          panel.grid.major.x = element_line(colour = "grey50"))
+
+wtagep
+
+```
+
+Diagnose Redfish problem (it was not negative increments, but inadequate lagging and leading):
+
+```{r redfishdebug, message=FALSE}
+
+annageRed <- NOBAom_redcodcap$truenumsage_ss %>%
+  filter(species == "Redfish")
+
+annageCod <- NOBAom_redcodcap$truenumsage_ss %>%
+  filter(species == "North_atl_cod")
+
+annageCap <- NOBAom_redcodcap$truenumsage_ss %>%
+  filter(species == "Capelin")
+
+annages <- annageRed
+wtagecl <- wtageRed
+fgs <- fgs
+
+  annages <- annages %>%
+    group_by(species, time, agecl) %>%
+    rename(trueage = agecl) %>%
+    summarise(truenatage = sum(atoutput))
+
+  # assuming fgs is data object already been read in by load_fgs elsewhere
+  # Figure out the groups that have multiple ages classes in each stage (or
+  # cohort)
+  multiple_ages <- fgs[fgs$NumAgeClassSize>1, c(1,4,10)]
+
+  # match species in wtagecl_data to those species in the fgs file
+  sppwt <- levels(wtagecl$species)
+  multiple_ages <- multiple_ages[multiple_ages$Name %in% sppwt, ]
+
+  # make a dataframe the right dimensions for all age classes
+  # has species, time from wtage and trueage, truenatage from annages
+  # add agecl column
+  # finds the average age of each agecl each timestep based on truenatage
+  wtage_out <- annages %>%
+    dplyr::semi_join(wtagecl, by = c("species", "time")) %>%
+    dplyr::inner_join(multiple_ages, by = c("species" = "Name")) %>%
+    dplyr::mutate(agecl = as.integer(ceiling(trueage/NumAgeClassSize))) %>%
+    dplyr::group_by(species, time, agecl) %>%
+    dplyr::mutate(avgage = weighted.mean(trueage, truenatage))
+
+  wtage_avgage <- wtage_out %>%
+    select(species, time, agecl, avgage) %>%
+    distinct()
+
+  # find weight increment for a timestep
+  # muweight for agecl+1 - muweight for agecl
+  wtagecl_inc <- wtagecl %>%
+    dplyr::arrange(species, time, agecl) %>%
+    #dplyr::left_join(multiple_ages, by = c("species" = "Name")) %>%
+    dplyr::group_by(species, time) %>%
+    dplyr::mutate(increment = case_when(
+      agecl==1 ~ atoutput,
+      agecl>1 ~ atoutput - dplyr::lag(atoutput)
+    )) %>%
+    dplyr::left_join(wtage_avgage) %>%
+    dplyr::mutate(avgageinc = case_when(
+      agecl==1 ~ avgage,
+      agecl>1 ~ avgage - dplyr::lag(avgage)
+    ))
+
+  # complex, but works for 2 ages/agecl, need to test more, need to fix oldest
+  # for more than 2 ages/agecl need to have a bigger lead/lag
+  # achieved by faking a scalar for lead/lag n using max() function
+  # should be ceiling I think for odd NumAgeClassSize, but not sure?
+  wtage_out <- wtage_out %>%
+    dplyr::select(species, time, trueage, NumAgeClassSize, agecl, avgage) %>%
+    dplyr::left_join(wtagecl_inc) %>%
+    dplyr::group_by(species, time) %>%
+    dplyr::mutate(wtIntage = case_when(
+      (agecl==1 & trueage<avgage) ~ (1-(avgage-trueage)/avgageinc)*increment,
+      (agecl>1 & trueage<avgage) ~
+        (1-(avgage-trueage)/avgageinc)*increment +
+        lag(atoutput, ceiling(max(NumAgeClassSize)/2)),
+      (trueage>avgage) ~
+        ((trueage-avgage)/lead(avgageinc, ceiling(max(NumAgeClassSize)/2)))*lead(increment, ceiling(max(NumAgeClassSize)/2)) + atoutput
+    ))
+
+  # fix oldest age
+
+  # diagnostic plot--are interpolations where we expect them?
+  wtageclann <- ggplot(wtagecl_inc, aes(avgage, atoutput)) +
+    geom_point() +
+    geom_line(aes(colour = factor(time))) +
+    scale_x_continuous(minor_breaks = c(0:max(wtage_out$trueage))) +
+    theme_tufte() +
+    theme(legend.position = "none",
+          panel.grid.minor.x = element_line(colour = "grey50"),
+          panel.grid.major.x = element_line(colour = "grey50"))
+
+  diag.p <- wtageclann + geom_point(data = wtage_out,
+                          mapping = aes(trueage, wtIntage))
+
+  # #clean up to have only standard columns
+  # finaldata <- data.frame("species" = wtage_out$species,
+  #                         "agecl" = wtage_out$trueage, "polygon" = NA, "layer" = NA,
+  #                         "time" = wtage_out$time, "atoutput" = wtage_out$wtIntage)
+
+  diag.p
+
+```
+
+Try the updated function again, any bugs with multiple species? OK so far.
+
+```{r test-correctedinterp, warning=FALSE, message=FALSE}
+
+omlist_ss <- NOBAom_redcodcap
+
+interp_survWtAtAge <- calc_avgwtstage2age(wtagecl = survObsWtAtAge[[1]],
+                                                     annages = omlist_ss$truenumsage_ss,
+                                                     fgs = omlist_ss$funct.group_ss)
+
+interp_survWtAtAge[[2]] + facet_wrap_paginate(~species, ncol=1, nrow = 1, page=1, scales = "free")
+
+interp_survWtAtAge[[2]] + facet_wrap_paginate(~species, ncol=1, nrow = 1, page=2, scales = "free")
+
+interp_survWtAtAge[[2]] + facet_wrap_paginate(~species, ncol=1, nrow = 1, page=3, scales = "free")
+
+```
+
+Last issue is extrapolating the oldest wt at age. Linear from the last one?

### 2020-06-05 InterpolateWeightAtAge.Rmd
now works for oldest ages

@@
 ---
-title: "Interpolate the weight at agecl output to annual age"
+title: "Interpolate weight at agecl output to annual age"
 author: "Sarah Gaichas"
@@
 library(ggthemes)
+library(ggforce)
 
@@
 # plot portion, just redfish, capelin is blank and cod looks the same
-library(ggforce)
+#library(ggforce)
 survObsFullWtAtAge[[1]][[2]] + facet_wrap_paginate(~species, ncol=1, nrow = 1, page=3, scales = "free")
@@
 wtagecl <- wtageRed
-fgs <- fgs
+fgs <- NOBAom_redcodcap$funct.group_ss
 
@@
   # should be ceiling I think for odd NumAgeClassSize, but not sure?
+  # for oldest age(s), perhaps default=last(x) in lead statement
   wtage_out <- wtage_out %>%
@@
         (1-(avgage-trueage)/avgageinc)*increment +
-        lag(atoutput, ceiling(max(NumAgeClassSize)/2)),
+        dplyr::lag(atoutput, ceiling(max(NumAgeClassSize)/2)),
       (trueage>avgage) ~
-        ((trueage-avgage)/lead(avgageinc, ceiling(max(NumAgeClassSize)/2)))*lead(increment, ceiling(max(NumAgeClassSize)/2)) + atoutput
+        (trueage-avgage)/dplyr::lead(avgageinc, ceiling(max(NumAgeClassSize)/2), default=last(avgageinc))*dplyr::lead(increment, ceiling(max(NumAgeClassSize)/2), default=last(increment)) + atoutput
     ))
-
-  # fix oldest age
 
@@
 
-Try the updated function again, any bugs with multiple species? OK so far.
+Try the updated function again, any bugs with multiple species? Is it getting the weight at max age for both? OK so far.
 
@@
 
-Last issue is extrapolating the oldest wt at age. Linear from the last one?
+Extrapolating the oldest wt at age is now linear from the last agecl weight at age using default=last() in the dplyr::lead() calls.

### 2020-06-05 NOBAcod.Rmd
update with new interpolated wtage

@@
 
-library(ggthemes)
 omlist_ss <- NOBAom_cod
@@
 
-#### Survey weight at age (annual ages) pending new function
+#### Survey iterpolated weight at age (annual ages) 
 ```{r plot-annwtage}
+
+wtage <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullWtAtAge.rds")))
+wtage <- wtage[[1]] #this still has second list component, diagnostic plot
+wtage <- wtage[[1]]
+
+wageplot <- ggplot(wtage, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  xlab("model timestep (5 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
 
@@
   Catageplot + facet_wrap(~time, ncol=6, scales="free_y")
+```
+
+#### Fishery iterpolated weight at age (annual ages) 
+```{r plot-fishannwtage}
+
+wtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullWtAtAge.rds")))
+wtage <- wtage[[1]] #this still has second list component, diagnostic plot
+wtage <- wtage[[1]]
+
+wageplot <- ggplot(wtage, aes(time, atoutput)) +
+  geom_line(aes(colour = factor(agecl))) +
+  theme_tufte() +
+  theme(legend.position = "bottom") +
+  xlab("model timestep (5 per year)") +
+  ylab("average individual weight (g)") +
+  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
+
+wageplot + facet_wrap(c("species"), scales="free_y")
+
 ```

### 2020-06-05 docs/index.Rmd
moderately successful wt at age interpolation
correct spelling
update index
update with new interpolated wtage

@@
 
-### Apply wrapper to Norwegian Barents cod, March 2020
+### Full age structured dataset for Norwegian Barents cod, March-June 2020
 
-<a href="https://sgaichas.github.io/poseidon-dev/NOBAcod.html" target="_blank">Apply and extend wrappers to get full age structure, updated June 1, 2020</a>
+<a href="https://sgaichas.github.io/poseidon-dev/NOBAcod.html" target="_blank">Apply and extend wrappers to get full age structure, updated June 5, 2020</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/InterpolateWeightAtAge.html" target="_blank">Interpolate to get average weight at true age, June 5, 2020</a>
+
+Also bug fixed config files for multispecies dataset. Interpolate wtage includes data generation for NOBA cod (max age 20), redfish (max age 40), and capelin (max age 5).
 

### 2020-06-05 config/codsurvey.R

@@ -35,9 +35,9 @@ surveffic <- data.frame(species=survspp,
 #                     selex=rep(1.0,length(survspp)*n_age_classes))
 
 # for annage output
-survselex <- data.frame(species=rep(survspp, each=n_annages),
-                        agecl=rep(c(1:n_annages),length(survspp)),
-                        selex=rep(1.0,length(survspp)*n_annages))
+survselex <- data.frame(species=rep(survspp, n_annages), #  
+                        agecl=unlist(sapply(n_annages,seq)),
+                        selex=rep(1.0,sum(n_annages)))
 
 # effective sample size needed for sample_fish
 # this effective N is high but not equal to total for numerous groups

### 2020-06-09 atlantisom_landingpage.Rmd
draft atlantisom landing page

---
title: "atlantisom: simulated ecosystem datasets for assessment model performance testing"
author: "Sarah Gaichas, Christine Stawitz, Kelli Johnson, Alexander Keth, Allan Hicks, Sean Lucey, Emma Hodgson, and Gavin Fay"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}

htmltools::img(src = knitr::image_uri(here::here("docs/images/atlantis-logo2.jpg")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; height:175px; padding-right:20px;')

knitr::opts_chunk$set(echo = F,
                      warning = F,
                      message = F)
#Plotting and data libraries
library(tidyverse)
library(here)
library(ggrepel)
library(stringr)

library(atlantisom)

data.dir <- here::here("data")

```

## Overview

The purpose of `atlantisom` is to use existing [Atlantis](https://github.com/Atlantis-Ecosystem-Model/Atlantis_example_and_instructions) ecosystem model output to generate input datasets for a variety of simpler population models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics. Atlantis models simulate a wide range of physical and ecological processes, include a full food web, and can be run using different climate forcing, fishing, and other scenarios.

Users of `atlantisom` specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes to simulate survey and fishery "data" from stored Atlantis scenario output. `atlantisom` outputs internally consistent multispecies and ecosystem datasets with known observation error characteristics for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against “true” Atlantis outputs.

Development of `atlantisom` began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA. Substantial progress since then was made possible by a NOAA NMFS International Fellowship and the Institute of Marine Research in Bergen, Norway.

## Get `atlantisom`  https://github.com/r4atlantis/atlantisom

```{r getit, echo=TRUE, eval=FALSE}
# install.packages("devtools")
devtools::install_github("r4atlantis/atlantisom")
```

## Background

![A wide range of models exist to address potential fishery management questions](docs/images/ecomods.png)

### How do we know our models are right?

 * Fits to historical data (hindcast)

 * Influence of data over time (retrospective diagnostics)

 * Keep as simple and focused as possible

 * Simulation testing

But what if data are noisy? What if we need to model complex interactions? What if conditions change over time?

### End-to-end ecosystem operating models as dataset generators

Atlantis is a spatially resolved mechanistic end-to-end ecosystem modeling framework: [Fulton et al. 2011](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1467-2979.2011.00412.x), [Fulton and Smith 2004](https://www.ajol.info/index.php/ajms/article/view/33182). Atlantis models have been implemented for regional ecosystems around the world, including:

**Norwegian-Barents Sea**

[Hansen et al. 2016](https://www.imr.no/filarkiv/2016/04/fh-2-2016_noba_atlantis_model_til_web.pdf/nn-no), 
[2018](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0210419)

![NoBa model areas](docs/images/journal.pone.0210419.g001.png)

**California Current**

[Marshall et al. 2017](https://onlinelibrary.wiley.com/doi/full/10.1111/gcb.13594), 
[Kaplan et al. 2017](https://www.sciencedirect.com/science/article/pii/S0304380016308262?via%3Dihub)

![CCA model areas](docs/images/CCAspatial.png)

Atlantis models can incoroporate physical drivers from global change projections and simulate complex biological responses throughout the ecosystem:  [Hodgson et al. 2018](https://www.sciencedirect.com/science/article/pii/S0304380018301856?via%3Dihub), 
[Olsen et al. 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00064/full)

## Why use Atlantis?

-  Mechanistic processes create internally consistent "truth"
-  Include cumulative effects of multiple processes:
    -  Climate drivers
    -  Species interactions 
    -  Spatial and seasonal variability
    -  Fisheries
    -  Oil spills, red tide, anything else Atlantis can do
-  Implemented for many ecosystems worldwide

## Why generate datasets instead of simulating within Atlantis?

-  Not all analyses need computationally expensive model interaction
-  Faster! 
-  Test many models or model configurations with the same dataset
-  Many dataset realizations from same "truth"; compare:
    -  Different observation error and bias
    -  Changing temporal and spatial survey coverage 
    -  Improved or degraded fishery observations

### Illustration: Climate impacts in the operating model 
![Changing sardine habitat](docs/images/sardinetempIsaac.png)

Hodgson, E. E., Kaplan, I. C., Marshall, K. N., Leonard, J., Essington, T. E., Busch, D. S., Fulton, E. A., et al. 2018. Consequences of spatially variable ocean acidification in the California Current: Lower pH drives strongest declines in benthic species in southern regions while greatest economic impacts occur in northern regions. Ecological Modelling, 383: 106–117.

Marshall, K. N., Kaplan, I. C., Hodgson, E. E., Hermann, A., Busch, D. S., McElhany, P., Essington, T. E., et al. 2017. Risks of ocean acidification in the California Current food web and fisheries: ecosystem model projections. Global Change Biology, 23: 1525–1539.

### Illustration: Climate + cumulative impacts in the operating model 
![Habitat changes lead to changes in both populations and community interactions over time](docs/images/guildbioIsaac.png)

## How does this affect the "data" for stock assessments?

The atlantisom user must specify uncertainty in assessment "data":

1. Survey specification:

    1. timing and spatial coverage?

    1. which species are captured?

    1. species-specific survey efficiency ("q")?

    1. selectivity at age for each species?

1. Survey uncertainty:

    1. additional observation error (survey cv for index)?

    1. effective sample size for biological samples?

1. Fishery uncertainty:

    1. additional observation error (catch cv for total)?

    1. catch sampled for length/age in all areas?

    1. effective sample size for biological samples?

## Make Atlantis output into assessment model input

Example atlantisom workflows:

1. Get true biomass, abundance, age composition, length composition, weight at age, fishery catch, fishery catch at age, fishery length composition, and fishery weight age age for a "sardine-like species": https://sgaichas.github.io/poseidon-dev/FullSardineTruthEx.html

1. Format these outputs and get other life history parameters for input into a stock assessment model ([Stock Synthesis](https://www.sciencedirect.com/science/article/abs/pii/S0165783612003293), using [`r4ss`](https://github.com/r4ss)):
https://sgaichas.github.io/poseidon-dev/CreateStockSynthesis.html

1. Get true and observed input data, format inputs, and run the assessment model: https://sgaichas.github.io/poseidon-dev/SardinesHakeatlantisom2SStest.html

1. In progress: compare assessment results with truth:
https://sgaichas.github.io/poseidon-dev/SkillAssessInit.html

1. Simplified dataset extraction with wrapper functions:
https://sgaichas.github.io/poseidon-dev/NOBAcod.html

## What can atlantisom do so far?

Survey census test NOBA 
```{r ex-output1, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/NOBAcensus.png")
```

True length composition NOBA
```{r ex-output3, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/NOBAherringtruelf.png")
```

Standard survey test CCA
```{r ex-output2, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/CCsurveyex.png")
```

Survey length composition CCA
```{r ex-output4, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/CCV3herringlfcomp.png")
```

## A "sardine" assessment

Need: assessment model data inputs and life history parameters

*(model based on actual Sardine assessment in Stock Synthesis 3)*

Data:

  * survey biomass index 
  * survey length composition 
  * survey age composition (conditional catch at age)
  * fishery catch (tons) 
  * fishery length composition 
  * fishery age composition

Parameters:

  * natural mortality (from total mortality)
  * growth curve (from survey length at age)
  * maturity at age (true)
  * unfished recruitment and steepness (true)
  * weight-length curve (true)

## A "sardine" assessment: setup

  * California Current Atlantis run with and without climate signal
  * Input data generated (e.g. sardine survey, below in green)
  * Parameters derived; simpler recruitment distribution

```{r sardsurv, echo = F, warning = F, message=F, out.width = "90%", fig.align = "left"}
knitr::include_graphics("docs/images/CCsurvObsBiom.png")
```

## A "sardine" assessment: fits to data

![survey index fit](docs/images/index2_cpuefit_Survey.png)

![length fit](docs/images/comp_lenfit_flt2mkt0_page1.png)

![catch at age fit 1](docs/images/comp_agefit_flt1mkt0_page1.png)
![catch at age fit 2](docs/images/comp_agefit_flt1mkt0_page2.png)

## A "sardine" assessment: skill? (proof of concept)

Biomass
```{r skillB, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/trueBss3B.png")
```

Fishing mortality
```{r skillF, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/trueFss3F.png")
```

Recruitment
```{r skillR, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/trueRss3R.png")
```

Key: <span style="color:teal">True</span> <span style="color:red">SS3 estimate</span>

## External Resources

* [Atlantis Model Documentation](https://github.com/Atlantis-Ecosystem-Model/Atlantis_example_and_instructions)

* [atlantisom R package](https://github.com/r4atlantis/atlantisom)

* [Testing atlantisom](https://github.com/sgaichas/poseidon-dev)

### 2020-06-09 atlantisom_landingpage.Rmd
streamline

@@
 
-![A wide range of models exist to address potential fishery management questions](docs/images/ecomods.png)
-
-### How do we know our models are right?
-
- * Fits to historical data (hindcast)
-
- * Influence of data over time (retrospective diagnostics)
-
- * Keep as simple and focused as possible
-
- * Simulation testing
-
-But what if data are noisy? What if we need to model complex interactions? What if conditions change over time?
+Assessing the ability of models to predict key processes is essential if the models are to be used in decision-making dependent on those processes.
+Skill assessment compares model predictions with observations; good agreement between predictions and observations indicates high skill (see, e.g.
+[Stow et al 2009](https://www.sciencedirect.com/science/article/abs/pii/S0924796308001103)).
+However, in many natural resrouce management contexts, the observations themselves are highly uncertain, so assessing model skill can be difficult.
+In these cases, creating an observational dataset with a simulation model can provide alternative information for model skill assessment.
 
@@
 
-### Illustration: Climate impacts in the operating model 
-![Changing sardine habitat](docs/images/sardinetempIsaac.png)
+## How does `atlantisom` make "data" for stock assessments?
 
-Hodgson, E.
-E., Kaplan, I.
-C., Marshall, K.
-N., Leonard, J., Essington, T.
-E., Busch, D.
-S., Fulton, E.
-A., et al.
-2018.
-Consequences of spatially variable ocean acidification in the California Current: Lower pH drives strongest declines in benthic species in southern regions while greatest economic impacts occur in northern regions.
-Ecological Modelling, 383: 106–117.
-
-Marshall, K.
-N., Kaplan, I.
-C., Hodgson, E.
-E., Hermann, A., Busch, D.
-S., McElhany, P., Essington, T.
-E., et al.
-2017.
-Risks of ocean acidification in the California Current food web and fisheries: ecosystem model projections.
-Global Change Biology, 23: 1525–1539.
-
-### Illustration: Climate + cumulative impacts in the operating model 
-![Habitat changes lead to changes in both populations and community interactions over time](docs/images/guildbioIsaac.png)
-
-## How does this affect the "data" for stock assessments?
-
-The atlantisom user must specify uncertainty in assessment "data":
+The `atlantisom` user must specify uncertainty in assessment "data":
 
@@
 
+Using these specifications, `atlantisom` generates survey data from the true biomass outputs of the Atlantis model, and generates fishery data from the true fishery outputs of the Atlantis model.
+The simulated data can then be used as inputs to a variety of single species or multispecies assessment models.
+
 ## Make Atlantis output into assessment model input
@@
 
-## What can atlantisom do so far?
+## Visualizing `atlantisom` outputs
 
@@
 
-## A "sardine" assessment
+## Example: a "sardine" assessment
 

### 2020-06-10 atlantisom_landingpage.Rmd
add detailed data flow section

@@
 ```
+
+## Detailed data flow
+
+Have these Atlantis input files in a directory:
+
+- The .bgm file defining model geometry
+- The initial conditions/biomass pools .nc 
+- The functional groups .csv
+- The fishery groups .csv
+- The biology .prm
+- The run prm (.xml version)
+
+At present,`atlantisom` uses the following Atlantis output files, where [...] is a model-specific prefix. They should be in the same directory as the input files above:
+
+- [...]BiomIndx.txt
+- [...]Catch.txt
+- [...]CatchPerFishery.txt
+- [...]DietCheck.txt
+- [...]YOY.txt
+- [...].nc
+- [...]CATCH.nc
+- [...]PROD.nc
+- [...]ANNAGEBIO.nc (if available)
+- [...]ANNAGECATCH.nc (if available)
+
+Get the truth:
+![](docs/images/atlantisomDataFlow_truth-link.png)
+
+Generate the data:
+![](docs/images/atlantisomDataFlow_data.png)
 
@@
 
-## External Resources
+## More information
 

### 2020-06-12 docs/index.Rmd
add landing page to index
revisited calc_Z, decided not to change it

@@
 Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.
+
+<a href="https://sgaichas.github.io/poseidon-dev/atlantisom_landingpage.html" target="_blank">Draft atlantisom landing page for review, June 9, 2020</a>
 
@@
 
+Further experiments with M, F, and Z across models and species confirm that mortality needs a lot more work:
+
+<a href="https://sgaichas.github.io/poseidon-dev/TestCalcFandM.html" target="_blank">Why is mortality all over the map? June 12, 2020</a>
+
 ### Rabbit holes that turned out ok

### 2020-06-15 Visualize_atlantisom.Rmd
hide r4ss text output
update plot for MSE paper

@@
 
+So I wound up doing this by hand using online flowchart software Lucidchart (see original [here](https://app.lucidchart.com/documents/edit/7af8dad3-0bc6-4000-9668-9bc1add809ba/0_0)):
+
+Get the truth:
+![](docs/images/atlantisomDataFlow_truth-link.png)
+
+Generate the data:
+![](docs/images/atlantisomDataFlow_data.png)
+
 ## Visualize results using wrappers
@@
 
-```{r ggplotSS}
+```{r ggplotSS, results="hide"}
 
@@
                          shape = c(16, NA)))) + 
+  xlim(0,80) +
   theme_tufte() +
@@
                          shape = c(NA, 16)))) + 
+  xlim(0,80) +
   theme_tufte() +

### 2020-06-15 docs/index.Rmd
update plot for MSE paper

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/Visualize_atlantisom.html" target="_blank">Different packages to visualize `atlantisom` functions, updated May 21, 2020</a>
+<a href="https://sgaichas.github.io/poseidon-dev/Visualize_atlantisom.html" target="_blank">Different packages to visualize `atlantisom` functions, updated June 15, 2020</a>
 

### 2020-06-18 NOBAmsdiets.Rmd
start working with multispecies surveys and diets

--- 2020-06-05 NOBAcod.Rmd
@@
 ---
-title: "Generate data for NOBA cod ensemble modeling"
+title: "Generate data for multispecies modeling, including diet!"
 author: "Sarah Gaichas"
@@
 library(here)
-```
-
-## Standard atlantisom output test (age class not age)
-
-Let's use the wrapper functions to get initial stock assessment model inputs.
-
-Then we need to adjust the base atlantisom functions for true age classes, and develop options to call those within the wrappers.
-
-We need config files for the NOBA model, our surveys and fisheries. Surveys and fisheries are based on the ones we used for sardine in the California Current as a test.
-
-**Special note:** *Atlantis output files formerly had the prefix "output" and `atlantisom` was coded to expect this.
-NOBA output files (and newer Atlantis models) do not universally use the "output" prefix naming convention.
-Annage branch `atlantisom` functions have been updated for user-specified output filenames.*
-
-**Special note 2:** *NOBA's nordic_groups_v04.csv file has an inFishery column, rather than the `atlantisom` expected InFishery column.
-I've changed the column name in the input file, but need to add robustness to atlantisom code to remove case sensitivity.*
-
-### Config files
-
-`NOBA2config.R` looks like this (adjusted from Alfonso's original):
-
-```{r, code = readLines("./config/NOBA2Config.R"), eval=F}
-```
-
-`omdimensions.R` standardizes timesteps, etc.:
-
-```{r, code = readLines("./config/omdimensions.R"), eval=F}
-```
-
-`codsurvey.R` configures the fishery independent survey and looks like this:
-
-```{r, code = readLines("./config/codsurvey.R"), eval=F}
-```
-
-`codfishery.R` configures the fishery dependent data and looks like this:
-
-```{r, code = readLines("./config/codfishery.R"), eval=F}
-```
-
-### Using the wrappers (now updated for annual age)
-
-All the config files go in the config folder and we'll try running this:
-
-```{r full-test, message=FALSE, warning=FALSE}
-
-#library(here) now above
 library(tidyverse)  
@@
 
+```
+
+## Which species?
+
+Let's use the wrapper functions to get initial multispecies assessment model inputs.
+
+We use config files for the NOBA model, and define multispecies surveys and fisheries. We will also test diet functions which are needed as multispecies and ecosystem model inputs.
+
+Our initial species selection includes 11 single species groups from the NOBA model:
+
+```{r spp-table}
+
+fgs <- load_fgs(here("atlantisoutput", "NOBA_March_2020"), "nordic_groups_v04.csv")
+
+lname <- data.frame(Latin = c("*Hippoglossoides platessoides*",
+                              "*Reinhardtius hippoglossoides*",
+                              "*Scomber scombrus*",
+                              "*Melongrammus aeglefinus*",
+                              "*Pollachius virens*",
+                              "*Sebastes mentella*",
+                              "*Micromesistius poutassou*",
+                              "*Clupea harengus*",
+                              "*Gadus morhua*",
+                              "*Boreogadus saida*",
+                              "*Mallotus villosus*"),
+                    Code = c("LRD", "GRH", "MAC", "HAD", "SAI", "RED", 
+                             "BWH", "SSH", "NCO", "PCO", "CAP")
+)
+
+sppsubset <- merge(fgs, lname, all.y = TRUE)
+spptable <- sppsubset %>% 
+  arrange(Index) %>%
+  select(Name, Long.Name, Latin)
+
+knitr::kable(spptable, col.names = c("Model name", "Full name", "Latin name"))
+
+```
+
+These represent a range of life histories which are similar to those on the Northeast US shelf (some are the same species), so they form a reasonable set of species for multispecies model testing.
+
+### Diet interactions
+
+How much do these species interact with each other? Now we test the diet composition functions in `atlantisom`.
+
+Diet composition is loaded with `load_diet_comp()`:
+
+```{r loaddiet}
+
+diettest <- load_diet_comp(here("atlantisoutput", "NOBA_march_2020"), 
+                           "nordic_runresults_01DietCheck.txt", fgs = fgs)
+
+ms_diet <- diettest %>%
+  filter(species %in% sppsubset$Name)
+
+```
+
+## Generating the dataset
+
+Now to get a range of data:
+
+### Config files
+
+`NOBA2config.R` looks like this (adjusted from Alfonso's original):
+
+```{r, code = readLines("./config/NOBA2Config.R"), eval=F}
+```
+
+`omdimensions.R` standardizes timesteps, etc. (this is part of atlantisom and should not need to be changed by the user):
+
+```{r, code = readLines("./config/omdimensions.R"), eval=F}
+```
+
+`mssurvey_spring.R` and `mssurvey_fall.R` configure the fishery independent surveys:
+
+```{r, code = readLines("./config/mssurvey_spring.R"), eval=F}
+```
+
+```{r, code = readLines("./config/mssurvey_fall.R"), eval=F}
+```
+
+`msfishery.R` configures the fishery dependent data:
+
+```{r, code = readLines("./config/msfishery.R"), eval=F}
+```
+
+### Using the wrappers (now updated for annual age)
+
+All the config files go in the config folder and we'll try running this:
+
+```{r full-test, message=FALSE, warning=FALSE}
+
 NOBAom <- om_init(here("config/NOBA2config.R"))
 
-NOBAom_cod <- om_species(c("North_atl_cod"), NOBAom)
-
-NOBAom_cod_ind <- om_index(usersurvey = here("config/codsurvey.R"), 
-                           userfishery = here("config/codfishery.R"),
-                           omlist_ss = NOBAom_cod, 
+NOBAom_ms <- om_species(sppsubset$Name, NOBAom)
+
+NOBAom_ms_ind <- om_index(usersurvey = here("config/mssurvey.R"), 
+                           userfishery = here("config/msfishery.R"),
+                           omlist_ss = NOBAom_ms, 
                            n_reps = 1, 
@@
 
-NOBAom_cod_comp <- om_comps(usersurvey = here("config/codsurvey.R"), 
-                           userfishery = here("config/codfishery.R"),
-                           omlist_ss = NOBAom_cod, 
+NOBAom_ms_comp <- om_comps(usersurvey = here("config/mssurvey.R"), 
+                           userfishery = here("config/msfishery.R"),
+                           omlist_ss = NOBAom_ms, 
                            n_reps = 1, 

### 2020-06-18 config/msppfishery.R

@@ -1,0 +1,42 @@
+# Default fishery configuration here is a census
+# Inherits species from input omlist_ss
+fishspp <- omlist_ss$species_ss 
+
+#Number of years of data to pull
+nyears <- 50
+
+#Atlantis initialization period in years
+burnin <- 30
+
+# fishery output: learned the hard way this can be different from ecosystem outputs
+fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
+
+
+# same time dimensioning parameters as in surveycensus.R
+#Vector of indices of catch in numbers to pull (by timestep to sum)
+fish_sample_full <- c(0:total_sample)  #total_sample defined in sardinesurvey.R
+fish_burnin <- burnin*fstepperyr+1
+fish_nyears <- nyears*fstepperyr
+fish_times <- fish_sample_full[fish_burnin:(fish_burnin+fish_nyears-1)]
+fish_timesteps <- seq(fish_times[fstepperyr], max(fish_times), by=fstepperyr) #last timestep
+#fish_years <- unique(floor(fish_times/fstepperyr)+1) # my original
+fish_years <- unique(floor(fish_times/fstepperyr)) #from Christine's new sardine_config.R
+
+fishtime <- fish_times
+
+
+# fishery sampling area
+# should return all model areas, this assumes you see everything that it caught
+fishboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+fisheffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+
+#this adjusts for subannual fishery output so original effN is for whole year
+fisheffN$effN <- fisheffN$effN/fstepperyr 
+
+# fishery catch cv can be used in sample_survey_biomass
+# perfect observation
+fish_cv <- data.frame(species=survspp, cv=rep(0.01,length(survspp)))
+

### 2020-06-18 config/mssurvey_fall.R

@@ -1,0 +1,58 @@
+# Default survey configuration here has a range of efficiencies and selectivities
+# To emulate a range of species in a single multispecies survey
+# Also now happens in "spring" and "fall"
+# Need to define survey season, area, efficiency, selectivity
+
+#Atlantis model timestep corresponding to the true output--now from census_spec.R
+timestep <- stepperyr #5
+
+#Which atlantis timestep does the survey run in?--now from census_spec.R
+# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
+# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
+
+survey_sample_time <- 3 # fall survey
+
+#The last timestep to sample
+total_sample <- noutsteps-1 #495
+
+#Vector of indices of survey times to pull
+survey_sample_full <- seq(survey_sample_time,
+                          total_sample, by=timestep)
+
+survtime <- survey_sample_full
+
+# survey area
+# should return all model areas
+survboxes <- allboxes
+
+# survey efficiency (q)
+# should return a perfectly efficient survey 
+surveffic <- data.frame(species=survspp,
+                     efficiency=rep(1.0,length(survspp)))
+
+# survey selectivity (agecl based)
+# this is by age class, need to change to use with ANNAGEBIO output
+#survselex <- data.frame(species=rep(survspp, each=n_age_classes),
+#                     agecl=rep(c(1:n_age_classes),length(survspp)),
+#                     selex=rep(1.0,length(survspp)*n_age_classes))
+
+# for annage output
+survselex <- data.frame(species=rep(survspp, n_annages), #  
+                        agecl=unlist(sapply(n_annages,seq)),
+                        selex=rep(1.0,sum(n_annages)))
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+
+# survey index cv needed for sample_survey_xxx
+# cv = 0.1
+surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp)))
+
+# length at age cv for input into calc_age2length function
+# function designed to take one cv for all species, need to change to pass it a vector
+lenage_cv <- 0.1
+
+# max size bin for length estimation, function defaults to 150 cm if not supplied
+maxbin <- 200
+

### 2020-06-18 config/mssurvey_spring.R

@@ -1,0 +1,58 @@
+# Default survey configuration here has a range of efficiencies and selectivities
+# To emulate a range of species in a single multispecies survey
+# Also now happens in "spring" and "fall"
+# Need to define survey season, area, efficiency, selectivity
+
+#Atlantis model timestep corresponding to the true output--now from census_spec.R
+timestep <- stepperyr #5
+
+#Which atlantis timestep does the survey run in?--now from census_spec.R
+# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
+# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
+
+survey_sample_time <- 1 # spring survey
+
+#The last timestep to sample
+total_sample <- noutsteps-1 #495
+
+#Vector of indices of survey times to pull
+survey_sample_full <- seq(survey_sample_time,
+                          total_sample, by=timestep)
+
+survtime <- survey_sample_full
+
+# survey area
+# should return all model areas
+survboxes <- allboxes
+
+# survey efficiency (q)
+# should return a perfectly efficient survey 
+surveffic <- data.frame(species=survspp,
+                     efficiency=rep(1.0,length(survspp)))
+
+# survey selectivity (agecl based)
+# this is by age class, need to change to use with ANNAGEBIO output
+#survselex <- data.frame(species=rep(survspp, each=n_age_classes),
+#                     agecl=rep(c(1:n_age_classes),length(survspp)),
+#                     selex=rep(1.0,length(survspp)*n_age_classes))
+
+# for annage output
+survselex <- data.frame(species=rep(survspp, n_annages), #  
+                        agecl=unlist(sapply(n_annages,seq)),
+                        selex=rep(1.0,sum(n_annages)))
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+
+# survey index cv needed for sample_survey_xxx
+# cv = 0.1
+surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp)))
+
+# length at age cv for input into calc_age2length function
+# function designed to take one cv for all species, need to change to pass it a vector
+lenage_cv <- 0.1
+
+# max size bin for length estimation, function defaults to 150 cm if not supplied
+maxbin <- 200
+

### 2020-06-19 NOBAmsdiets.Rmd
testing diet functions, plotting output

@@
 ms_diet <- diettest %>%
-  filter(species %in% sppsubset$Name)
+  filter(species %in% sppsubset$Name) %>%
+  filter(atoutput>0)
 
@@
 
-## Generating the dataset
+### Plotting diets {.tabset}
+
+```{r plotdietcomp}
+
+plist = lapply(split(ms_diet, ms_diet$species), function(d) {
+  ggplot(d, aes(time.days/365, atoutput, fill=prey)) + 
+    geom_bar(stat = "identity") +
+    facet_wrap(species~agecl) +
+    xlab("year") +
+    ylab("diet proportion") +
+    theme_tufte() +
+    theme(legend.position="bottom")
+})
+
+```
+
+#### Capelin
+```{r}
+plist$Capelin
+```
+
+#### Blue whiting
+```{r,  fig.width=9, fig.height=8}
+plist$Blue_whiting
+```
+
+#### Saithe
+```{r, fig.width=9, fig.height=8}
+plist$Saithe
+```
+
+#### Cod
+```{r, fig.width=9, fig.height=9}
+plist$North_atl_cod
+```
+
+#### Haddock
+```{r, fig.width=9, fig.height=9}
+plist$Haddock
+```
+
+#### Greenland halibut
+```{r, fig.width=9, fig.height=9}
+plist$Green_halibut
+```
+
+#### Herring
+```{r, fig.width=9, fig.height=8}
+plist$Norwegian_ssh
+```
+
+#### Redfish
+```{r, fig.width=9, fig.height=8}
+plist$Redfish
+```
+
+#### Mackerel
+```{r, fig.width=9, fig.height=8}
+plist$Mackerel
+```
+
+#### Long rough dab
+```{r, fig.width=9, fig.height=9}
+plist$Long_rough_dab
+```
+
+#### Polar cod
+```{r, fig.width=9, fig.height=8}
+plist$Polar_cod
+```
+
+### {-}
+
+### Interactions among these species
+
+## Generating the dataset *(PLACEHOLDER; not done yet June 19 2020)*
 
@@
 
-```{r full-test, message=FALSE, warning=FALSE}
+```{r full-test, message=FALSE, warning=FALSE, eval=FALSE}
 
@@
 #### Biomass index
-```{r bioind-1}
+```{r bioind-1, eval=FALSE}
 
@@
 #### Catch time series
-```{r catchind-1}
+```{r catchind-1, eval=FALSE}
 #read time series data
@@
 #### Survey length composition
-```{r lencomp-1}
+```{r lencomp-1, eval=FALSE}
 
@@
 #### Survey age composition (age classes)
-```{r agecomp-1}
+```{r agecomp-1, eval=FALSE}
 
@@
 #### Survey weight at age (age classes)
-```{r wtageclass-1}
+```{r wtageclass-1, eval=FALSE}
 
@@
 #### Fishery age composition (age classes)
-```{r fishagecomp-1}
+```{r fishagecomp-1, eval=FALSE}
 
@@
 #### Survey age composition (annual ages)
-```{r plot-annagecomp}
+```{r plot-annagecomp, eval=FALSE}
 #read in comp data
@@
 #### Survey iterpolated weight at age (annual ages) 
-```{r plot-annwtage}
+```{r plot-annwtage, eval=FALSE}
 
@@
 #### Fishery catch at age (annual ages)
-```{r plot-catchagecomp}
+```{r plot-catchagecomp, eval=FALSE}
 
@@
 #### Fishery iterpolated weight at age (annual ages) 
-```{r plot-fishannwtage}
+```{r plot-fishannwtage, eval=FALSE}
 
@@
 ### {-}
-
-Note: NOBA doesn't seem to have a selective fishery for cod. Lots of age 1 in the catch!
-
-# (Annage development details below)
-
-Now to modify the underlying code to get full age structure.
-
-## Updating functions for full age structure
-
-General strategy:
-
-1. add annual age option to run_truth function call, default it to FALSE
-  + if true, looks for ANNAGEBIO and ANNAGECATCH in designated directory, error out if missing
-  + if true, read these files in addition to existing set rather than replace existing
-    + read them in with new load_nc_annage function
-    + add them to the output object in addition to existing set
-1. write load_nc_annage
-  + need to understand annage nc structures
-    + if same as others just with full age structure, replace previous nc
-    + only numbers output? 
-      + how to match with resn and structn for lengths/wts?
-      + reconcile with bio parameters such as maturity ogive?
-    + only for some groups?
-
-What is in the ANNAGEBIO and ANNAGECATCH .nc files?
-
-```{r annagebio-contents}
-
-# taken from existing load_nc
-
-file.nc <- file.path(d.name, "outputnordic_runresults_01ANNAGEBIO.nc")
-
-  # Load ATLANTIS output!
-  at_out <- RNetCDF::open.nc(con = file.nc)
-
-  # Get info from netcdf file! (Filestructure and all variable names)
-  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
-    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
-  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
-  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
-  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
-
-RNetCDF::close.nc(at_out)
-
-```
-
-ANNAGEBIO.nc variable names (first hundred): `r var_names_ncdf[1:100]`
-
-ANNABEBIO.nc variable names (last hundred): `r var_names_ncdf[1590:1690]`
-
-ANNAGEBIO.nc timesteps: `r n_timesteps`
-
-ANNAGEBIO.nc boxes: `r n_boxes`
-
-ANNAGEBIO.nc layers: `r n_layers`
-
-Annual age output is in numbers and weight at each timestep, layer, and box, which is really good news.
-We should be able to get weight at true age more directly this way than the interpolation I was planning.
-Also, we may be able to do length estimation differently?
-
-To modify `load_nc_annage` we need to figure out how many true ages each species has from the biological parameter file using `load_biolprm`:
-
-```{r biolprm}
-
-biolprm <- load_biolprm(d.name, biol.prm.file)
-fgs <- load_fgs(d.name, functional.groups.file)
-
-#max age should be ages per cohort times the number of age classes (<10 for some NOBA spp)
-
-names(biolprm$agespercohort) <- c("code", "ageperagecl")
-
-maxage <- merge(biolprm$maturityogive, biolprm$agespercohort, all.x = T) %>%
-    select(code, nagecl, ageperagecl) %>%
-    mutate(maxage = nagecl * ageperagecl) %>%
-    mutate(name = fgs$Name[match(code, fgs$Code)])
-
-select_groups <- c('Mackerel', 'North_atl_cod', 'Capelin')
-
-for (i in seq_along(select_groups)) {
-  select_group_ages <- 1:maxage$maxage[match(select_groups[i], maxage$name)]
-}
-
-```
-
-Try this `load_nc_annage` function:
-
-```{r loadncannage, eval=FALSE}
-
-load_nc_annage <- function(dir = getwd(), file_nc, bps, fgs, biolprm, select_groups,
-  select_variable =
-  c("Nums", "Weight"),
-  check_acronyms = TRUE, bboxes = c(0), verbose = FALSE) {
-  # NOTE: The extraction procedure may look a bit complex...
-  # A different approach would be to
-  # create a dataframe for each variable (e.g. GroupAge_Nums)
-  # and combine all dataframes at the end.
-  # This alternative approach requires a lot more storage
-  # and the code wouldn't be vectorized.
-
-  # Check input of the nc file
-  if (tail(strsplit(file_nc, "\\.")[[1]], 1) != "nc") {
-    stop("The argument for file_nc,", file_nc, "does not end in nc")
-  }
-  if (is.null(dir)) {
-    file.nc <- file_nc
-  } else {
-    file.nc <- file.path(dir, file_nc)
-  }
-
-  # Check input of select_variable as only one value is allowed
-  select_variable <- match.arg(select_variable, several.ok = FALSE)
-
-  # Check input structure!
-  if (check_acronyms) {
-    active_groups <- as.vector(subset(fgs, IsTurnedOn == 1)$Name)
-    inactive_groups <- select_groups[which(
-      !is.element(select_groups, active_groups))]
-    if (length(inactive_groups) >= 1) {
-      select_groups <- select_groups[!is.element(select_groups, inactive_groups)]
-      warning(paste(paste("Some selected groups are not active in the model run.",
-        "Check 'IsTurnedOn' in fgs\n"),
-        paste(inactive_groups, collapse = "\n")))
-    }
-    if (all(!is.element(select_groups, active_groups))) {
-      stop(paste("None of the species selected are active in the model run.",
-        "Check spelling and Check 'IsTurnedOn' in fgs"))
-    }
-  }
-
-  # Deal with file structures
-
-  # Load ATLANTIS output!
-  at_out <- RNetCDF::open.nc(con = file.nc)
-  on.exit(RNetCDF::close.nc(at_out))
-
-  if (select_variable != "N" & all(is.element(select_groups, bps))) {
-    stop("The only output for Biomasspools is N.")
-  } else{
-    print(paste("Read", file.nc, "successfully"))
-  }
-
-  # Get info from netcdf file! (Filestructure and all variable names)
-  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
-    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
-  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
-  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
-  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
-
-  # Extract data from the ncdf file
-  # Create a vector of all potential variable names
-  # Only use names which are available in the ncdf-file as an
-  # extraction of missing variables is not possible
-  # Create vector of available species at the end using search_clean
-  # This is needed to create species-names later on
-
-  # CHANGED hardcoded cohort number to reflect true ages from
-
-  # To make the creation of variables as robust as possible
-  # we introduce different combinations of groups, variable, and cohort
-  # Only combinations present in the ncdf are used later on
-  # Loop over select_groups to use the same ordering
-
-  names(biolprm$agespercohort) <- c("code", "ageperagecl")
-
-  maxage <- merge(biolprm$maturityogive, biolprm$agespercohort, all.x = T) %>%
-    select(code, nagecl, ageperagecl) %>%
-    mutate(maxage = nagecl * ageperagecl) %>%
-    mutate(name = fgs$Name[match(code, fgs$Code)])
-
-  search <- list()
-  for (i in seq_along(select_groups)) {
-    select_group_ages <- 1:maxage$maxage[match(select_groups[i], maxage$name)]
-    search[[i]] <- c(
-      unlist(lapply(paste0(select_groups[i], select_group_ages),
-                    paste0, select_variable)),           # GroupCohortVariable
-      unlist(lapply(paste0(select_groups[i], select_variable),
-                    paste0, select_group_ages)),                   # GroupVariableCohort
-      unlist(lapply(paste0(select_groups[i], select_group_ages),
-                    paste, select_variable, sep = "_")), # GroupCohort_Variable
-      unlist(lapply(paste(select_groups[i], select_variable, sep = "_"),
-                    paste0, select_group_ages)),                   # Group_VariableCohort
-      unlist(lapply(paste(select_groups[i], select_group_ages, sep = "_"),
-                    paste, select_variable, sep = "_")), # Group_Cohort_Variable
-      unlist(lapply(paste(select_groups[i], select_variable, sep = "_"),
-                    paste, select_group_ages, sep = "_")),         # Group_Variable_Cohort
-      unlist(paste0(select_groups[i], select_variable)),                                                      # GroupVariable
-      unlist(paste(select_groups[i], select_variable,
-                   sep = "_"))                           # Group_Variable
-      )
-    search[[i]] <- search[[i]][is.element(search[[i]], var_names_ncdf)]
-    search[[i]] <- unique(search[[i]])
-  }
-  search_clean <- do.call(c, search)
-  # If the combination of select_groups and select_variable ends up not being found.
-  if (length(search_clean) == 0) return(0)
-
-  at_data <- lapply(search_clean, RNetCDF::var.get.nc, ncfile = at_out)
-
-  # Get final species and number of ageclasses per species
-  final_species <- select_groups[sapply(
-    lapply(select_groups, grepl, x = search_clean), any)]
-  final_agecl <- maxage$maxage[
-    sapply(final_species, function(x) which(x == maxage$name))]
-
-  num_layers <- RNetCDF::var.get.nc(ncfile = at_out, variable = "numlayers")[, 1]
-  # add sediment layer!
-  num_layers <- num_layers + ifelse(num_layers == 0, 0, 1)
-
-  # Create an array of layerids.
-  # Every entry in the array indicates if a layer is present (= 1) or not (= 0).
-  # Boxes without layers (= islands) have only 0s as id,
-  # used later on to remove data from non-existent layers!
-  # By default output should be 0 in those layers.
-  # Layers in boundary boxes are set to 0 if bboxes is anything other than NULL!
-  # Applying a boolean array to an array results in a vector!
-  for (i in seq_along(num_layers)) {
-    if (i == 1) layerid <- array(dim = c(n_layers, n_boxes))
-    if (num_layers[i] == 0) {
-      layerid[, i] <- 0
-    } else {
-      if (!is.null(bboxes) & is.element((i - 1), bboxes)) {
-        layerid[, i] <- 0
-      } else {
-        layerid[, i] <- c(rep(0, times = n_layers - num_layers[i]),
-                          rep(1, times = num_layers[i]))
-      }
-    }
-  }
-
-  # Create vectors for polygons and layers
-  # Each vector has the length equal to one time-step
-  # All data from islands and non-existent layers is removed
-  # Therefore the length of these
-  # vectors is equal for each extracted variable
-  boxes <- 0:(n_boxes - 1)
-  # Remove islands and boundary boxes
-  island_ids <- num_layers == 0
-  if (!is.null(bboxes)) {
-    boundary_ids <- is.element(boxes, bboxes)
-    island_ids <- island_ids | boundary_ids
-  }
-  boxes <- boxes[!island_ids]
-  num_layers <- num_layers[!island_ids]
-
-  polygons <- rep(boxes, times = num_layers)
-  layers <- sapply(num_layers[num_layers != 0] - 2,
-    function(x) c(seq(x, from = 0, by = 1), n_layers - 1))
-  if (any(sapply(layers, length) != num_layers[num_layers != 0])) {
-    stop("Number of layers incorrect.")
-  }
-  layers <- do.call(c, layers)
-  if (length(polygons) != length(layers)) {
-    stop("Number of polygons and layers do not match.")
-  }
-
-  # In the following section the data is transformed to a long dataframe
-  # I haven't found any solution to vectorize the creation of the dataframe
-  # columns (species, age, polygons,...)
-  # when data from 2d and 3d arrays
-  # (e.g. select_variable = "N" all biomasspools are only present in the
-  # sediment layer.) are read in simultaneously.
-  # Therefore the current "messy" solution splits the data
-  # in 2 subpopulations: 2d-data and 3d-data
-  at_data3d <- at_data[which(sapply(at_data, function(x) length(dim(x))) == 3)]
-  at_data2d <- at_data[which(sapply(at_data, function(x) length(dim(x))) == 2)]
-
-  int_fs <- final_species
-  int_fa <- final_agecl
-
-  if (length(at_data3d) >= 1) {
-    # Remove biomasspools if selected variable is "N"!
-    if (select_variable == "N") {
-      int_fs <- final_species[!is.element(final_species, bps)]
-      int_fa <- final_agecl[!is.element(final_species, bps)]
-      # Note this only works if age-structured vertebrates have 10 ageclasses
-      int_fa[int_fa == 10] <- 1
-    }
-    for (i in seq_along(at_data3d)) {# for loop over all variables
-      if (i == 1) result3d <- list()
-      for (j in 1:n_timesteps) {# loop over timesteps
-        if (j == 1) values <- array(dim = c(length(layers), n_timesteps))
-        values[, j] <- at_data3d[[i]][,, j][layerid == 1]
-      }
-      result3d[[i]] <- as.vector(values)
-    }
-    result3d <- data.frame(
-      species = unlist(sapply(
-        X = mapply(FUN = rep, x = int_fs, each = int_fa, SIMPLIFY = FALSE,
-        USE.NAMES = FALSE),
-        FUN = rep, each = length(layers) * n_timesteps, simplify = FALSE)),
-      agecl = unlist(sapply(
-        X = sapply(X = int_fa, FUN = seq, from = 1, by = 1, simplify = FALSE,
-        USE.NAMES = FALSE),
-        FUN = rep, each = length(layers) * n_timesteps, simplify = FALSE)),
-      polygon = unlist(sapply(
-        X = n_timesteps * int_fa, FUN = rep, x = polygons, simplify = F,
-        USE.NAMES = FALSE)),
-      layer = unlist(sapply(
-        X = n_timesteps * int_fa, FUN = rep, x = layers, simplify = FALSE,
-        USE.NAMES = FALSE)),
-      time = unlist(sapply(
-        X = int_fa, FUN = rep, x = rep(0:(n_timesteps - 1), each = length(layers)),
-        simplify = FALSE, USE.NAMES = FALSE)),
-      atoutput = do.call(c, result3d),
-      stringsAsFactors = FALSE)
-  }
-
-  if (length(at_data2d) >= 1) {
-    # Only select biomasspools if selected variable is "N"!
-    if (select_variable == "N") {
-      int_fs <- final_species[is.element(final_species, bps)]
-      int_fa <- final_agecl[is.element(final_species, bps)]
-    }
-    # age-structured invert groups are combined in ncdf file!
-    if (select_variable == "Grazing") int_fa <- 1
-    for (i in seq_along(at_data2d)) {# for loop over all variables!
-      if (i == 1) result2d <- list()
-      for (j in 1:n_timesteps) {# loop over timesteps
-        if (j == 1) values <- array(dim = c(length(boxes), n_timesteps))
-        values[, j] <- at_data2d[[i]][, j][boxes + 1]
-      }
-      result2d[[i]] <- as.vector(values)
-    }
-
-    # Order of the data in value column = "atoutput".
-    # 1. species  --> rep each with the number of
-    #                 ageclasses and n_timesteps * boxes
-    # 2. age      --> rep each (1:10 for each species) with n_timesteps * boxes
-    # 3. timestep --> rep each timestep (1:n_timesteps)
-    #                 with the number of boxes and final_agecl
-    #                 (num cohorts per species)
-    # 4. polygon  --> rep boxes times n_timesteps * final_agecl
-    #                 (num cohorts per species)
-    result2d <- data.frame(species = unlist(sapply(
-      X = mapply(FUN = rep, x = int_fs, each = int_fa, SIMPLIFY = FALSE,
-        USE.NAMES = FALSE),
-        FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
-      agecl = unlist(sapply(X = sapply(X = int_fa, FUN = seq, from = 1,
-        by = 1, simplify = FALSE, USE.NAMES = FALSE),
-        FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
-      polygon = unlist(sapply(X = n_timesteps * int_fa,
-        FUN = rep, x = boxes, simplify = FALSE, USE.NAMES = FALSE)),
-      time = unlist(sapply(X = int_fa, FUN = rep, x = rep(0:(n_timesteps - 1),
-        each = length(boxes)), simplify = FALSE, USE.NAMES = FALSE)),
-      atoutput = do.call(c, result2d),
-      stringsAsFactors = F)
-    if (select_variable == "N") result2d$layer <- n_layers - 1
-  }
-
-  # Combine dataframes if necessary!
-  if (all(sapply(lapply(at_data, dim), length) == 3) & select_variable != "N") {
-    result <- result3d
-  }
-  if (all(sapply(lapply(at_data, dim), length) == 2) & select_variable != "N") {
-    result <- result2d
-  }
-  if (select_variable == "N") {
-    if (length(at_data2d) >= 1 & length(at_data3d) == 0) result <- result2d
-    if (length(at_data2d) == 0 & length(at_data3d) >= 1) result <- result3d
-    if (length(at_data2d) >= 1 & length(at_data3d) >= 1) {
-      result <- rbind(result2d, result3d)
-    }
-  }
-
-  # Remove min_pools if existent (well, there always are min pools... ;)).
-  min_pools <- is.element(result$atoutput, c(0, 1e-08, 1e-16))
-  if (length(min_pools) > 0) {
-    # exclude 1st timestep and sediment layer from calculation
-    print_min_pools <- sum(min_pools) -
-      length(result[min_pools & result$time == 1, 1]) -
-      length(result[min_pools & result$time > 1 & result$layer == 7, 1])
-    if (print_min_pools > 0 & verbose) {
-      warning(paste0(round(print_min_pools/dim(result)[1] * 100),
-        "% of ", select_variable, " are true min-pools (0, 1e-08, 1e-16)"))
-    }
-    result <- result[!min_pools, ]
-  }
-
-  # Remove non-existent layers.
-  # WARNING: Biomass is build up (very few) in sediment layer for
-  # NON sediment groups (e.g. baleen whales)
-  # Therefore, I subset all data from that layer for non biomass groups and
-  # groups which cannot penetrate into the sediment!
-  # UPDATE: Doesn't work with layers as species are not distributed through
-  # the whole water column and do not appear in
-  # every polygon.
-
-  # Sum up N for invert cohorts if invert cohorts are present!
-  # NOTE: invert cohorts of size 10 are not considered!
-  if (select_variable == "N" & any(final_agecl != 10 & final_agecl > 1)) {
-    result <- result %>%
-      dplyr::group_by(polygon, layer, species, time) %>%
-      dplyr::summarise(atoutput = sum(atoutput))
-  }
-
-  return(result)
-}
-
-```
-
-Does it work? This snippet will go into `run_truth` if it does:
-
-```{r testloadnc}
-
-file_nc <- ("outputnordic_runresults_01ANNAGEBIO.nc")
-bps <- load_bps(dir = d.name, fgs = functional.groups.file, file_init = biomass.pools.file)
-allboxes <- load_box(dir = d.name, file_bgm = box.file)
-boxes <- get_boundary(allboxes)
-
-numsage <- load_nc_annage(dir = d.name,
-                  file_nc = file_nc,
-                  bps = bps,
-                  fgs = fgs,
-                  biolprm = biolprm,
-                  select_groups = select_groups,
-                  select_variable = "Nums",
-                  check_acronyms = TRUE,
-                  bboxes = boxes, 
-                  verbose = TRUE)
-  #if(verbose) message("Numbers read in from ANNAGEBIO.")
-
-weightage <- load_nc_annage(dir = d.name,
-                  file_nc = file_nc,
-                  bps = bps,
-                  fgs = fgs,
-                  biolprm = biolprm,
-                  select_groups = select_groups,
-                  select_variable = "Weight",
-                  check_acronyms = TRUE,
-                  bboxes = boxes, 
-                  verbose = TRUE)
-  #if(verbose) message("Weight read in from ANNAGEBIO.")
-
-```
-
-Need to aggreagate these to see if they look reasonable; this code snippet from `om_comps` with `numsage` output subsituted for `nums`.
-
-```{r fullagecomps-test}
-
-omlist_ss<-NOBAom_cod
-source(here("/config/NOBA2Config.R"))
-source(here("/config/omdimensions.R"))
-source(here("/config/codsurvey.R"))
-source(here("/config/codfishery.R"))
-
-n_reps <- 1
-
-  #numbers based fishery independent survey for age and length comps
-  # same user specifications as indices
-  survey_N <- atlantisom::create_survey(dat = numsage,
-                                        time = survtime,
-                                        species = survspp,
-                                        boxes = survboxes,
-                                        effic = surveffic,
-                                        selex = survselex)
-  #Sample fish for age composition
-  # if we want replicates for obs error this sample function will generate them
-  age_comp_data <- list()
-  for(i in 1:n_reps){
-    age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
-  }
-
-    # save age comps
-  #if(save){
-    saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
-  #}
-
-```
-
-Plot full age comps, have 20 years for cod now, no longer age class but annual ages.
-
-```{r plotfullagecomp}
-
-#read in comp data
-age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
-age_comp_data <- age_comp_data[[1]]
-
-#fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
-#fish_age_comp <- fish_age_comp[[1]]
-
-#add this to om_indices function so that this has years when read in
-#fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
-
-Natage <- filter(age_comp_data, time %in% c(55:175))
-
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Natage$species))
-
-  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
-
-```
-
-What is the weight output?
-Documentation only mentions numbers output.
-Looks like average weight at age for each time and polygon?
-If so, need to do weighted average weight using the numbers at age in each polygon?
-
-This output is described in Atlantis code, file atlantis/atphysics/atagetracerIO.c, function `writeBMAnnAgeBioData`:
-
-```
-if (propit == out_nums_id) {  // Numbers case
-                                for (b = 0; b < bm->nbox; b++) {
-                                    val[b] = 0.0;
-                                    for (k = 0; k < bm->wcnz; k++) {  // Sum over all individuals in the box
-                                        val[b] += bm->wctr[b][k][id] * FunctGroupArray[sp].boxPopRatio[b][k][cohort][ai];
-                                    }
-                                }
-                            } else  { // Size data - get weighted average of values per layer if numbers > min_dens
-                                for (b = 0; b < bm->nbox; b++) {
-                                    val[b] = 0.0;
-                                    tot = 0.0;
-                                    for (k = 0; k < bm->wcnz; k++) {  // Sum over all individuals in the box
-                                        if (bm->wctr[b][k][id] > bm->min_dens) {
-                                            val[b] += bm->wctr[b][k][id] * (bm->wctr[b][k][sn] + bm->wctr[b][k][rn]) * 5.7 * mg_2_g * FunctGroupArray[sp].boxPopRatio[b][k][cohort][ai];
-                                            tot += bm->wctr[b][k][id];
-                                        }
-                                    }
-                                    if (tot > 0.0)
-                                        val[b] /= tot;
-                                }
-                            }
-```
-
-So weight output appears to be average weight at age in grams?
-
-```{r wtage}
-
-annwtage <- weightage %>%
-  group_by(species, agecl, time) %>%
-  summarise(wtage = mean(atoutput)) %>%
-  filter(time %in% (survtime))
-
-boxwtage <- weightage %>%
-  mutate(wtage = atoutput) %>%
-  filter(species %in% c("North_atl_cod"),
-         polygon %in% c(1,7,24,30,38,45),
-         agecl %in% c(16:20))
-
-wtagedat <- boxwtage
-
-wageplot <- ggplot(wtagedat, aes(time, wtage)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (5 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " "))
-
-wageplot + facet_wrap(c("polygon"), scales="free_y")
-
-```
-
-So, to modify run_truth we need to add the annage output but not replace the regular output in case users want output from non-age structured groups (inverts, etc.).
-Also need to see if lengths can be generaged from the annage outputs, if not need to keep the full set of outputs from the .nc file for all species to generate them.
-
-UPDATE: there is a difference between the annage weight at age and that calculated using 'calc_age2length()'.
-While the results from the standard nc output look reasonable for NOBA cod, the annage weight at age is half that for the standard (and half what real cod would be).
-More worrisome, the oldest age classes have smaller average weight at age than some younger ones.
-I cannot think of why this would be.
-Until we are sure the annage weight output is correct, we will instead interpolate the weight at age output of 'calc_age2length" for use as assessment input.
-
-Looks like the annagecatch output is only numbers and not weights. We don't get weight at age from fisheries usually so this is fine.
-
-```{r loadannagecatch}
-# taken from existing load_nc
-
-file.nc <- file.path(d.name, "outputnordic_runresults_01ANNAGECATCH.nc")
-
-  # Load ATLANTIS output!
-  at_out <- RNetCDF::open.nc(con = file.nc)
-
-  # Get info from netcdf file! (Filestructure and all variable names)
-  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
-    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
-  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
-  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
-  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
-
-RNetCDF::close.nc(at_out)
-```
-
-ANNAGECATCH.nc variable names (first hundred): `r var_names_ncdf[1:100]`
-
-ANNAGECATCH.nc variable names (last hundred): `r tail(var_names_ncdf, 100)`
-
-ANNAGECATCH.nc timesteps: `r n_timesteps`
-
-ANNAGECATCH.nc boxes: `r n_boxes`
-
-ANNAGECATCH.nc layers: `r n_layers`
-
-Catch file contains "IsFished" species discard at age and catch at age by fishery in numbers.
-So now modifying load_nc_annage to get the fishery specific outputs, which requires additional atlantis fishery setup files and additions to the config files above.
-
-New function `load_fisheries()` needed to load the fishery specification file (a .csv), similar to `load_fgs()`:
-
-```{r load_fisheries, eval=FALSE}
-
-load_fisheries <- function(dir = getwd(), file_fish){
-  if (is.null(dir)) {
-    file.fish <- file_fish
-  } else {
-    file.fish <- file.path(dir, file_fish)
-  }
-  result <- read.table(file = file.fish, sep = ",",
-    header = TRUE, stringsAsFactors = FALSE)
-  return(result)
-}
-```
-
-Call `load_fisheries()` within `load_nc_annage()` to rename select_variable as used in ANNAGECATCH.nc:
-
-```{r addsnippet, eval=FALSE}
-
-select_variable <- c("Catch", "Discard")
-dir <- here("atlantisoutput","NOBA_march_2020")
-file_fish <- "NoBAFisheries.csv"
-
-#starting at line 115 in current load_nc_annage
-
-if(select_variable %in% c("Catch", "Discard")){
-    #read in fleet names
-    fleetnames <- load_fisheries(dir = dir, file_fish = file_fish)
-    svfish <- c()
-    for(i in select_variable){
-      sv <-paste0(i, "_", fleetnames$Code)
-      svfish <- c(svfish, sv)
-    }
-    select_variable <- svfish
-  }
-
-```
-
-Also added if statements to add fleets for fisheries outputs, if statements to separate output for annage and annagecatch.
-
-Test new `load_nc_annage()`:
-
-```{r testfishery}
-
-# source(here("R/load_nc_annage.R"))
-# source(here("R/load_fisheries.R"))
-
-source(here("config/NOBA2Config.R"))
-NOBAom_cod <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
-
-fgs <- NOBAom_cod$funct.group_ss
-biolprm <- NOBAom_cod$biol
-#select_groups <- c('Mackerel', 'North_atl_cod', 'Capelin')
-select_groups <- 'North_atl_cod'
-
-bps <- load_bps(dir = d.name, fgs = functional.groups.file, file_init = biomass.pools.file)
-allboxes <- load_box(dir = d.name, file_bgm = box.file)
-boxes <- get_boundary(allboxes)
-
-file_nc <- "outputnordic_runresults_01ANNAGECATCH.nc"
-file_fish <- "NoBAFisheries.csv"
-
-#for testing line by line
-bboxes <- boxes
-dir <- d.name
-select_variable <- "Catch"
-
-fishcatchage <- load_nc_annage(dir = d.name,
-                  file_nc = file_nc,
-                  file_fish = file_fish,
-                  bps = bps,
-                  fgs = fgs,
-                  biolprm = biolprm,
-                  select_groups = select_groups,
-                  select_variable = "Catch",
-                  check_acronyms = TRUE,
-                  bboxes = boxes, 
-                  verbose = TRUE)
-
-fishcatchage[1:100,]
-
-fishdiscage <- load_nc_annage(dir = d.name,
-                  file_nc = file_nc,
-                  file_fish = file_fish,
-                  bps = bps,
-                  fgs = fgs,
-                  biolprm = biolprm,
-                  select_groups = select_groups,
-                  select_variable = "Discard",
-                  check_acronyms = TRUE,
-                  bboxes = boxes, 
-                  verbose = TRUE)
-
-fishdiscage
-```
-
-Do we want catch by fishery? Yes. No fishery column here, check what it is doing.
-fishdiscage is empty, make sure there is no discard output from NOBA for these species (confirmed).
-
-Add a check for multiple fisheries per species, if only one then fleet irrelevant.
-
-NOBA has multiple fleets for cod (but they are not in output!), so add fleet column to output.
-
-Update: there is now a fleet column in output which should index correctly if a species is caught by multiple fleets.
-
-Plot catch at age; aggregate over polygons, apply sample_fish:
-
-```{r catchannage}
-
-omlist_ss<-NOBAom_cod
-source(here("/config/NOBA2Config.R"))
-source(here("/config/omdimensions.R"))
-source(here("/config/codsurvey.R"))
-source(here("/config/codfishery.R"))
-
-n_reps <- 1
-
-  #numbers based fishery independent survey for age and length comps
-  # same user specifications as indices
-  #fishery catch at age each observed timestep summed over observed polygons
-  # catch at age by area and timestep
-  catch_numbers <-  atlantisom::create_fishery_subset(dat = fishcatchage,
-                                                      time = fishtime,
-                                                      species = select_groups,
-                                                      boxes = fishboxes)
-
-  # if we want replicates for obs error this sample function will generate them
-  catch_age_comp <- list()
-  for(i in 1:n_reps){
-    catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
-  }
-
-  # save fishery age comps
-  #if(save){
-    saveRDS(catch_age_comp, file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
-  #}
-
-```
-
-NOTE that the fleet column is lost in the `atlantisom::sample_fish` function!
-
-Look at only every 10 years output (this is a sample, effN = 1000 for the year, which is 200 per timestep here; see calc below confirming):
-
-```{r plotfullcatchage}
-
-#read in comp data
-# age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
-# age_comp_data <- age_comp_data[[1]]
-
-fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
-fish_age_comp <- fish_age_comp[[1]]
-
-#add this to om_indices function so that this has years when read in
-fish_age_comp$time <- fish_age_comp$time/fstepperyr
-
-Catage <- filter(fish_age_comp, time %in% c(30.0, 40.0, 50.0, 60.0, 70.0, 79.8))
-
-Catageplot <- ggplot(Catage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Catage$species))
-
-  Catageplot + facet_wrap(~time, ncol=3, scales="free_y")
-
-```
-
-```{r testeffn}
-
-fish_age_comp %>%
-  filter(time<31) %>%
-  summarise(sum(atoutput))
-
-```
-
-Appears to be working for single fleet per species now.
-
-Quick test of `run_truth()` with annage option turned on for NOBA:
-
-```{r testruntruth, eval=FALSE}
-
-source(here("config/NOBA2Config.R"))
-
-  #Load functional groups
-  funct.groups <- atlantisom::load_fgs(dir=d.name,
-                           file_fgs = functional.groups.file)
-  #Get just the names of active functional groups
-  funct.group.names <- funct.groups %>% 
-    filter(IsTurnedOn == 1) %>%
-    select(Name) %>%
-    .$Name
-
-  # # load biolprm file
-  #biol <- atlantisom::load_biolprm(d.name, biol.prm.file)
-  # 
-  # # load run_prm
-  # runpar <- atlantisom::load_runprm(d.name, run.prm.file)
-  # 
-  # # load box info file
-  # boxpars <- atlantisom::load_box(d.name, box.file)
-
-testtruth <- atlantisom::run_truth(scenario = scenario.name,
-                       dir = d.name,
-                       file_fgs = functional.groups.file,
-                       file_bgm = box.file,
-                       select_groups = funct.group.names,
-                       file_init = initial.conditions.file,
-                       file_biolprm = biol.prm.file,
-                       file_runprm = run.prm.file,
-                       file_fish = fisheries.file,
-                       verbose = TRUE,
-                       annage = TRUE,
-                       save = TRUE
-    )
-
-```
-
-After much debugging to properly index multiple fleets per species in catch at annual age output, this produced a run_truth.Rdata file and took about 40 minutes to run.
-
-Now all of the wrappers need to read from annage to produce composition data.
-`om_init()`  now takes annage flag from the config file.
-`om_species()` looks for annage bio and catch output in the truth file and includes it in the species file (adding NULL if it isn't there).
-`om_comps()` looks for annage bio and catch in the species file and adds them to the comps list if there (adding NULL if they aren't there).
-
-Also still need the weight at age interpolation function. Incorporate this at the `om_comps()` step. Next rmd!
-
-Running `om_init()`, `om_species()`, `om_index()`, `om_comps()` in chunk above to make sure the original output is preserved, then plot the annage output [above](#annageoutput).
-
-## Write input data for specific assessment models
-
-We will need to alter this for SAM, SS3, etc. Right now just makes SS3.
-
-**THIS IS STILL THE CALIFORNIA CURRENT INFO, TO BE UPDATED**
-
-```{r writedat, eval=FALSE}
-
-require(r4ss)
-omlist_ss<-CC3om_sardine
-source(here("/config/CC3Config_sardLWcorr.R"))
-source(here("/config/omdimensions.R"))
-source(here("/config/sardinesurvey.R"))
-source(here("/config/sardinefishery.R"))
-
-#Directory with SS files
-model_dir <- "Sardine_SS_files/"
-
-#Name of SS data file
-datfile_name <- "sardEM_3_3.dat"
-
-#CVs for length at age, catch, and survey
-CVs <- list("lenage"=0.1, "fishery"=0.01, "survey"=0.1)
-
-#Month of survey/fishing
-survey_month <- 7
-fishing_month <- 1
-
-#Years to survey, assuming survey is once per year
-#survey_years <- survey_sample_full[(burnin+1):(burnin+nyears)] # my original
-survey_years <- survey_sample_full[burnin:(burnin+nyears-1)] #from Christine's new sardine_config.R
-
-#read time series data
-survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name, "surveyB.rds")))
-truecatchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
-
-survObsBiom <- survObsBiom[[1]]
-truecatchbio_ss <- truecatchbio_ss[[1]]
-
-#add this to om_indices function so that this has years when read in
-truecatchbio_ss$time <- as.integer(truecatchbio_ss$time/365)
-
-#load dummy dat file
-stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/",
-model_dir,
-datfile_name))
-
-#Test writing CPUE in biomass
-  stocksynthesis.data <- atlantisom::SS_write_ts(ss_data_list = stocksynthesis.data,
-                ts_data = list(survObsBiom$atoutput[fish_years],
-                  truecatchbio_ss$atoutput[truecatchbio_ss$time %in% fish_years]),
-                CVs = c(CVs$survey,
-                        CVs$fishery),
-                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             fish_years),
-            sampling_month = list(rep(survey_month,nyears),
-                                rep(fishing_month,nyears)),
-                units = c("biomass","biomass"),
-                data_type=c("CPUE","catch"),
-                fleets = c(2,1))
-
-#test  
-stocksynthesis.data$CPUE
-
-stocksynthesis.data$catch
-
-#read in comp data
-age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
-age_comp_data <- age_comp_data[[1]]
-
-fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
-fish_age_comp <- fish_age_comp[[1]]
-
-#Get the age bins
-age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
-age_bins <- sub("a","",age_bin_names)
-
-require(maditr)
-# add dependency on maditr::dcast to atlantisom
-age_comp_flat <- atlantisom::reformat_compositions(age_comp_data,
-                     round.places = 4,
-                     comp_type = "agecomp")
-
-## Write age composition data for survey
-stocksynthesis.data <- atlantisom::SS_write_comps(ss_data_list = stocksynthesis.data,
-               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
-               data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr+nyears-1)),
-               sampling_month = list(rep(survey_month,nyears)),
-               data_type = c("agecomp"),
-               fleet_number = c(1),
-               bins = list(age_bins),
-               caal_bool = c(FALSE))
-stocksynthesis.data$agecomp
-
-#length comps
-len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsLenComp.rds")))
-fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))
-
-len_comp_data <- len_comp_data[[1]]
-fish_len_comp_data <- fish_len_comp_data[[1]]
-
-#add this to om_indices function so that this has years when read in
-fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))
-fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
-
-if(fstepperyr>1){
-  fish_len_comp_anndata <- fish_len_comp_data %>%
-    #mutate(yr = floor(time/fstepperyr)) %>%
-    #group_by(species, agecl, lower.bins, upper.bins, time=as.integer(yr)) %>%
-    group_by(species, agecl, lower.bins, upper.bins, time) %>% 
-    summarise(annnatlength=sum(atoutput)) %>%
-    rename(atoutput = annnatlength)
-} else {
-  fish_len_comp_anndata <- fish_len_comp_data
-}
-
-caal_comp_flat <- atlantisom::reformat_compositions(len_comp_data,                                round.places=4,
-                comp_type="caalcomp")
-
-#remove burnin
-caal_comp_final <- filter(caal_comp_flat,
-                         time %in% survey_years)
-
-#Add over age classes to get sample size
-len_comp_flat <- atlantisom::reformat_compositions(len_comp_data,
-                                  round.places = 0,
-                           comp_type="lencomp")
-#remove burnin
-len_comp_final <- filter(len_comp_flat,
-                         time %in% survey_years)
-
-length_bins <- as.integer(names(len_comp_final))
-length_bins <- length_bins[!is.na(length_bins)]
-
-# fishery length comps are still 5 timesteps per year
-# need to aggregate to annual (done above)
-# also,  make effN annual goal/fstepsperyr (done above)
-fish_len_comp_flat <- atlantisom::reformat_compositions(fish_len_comp_anndata,
-                                  round.places = 0,
-                           comp_type="lencomp")
-
-#remove burnin works after adjustment above
-fish_len_comp_final <- filter(fish_len_comp_flat,
-                         time %in% fish_years)
-
-notbins <- c("time", "nsamp")
-
-# fish_length_bins <- as.integer(names(fish_len_comp_final))
-# fish_length_bins <- fish_length_bins[!is.na(fish_length_bins)]
-
-# need to fill empty length bins with 0s to have same bins as survey for SS_write_comps
-missing.lengths <- setdiff(length_bins, names(fish_len_comp_final)[!names(fish_len_comp_final) %in% notbins])
-fish_len_comp_final[as.character(missing.lengths)] <- 0                    # Add them, filled with '0's
-fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")] #
-
-# fishery age comps also 5 timesteps per year
-if(fstepperyr>1){
-  fish_age_comp_anndata <- fish_age_comp %>%
-    #mutate(yr = floor(time/fstepperyr)) %>%
-    #group_by(species, agecl, time=as.integer(yr)) %>%
-    group_by(species, agecl, time) %>%
-    summarise(annnatage=sum(atoutput)) %>%
-    rename(atoutput = annnatage)
-} else {
-  fish_age_comp_anndata <- fish_age_comp
-}
-
-fish_age_comp_flat <- atlantisom::reformat_compositions(fish_age_comp_anndata,
-                           comp_type="agecomp")
-
-#remove burnin (not necessary?fish comps made with fish_years only) 
-fish_age_comp_final <- filter(fish_age_comp_flat,
-                         time %in% fish_years)
-
-# #SS_write_comps breaking because fishery age bins start with 2 not 1; extracting bins from fish file may help?
-# fish_age_bins <- names(fish_age_comp_flat)[!names(fish_age_comp_flat) %in% notbins]
-
-# that leaves an empty column in data file, so instead fill with 0s
-missing.ages <- setdiff(age_bins, names(fish_age_comp_final)[!names(fish_age_comp_final) %in% notbins])
-fish_age_comp_final[missing.ages] <- 0                    # Add them, filled with '0's
-fish_age_comp_final <- fish_age_comp_final[c("time", age_bins, "nsamp")]
-
-comp_list <- list(caal_comp_final,len_comp_final, fish_age_comp_final, fish_len_comp_final)
-
-apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), 
-                    rep(survey_month, nrow(comp_list[[2]])),
-                    rep(fishing_month,nrow(comp_list[[3]])),
-                    rep(fishing_month,nrow(comp_list[[4]])))
-
-# This now runs by ensuring that survey and fishery compositions have the same bins 
-# (filled with 0s for missing bins in fishery relative to survey)
-
-# Write CAAL and length composition data
-stocksynthesis.data <- atlantisom::SS_write_comps(ss_data_list = stocksynthesis.data, 
-                                      comp_matrix = comp_list, 
-                                      data_rows = list((comp_list[[1]]$time-survey_sample_time)/timestep + 1 , (survey_years-survey_sample_time)/timestep + 1,fish_years,fish_years),
-                                      sampling_month = apply_month, 
-                                      data_type = rep(c("agecomp", "lencomp"),2), 
-                                      fleet_number = c(2,2,1,1),  
-                                      bins = list(age_bins, 
-                                                  length_bins, 
-                                                  age_bins, 
-                                                  length_bins), 
-                                      caal_bool = c(TRUE, rep(FALSE,3)))
-
-head(stocksynthesis.data$lencomp)
-head(stocksynthesis.data$agecomp)
-
-#Change length bin structure to match atlantis data
-stocksynthesis.data$lbin_vector <- length_bins
-
-#Get correct number of length bins
-stocksynthesis.data$N_lbins <- length(length_bins)
-
-#Set lbin_method to 1 - this makes the population length bins match the data bins 
-#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
-stocksynthesis.data$lbin_method <- 1
-stocksynthesis.data$binwidth <- "#"
-stocksynthesis.data$minimum_size <- "#"
-stocksynthesis.data$maximum_size <- "#"
-
-#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
-stocksynthesis.data$age_info$minsamplesize <- rep(0.001,2)
-
-SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
-datfile_name),
-                 overwrite=TRUE)
-
-```

### 2020-06-19 NOBAmsdiets.Rmd
add table with diet proportions within the species set

@@
 
-## Generating the dataset *(PLACEHOLDER; not done yet June 19 2020)*
+Do our selected species eat each other? What proportion of each species diet comes from species in our selected group?
+
+This table shows the proportion of diet for each species and age class where only our selected species are considered as prey.
+The statistics are across all time steps in this model run, and *do not* include timesteps where the diet composition was 0.
+
+```{r prop-in-group}
+
+prop <- ms_diet %>%
+  filter(prey %in% unique(species)) %>%
+  select(species, agecl, time.days, prey, atoutput) %>%
+  group_by(species, agecl, time.days) %>%
+  summarise(mspreyprop = sum(atoutput)) #sum over prey each timestep
+
+range <- prop %>%
+  group_by(species, agecl) %>%
+  summarise(minprop = min(mspreyprop),
+         medprop = median(mspreyprop),
+         meanprop = mean(mspreyprop),
+         maxprop = max(mspreyprop))
+
+#knitr::kable(range)
+library(DT)
+datatable(range, rownames = FALSE, options = list(pageLength = 25))
+
+```
+
+Mature age classes of cod and redfish average over a third of their diet compositions from other species in the selected group, suggesting fairly strong predation interactions.
+
+Many of the selected species share zooplankton prey, so there may be species interactions via prey as well.
+
+## Generating the dataset 
+*(PLACEHOLDER; not done yet June 19 2020)*
 

### 2020-08-04 NOBAmsdiets.Rmd
call new om_species with diet

@@
 
+### Total consumption time series
+
+The true data can be generated without updated config files (once I fix missing diets in wrappers!):
+
+```{r OMtruth}
+
+NOBAom <- om_init(here("config/NOBA2config.R")) #run and saved June 19,
+# run again 3 August to update biomass_eaten
+
+#NOBAom_ms <- om_species(sppsubset$Name, NOBAom) #run and saved June 19
+# run again 3 August to test new wrapper that should save biomass_eaten
+NOBAom_ms <- om_species(sppsubset$Name, NOBAom, diet=TRUE)
+
+#source(here("config", "NOBA2config.R"))
+
+# diet data are not in here, need to revise wrapper
+#NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+
+# get from full truth run--big file
+#truth <- get(load(file.path(d.name,
+#                                paste0(scenario.name, "run_truth.RData"))))
+
+# questions: does biomass_eaten have some diet comp info as dietcheck.txt?
+# load dietcheck with om_init using a switch in config file?
+# add or pass along switch to keep biomass_eaten truth output in species file
+# is biomass_eaten correct?
+#   revised load_diet_comp to *not include* NAs that merge poorly
+# polygon specific total mg N consumed per unit volume times volume of polygon
+# then split to prey using global diet fraction (and converted to t)
+
+#bioeaten_ss <- truth$biomass_eaten %>%
+#  filter(species %in% sppsubset$Name)
+
+# compare ms_diet$atoutput to bioeaten_ss$dietfrac
+# should match for species, agecl, time.days, prey
+
+```
+
 ## Generating the dataset 
@@
 
-NOBAom <- om_init(here("config/NOBA2config.R"))
+#NOBAom <- om_init(here("config/NOBA2config.R"))
 
-NOBAom_ms <- om_species(sppsubset$Name, NOBAom)
+#NOBAom_ms <- om_species(sppsubset$Name, NOBAom)
+
+NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
 

### 2020-08-06 NOBAmsdiets.Rmd
need to load detailed diet, a bear

@@
 
-The true data can be generated without updated config files (once I fix missing diets in wrappers!):
+The true data can be generated without updated config files (once I fix missing diets in wrappers!).
 
-```{r OMtruth}
+UPDATE: added an option to bring in biomass_eaten in the om_species wrapper.
+However, the way this is calculated is probably inappropriate for what we want.
+biomass_eaten is the output of box specific total tons consumed by predators from PROD.nc split by global diet comp from DietCheck.txt.
+This means predators will have the same diet comp in all boxes, whether prey were there or not.
+
+To do diet sampling using our survey design, we need to retain actual box-specific diet.
+This is found in the DetailedDietCheck.txt file, which outputs tons consumed since last timestep for each predator by box, layer, and prey.
+Because there is no conversion needed for tons, we can just read this in, get rid of all the zero diet comp values to make it a (possibly) manageable size, and apply the survey from there.
+A new function `load_detailed_diet_comp.R` is developed below.
+
+Bottom line is don't use biomass_eaten; maybe we should take it out of `run_truth()` output.
+
+```{r OMtruth, eval=FALSE}
 
@@
 # should match for species, agecl, time.days, prey
+
+cod5.365 <- NOBAom_ms$truebioeaten_ss %>%
+  filter(species=="North_atl_cod",
+         agecl==5,
+         time.days==365, 
+         polygon==4)
+
+cod5.365.check <- ms_diet %>%
+  filter(species=="North_atl_cod",
+         agecl==5,
+         time.days==365)
+
+#it does, but I think we need this by box
+#it makes no sense to apply global diet comp to boxes, not all prey 
+#may be in the box
+#DetailedDietCheck.txt is the file to read for using survey sampling.
+
+#this file is huge, try data.table::fread
+detaileddiettest <- data.table::fread(file.path(here("atlantisoutput", "NOBAwithAnnAgeOutput"), "outputnordic_runresults_01DetailedDietCheck.txt"),
+                                      data.table=FALSE
+                                      #nrows = 200
+                                      )
+
+diet <- detaileddiettest
+
+#remove rows that are all 0 prey
+#diet <- diet[apply(diet[,-c(1:5)], 1, function(x) !all(x==0)),]
+
+#should do the same thing faster
+diet <- diet[as.logical(abs(as.matrix(diet[,-c(1:5)])) %*% rep(1L,ncol(diet[,-c(1:5)]))), ]
+
+#test pieces of new load_detailed_diet function
+  if(length(grep("Group",colnames(diet)))>0)
+    colnames(diet) <- gsub("Group","Predator",colnames(diet))
+
+  # Change column order
+  diet <- diet[, c("Predator", "Cohort", "Time", "Box", "Layer",
+    names(diet)[which(!names(diet) %in% c("Predator", "Cohort", "Time", "Box", "Layer"))])]
+
+  # Convert to tidy dataframe to allow joining/merging with other dataframes.
+  diet <- tidyr::gather_(data = diet, key_col = "prey", value_col = "dietcomp",
+    #gather_cols = names(diet)[(which(names(diet) == "Updated") + 1):NCOL(diet)])
+    gather_cols = names(diet)[(which(names(diet) %in% fgs$Code))])
+
+  # Get rid of 0 dietcomp rows to make manageable size
+  diet <- filter(diet, dietcomp>0)
+
+  names(diet) <- tolower(names(diet))
+
+  diet$prey <- as.character(diet$prey)
+  diet$predator <- as.character(diet$predator)
+
+  # Change species acronyms to actual names.
+  species_names <- fgs[, c("Name", "Code")]
+  diet$species <- species_names$Name[match(diet$predator, species_names$Code)]
+  diet$prey <- species_names$Name[match(diet$prey, species_names$Code)]
+
+  # diet outputs are not all divisible by toutinc, leave in days
+  diet$time.days <- diet$time #/ toutinc
+
+  #fix cohort to agecl
+
+  diet <- diet %>%
+    dplyr::mutate(agecl = cohort + 1)
+
+  # drop redundant columns and reorder
+  diet <- diet[,c("species", "agecl", "time.days", "box", "layer", "prey", "dietcomp")]
 

### 2020-08-22 NOBAmsdiets.Rmd
dealing with the huge diet file

@@
 
-#source(here("config", "NOBA2config.R"))
+source(here("config", "NOBA2config.R"))
 
@@
   diet <- diet[,c("species", "agecl", "time.days", "box", "layer", "prey", "dietcomp")]
+
+```
+
+Try reading in the box-specific diet in weight and see if the proportions match dietcheck.txt outputs:
+
+Oy, have to deal with filesize first. Try stripping out all the 0 lines:
+
+Even running a "bash" chunk here is still running R `system2()`; I don't think this is worth it. Cant even count the columns without the fans kicking up.
+
+The awk code, for the record, is this, executed in terminal from the directory with the .gz file in it:
+```
+zcat < nordic_runresults_01DetailedDietCheck.txt.gz | awk 'NR > 1{s=0; for (i=6;i<=NF;i++) s+=$i; if (s!=0)print}' | gzip > DetDiet.gz
+```
+On linux it would just say zcat filename; macOS needs zcat < filename to work.
+The awk statement is saying after the header row (NR>1), sum up the values from column 6 to the last column and keep the row if they are >0.
+The last pipe sends the new file to a zip file.
+
+Links consulted: 
+[awk is the right thing](https://www.tim-dennis.com/data/tech/2016/08/09/using-awk-filter-rows.html)
+[awk to delete lines summing to 0](https://unix.stackexchange.com/questions/170588/delete-lines-that-sum-to-zero)  
+[awk a compressed file leaving it compressed](https://unix.stackexchange.com/questions/498587/how-to-extract-information-using-awk-on-gz-files-without-storing-the-uncompress)
+
+Started command at 5:35 pm. Finished at 6:58. Maybe I can fread the output smaller file?
+
+I can but it stripped off the header.
+So I need to grab that and cat it back together.
+cat works on two gzipped files, so check it with first line, make the zip version with second, and concatenate them with third.
+
+```
+zcat < nordic_runresults_01DetailedDietCheck.txt.gz | head -n1 > DetDietHead.txt
+zcat < nordic_runresults_01DetailedDietCheck.txt.gz | head -n1 | gzip > DetDietHead.gz
+cat DetDietHead.gz DetDiet.gz > NOBADetDiet.gz
+```
+
+```{r comparediets, eval=FALSE}
+
+#keeps failing with vector memory exhausted (limit reached?)
+#gets to the step of eliminating 0 rows
+#try awk in a system command
+# this test works but R overhead isnt worth it, faster run from terminal
+#system("zcat < atlantisoutput/NOBA_march_2020/nordic_runresults_01DetailedDietCheck.txt.gz | awk '{print NF}' | uniq")
+
+detaileddiettest <- load_detailed_diet_comp(here("atlantisoutput",
+                                                 "NOBA_march_2020"), 
+                           "NOBADetDiet.gz", 
+                           fgs = fgs)
+
+# detaileddiettest <- load_detailed_diet_comp(here("atlantisoutput",
+#                                                  "NOBA_march_2020"), 
+#                            "nordic_runresults_01DetailedDietCheck.txt.gz", 
+#                            fgs = fgs)
+
+# detailedpartdt <- data.table::fread(file.path(here("atlantisoutput",
+#                                                  "NOBA_march_2020"), 
+#                            "nordic_runresults_01DetailedDietCheck.txt.gz"),
+#                            nrows = 20000
+#                            )
+
+ms_detaileddiet <- detaileddiettest %>%
+  filter(species %in% sppsubset$Name)
 

### 2020-09-04 NOBAmsdiets.Rmd
testing code works function doesnt

@@
 
-detaileddiettest <- load_detailed_diet_comp(here("atlantisoutput",
+diet <- data.table::fread(file.path(here("atlantisoutput",
                                                   "NOBA_march_2020"), 
-                           "NOBADetDiet.gz", 
-                           fgs = fgs)
+                                                "NOBADetDiet.gz"))
+
+  # remove all 0 prey rows--done with awk statement first, file is too big for R to do this
+  #diet <- diet[as.logical(rowSums(diet[,-c(1:5)] != 0)), ]
+
+  # SKG June 2020: changing to the other way around. more new models now,
+  # only one diet function to change, and "Predator" seems clearer for users.
+  if(length(grep("Group",colnames(diet)))>0)
+    colnames(diet) <- gsub("Group","Predator",colnames(diet))
+
+  # Change column order--NOT DATA TABLE SAFE
+  #diet <- diet[, c("Predator", "Cohort", "Time", "Box", "Layer",
+  #  names(diet)[which(!names(diet) %in% c("Predator", "Cohort", "Time", "Box", "Layer"))])]
+
+  data.table::setcolorder(diet, c("Predator", "Cohort", "Time", "Box", "Layer"))
+
+  #remove rows that are all 0 prey
+  #diet <- diet[apply(diet[,-c(1:5)], 1, function(x) !all(x==0)),]
+
+  #should do the same thing faster
+  #diet <- diet[as.logical(abs(as.matrix(diet[,-c(1:5)])) %*% rep(1L,ncol(diet[,-c(1:5)]))), ]
+
+  # dataframe
+  #diet <- as.data.frame(diet)
+
+  diet <- data.table::melt(diet, id = 1:5,
+               variable.name="prey",
+               value.name = "dietcomp")
+
+  # Get rid of 0 dietcomp rows to make manageable size
+  diet <- diet[dietcomp>0]
+  #diet <- filter(diet, dietcomp>0)
+
+  names(diet) <- tolower(names(diet))
+
+  diet$prey <- as.character(diet$prey)
+  diet$predator <- as.character(diet$predator)
+
+  # Change species acronyms to actual names.
+  species_names <- fgs[, c("Name", "Code")]
+  diet$species <- species_names$Name[match(diet$predator, species_names$Code)]
+  diet$prey <- species_names$Name[match(diet$prey, species_names$Code)]
+  #diet <- diet[, -which(colnames(diet) == "predator")]
+
+  # diet outputs are not all divisible by toutinc, leave in days
+  diet$time.days <- diet$time #/ toutinc
+
+  #fix cohort to agecl
+  # diet <- diet %>%
+  #   dplyr::mutate(agecl = cohort + 1)
+  diet <- diet[, agecl := cohort + 1]
+
+  # drop redundant columns and reorder NOT DATA TABLE SAFE
+  #diet <- diet[,c("species", "agecl", "time.days", "box", "layer", "prey", "dietcomp")]
+
+  #leave layer info for now, this is a load function
+
+  #does the Update column matter? this assumes it doesn't
+  #in SETAS_Example, it doesn't make extra comps
+
+  dietcompwt <- data.frame(species = diet$species,
+                         agecl = diet$agecl,
+                         time.days = diet$time.days,
+                         polygon = diet$box,
+                         layer = diet$layer,
+                         atoutput = diet$dietcomp,
+                         prey = diet$prey)
+
+# function not working so test piecewise again
+# detaileddiettest <- load_detailed_diet_comp(here("atlantisoutput",
+#                                                  "NOBA_march_2020"), 
+#                            "NOBADetDiet.gz", 
+#                            fgs = fgs)
 
@@
 
-ms_detaileddiet <- detaileddiettest %>%
+```
+
+Still having problems with the function after editing
+
+```{r detdiettest, eval=FALSE}
+
+# still not working but works line by line
+detdiet <- load_detailed_diet_comp(here("atlantisoutput","NOBA_march_2020"), 
+                                   "NOBADetDiet.gz", 
+                                   fgs = fgs)
+
+ms_detaileddiet <- detdiet %>%
   filter(species %in% sppsubset$Name)

### 2020-09-11 NOBAmsdiets.Rmd
load function finally works

@@
 [awk is the right thing](https://www.tim-dennis.com/data/tech/2016/08/09/using-awk-filter-rows.html)
+
 [awk to delete lines summing to 0](https://unix.stackexchange.com/questions/170588/delete-lines-that-sum-to-zero)
+
 [awk a compressed file leaving it compressed](https://unix.stackexchange.com/questions/498587/how-to-extract-information-using-awk-on-gz-files-without-storing-the-uncompress)
@@
 ```
+Now we diagnose the load function again and try to optimize with data.table functions.
 
@@
 #system("zcat < atlantisoutput/NOBA_march_2020/nordic_runresults_01DetailedDietCheck.txt.gz | awk '{print NF}' | uniq")
+# see notes above for preprocessing step, here debug new load_detailed_diet function
 
-diet <- data.table::fread(file.path(here("atlantisoutput",
-                                                  "NOBA_march_2020"), 
-                                                "NOBADetDiet.gz"))
+dir <- here("atlantisoutput","NOBA_march_2020") 
+file_diet <-  "NOBADetDiet.gz"
+
+  if (is.null(dir)) {
+    diet.file <- file_diet
+  } else {
+    diet.file <- file.path(dir, file_diet)
+  }
+  diet <- data.table::fread(diet.file)
 
@@
 
-  # Change column order--NOT DATA TABLE SAFE
+  # Change column order NOT DATA TABLE SAFE
   #diet <- diet[, c("Predator", "Cohort", "Time", "Box", "Layer",
@@
 
+  # Change column order
   data.table::setcolorder(diet, c("Predator", "Cohort", "Time", "Box", "Layer"))
-
-  #remove rows that are all 0 prey
-  #diet <- diet[apply(diet[,-c(1:5)], 1, function(x) !all(x==0)),]
-
-  #should do the same thing faster
-  #diet <- diet[as.logical(abs(as.matrix(diet[,-c(1:5)])) %*% rep(1L,ncol(diet[,-c(1:5)]))), ]
-
-  # dataframe
-  #diet <- as.data.frame(diet)
 
@@
                value.name = "dietcomp")
+
+  # # Convert to tidy dataframe to allow joining/merging with other dataframes.
+  # diet <- tidyr::gather_(data = diet, key_col = "prey", value_col = "dietcomp",
+  #   #gather_cols = names(diet)[(which(names(diet) == "Updated") + 1):NCOL(diet)])
+  #   gather_cols = names(diet)[(which(names(diet) %in% fgs$Code))])
 
@@
 
-# function not working so test piecewise again
-# detaileddiettest <- load_detailed_diet_comp(here("atlantisoutput",
-#                                                  "NOBA_march_2020"), 
-#                            "NOBADetDiet.gz", 
-#                            fgs = fgs)
-
-# detaileddiettest <- load_detailed_diet_comp(here("atlantisoutput",
-#                                                  "NOBA_march_2020"), 
-#                            "nordic_runresults_01DetailedDietCheck.txt.gz", 
-#                            fgs = fgs)
-
-# detailedpartdt <- data.table::fread(file.path(here("atlantisoutput",
-#                                                  "NOBA_march_2020"), 
-#                            "nordic_runresults_01DetailedDietCheck.txt.gz"),
-#                            nrows = 20000
-#                            )
-
 ```
 
-Still having problems with the function after editing
+Using `data.table` improves speed for the `atantisom::load_detailed_diet_comp()` function.
+But it turns out you need to import data.table within a package in *both* the DESCRIPTION and NAMESPACE files [instructions!](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-importing.html) and now this beast of a function works.
+
+It takes minutes to load the detailed diet, even with pre-processing with awk as above, so saving the output is a good idea. We save only for our 10 species to streamline things a bit.
 
@@
 
-# still not working but works line by line
+# still not working but works line by line--fixed
+# N.B. data.table wants to be in both imports and namespace within a function!
 detdiet <- load_detailed_diet_comp(here("atlantisoutput","NOBA_march_2020"), 
@@
   filter(species %in% sppsubset$Name)
+
+saveRDS(ms_detaileddiet, file.path(d.name, paste0(scenario.name, "ms_detaileddiet.rds")))
+
+```
+
+All this was just to see what the regional diet data looks like compared with global diet comps from DietCheck.txt.
+
+```{r detdiet-vis}
+
+# look at a single species/age combo from above
+
+# how many polygons included? how many layers per polygon have data?
+
+# are there differences by layer within a polygon?
+
+# are there differences by polygon? (aggregate layers)
 

### 2020-09-11 NOBAmsdiets.Rmd
by layer plots, still breaking on by polygon

@@
 
-```{r detdiet-vis}
+```{r detdiet-vis1}
 
 # look at a single species/age combo from above
+#    agecl 5 (age 9-10) cod
+cod5det <- ms_detaileddiet %>%
+  filter(species=="North_atl_cod",
+         agecl==5)
 
 # how many polygons included? how many layers per polygon have data?
+cod5det %>% group_by(time.days, polygon, layer) %>% tally()
 
 # are there differences by layer within a polygon?
+#    polygons 12 and 19 are on the slope, have most layers with agecl 5 cod
+cod5det1219 <- cod5det %>%
+  filter(polygon %in% c(12,19))
 
+plist2 = lapply(split(cod5det1219, cod5det1219$polygon), function(d) {
+  ggplot(d, aes(time.days/365, atoutput, fill=prey)) + 
+    geom_bar(stat = "identity") +
+    facet_wrap(polygon~layer) +
+    xlab("year") +
+    ylab("diet tons") +
+    theme_tufte() +
+    theme(legend.position="bottom")
+})
+```
+
+### Diets by polygon and layer {.tabset}
+
+#### Cod agecl 5 polygon 12
+```{r, fig.width=9, fig.height=9}
+plist2$'12'
+```
+
+#### Cod agecl 5 polygon 19
+```{r, fig.width=9, fig.height=9}
+plist2$'19'
+```
+
+```{r detdiet-vis2}
 # are there differences by polygon? (aggregate layers)
+cod5det_agg <- cod5det %>%
+  group_by(species, agecl, time.days, polygon, prey) %>%
+  summarise(polytons = sum(atoutput))
+
+ggplot(cod5det_agg, aes(time.days/365, polytons, fill=prey)) + 
+    geom_bar(stat = "identity") +
+    facet_wrap(polygon) +
+    xlab("year") +
+    ylab("diet tons") +
+    theme_tufte() +
+    theme(legend.position="bottom")
 

### 2020-09-14 NOBAmsdiets.Rmd
compare polygon/layer diet, check global

@@
 
+```{r}
+# read in the saved file
+source(here("config", "NOBA2config.R"))
+
+ms_detaileddiet <- readRDS(file.path(d.name, paste0(scenario.name, "ms_detaileddiet.rds")))
+
+```
+
 ```{r detdiet-vis1}
@@
 # how many polygons included? how many layers per polygon have data?
-cod5det %>% group_by(time.days, polygon, layer) %>% tally()
+#cod5det %>% group_by(time.days, polygon, layer) %>% tally()
+# why are layers -1 and 0 included (must have trace data if 0 diet was eliminated)
 
@@
 cod5det1219 <- cod5det %>%
-  filter(polygon %in% c(12,19))
+  filter(polygon %in% c(1,6,7,12,19,25,29,37))
 
@@
 
-### Diets by polygon and layer {.tabset}
+As suspected, diet composition varies greatly by model area and layer.
+
+### Consumption by polygon and layer {.tabset}
+
+#### Cod agecl 5 polygon 1
+```{r, fig.width=9, fig.height=9}
+plist2$'1'
+```
+
+#### Cod agecl 5 polygon 6
+```{r, fig.width=9, fig.height=9}
+plist2$'6'
+```
+
+#### Cod agecl 5 polygon 7
+```{r, fig.width=9, fig.height=9}
+plist2$'7'
+```
 
@@
 
-```{r detdiet-vis2}
+#### Cod agecl 5 polygon 25
+```{r, fig.width=9, fig.height=9}
+plist2$'25'
+```
+
+#### Cod agecl 5 polygon 29
+```{r, fig.width=9, fig.height=9}
+plist2$'29'
+```
+
+#### Cod agecl 5 polygon 37
+```{r, fig.width=9, fig.height=9}
+plist2$'37'
+```
+
+```{r detdiet-poly}
 # are there differences by polygon? (aggregate layers)
-cod5det_agg <- cod5det %>%
+coddet_agg <- as.data.frame(
+  ms_detaileddiet %>%
+  filter(species=="North_atl_cod") %>%
   group_by(species, agecl, time.days, polygon, prey) %>%
   summarise(polytons = sum(atoutput))
+)
 
-ggplot(cod5det_agg, aes(time.days/365, polytons, fill=prey)) + 
+coddet_agg2 <- atlantisom::aggregateData(dat = ms_detaileddiet, 
+                                        time = unique(ms_detaileddiet$time.days), 
+                                        species = "North_atl_cod", 
+                                        boxes = unique(ms_detaileddiet$polygon),
+  keepColumns = c("species", "agecl", "polygon", "time.days", "prey"))
+
+names(coddet_agg2)[names(coddet_agg2)=="numAtAge"] <- "polytons"
+
+plist3 = lapply(split(coddet_agg2, coddet_agg2$agecl), function(d) {
+  ggplot(d, aes(time.days/365, polytons, fill=prey)) + 
     geom_bar(stat = "identity") +
-    facet_wrap(polygon) +
+    facet_wrap(species~polygon) +
     xlab("year") +
@@
     theme(legend.position="bottom")
+})
 
+```
+
+#### Cod agecl 5 all polygons
+```{r, fig.width=9, fig.height=9}
+plist3$'5'
+```
+
+#### Cod agecl 10 all polygons
+```{r, fig.width=9, fig.height=9}
+plist3$'10'
+```
+
+### {-}
+
+Aggregate across layers and polygons, compare to aggreate diet comp output (proportion) and biomass_eaten estimate:
+
+```{r detdiet-all, fig.width=9, fig.height=9}
+# are there differences by polygon? (aggregate layers)
+coddet_aggall <- as.data.frame(
+  coddet_agg %>%
+  filter(species=="North_atl_cod") %>%
+  group_by(species, agecl, time.days, prey) %>%
+  summarise(tottons = sum(polytons))
+)
+
+  ggplot(coddet_aggall, aes(time.days/365, tottons, fill=prey)) + 
+    #geom_bar(stat = "identity") +
+    geom_bar(position = "fill", stat = "identity") +
+    facet_wrap(species~agecl) +
+    xlab("year") +
+    ylab("diet tons") +
+    theme_tufte() +
+    theme(legend.position="bottom")
 ```

### 2020-09-15 NOBAmsdiets.Rmd
compare with biomass_eaten

@@
 
-Aggregate across layers and polygons, compare to aggreate diet comp output (proportion) and biomass_eaten estimate:
+Aggregate detailed diet across layers and polygons, compare to aggreate diet comp output (proportion). This looks identical to the DietCheck output:
 
@@
 ```
+
+This is detailed diet aggregated across layers and polygons in absolute tons consumed by age class (y axes are different--most consumption from age class 1):
+
+```{r, fig.width=9, fig.height=9}
+
+  ggplot(coddet_aggall, aes(time.days/365, tottons, fill=prey)) + 
+    geom_bar(stat = "identity") +
+    #geom_bar(position = "fill", stat = "identity") +
+    facet_wrap(species~agecl, scales = "free_y") +
+    xlab("year") +
+    ylab("diet tons") +
+    theme_tufte() +
+
+    theme(legend.position="bottom")
+```
+
+This is the output of biomass_eaten in absolute tons--a short cut that applies global diet comp to polygon-specific tons consumed.
+First we compare diet by polygon for age class 5 as above.
+As expected, the diet comp doesn't match DetailedDietCheck by polygon, but the amounts are also vastly different by polygon:
+
+```{r, fig.width=9, fig.height=9}
+
+NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+
+codbioeaten <- NOBAom_ms$truebioeaten_ss %>%
+  filter(species=="North_atl_cod")
+
+plist4 = lapply(split(codbioeaten, codbioeaten$agecl), function(d) {
+  ggplot(d, aes(time.days/365, bio_eaten, fill=prey)) + 
+    geom_bar(stat = "identity") +
+    facet_wrap(species~polygon) +
+    xlab("year") +
+    ylab("diet tons") +
+    theme_tufte() +
+    theme(legend.position="bottom")
+})
+
+plist4$'5'
+
+```
+
+Aggregate proportion from biomass_eaten should match, because diet proportions come from DietCheck:
+
+```{r, fig.width=9, fig.height=9}
+
+ggplot(codbioeaten, aes(time.days/365, bio_eaten, fill=prey)) + 
+    #geom_bar(stat = "identity") +
+    geom_bar(position = "fill", stat = "identity") +
+    facet_wrap(species~agecl) +
+    xlab("year") +
+    ylab("diet tons") +
+    theme_tufte() +
+    theme(legend.position="bottom")
+
+```
+
+Total biomass consumed from bio_eaten calculation is about a factor of 10 higher than from DetailedDietCheck. Patterns look a bit different too:
+
+```{r fig.width=9, fig.height=9}
+
+ggplot(codbioeaten, aes(time.days/365, bio_eaten, fill=prey)) + 
+    geom_bar(stat = "identity") +
+    #geom_bar(position = "fill", stat = "identity") +
+    facet_wrap(species~agecl, scales = "free_y") +
+    xlab("year") +
+    ylab("diet tons") +
+    theme_tufte() +
+    theme(legend.position="bottom")
+
+```
+
+Bottom line: biomass_eaten won't work for our purposes in generating truth for survey data that will select polygons for sampling.
+Also, I'm not sure why consumption estimates are an order of magnitude higher.
+
+So I'll make some functions for diet survey sampling that work on the beastly DetailedDietCheck.
+I won't try to bring this into run_truth due to the size of the file, and I will remove biomass_eaten from the run_truth output until I can troubleshoot the estimates.
 

### 2020-09-15 NOBAmsdiets.Rmd
refer to archived file

@@
 # run again 3 August to update biomass_eaten 
+# and output archived with _bio_eaten suffix
+# run again 15 Sept to get rid of biomass_eaten
 
@@
 # run again 3 August to test new wrapper that should save biomass_eaten
-NOBAom_ms <- om_species(sppsubset$Name, NOBAom, diet=TRUE)
+#NOBAom_ms <- om_species(sppsubset$Name, NOBAom, diet=TRUE)
+# and output archived with _bio_eaten suffix
+# run again 15 Sept to get rid of biomass eaten
+NOBAom_ms <- om_species(sppsubset$Name, NOBAom)
 
@@
 
-NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+NOBAom_ms_be <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss_bio_eaten.rds")))
 
-codbioeaten <- NOBAom_ms$truebioeaten_ss %>%
+codbioeaten <- NOBAom_ms_be$truebioeaten_ss %>%
   filter(species=="North_atl_cod")

### 2020-09-17 config/mssurvey_fall.R
add survey name to use multiple surveys in wrapper

@@ -3,6 +3,9 @@
 # Also now happens in "spring" and "fall"
 # Need to define survey season, area, efficiency, selectivity
 
+# Survey name
+survey.name="BTS_fall_allbox_effic1"
+
 #Atlantis model timestep corresponding to the true output--now from census_spec.R
 timestep <- stepperyr #5
 
@@ -43,7 +46,7 @@ survselex <- data.frame(species=rep(survspp, n_annages), #
 
 # effective sample size needed for sample_fish
 # this effective N is high but not equal to total for numerous groups
-surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+surveffN <- data.frame(species=survspp, effN=rep(100000, length(survspp)))
 
 # survey index cv needed for sample_survey_xxx
 # cv = 0.1

### 2020-09-17 config/mssurvey_spring.R

@@ -3,6 +3,9 @@
 # Also now happens in "spring" and "fall"
 # Need to define survey season, area, efficiency, selectivity
 
+# Survey name
+survey.name="BTS_spring_allbox_effic1"
+
 #Atlantis model timestep corresponding to the true output--now from census_spec.R
 timestep <- stepperyr #5
 
@@ -43,7 +46,7 @@ survselex <- data.frame(species=rep(survspp, n_annages), #
 
 # effective sample size needed for sample_fish
 # this effective N is high but not equal to total for numerous groups
-surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+surveffN <- data.frame(species=survspp, effN=rep(100000, length(survspp)))
 
 # survey index cv needed for sample_survey_xxx
 # cv = 0.1

### 2020-09-17 atlantisom_landingpage.Rmd
remove authors, add credits

@@
 title: "atlantisom: simulated ecosystem datasets for assessment model performance testing"
-author: "Sarah Gaichas, Christine Stawitz, Kelli Johnson, Alexander Keth, Allan Hicks, Sean Lucey, Emma Hodgson, and Gavin Fay"
 date: "`r Sys.Date()`"
@@
 Substantial progress since then was made possible by a NOAA NMFS International Fellowship and the Institute of Marine Research in Bergen, Norway.
+
+## Credits
+
+Sarah Gaichas, Christine Stawitz, Kelli Johnson, Alexander Keth, Allan Hicks, Sean Lucey, Emma Hodgson, and Gavin Fay
 

### 2020-09-17 NOBAmsdiets.Rmd
test multi-survey wrapper

@@
 
-## Generating the dataset 
-*(PLACEHOLDER; not done yet June 19 2020)*
+## Generating the dataset (September 2020)
 
@@
 
-```{r, code = readLines("./config/NOBA2Config.R"), eval=F}
+```{r, code = readLines("./config/NOBA2Config.R")}
 ```
@@
 
-```{r, code = readLines("./config/omdimensions.R"), eval=F}
+```{r, code = readLines("./config/omdimensions.R")}
 ```
 
-`mssurvey_spring.R` and `mssurvey_fall.R` configure the fishery independent surveys:
+`mssurvey_spring.R` and `mssurvey_fall.R` configure the fishery independent surveys (in this test, surveys sample all model polygons in all years and have efficiency of 1 for all species, with no size selectivity):
 
-```{r, code = readLines("./config/mssurvey_spring.R"), eval=F}
+```{r, code = readLines("./config/mssurvey_spring.R")}
 ```
 
-```{r, code = readLines("./config/mssurvey_fall.R"), eval=F}
+```{r, code = readLines("./config/mssurvey_fall.R")}
 ```
@@
 
-```{r, code = readLines("./config/msfishery.R"), eval=F}
+```{r, code = readLines("./config/msfishery.R")}
 ```
@@
 
-NOBAom_ms_ind <- om_index(usersurvey = here("config/mssurvey.R"), 
+# new wrapper for multiple surveys saves survey output separately
+# but returns a list with all surveys, not ideal; make consistent
+
+NOBAom_ms_ind <- om_index(usersurvey = c(here("config/mssurvey_spring.R"), 
+                                         here("config/mssurvey_fall.R")), 
                            userfishery = here("config/msfishery.R"),
@@
 
-NOBAom_ms_comp <- om_comps(usersurvey = here("config/mssurvey.R"), 
+NOBAom_ms_comp <- om_comps(usersurvey = c(here("config/mssurvey_spring.R"), 
+                                         here("config/mssurvey_fall.R")), 
                            userfishery = here("config/msfishery.R"),
@@
 
-omlist_ss <- NOBAom_cod
+omlist_ss <- NOBAom_ms
 source(here("config/omdimensions.R"))
 
-#read time series data
-survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name, "surveyB.rds")))
+#read time series data--now needs survey name from config files
+survname <- c("BTS_spring_allbox_effic1", "BTS_fall_allbox_effic1")
 
-survObsBiom <- survObsBiom[[1]]
+#need plotting/visualization wrappers
+#survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name, "surveyB.rds")))
+
+survObsBiom <- NOBAom_ms$survObsBiom[[survname]]
 

### 2020-09-17 NOBAmsdiets.Rmd
omdimensions is wrong problem found still needs fix

@@
 
-```{r, code = readLines("./config/NOBA2Config.R")}
+```{r, code = readLines("./config/NOBA2Config.R"), eval=F}
 ```
@@
 
-```{r, code = readLines("./config/omdimensions.R")}
+```{r, code = readLines("./config/omdimensions.R"), eval=F}
 ```
@@
 
-```{r, code = readLines("./config/mssurvey_spring.R")}
+```{r, code = readLines("./config/mssurvey_spring.R"), eval=F}
 ```
 
-```{r, code = readLines("./config/mssurvey_fall.R")}
+```{r, code = readLines("./config/mssurvey_fall.R"), eval=F}
 ```
@@
 
-```{r, code = readLines("./config/msfishery.R")}
+```{r, code = readLines("./config/msfishery.R"), eval=F}
 ```
@@
 #read time series data--now needs survey name from config files
-survname <- c("BTS_spring_allbox_effic1", "BTS_fall_allbox_effic1")
 
@@
 
-survObsBiom <- NOBAom_ms$survObsBiom[[survname]]
+survObsBiom <- NOBAom_ms_ind$survObsBiom
 
-plotB <-ggplot() +
-  geom_line(data=survObsBiom, aes(x=time/stepperyr,y=atoutput, color="survey Biomass"), 
+plotB <- function(dat){
+
+    ggplot() +
+    geom_line(data=dat, aes(x=time/stepperyr,y=atoutput, color="survey Biomass"), 
               alpha = 10/10) +
@@
     theme(legend.position = "top") +
-  labs(colour=scenario.name)
-
-plotB +
+    labs(colour=scenario.name) +
     facet_wrap(~species, scales="free")
 
+}
+
+# why is herring missing from both?
+herring <- NOBAom_ms$truebio_ss %>%
+  filter(species %in% c("Norwegian_ssh")) %>%
+  group_by(species, time) %>%
+  summarise(totB = sum(atoutput))
+
+# problem--omdimensions is doing species age accounting wrong
+# species alphabetical pulled separately but ages not
+# pull species name, n age classes and n ages all at once to avoid bad sorting
+```
+
+##### `r names(survObsBiom)[1]`
+``` {r}
+map(survObsBiom$BTS_spring_allbox_effic1, plotB)[[1]]
+```
+
+##### `r names(survObsBiom)[2]`
+```{r}
+map(survObsBiom$BTS_fall_allbox_effic1, plotB)[[1]]
 ```

### 2020-09-17 config/omdimensions.R

@@ -16,7 +16,7 @@ fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$t
 # survey selectivity (agecl based)
 # should return all age classes fully sampled (Atlantis output is 10 age groups per spp)
 n_age_classes <- omlist_ss$funct.group_ss$NumCohorts
-# changed below for multiple species
+# changed below for multiple species WRONG not in species order!! PULL ALL AT ONCE
 age_classes <- sapply(n_age_classes, seq)
 names(age_classes)<-survspp
 

### 2020-10-05 NOBAmsdiets.Rmd
start reading in multiple surveys with names WIP

@@
 #### Biomass index
-```{r bioind-1, eval=FALSE}
+```{r bioind-1}
 
-omlist_ss <- NOBAom_ms
+omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
 source(here("config/omdimensions.R"))
@@
 
-survObsBiom <- NOBAom_ms_ind$survObsBiom
+# object in memory is no longer the same as saved files, need to read back into same configuration
+# list of files in directory with surveyB.rds ending
+# append to survObsBiom
+survs <- list.files(path=d.name, pattern = "*surveyB.rds", full.names = TRUE)
+survey.name <-  
+survObsBiom <- lapply(survs, readRDS)
+
+#survObsBiom <- NOBAom_ms_ind$survObsBiom
+
+# survey.name="BTS_fall_allbox_effic1"
+# fallsurv <- readRDS(file.path(d.name, paste0(scenario.name, "_",survey.name,"surveyB.rds")))
 
@@
 # pull species name, n age classes and n ages all at once to avoid bad sorting
+# also fix same problem in om_index with wtatage hack!
 ```
@@
 #### Catch time series
-```{r catchind-1, eval=FALSE}
+```{r catchind-1}
 #read time series data

### 2020-10-06 NOBAmsdiets.Rmd
add survey names to saved file when reading in

@@
 # append to survObsBiom
+# apply survey name from filename
 survs <- list.files(path=d.name, pattern = "*surveyB.rds", full.names = TRUE)
-survey.name <-  
+survey.name <-  str_match(survs, paste0(scenario.name,"_\\s*(.*?)\\s*surveyB.rds"))[,2]
 survObsBiom <- lapply(survs, readRDS)
+names(survObsBiom) <- survey.name
 
@@
 # why is herring missing from both?
-herring <- NOBAom_ms$truebio_ss %>%
-  filter(species %in% c("Norwegian_ssh")) %>%
-  group_by(species, time) %>%
-  summarise(totB = sum(atoutput))
+# herring <- NOBAom_ms$truebio_ss %>%
+#   filter(species %in% c("Norwegian_ssh")) %>%
+#   group_by(species, time) %>%
+#   summarise(totB = sum(atoutput))
 

### 2020-10-20 NOBAmsdiets.Rmd
fns readsurvs, lencomp; add survey lencomp plots

@@
 
-### Using the wrappers (now updated for annual age)
+### Using the wrappers (now updated for annual age, multiple surveys, and multispecies)
 
 All the config files go in the config folder and we'll try running this:
+
+Run notes: om_comps output took 4 hours for this group of 11 species with 2 surveys and 1 fishery.
 
@@
 
+Now need a function to read survey-specific output files back into a single object with named surveys.
+
+```{r readsurvs-function}
+
+readsurvs <- function(path, type){
+
+  datlook <- data.frame(dattype = c('survB', 'survAge', 'survLen', 'survWtage', 'survAnnAge', 'survAnnWtage'),
+                        pattern = c("*surveyB.rds", "*survObsAgeComp.rds", "*survObsLenComp.rds", "*survObsWtAtAge.rds",
+                                    "*survObsFullAgeComp.rds", "*survObsFullWtAtAge.rds"),
+                        datname = c('survObsBiom', 'age_comp_data', 'len_comp_data', 'wtage', 'annage_comp_data', 'wtannage'))
+
+  survs <- list.files(path=path, pattern = as.character(datlook$pattern[datlook$dattype %in% type]), full.names = TRUE)
+
+  survey.name <-  str_match(survs, paste0(scenario.name,"_\\s*(.*?)\\s",datlook$pattern[datlook$dattype==type]))[,2]
+
+  datname <- lapply(survs, readRDS)
+
+  names(datname) <- survey.name
+
+  return(datname)
+}
+
+```
+
+Testing--read in the data for plotting:
+
+```{r test-readurvs}
+
+survObsBiom <- readsurvs(d.name, 'survB')
+age_comp_data <- readsurvs(d.name, 'survAge')
+len_comp_data <- readsurvs(d.name, 'survLen')
+wtage <- readsurvs(d.name, 'survWtage')
+annage_comp_data <- readsurvs(d.name, 'survAnnAge')
+annage_wtage <- readsurvs(d.name, 'survAnnWtage')
+
+```
+
+Also need new plot functions and updated input file-writing functions that expect this structure.
+
 ### Wrapper test results {.tabset}
@@
 # apply survey name from filename
-survs <- list.files(path=d.name, pattern = "*surveyB.rds", full.names = TRUE)
-survey.name <-  str_match(survs, paste0(scenario.name,"_\\s*(.*?)\\s*surveyB.rds"))[,2]
-survObsBiom <- lapply(survs, readRDS)
-names(survObsBiom) <- survey.name
+# survs <- list.files(path=d.name, pattern = "*surveyB.rds", full.names = TRUE)
+# survey.name <-  str_match(survs, paste0(scenario.name,"_\\s*(.*?)\\s*surveyB.rds"))[,2]
+# survObsBiom <- lapply(survs, readRDS)
+# names(survObsBiom) <- survey.name
 
@@
 #### Survey length composition
-```{r lencomp-1, eval=FALSE}
+```{r lencomp-1}
 
 #length comps
-len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsLenComp.rds")))
+#len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsLenComp.rds")))
+
+# now using readsurvs function above
+# survs <- list.files(path=d.name, pattern = "*survObsLenComp.rds", full.names = TRUE)
+# survey.name <-  str_match(survs, paste0(scenario.name,"_\\s*(.*?)\\s*survObsLenComp.rds"))[,2]
+# len_comp_data <- lapply(survs, readRDS)
+# names(len_comp_data) <- survey.name
+
 fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))
 
-len_comp_data <- len_comp_data[[1]]
+#len_comp_data <- len_comp_data[[1]]
 fish_len_comp_data <- fish_len_comp_data[[1]]
@@
 
-len <- filter(len_comp_data, time %in% c(55:175))
-
-  lfplot <- ggplot(len, aes(upper.bins)) +
+plotlen <- function(dat){
+
+    ggplot(dat, aes(upper.bins)) +
     geom_bar(aes(weight = atoutput)) +
@@
     labs(subtitle = paste(scenario.name,
-                          len$species))
-
-  lfplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
-
+                          dat$species)) +
+    facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+
+}
+
+# len <- filter(len_comp_data, time %in% c(55:175))
+# 
+#   lfplot <- ggplot(len, aes(upper.bins)) +
+#     geom_bar(aes(weight = atoutput)) +
+#     theme_tufte() +
+#     labs(subtitle = paste(scenario.name,
+#                           len$species))
+#   
+#   lfplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+
+```
+
+##### `r names(len_comp_data)[1]`
+``` {r, fig.show='hold'}
+lcompsub <- as.data.frame(len_comp_data$BTS_fall_allbox_effic1[1]) %>% filter(time %in% c(55:175)) %>%
+  group_by(species) %>%
+  group_map(~ plotlen(.x), keep = TRUE)
+
+for(i in 1:length(lcompsub)) {
+  print(lcompsub[[i]])
+}
+
+#map(len_comp_data$BTS_fall_allbox_effic1 %>% keep(time %in% c(55:175)), plotlen)[[1]]
+```
+
+##### `r names(len_comp_data)[2]`
+``` {r}
+lcompsub <- as.data.frame(len_comp_data$BTS_spring_allbox_effic1[1]) %>% filter(time %in% c(55:175)) %>%
+  group_by(species) %>%
+  group_map(~ plotlen(.x), keep = TRUE)
+
+for(i in 1:length(lcompsub)) {
+  print(lcompsub[[i]])
+}
 ```

### 2020-10-21 NOBAmsdiets.Rmd
corrected labels, add fishery lenage, nageplot

@@
 ``` {r}
-map(survObsBiom$BTS_spring_allbox_effic1, plotB)[[1]]
+map(survObsBiom$BTS_fall_allbox_effic1, plotB)[[1]]
 ```
@@
 ```{r}
-map(survObsBiom$BTS_fall_allbox_effic1, plotB)[[1]]
+map(survObsBiom$BTS_spring_allbox_effic1, plotB)[[1]]
 ```
@@
 
-#read in comp data
-age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
-age_comp_data <- age_comp_data[[1]]
-
-Natage <- filter(age_comp_data, time %in% c(150:270))
-
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
+#read in comp data--now done above with function
+# age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
+# age_comp_data <- age_comp_data[[1]]
+
+#Natage <- filter(age_comp_data, time %in% c(150:270))
+
+Natageplot <- <- function(dat){
+  ggplot(dat, aes(x=agecl, y=atoutput)) +
     geom_point() +
@@
     labs(subtitle = paste(scenario.name,
-                          Natage$species))
-
-  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+                          dat$species)) + 
+    facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+}
+
+#Natageplot 
 ```
@@
 wageplot + facet_wrap(c("species"), scales="free_y")
+```
+
+#### Fishery length composition
+``` {r}
+lcompsub <- as.data.frame(fish_len_comp_data) %>% filter(time %in% c(55:175)) %>%
+  group_by(species) %>%
+  group_map(~ plotlen(.x), keep = TRUE)
+
+for(i in 1:length(lcompsub)) {
+  print(lcompsub[[i]])
+}
 ```

### 2020-10-21 NOBAmsdiets.Rmd
agecomp and wtage--Blue whiting age comps problem?

@@
 
-Also need new plot functions and updated input file-writing functions that expect this structure.
+Also need new plot functions and updated input file-writing functions that expect this structure. Preliminary plot functions below.
 
@@
 #### Survey age composition (age classes)
-```{r agecomp-1, eval=FALSE}
+```{r agecomp-1}
 
@@
 
-Natageplot <- <- function(dat){
+Natageplot <- function(dat){
   ggplot(dat, aes(x=agecl, y=atoutput)) +
@@
 
+##### `r names(age_comp_data)[1]`
+``` {r}
+acompsub <- as.data.frame(age_comp_data$BTS_spring_allbox_effic1[1]) %>% filter(time %in% c(150:270)) %>%
+  group_by(species) %>%
+  group_map(~ Natageplot(.x), keep = TRUE)
+
+for(i in 1:length(acompsub)) {
+  print(acompsub[[i]])
+}
+```
+
+##### `r names(age_comp_data)[2]`
+``` {r}
+acompsub <- as.data.frame(age_comp_data$BTS_spring_allbox_effic1[1]) %>% filter(time %in% c(150:270)) %>%
+  group_by(species) %>%
+  group_map(~ Natageplot(.x), keep = TRUE)
+
+for(i in 1:length(acompsub)) {
+  print(acompsub[[i]])
+}
+```
+
 #### Survey weight at age (age classes)
-```{r wtageclass-1, eval=FALSE}
-
-wtage <- readRDS(file.path(d.name, paste0(scenario.name, "survObsWtAtAge.rds")))
-wtage <- wtage[[1]]
-
-wageplot <- ggplot(wtage, aes(time, atoutput)) +
+```{r wtageclass-1}
+
+# now done with function above
+# wtage <- readRDS(file.path(d.name, paste0(scenario.name, "survObsWtAtAge.rds")))
+# wtage <- wtage[[1]]
+
+wageplot <- function(dat){
+  ggplot(dat, aes(time, atoutput)) +
     geom_line(aes(colour = factor(agecl))) +
@@
     ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
+    labs(subtitle = paste0(scenario.name)) +
+    facet_wrap(c("species"), scales="free_y")
+}
+
+#wageplot 
+```
+
+##### `r names(wtage)[1]`
+``` {r}
+map(wtage$BTS_fall_allbox_effic1, wageplot)[[1]]
+```
+
+##### `r names(wtage)[2]`
+``` {r}
+map(wtage$BTS_spring_allbox_effic1, wageplot)[[1]]
 ```
@@
 #### Fishery age composition (age classes)
-```{r fishagecomp-1, eval=FALSE}
+```{r fishagecomp-1}
 
@@
 
-Natage <- filter(fish_age_comp, time %in% c(30:53))
-
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Natage$species))
-
-  Natageplot + facet_wrap(~time, ncol=6, scales="free_y")
+acompsub <- as.data.frame(age_comp_data$BTS_spring_allbox_effic1[1]) %>% filter(time %in% c(150:270)) %>%
+  group_by(species) %>%
+  group_map(~ Natageplot(.x), keep = TRUE)
+
+for(i in 1:length(acompsub)) {
+  print(acompsub[[i]])
+}
+
 ```

### 2020-10-22 NOBAmsdiets.Rmd
add fish wt age, ann age, ann wt age

@@
 
+#### Fishery weight at age (age classes)
+```{r}
+fishwtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsWtAtAge.rds")))
+fishwtage <- fishwtage[[1]]
+
+wageplot(fishwtage)
+
+```
+
 ### {-}
@@
 #### Survey age composition (annual ages)
-```{r plot-annagecomp, eval=FALSE}
-#read in comp data
-annage_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
-annage_comp_data <- annage_comp_data[[1]]
-
-#fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
-#fish_age_comp <- fish_age_comp[[1]]
-
-#add this to om_indices function so that this has years when read in
-#fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
-
-Natage <- filter(annage_comp_data, time %in% c(150:270))
-
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Natage$species))
-
-  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+```{r plot-annagecomp, results="asis"}
+
+# annage_comp_data read in above, apply plot function, quit hardcoding survey names
+
+for(s in names(annage_comp_data)){
+  cat("  \n##### ",  s,"  \n")
+  acompsub <- as.data.frame(annage_comp_data[[s]][[1]]) %>% filter(time %in% c(150:270)) %>%
+    group_by(species) %>%
+    group_map(~ Natageplot(.x), keep = TRUE)
+
+  for(i in 1:length(acompsub)) {
+    print(acompsub[[i]])
+  }
+  cat("  \n")
+}
 ```
@@
 #### Survey iterpolated weight at age (annual ages) 
-```{r plot-annwtage, eval=FALSE}
-
-wtage <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullWtAtAge.rds")))
-wtage <- wtage[[1]] #this still has second list component, diagnostic plot
-wtage <- wtage[[1]]
-
-wageplot <- ggplot(wtage, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (5 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-
+```{r plot-annwtage, results='asis'}
+
+# annage_wtage read in above 
+# wtage <- wtage[[1]] #this still has second list component, diagnostic plot
+# wtage <- wtage[[1]]
+
+for(s in names(annage_wtage)){
+  cat("  \n##### ",  s,"  \n")
+  print(wageplot(annage_wtage[[s]][[1]][[1]]))
+  cat("  \n")
+}
 ```
@@
 #### Fishery catch at age (annual ages)
-```{r plot-catchagecomp, eval=FALSE}
+```{r plot-catchagecomp}
 
@@
 
-Catage <- filter(fish_annage_comp, time %in% c(30:53))
-
-Catageplot <- ggplot(Catage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Catage$species))
-
-  Catageplot + facet_wrap(~time, ncol=6, scales="free_y")
+acompsub <- as.data.frame(fish_annage_comp) %>% filter(time %in% c(30:54)) %>%
+  group_by(species) %>%
+  group_map(~ Natageplot(.x), keep = TRUE)
+
+for(i in 1:length(acompsub)) {
+  print(acompsub[[i]])
+}
 ```
@@
 #### Fishery iterpolated weight at age (annual ages) 
-```{r plot-fishannwtage, eval=FALSE}
-
-wtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullWtAtAge.rds")))
-wtage <- wtage[[1]] #this still has second list component, diagnostic plot
-wtage <- wtage[[1]]
-
-wageplot <- ggplot(wtage, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (5 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
+```{r plot-fishannwtage}
+
+fish_annage_wtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullWtAtAge.rds")))
+
+wageplot(fish_annage_wtage[[1]][[1]])
 

### 2020-10-22 docs/index.Rmd
testing diet functions, plotting output
compare with biomass_eaten
add fish wt age, ann age, ann wt age

@@
 <a href="https://sgaichas.github.io/poseidon-dev/atlantisom_landingpage.html" target="_blank">Draft atlantisom landing page for review, June 9, 2020</a>
+
+### Full age structured data for multispecies modeling, June-October 2020
+
+<a href="https://sgaichas.github.io/poseidon-dev/NOBAmsdiets.html" target="_blank">Explore NOBA diet comps and multi-survey multispecies wrappers, October 21, 2020</a>
+
+Including detailed diet exploration, multiple surveys, survey function bug fixes for species with < 10 cohorts.
 

### 2020-10-22 NOBAmsdiets.Rmd
correct fishery age comp dataset, time still ugly fix elsewhere

@@
 #add this to om_indices function so that this has years when read in
-fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))
+#fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))
 
@@
 #add this to om_indices function so that this has years when read in
-fish_age_comp$time <- fish_age_comp$time/fstepperyr
-
-acompsub <- as.data.frame(age_comp_data$BTS_spring_allbox_effic1[1]) %>% filter(time %in% c(150:270)) %>%
+#fish_age_comp$time <- fish_age_comp$time/fstepperyr
+
+acompsub <- as.data.frame(fish_age_comp) %>% filter(time %in% c(150:270)) %>%
   group_by(species) %>%

### 2020-10-22 msSurveysTest.Rmd
initial commit survey specifications

---
title: "Multispecies survey specifications"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)  
library(atlantisom)
library(ggthemes)

```

## Species surveyed

Our initial species selection includes 11 single species groups from the NOBA model. We looked at diet overlap [here](https://sgaichas.github.io/poseidon-dev/NOBAmsdiets.html). Below we will double check survey census outputs against the true values, and then test some survey specifications that may be reasonable for multispecies model testing. Finally we will apply survey functions to the detailed diet data.

```{r spp-table}

fgs <- load_fgs(here("atlantisoutput", "NOBA_March_2020"), "nordic_groups_v04.csv")

lname <- data.frame(Latin = c("*Hippoglossoides platessoides*",
                              "*Reinhardtius hippoglossoides*",
                              "*Scomber scombrus*",
                              "*Melongrammus aeglefinus*",
                              "*Pollachius virens*",
                              "*Sebastes mentella*",
                              "*Micromesistius poutassou*",
                              "*Clupea harengus*",
                              "*Gadus morhua*",
                              "*Boreogadus saida*",
                              "*Mallotus villosus*"),
                    Code = c("LRD", "GRH", "MAC", "HAD", "SAI", "RED", 
                             "BWH", "SSH", "NCO", "PCO", "CAP")
)

sppsubset <- merge(fgs, lname, all.y = TRUE)
spptable <- sppsubset %>% 
  arrange(Index) %>%
  select(Name, Long.Name, Latin)

knitr::kable(spptable, col.names = c("Model name", "Full name", "Latin name"))

```

## Generating the census dataset (October 2020)

Note that survey config files must now name the survey.

### Unchanging config files

`NOBA2config.R` looks like this (adjusted from Alfonso's original):

```{r, code = readLines("./config/NOBA2Config.R"), eval=F}
```

`omdimensions.R` standardizes timesteps, etc. (this is part of atlantisom and should not need to be changed by the user):

```{r, code = readLines("./config/omdimensions.R"), eval=F}
```

### Change these survey and fishery config files

`mssurvey_spring.R` and `mssurvey_fall.R` configure the fishery independent surveys (in this census test, surveys sample all model polygons in all years and have efficiency of 1 for all species, with no size selectivity):

```{r, code = readLines("./config/mssurvey_spring.R"), eval=F}
```

```{r, code = readLines("./config/mssurvey_fall.R"), eval=F}
```

`msfishery.R` configures the fishery dependent data:

```{r, code = readLines("./config/msfishery.R"), eval=F}
```

These survey and fishery configurations were used [previously](https://sgaichas.github.io/poseidon-dev/NOBAmsdiets.html) to generate saved census survey output.

### Read in previously generated survey and fishery outputs and plot them against true outputs

Using `read_savedsurvs()` we get back the same objects generated by the wrapper functions `om_index()` and `om_comps()`.

```{r test-readurvs}

source(here("config", "NOBA2config.R"))
omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
source(here("config/omdimensions.R"))

survObsBiom <- read_savedsurvs(d.name, 'survB')
age_comp_data <- read_savedsurvs(d.name, 'survAge')
len_comp_data <- read_savedsurvs(d.name, 'survLen')
wtage <- read_savedsurvs(d.name, 'survWtage')
annage_comp_data <- read_savedsurvs(d.name, 'survAnnAge')
annage_wtage <- read_savedsurvs(d.name, 'survAnnWtage')

# no function needed for saved fishery output as long as there is only one fishery specification file

catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
catchbio_ss <- catchbio_ss[[1]] #in same time unit as surveys

fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
fish_age_comp <- fish_age_comp[[1]]

#add this to om_indices function so that this has years when read in
#fish_age_comp$time <- fish_age_comp$time/fstepperyr

fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))

#len_comp_data <- len_comp_data[[1]]
fish_len_comp_data <- fish_len_comp_data[[1]]

#add this to om_indices function so that this has years when read in
#fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))

fishwtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsWtAtAge.rds")))
fishwtage <- fishwtage[[1]] #need to change time units?

fish_annage_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
fish_annage_comp <- fish_annage_comp[[1]]

#add this to om_indices function so that this has years when read in
#fish_annage_comp$time <- fish_annage_comp$time/fstepperyr

fish_annage_wtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullWtAtAge.rds")))
fish_annage_wtage <- fish_annage_wtage[[1]] #need to change time units?

```

My plotting functions: too specialized to include in the package?

```{r plotfuns}

# plot biomass time series facet wrapped by species
plotB <- function(dat){

    ggplot() +
    geom_line(data=dat, aes(x=time/stepperyr,y=atoutput, color="Survey Biomass"), 
              alpha = 10/10) +
    theme_tufte() +
    theme(legend.position = "top") +
    xlab("model year") +
    ylab("tons") +
    labs(colour=scenario.name) +
    facet_wrap(~species, scales="free")

}

# make a catch series function that can be split by fleet? this doesnt
# also note different time (days) from model timestep in all other output
plotC <- function(dat){

    ggplot() +
    geom_line(data=dat, aes(x=time/365,y=atoutput, color="Catch biomass"), 
              alpha = 10/10) +
    theme_tufte() +
    theme(legend.position = "top") +
    xlab("model year") +
    ylab("tons") +
    labs(colour=scenario.name) +
    facet_wrap(~species, scales="free")

}

# plot length frequencies by timestep (one species)
plotlen <- function(dat){

    ggplot(dat, aes(upper.bins)) +
    geom_bar(aes(weight = atoutput)) +
    theme_tufte() +
    xlab("length (cm)") +
    ylab("number") +
    labs(subtitle = paste(scenario.name,
                          dat$species)) +
    facet_wrap(~time/stepperyr, ncol=6, scales="free_y")

}

# plot numbers at age by timestep (one species)
Natageplot <- function(dat){
  ggplot(dat, aes(x=agecl, y=atoutput)) +
    geom_point() +
    theme_tufte() +
    xlab("age/agecl") +
    ylab("number") +
    labs(subtitle = paste(scenario.name,
                          dat$species)) + 
    facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
}

# plot weight at age time series facet wrapped by species
wageplot <- function(dat){
  ggplot(dat, aes(time/stepperyr, atoutput)) +
    geom_line(aes(colour = factor(agecl))) +
    theme_tufte() +
    theme(legend.position = "bottom") +
    xlab("model year") +
    ylab("average individual weight (g)") +
    labs(subtitle = paste0(scenario.name)) +
    facet_wrap(c("species"), scales="free_y")
}

```

### Visualize census surveys {.tabset #census}

We are using the full time series for biomass and weight at age, and a subsample of years 30-53 for compositions.

#### Survey biomass
```{r, results="asis"}

for(s in names(survObsBiom)){
  cat("  \n##### ",  s,"  \n")
  print(plotB(survObsBiom[[s]][[1]]))
  cat("  \n")
}

```

#### Survey length composition
```{r, results="asis"}

for(s in names(len_comp_data)){
  cat("  \n##### ",  s,"  \n")
  lcompsub <- as.data.frame(len_comp_data[[s]][[1]]) %>% filter(time %in% c(150:270)) %>%
    group_by(species) %>%
    group_map(~ plotlen(.x), keep = TRUE)

  for(i in 1:length(lcompsub)) {
    print(lcompsub[[i]])
  }
  cat("  \n")
}

```

#### Survey age class composition (Atlantis age class)
```{r, results="asis"}

for(s in names(age_comp_data)){
  cat("  \n##### ",  s,"  \n")
  acompsub <- as.data.frame(age_comp_data[[s]][[1]]) %>% filter(time %in% c(150:270)) %>%
    group_by(species) %>%
    group_map(~ Natageplot(.x), keep = TRUE)

  for(i in 1:length(acompsub)) {
    print(acompsub[[i]])
  }
  cat("  \n")
}

```

#### Survey age composition (annual ages)
```{r plot-annagecomp, results="asis"}

for(s in names(annage_comp_data)){
  cat("  \n##### ",  s,"  \n")
  acompsub <- as.data.frame(annage_comp_data[[s]][[1]]) %>% filter(time %in% c(150:270)) %>%
    group_by(species) %>%
    group_map(~ Natageplot(.x), keep = TRUE)

  for(i in 1:length(acompsub)) {
    print(acompsub[[i]])
  }
  cat("  \n")
}
```

#### Survey weight at age (Atlantis age class)
```{r, results='asis'}

for(s in names(wtage)){
  cat("  \n##### ",  s,"  \n")
  print(wageplot(wtage[[s]][[1]]))
  cat("  \n")
}

```

#### Survey iterpolated weight at age (annual ages) 
```{r plot-annwtage, fig.height=10, results='asis'}

# annage_wtage read in above 
# annage_wtage <- annage_wtage[[1]] #this still has second list component, diagnostic plot
# annage_wtage <- annage_wtage[[1]] #still in this old version but just removed from function.

for(s in names(annage_wtage)){
  cat("  \n##### ",  s,"  \n")
  print(wageplot(annage_wtage[[s]][[1]][[1]]))
  cat("  \n")
}
```

### {-}

### Visualize perfectly observed fisheries {.tabset}

We are using the full time series for total catch and weight at age, and a subsample of years 30-35 for compositions. Fishery length and age composition is sampled every output timestep.

#### Fishery total catch
```{r}

plotC(catchbio_ss)

```

#### Fishery length composition 
```{r}

lcompsub <- as.data.frame(fish_len_comp_data) %>% filter(time %in% c(150:175)) %>%
  group_by(species) %>%
  group_map(~ plotlen(.x), keep = TRUE)

for(i in 1:length(lcompsub)) {
  print(lcompsub[[i]])
}

```

#### Fishery catch at age (Atlantis age class)
```{r}

acompsub <- as.data.frame(fish_age_comp) %>% filter(time %in% c(150:175)) %>%
  group_by(species) %>%
  group_map(~ Natageplot(.x), keep = TRUE)

for(i in 1:length(acompsub)) {
  print(acompsub[[i]])
}

```

#### Fishery catch at age (annual ages)
```{r}

acompsub <- as.data.frame(fish_annage_comp) %>% filter(time %in% c(150:175)) %>%
  group_by(species) %>%
  group_map(~ Natageplot(.x), keep = TRUE)

for(i in 1:length(acompsub)) {
  print(acompsub[[i]])
}

```

#### Fishery weight at age (Atlantis age class)
```{r}

wageplot(fishwtage)

```

#### Fishery interpolated weight at age (annual ages)
```{r, fig.height=10}

#diagnostic plot component still in this old version but just removed from function.
wageplot(fish_annage_wtage[[1]])

```

### {-}

### 2020-10-23 msSurveysTest.Rmd
add true bio and catch comparisons

@@
 # plot biomass time series facet wrapped by species
-plotB <- function(dat){
+plotB <- function(dat, truedat=NULL){
 
@@
               alpha = 10/10) +
+    {if(!is.null(truedat)) geom_line(data=truedat, aes(x=time/365,y=atoutput, color="True B"), alpha = 3/10)} + 
     theme_tufte() +
@@
 # also note different time (days) from model timestep in all other output
-plotC <- function(dat){
+plotC <- function(dat, truedat=NULL){
 
@@
               alpha = 10/10) +
+    {if(!is.null(truedat)) geom_line(data=truedat, aes(x=time/365,y=atoutput, color="True Catch"), alpha = 3/10)} + 
     theme_tufte() +
@@
 # plot length frequencies by timestep (one species)
-plotlen <- function(dat){
+plotlen <- function(dat, truedat=NULL){
 
@@
 # plot numbers at age by timestep (one species)
-Natageplot <- function(dat){
+Natageplot <- function(dat, truedat=NULL){
   ggplot(dat, aes(x=agecl, y=atoutput)) +
@@
 # plot weight at age time series facet wrapped by species
-wageplot <- function(dat){
+wageplot <- function(dat, truedat=NULL){
   ggplot(dat, aes(time/stepperyr, atoutput)) +
@@
 
+# compare with true output (all timesteps)
 for(s in names(survObsBiom)){
   cat("  \n##### ",  s,"  \n")
-  print(plotB(survObsBiom[[s]][[1]]))
+  print(plotB(survObsBiom[[s]][[1]], omlist_ss$truetotbio_ss))
   cat("  \n")
 }
+
+# plots survey only
+# for(s in names(survObsBiom)){
+#   print(plotB(survObsBiom[[s]][[1]]))
+# }
 
@@
 
-plotC(catchbio_ss)
+# compare with true catch, should match
+plotC(catchbio_ss, omlist_ss$truecatchbio_ss)
 

### 2020-10-23 msSurveysTest.Rmd
try to compare to true age comps WIP

@@
 
+Crunch a few things to get true age comps
+```{r}
+# age classes
+totNagecl <- omlist_ss$truenums_ss %>%
+  group_by(species, agecl, time) %>%
+  summarize(numAtAge = sum(atoutput))
+
+totN <- totNagecl %>%
+  group_by(species, time) %>%
+  summarize(totN = sum(numAtAge))
+
+trueNagecl <- merge(totNagecl, totN)
+
+# annual ages
+totNage <- omlist_ss$truenumsage_ss %>%
+  group_by(species, agecl, time) %>%
+  summarize(numAtAge = sum(atoutput))
+
+#this should be exactly the same as totN but just making sure...
+totN2 <- totNage %>%
+  group_by(species, time) %>%
+  summarize(totN = sum(numAtAge))
+
+trueNage <- merge(totNage, totN2)
+
+```
+
 My plotting functions: too specialized to include in the package?
@@
 # plot numbers at age by timestep (one species)
-Natageplot <- function(dat, truedat=NULL){
-  ggplot(dat, aes(x=agecl, y=atoutput)) +
-    geom_point() +
+Natageplot <- function(dat, effN=1, truedat=NULL){
+  ggplot() +
+    geom_point(dat, aes(x=agecl, y=atoutput/effN, color="Est Comp")) +
+    {if(!is.null(truedat)) geom_line(data=truedat, aes(x=agecl, y=numAtAge/totN, color="True Comp"), alpha = 3/10)} + 
     theme_tufte() +
@@
 We are using the full time series for biomass and weight at age, and a subsample of years 30-53 for compositions.
+In this test we see the effect of surveying in one season, with migratory species missedor only partially represented in some surveys, and with the overall surveyed biomass being at the lower or higher end of the range depending on the time of year.
+(Observation error is included in the survey biomass time series.) Samples for length and age assume a total of 100000 fish were randomly sampled for each species in each survey (surveffN defined in the survey config files), so older age classes may not be represented.
 
@@
 
-We are using the full time series for total catch and weight at age, and a subsample of years 30-35 for compositions. Fishery length and age composition is sampled every output timestep.
+We are using the full time series for total catch and weight at age, and a subsample of years 30-35 for compositions.
+Fishery length and age composition is sampled every output timestep.
+Fishery effN for length and age sampling was 1000 fish per time step, so this is reflected in the composition plots.
 

### 2020-10-28 msSurveysTest.Rmd
plot age comps relative to true

--- 2020-10-23 msSurveysTest.Rmd
@@
 
+Crunch a few things to get true age comps
+```{r}
+# age classes
+totNagecl <- omlist_ss$truenums_ss %>%
+  group_by(species, agecl, time) %>%
+  summarize(numAtAge = sum(atoutput))
+
+totN <- totNagecl %>%
+  group_by(species, time) %>%
+  summarize(totN = sum(numAtAge))
+
+trueNagecl <- merge(totNagecl, totN)
+
+# annual ages
+totNage <- omlist_ss$truenumsage_ss %>%
+  group_by(species, agecl, time) %>%
+  summarize(numAtAge = sum(atoutput))
+
+#this should be exactly the same as totN but just making sure...
+totN2 <- totNage %>%
+  group_by(species, time) %>%
+  summarize(totN = sum(numAtAge))
+
+trueNage <- merge(totNage, totN2)
+
+# fishery catch at age class
+totCagecl <- omlist_ss$truecatchnum_ss %>%
+  group_by(species, agecl, time) %>%
+  summarize(numAtAge = sum(atoutput))
+
+totCAAcl <- totCagecl %>%
+  group_by(species, time) %>%
+  summarize(totN = sum(numAtAge))
+
+trueCAAcl <- merge(totCagecl, totCAAcl)
+
+# fishery catch at annual age
+totCage <- omlist_ss$truecatchage_ss %>%
+  group_by(species, agecl, time) %>%
+  summarize(numAtAge = sum(atoutput))
+
+totCAA <- totCage %>%
+  group_by(species, time) %>%
+  summarize(totN = sum(numAtAge))
+
+trueCAA <- merge(totCage, totCAA)
+
+```
+
 My plotting functions: too specialized to include in the package?
@@
                           dat$species)) +
-    facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+    facet_wrap(~time, ncol=6, scales="free_y")
 
@@
 # plot numbers at age by timestep (one species)
-Natageplot <- function(dat, truedat=NULL){
-  ggplot(dat, aes(x=agecl, y=atoutput)) +
-    geom_point() +
+Natageplot <- function(dat, effN=1, truedat=NULL){
+  ggplot() +
+    geom_point(data=dat, aes(x=agecl, y=atoutput/effN, color="Est Comp")) +
+    {if(!is.null(truedat)) geom_line(data=dat, aes(x=agecl, y=numAtAge/totN, color="True Comp"))} + 
     theme_tufte() +
+    theme(legend.position = "bottom") +    
     xlab("age/agecl") +
@@
                           dat$species)) + 
-    facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+    facet_wrap(~time, ncol=6, scales="free_y")
 }
@@
 We are using the full time series for biomass and weight at age, and a subsample of years 30-53 for compositions.
+In this test we see the effect of surveying in one season, with migratory species missed or only partially represented in some surveys, and with the overall surveyed biomass being at the lower or higher end of the range depending on the time of year.
+(Observation error is included in the survey biomass time series.) Samples for length and age assume a total of 100000 fish were randomly sampled for each species in each survey (surveffN defined in the survey config files).
 
@@
 #### Survey age class composition (Atlantis age class)
-```{r, results="asis"}
+```{r, results="asis", message=FALSE, warning=FALSE}
 
@@
     group_by(species) %>%
-    group_map(~ Natageplot(.x), keep = TRUE)
+    left_join(., trueNagecl) %>%
+    #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+    group_map(~ Natageplot(.x, effN = 100000, truedat = 1), keep = TRUE) # plots merged true age comp
 
@@
 #### Survey age composition (annual ages)
-```{r plot-annagecomp, results="asis"}
+```{r plot-annagecomp, results="asis", message=FALSE, warning=FALSE}
 
@@
     group_by(species) %>%
-    group_map(~ Natageplot(.x), keep = TRUE)
+    left_join(., trueNage) %>%
+    #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+    group_map(~ Natageplot(.x, effN = 100000, truedat = 1), keep = TRUE) # plots merged true age comp
 
@@
 
-We are using the full time series for total catch and weight at age, and a subsample of years 30-35 for compositions. Fishery length and age composition is sampled every output timestep.
+We are using the full time series for total catch and weight at age, and a subsample of years 30-35 for compositions.
+Fishery length and age composition is sampled every output timestep.
+Fishery effN for length and age sampling was 1000 fish per year (200 per time step), so this is reflected in the composition plots.
 
@@
 #### Fishery catch at age (Atlantis age class)
-```{r}
+```{r, message=FALSE, warning=FALSE}
 
@@
   group_by(species) %>%
-  group_map(~ Natageplot(.x), keep = TRUE)
+  left_join(., trueCAAcl) %>%
+  #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+  group_map(~ Natageplot(.x, effN = 200, truedat = 1), keep = TRUE) # plots merged true age comp
 
@@
 #### Fishery catch at age (annual ages)
-```{r}
+```{r, message=FALSE, warning=FALSE}
 
@@
   group_by(species) %>%
-  group_map(~ Natageplot(.x), keep = TRUE)
+  left_join(., trueCAA) %>%
+  #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+  group_map(~ Natageplot(.x, effN = 200, truedat = 1), keep = TRUE) # plots merged true age comp
 

### 2020-10-29 msSurveysTest.Rmd
add new survey specifications, not run yet

@@
 ### {-}
+
+## Specifying a more realistic survey.
+
+### New survey and fishery config files
+
+`mssurvey_spring_01.R` and `mssurvey_fall_01.R` configure the fishery independent surveys.
+We keep spring and fall survey timing, but now we have efficiency defined for each species and a selectivity function for each species.
+Our surveys exclude offshore areas.
+These specifications are invented for demonstration purposes, not intended to realistically mimic a particular survey:
+
+```{r, code = readLines("./config/mssurvey_spring_01.R"), eval=F}
+```
+
+```{r, code = readLines("./config/mssurvey_fall_01.R"), eval=F}
+```
+
+`msfishery_01.R` configures the fishery dependent data:
+
+```{r, code = readLines("./config/msfishery_01.R"), eval=F}
+```
+
+Model boxes show where our survey works now. We are including boxes x...x in these surveys:
+```{r NOBAmap}
+knitr::include_graphics("images/NOBAboxes.png")
+```
+
+The way sampling is set up, effective N for each survey is the number of fish measured for length, age, and average weight at age.
+We could introduce further realism by using the `sample_ages()` function to take a subsample for age composition and optionally apply an ageing error matrix.
+This would require changing the `om_comps()` wrapper: see [atlantisom issue](https://github.com/r4atlantis/atlantisom/issues/42).
+
+### New survey and fishery outputs
+
+```{r full-test, message=FALSE, warning=FALSE, eval=FALSE}
+
+#NOBAom <- om_init(here("config/NOBA2config.R"))
+
+#NOBAom_ms <- om_species(sppsubset$Name, NOBAom)
+
+NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+
+# new wrapper for multiple surveys saves survey output separately
+# but returns a list with all surveys, not ideal; make consistent
+
+NOBAom_ms_ind <- om_index(usersurvey = c(here("config/mssurvey_spring_01.R"), 
+                                         here("config/mssurvey_fall_01.R")), 
+                           userfishery = here("config/msfishery_01.R"),
+                           omlist_ss = NOBAom_ms, 
+                           n_reps = 1, 
+                           save = TRUE)
+
+NOBAom_ms_comp <- om_comps(usersurvey = c(here("config/mssurvey_spring_01.R"), 
+                                         here("config/mssurvey_fall_01.R")), 
+                           userfishery = here("config/msfishery_01.R"),
+                           omlist_ss = NOBAom_ms, 
+                           n_reps = 1, 
+                           save = TRUE)
+
+```

### 2020-10-29 config/msfishery.R

@@ -30,7 +30,8 @@ fishtime <- fish_times
 fishboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
 
 # effective sample size needed for sample_fish
-# this effective N is high but not equal to total for numerous groups
+# this effective N is divided by the number of annual timesteps below, so 200 per time
+# use as input to the length samples, ages can be a subset
 fisheffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
 
 #this adjusts for subannual fishery output so original effN is for whole year

### 2020-10-29 config/mssurvey_fall_01.R

@@ -1,0 +1,132 @@
+# Default survey configuration here has a range of efficiencies and selectivities
+# To emulate a range of species in a single multispecies survey
+# Also now happens in "spring" and "fall"
+# Need to define survey season, area, efficiency, selectivity
+
+# Survey name
+survey.name="BTS_fall_nearbox_qmix_selmix"
+
+#Atlantis model timestep corresponding to the true output--now from census_spec.R
+timestep <- stepperyr #5
+
+#Which atlantis timestep does the survey run in?--now from census_spec.R
+# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
+# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
+
+survey_sample_time <- 3 # fall survey
+
+#The last timestep to sample
+total_sample <- noutsteps-1 #495
+
+#Vector of indices of survey times to pull
+survey_sample_full <- seq(survey_sample_time,
+                          total_sample, by=timestep)
+
+survtime <- survey_sample_full
+
+# survey area
+# should return all model areas
+#survboxes <- allboxes
+# shelf boxes excluding farthest east Barents
+survboxes <- c(1:8, 21, 23:31, 33, 39:44, 47:49)
+
+# survey efficiency (q)
+# should return a perfectly efficient survey 
+# surveffic <- data.frame(species=survspp,
+#                     efficiency=rep(1.0,length(survspp)))
+
+# group species to apply different efficiency (all NOBA fish and shark groups)
+nontrawl <- c("Sharks_other", "Pelagic_large","Mesop_fish")
+pelagics <- c("Pelagic_small","Redfish_other","Mackerel","Haddock",
+              "Saithe","Redfish","Blue_whiting","Norwegian_ssh","Capelin")
+demersals <- c("Demersals_other","Demersal_large","Flatfish_other","Skates_rays",
+               "Green_halibut","North_atl_cod","Polar_cod","Snow_crab")
+selflats <- c("Long_rough_dab")
+
+#   define bottom trawl mixed efficiency
+ef.nt <- 0.01 # for large pelagics, reef dwellers, others not in trawlable habitat
+ef.pl <- 0.1  # for pelagics
+ef.dm <- 0.6  # for demersals
+ef.fl <- 1.1  # for selected flatfish
+
+# bottom trawl survey efficiency specification by species group
[…]

### 2020-10-29 config/mssurvey_spring_01.R

@@ -1,0 +1,134 @@
+# Default survey configuration here has a range of efficiencies and selectivities
+# To emulate a range of species in a single multispecies survey
+# Also now happens in "spring" and "fall"
+# Need to define survey season, area, efficiency, selectivity
+
+# Survey name
+survey.name="BTS_spring_nearbox_qmix_selmix"
+
+#Atlantis model timestep corresponding to the true output--now from census_spec.R
+timestep <- stepperyr #5
+
+#Which atlantis timestep does the survey run in?--now from census_spec.R
+# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
+# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
+
+survey_sample_time <- 1 # spring survey
+
+#The last timestep to sample
+total_sample <- noutsteps-1 #495
+
+#Vector of indices of survey times to pull
+survey_sample_full <- seq(survey_sample_time,
+                          total_sample, by=timestep)
+
+survtime <- survey_sample_full
+
+# survey area
+# should return all model areas
+# survboxes <- allboxes
+
+# shelf boxes excluding farthest east Barents
+survboxes <- c(1:8, 21, 23:31, 33, 39:44, 47:49)
+
+# survey efficiency (q)
+# should return a perfectly efficient survey 
+# surveffic <- data.frame(species=survspp,
+#                     efficiency=rep(1.0,length(survspp)))
+
+# group species to apply different efficiency (all NOBA fish and shark groups)
+nontrawl <- c("Sharks_other", "Pelagic_large","Mesop_fish")
+pelagics <- c("Pelagic_small","Redfish_other","Mackerel","Haddock",
+              "Saithe","Redfish","Blue_whiting","Norwegian_ssh","Capelin")
+demersals <- c("Demersals_other","Demersal_large","Flatfish_other","Skates_rays",
+               "Green_halibut","North_atl_cod","Polar_cod","Snow_crab")
+selflats <- c("Long_rough_dab")
+
+#   define bottom trawl mixed efficiency
+ef.nt <- 0.01 # for large pelagics, reef dwellers, others not in trawlable habitat
+ef.pl <- 0.2  # for pelagics
+ef.dm <- 0.7  # for demersals
+ef.fl <- 1.1  # for selected flatfish
+
+# bottom trawl survey efficiency specification by species group
[…]

### 2020-11-02 msSurveysTest.Rmd
fixing selectivity WIP

@@
 
+# add this paste0(d.name, "/mscensus1") and add option to name save directory in wrappers
 survObsBiom <- read_savedsurvs(d.name, 'survB')
@@
 
-`msfishery_01.R` configures the fishery dependent data:
+`msfishery_01.R` configures the fishery dependent data, which has a higher observation error on the total catch and does not get biological sampling from all areas:
 
@@
 
-Model boxes show where our survey works now. We are including boxes x...x in these surveys:
+Model boxes show where our survey works now. We are including boxes 1:8, 21, 23:31, 33, 39:44, 47:49 in these surveys, and in fishery comps sampling:
+
 ```{r NOBAmap}
-knitr::include_graphics("images/NOBAboxes.png")
+knitr::include_graphics("visualization_Rmd/images/NOBAboxes.png")
 ```
@@
 
-```{r full-test, message=FALSE, warning=FALSE, eval=FALSE}
+```{r full-test, message=FALSE, warning=FALSE}
 
@@
 ```
+
+Diagnose failure of om_comps on annage data with new survey config file--selectivity function by 10 agecl instead of annual ages.
+
+```{r}
+
+# source(here("config/mssurvey_spring_01.R"))
+# source(here("config/mssurvey_fall_01.R"))
+# source(here("config/msfishery_01.R"))
+
+usersurvey <- c(here("config/mssurvey_spring_01.R"), 
+                here("config/mssurvey_fall_01.R"))
+
+n_reps <- 1
+
+if(!is.null(omlist_ss$truenumsage_ss)){
+    #numbers based fishery independent survey for age and length comps
+    #allows for mulitple surveys
+    annage_comp_datas <- list()
+
+    for (s in usersurvey)
+    {
+      source(s, local = TRUE)
+
+      # same user specifications as indices
+      survey_annageN <- atlantisom::create_survey(dat = omlist_ss$truenumsage_ss,
+                                                  time = survtime,
+                                                  species = survspp,
+                                                  boxes = survboxes,
+                                                  effic = surveffic,
+                                                  selex = survselex)
+      #Sample fish for age composition
+      # if we want replicates for obs error this sample function will generate them
+      annage_comp_data <- list()
+      for(i in 1:n_reps){
+        annage_comp_data[[i]] <- atlantisom::sample_fish(survey_annageN, surveffN)
+      }
+
+      # save survey annual age comps
+      if(save){
+        saveRDS(annage_comp_data, file.path(d.name, paste0(scenario.name, "_",
+                                                           survey.name, "survObsFullAgeComp.rds")))
+      }
+      annage_comp_datas[[survey.name]] <- annage_comp_data
+    }
+
+}else{annage_comp_datas <- NULL}
+
+#breaking at sample_fish; test:
+
+dat <- survey_annageN
+effN <- surveffN
+
+    #sum over boxes and assume sampling occurs coastwide (the sampled boxes were already subset in create functions)
+    dat2 <- aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time),sum)
+    names(dat2) <- c("species","agecl","time","numAtAge")
+
+    #TODO: Need error checking--SKG moved nn assignement of effN to inside y loop from outside
+	  dat2$numAtAgeSamp <- NA
+	  for(sp in unique(dat2$species)) {
+	    for(y in unique(dat2$time)) {
+	      nn <- effN[effN$species==sp,"effN"]
+	      ind <- dat2$species == sp & dat2$time == y
+	      totalNums <- sum(dat2[ind,]$numAtAge) #add na.rm to get totalNums for Greenland halibut
+
+	      if(nn > totalNums) {
+	        nn <- totalNums
+	        message("effN is greater than total numbers available, so nEff set equal to ", nn," for species ",sp," and time ",y,"\n")
+	      }
+	      probs <- matrix(dat2[ind,]$numAtAge,nrow=1)
+	      if(nn > 0){
+	        dat2[ind,]$numAtAgeSamp <- rmultinom(1,nn,probs)[,1]
+	      } else { # sample is 0 if probs vector all 0s, no fish that year
+	        dat2[ind,]$numAtAgeSamp <- rep(0, length(probs))
+	        message("total numAtAge ", nn,", assigning 0 sample for species ",sp," and time ",y,"\n")
+	      }
+	    }
+	  }
+
+# breaking at Greenland halibut, which has no fish in ages 11-20 in this survey
+
+	  survey_annageN %>% filter(species %in% "Green_halibut") %>% group_by(agecl) %>% tally(atoutput)
+
+	  # want the function to return 0 for missing ages
+	  # probs statement returns NA for ages 11-20
+
+```

### 2020-11-02 config/msfishery_01.R

@@ -1,0 +1,47 @@
+# Default fishery configuration here is a census
+fishery.name="nearbox"
+
+# Inherits species from input omlist_ss
+fishspp <- omlist_ss$species_ss 
+
+#Number of years of data to pull
+nyears <- 50
+
+#Atlantis initialization period in years
+burnin <- 30
+
+# fishery output: learned the hard way this can be different from ecosystem outputs
+fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc
+
+
+# same time dimensioning parameters as in surveycensus.R
+#Vector of indices of catch in numbers to pull (by timestep to sum)
+fish_sample_full <- c(0:total_sample)  #total_sample defined in survey configs
+fish_burnin <- burnin*fstepperyr+1
+fish_nyears <- nyears*fstepperyr
+fish_times <- fish_sample_full[fish_burnin:(fish_burnin+fish_nyears-1)]
+fish_timesteps <- seq(fish_times[fstepperyr], max(fish_times), by=fstepperyr) #last timestep
+#fish_years <- unique(floor(fish_times/fstepperyr)+1) # my original
+fish_years <- unique(floor(fish_times/fstepperyr)) #from Christine's new sardine_config.R
+
+fishtime <- fish_times
+
+
+# fishery sampling area (applies to comps, not total catch)
+# should return all model areas, this assumes you see everything that is caught
+#fishboxes <- c(0:(omlist_ss$boxpars$nbox - 1))
+# same as survey, only getting catch comps from nearby boxes
+fishboxes <- c(1:8, 21, 23:31, 33, 39:44, 47:49)
+
+# effective sample size needed for sample_fish
+# this effective N is divided by the number of annual timesteps below, so 200 per time
+# use as input to the length samples, ages can be a subset
+fisheffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+
+#this adjusts for subannual fishery output so original effN is for whole year
+fisheffN$effN <- fisheffN$effN/fstepperyr 
+
+# fishery catch cv can be used in sample_fishery_totcatch to
+# perfect observation is cv=0
+fish_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp)))
+

### 2020-11-02 config/mssurvey_spring_01.R

@@ -61,13 +61,15 @@ efficmix <- bind_rows(effnontrawl, effpelagics, effdemersals, effselflats)
 surveffic <- efficmix %>%
   filter(species %in% survspp)
 
-# survey selectivity (agecl based, flat)
+# survey selectivity (true age based, flat)
 # for annage output uses names(annages) NOT alphabetical survspp
-# survselex <- data.frame(species=rep(names(annages), n_annages), #  
-#                         agecl=unlist(sapply(n_annages,seq)),
-#                         selex=rep(1.0,sum(n_annages)))
+ survselex <- data.frame(species=rep(names(annages), n_annages), #  
+                         agecl=unlist(sapply(n_annages,seq)),
+                         selex=rep(1.0,sum(n_annages)))
 
-#   mixed selectivity (using 10 agecl for all species)
+#   mixed selectivity (using 10 agecl for all species)--breaks for annual ages! specify for annual
+#   then apply that curve to agecl keeping every nth selectivity index n=NumAgeClassSize
+#   ageclsel <- fullsel[seq(NumAgeClassSize, n_annages, length.out=NumCohorts)]
 #     flat=1 for large pelagics, reef dwellers, others not in trawlable habitat
 #     sigmoid 0 to 1 with 0.5 inflection at agecl 3 for pelagics, reaching 1 at agecl 5, flat top
 #     sigmoid 0 to 1 with 0.5 inflection at agecl 5 for most demersals and flatfish, reaching 1 at agecl 7, flat top
@@ -77,24 +79,24 @@ sigmoid <- function(a,b,x) {
   1 / (1 + exp(-a-b*x))
 }
 
-# survey selectivity specification by species group
-selnontrawl <- data.frame(species=rep(nontrawl, each=10),
-                          agecl=rep(c(1:10),length(nontrawl)),
-                          selex=rep(1.0,length(nontrawl)*10))
-selpelagics <- data.frame(species=rep(pelagics, each=10),
-                          agecl=rep(c(1:10),length(pelagics)),
-                          selex=sigmoid(5,1,seq(-10,10,length.out=10)))
-seldemersals <- data.frame(species=rep(demersals, each=10),
-                           agecl=rep(c(1:10),length(demersals)),
-                           selex=sigmoid(1,1,seq(-10,10,length.out=10)))
-selselflats <- data.frame(species=rep(selflats, each=10),
[…]

### 2020-11-03 msSurveysTest.Rmd
selectivity for full annage, now cut to agecl WIP

@@
     facet_wrap(c("species"), scales="free_y")
+}
+
+# compare N at age and C at age between standard and ANNAGE outputs
+totNageplot <- function(dat, anndat){
+
+    ggplot() +
+    geom_line(data=dat, aes(x=time/stepperyr,y=totN, color="Tot N cohorts"), 
+              alpha = 5/10) +
+    geom_line(data=anndat, aes(x=time/stepperyr,y=totN, color="Tot N annage"), alpha = 5/10) + 
+    theme_tufte() +
+    theme(legend.position = "top") +
+    xlab("model year") +
+    ylab("N") +
+    labs(colour=scenario.name) +
+    facet_wrap(~species, scales="free")
+
 }

### 2020-11-03 config/mssurvey_spring_01.R

@@ -79,24 +79,48 @@ sigmoid <- function(a,b,x) {
   1 / (1 + exp(-a-b*x))
 }
 
-selgroups <- as.data.frame(pelagics) %>% mutate(survgroup="pelagics") %>% rename(species = pelagics) %>%
-  bind_rows(as.data.frame(demersals) %>% mutate(survgroup="demersals") %>% rename(species = demersals)) %>%
-  bind_rows(as.data.frame(selflats) %>% mutate(survgroup="selflats") %>% rename(species = selflats))
+# selgroups <- as.data.frame(pelagics) %>% mutate(survgroup="pelagics") %>% rename(species = pelagics) %>%
+#   bind_rows(as.data.frame(demersals) %>% mutate(survgroup="demersals") %>% rename(species = demersals)) %>%
+#   bind_rows(as.data.frame(selflats) %>% mutate(survgroup="selflats") %>% rename(species = selflats))
+# 
+# selage <- sp_age %>%
+#   mutate(n_annages = NumCohorts * NumAgeClassSize) %>%
+#   left_join(selgroups, by=c("Name"="species"))
+# 
+# # survey selectivity specification by species group--replace for each group in surselex
+# 
+# selpelagics <- sigmoid(2*n_annages/5,1,seq(-10,10,length.out=n_annages))
+# seldemersals <- sigmoid(n_annages/4,1,seq(-10,10,length.out=n_annages))
+# selselflats <- sigmoid(1,1,seq(-10,10,length.out=n_annages))
 
-selage <- sp_age %>%
-  mutate(n_annages = NumCohorts * NumAgeClassSize) %>%
-  left_join(selgroups, by=c("Name"="species"))
+selnontrawl <- data.frame(species=rep(nontrawl, each=10),
+                          agecl=rep(c(1:10),length(nontrawl)),
+                          selex=rep(1.0,length(nontrawl)*10))
+selpelagics <- data.frame(species=rep(pelagics, each=10),
+                          agecl=rep(c(1:10),length(pelagics)),
+                          selex=sigmoid(5,1,seq(-10,10,length.out=10)))
+seldemersals <- data.frame(species=rep(demersals, each=10),
+                           agecl=rep(c(1:10),length(demersals)),
+                           selex=sigmoid(1,1,seq(-10,10,length.out=10)))
+selselflats <- data.frame(species=rep(selflats, each=10),
+                          agecl=rep(c(1:10),length(selflats)),
[…]

### 2020-11-05 msSurveysTest.Rmd
update survey configs and rerun

--- 2020-10-29 msSurveysTest.Rmd
@@
 
-survObsBiom <- read_savedsurvs(d.name, 'survB')
-age_comp_data <- read_savedsurvs(d.name, 'survAge')
-len_comp_data <- read_savedsurvs(d.name, 'survLen')
-wtage <- read_savedsurvs(d.name, 'survWtage')
-annage_comp_data <- read_savedsurvs(d.name, 'survAnnAge')
-annage_wtage <- read_savedsurvs(d.name, 'survAnnWtage')
+# add this paste0(d.name, "/mscensus1") and add option to name save directory in wrappers
+survObsBiom <- read_savedsurvs(paste0(d.name, "/mscensus1"), 'survB')
+age_comp_data <- read_savedsurvs(paste0(d.name, "/mscensus1"), 'survAge')
+len_comp_data <- read_savedsurvs(paste0(d.name, "/mscensus1"), 'survLen')
+wtage <- read_savedsurvs(paste0(d.name, "/mscensus1"), 'survWtage')
+annage_comp_data <- read_savedsurvs(paste0(d.name, "/mscensus1"), 'survAnnAge')
+annage_wtage <- read_savedsurvs(paste0(d.name, "/mscensus1"), 'survAnnWtage')
 
@@
 
-catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
+catchbio_ss <- readRDS(file.path(paste0(d.name, "/mscensus1"), paste0(scenario.name, "fishCatch.rds")))
 catchbio_ss <- catchbio_ss[[1]] #in same time unit as surveys
 
-fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
+fish_age_comp <- readRDS(file.path(paste0(d.name, "/mscensus1"), paste0(scenario.name, "fishObsAgeComp.rds")))
 fish_age_comp <- fish_age_comp[[1]]
@@
 
-fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))
+fish_len_comp_data <- readRDS(file.path(paste0(d.name, "/mscensus1"), paste0(scenario.name, "fishObsLenComp.rds")))
 
@@
 
-fishwtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsWtAtAge.rds")))
+fishwtage <- readRDS(file.path(paste0(d.name, "/mscensus1"), paste0(scenario.name, "fishObsWtAtAge.rds")))
 fishwtage <- fishwtage[[1]] #need to change time units?
 
-fish_annage_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
+fish_annage_comp <- readRDS(file.path(paste0(d.name, "/mscensus1"), paste0(scenario.name, "fishObsFullAgeComp.rds")))
 fish_annage_comp <- fish_annage_comp[[1]]
@@
 
-fish_annage_wtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullWtAtAge.rds")))
+fish_annage_wtage <- readRDS(file.path(paste0(d.name, "/mscensus1"), paste0(scenario.name, "fishObsFullWtAtAge.rds")))
 fish_annage_wtage <- fish_annage_wtage[[1]] #need to change time units?
@@
 ```
+
+NOTE!
+Need to have selectivities defined *both* for annual ages and for age classes.
+Worked in census because flat selectivity was defined for full set of annual ages and it just didn't use past age 10 for age class output.
+They were all 1 anyway so it didnt matter.
+The updated survey config file first specifies selectivity by true age and then backs out an approximate equivalent for age classes.
+We will use true age classes for the real datasets but wanted to check both outputs here.
+`om_comps()` has now changed to take `survselex.agecl` as input to the age class comps and survselex as input to the full comps.
+
+Both `om_index()` and `om_comps()` wrappers now name fishery output based on a `fishery.name` in the fishery config file.

### 2020-11-05 config/mssurvey_fall_01.R

@@ -60,23 +60,29 @@ efficmix <- bind_rows(effnontrawl, effpelagics, effdemersals, effselflats)
 surveffic <- efficmix %>%
   filter(species %in% survspp)
 
-# survey selectivity (agecl based, flat)
+# survey selectivity (true age based, flat)
 # for annage output uses names(annages) NOT alphabetical survspp
-# survselex <- data.frame(species=rep(names(annages), n_annages), #  
-#                         agecl=unlist(sapply(n_annages,seq)),
-#                         selex=rep(1.0,sum(n_annages)))
+survselex <- data.frame(species=rep(names(annages), n_annages), #  
+                        agecl=unlist(sapply(n_annages,seq)),
+                        selex=rep(1.0,sum(n_annages)))
 
-#   mixed selectivity (using 10 agecl for all species)
+#   mixed selectivity: specify for annual ages 0-10 to 
+#   then apply that curve to agecl keeping every nth selectivity index n=NumAgeClassSize
+#   ageclsel <- fullsel[seq(NumAgeClassSize, n_annages, length.out=NumCohorts)]
 #     flat=1 for large pelagics, reef dwellers, others not in trawlable habitat
-#     sigmoid 0 to 1 with 0.5 inflection at agecl 3 for pelagics, reaching 1 at agecl 5, flat top
-#     sigmoid 0 to 1 with 0.5 inflection at agecl 5 for most demersals and flatfish, reaching 1 at agecl 7, flat top
-#     dome shaped 0 to 1 at agecl 6&7 for selected demersals, falling off to 0.7 by agecl 10
+#     sigmoid 0 to 1 with 0.5 inflection ~ age 3 for pelagics, reaching 1 at age 5, flat top
+#     sigmoid 0 to 1 with 0.5 inflection ~ age 5 for most demersals and flatfish, reaching 1 at age 7, flat top
+#     dome shaped 0 to 1 at agecl 6&7 for selected demersals, falling off to 0.7 by agecl 10--didn't do
 
 sigmoid <- function(a,b,x) {
   1 / (1 + exp(-a-b*x))
 }
 
-# survey selectivity specification by species group
+sp_age <- sp_age %>%
+  mutate(n_annages = NumCohorts * NumAgeClassSize) 
+# 
+# survey selectivity specification for true ages 1-10 by species group--replace for each group in surselex
+
 selnontrawl <- data.frame(species=rep(nontrawl, each=10),
[…]

### 2020-11-05 config/mssurvey_spring_01.R

@@ -67,31 +67,22 @@ surveffic <- efficmix %>%
                          agecl=unlist(sapply(n_annages,seq)),
                          selex=rep(1.0,sum(n_annages)))
 
-#   mixed selectivity (using 10 agecl for all species)--breaks for annual ages! specify for annual
+#   mixed selectivity: specify for annual ages 0-10 to 
 #   then apply that curve to agecl keeping every nth selectivity index n=NumAgeClassSize
 #   ageclsel <- fullsel[seq(NumAgeClassSize, n_annages, length.out=NumCohorts)]
 #     flat=1 for large pelagics, reef dwellers, others not in trawlable habitat
-#     sigmoid 0 to 1 with 0.5 inflection at agecl 3 for pelagics, reaching 1 at agecl 5, flat top
-#     sigmoid 0 to 1 with 0.5 inflection at agecl 5 for most demersals and flatfish, reaching 1 at agecl 7, flat top
-#     dome shaped 0 to 1 at agecl 6&7 for selected demersals, falling off to 0.7 by agecl 10
+#     sigmoid 0 to 1 with 0.5 inflection ~ age 3 for pelagics, reaching 1 at age 5, flat top
+#     sigmoid 0 to 1 with 0.5 inflection ~ age 5 for most demersals and flatfish, reaching 1 at age 7, flat top
+#     dome shaped 0 to 1 at agecl 6&7 for selected demersals, falling off to 0.7 by agecl 10--didn't do
 
 sigmoid <- function(a,b,x) {
   1 / (1 + exp(-a-b*x))
 }
 
-# selgroups <- as.data.frame(pelagics) %>% mutate(survgroup="pelagics") %>% rename(species = pelagics) %>%
-#   bind_rows(as.data.frame(demersals) %>% mutate(survgroup="demersals") %>% rename(species = demersals)) %>%
-#   bind_rows(as.data.frame(selflats) %>% mutate(survgroup="selflats") %>% rename(species = selflats))
+ sp_age <- sp_age %>%
+   mutate(n_annages = NumCohorts * NumAgeClassSize) 
 # 
-# selage <- sp_age %>%
-#   mutate(n_annages = NumCohorts * NumAgeClassSize) %>%
-#   left_join(selgroups, by=c("Name"="species"))
-# 
-# # survey selectivity specification by species group--replace for each group in surselex
-# 
-# selpelagics <- sigmoid(2*n_annages/5,1,seq(-10,10,length.out=n_annages))
[…]

### 2020-11-05 msSurveysTest.Rmd
plot new surveys--not yet run

@@
 Both `om_index()` and `om_comps()` wrappers now name fishery output based on a `fishery.name` in the fishery config file.
+
+```{r read-newsurvs}
+
+# source(here("config", "NOBA2config.R"))
+# omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+# source(here("config/omdimensions.R"))
+
+# add this paste0(d.name, "/mscensus1") and add option to name save directory in wrappers
+survObsBiom <- read_savedsurvs(d.name, 'survB')
+age_comp_data <- read_savedsurvs(d.name, 'survAge')
+len_comp_data <- read_savedsurvs(d.name, 'survLen')
+wtage <- read_savedsurvs(d.name, 'survWtage')
+annage_comp_data <- read_savedsurvs(d.name, 'survAnnAge')
+annage_wtage <- read_savedsurvs(d.name, 'survAnnWtage')
+
+# no function needed for saved fishery output as long as there is only one fishery specification file
+
+catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
+catchbio_ss <- catchbio_ss[[1]] #in same time unit as surveys
+
+fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
+fish_age_comp <- fish_age_comp[[1]]
+
+#add this to om_indices function so that this has years when read in
+#fish_age_comp$time <- fish_age_comp$time/fstepperyr
+
+fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))
+
+#len_comp_data <- len_comp_data[[1]]
+fish_len_comp_data <- fish_len_comp_data[[1]]
+
+#add this to om_indices function so that this has years when read in
+#fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))
+
+fishwtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsWtAtAge.rds")))
+fishwtage <- fishwtage[[1]] #need to change time units?
+
+fish_annage_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
+fish_annage_comp <- fish_annage_comp[[1]]
+
+#add this to om_indices function so that this has years when read in
+#fish_annage_comp$time <- fish_annage_comp$time/fstepperyr
+
+fish_annage_wtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullWtAtAge.rds")))
+fish_annage_wtage <- fish_annage_wtage[[1]] #need to change time units?
+
+```
+
+### Visualize surveys with different q and selectivity {.tabset #census}
+
+We are using the full time series for biomass and weight at age, and a subsample of years 30-53 for compositions.
+In this test we see the effect of survey efficiency and selectivity as well as the effect of surveying in one season.
+(Observation error that differs by species group is included in the survey biomass time series.) Samples for length and age assume a total of 10000 fish were randomly sampled for each species in each survey (surveffN defined in the survey config files).
+
+#### Survey biomass
+```{r, results="asis"}
+
+# compare with true output (all timesteps)
+for(s in names(survObsBiom)){
+  cat("  \n##### ",  s,"  \n")
+  print(plotB(survObsBiom[[s]][[1]], omlist_ss$truetotbio_ss))
+  cat("  \n")
+}
+
+# plots survey only
+# for(s in names(survObsBiom)){
+#   print(plotB(survObsBiom[[s]][[1]]))
+# }
+
+```
+
+#### Survey length composition
+```{r, results="asis"}
+
+for(s in names(len_comp_data)){
+  cat("  \n##### ",  s,"  \n")
+  lcompsub <- as.data.frame(len_comp_data[[s]][[1]]) %>% filter(time %in% c(150:270)) %>%
+    group_by(species) %>%
+    group_map(~ plotlen(.x), keep = TRUE)
+
+  for(i in 1:length(lcompsub)) {
+    print(lcompsub[[i]])
+  }
+  cat("  \n")
+}
+
+```
+
+#### Survey age class composition (Atlantis age class)
+```{r, results="asis", message=FALSE, warning=FALSE}
+
+for(s in names(age_comp_data)){
+  cat("  \n##### ",  s,"  \n")
+  acompsub <- as.data.frame(age_comp_data[[s]][[1]]) %>% filter(time %in% c(150:270)) %>%
+    group_by(species) %>%
+    left_join(., trueNagecl) %>%
+    #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+    group_map(~ Natageplot(.x, effN = 100000, truedat = 1), keep = TRUE) # plots merged true age comp
+
+  for(i in 1:length(acompsub)) {
+    print(acompsub[[i]])
+  }
+  cat("  \n")
+}
+
+```
+
+#### Survey age composition (annual ages)
+```{r plot-annagecomp, results="asis", message=FALSE, warning=FALSE}
+
+for(s in names(annage_comp_data)){
+  cat("  \n##### ",  s,"  \n")
+  acompsub <- as.data.frame(annage_comp_data[[s]][[1]]) %>% filter(time %in% c(150:270)) %>%
+    group_by(species) %>%
+    left_join(., trueNage) %>%
+    #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+    group_map(~ Natageplot(.x, effN = 100000, truedat = 1), keep = TRUE) # plots merged true age comp
+
+  for(i in 1:length(acompsub)) {
+    print(acompsub[[i]])
+  }
+  cat("  \n")
+}
+```
+
+#### Survey weight at age (Atlantis age class)
+```{r, results='asis'}
+
+for(s in names(wtage)){
+  cat("  \n##### ",  s,"  \n")
+  print(wageplot(wtage[[s]][[1]]))
+  cat("  \n")
+}
+
+```
+
+#### Survey iterpolated weight at age (annual ages) 
+```{r plot-annwtage, fig.height=10, results='asis'}
+
+# annage_wtage read in above 
+# annage_wtage <- annage_wtage[[1]] #this still has second list component, diagnostic plot
+# annage_wtage <- annage_wtage[[1]] #still in this old version but just removed from function.
+
+for(s in names(annage_wtage)){
+  cat("  \n##### ",  s,"  \n")
+  print(wageplot(annage_wtage[[s]][[1]][[1]]))
+  cat("  \n")
+}
+```
+
+### {-}
+
+### Visualize fisheries observed with some error {.tabset}
+
+Here we are not sampling all fished areas, and the observation error on the total catch index is higher.
+We are using the full time series for total catch and weight at age, and a subsample of years 30-35 for compositions.
+Fishery length and age composition is sampled every output timestep.
+Fishery effN for length and age sampling was 1000 fish per year (200 per time step), so this is reflected in the composition plots.
+
+#### Fishery total catch
+```{r}
+
+# compare with true catch, should match
+plotC(catchbio_ss, omlist_ss$truecatchbio_ss)
+
+```
+
+#### Fishery length composition 
+```{r}
+
+lcompsub <- as.data.frame(fish_len_comp_data) %>% filter(time %in% c(150:175)) %>%
+  group_by(species) %>%
+  group_map(~ plotlen(.x), keep = TRUE)
+
+for(i in 1:length(lcompsub)) {
+  print(lcompsub[[i]])
+}
+
+```
+
+#### Fishery catch at age (Atlantis age class)
+```{r, message=FALSE, warning=FALSE}
+
+acompsub <- as.data.frame(fish_age_comp) %>% filter(time %in% c(150:175)) %>%
+  group_by(species) %>%
+  left_join(., trueCAAcl) %>%
+  #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+  group_map(~ Natageplot(.x, effN = 200, truedat = 1), keep = TRUE) # plots merged true age comp
+
+for(i in 1:length(acompsub)) {
+  print(acompsub[[i]])
+}
+
+```
+
+#### Fishery catch at age (annual ages)
+```{r, message=FALSE, warning=FALSE}
+
+acompsub <- as.data.frame(fish_annage_comp) %>% filter(time %in% c(150:175)) %>%
+  group_by(species) %>%
+  left_join(., trueCAA) %>%
+  #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+  group_map(~ Natageplot(.x, effN = 200, truedat = 1), keep = TRUE) # plots merged true age comp
+
+for(i in 1:length(acompsub)) {
+  print(acompsub[[i]])
+}
+
+```
+
+#### Fishery weight at age (Atlantis age class)
+```{r}
+
+wageplot(fishwtage)
+
+```
+
+#### Fishery interpolated weight at age (annual ages)
+```{r, fig.height=10}
+
+#diagnostic plot component still in this old version but just removed from function.
+wageplot(fish_annage_wtage[[1]])
+
+```
+
+### {-}

### 2020-11-05 msSurveysTest.Rmd
add prelim comparison plots

@@
 
-Crunch a few things to get true age comps
+Crunch a few things to get true age comps and census length comp for later
 ```{r}
@@
 
+# census length comps each survey
+lf_census_prop <- list()
+for(s in names(len_comp_data)){
+  lcomp <- as.data.frame(len_comp_data[[s]][[1]])
+  lf_census_tot <- aggregate(lcomp$atoutput,list(lcomp$species, lcomp$time),sum )
+  names(lf_census_tot) <- c("species", "time","totlen")
+  lf_census <- left_join(lcomp, lf_census_tot) %>%
+    rename(censuslen = atoutput)
+  lf_census_prop[[s]] <- lf_census
+}
+
+# fishery length comps (haven't changed here)
+
 ```
@@
 # plot length frequencies by timestep (one species)
-plotlen <- function(dat, truedat=NULL){
+plotlen <- function(dat, effN=1, truedat=NULL){
 
-    ggplot(dat, aes(upper.bins)) +
-    geom_bar(aes(weight = atoutput)) +
+    ggplot(mapping=aes(x=upper.bins)) +
+    {if(!is.null(truedat)) geom_bar(data=truedat, aes(weight = censuslen/totlen), fill="Census Len Comp", alpha = 5/10)} +
+    geom_bar(data=dat, aes(weight = atoutput/effN), fill="Len Comp") +
     theme_tufte() +
     xlab("length (cm)") +
-    ylab("number") +
+    {if(is.null(truedat)) ylab("number")} +
+    {if(!is.null(truedat)) ylab("proportion")} +
     labs(subtitle = paste(scenario.name,
@@
     xlab("age/agecl") +
-    ylab("number") +
+    {if(is.null(truedat)) ylab("number")} +
+    {if(!is.null(truedat)) ylab("proportion")} +
     labs(subtitle = paste(scenario.name,
@@
 
-```{r read-newsurvs}
+```{r read-newsurvs, message=FALSE, warning=FALSE}
 
@@
 # source(here("config/omdimensions.R"))
+
+source(here("config/mssurvey_spring_01.R"))
+source(here("config/msfishery_01.R"))
 
@@
 
-catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
+catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "_",
+                                                     fishery.name, "fishCatch.rds")))
 catchbio_ss <- catchbio_ss[[1]] #in same time unit as surveys
 
-fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
+fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "_",
+                                                     fishery.name, "fishObsAgeComp.rds")))
 fish_age_comp <- fish_age_comp[[1]]
@@
 
-fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))
+fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "_",
+                                                     fishery.name, "fishObsLenComp.rds")))
 
@@
 
-fishwtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsWtAtAge.rds")))
+fishwtage <- readRDS(file.path(d.name, paste0(scenario.name, "_",
+                                                     fishery.name, "fishObsWtAtAge.rds")))
 fishwtage <- fishwtage[[1]] #need to change time units?
 
-fish_annage_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
+fish_annage_comp <- readRDS(file.path(d.name, paste0(scenario.name, "_",
+                                                     fishery.name, "fishObsFullAgeComp.rds")))
 fish_annage_comp <- fish_annage_comp[[1]]
@@
 
-fish_annage_wtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullWtAtAge.rds")))
+fish_annage_wtage <- readRDS(file.path(d.name, paste0(scenario.name, "_",
+                                                     fishery.name, "fishObsFullWtAtAge.rds")))
 fish_annage_wtage <- fish_annage_wtage[[1]] #need to change time units?
@@
 
-### Visualize surveys with different q and selectivity {.tabset #census}
+### Visualize surveys with different q and selectivity {.tabset #survey}
 
@@
     group_by(species) %>%
-    group_map(~ plotlen(.x), keep = TRUE)
+    left_join(., lf_census_prop[[s]]) %>%
+    #group_map(~ plotlen(.x), keep = TRUE)
+    group_map(~ plotlen(.x, effN = 10000, truedat = 1), keep = TRUE) # plots merged true length comp
 
@@
 #### Survey age composition (annual ages)
-```{r plot-annagecomp, results="asis", message=FALSE, warning=FALSE}
+```{r plot-annagecomp-2, results="asis", message=FALSE, warning=FALSE}
 
@@
 #### Survey iterpolated weight at age (annual ages) 
-```{r plot-annwtage, fig.height=10, results='asis'}
+```{r plot-annwtage-2, fig.height=10, results='asis'}
 
@@
   cat("  \n##### ",  s,"  \n")
-  print(wageplot(annage_wtage[[s]][[1]][[1]]))
+  print(wageplot(annage_wtage[[s]][[1]]))
   cat("  \n")
@@
 #diagnostic plot component still in this old version but just removed from function.
-wageplot(fish_annage_wtage[[1]])
+wageplot(fish_annage_wtage)
 

### 2020-11-05 msSurveysTest.Rmd
all new plots and add to index

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
 library(here)
@@
 }
+#lf_census_prop <- bind_rows(lf_census_prop) flatten, only works if surveys at different times
 
@@
 
+# note on ggplot default colors, can get the first and second using this
+# library(scales)
+# show_col(hue_pal()(2))
+
 # plot length frequencies by timestep (one species)
@@
 
+  cols <- c("Census Lcomp"="#00BFC4","Sample Lcomp"="#F8766D")  
   ggplot(mapping=aes(x=upper.bins)) +
-    {if(!is.null(truedat)) geom_bar(data=truedat, aes(weight = censuslen/totlen), fill="Census Len Comp", alpha = 5/10)} +
-    geom_bar(data=dat, aes(weight = atoutput/effN), fill="Len Comp") +
+    {if(is.null(truedat)) geom_bar(data=dat, aes(weight = atoutput/effN))} +
+    {if(!is.null(truedat)) geom_bar(data=dat, aes(weight = censuslen/totlen, fill="Census Lcomp"), alpha = 5/10)} +
+    {if(!is.null(truedat)) geom_bar(data=dat, aes(weight = atoutput/effN, fill="Sample Lcomp"), alpha = 5/10)} +
     theme_tufte() +
+    theme(legend.position = "bottom") +
     xlab("length (cm)") +
@@
     {if(!is.null(truedat)) ylab("proportion")} +
+    scale_colour_manual(name="", values=cols) +
     labs(subtitle = paste(scenario.name,
@@
 #### Survey length composition
-```{r, results="asis"}
+```{r, results="asis", message=FALSE}
 
@@
     group_by(species) %>%
-    left_join(., lf_census_prop[[s]]) %>%
+    left_join(., lf_census_prop[[which(names(len_comp_data) == s)]]) %>%
     #group_map(~ plotlen(.x), keep = TRUE)
@@
     #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
-    group_map(~ Natageplot(.x, effN = 100000, truedat = 1), keep = TRUE) # plots merged true age comp
+    group_map(~ Natageplot(.x, effN = 10000, truedat = 1), keep = TRUE) # plots merged true age comp
 

### 2020-11-10 msSurveysTest.Rmd
rerun fishery comps with smaller effN

@@
 Fishery length and age composition is sampled every output timestep.
-Fishery effN for length and age sampling was 1000 fish per year (200 per time step), so this is reflected in the composition plots.
+Fishery effN for length and age sampling was 500 fish per year (100 per time step), so this is reflected in the composition plots.
+
+```{r rerun-fishery-comps-only, eval=FALSE}
+
+# I changed the effN based on NE sampling and didn't want to run the full om_comps again because surveys aren't changing
+# these are the fishery comps snippets only
+
+source(here("config", "NOBA2config.R"))
+omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+source(here("config/omdimensions.R"))
+
+#The last timestep to sample
+total_sample <- noutsteps-1 #495
+n_reps <- 1
+save = TRUE
+lenage_cv <- 0.1
+maxbin <- 200
+
+userfishery <- here("config/msfishery_01.R")
+
+# snippet of om_comps
+  #now do fishery comps
+  # user options for fishery--default is a census with mid-year sample
+  # only one fishery, but multiple fleets possible within it
+  source(userfishery, local = TRUE)
+
+  #fishery catch at age each observed timestep summed over observed polygons
+  # catch at age by area and timestep
+  catch_numbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchnum_ss,
+                                                      time = fishtime,
+                                                      species = survspp,
+                                                      boxes = fishboxes)
+
+  # if we want replicates for obs error this sample function will generate them
+  catch_age_comp <- list()
+  for(i in 1:n_reps){
+    catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
+  }
+
+  # save fishery age comps
+  if(save){
+    saveRDS(catch_age_comp, file.path(d.name, paste0(scenario.name, "_",
+                                                     fishery.name, "fishObsAgeComp.rds")))
+  }
+
+  #Get catch weights for length comp calc
+  # aggregate true resn per fishery subset design
+  catch_aggresnss <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
+                                                      time = fishtime,
+                                                      species = survspp,
+                                                      boxes = fishboxes)
+
+  # aggregate true structn fishery subsetdesign
+  catch_aggstructnss <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
+                                                         time = fishtime,
+                                                         species = survspp,
+                                                         boxes = fishboxes)
+
+  #dont sample these, just aggregate them using median
+  catch_structnss <- atlantisom::sample_fish(catch_aggstructnss, fisheffN, sample = FALSE)
+
+  catch_resnss <- atlantisom::sample_fish(catch_aggresnss, fisheffN, sample = FALSE)
+
+  # these fishery lengths and weight at age are each output timestep
+  #same structure as above for surveys, replicates follow age comp reps
+  #  separating the length comps from the weight at age here
+  fishery_lenwt <- list()
+  fishObsLenComp <- list()
+  fishObsWtAtAge <- list()
+
+  for(i in 1:n_reps){
+    fishery_lenwt[[i]] <- atlantisom::calc_age2length(structn = catch_structnss,
+                                                      resn = catch_resnss,
+                                                      nums = catch_age_comp[[i]],
+                                                      biolprm = omlist_ss$biol,
+                                                      fgs = omlist_ss$funct.group_ss,
+                                                      maxbin = maxbin,
+                                                      CVlenage = lenage_cv,
+                                                      remove.zeroes=TRUE)
+
+    fishObsLenComp[[i]] <- fishery_lenwt[[i]]$natlength
+    fishObsWtAtAge[[i]] <- fishery_lenwt[[i]]$muweight
+  }
+
+  if(save){
+    saveRDS(fishObsLenComp, file.path(d.name, paste0(scenario.name, "_",
+                                                     fishery.name, "fishObsLenComp.rds")))
+    saveRDS(fishObsWtAtAge, file.path(d.name, paste0(scenario.name, "_",
+                                                     fishery.name, "fishObsWtAtAge.rds")))
+  }
+
+    #fishery catch at age each observed timestep summed over observed polygons
+    # catch at age by area and timestep
+    catch_annagenumbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchage_ss,
+                                                        time = fishtime,
+                                                        species = survspp,
+                                                        boxes = fishboxes)
+
+    # if we want replicates for obs error this sample function will generate them
+    # WARNING THIS AGGREGATES ACROSS FLEETS
+    # TODO: need to change sample_fish to fix
+    catch_annage_comp <- list()
+    for(i in 1:n_reps){
+      catch_annage_comp[[i]] <- atlantisom::sample_fish(catch_annagenumbers, fisheffN)
+    }
+
+    # save fishery annual age comps
+    if(save){
+    saveRDS(catch_annage_comp, file.path(d.name, paste0(scenario.name,"_",
+                                                        fishery.name, "fishObsFullAgeComp.rds")))
+    }
+
+    interp_fishWtAtAge <- list()
+    for(i in 1:n_reps){
+      interp_fishWtAtAge[[i]] <- calc_avgwtstage2age(wtagecl = fishObsWtAtAge[[i]],
+                                                     annages = omlist_ss$truecatchage_ss,
+                                                     fgs = omlist_ss$funct.group_ss)
+    }
+    if(save){
+      saveRDS(interp_fishWtAtAge, file.path(d.name, paste0(scenario.name,"_",
+                                                           fishery.name, "fishObsFullWtAtAge.rds")))
+    }
+
+```
 
@@
   #group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
-  group_map(~ Natageplot(.x, effN = 200, truedat = 1), keep = TRUE) # plots merged true age comp
+  group_map(~ Natageplot(.x, effN = 100, truedat = 1), keep = TRUE) # plots merged true age comp
 

### 2020-11-10 config/msfishery_01.R

@@ -34,9 +34,10 @@ fishtime <- fish_times
 fishboxes <- c(1:8, 21, 23:31, 33, 39:44, 47:49)
 
 # effective sample size needed for sample_fish
+# based on fishery lengths in Northeast US, my original effN of 1000 was too high
 # this effective N is divided by the number of annual timesteps below, so 200 per time
 # use as input to the length samples, ages can be a subset
-fisheffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))
+fisheffN <- data.frame(species=survspp, effN=rep(500, length(survspp)))
 
 #this adjusts for subannual fishery output so original effN is for whole year
 fisheffN$effN <- fisheffN$effN/fstepperyr 

### 2020-11-10 SurveyDietCompTest.Rmd
starting om_diet function

---
title: "Multispecies survey for diet comp"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)  
library(atlantisom)
library(ggthemes)

```

## Species surveyed

Diet composition sampling will need to include more than the species in the set of 11... for now lets get the diet comp for just those species. In the next step we will figure out how to deal with biomass pool groups in surveys and for diets.

We will want a new `atlantisom` wrapper function to deal with diets separately. The current `om_comps()` already takes a long time to run generating length compositions and adding the huge diet file here will be really cumbersome for users only looking for ages and lengths.

The `om_diet()` function will either call `load_detailed_diet_comp()` on a pre-processed zipped `DetailedDietCheck.txt.gz` file as explained [here](https://sgaichas.github.io/poseidon-dev/NOBAmsdiets.html) or will read in an .rds file saved from its output.

It will then apply `create_survey()` with the diet file as dat and the survey specifications from the config file:

```{r}

om_diet<- function(config = configfile,
                   dietfile = file_diet,
                   usersurvey = usersurvey_file,
                   userfishery = NULL,
                   omlist_ss,
                   n_reps = n_reps,
                   save = TRUE){

  if(!file.exists(file.path(d.name,
                            paste0(scenario.name, "ms_detaileddiet.rds")))){
    ms_detaileddiet <- load_detailed_diet_comp(dir = d.name, 
                                               file_diet, 
                                               fgs = om_list_ss$fgs)

    # passing fgs for the subset so shouldn't need this filter
    # ms_detaileddiet <- detdiet %>%
    #   filter(species %in% sppsubset$Name)

    if(save){
      saveRDS(ms_detaileddiet, file.path(d.name, paste0(scenario.name, "ms_detaileddiet.rds")))
    }

  } else {
    ms_detaileddiet <- get(load(file.path(d.name,
                                paste0(scenario.name, "ms_detaileddiet.rds"))))
  }

  survey_diet = create_survey(dat = ms_detaileddiet,
                              time = survtime,
                              species = survspp,
                              boxes = survboxes,
                              effic = surveffic,
                              selex = survselex)

}
```

### 2020-11-10 SurveyDietCompTest.Rmd
initial commit survey diet testing

@@
 
-## Species surveyed
+## Apply survey design to diet outputs
 
@@
 
-```{r}
+```{r omdiet-test, eval=FALSE}
 
@@
                    dietfile = file_diet,
+                   fgs = fgs,
                    usersurvey = usersurvey_file,
@@
 
+  # load or read in saved detailed diet
   if(!file.exists(file.path(d.name,
-                            paste0(scenario.name, "ms_detaileddiet.rds")))){
-    ms_detaileddiet <- load_detailed_diet_comp(dir = d.name, 
+                            paste0(scenario.name, "detaileddiet.rds")))){
+    detaileddiet <- load_detailed_diet_comp(dir = d.name, 
                                                file_diet, 
-                                               fgs = om_list_ss$fgs)
-
-    # passing fgs for the subset so shouldn't need this filter
-    # ms_detaileddiet <- detdiet %>%
-    #   filter(species %in% sppsubset$Name)
+                                               fgs = fgs)
 
     if(save){
-      saveRDS(ms_detaileddiet, file.path(d.name, paste0(scenario.name, "ms_detaileddiet.rds")))
+      saveRDS(detaileddiet, file.path(d.name, paste0(scenario.name, "detaileddiet.rds")))
     }
@@
   } else {
-    ms_detaileddiet <- get(load(file.path(d.name,
-                                paste0(scenario.name, "ms_detaileddiet.rds"))))
+    detaileddiet <- readRDS(file.path(d.name,
+                                paste0(scenario.name, "detaileddiet.rds")))
   }
 
-  survey_diet = create_survey(dat = ms_detaileddiet,
+  #one script for dimension parameters to be used in multiple functions
+  source("config/omdimensions.R", local = TRUE)
+
+  survObsDiets <- list()
+
+  for (s in usersurvey)
+  {
+    source(s, local = TRUE)
+
+    # survtime doesn't match units of time.days in detaileddiet
+    survtime <- survey_sample_full*omlist_ss$runpar$outputstep
+
+    # apply survey design to detailed diet
+    survey_cons <- create_survey_diet(dat = detaileddiet,
                                       time = survtime,
@@
 
+    # add observation error to survey diet
+    survObsDiet <- list()
+    for(i in 1:n_reps){
+      survObsDiet[[i]] <- atlantisom::sample_diet(survey_diet, surv_cv)
+    }
+
+    #save survey diets, takes a long time to generate with lots of reps/species
+    if(save){
+      saveRDS(survObsDiet, file.path(d.name, paste0(scenario.name, "_",
+                                                     survey.name, "surveydiet.rds")))
+    }
+
+    survObsDiets[[survey.name]] <- survObsDiet
+
+  }
+
 }
 ```
+
+Reading in the full diet file works. Saved as `[...]detaileddiet.rds`, a huge file.
+
+```{r, readdiet-test}
+
+fgs <- load_fgs(here("atlantisoutput", "NOBA_March_2020"), "nordic_groups_v04.csv")
+
+source(here("config", "NOBA2config.R"))
+
+file_diet <- "NOBADetDiet.gz"
+save = TRUE
+
+  if(!file.exists(file.path(d.name,
+                            paste0(scenario.name, "detaileddiet.rds")))){
+    detaileddiet <- load_detailed_diet_comp(dir = d.name, 
+                                               file_diet, 
+                                               fgs = fgs)
+
+    if(save){
+      saveRDS(detaileddiet, file.path(d.name, paste0(scenario.name, "detaileddiet.rds")))
+    }
+
+  } else {
+    detaileddiet <- readRDS(file.path(d.name,
+                                paste0(scenario.name, "detaileddiet.rds")))
+  }
+
+```
+
+Fixing mismatches between diet data objects and survey function inputs: align time units and column name, add prey column to aggregateData function call.
+It was hard-coded without the prey column in `create_survey()` and also didn't output prey so a new function, `create_survey_diet()` fixes this.
+
+If we modify the existing `create_survey()`, the total consumption in tons aggregated over layers (output of aggregateData) will be multiplied by our survey efficiency (q) and selectivity at agecl.
+Is this appropriate?
+
+In a global diet composition, applying selectivity may make sense because it excludes or downweights consumption by less selected age classes.
+However, if we are looking at agecl specific diet, the composition is the same, just scaled down.
+Similar thoughts for survey q, all it does is scale the total consumption in tons across all ages so the comp is the same but tonnage smaller.
+Would this be double counting the effects of q if put back into a model and scaled up somehow?
+Need to see how consumption in tons is used in our models.
+For now `create_survey_diet()` applies q and selectivity exactly like `create_survey()`.
+
+```{r, surveydiet-test, eval=FALSE}
+
+omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+source(here("config/omdimensions.R"))
+
+# first a census in spring
+source(here("config/mssurvey_spring.R"))
+
+# survtime doesn't match units of time.days in detaileddiet, add to om_diet
+survtime <- survey_sample_full*omlist_ss$runpar$outputstep
+
+# aggregateData called by create_survey expects a column "time", add to function
+#  if("time.days" %in% names(dat) & !("time" %in% names(dat))) {
+#    names(dat)[names(dat) == 'time.days'] <- 'time'
+#  }
+
+# aggregateData doesn't keep the prey column
+# create survey also won't output prey column and hardcodes keep columns for aggegateData
+# use functions separately
+
+# aggregate over layer, adding prey column (names resulting column "numAtAge", should generalize)
+	aggDat <- aggregateData(dat = detaileddiet, 
+	                        time = survtime, 
+	                        species = survspp, 
+	                        boxes = survboxes, 
+	                        keepColumns=c("species","agecl","polygon","time", "prey"))
+
+# should we be multiplying true consumption by q and selectivity? create_survey does
+# selectivity misses consumption from younger age classes
+
+	effic <- surveffic
+	selex <- survselex # should be survselex.agecl but identical in census survey config
+
+	surv <- aggDat   #this can be removed and uncomment density stuff above. Make sure to think how this plays into sampling
+	#merge in efficiency
+	surv <- merge(surv,effic,by="species",all.x=T)
+	#merge in selex
+	surv <- merge(surv,selex,by=c("species","agecl"),all.x=T)
+	#should I change any missing selex to zero???
+	#should I scale or check selex maximum at 1?
+
+	#surv$numAtAgeSurv <- surv$density * surv$survArea * surv$efficiency * surv$selex
+	surv$numAtAgeSurv <- surv$numAtAge * surv$efficiency * surv$selex
+
+	#Should I be checking for NA's along the way to identify problems?
+
+	#Create final dataframe in same format as input
+	#put time (mean) and layers (NA) back in the dataframe for completeness
+	out <- data.frame(species = surv$species,
+		              agecl = surv$agecl,
+		              polygon = surv$polygon,
+		              layer = NA,
+		              time = surv$time,
+		              prey = surv$prey,
+		              atoutput = surv$numAtAgeSurv)
+
+	survey_cons_test <- out[order(out$species,out$time,out$polygon,out$agecl),]
+
+# survey_diet <- create_survey(dat = detaileddiet,
+#                              time = survtime,
+#                              species = survspp,
+#                              boxes = survboxes,
+#                              effic = surveffic,
+#                              selex = survselex)
+
+```
+
+```{r, test-createsurveydiet}
+
+omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+source(here("config/omdimensions.R"))
+
+# first a census in spring
+source(here("config/mssurvey_spring.R"))
+
+# survtime doesn't match units of time.days in detaileddiet, add to om_diet
+survtime <- survey_sample_full*omlist_ss$runpar$outputstep
+
+survey_cons <- create_survey_diet(dat = detaileddiet,
+                                  time = survtime,
+                                  species = survspp,
+                                  boxes = survboxes,
+                                  effic = surveffic,
+                                  selex = survselex)
+
+```

### 2020-11-10 docs/index.Rmd
add to index
plot age comps relative to true
all new plots and add to index
initial commit survey diet testing

@@
 
-### Full age structured data for multispecies modeling, June-October 2020
+### Full age structured data for multispecies modeling, June-November 2020
+
+<a href="https://sgaichas.github.io/poseidon-dev/SurveyDietCompTest.html" target="_blank">Apply survey specifications to diet outputs, November 10, 2020</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/msSurveysTest.html" target="_blank">Compare multispecies survey specifications, November 9, 2020</a>
 

### 2020-11-18 SurveyDietCompTest.Rmd
add Kerim's diet code and outline of new sample_diet

@@
 ```
+
+We want `sample_diet()` to give us something like we would get out of the survey sampling process.
+The diet sampling function could return diet proportions by predator, including unknown/family level id, with observation error.
+This would then be input into the models.
+True diets will be the actual consumption in tons output from `load_detailed_diet_comp()`.
+
+The current partial `sample_diet()` function needs to change because it expects a wide format file of diet proportions, rather than a long-format file of consumption in tons which is the output of `create_survey_diet()`.
+
+It would be nice if `sample_diet()` could take either the output of `load_diet_comp()` (global proportions not surveyed, long format) or `create_survey_diet()`.
+
+The current function removes predators not expected on a fish trawl survey, removes prey types that are normally not counted, then adds uniform error to half the remaining diet proportions, removes a random subet of prey accounting for <0.25 diet proportion, then renormalizes.
+I would rather have a function that could be used on any "survey" so have already removed hardcoded predator selection.
+I think also the distributional assumptions need revision; a Dirichlet seems most appropriate to simulate continuous proportions.
+
+```{r, scratch-samplediet, eval=FALSE}
+
+sample_diet <- function(dat, fgs) {
+
+  # read in dat, long format output of load_diet_comp, or create_survey_diet, both have layers aggregated
+
+  # load_diet_comp has no polygon information and is proportion, create_survey_diet has polygons and is consumption in tons
+
+  # first aggregate to global diet comp in proportion
+
+  # then apply aggregation bias
+
+  # then apply Dirichlet obs error; Dirichlet equivalent of rmultinom(1, effN, trueagecomp)?   
+  # see e.g., https://rdrr.io/cran/DirichletReg/man/Dirichlet.html
+
+  # returns diet proportions with observation error: Dirichlet, and bias from unknown id and/or family level aggregation
+
+  return(dietprop)
+}
+
+  # old function below
+  # # first remove species not sampled and not quantified in gut analyses
+  # colnames(fgs) <- tolower(colnames(fgs))
+  #
+  # # check for GroupType or InvertType
+  # if (!("grouptype" | "inverttype") %in% colnames(fgs)) {
+  #   stop(paste("The columns GroupType  or InvertType ars not in your functional\n",
+  #     "groups file and thus sample_diet does not know which groups to sample."))
+  # }
+  #
+  # # change inverttype to grouptype, contents should be the same
+  # if("inverttype" %in% names(fgs)) names(fgs)[names(fgs) == 'inverttype'] <- 'grouptype'
+  #
+  # fgs$grouptype <- tolower(fgs$grouptype)
+  # nonsampledtypes <- c("bird", "mammal", "cep", "sed_ep_ff", "sed_ep_other",
+  #   "mob_ep_other", "pwn", "lg_zoo", "lg_inf", "phytoben")
+  # nonSampled <- subset(fgs, isfished == 0 | grouptype %in% nonsampledtypes |
+  #   code == "REP")
+  # notenumeratedtypes <- c("bird", "mammal", "cep", "sed_ep_other", "lg_zoo",
+  #   "lg_inf", "phytoben")
+  # notEnum <- subset(fgs, grouptype %in% notenumeratedtypes | code == "BFF")
+  #
+  # dat <- dat[!(dat$Predator %in% nonSampled), !(colnames(dat) %in% notEnum)]
+  #
+  # # add uniform error to half of the "observations"
+  # nPreyObs <- NROW(dat) * (NCOL(dat)-1)
+  # for(obs in 1:(nPreyObs/2)){
+  #   # determine row and column indices
+  #   rowR <- sample(1:NROW(dat), 1)
+  #   colC <- sample(2:NCOL(dat), 1)
+  # 
+  #   dat[rowR, colC] <- dat[rowR, colC] + runif(1, -0.1, 0.1)
+  # }
+  # 
+  # # add bias by removing little-observed prey at random
+  # for(i in 1:NROW(dat)){
+  #   for(j in 2:NCOL(dat)){
+  #     if(dat[i,j] < 0.25 & runif(1) < 0.15){
+  #       dat[i,j] <- 0
+  #     }
+  #   }
+  # }
+  # 
+  # # recalibrate so that rows add to 1
+  # # first need to adjust/account for negative values
+  # baseAdd <- min(dat[,2:NCOL(dat)])
+  # if(baseAdd < 0){
+  # 
+  #   for(i in 1:NROW(dat)){
+  #     for(j in 2:NCOL(dat)){
+  #       if(dat[i,j] != 0){
+  #         dat[i,j] <- dat[i,j]-baseAdd
+  #       }
+  #     }
+  #   }
+  # 
+  #   }
+  # 
+  # 
+  #   for(r in 1:NROW(dat)){
+  # 
+  #     denom <- rowSums(dat[r,2:NCOL(dat)])
+  #     dat[r,2:NCOL(dat)] <- (dat[r,2:NCOL(dat)])/denom
+  #   }
+
+```
+
+```{r, test-samplediet}
+
+```

### 2020-11-18 R/Dirichlet_2019.r

@@ -1,0 +1,125 @@
+# Code from Kerim Aydin for diet sampling in AK, 18 Nov 2020
+
+setwd("C:/src/octo2016")
+              
+require(VGAM)
+
+rawdat   <- read.csv("BS_allPrey_out.csv")
+preylook <- read.csv("Ecopath_diet_lookup_octo.csv")
+preycol <- "OCTO"
+
+LCLASS <- c(0,40,60,150)
+preds    <- list("P.Cod" = "P. Cod")
+lenlist  <- c("[0,40)", "[40,60)", "[60,150)")
+yrs      <- c(1984:2011,2014:2015)
+DOMS     <- c("InnerNW","InnerSE","MiddleNW","MiddleSE","OuterNW","OuterSE")
+
+DETECT <- 0.0001 # Grams of empty prey for proportioning
+SAMPLES <- 10000
+
+# Build a crosstab-query (per-predator) using names in GOA_MED column
+  guild <- as.character(preylook[,preycol])
+  names(guild) <- as.character(preylook$PREYNAME)
+  glist <- guild[as.character(rawdat$ECOPATH_PREY)]     
+  gdat <- aggregate(rawdat$TWT,list(as.character(rawdat$PREDJOIN),glist),sum)
+  gtab <- tapply(rawdat$TWT,list(as.character(rawdat$PREDJOIN),glist),sum)
+  gtab[is.na(gtab)]<-0
+  predtab <- unique(rawdat[,1:23])
+  preycross <- gtab[as.character(predtab$PREDJOIN),]
+  allprey   <- colnames(preycross)
+  lbin  <- cut(predtab$PRED_LEN,LCLASS,right=F)
+  GDAT<-cbind(predtab,lbin,preycross)
+
+# Halibut mm to cm regression for a (6.291e-6)*10^3.24 =  0.01093251
+A_L <- c("W.Pollock"=0.00553096, "P.Cod"=0.00411781, "Arrowtooth"=0.00443866,"P.Halibut"=0.01093251)
+B_L <- c("W.Pollock"=3.044172,   "P.Cod"=3.25325765, "Arrowtooth"=3.19894001,"P.Halibut"=3.24)
+
+#FinalEst  <- NULL
+Nests     <- length(yrs)*length(preds)*length(lenlist)*length(DOMS)
+alphalist <- matrix(NA,Nests,length(allprey))
+colnames(alphalist)<-allprey
[…]

### 2020-11-18 SurveyDietCompTest.Rmd
aggregate survey diet dat

@@
 I would rather have a function that could be used on any "survey" so have already removed hardcoded predator selection.
-I think also the distributional assumptions need revision; a Dirichlet seems most appropriate to simulate continuous proportions.
+I think also the distributional assumptions need revision; a Dirichlet is most appropriate to simulate continuous proportions.
 
@@
 
-  # read in dat, long format output of load_diet_comp, or create_survey_diet, both have layers aggregated
+  # dat can be long format output of load_diet_comp, or create_survey_diet, both have layers aggregated
+  # load_diet_comp output has no polygon column and atoutput is proportion, 
+  # create_survey_diet output has polygon column and atoutput is consumption in tons
+  # first aggregate to get global diet comp in proportion, if not already in that form
 
-  # load_diet_comp has no polygon information and is proportion, create_survey_diet has polygons and is consumption in tons
+  if("polygon" %in% names(dat)){
+    #sum over boxes and assume sampling occurs coastwide (the sampled boxes were already subset in create functions)
+    dat2 <- aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time, dat$prey),sum)
+    names(dat2) <- c("species","agecl","time","prey", "sumcons")
 
-  # first aggregate to global diet comp in proportion
+  }
 
-  # then apply aggregation bias
+  # then apply id/aggregation bias--this adds prey categories that don't exist??
 
@@
 
+# global proportion diet comp for comparison with survey_cons
+diettest <- load_diet_comp(here("atlantisoutput", "NOBA_march_2020"), 
+                           "nordic_runresults_01DietCheck.txt", fgs = fgs) %>%
+  filter(atoutput>0)
+
 ```

### 2020-11-19 SurveyDietCompTest.Rmd
diet prop from survey, outline bias addition

@@
 
+    dat2 <- dat2 %>%
+      group_by(species, agecl, time) %>%
+      mutate(totcons = sum(sumcons)) %>%
+      mutate(dietprop = sumcons/totcons) %>%
+      arrange(species, agecl, time)
+
   }
@@
   # then apply id/aggregation bias--this adds prey categories that don't exist??
+  # user specified max proportion to reallocate
+  # apply taxonomically, allocating fish prey to fish unid, invert prey to invert unid, then some portion unid
+  # phytoplankton and zooplankton prey direct to unid?
 

### 2020-11-20 SurveyDietCompTest.Rmd
scaling diet and sampling with rdirichlet

@@
 
-sample_diet <- function(dat, fgs) {
+sample_diet <- function(dat, fgs, alphamult = 10000000) {
 
@@
   # create_survey_diet output has polygon column and atoutput is consumption in tons
-  # first aggregate to get global diet comp in proportion, if not already in that form
+  # first aggregate to get global diet comp by species, agecl, time in proportion, if not already in that form
 
@@
 
+  # aggregate over age? which ones? leave as age class specific?
+
   # then apply id/aggregation bias--this adds prey categories that don't exist??
@@
   # see e.g., https://rdrr.io/cran/DirichletReg/man/Dirichlet.html
+  # alpha vector is the diet comp scaled by a user-supplied multiplier
+  # default to a very high multiplier for close-to-true diet comp
+  # lower multipliers stray further from true diet comp
+
+  # testdiet <- c(.1, .32, .06, .02, .49, .01)
+  # library(DirichletReg)
+  # rdirichlet(10, 10000000*testdiet)
+  # rdirichlet(10, 100*testdiet)
+
+  alphavec <- alphamult *
+
+  sampdiet <- DirichletReg::rdirichlet(1, alphavec)
 

### 2020-11-20 SurveyDietCompTest.Rmd
renaming for consistency

@@
 
-The current partial `sample_diet()` function needs to change because it expects a wide format file of diet proportions, rather than a long-format file of consumption in tons which is the output of `create_survey_diet()`.
+The intial partial `sample_diet()` function in `atlantisom` needs to change because it expects a wide format file of diet proportions, rather than a long-format file of consumption in tons which is the output of `create_survey_diet()`.
 
@@
 
-The current function removes predators not expected on a fish trawl survey, removes prey types that are normally not counted, then adds uniform error to half the remaining diet proportions, removes a random subet of prey accounting for <0.25 diet proportion, then renormalizes.
+The initial function removed predators not expected on a fish trawl survey, removed prey types that are normally not counted, then adds uniform error to half the remaining diet proportions, removes a random subet of prey accounting for <0.25 diet proportion, then renormalizes.
 I would rather have a function that could be used on any "survey" so have already removed hardcoded predator selection.
+
 I think also the distributional assumptions need revision; a Dirichlet is most appropriate to simulate continuous proportions.
+See [Douma and Weedon 2019](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210X.13234) for an excellent overview, as well as [a description of the Dirichlet alpha parameter](https://stats.stackexchange.com/questions/244917/what-exactly-is-the-alpha-in-the-dirichlet-distribution) and [Lin 2016](https://mast.queensu.ca/~communications/Papers/msc-jiayu-lin.pdf).
+
+Here I'll use the random number generation function `rdirichlet()` in the package `DirichletReg`but others may work too.
+
+```{r, test-samplediet}
+
+# global proportion diet comp for comparison with survey_cons
+diettest <- load_diet_comp(here("atlantisoutput", "NOBA_march_2020"), 
+                           "nordic_runresults_01DietCheck.txt", fgs = fgs) %>%
+  filter(atoutput>0) %>%
+  filter(species %in% survspp)
+
+```
 
 ```{r, scratch-samplediet, eval=FALSE}
+
+dat <- survey_cons
+#dat <- diettest
 
@@
 
-  }
+    # rename to match output of load_diet_comp
+    names(dat2)[names(dat2) == 'dietprop'] <- 'atoutput'
+    names(dat2)[names(dat2) == 'time'] <- 'time.days'
+
+  } else dat2 <- dat
 
@@
 ```
-
-```{r, test-samplediet}
-
-# global proportion diet comp for comparison with survey_cons
-diettest <- load_diet_comp(here("atlantisoutput", "NOBA_march_2020"), 
-                           "nordic_runresults_01DietCheck.txt", fgs = fgs) %>%
-  filter(atoutput>0)
-
-```

### 2020-11-30 SurveyDietCompTest.Rmd
apply dirichlet to diet, start taxonomic bias

@@
 
+  # use grouptype column to allocate
+  colnames(fgs) <- tolower(colnames(fgs))
+
+  # check for GroupType or InvertType
+  if (!("grouptype" | "inverttype") %in% colnames(fgs)) {
+    stop(paste("The columns GroupType or InvertType ars not in your functional\n",
+      "groups file."))
+  }
+
+  # change inverttype to grouptype, contents should be the same
+  if("inverttype" %in% names(fgs)) names(fgs)[names(fgs) == 'inverttype'] <- 'grouptype'
+
+  fgs$grouptype <- tolower(fgs$grouptype)
+
   # then apply Dirichlet obs error; Dirichlet equivalent of rmultinom(1, effN, trueagecomp)?   
@@
 
-  alphavec <- alphamult *
-
-  sampdiet <- DirichletReg::rdirichlet(1, alphavec)
+	#tidy
+	dietprop <- dat2 %>%
+	  group_by(species, time.days, agecl) %>%
+	  do(mutate(., dietSamp = DirichletReg::rdirichlet(1, alphamult * atoutput)))
 

### 2020-12-01 SurveyDietCompTest.Rmd
outline bias steps WIP

@@
 
-sample_diet <- function(dat, fgs, alphamult = 10000000) {
+sample_diet <- function(dat, fgs, unidprey = 0, alphamult = 10000000) {
 
@@
   fgs$grouptype <- tolower(fgs$grouptype)
+
+  # associate prey categories with unid categories
+  fgs<- fgs%>% 
+       mutate(unidtype = case_when((grouptype %in% c("mammal", "bird")) ~ "Unid_Vert",
+                                   (grouptype %in% "fish") ~ "Unid_Fish",
+                                   (grouptype %in% c("mammal")) ~ "Unid_Invert",
+                                   (grouptype %in% c("lg_zoo", "med_zoo")) ~ "Unid_Plankton",
+                                          TRUE ~ 0))
+
+  # associate prey with prey category
+
+  # add rows with unid categories
+
+  # randomly assign up to unidprey % of prey in each prey category to associated unid category
 

### 2020-12-03 SurveyDietCompTest.Rmd
random proportion to unid category

--- 2020-11-30 SurveyDietCompTest.Rmd
@@
   # check for GroupType or InvertType
-  if (!("grouptype" | "inverttype") %in% colnames(fgs)) {
+  if (!("grouptype" %in% colnames(fgs) | "inverttype"%in% colnames(fgs))) {
     stop(paste("The columns GroupType or InvertType ars not in your functional\n",
@@
   fgs$grouptype <- tolower(fgs$grouptype)
+
+  # associate prey categories (atlantis user guide I Table 8) with unid categories
+  fgs<- fgs%>% 
+       mutate(unidtype = case_when((grouptype %in% c("mammal", 
+                                                     "bird", 
+                                                     "reptile")) ~ "Unid_Vert",
+                                   (grouptype %in% c("fish", 
+                                                     "shark")) ~ "Unid_Fish",
+                                   (grouptype %in% c("fish_invert", 
+                                                     "cep", 
+                                                     "pwn", 
+                                                     "lg_inf",
+                                                     "sm_inf",
+                                                     "sed_ep_ff",
+                                                     "sed_ep_other",
+                                                     "mobile_ep_other",
+                                                     "coral",
+                                                     "sponge")) ~ "Unid_Invert",
+                                   (grouptype %in% c("lg_zoo", 
+                                                     "med_zoo", 
+                                                     "sm_zoo",
+                                                     "dinoflag",
+                                                     "lg_phy",
+                                                     "sm_phy")) ~ "Unid_Plankton",
+                                   TRUE ~ "Unid"))
+
+  # associate prey with prey category
+  # randomly assign up to unidprey % of prey in each prey category to associated unid category
+
+  fgsprey <- fgs[,c("name", "unidtype")]
+  names(fgsprey)[names(fgsprey) == 'name'] <- 'prey'
+
+  dat2 <- left_join(dat2, fgsprey) %>%
+    group_by(species, time.days, agecl, prey) %>%
+    mutate(unidprop = runif(1, 0, unidprey*atoutput))
+
+  # add rows with unid categories and sum prop assigned to the category
 

### 2020-12-04 SurveyDietCompTest.Rmd
sum unid prey

@@
   # randomly assign up to unidprey % of prey in each prey category to associated unid category
+  # take remainder for identified prey
 
@@
 
-  dat2 <- left_join(dat2, fgsprey) %>%
-    group_by(species, time.days, agecl, prey) %>%
-    mutate(unidprop = runif(1, 0, unidprey*atoutput))
+  dat3 <- left_join(dat2, fgsprey) %>%
+    rowwise() %>%
+    mutate(unidprop = runif(1, 0, unidprey*atoutput),
+           sampprop = atoutput - unidprop)
 
   # add rows with unid categories and sum prop assigned to the category
+
+  unidrows <- dat3 %>%
+    #select(-atoutput, -prey, -sampprop) %>%
+    group_by(species, time.days, agecl, unidtype) %>%
+    summarise(unidtot = sum(unidprop)) %>%
+    ungroup() %>%
+    rename(prey = unidtype, sampprop = unidtot)
 

### 2020-12-04 SurveyDietCompTest.Rmd
draft sample_diet seems to work

@@
 #dat <- diettest
+unidprey <- 0.5 #0 makes none unid, 1 could be all, reflects max proportion to unid
+alphamult <- 10000000 #default is almost identical to true, smaller more variable
 
@@
 
-  dat3 <- left_join(dat2, fgsprey) %>%
+  dat2 <- left_join(dat2, fgsprey) %>%
     rowwise() %>%
@@
 
-  unidrows <- dat3 %>%
+  unidrows <- dat2 %>%
     #select(-atoutput, -prey, -sampprop) %>%
@@
     rename(prey = unidtype, sampprop = unidtot)
+
+  dat2 <- dat2 %>%
+    select(species, time.days, agecl, prey, sampprop) %>%
+    full_join(unidrows)
 
@@
 	#tidy
-	dietprop <- dat2 %>%
+	dat2 <- dat2 %>%
+	  filter(sampprop>0) %>%
 	  group_by(species, time.days, agecl) %>%
-	  do(mutate(., dietSamp = DirichletReg::rdirichlet(1, alphamult * atoutput)))
+	  do(mutate(., dietSamp = DirichletReg::rdirichlet(1, alphamult * sampprop))) %>%
+    ungroup() %>%
+    select(-sampprop)
 
-  # returns diet proportions with observation error: Dirichlet, and bias from unknown id and/or family level aggregation
+  # returns diet proportions with observation error: Dirichlet, and bias from unknown id
 
-  return(dietprop)
+  return(dat2)
 }

### 2020-12-07 SurveyDietCompTest.Rmd
function works on both input types
add comparison plots

@@
 
-## Apply survey design to diet outputs
+## Apply survey design to diet outputs--develop functions and wrapper
 
@@
 
+## Test `create_survey_diet()`
+
 ```{r, test-createsurveydiet}
@@
 ```
+
+## Create `sample_diet()`
 
@@
 ```
+
+## Test `sample_diet()`
+
+Testing the new `sample_diet()` function with both surveyed diet (survey_cons, output of create_survey_diet) and global diet (diettest, output of load_diet_comp), with both unidentified prey bias and Dirichlet observation error.
+
+```{r test-sample-diet}
+
+samp1 <- sample_diet(survey_cons, fgs, unidprey = 0, alphamult = 10000000)
+
+samp2 <- sample_diet(diettest, fgs, unidprey = 0.5, alphamult = 10000000)
+
+samp3 <- sample_diet(survey_cons, fgs, unidprey = 0.3, alphamult = 10000000)
+
+samp4 <- sample_diet(survey_cons, fgs, unidprey = 0.3, alphamult = 10000)
+
+samp5 <- sample_diet(diettest, fgs, unidprey = 0, alphamult = 10000000)
+
+```
+
+Plotting functions and colors for everyone
+
+```{r color-plotfun}
+
+# length(unique(samp2$prey)) #39 prey categories max 
+# 35 categories in samp5 without unid length(unique) #http://medialab.github.io/iwanthue/
+preycol <- c("#7b7927",
+             "#746dd8",
+             "#a6bc3a",
+             "#4f3587",
+             "#86af42",
+             "#c666c7",
+             "#42c87f",
+             "#cb417f",
+             "#6fb95c",
+             "#882e7b",
+             "#67b271",
+             "#c83c63",
+             "#43c8ac",
+             "#da4953",
+             "#3ba7e5",
+             "#caa432",
+             "#5081db",
+             "#c8802f",
+             "#6c79c2",
+             "#c7b961",
+             "#bb86d6",
+             "#3b7125",
+             "#d672b9",
+             "#b57e43",
+             "#872957",
+             "#dd6741",
+             "#db75a2",
+             "#9e3d14",
+             "#db6f81",
+             "#7b3015",
+             "#e08262",
+             "#8d2a41",
+             "#d56467",
+             "#8c242e",
+             "#a73830")
+names(preycol) <- as.factor(sort(unique(samp5$prey)))
+
+# going for more greyscale for unident categories, same website
+unidcol <- c("#b8b8b2",
+             "#302a1d",
+             "#6b7069",
+             "#1e3430")
+names(unidcol) <- as.factor(c("Unid", "Unid_Fish",  "Unid_Invert", "Unid_Plankton"))
+
+col <- c(preycol, unidcol)
+
+# plot diet comp over time at age by species
+plotdiet <- function(dat, compdat=NULL, namedat, namecomp=NULL){
+
+  dat <- dat %>% add_column(run = namedat)
+  if(!is.null(compdat)) compdat <- compdat %>% add_column(run = namecomp)
+
+    ggplot() +
+    geom_bar(data=dat, aes(time.days/365, dietSamp, fill=prey), stat = "identity") +
+    {if(!is.null(compdat)) geom_bar(data=compdat, aes(time.days/365, dietSamp, fill=prey), stat = "identity")} + 
+    theme_tufte() +
+    theme(legend.position = "bottom") +
+    xlab("year") +
+    ylab("diet proportion") +
+    facet_grid(agecl~run) + 
+    scale_fill_manual(values=col) + 
+    ggtitle(dat$species)
+
+}
+
+# method for a single species diet, no comparisons
+# plist = lapply(split(ms_diet, ms_diet$species), function(d) {
+#   ggplot(d, aes(time.days/365, atoutput, fill=prey)) + 
+#     geom_bar(stat = "identity") +
+#     facet_wrap(species~agecl) +
+#     xlab("year") +
+#     ylab("diet proportion") +
+#     theme_tufte() +
+#     theme(legend.position="bottom")
+# })
+
+```
+
+## Visualize diet comparisons
+
+### Surveyed and Global diet {.tabset}
+
+Here we compare the surveyed diet comp with the global diet comp, both with no prey bias (unidprey = 0) and extremely low Diriclet observation error (alphamult = 10000000):
+
+```{r, fig.width=9, fig.height=12, results='asis'}
+
+preds <- unique(samp1$species)
+
+for(i in 1:length(preds)) {
+  cat("  \n####",  as.character(preds[i]),"  \n")
+  print(plotdiet(dat = filter(samp1, species %in% preds[i]), namedat = "Survey diet", compdat = filter(samp5, species %in% preds[i]), namecomp = "Global diet")) 
+  cat("  \n")
+}
+
+```
+
+### {-}
+
+### Unbiased and biased global diet {.tabset}
+
+Here we compare the global diet comp without and with unidentified prey bias (unidprey = 0 and unidprey = 0.5, or up to 50% of each prey proportion could be set to unidentified) using extremely low Diriclet observation error (alphamult = 10000000):
+
+```{r, fig.width=9, fig.height=12, results='asis'}
+
+for(i in 1:length(preds)) {
+  cat("  \n####",  as.character(preds[i]),"  \n")
+  print(plotdiet(dat = filter(samp5, species %in% preds[i]), namedat = "Global diet", compdat = filter(samp2, species %in% preds[i]), namecomp = "Biased global diet")) 
+  cat("  \n")
+}
+```
+
+### {-}
+
+### Unbiased and biased survey diet {.tabset}
+
+Here we compare survey diet comp with no bias and variance, with some bias (unidprey = 0.3 or up to 30% of each prey proportion could be set to unidentified) and extremely low to moderate Dirichlet observation error (alphamult = 10000000 and alphamult = 10000):
+
+```{r, fig.width=9, fig.height=12, results='asis'}
+
+for(i in 1:length(preds)) {
+  cat("  \n####",  as.character(preds[i]),"  \n")
+  print(plotdiet(dat = filter(samp1, species %in% preds[i]), namedat = "Survey diet", compdat = filter(samp3, species %in% preds[i]), namecomp = "Biased survey diet")) 
+  cat("  \n")
+}
+```
+
+### {-}
+
+### Biased survey diet, low and high observation error {.tabset}
+
+```{r, fig.width=9, fig.height=12, results='asis'}
+
+for(i in 1:length(preds)) {
+  cat("  \n####",  as.character(preds[i]),"  \n")
+  print(plotdiet(dat = filter(samp3, species %in% preds[i]), namedat = "Biased survey diet", compdat = filter(samp4, species %in% preds[i]), namecomp = "Bias and variance survey diet")) 
+  cat("  \n")
+}
+```
+
+### {-}

### 2020-12-07 docs/index.Rmd
update index

--- 2020-10-22 docs/index.Rmd
@@
 
-### Full age structured data for multispecies modeling, June-October 2020
+### Full age structured data for multispecies modeling, June-December 2020
+
+<a href="https://sgaichas.github.io/poseidon-dev/SurveyDietCompTest.html" target="_blank">Apply survey specifications to diet outputs, December 7, 2020</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/msSurveysTest.html" target="_blank">Compare multispecies survey specifications, November 9, 2020</a>
 

### 2020-12-08 SurveyDietCompTest.Rmd
add comparison of low and high variance without bias

@@
 
-samp4 <- sample_diet(survey_cons, fgs, unidprey = 0.3, alphamult = 10000)
+samp4 <- sample_diet(survey_cons, fgs, unidprey = 0.3, alphamult = 10)
 
 samp5 <- sample_diet(diettest, fgs, unidprey = 0, alphamult = 10000000)
+
+samp6 <- sample_diet(survey_cons, fgs, unidprey = 0, alphamult = 10)
 
@@
 
-Here we compare survey diet comp with no bias and variance, with some bias (unidprey = 0.3 or up to 30% of each prey proportion could be set to unidentified) and extremely low to moderate Dirichlet observation error (alphamult = 10000000 and alphamult = 10000):
+Here we compare survey diet comp with no bias and variance (unidprey = 0 and alphamult = 10000000) to survey diet with some bias (unidprey = 0.3 or up to 30% of each prey proportion could be set to unidentified, and alphamult = 10000000):
 
@@
 
-### Biased survey diet, low and high observation error {.tabset}
+### Unbiased survey diet, low and high observation error {.tabset}
+Here we compare survey diet comp with no bias or variance (unidprey = 0 and alphamult = 10000000), with no bias and high Dirichlet observation error (unidprey = 0 and alphamult = 10):
 
@@
   cat("  \n####",  as.character(preds[i]),"  \n")
-  print(plotdiet(dat = filter(samp3, species %in% preds[i]), namedat = "Biased survey diet", compdat = filter(samp4, species %in% preds[i]), namecomp = "Bias and variance survey diet")) 
+  print(plotdiet(dat = filter(samp1, species %in% preds[i]), namedat = "No variance survey diet", compdat = filter(samp6, species %in% preds[i]), namecomp = "High variance survey diet")) 
+  cat("  \n")
+}
+
+```
+
+### {-}
+
+### Biased survey diet, low and high observation error {.tabset}
+Last we compare survey diet comp with some bias (unidprey = 0.3 or up to 30% of each prey proportion could be set to unidentified) and low observation error (alphamult = 1000000) to the same bias and high Dirichlet observation error (alphamult = 10):
+
+```{r, fig.width=9, fig.height=12, results='asis'}
+
+for(i in 1:length(preds)) {
+  cat("  \n####",  as.character(preds[i]),"  \n")
+  print(plotdiet(dat = filter(samp3, species %in% preds[i]), namedat = "Biased survey diet", compdat = filter(samp4, species %in% preds[i]), namecomp = "Bias and high variance survey diet")) 
   cat("  \n")

### 2020-12-18 SurveyDietCompTest.Rmd
test wrapper

@@
                    n_reps = n_reps,
+                   unidprey = 0,
+                   alphamult = 10000000,
                    save = TRUE){
@@
     for(i in 1:n_reps){
-      survObsDiet[[i]] <- atlantisom::sample_diet(survey_diet, surv_cv)
+      survObsDiet[[i]] <- atlantisom::sample_diet(survey_cons, fgs, unidprey = unidprey, alphamult = alphamult)
     }

### 2020-12-18 SurveyDietCompTest.Rmd
test new wrapper and compare surveys

@@
                    dietfile = file_diet,
-                   fgs = fgs,
                    usersurvey = usersurvey_file,
-                   userfishery = NULL,
                    omlist_ss,
                    n_reps = n_reps,
-                   unidprey = 0,
-                   alphamult = 10000000,
                    save = TRUE){
+
+  source(config)
+
+  #Load functional groups
+  fgs <- atlantisom::load_fgs(dir=d.name,
+                                       file_fgs = functional.groups.file)
 
@@
   }
+
+  return(survObsDiets)
 
@@
 ### {-}
+
+## Test wrapper
+
+The new `om_diet` function should load the detailed diet file, run the survey, then sample the diets and save outputs for each survey. This takes 3 minutes to run with two surveys and 10 species.
+
+```{r testwrapper, eval=FALSE}
+
+NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+
+NOBAom_ms_diet <- om_diet(config = here("config", "NOBA2config.R"),
+                          dietfile = "NOBADetDiet.gz",
+                          usersurvey = c(here("config/mssurvey_spring_01.R"), 
+                                         here("config/mssurvey_fall_01.R")), 
+                          omlist_ss = NOBAom_ms, 
+                          n_reps = 1, 
+                          save = TRUE)
+
+```
+
+### Compare spring and fall surveys {.tabset}
+
+The diet was sampled with alphamult = 10 and unidprey = 0.3 using the survey configuration files.
+
+```{r, fig.width=9, fig.height=12, results='asis'}
+
+NOBAom_ms_diet <- read_savedsurvs(d.name, 'survDiet')
+
+preds <- unique(unique(NOBAom_ms_diet[[2]][[1]]$species))
+
+fall <- as.data.frame(NOBAom_ms_diet$BTS_fall_nearbox_qmix_selmix[[1]])
+spring <- as.data.frame(NOBAom_ms_diet$BTS_spring_nearbox_qmix_selmix[[1]])
+
+for(i in 1:length(preds)) {
+  cat("  \n####",  as.character(preds[i]),"  \n")
+  print(plotdiet(dat = filter(spring, species %in% preds[i]), namedat = "Spring survey diet", compdat = filter(fall, species %in% preds[i]), namecomp = "Fall survey diet")) 
+  cat("  \n")
+}
+
+```
+
+### {-}

### 2020-12-18 config/mssurvey_fall_01.R

@@ -156,3 +156,7 @@ lenage_cv <- 0.1
 # max size bin for length estimation, function defaults to 150 cm if not supplied
 maxbin <- 200
 
+# diet sampling parameters
+alphamult <- 10
+unidprey <- 0.3
+

### 2020-12-18 config/mssurvey_spring.R

@@ -58,4 +58,3 @@ lenage_cv <- 0.1
 
 # max size bin for length estimation, function defaults to 150 cm if not supplied
 maxbin <- 200
-

### 2020-12-18 config/mssurvey_spring_01.R

@@ -157,3 +157,7 @@ lenage_cv <- 0.1
 # max size bin for length estimation, function defaults to 150 cm if not supplied
 maxbin <- 200
 
+# diet sampling parameters
+alphamult <- 10
+unidprey <- 0.3
+

### 2021-01-29 Visualize_atlantisom.Rmd
changed year range in MSE paper figure

@@
                          shape = c(16, NA)))) + 
-  xlim(0,80) +
+  xlim(20,80) +
   theme_tufte() +
@@
                          shape = c(NA, 16)))) + 
-  xlim(0,80) +
+  xlim(20,80) +
   theme_tufte() +

### 2021-05-05 NOBAcod.Rmd
modify NOBAcod script to create SSmodel

@@
 require(r4ss)
-omlist_ss<-CC3om_sardine
-source(here("/config/CC3Config_sardLWcorr.R"))
-source(here("/config/omdimensions.R"))
-source(here("/config/sardinesurvey.R"))
-source(here("/config/sardinefishery.R"))
+#omlist_ss<- ?
+source(here("config/omdimensions.R"))
+source(here("/config/codsurvey.R"))
+source(here("/config/codfishery.R"))
 
 #Directory with SS files
-model_dir <- "Sardine_SS_files/"
+model_dir <- "NOBA_cod_files/"
 
 #Name of SS data file
-datfile_name <- "sardEM_3_3.dat"
+datfile_name <- "ss3.dat"
 
@@
 
+#Load in model object from SAM
+noba_cod_out <- system.file("inst","extdata","NOBA_cod_files","NEAcod-2020","run","model.RData",
+                            package = "atlantisom") 
+#This is the SAM object
+noba_fit <- noba_cod_out$fit
+
+properties <- c("minAge","maxAge", "maxAgePlusGroup")
+
+age_info <- lapply(properties, get, noba_fit$conf)
+names(age_info) <- properties
+
+#install SAM
+remotes::install_github("fishfollower/SAM/stockassessment")
+require(stockassessment)
+
 #Month of survey/fishing
-survey_month <- 7
-fishing_month <- 1
+#survey_month <- 7
+#fishing_month <- 1
 

### 2021-05-06 NOBAcodSAMtoSS.R
add script to convert fromSAMmodel to SS

@@ -1,0 +1,31 @@
+#Directory with SS files
+model_dir <- "NOBA_cod_files/"
+
+#Name of SS data file
+datfile_name <- "ss3.dat"
+
+#CVs for length at age, catch, and survey
+CVs <- list("lenage"=0.1, "fishery"=0.01, "survey"=0.1)
+
+#Load in model object from SAM
+noba_cod_out <- system.file("inst","extdata","NOBA_cod_files","NEAcod-2020","run","model.RData",
+                            package = "atlantisom") 
+#This is the SAM object
+noba_fit <- noba_cod_out$fit
+
+noba_fit
+
+
+properties <- c("minAge","maxAge", "maxAgePlusGroup")
+
+age_info <- lapply(properties, get, noba_fit$conf)
+names(age_info) <- properties
+
+idx <- which(names(noba_fit$sdrep$value)=="logfbar")
+names(noba_fit$sdrep$value)
+
+plot(exp(noba_fit$sdrep$value[idx]))
+
+#install SAM
+remotes::install_github("fishfollower/SAM/stockassessment")
+require(stockassessment)

### 2021-05-06 NOBAcod.Rmd
add noba files and script

@@
                             package = "atlantisom") 
-#This is the SAM object
-noba_fit <- noba_cod_out$fit
-
-properties <- c("minAge","maxAge", "maxAgePlusGroup")
-
-age_info <- lapply(properties, get, noba_fit$conf)
-names(age_info) <- properties
-
-#install SAM
-remotes::install_github("fishfollower/SAM/stockassessment")
-require(stockassessment)
-
 #Month of survey/fishing

### 2021-05-06 NOBAcodSAMtoSS.R

@@ -1,31 +1,0 @@
-#Directory with SS files
-model_dir <- "NOBA_cod_files/"
-
-#Name of SS data file
-datfile_name <- "ss3.dat"
-
-#CVs for length at age, catch, and survey
-CVs <- list("lenage"=0.1, "fishery"=0.01, "survey"=0.1)
-
-#Load in model object from SAM
-noba_cod_out <- system.file("inst","extdata","NOBA_cod_files","NEAcod-2020","run","model.RData",
-                            package = "atlantisom") 
-#This is the SAM object
-noba_fit <- noba_cod_out$fit
-
-noba_fit
-
-
-properties <- c("minAge","maxAge", "maxAgePlusGroup")
-
-age_info <- lapply(properties, get, noba_fit$conf)
-names(age_info) <- properties
-
-idx <- which(names(noba_fit$sdrep$value)=="logfbar")
-names(noba_fit$sdrep$value)
-
-plot(exp(noba_fit$sdrep$value[idx]))
-
-#install SAM
-remotes::install_github("fishfollower/SAM/stockassessment")
-require(stockassessment)

### 2021-05-06 NOBAcodSAMtoSSdat.R

@@ -1,0 +1,32 @@
+#Directory with SS files
+model_dir <- "./NOBA_cod_files"
+#Name of SS data file
+#read in existing dat file (Atlantic cod from NE US)
+datfile_name <- "ss3.dat"
+dat_in <- r4ss::SS_readdat(file.path(model_dir,datfile_name))
+
+#Load in model object from SAM
+noba_cod_out <- get(load(file.path(model_dir,"NEAcod-2020","run","model.RData")))
+#This is the SAM object
+#Extract the estimated values
+noba_estimates <- broom::tidy(noba_cod_out$sdrep$value)
+
+#Get estimated F
+filter(noba_estimates, names=="logfbar")
+#Get estimated R
+filter(noba_estimates, names=="logR")
+#Get catch
+
+#Get the ages from the conf file
+properties <- c("minAge","maxAge", "maxAgePlusGroup")
+age_info <- lapply(properties, get, noba_cod_out$conf)
+names(age_info) <- properties
+
+dat_in$Nages <- length(age_info$minAge:age_info$maxAge)
+
+
+plot(exp(noba_fit$sdrep$value[idx]))
+
+#install SAM
+remotes::install_github("fishfollower/SAM/stockassessment")
+require(stockassessment)

### 2021-05-25 PerCapitaConsumption.Rmd
initial per capita consumption testing

---
title: "Per captia consumption"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)  
library(atlantisom)
library(ggthemes)
library(FSA)

```

## Diet info for multispecies modeling

Multispecies models rely on either estimated or input parameters governing predator consumption rates to estimate predation mortality on consumed species.

Stomach weights are needed as input to the length-based multispecies model hydra. At present each species has a stomach weight for each size bin that is repeated for each year. We have total consumed weight for each predator age class at each timestep in the detailed diet atlantis output, so would need to map age classes to length bins, sum consumed weight by length bin and divide by total numbers in that length bin to get stomach weight.

What we have saved so far is a diet comp in proportion that has had bias and observation error added to proportions. However, we need the stomach weight portion of the equation, which has the survey timing/area/efficiency/selectivity applied to true consumed weight. The hydra input is mean daily stomach weight in grams. From the Atlantis manual, "DetailedDietCheck.txt returns the total consumed biomass since the last output given for each cell (box and layer) of each age group of each functional group." So daily per capita consumed biomass is this output, summed over prey for total consumption, divided by the numbers from the same survey design, divided by the output step `omlist_ss$runpar$outputstep`.

The function below should return per-capita consumption by age class. This can then be converted to lengthbin as necessary.

```{r stomwt, eval=FALSE}

om_cons<- function(config = configfile,
                   dietfile = file_diet,
                   usersurvey = usersurvey_file,
                   omlist_ss,
                   n_reps = n_reps,
                   save = TRUE){

  source(config)

  #Load functional groups
  fgs <- atlantisom::load_fgs(dir=d.name,
                              file_fgs = functional.groups.file)

  # load or read in saved detailed diet
  if(!file.exists(file.path(d.name,
                            paste0(scenario.name, "detaileddiet.rds")))){
    detaileddiet <- load_detailed_diet_comp(dir = d.name, 
                                            file_diet, 
                                            fgs = fgs)

    if(save){
      saveRDS(detaileddiet, file.path(d.name, paste0(scenario.name, "detaileddiet.rds")))
    }

  } else {
    detaileddiet <- readRDS(file.path(d.name,
                                      paste0(scenario.name, "detaileddiet.rds")))
  }

  #one script for dimension parameters to be used in multiple functions
  source("config/omdimensions.R", local = TRUE)

  survObsStomWt <- list()

  for (s in usersurvey)
  {
    source(s, local = TRUE)

    # survtime doesn't match units of time.days in detaileddiet
    survtime <- survey_sample_full*omlist_ss$runpar$outputstep

    # apply survey design to detailed diet
    survey_cons <- create_survey_diet(dat = detaileddiet,
                                      time = survtime,
                                      species = survspp,
                                      boxes = survboxes,
                                      effic = surveffic,
                                      selex = survselex)

    # get numbers at ageclass for same survey design
    # note different time units!
    survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
                                          time = survey_sample_full,
                                          species = survspp,
                                          boxes = survboxes,
                                          effic = surveffic,
                                          selex = survselex.agecl)

    # convert survey N times to cons times in days
    survey_N$time <- survey_N$time*omlist_ss$runpar$outputstep

    # get rid of polygon
    survey_totN <- survey_N %>%
      group_by(species, agecl, time) %>%
      summarise(totN = sum(atoutput)) %>%
      ungroup()

    # sum over prey to get total consumption in t, divide by N and timestep, convert to g
    survey_totcons <- survey_cons %>%
      group_by(species, agecl, time) %>%
      summarise(totcons = sum(atoutput)) %>%
      ungroup() %>%
      left_join(survey_totN) %>%
      mutate(percap_cons = totcons/totN) %>%
      #mutate(daily_percap_g = percap_cons/omlist_ss$runpar$outputstep*1000000)
      mutate(daily_percap_g = percap_cons*1000000) #try assuming cons is snapshot not cumulative since last timestep

    #save survey consumption, takes a long time to generate with lots of reps/species
    #if(save){
    saveRDS(survey_cons, file.path(d.name, paste0(scenario.name, "_",
                                                  survey.name, "surveycons.rds")))
    #}

    survObsStomWt[[survey.name]] <- survey_cons

  }

}

```

## NOBA

The configuration files for NOBA are in posiedon-dev/config and see also ms-keyrun project repo [illustrating full simulated multispecies modeling data](https://noaa-edab.github.io/ms-keyrun/SimData.html). The NOBA model output was producing fairly low per capita consumption using the code above, so we will compare that with what we get from the CC model to see if we have an output file problem across models or if this is a characteristic of the NOBA model.

```{r, eval=FALSE}

source(here("config/NOBA2Config.R"))

# apply surveys to true total consumption (snipped from atlantosom om_diet.R and om_comps.R)
detaileddiet <- readRDS(file.path(d.name, paste0(scenario.name, "detaileddiet.rds")))

usersurvey <- c(here("config/mssurvey_spring.R"),
                here("config/mssurvey_fall.R"),
                here("config/mssurvey_spring_01.R"),
                here("config/mssurvey_fall_01.R"))

omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))

source(here("config/omdimensions.R"))

```

## CCA

This is a short run of the CCA model implemented in the latest codebase, so it can produce full age structured output. First we need to process the detailed diet file, and also get the output of `run_truth` for numbers.

```{r, eval=FALSE}

source(here("config/CCConfig_constest.R"))

file_diet <- paste0(scenario.name,"DetailedDietCheck.txt")

#Load functional groups
  fgs <- atlantisom::load_fgs(dir=d.name,
                              file_fgs = functional.groups.file)

  # load or read in saved detailed diet
  if(!file.exists(file.path(d.name,
                            paste0(scenario.name, "detaileddiet.rds")))){
    detaileddiet <- load_detailed_diet_comp(dir = d.name, 
                                            file_diet, 
                                            fgs = fgs)

    if(save){
      saveRDS(detaileddiet, file.path(d.name, paste0(scenario.name, "detaileddiet.rds")))
    }

  } else {
    detaileddiet <- readRDS(file.path(d.name,
                                      paste0(scenario.name, "detaileddiet.rds")))
  }

```

### 2021-05-25 docs/index.Rmd
initial per capita consumption testing

@@
 
+### Diagnosing diet and consumption outputs, May 2021
+
+<a href="https://noaa-edab.github.io/poseidon-dev/PerCapitaConsumption.html" target="_blank"> Compare NOBA and CCA consumption outputs, May 25, 2021</a>
+
 ### Full age structured data for multispecies modeling, June-December 2020
+
+Code converting these outputs to multispecies production model and length structured model inputs is in development as of spring 2021 and is on the [ms-keyrun repository](https://github.com/NOAA-EDAB/ms-keyrun/blob/master/docs/SimData.Rmd).
+See also <a href="https://noaa-edab.github.io/ms-keyrun/SimData.html" target="_blank"> Initial simulated data, May 24, 2021</a>
 

### 2021-05-25 config/CCConfig_constest.R

@@ -1,0 +1,12 @@
+d.name <- here("atlantisoutput","CCconstest")
+functional.groups.file <- "CalCurrentV3Groups.csv"
+biomass.pools.file <- "DIVCalCurrentV3_Biol.nc"
+biol.prm.file <- "CalCurrentV3_Biol.prm"
+box.file <- "CalCurrentV3_utm.bgm"
+initial.conditions.file <- "DIVCalCurrentV3_Biol.nc"
+run.prm.file <- "CalCurrentV3_run.xml"
+scenario.name <- "outputCCV3"
+bioind.file <- "outputCCV3BiomIndx.txt"
+catch.file <- "outputCCV3Catch.txt"
+annage <- TRUE
+fisheries.file <- "CalCurrentV3Fisheries.csv"

### 2021-05-26 docs/index.Rmd
fix link

@@
 
-<a href="https://noaa-edab.github.io/poseidon-dev/PerCapitaConsumption.html" target="_blank"> Compare NOBA and CCA consumption outputs, May 25, 2021</a>
+<a href="https://sgaichas.github.io/poseidon-dev/PerCapitaConsumption.html" target="_blank"> Compare NOBA and CCA consumption outputs, May 25, 2021</a>
 

### 2021-06-14 NOBA_cod_files/Rscript/noba_cod.R
clean up rscript

@@ -96,13 +96,18 @@ for (i in 1:length(surveys)) {
 # SAM2SS --------------------------------------------------------------------------------------
 
 #### Recruitment age is 0 and bias adjustment in recruitment = TRUE
-output_path <- here::here("NOBA_cod_files", "output", "age0")
+i <- 1
+path_final <- c("age0","age0_slx20", "age1")
+output_path <- here::here("NOBA_cod_files", "output", path_final[i])
+slx_pattern <- c(12,20,12)
+start_age <- c(0,0,1)
+bias_adj <- c(TRUE, TRUE, FALSE)
 
 saconvert::ICES2SS(
   user.wd = sam_input_path,
   user.od = output_path,
   ices.id = "",
-  slx = 12,
+  slx = slx_pattern[i],
   tvslx = FALSE,
   ages = NULL,
   nsexes = 1, # 1: one sex; 2: two sex; -1: one sex and multiply the spawning biomass by the fraction female in the control file
@@ -111,55 +116,9 @@ saconvert::ICES2SS(
   q.float = FALSE,
   sigma.init = 0.5,
   steep.init = 1,
-  start.age = 0,
-  bias.adj = TRUE
+  start.age = start_age[i],
+  bias.adj = bias_adj[i]
 ) # steep.init: http://sedarweb.org/docs/wpapers/SEDAR19_DW_06_SteepnessInference.pdf
-
-ss_ouput <- SS_output(dir = output_path, verbose = TRUE, printstats = TRUE)
-SS_plots(ss_ouput)
-
-#### Recruitment age is 0, bias adjustment in recruitment = TRUE, selectivity pattern = 20
-output_path <- here::here("NOBA_cod_files", "output", "age0_slx20")
-
-saconvert::ICES2SS(
-  user.wd = sam_input_path,
-  user.od = output_path,
-  ices.id = "",
-  slx = 20,
-  tvslx = FALSE,
-  ages = NULL,
-  nsexes = 1, # 1: one sex; 2: two sex; -1: one sex and multiply the spawning biomass by the fraction female in the control file
-  forN = 2,
-  q.extra.se = FALSE,
-  q.float = FALSE,
-  sigma.init = 0.5,
-  steep.init = 1,
-  start.age = 0,
-  bias.adj = TRUE
-) # steep.init: http://sedarweb.org/docs/wpapers/SEDAR19_DW_06_SteepnessInference.pdf
-
[…]

### 2021-06-14 noba_cod_files/Rscript/noba_cod_ccs.R

@@ -1,0 +1,113 @@
+# devtools::load_all("C:/Users/bai.li/Documents/saconvert")
+remotes::install_github("Bai-Li-NOAA/saconvert")
+
+
+if(here()){
+  args <- c("--no-multiarch")
+} else{
+  args <- c("")
+}
+
+remotes::install_github("r4ss/r4ss", branch ="development")
+devtools::install_github("fishfollower/SAM/stockassessment", args)
+
+
+library(here)
+library(stockassessment) # For using SAM
+library(r4ss) # For using SS
+library(saconvert)
+
+# Read SAM input data (NAE cod) ---------------------------------------------------------------
+
+
+sam_input_path <- here::here("NOBA_cod_files", "NEAcod-2020", "data")
+
+# Read in data files
+regdat <- grep(".dat",list.files(sam_input_path))
+filenames_ICES <- list.files(sam_input_path)[regdat]
+
+#All objects are now on input_file_list
+input_file_list <- lapply(file.path(sam_input_path,filenames_ICES), read.ices)
+names(input_file_list) <- gsub(".dat","",filenames_ICES)
+
+#Load into function environment as objects
+list2env(input_file_list, environment())
+
+# Create a collected data object
+sam_dat <- setup.sam.data(
+  surveys = survey,
+  residual.fleet = cn,
+  prop.mature = mo,
+  stock.mean.weight = sw,
+  catch.mean.weight = cw,
+  dis.mean.weight = dw,
+  land.mean.weight = lw,
+  prop.f = pf,
+  prop.m = pm,
+  natural.mortality = nm,
+  land.frac = lf
+)
+
+# Generate a default/minimalistic model configuration
+sam_conf <- defcon(sam_dat)
+
+# Generate default initial values for all our model parameters
+sam_par <- defpar(sam_dat, sam_conf)
+
+# Fit the SAM model
+# fit <- sam.fit(sam_dat, sam_conf, sam_par)
+
+# Plot figures 
+# par(mfrow=c(2, 2))
+# ssbplot(fit)
+# fbarplot(fit)
+# recplot(fit)
+# catchplot(fit)
+
+# SAM2SS --------------------------------------------------------------------------------------
+
+#### Recruitment age is 0 and bias adjustment in recruitment = TRUE
[…]

### 2021-06-15 NOBA_cod_files/Rscript/noba_cod.R
Update noba_cod.R: update scenarios and comparison code

@@ -1,23 +1,30 @@
+# Install and library packages --------------------------------------------------------
+
+# install.packages("here")
+#
+# install_64bit <- TRUE
+#
+# if(install_64bit){
+#   args <- c("--no-multiarch")
+# } else{
+#   args <- c("")
+# }
+#
+# devtools::install_github("fishfollower/SAM/stockassessment", INSTALL_opts=args)
+#
+# remotes::install_github("r4ss/r4ss", branch ="development")
+#
 # remotes::install_github("Bai-Li-NOAA/saconvert")
-
-
-if(here()){
-  args <- c("--no-multiarch")
-} else{
-  args <- c("")
-}
-
-remotes::install_github("r4ss/r4ss", branch ="development")
-devtools::install_github("fishfollower/SAM/stockassessment", args)
-
+#
+# remotes::install_github("nmfs-general-modeling-tools/nmfspalette")
 
 library(here)
 library(stockassessment) # For using SAM
 library(r4ss) # For using SS
 library(saconvert)
+library(nmfspalette)
 
 # Read SAM input data (NAE cod) ---------------------------------------------------------------
-
 
 sam_input_path <- here::here("NOBA_cod_files", "NEAcod-2020", "data")
 
@@ -63,103 +70,91 @@ fbarplot(sam_fit)
 recplot(sam_fit)
 catchplot(sam_fit)
 
-# Plot SAM selectivity
-par(mfrow = c(1,1))
-slex_age <- sam_conf$minAge:sam_conf$maxAge
-plot(slex_age, apply(cn, 2, mean, na.rm = TRUE), 
-     type="o", 
-     xlab="Age",
-     ylab="Number of fish")
-legend("topleft", "Catch",
-       bty="n")
+# SAM2SS --------------------------------------------------------------------------------------
+scenario_path <- c("A", "B")
+slx_pattern <- c(12, 12)
+f_method <- c(3, 2)
 
-par(mfrow = c(2, 2))
-slex_age <- sam_conf$minAge:sam_conf$maxAge
-survey_slex <- matrix(NA, ncol=length(surveys), nrow=length(slex_age))
-row.names(survey_slex) <- slex_age
 
-for (i in 1:length(surveys)) {
+for (scenario_id in seq_along(scenario_path)){
[…]

### 2021-06-15 NOBA_cod_files/README.Rmd
add README.Rmd to update README.MD

---
output: 
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
### NOBA Cod SAM2SS conversion
In this example, we convert NOBA Cod State-space assessment model (SAM) inputs to Stock Synthesis (SS) inputs and compare the outputs from the two packages.

### Install and library packages
```{r, warning = FALSE, message = FALSE}
# install.packages("here")
# 
# install_64bit <- TRUE
# 
# if (install_64bit) {
#   args <- c("--no-multiarch")
# } else {
#   args <- c("")
# }
# 
# devtools::install_github("fishfollower/SAM/stockassessment", INSTALL_opts = args)
# 
# remotes::install_github("r4ss/r4ss", branch = "development")
# 
# remotes::install_github("Bai-Li-NOAA/saconvert") # Need to install "kellijohnson-NOAA/saconvert" after doing pull request.
# 
# remotes::install_github("nmfs-general-modeling-tools/nmfspalette")

library(here)
library(stockassessment) # For using SAM
library(r4ss) # For using SS
library(saconvert)
library(nmfspalette)
```

### Read SAM input data and run SAM
```{r, warning = FALSE, message = FALSE}
sam_input_path <- here::here("NOBA_cod_files", "NEAcod-2020", "data")

# Read in data files
regdat <- grep(".dat",list.files(sam_input_path))
filenames_ICES <- list.files(sam_input_path)[regdat]

#All objects are now on input_file_list
input_file_list <- lapply(file.path(sam_input_path,filenames_ICES), read.ices)
names(input_file_list) <- gsub(".dat","",filenames_ICES)

#Load into function environment as objects
list2env(input_file_list, environment())

# Create a collected data object
sam_dat <- setup.sam.data(
  surveys = survey,
  residual.fleet = cn,
  prop.mature = mo,
  stock.mean.weight = sw,
  catch.mean.weight = cw,
  dis.mean.weight = dw,
  land.mean.weight = lw,
  prop.f = pf,
  prop.m = pm,
  natural.mortality = nm,
  land.frac = lf
)

# Generate a default/minimalistic model configuration
sam_conf <- defcon(sam_dat)

# Generate default initial values for all our model parameters
sam_par <- defpar(sam_dat, sam_conf)

# Fit the SAM model
sam_fit <- sam.fit(sam_dat, sam_conf, sam_par, silent=TRUE)

```

### Plot SAM fits
```{r, fig.height = 2, sam_fit}
par(mfrow = c(1, 1))
ssbplot(sam_fit)
fbarplot(sam_fit)
recplot(sam_fit)
catchplot(sam_fit)
```

### Convert SAM inputs to SS inputs
```{r, warning = FALSE, message = FALSE}
# Set up two scenarios
# Scenario A: use selectivity pattern 12 and F method 3 in SS
# Scenario B: use selectivity pattern 12 and F method 2 in SS

scenario_path <- c("A", "B")
slx_pattern <- c(12, 12)
f_method <- c(3, 2)

for (scenario_id in seq_along(scenario_path)){
  output_path <- here::here("NOBA_cod_files", "output", scenario_path[scenario_id])
  if(!dir.exists(output_path)) dir.create(output_path)

  saconvert::ICES2SS(
    user.wd = sam_input_path,
    user.od = output_path,
    ices.id = "",
    tvslx = FALSE,
    ages = NULL,
    nsexes = 1, # 1: one sex; 2: two sex; -1: one sex and multiply the spawning biomass by the fraction female in the control file
    forN = 2,
    q.extra.se = FALSE,
    q.float = FALSE,
    slx = slx_pattern[scenario_id],
    f.method = f_method[scenario_id]
  ) # steep.init: http://sedarweb.org/docs/wpapers/SEDAR19_DW_06_SteepnessInference.pdf

}
```

### Compare SAM outputs and SS outputs
```{r, warning = FALSE, message = FALSE, fig.height = 2, sam_ss_comparison}
sam_output <- data.frame(
  "Year" = c(sam_dat$years),
  "SSB" = exp(sam_fit$sdrep$value[names(sam_fit$sdrep$value) %in% "logssb"]),
  "Recruits" = exp(sam_fit$sdrep$value[names(sam_fit$sdrep$value) %in% "logR"]),
  "F" = exp(sam_fit$sdrep$value[names(sam_fit$sdrep$value) %in% "logfbar"]),
  "Catch" = c(exp(sam_fit$sdrep$value[names(sam_fit$sdrep$value) %in% "logCatch"]), NA)
)

ss_output <- list()
for (scenario_id in seq_along(scenario_path)){
  ss_output_path <- here::here("NOBA_cod_files", "output", scenario_path[scenario_id])
  ss_output_data <- SS_output(dir = ss_output_path, verbose = F, printstats = F)
  ss_output[[scenario_id]] <- data.frame(
    "Year" = sam_output$Year,
    "SSB" = ss_output_data$timeseries$SpawnBio[ss_output_data$timeseries$Yr %in% sam_output$Year],
    "Recruits" = ss_output_data$timeseries$Recruit_0[ss_output_data$timeseries$Yr %in% sam_output$Year],
    "F" = ss_output_data$timeseries$`F:_1`[ss_output_data$timeseries$Yr %in% sam_output$Year],
    "Catch" = ss_output_data$timeseries$`sel(B):_1`[ss_output_data$timeseries$Yr %in% sam_output$Year]
  )
}

var <- c("SSB", "Recruits", "F", "Catch")
var_label <- c("SSB", "Recruits (Age 3)", expression(F[9-13]), "Catch (in weight)")
colors <- nmfspalette::nmfs_palette("regional web")(length(scenario_path)+1)

par(mfrow = c(2, 2), mar = c(4, 4, 1, 1))

for (i in seq_along(var)){
  plot(NA,
       type="n",
       xlim = range(sam_output$Year),
       ylim = range(sam_output[, var[i]], na.rm = T),
       xlab = "Year",
       ylab = var_label[i]
  )
  lines(sam_output$Year,
        sam_output[, var[i]],
        lty=1,
        col=colors[1])
  for (j in seq_along(scenario_path)){
    lines(ss_output[[j]]$Year,
          ss_output[[j]][, var[i]],
          lty=j+1,
          col=colors[j+1])
  }
}

legend("top", 
       c("SAM", paste("SS_", scenario_path)),
       lty = seq_along(colors),
       col = colors,
       bty="n")
```

### Userful resources
  - [SAM GitHub repository](https://github.com/fishfollower/SAM)
	- Aldrin et al., 2020: [The specification of the data model part in the SAM model matters](https://www.sciencedirect.com/science/article/pii/S0165783620301028#bib0020)
	- Nielsen and Berg, 2014: [Estimation of time-varying selectivity in stock assessments using state-space models](https://www.sciencedirect.com/science/article/abs/pii/S0165783614000228?via%3Dihub)
	- Berg and Nielsen, 2016: [Accounting for correlated observations in an age-based state-space stock assessment model](https://academic.oup.com/icesjms/article/73/7/1788/2458744)
	- [ICES Report of the Arctic Fisheries Working Group (AFWG)](https://www.ices.dk/sites/pub/Publication%20Reports/Expert%20Group%20Report/acom/2018/AFWG/00-AFWG%202018%20Report.pdf)

### 2021-06-15 NOBA_cod_files/README.Rmd

@@
 ### Plot SAM fits
-```{r, fig.height = 2, sam_fit}
+```{r, fig.height = 6, sam_fit}
 par(mfrow = c(1, 1))
@@
 ### Compare SAM outputs and SS outputs
-```{r, warning = FALSE, message = FALSE, fig.height = 2, sam_ss_comparison}
+```{r, warning = FALSE, message = FALSE, fig.height = 6, sam_ss_comparison}
 sam_output <- data.frame(
@@
 
-### Userful resources
+### Resources
   - [SAM GitHub repository](https://github.com/fishfollower/SAM)

### 2021-06-15 NOBA_cod_files/Rscript/noba_cod.R
Update selectivity pattern and recruitment output

@@ -1,21 +1,21 @@
 # Install and library packages --------------------------------------------------------
 
 # install.packages("here")
-#
+# 
 # install_64bit <- TRUE
-#
+# 
 # if(install_64bit){
 #   args <- c("--no-multiarch")
 # } else{
 #   args <- c("")
 # }
-#
+# 
 # devtools::install_github("fishfollower/SAM/stockassessment", INSTALL_opts=args)
-#
+# 
 # remotes::install_github("r4ss/r4ss", branch ="development")
-#
+# 
 # remotes::install_github("Bai-Li-NOAA/saconvert")
-#
+# 
 # remotes::install_github("nmfs-general-modeling-tools/nmfspalette")
 
 library(here)
@@ -72,7 +72,7 @@ catchplot(sam_fit)
 
 # SAM2SS --------------------------------------------------------------------------------------
 scenario_path <- c("A", "B")
-slx_pattern <- c(12, 12)
+slx_pattern <- c(20, 20)
 f_method <- c(3, 2)
 
 
@@ -112,7 +112,7 @@ for (scenario_id in seq_along(scenario_path)){
   ss_output[[scenario_id]] <- data.frame(
     "Year" = sam_output$Year,
     "SSB" = ss_output_data$timeseries$SpawnBio[ss_output_data$timeseries$Yr %in% sam_output$Year],
-    "Recruits" = ss_output_data$timeseries$Recruit_0[ss_output_data$timeseries$Yr %in% sam_output$Year],
+    "Recruits" = ss_output_data$timeseries$Recruit_0[ss_output_data$timeseries$Yr %in% (sam_output$Year-1)],
     "F" = ss_output_data$timeseries$`F:_1`[ss_output_data$timeseries$Yr %in% sam_output$Year],
     "Catch" = ss_output_data$timeseries$`sel(B):_1`[ss_output_data$timeseries$Yr %in% sam_output$Year]
   )

### 2021-06-15 PerCapitaConsumption.Rmd
results from CCA

@@
 usersurvey <- c(here("config/mssurvey_spring.R"),
-                here("config/mssurvey_fall.R"),
-                here("config/mssurvey_spring_01.R"),
-                here("config/mssurvey_fall_01.R"))
+                here("config/mssurvey_fall.R"))
 
@@
 First we need to process the detailed diet file, and also get the output of `run_truth` for numbers.
+At present it seems the CC model is also producing very low per capita consumpton.
+This suggests that interpretation of the DetailedDietCheck.txt output is incorrect?
+It doesn't seem like it is tons consumed at by the population of predators at agecl.
 
@@
 file_diet <- paste0(scenario.name,"DetailedDietCheck.txt")
+
+save <- TRUE
 
@@
 
+# Age structured single species subset
+  sppsubset <- fgs %>%
+    filter(Name %in% c("Yelloweye_rockfish",
+                       "Darkblotched_rockfish",
+                       "Bocaccio_rockfish",
+                       "Pacific_Ocean_Perch",
+                       "Arrowtooth_flounder",
+                       "Petrale_sole",
+                       "Jack_mackerel",
+                       "Pacific_sardine",
+                       "Anchovy",
+                       "Herring",
+                       "Spiny_dogfish")
+    )
+
+# Get true data and species subset
+
+CCAtestom <- om_init(here("config/CCConfig_constest.R"))
+
+CCAtestom_ms <- om_species(sppsubset$Name, CCAtestom)
+
+# fall census survey
+source(here("config/omdimensions.R"))
+source(here("config/mssurvey_CCtest.R"))
+
+omlist_ss <- CCAtestom_ms
+
+    # survtime doesn't match units of time.days in detaileddiet
+    survtime <- survey_sample_full*omlist_ss$runpar$outputstep
+
+    # apply survey design to detailed diet
+    survey_cons <- create_survey_diet(dat = detaileddiet,
+                                      time = survtime,
+                                      species = survspp,
+                                      boxes = survboxes,
+                                      effic = surveffic,
+                                      selex = survselex)
+
+    # get numbers at ageclass for same survey design
+    # note different time units!
+    survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
+                                          time = survey_sample_full,
+                                          species = survspp,
+                                          boxes = survboxes,
+                                          effic = surveffic,
+                                          selex = survselex.agecl)
+
+    # convert survey N times to cons times in days
+    survey_N$time <- survey_N$time*omlist_ss$runpar$outputstep
+
+    # get rid of polygon
+    survey_totN <- survey_N %>%
+      group_by(species, agecl, time) %>%
+      summarise(totN = sum(atoutput)) %>%
+      ungroup()
+
+    # sum over prey to get total consumption in t, divide by N and timestep, convert to g
+    survey_totcons <- survey_cons %>%
+      group_by(species, agecl, time) %>%
+      summarise(totcons = sum(atoutput)) %>%
+      ungroup() %>%
+      left_join(survey_totN) %>%
+      mutate(percap_cons = totcons/totN) %>%
+      #mutate(daily_percap_g = percap_cons/omlist_ss$runpar$outputstep*1000000)
+      mutate(daily_percap_g = percap_cons*1000000) #try assuming cons is snapshot not cumulative since last timestep
+
 ```

### 2021-06-15 config/CCConfig_constest.R

@@ -8,5 +8,5 @@ run.prm.file <- "CalCurrentV3_run.xml"
 scenario.name <- "outputCCV3"
 bioind.file <- "outputCCV3BiomIndx.txt"
 catch.file <- "outputCCV3Catch.txt"
-annage <- TRUE
+annage <- FALSE
 fisheries.file <- "CalCurrentV3Fisheries.csv"

### 2021-06-15 config/mssurvey_CCtest.R

@@ -1,0 +1,67 @@
+# Default survey configuration here has a range of efficiencies and selectivities
+# To emulate a range of species in a single multispecies survey
+# Also now happens in "spring" and "fall"
+# Need to define survey season, area, efficiency, selectivity
+
+# Survey name
+survey.name="BTS_fall_allbox_effic1"
+
+#Atlantis model timestep corresponding to the true output--now from census_spec.R
+timestep <- stepperyr #5
+
+#Which atlantis timestep does the survey run in?--now from census_spec.R
+# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
+# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)
+
+survey_sample_time <- 5 # fall survey
+
+#The last timestep to sample
+total_sample <- noutsteps-1 #495
+
+#Vector of indices of survey times to pull
+survey_sample_full <- seq(survey_sample_time,
+                          total_sample, by=timestep)
+
+survtime <- survey_sample_full
+
+# survey area
+# should return all model areas
+survboxes <- allboxes
+
+# survey efficiency (q)
+# should return a perfectly efficient survey 
+surveffic <- data.frame(species=survspp,
+                        efficiency=rep(1.0,length(survspp)))
+
+# survey selectivity (agecl based)
+# this is by age class, need to change to use with ANNAGEBIO output
+#survselex <- data.frame(species=rep(survspp, each=n_age_classes),
+#                     agecl=rep(c(1:n_age_classes),length(survspp)),
+#                     selex=rep(1.0,length(survspp)*n_age_classes))
+
+# for annage output
+survselex <- data.frame(species=rep(names(annages), n_annages), #  
+                        agecl=unlist(sapply(n_annages,seq)),
+                        selex=rep(1.0,sum(n_annages)))
+
+survselex.agecl <- survselex 
+
+
+# effective sample size needed for sample_fish
+# this effective N is high but not equal to total for numerous groups
+surveffN <- data.frame(species=survspp, effN=rep(100000, length(survspp)))
+
+# survey index cv needed for sample_survey_xxx
+# cv = 0.1
+surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp)))
+
+# length at age cv for input into calc_age2length function
+# function designed to take one cv for all species, need to change to pass it a vector
+lenage_cv <- 0.1
+
[…]

### 2021-06-15 PerCapitaConsumption.Rmd
add consumption tables for NOBA and CCA

@@
 
-```{r, eval=FALSE}
+```{r}
 
@@
 source(here("config/omdimensions.R"))
+source(usersurvey[2])
 
+    # survtime doesn't match units of time.days in detaileddiet
+    survtime <- survey_sample_full*omlist_ss$runpar$outputstep
+
+    # apply survey design to detailed diet
+    survey_cons <- create_survey_diet(dat = detaileddiet,
+                                      time = survtime,
+                                      species = survspp,
+                                      boxes = survboxes,
+                                      effic = surveffic,
+                                      selex = survselex)
+
+    # get numbers at ageclass for same survey design
+    # note different time units!
+    survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
+                                          time = survey_sample_full,
+                                          species = survspp,
+                                          boxes = survboxes,
+                                          effic = surveffic,
+                                          selex = survselex.agecl)
+
+    # convert survey N times to cons times in days
+    survey_N$time <- survey_N$time*omlist_ss$runpar$outputstep
+
+    # get rid of polygon
+    survey_totN <- survey_N %>%
+      group_by(species, agecl, time) %>%
+      summarise(totN = sum(atoutput)) %>%
+      ungroup()
+
+    # sum over prey to get total consumption in t, divide by N and timestep, convert to g
+    survey_totcons <- survey_cons %>%
+      group_by(species, agecl, time) %>%
+      summarise(totcons = sum(atoutput)) %>%
+      ungroup() %>%
+      left_join(survey_totN) %>%
+      mutate(percap_cons = totcons/totN) %>%
+      #mutate(daily_percap_g = percap_cons/omlist_ss$runpar$outputstep*1000000)
+      mutate(daily_percap_g = percap_cons*1000000) #try assuming cons is snapshot not cumulative since last timestep
+
+```
+
+For example, NOBA cod daily per capita intake is tens of grams for older age classes, but is 3-4x lower than expected:
+
+```{r}
+DT::datatable(as.data.frame(survey_totcons%>%filter(species=="North_atl_cod")),
+              rownames = FALSE,
+              options = list(pageLength = 25, 
+                         order = list(list(0, 'asc')))
+)
 ```
@@
 
+omlist_ss <- CCAtestom_ms
+
 # fall census survey
@@
 source(here("config/mssurvey_CCtest.R"))
-
-omlist_ss <- CCAtestom_ms
 
@@
 ```
+
+For example, CCA arrowtooth flounder should have a daily per capita intake in at least tens of grams for older age classes:
+
+```{r}
+knitr::kable(as.data.frame(survey_totcons%>%filter(species=="Arrowtooth_flounder")))
+```

### 2021-06-15 config/mssurvey_fall.R

@@ -44,6 +44,9 @@ survselex <- data.frame(species=rep(names(annages), n_annages), #
                         agecl=unlist(sapply(n_annages,seq)),
                         selex=rep(1.0,sum(n_annages)))
 
+survselex.agecl <- survselex 
+
+
 # effective sample size needed for sample_fish
 # this effective N is high but not equal to total for numerous groups
 surveffN <- data.frame(species=survspp, effN=rep(100000, length(survspp)))
@@ -58,4 +61,8 @@ lenage_cv <- 0.1
 
 # max size bin for length estimation, function defaults to 150 cm if not supplied
 maxbin <- 200
+
+# diet sampling parameters
+alphamult <- 10000000
+unidprey <- 0
 

### 2021-06-15 docs/index.Rmd
update index

@@
 
-<a href="https://sgaichas.github.io/poseidon-dev/PerCapitaConsumption.html" target="_blank"> Compare NOBA and CCA consumption outputs, May 25, 2021</a>
+<a href="https://sgaichas.github.io/poseidon-dev/PerCapitaConsumption.html" target="_blank"> Compare NOBA and CCA consumption outputs, June 15, 2021</a>
 

### 2021-06-15 NOBA_cod_files/README.Rmd

@@
 # Set up two scenarios
-# Scenario A: use selectivity pattern 12 and F method 3 in SS
-# Scenario B: use selectivity pattern 12 and F method 2 in SS
+# Scenario A: use selectivity pattern 20 and F method 3 in SS
+# Scenario B: use selectivity pattern 20 and F method 2 in SS
 
 scenario_path <- c("A", "B")
-slx_pattern <- c(12, 12)
+slx_pattern <- c(20, 20)
 f_method <- c(3, 2)
@@
     "SSB" = ss_output_data$timeseries$SpawnBio[ss_output_data$timeseries$Yr %in% sam_output$Year],
-    "Recruits" = ss_output_data$timeseries$Recruit_0[ss_output_data$timeseries$Yr %in% sam_output$Year],
+    "Recruits" = ss_output_data$timeseries$Recruit_0[ss_output_data$timeseries$Yr %in% (sam_output$Year-1)],
     "F" = ss_output_data$timeseries$`F:_1`[ss_output_data$timeseries$Yr %in% sam_output$Year],

### 2021-08-25 NOBAcod2.Rmd
WIP new dataset for NOBA cod from sacc_30

--- 2021-05-06 NOBAcod.Rmd
@@
 ---
-title: "Generate data for NOBA cod ensemble modeling"
+title: "Generate data for NOBA cod SS3 modeling"
 author: "Sarah Gaichas"
@@
 
-## Standard atlantisom output test (age class not age)
+## New NOBA run with updated selectivity
 
-Let's use the wrapper functions to get initial stock assessment model inputs.
+For this work we want a NOBA run with the climate signal, the planned fishery forcing, and more realistic fishery selectivities for key species (fisheries that don't catch the smallest age classes).
 
-Then we need to adjust the base atlantisom functions for true age classes, and develop options to call those within the wrappers.
-
-We need config files for the NOBA model, our surveys and fisheries. Surveys and fisheries are based on the ones we used for sardine in the California Current as a test.
-
-**Special note:** *Atlantis output files formerly had the prefix "output" and `atlantisom` was coded to expect this.
-NOBA output files (and newer Atlantis models) do not universally use the "output" prefix naming convention.
-Annage branch `atlantisom` functions have been updated for user-specified output filenames.*
-
-**Special note 2:** *NOBA's nordic_groups_v04.csv file has an inFishery column, rather than the `atlantisom` expected InFishery column.
-I've changed the column name in the input file, but need to add robustness to atlantisom code to remove case sensitivity.*
+This dataset will come from NOBA sacc_30, with original biology, planned F + changes in selectivity for multiple groups, forced recruitment for groups	SCR+SSH+NCO, and climate change from	1981-2068+loop 2067..
 
@@
 
-`NOBA2config.R` looks like this (adjusted from Alfonso's original):
+`NOBAsacc30config.R` looks like this (adjusted from Alfonso's original):
 
-```{r, code = readLines("./config/NOBA2Config.R"), eval=F}
+```{r, code = readLines("./config/NOBAsacc30Config.R"), eval=F}
 ```
@@
 
-NOBAom <- om_init(here("config/NOBA2config.R"))
+NOBAom <- om_init(here("config/NOBAsacc30config.R"))
 
@@
 
-Note: NOBA doesn't seem to have a selective fishery for cod. Lots of age 1 in the catch!
-
-# (Annage development details below)
-
-Now to modify the underlying code to get full age structure.
-
-## Updating functions for full age structure
-
-General strategy:
-
-1. add annual age option to run_truth function call, default it to FALSE
-  + if true, looks for ANNAGEBIO and ANNAGECATCH in designated directory, error out if missing
-  + if true, read these files in addition to existing set rather than replace existing
-    + read them in with new load_nc_annage function
-    + add them to the output object in addition to existing set
-1. write load_nc_annage
-  + need to understand annage nc structures
-    + if same as others just with full age structure, replace previous nc
-    + only numbers output? 
-      + how to match with resn and structn for lengths/wts?
-      + reconcile with bio parameters such as maturity ogive?
-    + only for some groups?
-
-What is in the ANNAGEBIO and ANNAGECATCH .nc files?
-
-```{r annagebio-contents}
-
-# taken from existing load_nc
-
-file.nc <- file.path(d.name, "outputnordic_runresults_01ANNAGEBIO.nc")
-
-  # Load ATLANTIS output!
-  at_out <- RNetCDF::open.nc(con = file.nc)
-
-  # Get info from netcdf file! (Filestructure and all variable names)
-  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
-    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
-  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
-  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
-  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
-
-RNetCDF::close.nc(at_out)
-
-```
-
-ANNAGEBIO.nc variable names (first hundred): `r var_names_ncdf[1:100]`
-
-ANNABEBIO.nc variable names (last hundred): `r var_names_ncdf[1590:1690]`
-
-ANNAGEBIO.nc timesteps: `r n_timesteps`
-
-ANNAGEBIO.nc boxes: `r n_boxes`
-
-ANNAGEBIO.nc layers: `r n_layers`
-
-Annual age output is in numbers and weight at each timestep, layer, and box, which is really good news.
-We should be able to get weight at true age more directly this way than the interpolation I was planning.
-Also, we may be able to do length estimation differently?
-
-To modify `load_nc_annage` we need to figure out how many true ages each species has from the biological parameter file using `load_biolprm`:
-
-```{r biolprm}
-
-biolprm <- load_biolprm(d.name, biol.prm.file)
-fgs <- load_fgs(d.name, functional.groups.file)
-
-#max age should be ages per cohort times the number of age classes (<10 for some NOBA spp)
-
-names(biolprm$agespercohort) <- c("code", "ageperagecl")
-
-maxage <- merge(biolprm$maturityogive, biolprm$agespercohort, all.x = T) %>%
-    select(code, nagecl, ageperagecl) %>%
-    mutate(maxage = nagecl * ageperagecl) %>%
-    mutate(name = fgs$Name[match(code, fgs$Code)])
-
-select_groups <- c('Mackerel', 'North_atl_cod', 'Capelin')
-
-for (i in seq_along(select_groups)) {
-  select_group_ages <- 1:maxage$maxage[match(select_groups[i], maxage$name)]
-}
-
-```
-
-Try this `load_nc_annage` function:
-
-```{r loadncannage, eval=FALSE}
-
-load_nc_annage <- function(dir = getwd(), file_nc, bps, fgs, biolprm, select_groups,
-  select_variable =
-  c("Nums", "Weight"),
-  check_acronyms = TRUE, bboxes = c(0), verbose = FALSE) {
-  # NOTE: The extraction procedure may look a bit complex...
-  # A different approach would be to
-  # create a dataframe for each variable (e.g. GroupAge_Nums)
-  # and combine all dataframes at the end.
-  # This alternative approach requires a lot more storage
-  # and the code wouldn't be vectorized.
-
-  # Check input of the nc file
-  if (tail(strsplit(file_nc, "\\.")[[1]], 1) != "nc") {
-    stop("The argument for file_nc,", file_nc, "does not end in nc")
-  }
-  if (is.null(dir)) {
-    file.nc <- file_nc
-  } else {
-    file.nc <- file.path(dir, file_nc)
-  }
-
-  # Check input of select_variable as only one value is allowed
-  select_variable <- match.arg(select_variable, several.ok = FALSE)
-
-  # Check input structure!
-  if (check_acronyms) {
-    active_groups <- as.vector(subset(fgs, IsTurnedOn == 1)$Name)
-    inactive_groups <- select_groups[which(
-      !is.element(select_groups, active_groups))]
-    if (length(inactive_groups) >= 1) {
-      select_groups <- select_groups[!is.element(select_groups, inactive_groups)]
-      warning(paste(paste("Some selected groups are not active in the model run.",
-        "Check 'IsTurnedOn' in fgs\n"),
-        paste(inactive_groups, collapse = "\n")))
-    }
-    if (all(!is.element(select_groups, active_groups))) {
-      stop(paste("None of the species selected are active in the model run.",
-        "Check spelling and Check 'IsTurnedOn' in fgs"))
-    }
-  }
-
-  # Deal with file structures
-
-  # Load ATLANTIS output!
-  at_out <- RNetCDF::open.nc(con = file.nc)
-  on.exit(RNetCDF::close.nc(at_out))
-
-  if (select_variable != "N" & all(is.element(select_groups, bps))) {
-    stop("The only output for Biomasspools is N.")
-  } else{
-    print(paste("Read", file.nc, "successfully"))
-  }
-
-  # Get info from netcdf file! (Filestructure and all variable names)
-  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
-    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
-  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
-  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
-  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
-
-  # Extract data from the ncdf file
-  # Create a vector of all potential variable names
-  # Only use names which are available in the ncdf-file as an
-  # extraction of missing variables is not possible
-  # Create vector of available species at the end using search_clean
-  # This is needed to create species-names later on
-
-  # CHANGED hardcoded cohort number to reflect true ages from
-
-  # To make the creation of variables as robust as possible
-  # we introduce different combinations of groups, variable, and cohort
-  # Only combinations present in the ncdf are used later on
-  # Loop over select_groups to use the same ordering
-
-  names(biolprm$agespercohort) <- c("code", "ageperagecl")
-
-  maxage <- merge(biolprm$maturityogive, biolprm$agespercohort, all.x = T) %>%
-    select(code, nagecl, ageperagecl) %>%
-    mutate(maxage = nagecl * ageperagecl) %>%
-    mutate(name = fgs$Name[match(code, fgs$Code)])
-
-  search <- list()
-  for (i in seq_along(select_groups)) {
-    select_group_ages <- 1:maxage$maxage[match(select_groups[i], maxage$name)]
-    search[[i]] <- c(
-      unlist(lapply(paste0(select_groups[i], select_group_ages),
-                    paste0, select_variable)),           # GroupCohortVariable
-      unlist(lapply(paste0(select_groups[i], select_variable),
-                    paste0, select_group_ages)),                   # GroupVariableCohort
-      unlist(lapply(paste0(select_groups[i], select_group_ages),
-                    paste, select_variable, sep = "_")), # GroupCohort_Variable
-      unlist(lapply(paste(select_groups[i], select_variable, sep = "_"),
-                    paste0, select_group_ages)),                   # Group_VariableCohort
-      unlist(lapply(paste(select_groups[i], select_group_ages, sep = "_"),
-                    paste, select_variable, sep = "_")), # Group_Cohort_Variable
-      unlist(lapply(paste(select_groups[i], select_variable, sep = "_"),
-                    paste, select_group_ages, sep = "_")),         # Group_Variable_Cohort
-      unlist(paste0(select_groups[i], select_variable)),                                                      # GroupVariable
-      unlist(paste(select_groups[i], select_variable,
-                   sep = "_"))                           # Group_Variable
-      )
-    search[[i]] <- search[[i]][is.element(search[[i]], var_names_ncdf)]
-    search[[i]] <- unique(search[[i]])
-  }
-  search_clean <- do.call(c, search)
-  # If the combination of select_groups and select_variable ends up not being found.
-  if (length(search_clean) == 0) return(0)
-
-  at_data <- lapply(search_clean, RNetCDF::var.get.nc, ncfile = at_out)
-
-  # Get final species and number of ageclasses per species
-  final_species <- select_groups[sapply(
-    lapply(select_groups, grepl, x = search_clean), any)]
-  final_agecl <- maxage$maxage[
-    sapply(final_species, function(x) which(x == maxage$name))]
-
-  num_layers <- RNetCDF::var.get.nc(ncfile = at_out, variable = "numlayers")[, 1]
-  # add sediment layer!
-  num_layers <- num_layers + ifelse(num_layers == 0, 0, 1)
-
-  # Create an array of layerids.
-  # Every entry in the array indicates if a layer is present (= 1) or not (= 0).
-  # Boxes without layers (= islands) have only 0s as id,
-  # used later on to remove data from non-existent layers!
-  # By default output should be 0 in those layers.
-  # Layers in boundary boxes are set to 0 if bboxes is anything other than NULL!
-  # Applying a boolean array to an array results in a vector!
-  for (i in seq_along(num_layers)) {
-    if (i == 1) layerid <- array(dim = c(n_layers, n_boxes))
-    if (num_layers[i] == 0) {
-      layerid[, i] <- 0
-    } else {
-      if (!is.null(bboxes) & is.element((i - 1), bboxes)) {
-        layerid[, i] <- 0
-      } else {
-        layerid[, i] <- c(rep(0, times = n_layers - num_layers[i]),
-                          rep(1, times = num_layers[i]))
-      }
-    }
-  }
-
-  # Create vectors for polygons and layers
-  # Each vector has the length equal to one time-step
-  # All data from islands and non-existent layers is removed
-  # Therefore the length of these
-  # vectors is equal for each extracted variable
-  boxes <- 0:(n_boxes - 1)
-  # Remove islands and boundary boxes
-  island_ids <- num_layers == 0
-  if (!is.null(bboxes)) {
-    boundary_ids <- is.element(boxes, bboxes)
-    island_ids <- island_ids | boundary_ids
-  }
-  boxes <- boxes[!island_ids]
-  num_layers <- num_layers[!island_ids]
-
-  polygons <- rep(boxes, times = num_layers)
-  layers <- sapply(num_layers[num_layers != 0] - 2,
-    function(x) c(seq(x, from = 0, by = 1), n_layers - 1))
-  if (any(sapply(layers, length) != num_layers[num_layers != 0])) {
-    stop("Number of layers incorrect.")
-  }
-  layers <- do.call(c, layers)
-  if (length(polygons) != length(layers)) {
-    stop("Number of polygons and layers do not match.")
-  }
-
-  # In the following section the data is transformed to a long dataframe
-  # I haven't found any solution to vectorize the creation of the dataframe
-  # columns (species, age, polygons,...)
-  # when data from 2d and 3d arrays
-  # (e.g. select_variable = "N" all biomasspools are only present in the
-  # sediment layer.) are read in simultaneously.
-  # Therefore the current "messy" solution splits the data
-  # in 2 subpopulations: 2d-data and 3d-data
-  at_data3d <- at_data[which(sapply(at_data, function(x) length(dim(x))) == 3)]
-  at_data2d <- at_data[which(sapply(at_data, function(x) length(dim(x))) == 2)]
-
-  int_fs <- final_species
-  int_fa <- final_agecl
-
-  if (length(at_data3d) >= 1) {
-    # Remove biomasspools if selected variable is "N"!
-    if (select_variable == "N") {
-      int_fs <- final_species[!is.element(final_species, bps)]
-      int_fa <- final_agecl[!is.element(final_species, bps)]
-      # Note this only works if age-structured vertebrates have 10 ageclasses
-      int_fa[int_fa == 10] <- 1
-    }
-    for (i in seq_along(at_data3d)) {# for loop over all variables
-      if (i == 1) result3d <- list()
-      for (j in 1:n_timesteps) {# loop over timesteps
-        if (j == 1) values <- array(dim = c(length(layers), n_timesteps))
-        values[, j] <- at_data3d[[i]][,, j][layerid == 1]
-      }
-      result3d[[i]] <- as.vector(values)
-    }
-    result3d <- data.frame(
-      species = unlist(sapply(
-        X = mapply(FUN = rep, x = int_fs, each = int_fa, SIMPLIFY = FALSE,
-        USE.NAMES = FALSE),
-        FUN = rep, each = length(layers) * n_timesteps, simplify = FALSE)),
-      agecl = unlist(sapply(
-        X = sapply(X = int_fa, FUN = seq, from = 1, by = 1, simplify = FALSE,
-        USE.NAMES = FALSE),
-        FUN = rep, each = length(layers) * n_timesteps, simplify = FALSE)),
-      polygon = unlist(sapply(
-        X = n_timesteps * int_fa, FUN = rep, x = polygons, simplify = F,
-        USE.NAMES = FALSE)),
-      layer = unlist(sapply(
-        X = n_timesteps * int_fa, FUN = rep, x = layers, simplify = FALSE,
-        USE.NAMES = FALSE)),
-      time = unlist(sapply(
-        X = int_fa, FUN = rep, x = rep(0:(n_timesteps - 1), each = length(layers)),
-        simplify = FALSE, USE.NAMES = FALSE)),
-      atoutput = do.call(c, result3d),
-      stringsAsFactors = FALSE)
-  }
-
-  if (length(at_data2d) >= 1) {
-    # Only select biomasspools if selected variable is "N"!
-    if (select_variable == "N") {
-      int_fs <- final_species[is.element(final_species, bps)]
-      int_fa <- final_agecl[is.element(final_species, bps)]
-    }
-    # age-structured invert groups are combined in ncdf file!
-    if (select_variable == "Grazing") int_fa <- 1
-    for (i in seq_along(at_data2d)) {# for loop over all variables!
-      if (i == 1) result2d <- list()
-      for (j in 1:n_timesteps) {# loop over timesteps
-        if (j == 1) values <- array(dim = c(length(boxes), n_timesteps))
-        values[, j] <- at_data2d[[i]][, j][boxes + 1]
-      }
-      result2d[[i]] <- as.vector(values)
-    }
-
-    # Order of the data in value column = "atoutput".
-    # 1. species  --> rep each with the number of
-    #                 ageclasses and n_timesteps * boxes
-    # 2. age      --> rep each (1:10 for each species) with n_timesteps * boxes
-    # 3. timestep --> rep each timestep (1:n_timesteps)
-    #                 with the number of boxes and final_agecl
-    #                 (num cohorts per species)
-    # 4. polygon  --> rep boxes times n_timesteps * final_agecl
-    #                 (num cohorts per species)
-    result2d <- data.frame(species = unlist(sapply(
-      X = mapply(FUN = rep, x = int_fs, each = int_fa, SIMPLIFY = FALSE,
-        USE.NAMES = FALSE),
-        FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
-      agecl = unlist(sapply(X = sapply(X = int_fa, FUN = seq, from = 1,
-        by = 1, simplify = FALSE, USE.NAMES = FALSE),
-        FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
-      polygon = unlist(sapply(X = n_timesteps * int_fa,
-        FUN = rep, x = boxes, simplify = FALSE, USE.NAMES = FALSE)),
-      time = unlist(sapply(X = int_fa, FUN = rep, x = rep(0:(n_timesteps - 1),
-        each = length(boxes)), simplify = FALSE, USE.NAMES = FALSE)),
-      atoutput = do.call(c, result2d),
-      stringsAsFactors = F)
-    if (select_variable == "N") result2d$layer <- n_layers - 1
-  }
-
-  # Combine dataframes if necessary!
-  if (all(sapply(lapply(at_data, dim), length) == 3) & select_variable != "N") {
-    result <- result3d
-  }
-  if (all(sapply(lapply(at_data, dim), length) == 2) & select_variable != "N") {
-    result <- result2d
-  }
-  if (select_variable == "N") {
-    if (length(at_data2d) >= 1 & length(at_data3d) == 0) result <- result2d
-    if (length(at_data2d) == 0 & length(at_data3d) >= 1) result <- result3d
-    if (length(at_data2d) >= 1 & length(at_data3d) >= 1) {
-      result <- rbind(result2d, result3d)
-    }
-  }
-
-  # Remove min_pools if existent (well, there always are min pools... ;)).
-  min_pools <- is.element(result$atoutput, c(0, 1e-08, 1e-16))
-  if (length(min_pools) > 0) {
-    # exclude 1st timestep and sediment layer from calculation
-    print_min_pools <- sum(min_pools) -
-      length(result[min_pools & result$time == 1, 1]) -
-      length(result[min_pools & result$time > 1 & result$layer == 7, 1])
-    if (print_min_pools > 0 & verbose) {
-      warning(paste0(round(print_min_pools/dim(result)[1] * 100),
-        "% of ", select_variable, " are true min-pools (0, 1e-08, 1e-16)"))
-    }
-    result <- result[!min_pools, ]
-  }
-
-  # Remove non-existent layers.
-  # WARNING: Biomass is build up (very few) in sediment layer for
-  # NON sediment groups (e.g. baleen whales)
-  # Therefore, I subset all data from that layer for non biomass groups and
-  # groups which cannot penetrate into the sediment!
-  # UPDATE: Doesn't work with layers as species are not distributed through
-  # the whole water column and do not appear in
-  # every polygon.
-
-  # Sum up N for invert cohorts if invert cohorts are present!
-  # NOTE: invert cohorts of size 10 are not considered!
-  if (select_variable == "N" & any(final_agecl != 10 & final_agecl > 1)) {
-    result <- result %>%
-      dplyr::group_by(polygon, layer, species, time) %>%
-      dplyr::summarise(atoutput = sum(atoutput))
-  }
-
-  return(result)
-}
-
-```
-
-Does it work? This snippet will go into `run_truth` if it does:
-
-```{r testloadnc}
-
-file_nc <- ("outputnordic_runresults_01ANNAGEBIO.nc")
-bps <- load_bps(dir = d.name, fgs = functional.groups.file, file_init = biomass.pools.file)
-allboxes <- load_box(dir = d.name, file_bgm = box.file)
-boxes <- get_boundary(allboxes)
-
-numsage <- load_nc_annage(dir = d.name,
-                  file_nc = file_nc,
-                  bps = bps,
-                  fgs = fgs,
-                  biolprm = biolprm,
-                  select_groups = select_groups,
-                  select_variable = "Nums",
-                  check_acronyms = TRUE,
-                  bboxes = boxes, 
-                  verbose = TRUE)
-  #if(verbose) message("Numbers read in from ANNAGEBIO.")
-
-weightage <- load_nc_annage(dir = d.name,
-                  file_nc = file_nc,
-                  bps = bps,
-                  fgs = fgs,
-                  biolprm = biolprm,
-                  select_groups = select_groups,
-                  select_variable = "Weight",
-                  check_acronyms = TRUE,
-                  bboxes = boxes, 
-                  verbose = TRUE)
-  #if(verbose) message("Weight read in from ANNAGEBIO.")
-
-```
-
-Need to aggreagate these to see if they look reasonable; this code snippet from `om_comps` with `numsage` output subsituted for `nums`.
-
-```{r fullagecomps-test}
-
-omlist_ss<-NOBAom_cod
-source(here("/config/NOBA2Config.R"))
-source(here("/config/omdimensions.R"))
-source(here("/config/codsurvey.R"))
-source(here("/config/codfishery.R"))
-
-n_reps <- 1
-
-  #numbers based fishery independent survey for age and length comps
-  # same user specifications as indices
-  survey_N <- atlantisom::create_survey(dat = numsage,
-                                        time = survtime,
-                                        species = survspp,
-                                        boxes = survboxes,
-                                        effic = surveffic,
-                                        selex = survselex)
-  #Sample fish for age composition
-  # if we want replicates for obs error this sample function will generate them
-  age_comp_data <- list()
-  for(i in 1:n_reps){
-    age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
-  }
-
-    # save age comps
-  #if(save){
-    saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
-  #}
-
-```
-
-Plot full age comps, have 20 years for cod now, no longer age class but annual ages.
-
-```{r plotfullagecomp}
-
-#read in comp data
-age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
-age_comp_data <- age_comp_data[[1]]
-
-#fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
-#fish_age_comp <- fish_age_comp[[1]]
-
-#add this to om_indices function so that this has years when read in
-#fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
-
-Natage <- filter(age_comp_data, time %in% c(55:175))
-
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Natage$species))
-
-  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
-
-```
-
-What is the weight output?
-Documentation only mentions numbers output.
-Looks like average weight at age for each time and polygon?
-If so, need to do weighted average weight using the numbers at age in each polygon?
-
-This output is described in Atlantis code, file atlantis/atphysics/atagetracerIO.c, function `writeBMAnnAgeBioData`:
-
-```
-if (propit == out_nums_id) {  // Numbers case
-                                for (b = 0; b < bm->nbox; b++) {
-                                    val[b] = 0.0;
-                                    for (k = 0; k < bm->wcnz; k++) {  // Sum over all individuals in the box
-                                        val[b] += bm->wctr[b][k][id] * FunctGroupArray[sp].boxPopRatio[b][k][cohort][ai];
-                                    }
-                                }
-                            } else  { // Size data - get weighted average of values per layer if numbers > min_dens
-                                for (b = 0; b < bm->nbox; b++) {
-                                    val[b] = 0.0;
-                                    tot = 0.0;
-                                    for (k = 0; k < bm->wcnz; k++) {  // Sum over all individuals in the box
-                                        if (bm->wctr[b][k][id] > bm->min_dens) {
-                                            val[b] += bm->wctr[b][k][id] * (bm->wctr[b][k][sn] + bm->wctr[b][k][rn]) * 5.7 * mg_2_g * FunctGroupArray[sp].boxPopRatio[b][k][cohort][ai];
-                                            tot += bm->wctr[b][k][id];
-                                        }
-                                    }
-                                    if (tot > 0.0)
-                                        val[b] /= tot;
-                                }
-                            }
-```
-
-So weight output appears to be average weight at age in grams?
-
-```{r wtage}
-
-annwtage <- weightage %>%
-  group_by(species, agecl, time) %>%
-  summarise(wtage = mean(atoutput)) %>%
-  filter(time %in% (survtime))
-
-boxwtage <- weightage %>%
-  mutate(wtage = atoutput) %>%
-  filter(species %in% c("North_atl_cod"),
-         polygon %in% c(1,7,24,30,38,45),
-         agecl %in% c(16:20))
-
-wtagedat <- boxwtage
-
-wageplot <- ggplot(wtagedat, aes(time, wtage)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (5 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " "))
-
-wageplot + facet_wrap(c("polygon"), scales="free_y")
-
-```
-
-So, to modify run_truth we need to add the annage output but not replace the regular output in case users want output from non-age structured groups (inverts, etc.).
-Also need to see if lengths can be generaged from the annage outputs, if not need to keep the full set of outputs from the .nc file for all species to generate them.
-
-UPDATE: there is a difference between the annage weight at age and that calculated using 'calc_age2length()'.
-While the results from the standard nc output look reasonable for NOBA cod, the annage weight at age is half that for the standard (and half what real cod would be).
-More worrisome, the oldest age classes have smaller average weight at age than some younger ones.
-I cannot think of why this would be.
-Until we are sure the annage weight output is correct, we will instead interpolate the weight at age output of 'calc_age2length" for use as assessment input.
-
-Looks like the annagecatch output is only numbers and not weights. We don't get weight at age from fisheries usually so this is fine.
-
-```{r loadannagecatch}
-# taken from existing load_nc
-
-file.nc <- file.path(d.name, "outputnordic_runresults_01ANNAGECATCH.nc")
-
-  # Load ATLANTIS output!
-  at_out <- RNetCDF::open.nc(con = file.nc)
-
-  # Get info from netcdf file! (Filestructure and all variable names)
-  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
-    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
-  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
-  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
-  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
-
-RNetCDF::close.nc(at_out)
-```
-
-ANNAGECATCH.nc variable names (first hundred): `r var_names_ncdf[1:100]`
-
-ANNAGECATCH.nc variable names (last hundred): `r tail(var_names_ncdf, 100)`
-
-ANNAGECATCH.nc timesteps: `r n_timesteps`
-
-ANNAGECATCH.nc boxes: `r n_boxes`
-
-ANNAGECATCH.nc layers: `r n_layers`
-
-Catch file contains "IsFished" species discard at age and catch at age by fishery in numbers.
-So now modifying load_nc_annage to get the fishery specific outputs, which requires additional atlantis fishery setup files and additions to the config files above.
-
-New function `load_fisheries()` needed to load the fishery specification file (a .csv), similar to `load_fgs()`:
-
-```{r load_fisheries, eval=FALSE}
-
-load_fisheries <- function(dir = getwd(), file_fish){
-  if (is.null(dir)) {
-    file.fish <- file_fish
-  } else {
-    file.fish <- file.path(dir, file_fish)
-  }
-  result <- read.table(file = file.fish, sep = ",",
-    header = TRUE, stringsAsFactors = FALSE)
-  return(result)
-}
-```
-
-Call `load_fisheries()` within `load_nc_annage()` to rename select_variable as used in ANNAGECATCH.nc:
-
-```{r addsnippet, eval=FALSE}
-
-select_variable <- c("Catch", "Discard")
-dir <- here("atlantisoutput","NOBA_march_2020")
-file_fish <- "NoBAFisheries.csv"
-
-#starting at line 115 in current load_nc_annage
-
-if(select_variable %in% c("Catch", "Discard")){
-    #read in fleet names
-    fleetnames <- load_fisheries(dir = dir, file_fish = file_fish)
-    svfish <- c()
-    for(i in select_variable){
-      sv <-paste0(i, "_", fleetnames$Code)
-      svfish <- c(svfish, sv)
-    }
-    select_variable <- svfish
-  }
-
-```
-
-Also added if statements to add fleets for fisheries outputs, if statements to separate output for annage and annagecatch.
-
-Test new `load_nc_annage()`:
-
-```{r testfishery}
-
-# source(here("R/load_nc_annage.R"))
-# source(here("R/load_fisheries.R"))
-
-source(here("config/NOBA2Config.R"))
-NOBAom_cod <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
-
-fgs <- NOBAom_cod$funct.group_ss
-biolprm <- NOBAom_cod$biol
-#select_groups <- c('Mackerel', 'North_atl_cod', 'Capelin')
-select_groups <- 'North_atl_cod'
-
-bps <- load_bps(dir = d.name, fgs = functional.groups.file, file_init = biomass.pools.file)
-allboxes <- load_box(dir = d.name, file_bgm = box.file)
-boxes <- get_boundary(allboxes)
-
-file_nc <- "outputnordic_runresults_01ANNAGECATCH.nc"
-file_fish <- "NoBAFisheries.csv"
-
-#for testing line by line
-bboxes <- boxes
-dir <- d.name
-select_variable <- "Catch"
-
-fishcatchage <- load_nc_annage(dir = d.name,
-                  file_nc = file_nc,
-                  file_fish = file_fish,
-                  bps = bps,
-                  fgs = fgs,
-                  biolprm = biolprm,
-                  select_groups = select_groups,
-                  select_variable = "Catch",
-                  check_acronyms = TRUE,
-                  bboxes = boxes, 
-                  verbose = TRUE)
-
-fishcatchage[1:100,]
-
-fishdiscage <- load_nc_annage(dir = d.name,
-                  file_nc = file_nc,
-                  file_fish = file_fish,
-                  bps = bps,
-                  fgs = fgs,
-                  biolprm = biolprm,
-                  select_groups = select_groups,
-                  select_variable = "Discard",
-                  check_acronyms = TRUE,
-                  bboxes = boxes, 
-                  verbose = TRUE)
-
-fishdiscage
-```
-
-Do we want catch by fishery? Yes. No fishery column here, check what it is doing.
-fishdiscage is empty, make sure there is no discard output from NOBA for these species (confirmed).
-
-Add a check for multiple fisheries per species, if only one then fleet irrelevant.
-
-NOBA has multiple fleets for cod (but they are not in output!), so add fleet column to output.
-
-Update: there is now a fleet column in output which should index correctly if a species is caught by multiple fleets.
-
-Plot catch at age; aggregate over polygons, apply sample_fish:
-
-```{r catchannage}
-
-omlist_ss<-NOBAom_cod
-source(here("/config/NOBA2Config.R"))
-source(here("/config/omdimensions.R"))
-source(here("/config/codsurvey.R"))
-source(here("/config/codfishery.R"))
-
-n_reps <- 1
-
-  #numbers based fishery independent survey for age and length comps
-  # same user specifications as indices
-  #fishery catch at age each observed timestep summed over observed polygons
-  # catch at age by area and timestep
-  catch_numbers <-  atlantisom::create_fishery_subset(dat = fishcatchage,
-                                                      time = fishtime,
-                                                      species = select_groups,
-                                                      boxes = fishboxes)
-
-  # if we want replicates for obs error this sample function will generate them
-  catch_age_comp <- list()
-  for(i in 1:n_reps){
-    catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
-  }
-
-  # save fishery age comps
-  #if(save){
-    saveRDS(catch_age_comp, file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
-  #}
-
-```
-
-NOTE that the fleet column is lost in the `atlantisom::sample_fish` function!
-
-Look at only every 10 years output (this is a sample, effN = 1000 for the year, which is 200 per timestep here; see calc below confirming):
-
-```{r plotfullcatchage}
-
-#read in comp data
-# age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
-# age_comp_data <- age_comp_data[[1]]
-
-fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
-fish_age_comp <- fish_age_comp[[1]]
-
-#add this to om_indices function so that this has years when read in
-fish_age_comp$time <- fish_age_comp$time/fstepperyr
-
-Catage <- filter(fish_age_comp, time %in% c(30.0, 40.0, 50.0, 60.0, 70.0, 79.8))
-
-Catageplot <- ggplot(Catage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Catage$species))
-
-  Catageplot + facet_wrap(~time, ncol=3, scales="free_y")
-
-```
-
-```{r testeffn}
-
-fish_age_comp %>%
-  filter(time<31) %>%
-  summarise(sum(atoutput))
-
-```
-
-Appears to be working for single fleet per species now.
-
-Quick test of `run_truth()` with annage option turned on for NOBA:
-
-```{r testruntruth, eval=FALSE}
-
-source(here("config/NOBA2Config.R"))
-
-  #Load functional groups
-  funct.groups <- atlantisom::load_fgs(dir=d.name,
-                           file_fgs = functional.groups.file)
-  #Get just the names of active functional groups
-  funct.group.names <- funct.groups %>% 
-    filter(IsTurnedOn == 1) %>%
-    select(Name) %>%
-    .$Name
-
-  # # load biolprm file
-  #biol <- atlantisom::load_biolprm(d.name, biol.prm.file)
-  # 
-  # # load run_prm
-  # runpar <- atlantisom::load_runprm(d.name, run.prm.file)
-  # 
-  # # load box info file
-  # boxpars <- atlantisom::load_box(d.name, box.file)
-
-testtruth <- atlantisom::run_truth(scenario = scenario.name,
-                       dir = d.name,
-                       file_fgs = functional.groups.file,
-                       file_bgm = box.file,
-                       select_groups = funct.group.names,
-                       file_init = initial.conditions.file,
-                       file_biolprm = biol.prm.file,
-                       file_runprm = run.prm.file,
-                       file_fish = fisheries.file,
-                       verbose = TRUE,
-                       annage = TRUE,
-                       save = TRUE
-    )
-
-```
-
-After much debugging to properly index multiple fleets per species in catch at annual age output, this produced a run_truth.Rdata file and took about 40 minutes to run.
-
-Now all of the wrappers need to read from annage to produce composition data.
-`om_init()`  now takes annage flag from the config file.
-`om_species()` looks for annage bio and catch output in the truth file and includes it in the species file (adding NULL if it isn't there).
-`om_comps()` looks for annage bio and catch in the species file and adds them to the comps list if there (adding NULL if they aren't there).
-
-Also still need the weight at age interpolation function. Incorporate this at the `om_comps()` step. Next rmd!
-
-Running `om_init()`, `om_species()`, `om_index()`, `om_comps()` in chunk above to make sure the original output is preserved, then plot the annage output [above](#annageoutput).
+We previously noted that NOBA didn't have a selective fishery for cod; there were lots of age 1 in the catch! With this new run, we see the effects of selective fisheries
 
@@
 require(r4ss)
-#omlist_ss<- ?
+#omlist_ss<-CC3om_sardine
 source(here("config/omdimensions.R"))
@@
 
-#Load in model object from SAM
-noba_cod_out <- system.file("inst","extdata","NOBA_cod_files","NEAcod-2020","run","model.RData",
-                            package = "atlantisom") 
 #Month of survey/fishing
-#survey_month <- 7
-#fishing_month <- 1
+survey_month <- 7
+fishing_month <- 1
 

### 2021-08-25 config/NOBAsacc30Config.R

@@ -1,0 +1,12 @@
+d.name <- here("atlantisoutput","NOBA_sacc_30")
+functional.groups.file <- "nordic_groups_v04.csv"
+biomass.pools.file <- "nordic_biol_v23.nc"
+biol.prm.file <- "nordic_biol_incl_harv_v_011_1skg.prm"
+box.file <- "Nordic02.bgm"
+initial.conditions.file <- "nordic_biol_v23.nc"
+run.prm.file <- "nordic_run_v04.xml"
+scenario.name <- "nordic_runresults_01"
+bioind.file <- "nordic_runresults_01BiomIndx.txt"
+catch.file <- "nordic_runresults_01Catch.txt"
+annage <- TRUE
+fisheries.file <- "NoBAFisheries.csv"

### 2021-08-25 config/codfishery.R

@@ -1,4 +1,6 @@
 # Default fishery configuration here is a census
+fishery.name="cod_census"
+
 # Inherits species from input omlist_ss
 fishspp <- omlist_ss$species_ss 
 

### 2021-08-25 config/codsurvey.R

@@ -1,6 +1,9 @@
 # Default survey configuration here is a census
 # Need to define survey season, area, efficiency, selectivity
 
+# Survey name
+survey.name="cod_midyear_allbox_effic1"
+
 #Atlantis model timestep corresponding to the true output--now from census_spec.R
 timestep <- stepperyr #5
 
@@ -11,7 +14,7 @@ timestep <- stepperyr #5
 survey_sample_time <- midptyr #2; defined in omdimensions.R
 
 #The last timestep to sample
-total_sample <- noutsteps-1 #495
+total_sample <- noutsteps-1 
 
 #Vector of indices of survey times to pull
 survey_sample_full <- seq(survey_sample_time,
@@ -35,9 +38,12 @@ surveffic <- data.frame(species=survspp,
 #                     selex=rep(1.0,length(survspp)*n_age_classes))
 
 # for annage output
-survselex <- data.frame(species=rep(survspp, n_annages), #  
+survselex <- data.frame(species=rep(names(annages), n_annages), #  
                         agecl=unlist(sapply(n_annages,seq)),
                         selex=rep(1.0,sum(n_annages)))
+
+survselex.agecl <- survselex 
+
 
 # effective sample size needed for sample_fish
 # this effective N is high but not equal to total for numerous groups

### 2021-09-29 NOBAcod2.Rmd
first attempt at new NOBA cod dataset

@@
 library(here)
+library(atlantisom)
+library(tidyverse)
+library(ggthemes)
+
 ```
@@
 
-`NOBAsacc30config.R` looks like this (adjusted from Alfonso's original):
+`NOBAsacc30config.R` looks like this (this is the config file used to produce the outputs posted to google drive, but note that it has a different directory structure than used here):
 
-```{r, code = readLines("./config/NOBAsacc30Config.R"), eval=F}
+```{r, code = readLines("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/SkillAssessment/config/NOBAsacc30Config.R"), eval=F}
 ```
@@
 
-```{r, code = readLines("./config/omdimensions.R"), eval=F}
+```{r, code = readLines("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/SkillAssessment/config/omdimensions.R"), eval=F}
 ```
 
-`codsurvey.R` configures the fishery independent survey and looks like this:
+`mssurvey_fall.R` and `mssurvey_spring.R`configure the fishery independent surveys and look like this:
 
-```{r, code = readLines("./config/codsurvey.R"), eval=F}
+```{r, code = readLines("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/SkillAssessment/config/mssurvey_fall.R"), eval=F}
 ```
 
-`codfishery.R` configures the fishery dependent data and looks like this:
-
-```{r, code = readLines("./config/codfishery.R"), eval=F}
+```{r, code = readLines("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/SkillAssessment/config/mssurvey_spring.R"), eval=F}
 ```
 
-### Using the wrappers (now updated for annual age)
+`msfishery.R` configures the fishery dependent data and looks like this:
 
-All the config files go in the config folder and we'll try running this:
+```{r, code = readLines("https://raw.githubusercontent.com/ices-eg/wg_WGSAM/master/SkillAssessment/config/msfishery.R"), eval=F}
+```
 
-```{r full-test, message=FALSE, warning=FALSE}
+### Running atlantisom and location of output files
 
-#library(here) now above
-library(tidyverse)
-library(atlantisom)
-library(ggthemes)
+I ran this for the full 11 species system on wg-WGSAM Skill Assessment using code [here](https://ices-eg.github.io/wg_WGSAM/SkillAssessProject.html), in the section **Run atlantisom and save outputs** and uploaded the outputs to google drive in the [WGSAM skill assessment folder](https://drive.google.com/drive/folders/1Pv4dSdMKSHUUSUY2bhPOE3Sm1Xe6m4zD).
 
-NOBAom <- om_init(here("config/NOBAsacc30config.R"))
+NOBA atlantisom outputs for run sacc_30 with climate are [here](https://drive.google.com/drive/folders/1PBpgS36YYeQNwyOBzo9HZlzMDIW8DLqI).
 
-NOBAom_cod <- om_species(c("North_atl_cod"), NOBAom)
+NOBA Atlantis model outputs and input/parameter files for run sacc_30 with climate are [here](https://drive.google.com/drive/folders/1zyGSBIZze3-5ejXNdqrpa9YxQbqcOK2G).
 
-NOBAom_cod_ind <- om_index(usersurvey = here("config/codsurvey.R"), 
-                           userfishery = here("config/codfishery.R"),
-                           omlist_ss = NOBAom_cod, 
-                           n_reps = 1, 
-                           save = TRUE)
+For the scripts below to work, ensure that the atlantisom and Atlantis files are available locally in the folder designated `d.name` in the `NOBAsacc30Config.R` script.
+Mine were produced in a `SkillAssessment` directory with subdirectories `atlantisoutput/NOBA_sacc_30` .
+The `config` folder was also in the `SkillAssessment` directory.
+On this repository, there is no interim `SkillAssessment` folder, just the `atlantisoutput` and `config` directories on the same level.
 
-NOBAom_cod_comp <- om_comps(usersurvey = here("config/codsurvey.R"), 
-                           userfishery = here("config/codfishery.R"),
-                           omlist_ss = NOBAom_cod, 
-                           n_reps = 1, 
-                           save = TRUE)
+Future work: integrate with the [`atlantisdrive` package](https://noaa-edab.github.io/atlantisdrive/index.html) Andy Beet has developed to upload and download specific Atlantis files.
+For now just use `googledrive` functions.
+
+```{r, eval=FALSE}
+
+# from https://community.rstudio.com/t/how-to-download-a-google-drives-contents-based-on-drive-id-or-url/16896/12
+
+library(googledrive)
+library(purrr)
+
+## store the URL you have (example, the one with atlantisom output above)
+folder_url <- "https://drive.google.com/drive/folders/1PBpgS36YYeQNwyOBzo9HZlzMDIW8DLqI"
+
+## identify this folder on Drive
+## let googledrive know this is a file ID or URL, as opposed to file name
+folder <- drive_get(as_id(folder_url))
+
+## identify the csv files in that folder
+all_files <- drive_ls(folder)
+
+setwd(here("atlantisoutput/NOBA_sacc_30")) #a hack to get them to download to a folder
+
+## download them
+walk(all_files$id, ~ drive_download(as_id(.x))
+
+#working directory should reset when chunk exits
+```
+
+### Trim to smaller time series
+
+Here I'm just suggesting how to cut the dataset down to the right time period to miss squirrely stuff on either end of the Atlantis run.
+The "data" has been pulled using the full Atlantis run so in the future we can make this part of the survey and fishery design, but for now I wanted the flexibility to use different parts of the full time series.
+
+Source the config file to get the right directories
+```{r}
+source(here("config/NOBAsacc30Config.R"))
+```
+
+Read in the om_list file, here called NOBAom_ms
+```{r}
+NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+```
+
+Adjust this script as need be for start and end times.
+Years 40 to 110 capture the climate trend and leave out the strangest fishery anomalies from spinup.
+This still starts with a period of 0 fishing.
+
+```{r modtime}
+
+fitstartyr <- 40
+fitendyr <- 110  #used 120 for sacc38
+
+#Number of years of data to pull
+nyears <- NOBAom_ms$runpar$nyears
+total_sample <- NOBAom_ms$runpar$tstop/NOBAom_ms$runpar$outputstep
+stepperyr <- if(NOBAom_ms$runpar$outputstepunit=="days") 365/NOBAom_ms$runpar$toutinc
+
+atlantis_full <- c(0:total_sample)  
+mod_burnin <- fitstartyr*stepperyr+1
+fit_nyears <- fitendyr-fitstartyr
+fit_ntimes <- fit_nyears*stepperyr
+fittimes <- atlantis_full[mod_burnin:(mod_burnin+fit_ntimes-1)]
+#fit_timesteps <- seq(fittimes[stepperyr], max(fittimes), by=stepperyr) #last timestep
+fit_years <- unique(floor(fittimes/stepperyr)) #from Christine's new sardine_config.R
+fittimes.days <- fittimes*73
 
@@
 
-### Wrapper test results {.tabset}
+### Plotting functions
 
-#### Biomass index
-```{r bioind-1}
+A collection of functions used previosuly that may be harvested and modified for diagnostics or visualizations in the ms-keyrun and ICES WGSAM skill assessment projects.
 
-omlist_ss <- NOBAom_cod
-source(here("config/omdimensions.R"))
+These functions assume you are using `atlantisom` output datasets but they could be generalized.
 
-#read time series data
-survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name, "surveyB.rds")))
+```{r plotfuns}
 
-survObsBiom <- survObsBiom[[1]]
+# plot biomass time series facet wrapped by species
+plotB <- function(dat, truedat=NULL){
 
-plotB <-ggplot() +
-  geom_line(data=survObsBiom, aes(x=time/stepperyr,y=atoutput, color="survey Biomass"), 
+    ggplot() +
+    geom_line(data=dat, aes(x=time/stepperyr,y=atoutput, color="Survey Biomass"), 
               alpha = 10/10) +
+    {if(!is.null(truedat)) geom_line(data=truedat, aes(x=time/365,y=atoutput, color="True B"), alpha = 3/10)} + 
     theme_tufte() +
     theme(legend.position = "top") +
-  labs(colour=scenario.name)
+    xlab("model year") +
+    ylab("tons") +
+    labs(colour=scenario.name) +
+    facet_wrap(~species, scales="free")
 
-plotB +
+}
+
+# make a catch series function that can be split by fleet? this doesnt
+# also note different time (days) from model timestep in all other output
+plotC <- function(dat, truedat=NULL){
+
+    ggplot() +
+    geom_line(data=dat, aes(x=time/365,y=atoutput, color="Catch biomass"), 
+              alpha = 10/10) +
+    {if(!is.null(truedat)) geom_line(data=truedat, aes(x=time/365,y=atoutput, color="True Catch"), alpha = 3/10)} + 
+    theme_tufte() +
+    theme(legend.position = "top") +
+    xlab("model year") +
+    ylab("tons") +
+    labs(colour=scenario.name) +
     facet_wrap(~species, scales="free")
+
+}
+
+# note on ggplot default colors, can get the first and second using this
+# library(scales)
+# show_col(hue_pal()(2))
+
+# plot length frequencies by timestep (one species)
+plotlen <- function(dat, effN=1, truedat=NULL){
+
+  cols <- c("Census Lcomp"="#00BFC4","Sample Lcomp"="#F8766D")  
+  ggplot(mapping=aes(x=upper.bins)) +
+    {if(is.null(truedat)) geom_bar(data=dat, aes(weight = atoutput/effN))} +
+    {if(!is.null(truedat)) geom_bar(data=dat, aes(weight = censuslen/totlen, fill="Census Lcomp"), alpha = 5/10)} +
+    {if(!is.null(truedat)) geom_bar(data=dat, aes(weight = atoutput/effN, fill="Sample Lcomp"), alpha = 5/10)} +
+    theme_tufte() +
+    theme(legend.position = "bottom") +
+    xlab("length (cm)") +
+    {if(is.null(truedat)) ylab("number")} +
+    {if(!is.null(truedat)) ylab("proportion")} +
+    scale_colour_manual(name="", values=cols) +
+    labs(subtitle = paste(scenario.name,
+                          dat$species)) +
+    facet_wrap(~time, ncol=6, scales="free_y")
+
+}
+
+# plot numbers at age by timestep (one species)
+Natageplot <- function(dat, effN=1, truedat=NULL){
+  ggplot() +
+    geom_point(data=dat, aes(x=agecl, y=atoutput/effN, color="Est Comp")) +
+    {if(!is.null(truedat)) geom_line(data=dat, aes(x=agecl, y=numAtAge/totN, color="True Comp"))} + 
+    theme_tufte() +
+    theme(legend.position = "bottom") +    
+    xlab("age/agecl") +
+    {if(is.null(truedat)) ylab("number")} +
+    {if(!is.null(truedat)) ylab("proportion")} +
+    labs(subtitle = paste(scenario.name,
+                          dat$species)) + 
+    facet_wrap(~time, ncol=6, scales="free_y")
+}
+
+# plot weight at age time series facet wrapped by species
+wageplot <- function(dat, truedat=NULL){
+  ggplot(dat, aes(time/stepperyr, atoutput)) +
+    geom_line(aes(colour = factor(agecl))) +
+    theme_tufte() +
+    theme(legend.position = "bottom") +
+    xlab("model year") +
+    ylab("average individual weight (g)") +
+    labs(subtitle = paste0(scenario.name)) +
+    facet_wrap(c("species"), scales="free_y")
+}
+
+# compare N at age and C at age between standard and ANNAGE outputs
+totNageplot <- function(dat, anndat){
+
+    ggplot() +
+    geom_line(data=dat, aes(x=time/stepperyr,y=totN, color="Tot N cohorts"), 
+              alpha = 5/10) +
+    geom_line(data=anndat, aes(x=time/stepperyr,y=totN, color="Tot N annage"), alpha = 5/10) + 
+    theme_tufte() +
+    theme(legend.position = "top") +
+    xlab("model year") +
+    ylab("N") +
+    labs(colour=scenario.name) +
+    facet_wrap(~species, scales="free")
+
+}
 
@@
 
-#### Catch time series
-```{r catchind-1}
-#read time series data
-catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))
+### Read in the "data" to check with plots
 
-catchbio_ss <- catchbio_ss[[1]]
+Note that these are multispecies survey and fishery datasets.
 
-plotC <-ggplot() +
-  geom_line(data=catchbio_ss, aes(x=time/365,y=atoutput, color="observed Catch"), 
-            alpha = 10/10) +
-  theme_tufte() +
-  theme(legend.position = "top") +
-  labs(colour=scenario.name)
+They can be filtered for species=="North_atl_cod".
 
-plotC +
-  facet_wrap(~species, scales="free")
+Each object is a list of lists with the survey or fishery names being the first list element and the data being the first element (a dataframe) under the name (this is because we can generate multiple replicates but we only have one saved here).
+
+```{r}
+
+survObsBiom <- atlantisom::read_savedsurvs(d.name, 'survB')
+age_comp_data <- atlantisom::read_savedsurvs(d.name, 'survAge') #not using in assessment
+len_comp_data <- atlantisom::read_savedsurvs(d.name, 'survLen')
+#wtage <- atlantisom::read_savedsurvs(d.name, 'survWtage')  #not using in assessment
+annage_comp_data <- atlantisom::read_savedsurvs(d.name, 'survAnnAge')
+annage_wtage <- atlantisom::read_savedsurvs(d.name, 'survAnnWtage')
+
+#all_diets <- atlantisom::read_savedsurvs(d.name, 'survDiet') #not using in assessment
+
+catchbio_ss <- atlantisom::read_savedfisheries(d.name, 'Catch')
+catchlen_ss <- atlantisom::read_savedfisheries(d.name, "catchLen")
+fish_age_comp <- atlantisom::read_savedfisheries(d.name, "catchAge")
+fish_annage_comp <- atlantisom::read_savedfisheries(d.name, 'catchAnnAge')
+fish_annage_wtage <- atlantisom::read_savedfisheries(d.name, 'catchAnnWtage')
+
+```
+
+### Visualize survey outputs {.tabset}
+
+We filter the full datasets using the times defined above. These plots represent the selected time series for the survey biomass index and weight at age, and a subset for length and age composition.
+
+#### Survey biomass index
+```{r, results="asis"}
+
+# compare with true output (all timesteps)
+# for(s in names(survObsBiom)){
+#   cat("  \n##### ",  s,"  \n")
+#   print(plotB(survObsBiom[[s]][[1]], omlist_ss$truetotbio_ss))
+#   cat("  \n")
+# }
+
+# plots survey only
+ for(s in names(survObsBiom)){
+   cat("  \n##### ",  s,"  \n")
+   print(plotB(survObsBiom[[s]][[1]] %>%
+                 filter((time %in% fittimes))))
+   cat("  \n")
+ }
 
@@
 #### Survey length composition
-```{r lencomp-1}
+```{r lencomp-1, results="asis", message=FALSE, warning=FALSE}
 
-#length comps
-len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsLenComp.rds")))
-fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))
+# not the full time series, just first 24 yrs
+for(s in names(len_comp_data)){
+  cat("  \n##### ",  s,"  \n")
+  lcompsub <- as.data.frame(len_comp_data[[s]][[1]]) %>% filter(time %in% c(200:320)) %>%
+    group_by(species) %>%
+    group_map(~ plotlen(.x), keep = TRUE)
 
-len_comp_data <- len_comp_data[[1]]
-fish_len_comp_data <- fish_len_comp_data[[1]]
-
-#add this to om_indices function so that this has years when read in
-fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))
-
-len <- filter(len_comp_data, time %in% c(55:175))
-
-  lfplot <- ggplot(len, aes(upper.bins)) +
-    geom_bar(aes(weight = atoutput)) +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          len$species))
-
-  lfplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+  for(i in 1:length(lcompsub)) {
+    print(lcompsub[[i]])
+  }
+  cat("  \n")
+}
 
@@
 
-#### Survey age composition (age classes)
-```{r agecomp-1}
+#### Survey age composition (annual ages)
+```{r plot-annagecomp, results="asis", message=FALSE, warning=FALSE}
 
-#read in comp data
-age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
-age_comp_data <- age_comp_data[[1]]
+for(s in names(annage_comp_data)){
+  cat("  \n##### ",  s,"  \n")
+  acompsub <- as.data.frame(annage_comp_data[[s]][[1]]) %>% filter(time %in% c(200:320)) %>%
+    group_by(species) %>%
+    #left_join(., trueNage) %>%
+    group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+    #group_map(~ Natageplot(.x, effN = 100000, truedat = 1), keep = TRUE) # plots merged true age comp
 
-Natage <- filter(age_comp_data, time %in% c(150:270))
-
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Natage$species))
-
-  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
-```
-
-#### Survey weight at age (age classes)
-```{r wtageclass-1}
-
-wtage <- readRDS(file.path(d.name, paste0(scenario.name, "survObsWtAtAge.rds")))
-wtage <- wtage[[1]]
-
-wageplot <- ggplot(wtage, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (5 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-```
-
-#### Fishery age composition (age classes)
-```{r fishagecomp-1}
-
-#read in comp data
-fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
-fish_age_comp <- fish_age_comp[[1]]
-
-#add this to om_indices function so that this has years when read in
-fish_age_comp$time <- fish_age_comp$time/fstepperyr
-
-Natage <- filter(fish_age_comp, time %in% c(30:53))
-
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Natage$species))
-
-  Natageplot + facet_wrap(~time, ncol=6, scales="free_y")
-```
-
-### {-}
-
-Wrappers still work for NOBA cod using standard age classes.
-
-Visualize new annual age output for numbers at age, catch at age, and weight at age (interpolated):
-
-### Wrapper test results for annual ages {.tabset #annageoutput}
-
-#### Survey age composition (annual ages)
-```{r plot-annagecomp}
-#read in comp data
-annage_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
-annage_comp_data <- annage_comp_data[[1]]
-
-#fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
-#fish_age_comp <- fish_age_comp[[1]]
-
-#add this to om_indices function so that this has years when read in
-#fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))
-
-Natage <- filter(annage_comp_data, time %in% c(150:270))
-
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Natage$species))
-
-  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+  for(i in 1:length(acompsub)) {
+    print(acompsub[[i]])
+  }
+  cat("  \n")
+}
 ```
@@
 #### Survey iterpolated weight at age (annual ages) 
-```{r plot-annwtage}
+```{r plot-annwtage, fig.height=10, results='asis'}
 
-wtage <- readRDS(file.path(d.name, paste0(scenario.name, "survObsFullWtAtAge.rds")))
-wtage <- wtage[[1]] #this still has second list component, diagnostic plot
-wtage <- wtage[[1]]
-
-wageplot <- ggplot(wtage, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (5 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
-
+for(s in names(annage_wtage)){
+  cat("  \n##### ",  s,"  \n")
+  print(wageplot(annage_wtage[[s]][[1]] %>%
+                 filter((time %in% fittimes))))
+  cat("  \n")
+}
 ```
 
-#### Fishery catch at age (annual ages)
-```{r plot-catchagecomp}
+#### Survey age class (Atlantis agecl--check against full survey age comp)
+```{r, results="asis", message=FALSE, warning=FALSE}
 
-fish_annage_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullAgeComp.rds")))
-fish_annage_comp <- fish_annage_comp[[1]]
+for(s in names(age_comp_data)){
+  cat("  \n##### ",  s,"  \n")
+  acompsub <- as.data.frame(age_comp_data[[s]][[1]]) %>% filter(time %in% c(200:320)) %>%
+    group_by(species) %>%
+    #left_join(., trueNagecl) %>%
+    group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+    #group_map(~ Natageplot(.x, effN = 100000, truedat = 1), keep = TRUE) # plots merged true age comp
 
-#add this to om_indices function so that this has years when read in
-fish_annage_comp$time <- fish_annage_comp$time/fstepperyr
-
-Catage <- filter(fish_annage_comp, time %in% c(30:53))
-
-Catageplot <- ggplot(Catage, aes(x=agecl, y=atoutput)) +
-    geom_point() +
-    theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Catage$species))
-
-  Catageplot + facet_wrap(~time, ncol=6, scales="free_y")
-```
-
-#### Fishery iterpolated weight at age (annual ages) 
-```{r plot-fishannwtage}
-
-wtage <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsFullWtAtAge.rds")))
-wtage <- wtage[[1]] #this still has second list component, diagnostic plot
-wtage <- wtage[[1]]
-
-wageplot <- ggplot(wtage, aes(time, atoutput)) +
-  geom_line(aes(colour = factor(agecl))) +
-  theme_tufte() +
-  theme(legend.position = "bottom") +
-  xlab("model timestep (5 per year)") +
-  ylab("average individual weight (g)") +
-  labs(subtitle = paste0(scenario.name, " annual mid year sample"))
-
-wageplot + facet_wrap(c("species"), scales="free_y")
+  for(i in 1:length(acompsub)) {
+    print(acompsub[[i]])
+  }
+  cat("  \n")
+}
 
@@
 
-We previously noted that NOBA didn't have a selective fishery for cod; there were lots of age 1 in the catch! With this new run, we see the effects of selective fisheries
+### Visualize fishery data {.tabset}
+
+These plots represent the selected time series for fishery catch and weight at age, and a subset for length and age composition.
+Fishery lengths and ages are sampled 5x per year from this model, while the surveys above sample 1x per year.
+
+#### Fishery catch time series
+```{r catchind-1, results="asis", message=FALSE, warning=FALSE}
+# observed catch only
+plotC(catchbio_ss$census[[1]] %>%
+        filter(time %in% fittimes.days))
+```
+
+#### Fishery length composition
+```{r, results="asis", message=FALSE, warning=FALSE}
+
+lcompsub <- as.data.frame(catchlen_ss$census[[1]]) %>% filter(time %in% c(291:320)) %>%
+  group_by(species) %>%
+  group_map(~ plotlen(.x), keep = TRUE)
+
+for(i in 1:length(lcompsub)) {
+  print(lcompsub[[i]])
+}
+
+```
+
+#### Fishery catch at age (annual ages)
+```{r plot-catchagecomp, results='asis'}
+
+acompsub <- as.data.frame(fish_annage_comp$census[[1]]) %>% filter(time %in% c(291:320)) %>%
+  group_by(species) %>%
+  #left_join(., trueCAA) %>%
+  group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+  #group_map(~ Natageplot(.x, effN = 200, truedat = 1), keep = TRUE) # plots merged true age comp
+
+for(i in 1:length(acompsub)) {
+  print(acompsub[[i]])
+}
+
+```
+
+#### Fishery interpolated weight at age (annual ages)
+```{r, fig.height=10}
+
+wageplot(fish_annage_wtage$census[[1]] %>%
+                 filter((time %in% fittimes)))
+
+```
+
+#### Fishery catch at age class (Atlantis agecl--check against full age comp)
+```{r, message=FALSE, warning=FALSE}
+
+acompsub <- as.data.frame(fish_age_comp$census[[1]]) %>% filter(time %in% c(291:320)) %>%
+  group_by(species) %>%
+  #left_join(., trueCAAcl) %>%
+  group_map(~ Natageplot(.x), keep = TRUE) # plots only sampled age comp
+  #group_map(~ Natageplot(.x, effN = 200, truedat = 1), keep = TRUE) # plots merged true age comp
+
+for(i in 1:length(acompsub)) {
+  print(acompsub[[i]])
+}
+
+```
+
+### {-}
+
+We previously noted that NOBA didn't have a selective fishery for cod; there were lots of age 1 in the catch!
+With this run, we see some effects of selective fisheries, and 4 different fleets targeting cod.
 
@@
 
-**THIS IS STILL THE CALIFORNIA CURRENT INFO, TO BE UPDATED**
+**I HAVE NOT TOUCHED ANYTHING BELOW**
 

### 2021-09-30 CheckFisheryAnnageComp.Rmd
cod CAA from ANNAGECATCH diagnosis part I

--- 2021-09-29 NOBAcod2.Rmd
@@
 ---
-title: "Generate data for NOBA cod SS3 modeling"
+title: "Investigate fishery annual age comp issues"
 author: "Sarah Gaichas"
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE,
+                      warning = FALSE,
+                      message = FALSE)
 library(here)
@@
 
+There appears to be a problem with the fishery annual age output, especially for NOBA cod which has multiple fleets fishing it.
+Fishery annual age comps generated for "North_atl_cod" in Sept 2021 from NOBA sacc_38 and sacc_30 (without and with climate signal, but otherwise the same run) have the highest numbers for the oldest ages, which matches neither the aggregated fishery agecl age output nor the length comps generated from atlantisom.
+We'll try to figure out the problem and a fix here.
+
 ## New NOBA run with updated selectivity
@@
 
-This dataset will come from NOBA sacc_30, with original biology, planned F + changes in selectivity for multiple groups, forced recruitment for groups	SCR+SSH+NCO, and climate change from	1981-2068+loop 2067..
+This dataset will come from NOBA sacc_30, with original biology, planned F + changes in selectivity for multiple groups, forced recruitment for groups	SCR+SSH+NCO, and climate change from	1981-2068+loop 2067.
+
+### What is in the ANNAGECATCH.nc file?
+
+Source the config file to get the right directories
+```{r}
+source(here("config/NOBAsacc30Config.R"))
+```
+
+Load the `ANNAGECATCH.nc` file only for cod.
+```{r}
+
+fgs <- atlantisom::load_fgs(d.name, functional.groups.file)
+
+bps <- atlantisom::load_bps(d.name, functional.groups.file, biomass.pools.file)
+
+biol <- atlantisom::load_biolprm(d.name, biol.prm.file)
+
+allboxes <- atlantisom::load_box(d.name, box.file)
+boxes <- atlantisom::get_boundary(allboxes)
+
+annagecatchnc <- atlantisom::load_nc_annage(dir = d.name, 
+                                            file_nc = paste0(scenario.name, "ANNAGECATCH.nc"),
+                                            file_fish = fisheries.file, 
+                                            bps = bps,
+                                            fgs = fgs,
+                                            biolprm = biol,
+                                            select_groups = "North_atl_cod",
+                                            select_variable = "Catch",
+                                            bboxes = boxes)
+
+```
+
+Visualize catch at age by fleet from full "true" output
+```{r}
+
+plotCAA <- function(dat){
+  ggplot(dat, aes(x=agecl, y=atoutput)) +
+  geom_point() +
+  theme_tufte() +
+  facet_wrap(~fleet) +
+  labs(subtitle = dat$species)
+}
+
+```
+
+At time = 1
+```{r}
+plotCAA(annagecatchnc %>%
+          filter(time==1) %>%
+          group_by(species, agecl, fleet, time) %>%
+          summarise(atoutput = sum(atoutput)))
+```
+
+At time = 200
+```{r}
+plotCAA(annagecatchnc %>%
+          filter(time==200) %>%
+          group_by(species, agecl, fleet, time) %>%
+          summarise(atoutput = sum(atoutput)))
+```
+
+At time = 320
+```{r}
+plotCAA(annagecatchnc %>%
+          filter(time==320) %>%
+          group_by(species, agecl, fleet, time) %>%
+          summarise(atoutput = sum(atoutput)))
+```
+
+At time = 490
+```{r}
+plotCAA(annagecatchnc %>%
+          filter(time==490) %>%
+          group_by(species, agecl, fleet, time) %>%
+          summarise(atoutput = sum(atoutput)))
+```
+Patterns here look identical over time, with a huge amount of catch of 17-20 year olds in the dtrawlNCO fishery.
+
+Is this the same as what we would see in the omlist_ss output?
+
+Read in the om_list file, here called NOBAom_ms
+```{r}
+NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
+```
+
+The true catch at annual age in this list is `NOBAom_ms$truecatchage_ss`, lets plot it
+At time = 1
+```{r}
+plotCAA(NOBAom_ms$truecatchage_ss %>%
+          filter(time==1,
+                 species=="North_atl_cod") %>%
+          group_by(species, agecl, fleet, time) %>%
+          summarise(atoutput = sum(atoutput)))
+```
+At time = 200
+```{r}
+plotCAA(NOBAom_ms$truecatchage_ss %>%
+          filter(time==200,
+                 species=="North_atl_cod") %>%
+          group_by(species, agecl, fleet, time) %>%
+          summarise(atoutput = sum(atoutput)))
+```
+
+At time = 320
+```{r}
+plotCAA(NOBAom_ms$truecatchage_ss %>%
+          filter(time==320,
+                 species=="North_atl_cod") %>%
+          group_by(species, agecl, fleet, time) %>%
+          summarise(atoutput = sum(atoutput)))
+```
+
+At time = 490
+```{r}
+plotCAA(NOBAom_ms$truecatchage_ss %>%
+          filter(time==490,
+                 species=="North_atl_cod") %>%
+          group_by(species, agecl, fleet, time) %>%
+          summarise(atoutput = sum(atoutput)))
+```
+Yes, these are identical.
+
+Look at the file after create_fishery_subset
+
+Look at the file after sample_fish
+
+Compare with atlantisom output below
 
@@
 The "data" has been pulled using the full Atlantis run so in the future we can make this part of the survey and fishery design, but for now I wanted the flexibility to use different parts of the full time series.
-
-Source the config file to get the right directories
-```{r}
-source(here("config/NOBAsacc30Config.R"))
-```
-
-Read in the om_list file, here called NOBAom_ms
-```{r}
-NOBAom_ms <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))
-```
 

### 2021-09-30 docs/index.Rmd
add noba cod2 to index
cod CAA from ANNAGECATCH diagnosis part I

@@
 <a href="https://sgaichas.github.io/poseidon-dev/atlantisom_landingpage.html" target="_blank">Draft atlantisom landing page for review, June 9, 2020</a>
+
+### NOBA cod dataset take 2
+
+Need to diagnose the annage outputs; they look strange
+
+<a href="https://sgaichas.github.io/poseidon-dev/CheckFisheryAnnageComp.html" target="_blank"> Starting cod catch at annual age diagnosis, Sept 30, 2021</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NOBAcod2.html" target="_blank"> Multispecies dataset from NOBA sacc_30 with climate and fishery selectivity, Sept 29, 2021</a>
 

### 2022-06-23 PerCapitaConsumption.Rmd
added new totcons but wont knit

@@
 ---
-title: "Per captia consumption"
+title: "Per capita consumption"
 author: "Sarah Gaichas"
@@
 
-## Diet info for multispecies modeling
+# Consumption info for multispecies modeling
 
@@
 
+## Approach 1 based on detailedDietCheck.txt
+
 What we have saved so far is a diet comp in proportion that has had bias and observation error added to proportions.
@@
 
-The function below should return per-capita consumption by age class. This can then be converted to lengthbin as necessary.
+The function below should return per-capita consumption by age class using the `detailedDietCheck.txt` output. This can then be converted to lengthbin as necessary.
 
@@
 
-## NOBA
+### NOBA
 
@@
 
-## CCA
+### CCA
 
@@
 ```
+
+## Approach 2 based on PROD.nc
+
+An existing `atlantisom` function `calc_pred_diet()` was [tested and found to produce consumption estimates orders of magnitude higher](https://sgaichas.github.io/poseidon-dev/NOBAmsdiets.html) than those derived from `detailedDietCheck.txt`.
+
+Here we develop new functions that don't apply global diet composition to consumption but rather just get total consumption from PROD.nc.
+The challenge is figuring out what the units are of "Eat" and "Grazing" and how to appropriately expand them to the population level (water volume or numbers in a polygon).
+
+First we need to get "Eat" and "Grazing" from PROD.nc and "Vol" from the .nc file.
+Here is our model setup from the config file.
+Lets look at NOBA_sacc_38, which is being used for mskeyrun simulated dataset generation.
+**THIS OUTPUT IS NOT COMPARABLE TO NOBA_March_2020 above**.
+
+```{r}
+
+# WARNING!
+# hardcoding because making copies of this in each repo is too cumbersome
+# this currently lives in mskeyrun to make the simulated datasets with no climate
+# change as appropriate!
+
+dir <- "/Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38"
+
+file_fgs <- "nordic_groups_v04.csv"
+file_biolprm <- "nordic_biol_incl_harv_v_011_1skg.prm"
+file_bgm <- "Nordic02.bgm"
+file_init <- "nordic_biol_v23.nc"
+file_runprm <- "nordic_run_v01.xml"
+scenario <- "nordic_runresults_01"
+
+select_groups <- mskeyrun::simFocalSpecies$Name # 11 mskeyrun fish species
+
+verbose <- FALSE
+
+```
+
+This code is part of `atlantisom::run_truth()`. With our selected species from `mskeyrun`, only `eat` will be created. `grazing` applies to non-age structured groups.
+
+**If this works for fish, test for inverts too**
+
+```{r}
+
+  # Read in the functional groups csv since that is used by many functions
+  fgs <- load_fgs(dir = dir, file_fgs = file_fgs)
+  # Read in the biomass pools
+  bps <- load_bps(dir = dir, fgs = file_fgs, file_init = file_init)
+  # Read in the biological parameters
+  biol <- load_biolprm(dir = dir, file_biolprm = file_biolprm)
+  # Read in the run parameters
+  runprm <- load_runprm(dir = dir, file_runprm = file_runprm)
+
+  nc_out <- paste0(scenario, ".nc")
+  nc_prod <- paste0(scenario, "PROD.nc")
+
+  # Get the boundary boxes
+  allboxes <- load_box(dir = dir, file_bgm = file_bgm)
+  boxes <- get_boundary(allboxes)
+
+  eat <- load_nc(dir = dir,
+                     file_nc = nc_prod,
+                     bps = bps,
+                     fgs = fgs,
+                     select_groups = select_groups,
+                     select_variable = "Eat",
+                     check_acronyms = TRUE,
+                     bboxes = boxes)
+  if(verbose) message("Eaten read in.")
+
+  grazing <- load_nc(dir = dir,
+                 file_nc = nc_prod,
+                 bps = bps,
+                 fgs = fgs,
+                 select_groups = select_groups,
+                 select_variable = "Grazing",
+                 check_acronyms = TRUE,
+                 bboxes = boxes)
+  if(verbose) message("Grazing read in.")
+
+  vol <- load_nc_physics(dir = dir,
+                         file_nc = nc_out,
+                         physic_variables = "volume",
+                         aggregate_layers = FALSE,
+                         bboxes = boxes)
+  if(verbose) message("Volume read in.")
+
+```
+
+Now try the new function `atlantisom::calc_pred_cons()` which uses the volume of all water column layers to expand the atoutput (assumed to be mg N per cubic m at that timestep) to bio_eaten in tons.
+
+This is the function code
+
+```{r, code = readLines("https://github.com/r4atlantis/atlantisom/raw/master/R/calc_pred_cons.R"), eval=F}
+```
+
+This is the current result
+
+```{r}
+
+biomass_eaten <- calc_pred_cons(eat = eat, grazing = grazing, vol = vol, biolprm = biol, runprm = runprm)
+
+codtotcons <- biomass_eaten %>% 
+  dplyr::filter(species=="North_atl_cod") %>%
+  dplyr::group_by(time, agecl) %>% 
+  dplyr::summarise(totcons = sum(bio_eaten))
+
+```
+
+Plotting a cod example total consumption by age class
+
+```{r}
+ggplot2::ggplot(codtotcons, aes(x=time, y=totcons)) + geom_line() + facet_wrap(~agecl, scales="free")
+
+```

### 2022-06-30 NewCCASardines.Rmd
add new CC run

---
title: "New CCA sardines"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)  
library(atlantisom)
library(ggthemes)
library(FSA)

```

A test run for sardines with the updated California Current model, run 44:

```{r full-test, message=FALSE, warning=FALSE, eval=FALSE}

CCom <- atlantisom::om_init(here("config/CCConfigv4.R"))

CCom_sard <- atlantisom::om_species(c("Pacific_sardine"), CCom)

CCom_sard_ind <- atlantisom::om_index(usersurvey = here("config/sardinesurvey.R"), 
                           userfishery = here("config/sardinefishery.R"),
                           omlist_ss = CCom_sard, 
                           n_reps = 1, 
                           save = TRUE)

CCom_sard_comp <- atlantisom::om_comps(usersurvey = here("config/sardinesurvey.R"), 
                           userfishery = here("config/sardinefishery.R"),
                           omlist_ss = CCom_sard, 
                           n_reps = 1, 
                           save = TRUE)

```

Diagnose why survey functions resulting in NA outputs. Found it: change `sapply` to `lapply` in omdimensions to make the survey selectivity dataframe. The `sapply` function behaves inconsistently when the age classes are all 10, it orients the list incorrectly; `lapply` keeps the structure correct. This only worked for NOBA because I selected a species with fewer than 10 age classes.

```{r testomindex, eval=FALSE}

usersurvey = here("config/sardinesurvey.R")

omlist_ss <- CCom_sard

#this is om_index

#one script for dimension parameters to be used in multiple functions
  source("config/omdimensions.R", local = TRUE)

  # user options for survey--default is a census with mid-year sample
  # allows muliple surveys
  survObsBiomBs <- list()

  for (s in usersurvey)
  {
    source(s, local = TRUE)

    #biomass based fishery independent survey index
    # this uses result$biomass_ages to sample biomass directly, no need for wt@age est
    survey_B <- atlantisom::create_survey(dat = omlist_ss$truebio_ss,
                                          time = survtime,
                                          species = survspp,
                                          boxes = survboxes,
                                          effic = surveffic,
                                          selex = survselex)

    # call sample_survey_biomass with a bunch of 1000s for weight at age
    # in the code it multiplies atoutput by wtatage/1000 so this allows us to use
    # biomass directly
    wtage <- data.frame(species=rep(names(age_classes), n_age_classes),
                        agecl=unlist(sapply(n_age_classes,seq)),
                        wtAtAge=rep(1000.0,sum(n_age_classes)))

    # this is the step to repeat n_reps time if we want different realizations
    # of the same survey design specified above; only observation error differs
    # using the census cv of 0 will produce identical reps!
    survObsBiomB <- list()
    for(i in 1:n_reps){
      survObsBiomB[[i]] <- atlantisom::sample_survey_biomass(survey_B, surv_cv, wtage)
    }

    #save survey indices, takes a long time to generate with lots of reps/species
    if(save){
      saveRDS(survObsBiomB, file.path(d.name, paste0(scenario.name, "_",
                                                     survey.name, "surveyB.rds")))
    }

    survObsBiomBs[[survey.name]] <- survObsBiomB
  }

  #configure the fishery, a default is in config/fisherycensus.R
  #fishery configuration can specify only area and time of observation
  #fishery species inherited from omlist_ss
  #this is total catch not by fleet, so only one "fishery"
  source(userfishery, local = TRUE)

  #we are not currently subsetting fishery catch because we cannot correct catch.nc
  #  instead the catch in biomass from catch.txt is read in for the index
  #  we do not apply any cv to this, but we could this way (default cv=0)

  fishObsCatchB <- list()
  for(i in 1:n_reps){
    fishObsCatchB[[i]] <- atlantisom::sample_fishery_totcatch(omlist_ss$truecatchbio_ss, fish_cv)
  }

  if(save){
    saveRDS(fishObsCatchB, file.path(d.name, paste0(scenario.name, "_",
                                                    fishery.name, "fishCatch.rds")))
  }

  indices <- list("survObsBiomB" = survObsBiomBs,
                  "fishObsCatchB" = fishObsCatchB)

  return(indices)
}

```

Diagnose why comps aren't working. Survey worked, fishery didn't because the burnin period was longer than this run. Changed to 0 in the sardinefishery.R config file and runs.

```{r testomcomps, eval=FALSE}

usersurvey  <-  here("config/sardinesurvey.R")
userfishery <- here("config/sardinefishery.R")

omlist_ss <- CCom_sard

n_reps <- 1

#this is om_comps

 #one script for dimension parameters to be used in multiple functions
  source("config/omdimensions.R", local = TRUE)

  # user options for survey--default is a census with mid-year sample
  # allows muliple surveys
  age_comp_datas <- list()
  survObsLenComps <- list()
  survObsWtAtAges <- list()

  for (s in usersurvey)
  {
    source(s, local = TRUE)

    #numbers based fishery independent survey for age and length comps
    # same user specifications as indices
    survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
                                          time = survtime,
                                          species = survspp,
                                          boxes = survboxes,
                                          effic = surveffic,
                                          selex = survselex.agecl)

    #Sample fish for age composition
    # if we want replicates for obs error this sample function will generate them
    age_comp_data <- list()
    for(i in 1:n_reps){
      age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
    }

    # save age comps
    if(save){
      saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "_",
                                                      survey.name, "survObsAgeComp.rds")))
    }

    #weights needed for weight at age and length comp calcs
    # aggregate true resn per survey design
    survey_aggresn <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
                                                       time = survtime,
                                                       species = survspp,
                                                       boxes = survboxes)

    # aggregate true structn per survey design
    survey_aggstructn <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
                                                          time = survtime,
                                                          species = survspp,
                                                          boxes = survboxes)

    #dont sample these, just aggregate them using median
    structnss <- atlantisom::sample_fish(survey_aggstructn, surveffN, sample = FALSE)

    resnss <- atlantisom::sample_fish(survey_aggresn, surveffN, sample = FALSE)

    #this is all input into the length function, replicates follow age comp reps
    #  separating the length comps from the weight at age here
    survey_lenwt <- list()
    survObsLenComp <- list()
    survObsWtAtAge <- list()

    for(i in 1:n_reps){
      survey_lenwt[[i]] <- atlantisom::calc_age2length(structn = structnss,
                                                       resn = resnss,
                                                       nums = age_comp_data[[i]],
                                                       biolprm = omlist_ss$biol,
                                                       fgs = omlist_ss$funct.group_ss,
                                                       maxbin = maxbin,
                                                       CVlenage = lenage_cv,
                                                       remove.zeroes=TRUE)

      survObsLenComp[[i]] <- survey_lenwt[[i]]$natlength
      survObsWtAtAge[[i]] <- survey_lenwt[[i]]$muweight
    }

    if(save){
      saveRDS(survObsLenComp, file.path(d.name, paste0(scenario.name, "_",
                                                       survey.name, "survObsLenComp.rds")))
      saveRDS(survObsWtAtAge, file.path(d.name, paste0(scenario.name, "_",
                                                       survey.name, "survObsWtAtAge.rds")))
    }
    # add each survey to master list objects for survey data
    age_comp_datas[[survey.name]] <- age_comp_data
    survObsLenComps[[survey.name]] <- survObsLenComp
    survObsWtAtAges[[survey.name]] <- survObsWtAtAge

  }

  #now do fishery comps
  # user options for fishery--default is a census with mid-year sample
  # only one fishery, but multiple fleets possible within it
  source(userfishery, local = TRUE)

  #fishery catch at age each observed timestep summed over observed polygons
  # catch at age by area and timestep
  catch_numbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchnum_ss,
                                                      time = fishtime,
                                                      species = survspp,
                                                      boxes = fishboxes)

  # if we want replicates for obs error this sample function will generate them
  catch_age_comp <- list()
  for(i in 1:n_reps){
    catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
  }

  # save fishery age comps
  if(save){
    saveRDS(catch_age_comp, file.path(d.name, paste0(scenario.name, "_",
                                                     fishery.name, "fishObsAgeComp.rds")))
  }

  #Get catch weights for length comp calc
  # aggregate true resn per fishery subset design
  catch_aggresnss <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
                                                      time = fishtime,
                                                      species = survspp,
                                                      boxes = fishboxes)

  # aggregate true structn fishery subsetdesign
  catch_aggstructnss <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
                                                         time = fishtime,
                                                         species = survspp,
                                                         boxes = fishboxes)

  #dont sample these, just aggregate them using median
  catch_structnss <- atlantisom::sample_fish(catch_aggstructnss, fisheffN, sample = FALSE)

  catch_resnss <- atlantisom::sample_fish(catch_aggresnss, fisheffN, sample = FALSE)

  # these fishery lengths and weight at age are each output timestep
  #same structure as above for surveys, replicates follow age comp reps
  #  separating the length comps from the weight at age here
  fishery_lenwt <- list()
  fishObsLenComp <- list()
  fishObsWtAtAge <- list()

  for(i in 1:n_reps){
    fishery_lenwt[[i]] <- atlantisom::calc_age2length(structn = catch_structnss,
                                                      resn = catch_resnss,
                                                      nums = catch_age_comp[[i]],
                                                      biolprm = omlist_ss$biol,
                                                      fgs = omlist_ss$funct.group_ss,
                                                      maxbin = maxbin,
                                                      CVlenage = lenage_cv,
                                                      remove.zeroes=TRUE)

    fishObsLenComp[[i]] <- fishery_lenwt[[i]]$natlength
    fishObsWtAtAge[[i]] <- fishery_lenwt[[i]]$muweight
  }

  if(save){
    saveRDS(fishObsLenComp, file.path(d.name, paste0(scenario.name, "_",
                                                     fishery.name, "fishObsLenComp.rds")))
    saveRDS(fishObsWtAtAge, file.path(d.name, paste0(scenario.name, "_",
                                                     fishery.name, "fishObsWtAtAge.rds")))
  }

  if(!is.null(omlist_ss$truenumsage_ss)){
    #numbers based fishery independent survey for age and length comps
    #allows for mulitple surveys
    annage_comp_datas <- list()

    for (s in usersurvey)
    {
      source(s, local = TRUE)

      # same user specifications as indices
      survey_annageN <- atlantisom::create_survey(dat = omlist_ss$truenumsage_ss,
                                                  time = survtime,
                                                  species = survspp,
                                                  boxes = survboxes,
                                                  effic = surveffic,
                                                  selex = survselex)
      #Sample fish for age composition
      # if we want replicates for obs error this sample function will generate them
      annage_comp_data <- list()
      for(i in 1:n_reps){
        annage_comp_data[[i]] <- atlantisom::sample_fish(survey_annageN, surveffN)
      }

      # save survey annual age comps
      if(save){
        saveRDS(annage_comp_data, file.path(d.name, paste0(scenario.name, "_",
                                                           survey.name, "survObsFullAgeComp.rds")))
      }
      annage_comp_datas[[survey.name]] <- annage_comp_data
    }

  }else{annage_comp_datas <- NULL}

  if(!is.null(omlist_ss$truecatchage_ss)){
    #fishery catch at age each observed timestep summed over observed polygons
    # catch at age by area and timestep
    catch_annagenumbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchage_ss,
                                                        time = fishtime,
                                                        species = survspp,
                                                        boxes = fishboxes)

    # if we want replicates for obs error this sample function will generate them
    # WARNING THIS AGGREGATES ACROSS FLEETS
    # TODO: need to change sample_fish to fix
    catch_annage_comp <- list()
    for(i in 1:n_reps){
      catch_annage_comp[[i]] <- atlantisom::sample_fish(catch_annagenumbers, fisheffN)
    }

    # save fishery annual age comps
    if(save){
    saveRDS(catch_annage_comp, file.path(d.name, paste0(scenario.name,"_",
                                                        fishery.name, "fishObsFullAgeComp.rds")))
    }
  }else{catch_annage_comp <- NULL}

  # call interpolate weight at age function to get survObsFullWtAtAge
  if(!is.null(omlist_ss$truenumsage_ss)){
    interp_survWtAtAges <- list()
    for (s in usersurvey)
    {
      source(s, local = TRUE)

      interp_survWtAtAge <- list()
      for(i in 1:n_reps){
        interp_survWtAtAge[[i]] <- calc_avgwtstage2age(wtagecl = survObsWtAtAges[[survey.name]][[i]],
                                                       annages = omlist_ss$truenumsage_ss,
                                                       fgs = omlist_ss$funct.group_ss)
      }
      if(save){
        saveRDS(interp_survWtAtAge, file.path(d.name, paste0(scenario.name, "_",
                                                             survey.name, "survObsFullWtAtAge.rds")))
      }
      interp_survWtAtAges[[survey.name]] <- interp_survWtAtAge
    }
  }else{interp_survWtAtAges <- NULL}

  # do we want fishery average weight at true age too? why not
  # call interpolate weight at age function to get fishObsFullWtAtAge
  # WARNING currently aggregates out fleet info, but no fleets in aggregate wtage
  if(!is.null(omlist_ss$truecatchage_ss)){
    interp_fishWtAtAge <- list()
    for(i in 1:n_reps){
      interp_fishWtAtAge[[i]] <- calc_avgwtstage2age(wtagecl = fishObsWtAtAge[[i]],
                                                     annages = omlist_ss$truecatchage_ss,
                                                     fgs = omlist_ss$funct.group_ss)
    }
    if(save){
      saveRDS(interp_fishWtAtAge, file.path(d.name, paste0(scenario.name,"_",
                                                           fishery.name, "fishObsFullWtAtAge.rds")))
    }
  }else{interp_fishWtAtAge <- NULL}

  comps <- list("survObsAgeComp" = age_comp_datas,
                "survObsLenComp" = survObsLenComps,
                "survObsWtAtAge" = survObsWtAtAges,
                "fishObsAgeComp" = catch_age_comp,
                "fishObsLenComp" = fishObsLenComp,
                "fishObsWtAtAge" = fishObsWtAtAge,
                "survObsFullAgeComp" = annage_comp_datas,
                "fishObsFullAgeComp" = catch_annage_comp,
                "survObsFullWtAtAge" = interp_survWtAtAges,
                "fishObsFullWtAtAge" = interp_fishWtAtAge
                )

  return(comps)
}

```

### Wrapper test results {.tabset}

#### Biomass index
```{r bioind-1}

source(here("config/CCConfigv4.R"))

omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))

source(here("config/omdimensions.R"))
source(here("config/sardinesurvey.R"))
source(here("config/sardinefishery.R"))

#read time series data
survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name,
                                                "surveyB.rds")))

survObsBiom <- survObsBiom[[1]]

plotB <-ggplot() +
  geom_line(data=survObsBiom, aes(x=time/stepperyr,y=atoutput, color="survey Biomass"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotB +
  facet_wrap(~species, scales="free")

```

#### Catch time series
```{r catchind-1}
#read time series data
catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                fishery.name, "fishCatch.rds")))

catchbio_ss <- catchbio_ss[[1]]

plotC <-ggplot() +
  geom_line(data=catchbio_ss, aes(x=time/365,y=atoutput, color="observed Catch"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotC +
  facet_wrap(~species, scales="free")

```

#### Survey length composition
```{r lencomp-1}

#length comps
len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name, "survObsLenComp.rds")))
fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                fishery.name,"fishObsLenComp.rds")))

len_comp_data <- len_comp_data[[1]]
fish_len_comp_data <- fish_len_comp_data[[1]]

#add this to om_indices function so that this has years when read in
fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))

len <- len_comp_data

  lfplot <- ggplot(len, aes(upper.bins)) +
    geom_bar(aes(weight = atoutput)) +
    theme_tufte() +
    labs(subtitle = paste(scenario.name,
                          len$species))

  lfplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")

```

#### Survey age composition (age classes)
```{r agecomp-1}

#read in comp data
age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name, "survObsAgeComp.rds")))
age_comp_data <- age_comp_data[[1]]

Natage <- age_comp_data

Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
    geom_point() +
    theme_tufte() +
    labs(subtitle = paste(scenario.name,
                          Natage$species))

  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
```

#### Survey weight at age (age classes)
```{r wtageclass-1}

wtage <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name, "survObsWtAtAge.rds")))
wtage <- wtage[[1]]

wageplot <- ggplot(wtage, aes(time, atoutput)) +
  geom_line(aes(colour = factor(agecl))) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  xlab("model timestep (5 per year)") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste0(scenario.name, " annual mid year sample"))

wageplot + facet_wrap(c("species"), scales="free_y")
```

#### Fishery age composition (age classes)
```{r fishagecomp-1}

#read in comp data
fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                fishery.name, "fishObsAgeComp.rds")))
fish_age_comp <- fish_age_comp[[1]]

#add this to om_indices function so that this has years when read in
fish_age_comp$time <- fish_age_comp$time/fstepperyr

Natage <- fish_age_comp

Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
    geom_point() +
    theme_tufte() +
    labs(subtitle = paste(scenario.name,
                          Natage$species))

  Natageplot + facet_wrap(~time, ncol=6, scales="free_y")
```

### {-}

### 2022-06-30 config/CCConfigv4.R

@@ -1,0 +1,12 @@
+d.name <- here("atlantisoutput","AtlantisCalCurrV4_044")
+functional.groups.file <- "CalCurrentV4Groups_noFisheries.csv"
+biomass.pools.file <- "CalCurrentV4_Biol3Feb2022.nc"
+biol.prm.file <- "CalCurrentV4_Biol3Feb2022.prm"
+box.file <- "CalCurrentV3_utm.bgm"
+initial.conditions.file <- "CalCurrentV4_Biol3Feb2022.nc"
+run.prm.file <- "outputFolder/CalCurrentV4_run_test_withFisheries.xml"
+scenario.name <- "outputFolder/outputCCV4_test"
+bioind.file <- "outputFolder/outputCCV4_testBiomIndx.txt"
+catch.file <- "outputFolder/outputCCV4_testCatch.txt"
+annage <- FALSE
+fisheries.file <- "CalCurrentV4Fisheries.csv"

### 2022-06-30 config/sardinefishery.R

@@ -1,4 +1,6 @@
 # Default fishery configuration here is a census
+fishery.name="sardine_census"
+
 # Inherits species from input omlist_ss
 fishspp <- omlist_ss$species_ss 
 
@@ -6,7 +8,7 @@ fishspp <- omlist_ss$species_ss
 nyears <- 50
 
 #Atlantis initialization period in years
-burnin <- 30
+burnin <- 0
 
 
 # same time dimensioning parameters as in surveycensus.R

### 2022-06-30 config/sardinesurvey.R

@@ -1,6 +1,9 @@
 # Default survey configuration here is a census
 # Need to define survey season, area, efficiency, selectivity
 
+# Survey name
+survey.name="sardine_midyear_allbox_effic.5"
+
 #Atlantis model timestep corresponding to the true output--now from census_spec.R
 timestep <- stepperyr #5
 
@@ -28,11 +31,12 @@ survboxes <- allboxes
 surveffic <- data.frame(species=survspp,
                      efficiency=rep(0.5,length(survspp)))
 
-# survey selectivity (agecl based)
-# this is by age class, need to change to use with ANNAGEBIO output
-survselex <- data.frame(species=rep(survspp, each=n_age_classes),
-                     agecl=rep(c(1:n_age_classes),length(survspp)),
-                     selex=rep(1.0,length(survspp)*n_age_classes))
+# for annage output
+survselex <- data.frame(species=rep(names(annages), n_annages), #  
+                        agecl=unlist(sapply(n_annages,seq)),
+                        selex=rep(1.0,sum(n_annages)))
+
+survselex.agecl <- survselex 
 
 # effective sample size needed for sample_fish
 # this effective N is high but not equal to total for numerous groups

### 2022-09-16 NewCCASardines.Rmd
update CCA results WIP

@@
 ---
-title: "New CCA sardines"
+title: "New CCA sardines, Sept 2022"
 author: "Sarah Gaichas"
@@
 
-A test run for sardines with the updated California Current model, run 44:
+A test run for sardines with the updated California Current model, September 2022:
 
@@
 
-CCom <- atlantisom::om_init(here("config/CCConfigv4.R"))
+CCom <- atlantisom::om_init(here("config/CCConfigSep22.R"))
 
@@
 
-source(here("config/CCConfigv4.R"))
+source(here("config/CCConfigSep22.R"))
 
@@
 
+len <- filter(len_comp_data, time %in% c(55:175))
+
   lfplot <- ggplot(len, aes(upper.bins)) +
@@
 Natage <- age_comp_data
+
+Natage <- filter(age_comp_data, time %in% c(150:270))
 
@@
 
+Natage <- filter(fish_age_comp, time %in% c(30:53))
+
 Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
@@
 ### {-}
+
+## Write data for Stock Synthesis
+
+To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`.
+You will need to ensure the relevant Stock Synthesis files are stored in the inst/extdata/ folder, with each species example in its own folder.
+
+```{r readdata}
+#devtools::install_github("r4ss/r4ss", dependencies = FALSE)
+require(r4ss)
+#source("./config/sardine_config.R") this is above for illustration
+
+#Directory with SS files
+model_dir <- "Sardine_SS_files/"
+
+#Name of SS data file
+datfile_name <- "sardEM_3_3.dat"
+
+stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/",
+model_dir,
+datfile_name))
+```
+
+*Now changing catch units to tons!*
+
+Write time series biomass (tons) and catch (tons--taken directly from truth without error) data for SS using `SS_write_ts`:
+
+```{r SS-write-ts}
+
+#Month of survey/fishing
+survey_month <- 7
+fishing_month <- 1
+
+#Test writing CPUE in biomass
+  stocksynthesis.data <- SS_write_ts(ss_data_list = stocksynthesis.data,
+                ts_data = list(survObsBiom$atoutput[fish_years],
+                  omlist_ss$truecatchbio_ss$atoutput[omlist_ss$truecatchbio_ss$time %in% fish_years]),
+                CVs = c(CVs$survey,
+                        CVs$fishery),
+                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             fish_years),
+            sampling_month = list(rep(survey_month,nyears),
+                                rep(fishing_month,nyears)),
+                units = c("biomass","biomass"),
+                data_type=c("CPUE","catch"),
+                fleets = c(2,1))
+
+stocksynthesis.data$CPUE
+
+stocksynthesis.data$catch
+```
+
+Writing age composition data
+
+```{r SS-write-comps}
+#age_comp_data
+
+#Get the age bins
+age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
+age_bins <- sub("a","",age_bin_names)
+
+age_comp_flat <- reformat_compositions(age_comp_data,
+                     round.places = 4,
+                     comp_type = "agecomp")
+
+## Write age composition data for survey
+stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data,
+               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
+               data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr+nyears-1)),
+               sampling_month = list(rep(survey_month,nyears)),
+               data_type = c("agecomp"),
+               fleet_number = c(1),
+               bins = list(age_bins),
+               caal_bool = c(FALSE))
+stocksynthesis.data$agecomp
+
+```
+
+Once we have the natlength matrix, we can munge the data into the proper CAAL and length bin format for SS
+
+```{r SS-comps-formatting}
+
+#Check length compositions match age compositions
+#ss_length_stdsurv$natlength %>%
+#  group_by(time,agecl) %>%
+#  summarise(sum(atoutput))
+
+len_comp_data <- ss_length_stdsurv$natlength
+fish_len_comp_data <- catch_lengthwt_samp$natlength
+
+if(fstepperyr>1){
+  fish_len_comp_anndata <- fish_len_comp_data %>%
+    mutate(yr = floor(time/fstepperyr)) %>%
+    group_by(species, agecl, lower.bins, upper.bins, time=as.integer(yr)) %>%
+    summarise(annnatlength=sum(atoutput)) %>%
+    rename(atoutput = annnatlength)
+} else {
+  fish_len_comp_anndata <- fish_len_comp_data
+}
+
+caal_comp_flat <- reformat_compositions(len_comp_data,                                round.places=4,
+                comp_type="caalcomp")
+
+#remove burnin
+caal_comp_final <- filter(caal_comp_flat,
+                         time %in% survey_years)
+
+#Add over age classes to get sample size
+len_comp_flat <- reformat_compositions(len_comp_data,
+                                  round.places = 0,
+                           comp_type="lencomp")
+#remove burnin
+len_comp_final <- filter(len_comp_flat,
+                         time %in% survey_years)
+
+length_bins <- as.integer(names(len_comp_final))
+length_bins <- length_bins[!is.na(length_bins)]
+
+# fishery length comps are still 5 timesteps per year
+# need to aggregate to annual (done above)
+# also,  make effN annual goal/fstepsperyr (done above)
+fish_len_comp_flat <- reformat_compositions(fish_len_comp_anndata,
+                                  round.places = 0,
+                           comp_type="lencomp")
+
+#remove burnin works after adjustment above
+fish_len_comp_final <- filter(fish_len_comp_flat,
+                         time %in% fish_years)
+
+notbins <- c("time", "nsamp")
+
+# fish_length_bins <- as.integer(names(fish_len_comp_final))
+# fish_length_bins <- fish_length_bins[!is.na(fish_length_bins)]
+
+# need to fill empty length bins with 0s to have same bins as survey for SS_write_comps
+missing.lengths <- setdiff(length_bins, names(fish_len_comp_final)[!names(fish_len_comp_final) %in% notbins])
+fish_len_comp_final[as.character(missing.lengths)] <- 0                    # Add them, filled with '0's
+fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")]
+
+# fishery age comps also 5 timesteps per year
+if(fstepperyr>1){
+  fish_age_comp_anndata <- catch_numsss_samp %>%
+    mutate(yr = floor(time/fstepperyr)) %>%
+    group_by(species, agecl, time=as.integer(yr)) %>%
+    summarise(annnatage=sum(atoutput)) %>%
+    rename(atoutput = annnatage)
+} else {
+  fish_age_comp_anndata <- catch_numsss_samp
+}
+
+fish_age_comp_flat <- reformat_compositions(fish_age_comp_anndata,
+                           comp_type="agecomp")
+
+#remove burnin (not necessary?fish comps made with fish_years only) 
+fish_age_comp_final <- filter(fish_age_comp_flat,
+                         time %in% fish_years)
+
+# #SS_write_comps breaking because fishery age bins start with 2 not 1; extracting bins from fish file may help?
+# fish_age_bins <- names(fish_age_comp_flat)[!names(fish_age_comp_flat) %in% notbins]
+
+# that leaves an empty column in data file, so instead fill with 0s
+missing.ages <- setdiff(age_bins, names(fish_age_comp_final)[!names(fish_age_comp_final) %in% notbins])
+fish_age_comp_final[missing.ages] <- 0                    # Add them, filled with '0's
+fish_age_comp_final <- fish_age_comp_final[c("time", age_bins, "nsamp")]
+
+comp_list <- list(caal_comp_final,len_comp_final, fish_age_comp_final, fish_len_comp_final)
+
+apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), 
+                    rep(survey_month, nrow(comp_list[[2]])),
+                    rep(fishing_month,nrow(comp_list[[3]])),
+                    rep(fishing_month,nrow(comp_list[[4]])))
+
+# This now runs by ensuring that survey and fishery compositions have the same bins 
+# (filled with 0s for missing bins in fishery relative to survey)
+
+# Write CAAL and length composition data
+stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data, 
+                                      comp_matrix = comp_list, 
+                                      data_rows = list((comp_list[[1]]$time-survey_sample_time)/timestep + 1 , (survey_years-survey_sample_time)/timestep + 1,fish_years,fish_years),
+                                      sampling_month = apply_month, 
+                                      data_type = rep(c("agecomp", "lencomp"),2), 
+                                      fleet_number = c(2,2,1,1),  
+                                      bins = list(age_bins, 
+                                                  length_bins, 
+                                                  age_bins, 
+                                                  length_bins), 
+                                      caal_bool = c(TRUE, rep(FALSE,3)))
+
+head(stocksynthesis.data$lencomp)
+head(stocksynthesis.data$agecomp)
+
+#Change length bin structure to match atlantis data
+stocksynthesis.data$lbin_vector <- length_bins
+
+#Get correct number of length bins
+stocksynthesis.data$N_lbins <- length(length_bins)
+
+#Set lbin_method to 1 - this makes the population length bins match the data bins 
+#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
+stocksynthesis.data$lbin_method <- 1
+stocksynthesis.data$binwidth <- "#"
+stocksynthesis.data$minimum_size <- "#"
+stocksynthesis.data$maximum_size <- "#"
+
+#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
+stocksynthesis.data$age_info$minsamplesize <- rep(0.001,2)
+
+SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
+datfile_name),
+                 overwrite=TRUE)
+```
+
+## Coming up with dynamic life history parameters (partial from Christine)
+
+Below we get the weight-length relationship parameters from atlantis, calculate h and R0 from Atlantis $\alpha$ and $\beta$ parameters, use the \code{calc_Z} function to back out natural mortality (M) from Atlantis, and use the survey length-at-age from Atlantis to estimate a growth curve.
+
+The following is under construction from CreateStockSynthesis.Rmd and not evaluated.
+Existing control file was modified by hand to make the number of recruitment settlement assignments 1 instead of two, with GPattern month area age vector 1 1 1 1.
+
+```{r, eval=FALSE}
+#Get biological parameters
+
+#Load needed inputs for biological parameters
+source(here("config/CC3Config.R"))
+#biological parameters
+biolprm <- load_biolprm(dir = d.name, file_biolprm = biol.prm.file)#"C:/Users/chris/Documents/GitHub/atlantisom", 
+#d.name <- "C:/Users/chris/Documents/GitHub/atlantisom"
+#functional groups
+fgs <- truth$fgs
+runprm <- load_runprm(dir = d.name, file_runprm = run.prm.file)
+YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))
+truenums_ss <- truth$nums[results$nums$species==species,]
+
+YOY_ss <- YOY %>%
+  select(Time, "SAR.0")
+
+fullresZ <- calc_Z(yoy = YOY_ss,
+                   nums = truenums_ss,
+                   fgs = fgs,
+                   biolprm = biolprm,
+                   toutinc = runprm$toutinc)
+
+meanwt_spp <- ss_length_stdsurv$muweight %>% 
+  filter(time>burnin) %>%
+  group_by(agecl) %>%
+  summarize(meanwt = mean(atoutput))
+
+ctlfile_name <-  "sardEM_3_3.ctl"
+sardine.ctl <-r4ss::SS_readctl_3.30(paste0("./inst/extdata/",
+model_dir,
+ctlfile_name))
+sardine.ctl <- SS_write_biol(sardine.ctl, biolprm, "SAR", Z = mean(fullresZ$atoutput), wtsage=meanwt_spp)
+
+li_a_use <- biolprm$wl[match(fgs$Code[match(species,fgs$Name)],biolprm$wl[, 1]), 2]/1000
+li_b_use <- biolprm$wl[match(fgs$Code[match(species,fgs$Name)],biolprm$wl[, 1]), 3]
+
+#maturity ogive
+#check what R0 is
+
+len_nonburn <- ss_length_stdsurv$mulen %>%
+  filter(time > burnin*timestep)
+plot(len_nonburn$atoutput~len_nonburn$agecl, col=len_nonburn$time)
+
+length.data <- data.frame("Year"=(len_nonburn$time-survey_sample_time)/timestep+1, length=len_nonburn$atoutput, Weight=NA, Sex="Female", age=as.integer(len_nonburn$agecl))
+
+#require(bbmle)
+#vb_estimates <-sample_fit_vbgf(length.data, 25, 45, 0.4,
+#  0.1, 0.1, 20, 40, 0.05, 0.01, 0.01,
+#  30, 50, 0.6, 0.3, 0.3, 0.5, 10)
+
+```
+
+## Run Stock Synthesis
+After the data and control file modifications, ensure you copy the local SS 3.3 executable into the folder containing the model files. Then run the following chunk.
+
+```{r runSS}
+#Run stock synthesis model 
+# default runs with admb_options = "-noest"
+#run_stocksynthesis(species = species, model_dir = model_dir, show_output = TRUE)
+
+# for estimation
+run_stocksynthesis(species = species, model_dir = model_dir, admb_options = "", show_output = TRUE)
+
+```

### 2022-09-16 config/CCConfigSep22.R

@@ -1,0 +1,12 @@
+d.name <- here("atlantisoutput","CC2063_Sep2022")
+functional.groups.file <- "CalCurrentV3Groups.csv"
+biomass.pools.file <- "DIVCalCurrentV3_Biol.nc"
+biol.prm.file <- "CalCurrentV3_Biol.prm"
+box.file <- "CalCurrentV3_utm.bgm"
+initial.conditions.file <- "DIVCalCurrentV3_Biol.nc"
+run.prm.file <- "CalCurrentV3_run.xml"
+scenario.name <- "outputCCV3"
+bioind.file <- "outputCCV3BiomIndx.txt"
+catch.file <- "outputCCV3Catch.txt"
+annage <- FALSE
+fisheries.file <- "CalCurrentV3Fisheries.csv"

### 2022-10-28 NewCCASardines.Rmd

--- 2022-06-30 NewCCASardines.Rmd
@@
 library(FSA)
-
 ```
 
-A test run for sardines with the updated California Current model, run 44:
+A test run for sardines with the updated California Current model, September 2022:
 
 ```{r full-test, message=FALSE, warning=FALSE, eval=FALSE}
-
-CCom <- atlantisom::om_init(here("config/CCConfigv4.R"))
+CCom <- atlantisom::om_init(here("config/CCConfigSep22.R"))
 
@@
 
-CCom_sard_ind <- atlantisom::om_index(usersurvey = here("config/sardinesurvey.R"), 
+CCom_sard_ind <- atlantisom::om_index(
+  usersurvey = here("config/sardinesurvey.R"),
   userfishery = here("config/sardinefishery.R"),
@@
   n_reps = 1,
-                           save = TRUE)
+  save = TRUE
+)
 
-CCom_sard_comp <- atlantisom::om_comps(usersurvey = here("config/sardinesurvey.R"), 
+CCom_sard_comp <- atlantisom::om_comps(
+  usersurvey = here("config/sardinesurvey.R"),
   userfishery = here("config/sardinefishery.R"),
@@
   n_reps = 1,
-                           save = TRUE)
-
+  save = TRUE
+)
 ```
@@
 
-usersurvey = here("config/sardinesurvey.R")
+usersurvey <- here("config/sardinesurvey.R")
 
@@
 
-#this is om_index
+# this is om_index
 
-#one script for dimension parameters to be used in multiple functions
+# one script for dimension parameters to be used in multiple functions
 source("config/omdimensions.R", local = TRUE)
+
+###############################################
+n_reps <- 1 # not sure about it, used the value from line 31
+save <- FALSE # Sarah already provided output data
+###############################################
 
@@
 
-    #biomass based fishery independent survey index
+  # biomass based fishery independent survey index
   # this uses result$biomass_ages to sample biomass directly, no need for wt@age est
-    survey_B <- atlantisom::create_survey(dat = omlist_ss$truebio_ss,
+  survey_B <- atlantisom::create_survey(
+    dat = omlist_ss$truebio_ss,
     time = survtime,
@@
     effic = surveffic,
-                                          selex = survselex)
+    selex = survselex
+  )
 
@@
   # biomass directly
-    wtage <- data.frame(species=rep(names(age_classes), n_age_classes),
-                        agecl=unlist(sapply(n_age_classes,seq)),
-                        wtAtAge=rep(1000.0,sum(n_age_classes)))
+  wtage <- data.frame(
+    species = rep(names(age_classes), n_age_classes),
+    agecl = unlist(sapply(n_age_classes, seq)),
+    wtAtAge = rep(1000.0, sum(n_age_classes))
+  )
 
@@
   survObsBiomB <- list()
-    for(i in 1:n_reps){
+  for (i in 1:n_reps) {
     survObsBiomB[[i]] <- atlantisom::sample_survey_biomass(survey_B, surv_cv, wtage)
@@
 
-    #save survey indices, takes a long time to generate with lots of reps/species
-    if(save){
-      saveRDS(survObsBiomB, file.path(d.name, paste0(scenario.name, "_",
-                                                     survey.name, "surveyB.rds")))
+  # save survey indices, takes a long time to generate with lots of reps/species
+  if (save) {
+    saveRDS(survObsBiomB, file.path(d.name, paste0(
+      scenario.name, "_",
+      survey.name, "surveyB.rds"
+    )))
   }
@@
 
-  #configure the fishery, a default is in config/fisherycensus.R
-  #fishery configuration can specify only area and time of observation
-  #fishery species inherited from omlist_ss
-  #this is total catch not by fleet, so only one "fishery"
+# configure the fishery, a default is in config/fisherycensus.R
+# fishery configuration can specify only area and time of observation
+# fishery species inherited from omlist_ss
+# this is total catch not by fleet, so only one "fishery"
+
+##################################################
+# Use information from lines 29 and 35
+userfishery <- here("config/sardinefishery.R")
+##################################################
 source(userfishery, local = TRUE)
 
-  #we are not currently subsetting fishery catch because we cannot correct catch.nc
+# we are not currently subsetting fishery catch because we cannot correct catch.nc
 #  instead the catch in biomass from catch.txt is read in for the index
@@
 fishObsCatchB <- list()
-  for(i in 1:n_reps){
+for (i in 1:n_reps) {
   fishObsCatchB[[i]] <- atlantisom::sample_fishery_totcatch(omlist_ss$truecatchbio_ss, fish_cv)
@@
 
-  if(save){
-    saveRDS(fishObsCatchB, file.path(d.name, paste0(scenario.name, "_",
-                                                    fishery.name, "fishCatch.rds")))
+if (save) {
+  saveRDS(fishObsCatchB, file.path(d.name, paste0(
+    scenario.name, "_",
+    fishery.name, "fishCatch.rds"
+  )))
 }
 
-  indices <- list("survObsBiomB" = survObsBiomBs,
-                  "fishObsCatchB" = fishObsCatchB)
+indices <- list(
+  "survObsBiomB" = survObsBiomBs,
+  "fishObsCatchB" = fishObsCatchB
+)
 
-  return(indices)
-}
-
+#  return(indices)
+# }
 ```
@@
 
-#this is om_comps
+# this is om_comps
 
- #one script for dimension parameters to be used in multiple functions
+# one script for dimension parameters to be used in multiple functions
 source("config/omdimensions.R", local = TRUE)
@@
 
-    #numbers based fishery independent survey for age and length comps
+  # numbers based fishery independent survey for age and length comps
   # same user specifications as indices
-    survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
+  survey_N <- atlantisom::create_survey(
+    dat = omlist_ss$truenums_ss,
     time = survtime,
@@
     effic = surveffic,
-                                          selex = survselex.agecl)
+    selex = survselex.agecl
+  )
 
-    #Sample fish for age composition
+  # Sample fish for age composition
   # if we want replicates for obs error this sample function will generate them
   age_comp_data <- list()
-    for(i in 1:n_reps){
+  for (i in 1:n_reps) {
     age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
@@
   # save age comps
-    if(save){
-      saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "_",
-                                                      survey.name, "survObsAgeComp.rds")))
+  if (save) {
+    saveRDS(age_comp_data, file.path(d.name, paste0(
+      scenario.name, "_",
+      survey.name, "survObsAgeComp.rds"
+    )))
   }
 
-    #weights needed for weight at age and length comp calcs
+  # weights needed for weight at age and length comp calcs
   # aggregate true resn per survey design
-    survey_aggresn <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
+  survey_aggresn <- atlantisom::aggregateDensityData(
+    dat = omlist_ss$trueresn_ss,
     time = survtime,
     species = survspp,
-                                                       boxes = survboxes)
+    boxes = survboxes
+  )
 
   # aggregate true structn per survey design
-    survey_aggstructn <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
+  survey_aggstructn <- atlantisom::aggregateDensityData(
+    dat = omlist_ss$truestructn_ss,
     time = survtime,
     species = survspp,
-                                                          boxes = survboxes)
+    boxes = survboxes
+  )
 
-    #dont sample these, just aggregate them using median
+  # dont sample these, just aggregate them using median
   structnss <- atlantisom::sample_fish(survey_aggstructn, surveffN, sample = FALSE)
@@
 
-    #this is all input into the length function, replicates follow age comp reps
+  # this is all input into the length function, replicates follow age comp reps
   #  separating the length comps from the weight at age here
@@
 
-    for(i in 1:n_reps){
-      survey_lenwt[[i]] <- atlantisom::calc_age2length(structn = structnss,
+  for (i in 1:n_reps) {
+    survey_lenwt[[i]] <- atlantisom::calc_age2length(
+      structn = structnss,
       resn = resnss,
@@
       CVlenage = lenage_cv,
-                                                       remove.zeroes=TRUE)
+      remove.zeroes = TRUE
+    )
 
@@
 
-    if(save){
-      saveRDS(survObsLenComp, file.path(d.name, paste0(scenario.name, "_",
-                                                       survey.name, "survObsLenComp.rds")))
-      saveRDS(survObsWtAtAge, file.path(d.name, paste0(scenario.name, "_",
-                                                       survey.name, "survObsWtAtAge.rds")))
+  if (save) {
+    saveRDS(survObsLenComp, file.path(d.name, paste0(
+      scenario.name, "_",
+      survey.name, "survObsLenComp.rds"
+    )))
+    saveRDS(survObsWtAtAge, file.path(d.name, paste0(
+      scenario.name, "_",
+      survey.name, "survObsWtAtAge.rds"
+    )))
   }
@@
   survObsWtAtAges[[survey.name]] <- survObsWtAtAge
-
 }
 
-  #now do fishery comps
+# now do fishery comps
 # user options for fishery--default is a census with mid-year sample
@@
 
-  #fishery catch at age each observed timestep summed over observed polygons
+# fishery catch at age each observed timestep summed over observed polygons
 # catch at age by area and timestep
-  catch_numbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchnum_ss,
+catch_numbers <- atlantisom::create_fishery_subset(
+  dat = omlist_ss$truecatchnum_ss,
   time = fishtime,
   species = survspp,
-                                                      boxes = fishboxes)
+  boxes = fishboxes
+)
 
@@
 catch_age_comp <- list()
-  for(i in 1:n_reps){
+for (i in 1:n_reps) {
   catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
@@
 # save fishery age comps
-  if(save){
-    saveRDS(catch_age_comp, file.path(d.name, paste0(scenario.name, "_",
-                                                     fishery.name, "fishObsAgeComp.rds")))
+if (save) {
+  saveRDS(catch_age_comp, file.path(d.name, paste0(
+    scenario.name, "_",
+    fishery.name, "fishObsAgeComp.rds"
+  )))
 }
 
-  #Get catch weights for length comp calc
+# Get catch weights for length comp calc
 # aggregate true resn per fishery subset design
-  catch_aggresnss <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
+catch_aggresnss <- atlantisom::aggregateDensityData(
+  dat = omlist_ss$trueresn_ss,
   time = fishtime,
   species = survspp,
-                                                      boxes = fishboxes)
+  boxes = fishboxes
+)
 
 # aggregate true structn fishery subsetdesign
-  catch_aggstructnss <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
+catch_aggstructnss <- atlantisom::aggregateDensityData(
+  dat = omlist_ss$truestructn_ss,
   time = fishtime,
   species = survspp,
-                                                         boxes = fishboxes)
+  boxes = fishboxes
+)
 
-  #dont sample these, just aggregate them using median
+# dont sample these, just aggregate them using median
 catch_structnss <- atlantisom::sample_fish(catch_aggstructnss, fisheffN, sample = FALSE)
@@
 # these fishery lengths and weight at age are each output timestep
-  #same structure as above for surveys, replicates follow age comp reps
+# same structure as above for surveys, replicates follow age comp reps
 #  separating the length comps from the weight at age here
@@
 
-  for(i in 1:n_reps){
-    fishery_lenwt[[i]] <- atlantisom::calc_age2length(structn = catch_structnss,
+for (i in 1:n_reps) {
+  fishery_lenwt[[i]] <- atlantisom::calc_age2length(
+    structn = catch_structnss,
     resn = catch_resnss,
@@
     CVlenage = lenage_cv,
-                                                      remove.zeroes=TRUE)
+    remove.zeroes = TRUE
+  )
 
@@
 
-  if(save){
-    saveRDS(fishObsLenComp, file.path(d.name, paste0(scenario.name, "_",
-                                                     fishery.name, "fishObsLenComp.rds")))
-    saveRDS(fishObsWtAtAge, file.path(d.name, paste0(scenario.name, "_",
-                                                     fishery.name, "fishObsWtAtAge.rds")))
+if (save) {
+  saveRDS(fishObsLenComp, file.path(d.name, paste0(
+    scenario.name, "_",
+    fishery.name, "fishObsLenComp.rds"
+  )))
+  saveRDS(fishObsWtAtAge, file.path(d.name, paste0(
+    scenario.name, "_",
+    fishery.name, "fishObsWtAtAge.rds"
+  )))
 }
 
-  if(!is.null(omlist_ss$truenumsage_ss)){
-    #numbers based fishery independent survey for age and length comps
-    #allows for mulitple surveys
+if (!is.null(omlist_ss$truenumsage_ss)) {
+  # numbers based fishery independent survey for age and length comps
+  # allows for mulitple surveys
   annage_comp_datas <- list()
@@
     # same user specifications as indices
-      survey_annageN <- atlantisom::create_survey(dat = omlist_ss$truenumsage_ss,
+    survey_annageN <- atlantisom::create_survey(
+      dat = omlist_ss$truenumsage_ss,
       time = survtime,
@@
       effic = surveffic,
-                                                  selex = survselex)
-      #Sample fish for age composition
+      selex = survselex
+    )
+    # Sample fish for age composition
     # if we want replicates for obs error this sample function will generate them
     annage_comp_data <- list()
-      for(i in 1:n_reps){
+    for (i in 1:n_reps) {
       annage_comp_data[[i]] <- atlantisom::sample_fish(survey_annageN, surveffN)
@@
     # save survey annual age comps
-      if(save){
-        saveRDS(annage_comp_data, file.path(d.name, paste0(scenario.name, "_",
-                                                           survey.name, "survObsFullAgeComp.rds")))
+    if (save) {
+      saveRDS(annage_comp_data, file.path(d.name, paste0(
+        scenario.name, "_",
+        survey.name, "survObsFullAgeComp.rds"
+      )))
     }
@@
   }
+} else {
+  annage_comp_datas <- NULL
+}
 
-  }else{annage_comp_datas <- NULL}
-
-  if(!is.null(omlist_ss$truecatchage_ss)){
-    #fishery catch at age each observed timestep summed over observed polygons
+if (!is.null(omlist_ss$truecatchage_ss)) {
+  # fishery catch at age each observed timestep summed over observed polygons
   # catch at age by area and timestep
-    catch_annagenumbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchage_ss,
+  catch_annagenumbers <- atlantisom::create_fishery_subset(
+    dat = omlist_ss$truecatchage_ss,
     time = fishtime,
     species = survspp,
-                                                        boxes = fishboxes)
+    boxes = fishboxes
+  )
 
@@
   catch_annage_comp <- list()
-    for(i in 1:n_reps){
+  for (i in 1:n_reps) {
     catch_annage_comp[[i]] <- atlantisom::sample_fish(catch_annagenumbers, fisheffN)
@@
   # save fishery annual age comps
-    if(save){
-    saveRDS(catch_annage_comp, file.path(d.name, paste0(scenario.name,"_",
-                                                        fishery.name, "fishObsFullAgeComp.rds")))
+  if (save) {
+    saveRDS(catch_annage_comp, file.path(d.name, paste0(
+      scenario.name, "_",
+      fishery.name, "fishObsFullAgeComp.rds"
+    )))
   }
-  }else{catch_annage_comp <- NULL}
+} else {
+  catch_annage_comp <- NULL
+}
 
 # call interpolate weight at age function to get survObsFullWtAtAge
-  if(!is.null(omlist_ss$truenumsage_ss)){
+if (!is.null(omlist_ss$truenumsage_ss)) {
   interp_survWtAtAges <- list()
@@
     interp_survWtAtAge <- list()
-      for(i in 1:n_reps){
-        interp_survWtAtAge[[i]] <- calc_avgwtstage2age(wtagecl = survObsWtAtAges[[survey.name]][[i]],
+    for (i in 1:n_reps) {
+      interp_survWtAtAge[[i]] <- calc_avgwtstage2age(
+        wtagecl = survObsWtAtAges[[survey.name]][[i]],
         annages = omlist_ss$truenumsage_ss,
-                                                       fgs = omlist_ss$funct.group_ss)
+        fgs = omlist_ss$funct.group_ss
+      )
     }
-      if(save){
-        saveRDS(interp_survWtAtAge, file.path(d.name, paste0(scenario.name, "_",
-                                                             survey.name, "survObsFullWtAtAge.rds")))
+    if (save) {
+      saveRDS(interp_survWtAtAge, file.path(d.name, paste0(
+        scenario.name, "_",
+        survey.name, "survObsFullWtAtAge.rds"
+      )))
     }
@@
   }
-  }else{interp_survWtAtAges <- NULL}
+} else {
+  interp_survWtAtAges <- NULL
+}
 
@@
 # WARNING currently aggregates out fleet info, but no fleets in aggregate wtage
-  if(!is.null(omlist_ss$truecatchage_ss)){
+if (!is.null(omlist_ss$truecatchage_ss)) {
   interp_fishWtAtAge <- list()
-    for(i in 1:n_reps){
-      interp_fishWtAtAge[[i]] <- calc_avgwtstage2age(wtagecl = fishObsWtAtAge[[i]],
+  for (i in 1:n_reps) {
+    interp_fishWtAtAge[[i]] <- calc_avgwtstage2age(
+      wtagecl = fishObsWtAtAge[[i]],
       annages = omlist_ss$truecatchage_ss,
-                                                     fgs = omlist_ss$funct.group_ss)
+      fgs = omlist_ss$funct.group_ss
+    )
   }
-    if(save){
-      saveRDS(interp_fishWtAtAge, file.path(d.name, paste0(scenario.name,"_",
-                                                           fishery.name, "fishObsFullWtAtAge.rds")))
+  if (save) {
+    saveRDS(interp_fishWtAtAge, file.path(d.name, paste0(
+      scenario.name, "_",
+      fishery.name, "fishObsFullWtAtAge.rds"
+    )))
   }
-  }else{interp_fishWtAtAge <- NULL}
+} else {
+  interp_fishWtAtAge <- NULL
+}
 
-  comps <- list("survObsAgeComp" = age_comp_datas,
+comps <- list(
+  "survObsAgeComp" = age_comp_datas,
   "survObsLenComp" = survObsLenComps,
@@
 
-  return(comps)
-}
-
+#  return(comps)
+# }
 ```
@@
 
-#read time series data
-survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name,
+# read time series data
+survObsBiom <- readRDS(file.path(d.name, paste0(
+  scenario.name,
   "_",
   survey.name,
-                                                "surveyB.rds")))
+  "surveyB.rds"
+)))
 
@@
 
-plotB <-ggplot() +
-  geom_line(data=survObsBiom, aes(x=time/stepperyr,y=atoutput, color="survey Biomass"), 
-            alpha = 10/10) +
+plotB <- ggplot() +
+  geom_line(
+    data = survObsBiom, aes(x = time / stepperyr, y = atoutput, color = "survey Biomass"),
+    alpha = 10 / 10
+  ) +
   theme_tufte() +
   theme(legend.position = "top") +
-  labs(colour=scenario.name)
+  labs(colour = scenario.name)
 
 plotB +
-  facet_wrap(~species, scales="free")
-
+  facet_wrap(~species, scales = "free")
 ```
@@
 ```{r catchind-1}
-#read time series data
-catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name,
+# read time series data
+catchbio_ss <- readRDS(file.path(d.name, paste0(
+  scenario.name,
   "_",
-                                                fishery.name, "fishCatch.rds")))
+  fishery.name, "fishCatch.rds"
+)))
 
@@
 
-plotC <-ggplot() +
-  geom_line(data=catchbio_ss, aes(x=time/365,y=atoutput, color="observed Catch"), 
-            alpha = 10/10) +
+plotC <- ggplot() +
+  geom_line(
+    data = catchbio_ss, aes(x = time / 365, y = atoutput, color = "observed Catch"),
+    alpha = 10 / 10
+  ) +
   theme_tufte() +
   theme(legend.position = "top") +
-  labs(colour=scenario.name)
+  labs(colour = scenario.name)
 
 plotC +
-  facet_wrap(~species, scales="free")
-
+  facet_wrap(~species, scales = "free")
 ```
@@
 
-#length comps
-len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
+# length comps
+len_comp_data <- readRDS(file.path(d.name, paste0(
+  scenario.name,
   "_",
-                                                survey.name, "survObsLenComp.rds")))
-fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
+  survey.name, "survObsLenComp.rds"
+)))
+fish_len_comp_data <- readRDS(file.path(d.name, paste0(
+  scenario.name,
   "_",
-                                                fishery.name,"fishObsLenComp.rds")))
+  fishery.name, "fishObsLenComp.rds"
+)))
 
@@
 
-#add this to om_indices function so that this has years when read in
-fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))
+# add this to om_indices function so that this has years when read in
+fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time / fstepperyr))
 
 len <- len_comp_data
+
+len <- filter(len_comp_data, time %in% c(55:175))
 
@@
   theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          len$species))
+  labs(subtitle = paste(
+    scenario.name,
+    len$species
+  ))
 
-  lfplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
-
+lfplot + facet_wrap(~ time / stepperyr, ncol = 6, scales = "free_y")
 ```
@@
 
-#read in comp data
-age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
+# read in comp data
+age_comp_data <- readRDS(file.path(d.name, paste0(
+  scenario.name,
   "_",
-                                                survey.name, "survObsAgeComp.rds")))
+  survey.name, "survObsAgeComp.rds"
+)))
 age_comp_data <- age_comp_data[[1]]
@@
 
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
+Natage <- filter(age_comp_data, time %in% c(150:270))
+
+Natageplot <- ggplot(Natage, aes(x = agecl, y = atoutput)) +
   geom_point() +
   theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Natage$species))
+  labs(subtitle = paste(
+    scenario.name,
+    Natage$species
+  ))
 
-  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
+Natageplot + facet_wrap(~ time / stepperyr, ncol = 6, scales = "free_y")
 ```
@@
 
-wtage <- readRDS(file.path(d.name, paste0(scenario.name,
+wtage <- readRDS(file.path(d.name, paste0(
+  scenario.name,
   "_",
-                                                survey.name, "survObsWtAtAge.rds")))
+  survey.name, "survObsWtAtAge.rds"
+)))
 wtage <- wtage[[1]]
@@
 
-wageplot + facet_wrap(c("species"), scales="free_y")
+wageplot + facet_wrap(c("species"), scales = "free_y")
 ```
@@
 
-#read in comp data
-fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name,
+# read in comp data
+fish_age_comp <- readRDS(file.path(d.name, paste0(
+  scenario.name,
   "_",
-                                                fishery.name, "fishObsAgeComp.rds")))
+  fishery.name, "fishObsAgeComp.rds"
+)))
 fish_age_comp <- fish_age_comp[[1]]
 
-#add this to om_indices function so that this has years when read in
-fish_age_comp$time <- fish_age_comp$time/fstepperyr
+# add this to om_indices function so that this has years when read in
+fish_age_comp$time <- fish_age_comp$time / fstepperyr
 
@@
 
-Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
+Natage <- filter(fish_age_comp, time %in% c(30:53))
+
+Natageplot <- ggplot(Natage, aes(x = agecl, y = atoutput)) +
   geom_point() +
   theme_tufte() +
-    labs(subtitle = paste(scenario.name,
-                          Natage$species))
+  labs(subtitle = paste(
+    scenario.name,
+    Natage$species
+  ))
 
-  Natageplot + facet_wrap(~time, ncol=6, scales="free_y")
+Natageplot + facet_wrap(~time, ncol = 6, scales = "free_y")
 ```
@@
 ### {-}
+
+## Write data for Stock Synthesis
+
+To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`.
+You will need to ensure the relevant Stock Synthesis files are stored in the inst/extdata/ folder, with each species example in its own folder.
+
+```{r readdata}
+# devtools::install_github("r4ss/r4ss", dependencies = FALSE)
+require(r4ss)
+# source("./config/sardine_config.R") this is above for illustration
+
+# Set species name
+species <- "Pacific_sardine" # change to species_ss if we want both
+
+# Directory with SS files
+model_dir <- "Sardine_SS_files_Sep2022/"
+
+# Name of SS data file
+datfile_name <- "sardEM_3_3.dat"
+
+stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0(
+  "./inst/extdata/",
+  model_dir,
+  datfile_name
+))
+```
+
+```{r sardine-survey-spec}
+# Years to survey, assuming survey is once per year
+# survey_years <- survey_sample_full[(burnin+1):(burnin+nyears)] # my original
+survey_years <- survey_sample_full[burnin:(burnin + nyears - 1)] # from Christine's new sardine_config.R
+
+# Month of survey/fishing
+survey_month <- 7
+fishing_month <- 1
+```
+
+*Now changing catch units to tons!*
+
+Write time series biomass (tons) and catch (tons--taken directly from truth without error) data for SS using `SS_write_ts`:
+
+```{r SS-write-ts}
+
+# Test writing CPUE in biomass
+
+stocksynthesis.data <- SS_write_ts(
+  ss_data_list = stocksynthesis.data,
+  ts_data = list(
+    survObsBiom$atoutput[fish_years],
+    omlist_ss$truecatchbio_ss$atoutput[(omlist_ss$truecatchbio_ss$time / 365) %in% fish_years]
+  ),
+  CVs = c(
+    surv_cv$cv,
+    fish_cv$cv
+  ),
+  data_years = list((survObsBiom$time[fish_years] - survey_sample_time) / timestep + 1, fish_years),
+  sampling_month = list(
+    rep(survey_month, nyears),
+    rep(fishing_month, nyears)
+  ),
+  units = c("biomass", "biomass"),
+  data_type = c("CPUE", "catch"),
+  fleets = c(2, 1)
+)
+
+stocksynthesis.data$CPUE
+
+stocksynthesis.data$catch
+```
+
+Writing age composition data
+
+```{r SS-write-comps}
+# age_comp_data
+
+# Get the age bins
+age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
+age_bins <- sub("a", "", age_bin_names)
+
+age_comp_flat <- reformat_compositions(age_comp_data,
+  round.places = 4,
+  comp_type = "agecomp"
+)
+
+## Write age composition data for survey
+stocksynthesis.data <- SS_write_comps(
+  ss_data_list = stocksynthesis.data,
+  comp_matrix = list(age_comp_flat[burnin:(burnin + nyears - 1), ]),
+  data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr + nyears - 1)),
+  sampling_month = list(rep(survey_month, nyears)),
+  data_type = c("agecomp"),
+  fleet_number = c(2),
+  bins = list(age_bins),
+  caal_bool = c(FALSE)
+)
+
+stocksynthesis.data$agecomp[, age_bin_names] <- as.data.frame(age_comp_flat)[burnin:(burnin + nyears - 1), age_bins]
+```
+
+Once we have the natlength matrix, we can munge the data into the proper CAAL and length bin format for SS
+
+```{r SS-comps-formatting}
+
+# Check length compositions match age compositions
+# ss_length_stdsurv$natlength %>%
+#  group_by(time,agecl) %>%
+#  summarise(sum(atoutput))
+
+len_comp_data <- survey_lenwt[[1]]$natlength
+fish_len_comp_data <- fishery_lenwt[[1]]$natlength
+
+if (fstepperyr > 1) {
+  fish_len_comp_anndata <- fish_len_comp_data %>%
+    mutate(yr = floor(time / fstepperyr)) %>%
+    group_by(species, agecl, lower.bins, upper.bins, time = as.integer(yr)) %>%
+    summarise(annnatlength = sum(atoutput)) %>%
+    rename(atoutput = annnatlength)
+} else {
+  fish_len_comp_anndata <- fish_len_comp_data
+}
+
+caal_comp_flat <- reformat_compositions(len_comp_data,
+  round.places = 4,
+  comp_type = "caalcomp"
+)
+
+# remove burnin
+caal_comp_final <- filter(
+  caal_comp_flat,
+  time %in% survey_years
+)
+caal_comp_final <- as.data.frame(caal_comp_final)
+
+# Add over age classes to get sample size
+len_comp_flat <- reformat_compositions(len_comp_data,
+  round.places = 0,
+  comp_type = "lencomp"
+)
+# remove burnin
+len_comp_final <- filter(
+  len_comp_flat,
+  time %in% survey_years
+)
+len_comp_final <- as.data.frame(len_comp_final)
+
+length_bins <- as.integer(names(len_comp_final))
+length_bins <- length_bins[!is.na(length_bins)]
+
+# fishery length comps are still 5 timesteps per year
+# need to aggregate to annual (done above)
+# also,  make effN annual goal/fstepsperyr (done above)
+fish_len_comp_flat <- reformat_compositions(fish_len_comp_anndata,
+  round.places = 0,
+  comp_type = "lencomp"
+)
+
+# remove burnin works after adjustment above
+fish_len_comp_final <- filter(
+  fish_len_comp_flat,
+  time %in% fish_years
+)
+fish_len_comp_final <- as.data.frame(fish_len_comp_final)
+
+notbins <- c("time", "nsamp")
+
+# fish_length_bins <- as.integer(names(fish_len_comp_final))
+# fish_length_bins <- fish_length_bins[!is.na(fish_length_bins)]
+
+# need to fill empty length bins with 0s to have same bins as survey for SS_write_comps
+missing.lengths <- setdiff(length_bins, names(fish_len_comp_final)[!names(fish_len_comp_final) %in% notbins])
+fish_len_comp_final[as.character(missing.lengths)] <- 0 # Add them, filled with '0's
+fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")]
+
+# fishery age comps also 5 timesteps per year
+if (fstepperyr > 1) {
+  fish_age_comp_anndata <- catch_age_comp[[1]] %>%
+    mutate(yr = floor(time / fstepperyr)) %>%
+    group_by(species, agecl, time = as.integer(yr)) %>%
+    summarise(annnatage = sum(atoutput)) %>%
+    rename(atoutput = annnatage)
+} else {
+  fish_age_comp_anndata <- catch_age_comp[[1]]
+}
+
+fish_age_comp_flat <- reformat_compositions(fish_age_comp_anndata,
+  comp_type = "agecomp"
+)
+
+# remove burnin (not necessary?fish comps made with fish_years only)
+fish_age_comp_final <- filter(
+  fish_age_comp_flat,
+  time %in% fish_years
+)
+fish_age_comp_final <- as.data.frame(fish_age_comp_final)
+
+# #SS_write_comps breaking because fishery age bins start with 2 not 1; extracting bins from fish file may help?
+# fish_age_bins <- names(fish_age_comp_flat)[!names(fish_age_comp_flat) %in% notbins]
+
+# that leaves an empty column in data file, so instead fill with 0s
+missing.ages <- setdiff(age_bins, names(fish_age_comp_final)[!names(fish_age_comp_final) %in% notbins])
+fish_age_comp_final[missing.ages] <- 0 # Add them, filled with '0's
+fish_age_comp_final <- fish_age_comp_final[c("time", age_bins, "nsamp")]
+
+comp_list <- list(caal_comp_final, len_comp_final, fish_age_comp_final, fish_len_comp_final)
+
+apply_month <- list(
+  rep(survey_month, nrow(comp_list[[1]])),
+  rep(survey_month, nrow(comp_list[[2]])),
+  rep(fishing_month, nrow(comp_list[[3]])),
+  rep(fishing_month, nrow(comp_list[[4]]))
+)
+
+# This now runs by ensuring that survey and fishery compositions have the same bins
+# (filled with 0s for missing bins in fishery relative to survey)
+
+# Write CAAL and length composition data
+stocksynthesis.data <- SS_write_comps(
+  ss_data_list = stocksynthesis.data,
+  comp_matrix = comp_list,
+  data_rows = list((comp_list[[1]]$time - survey_sample_time) / timestep + 1, (survey_years - survey_sample_time) / timestep + 1, fish_years, fish_years),
+  sampling_month = apply_month,
+  data_type = rep(c("agecomp", "lencomp"), 2),
+  fleet_number = c(2, 2, 1, 1),
+  bins = list(
+    age_bins,
+    length_bins,
+    age_bins,
+    length_bins
+  ),
+  caal_bool = c(TRUE, rep(FALSE, 3))
+)
+
+head(stocksynthesis.data$lencomp)
+head(stocksynthesis.data$agecomp)
+
+# Change length bin structure to match atlantis data
+stocksynthesis.data$lbin_vector <- length_bins
+
+# Get correct number of length bins
+stocksynthesis.data$N_lbins <- length(length_bins)
+
+# Set lbin_method to 1 - this makes the population length bins match the data bins
+# When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
+stocksynthesis.data$lbin_method <- 1
+stocksynthesis.data$binwidth <- "#"
+stocksynthesis.data$minimum_size <- "#"
+stocksynthesis.data$maximum_size <- "#"
+
+# Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
+stocksynthesis.data$age_info$minsamplesize <- rep(0.001, 2)
+
+SS_writedat_3.30(
+  datlist = stocksynthesis.data,
+  outfile = paste0(
+    "./inst/extdata/", model_dir,
+    datfile_name
+  ),
+  overwrite = TRUE
+)
+```
+
+## Coming up with dynamic life history parameters (partial from Christine)
+
+Below we get the weight-length relationship parameters from atlantis, calculate h and R0 from Atlantis $\alpha$ and $\beta$ parameters, use the \code{calc_Z} function to back out natural mortality (M) from Atlantis, and use the survey length-at-age from Atlantis to estimate a growth curve.
+
+The following is under construction from CreateStockSynthesis.Rmd and not evaluated.
+Existing control file was modified by hand to make the number of recruitment settlement assignments 1 instead of two, with GPattern month area age vector 1 1 1 1.
+
+```{r, eval=FALSE}
+
+# # Need to clone atlantisom and load the package locally because the
+# # installed package from remote repository does not include SS_write_biol()
+# devtools::load_all("C:/Users/bai.li/Documents/sardine/atlantisom")
+# 
+# # Get biological parameters
+# 
+# # Load needed inputs for biological parameters
+# 
+# # biological parameters
+# biolprm <- load_biolprm(dir = d.name, file_biolprm = biol.prm.file) # "C:/Users/chris/Documents/GitHub/atlantisom"
+# # functional groups
+# 
+# fgs <- CCom$truth$fgs
+# runprm <- load_runprm(dir = d.name, file_runprm = run.prm.file)
+# YOY <- load_yoy(d.name, paste0(scenario.name, "YOY.txt"))
+# truenums_ss <- CCom$truth$nums[CCom$truth$nums$species == species, ]
+# 
+# YOY_ss <- YOY %>%
+#   select(Time, "SAR.0")
+# 
+# fullresZ <- calc_Z(
+#   yoy = YOY_ss,
+#   nums = truenums_ss,
+#   fgs = fgs,
+#   biolprm = biolprm,
+#   toutinc = runprm$toutinc
+# )
+# Z <- mean(fullresZ$atoutput)
+# 
+# meanwt_spp <- survey_lenwt[[1]]$muweight %>%
+#   filter(time > burnin) %>%
+#   group_by(agecl) %>%
+#   summarize(meanwt = mean(atoutput))
+# 
+# ctlfile_name <- "sardEM_3_3.ctl"
+# sardine.ctl <- r4ss::SS_readctl_3.30(paste0(
+#   "./inst/extdata/",
+#   model_dir,
+#   ctlfile_name
+# ),
+# datlist = paste0(
+#   "./inst/extdata/", model_dir,
+#   datfile_name
+# )
+# )
+# 
+# lensage <- data.frame(
+#   agecl = age_classes[[species]],
+#   meanln = seq(15.188, 48.39, length = length(age_classes[[species]])),
+#   cvln = seq(15.188 * 0.25, 0.9, length = length(age_classes[[species]]))
+# )
+# cvo_guess <- 48.39
+# 
+# sardine.ctl <- atlantisom::SS_write_biol(sardine.ctl,
+#   biolprm,
+#   "SAR",
+#   wtsage = meanwt_spp,
+#   M_est = rep(0.4, length(age_classes[[species]])),
+#   lensage = lensage
+# )
+# 
+# li_a_use <- biolprm$wl[match(fgs$Code[match(species, fgs$Name)], biolprm$wl[, 1]), 2] / 1000
+# li_b_use <- biolprm$wl[match(fgs$Code[match(species, fgs$Name)], biolprm$wl[, 1]), 3]
+# 
+# # maturity ogive
+# # check what R0 is
+# 
+# 
+# 
+# len_nonburn <- survey_lenwt[[1]]$mulen %>%
+#   filter(time > burnin * timestep)
+# plot(len_nonburn$atoutput ~ len_nonburn$agecl, col = len_nonburn$time)
+# 
+# length.data <- data.frame("Year" = (len_nonburn$time - survey_sample_time) / timestep + 1, length = len_nonburn$atoutput, Weight = NA, Sex = "Female", age = as.integer(len_nonburn$agecl))
+# 
+# # require(bbmle)
+# # vb_estimates <-sample_fit_vbgf(length.data, 25, 45, 0.4,
+# #  0.1, 0.1, 20, 40, 0.05, 0.01, 0.01,
+# #  30, 50, 0.6, 0.3, 0.3, 0.5, 10)
+```
+
+## Run Stock Synthesis
+After the data and control file modifications, ensure you copy the local SS 3.3 executable into the folder containing the model files. Then run the following chunk.
+
+```{r runSS}
+# Run stock synthesis model
+# default runs with admb_options = "-noest"
+# run_stocksynthesis(species = species, model_dir = model_dir, show_output = TRUE)
+
+# for estimation
+run_stocksynthesis(species = species, model_dir = model_dir, admb_options = "", show_output = TRUE)
+```

### 2023-05-17 NewCatchWorkflow.Rmd
get organized on new atlantisom catch workflow

---
title: "New catch functions from CATCH.nc file"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), ''%d %B %Y'')`"
output: html_document
---

## Intro

Total catch in tons has been pulled from the Atlantis `CATCH.txt` output file from 2019-2023 due to difficulties scaling catch in numbers from the `CATCH.nc` file to total weight. See comparisons [here](https://sgaichas.github.io/poseidon-dev/TestCatchNOBA.html).

For testing multispecies models, there is a need for subannual catch information, which is not provided in the annual `CATCH.txt` output. In addition, there is no polygon information in the `CATCH.txt` output, so spatial subsetting is not possible.

In order to get catch in tons subannually and by area, we can use the fleet-specific catch information in the `CATCH.nc` file. These outputs are reported in tons by species in aggregate over all age classes for each defined fleet.

## Methods

Workflow

1.  Modify `atlantisom::load_nc` to get catch in tons from different `CATCH.nc` outputs. New function because current function already gets catch in numbers, would be complicated to use the same "Catch" variable to get tons. Name new function `atlantisom::load_nc_catchtons`

2.  Test new `atlantisom::load_nc_catchtons` with NOBA, NEUS, CC by summing output over polygons and fleets to each year and comparing with annual output of current `atlantisom::load_catch` based on `CATCH.txt`

3.  Incorporate `atlantisom::load_nc_catchtons` into `atlantisom::run_truth`. Add a new catchtons object keeping polygon, timestep, and fleet information to the output list. Consider removing the `CATCH.txt` output in this function, currently called `catch_all` but not used? Or keep it, since catchtons can be summed to catch_all this could be used as an internal test as was started in the run_truth function.

4.  Adjust `atlantisom::om_init` for the new catchtons output. Should we keep reading in the `CATCH.txt` here? It may be ok to keep it just to keep existing functions for aggregate annual catch working.

5.  Adjust `atlantisom::om_species` for the new catchtons output.

6.  Adjust `atlantisom::om_index` for the new catchtons output. Create a new index that will be consistent across the subannual and annual datasets: this means applying the cv at the level of reporting? So possibly by fleet and timestep, then summing to annual? This involves including and updating two functions: `atlantisom::create_fishery_subset` parallel to the `atlantisom::create_survey` function to subset areas, currently used only on fishery comps but can now be applied to the index as well, and updating another `atlantisom::sample_fishery_totalcatch` for the cv application. This will save perhaps one new output and update one existing output, or just one output? Will we have the same current `censusfishCatch.rds` output that is annually aggregated plus a disaggregated version? Or have a single output with the same name `censusfishCatch.rds` that is disaggregated, to be summed separately in mskeyrun?

7. Adjust `mskeyrun` functions creating aggregated and subannual catch datasets. Do we need a fleet specific index?

### 1. New `atlantisom::load_nc_catchtons` function

Here is the new script, tested with snippets in mskeyrun sarah_wgsamsim branch and implemented in atlantisom dev branch

```{r}

```

### Test `load_nc_catchtons` with the NOBA, NEUS, CC model outputs

```{r}

```

### 2023-05-17 Test_load_biolprm.R

@@ -1,0 +1,29 @@
+# test new load_biolprm to bring in gape width pars
+
+dir <- "/Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38"
+
+file_biolprm <- "nordic_biol_incl_harv_v_011_1skg.prm"
+
+load_biolprm(dir, file_biolprm)
+
+dir <- "/Users/sarah.gaichas/Documents/0_Data/Atlantis/Poseidon/poseidon-dev/atlantisoutput/CC2063_Sep2022"
+
+file_biolprm <- "CalCurrentV3_Biol.prm"
+
+load_biolprm(dir, file_biolprm)
+
+config <- "/Users/sarah.gaichas/Documents/0_Data/ms-keyrun/data-raw/simulated-data/config/NOBA_sacc38Config.R"
+
+d.name <- "/Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38"
+functional.groups.file <- "nordic_groups_v04.csv"
+biomass.pools.file <- "nordic_biol_v23.nc"
+biol.prm.file <- "nordic_biol_incl_harv_v_011_1skg.prm"
+box.file <- "Nordic02.bgm"
+initial.conditions.file <- "nordic_biol_v23.nc"
+run.prm.file <- "nordic_run_v01.xml"
+scenario.name <- "nordic_runresults_01"
+bioind.file <- "nordic_runresults_01BiomIndx.txt"
+catch.file <- "nordic_runresults_01Catch.txt"
+annage <- TRUE
+fisheries.file <- "NoBAFisheries.csv"
+

### 2023-05-17 NewCatchWorkflow.Rmd
test with NOBA sacc30-climate run

@@
 author: "Sarah Gaichas"
-date: "`r format(Sys.time(), ''%d %B %Y'')`"
-output: html_document
+date: "`r format(Sys.time(), '%d %B %Y')`"
+output:
+  html_document:
+    code_fold: hide
 ---
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r message=FALSE, warning=FALSE}
+library(here)
+library(stringr)
+library(magrittr)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(atlantisom)
+```
 
@@
 
-Here is the new script, tested with snippets in mskeyrun sarah_wgsamsim branch and implemented in atlantisom dev branch
+Here is the new function, tested with snippets in mskeyrun sarah_wgsamsim branch and implemented in atlantisom dev branch
+
+```{r, code = readLines("https://raw.githubusercontent.com/r4atlantis/atlantisom/dev/R/load_nc_catchtons.R"), eval=F}
+```
+
+### 2. Test `load_nc_catchtons` with the NOBA, NEUS, CC model outputs
+
+Only NOBA first, this one sacc30 with climate, different from sacc38 in mskeyrun without
 
@@
 
+# define for each model
+
+## NOBA (different from mskeyrun version)
+
+atlmod <- here("config/NOBAsacc30Config.R")
+
+select_groups <- c("Long_rough_dab",
+                   "Green_halibut",
+                   "Mackerel",
+                   "Haddock",
+                   "Saithe",
+                   "Redfish",
+                   "Blue_whiting",
+                   "Norwegian_ssh",
+                   "North_atl_cod",
+                   "Polar_cod",
+                   "Capelin")
+
+stepperyr <- 5
+
+## CC
+
+## NEUS
+
+##### aggregate over polygons to fleets #######
+# make this a function
+
+source(atlmod)
+
+dir <- d.name
+nc_catch <- paste0(scenario.name, 'CATCH.nc')
+file_fgs <- functional.groups.file
+file_init <- initial.conditions.file
+file_fish <- fisheries.file
+
+# Get the boundary boxes
+allboxes <- atlantisom::load_box(dir = d.name, file_bgm = box.file)
+boxes <- atlantisom::get_boundary(allboxes)
+
+  # Read in information
+  # Read in the functional groups csv since that is used by many functions
+  fgs <- load_fgs(dir = dir, file_fgs = file_fgs)
+  # Read in the biomass pools
+  bps <- load_bps(dir = dir, fgs = file_fgs, file_init = file_init)
+
+  catchtons <- load_nc_catchtons(dir = dir,
+                                 file_nc = nc_catch,
+                                 file_fish = file_fish,
+                                 bps = bps,
+                                 fgs = fgs,
+                                 select_groups = select_groups,
+                                 select_variable = "Catch",
+                                 check_acronyms = TRUE,
+                                 bboxes = boxes)
+  #if(verbose) message("Catch tons read in.")
+
+# result is output of load_nc_catchtons
+
+aggcatchtons <- catchtons %>%
+  dplyr::select(-agecl) %>%
+  dplyr::group_by(species, fleet, time) %>%
+  dplyr::summarise(totcatch = sum(atoutput)) %>%
+  dplyr::mutate(year = ceiling(time/stepperyr))
+
+# compare annual catch in tons to CATCH.txt output
+yearcatchtons <- aggcatchtons %>%
+  dplyr::group_by(species, year) %>%
+  dplyr::summarise(yeartot = sum(totcatch))
+
+txtCatchtons <- atlantisom::load_catch(d.name, catch.file, fgs) %>%
+  dplyr::filter(species %in% select_groups) %>%
+  dplyr::mutate(year = time/365) %>%
+  dplyr::left_join(dplyr::select(fgs, Code, Name), by = c("species" = "Name"))
+
+comparetons <- txtCatchtons %>%
+  dplyr::left_join(yearcatchtons, by = c("Code" = "species", "year" = "year")) %>%
+  ggplot2::ggplot() +
+  geom_line(aes(year, atoutput)) +
+  geom_point(aes(year, yeartot), color = "blue", alpha = 0.2) +
+  facet_wrap(~species, scales="free_y")
+
+# looks like a match!               
+comparetons
 ```
-
-### Test `load_nc_catchtons` with the NOBA, NEUS, CC model outputs
-
-```{r}
-
-```

### 2023-05-17 docs/index.Rmd
add new CC run
test with NOBA sacc30-climate run

@@
 <a href="https://sgaichas.github.io/poseidon-dev/atlantisom_landingpage.html" target="_blank">Draft atlantisom landing page for review, June 9, 2020</a>
+
+### New workflow testing for WGSAM simulated datasets
+
+<a href="https://sgaichas.github.io/poseidon-dev/NewCatchWorkflow.html" target="_blank"> Total catch in tons now by polygon, fleet, and subannual timestep, in honor of Norwegian Constitution Day, May 17, 2023</a>
+
+### New CCA run
+
+<a href="https://sgaichas.github.io/poseidon-dev/NewCCASardines.html" target="_blank"> Test run from Isaac, June 30, 2022</a>
 

### 2023-05-17 NewCatchWorkflow.Rmd
update with all 3 models matching

@@
 
-Only NOBA first, this one sacc30 with climate, different from sacc38 in mskeyrun without
+For all plots, the total catch derived from CATCH.txt is considered correct, and is represented by the black line. The total catch summed across polygon, subannual timestep and fleet is in blue.
 
-```{r}
+Only NOBA first, this one is model sacc30 with climate, different from sacc38 in `mskeyrun` without climate:
+
+```{r, message=FALSE, warning=FALSE}
 
@@
 
-## CC
-
 ## NEUS
@@
 
-source(atlmod)
+comparecatch <- function(atlmod, select_groups, stepperyr) {
+  source(atlmod, local = TRUE)
 
@@
   # Read in the biomass pools
-  bps <- load_bps(dir = dir, fgs = file_fgs, file_init = file_init)
+  bps <- load_bps(dir = dir,
+                  fgs = file_fgs,
+                  file_init = file_init)
 
-  catchtons <- load_nc_catchtons(dir = dir,
+  catchtons <- load_nc_catchtons(
+    dir = dir,
     file_nc = nc_catch,
@@
     check_acronyms = TRUE,
-                                 bboxes = boxes)
+    bboxes = boxes
+  )
   #if(verbose) message("Catch tons read in.")
@@
     dplyr::summarise(totcatch = sum(atoutput)) %>%
-  dplyr::mutate(year = ceiling(time/stepperyr))
+    dplyr::mutate(year = ceiling(time / stepperyr))
 
@@
     dplyr::filter(species %in% select_groups) %>%
-  dplyr::mutate(year = time/365) %>%
+    dplyr::mutate(year = time / 365) %>%
     dplyr::left_join(dplyr::select(fgs, Code, Name), by = c("species" = "Name"))
@@
     geom_point(aes(year, yeartot), color = "blue", alpha = 0.2) +
-  facet_wrap(~species, scales="free_y")
+    facet_wrap( ~ species, scales = "free_y")
 
@@
   comparetons
+
+}
+
+comparecatch(atlmod = atlmod, 
+             select_groups = select_groups,
+             stepperyr = stepperyr)
+
 ```
+
+Now try the Isaac's recent CC model:
+
+```{r, message=FALSE, warning=FALSE}
+## CC
+
+atlmod <- here("config/CCConfigSep22.R")
+
+select_groups <- c("Arrowtooth_flounder",
+                   "Petrale_sole",
+                   "Pacific_sardine",
+                   "Anchovy",
+                   "Herring",
+                   "Pacific_Ocean_Perch",
+                   "Bocaccio_rockfish",
+                   "Yelloweye_rockfish",
+                   "Mesopel_M_Fish", #Pacific hake
+                   "Mesopel_N_Fish", #Sablefish
+                   "Demersal_P_Fish") #Dover sole
+
+stepperyr <- 5
+
+comparecatch(atlmod = atlmod, 
+             select_groups = select_groups,
+             stepperyr = stepperyr)
+```
+
+And finally the NEUS model, thanks to Joe for output files:
+
+```{r, message=FALSE, warning=FALSE}
+## NEUS
+atlmod <- here("config/NEUStestConfigMay23.R")
+
+select_groups <- c("Summerflounder",
+                   "Winterflounder",
+                   "Mackerel",
+                   "Herring",
+                   "Cod",
+                   "Monkfish",
+                   "Haddock",
+                   "Spiny_Dogfish",
+                   "Winter_Skate",
+                   "Yellowtail_Flounder",
+                   "Silver_Hake")
+
+stepperyr <- 5
+
+comparecatch(atlmod = atlmod, 
+             select_groups = select_groups,
+             stepperyr = stepperyr)
+
+```
+
+Looks like we have a match to CATCH.txt output for all systems using the same function and inputs from the CATCH.nc file.
+However, it looks like there is a final value in CATCH.txt that is not present, or not being summed correctly in the CATCH.nc output in the CC and NEUS models.
+Worth investigating...
+
+### 3. Incorporate new `load_nc_catchtons` function into `run_truth`

### 2023-05-17 config/NEUStestConfigMay23.R

@@ -1,0 +1,12 @@
+d.name <- here("atlantisoutput","NEUStest_May2023")
+functional.groups.file <- "neus_groups_noturtles.csv"
+biomass.pools.file <- "neus_init.nc"
+biol.prm.file <- "at_biology.prm"
+box.file <- "neus_tmerc_RM2.bgm"
+initial.conditions.file <- "neus_init.nc"
+run.prm.file <- "at_run.xml" #currently missing, only prm version
+scenario.name <- "neus_output"
+bioind.file <- ""
+catch.file <- "neus_outputCatch.txt"
+annage <- TRUE
+fisheries.file <- "neus_fisheries.csv"

### 2023-05-18 NewCatchWorkflow.Rmd
links to new run_truth which works

@@
 
-Total catch in tons has been pulled from the Atlantis `CATCH.txt` output file from 2019-2023 due to difficulties scaling catch in numbers from the `CATCH.nc` file to total weight.
+Total catch in tons has been pulled from the Atlantis `Catch.txt` output file from 2019-2023 due to difficulties scaling catch in numbers from the `CATCH.nc` file to total weight.
 See comparisons [here](https://sgaichas.github.io/poseidon-dev/TestCatchNOBA.html).
 
-For testing multispecies models, there is a need for subannual catch information, which is not provided in the annual `CATCH.txt` output.
-In addition, there is no polygon information in the `CATCH.txt` output, so spatial subsetting is not possible.
+For testing multispecies models, there is a need for subannual catch information, which is not provided in the annual `Catch.txt` output.
+In addition, there is no polygon information in the `Catch.txt` output, so spatial subsetting is not possible.
 
@@
 2.
-Test new `atlantisom::load_nc_catchtons` with NOBA, NEUS, CC by summing output over polygons and fleets to each year and comparing with annual output of current `atlantisom::load_catch` based on `CATCH.txt`
+Test new `atlantisom::load_nc_catchtons` with NOBA, NEUS, CC by summing output over polygons and fleets to each year and comparing with annual output of current `atlantisom::load_catch` based on `Catch.txt`
 
@@
 Add a new catchtons object keeping polygon, timestep, and fleet information to the output list.
-Consider removing the `CATCH.txt` output in this function, currently called `catch_all` but not used?
+Consider removing the `Catch.txt` output in this function, currently called `catch_all` but not used?
 Or keep it, since catchtons can be summed to catch_all this could be used as an internal test as was started in the run_truth function.
@@
 Adjust `atlantisom::om_init` for the new catchtons output.
-Should we keep reading in the `CATCH.txt` here?
+Should we keep reading in the `Catch.txt` here?
 It may be ok to keep it just to keep existing functions for aggregate annual catch working.
@@
 
-For all plots, the total catch derived from CATCH.txt is considered correct, and is represented by the black line. The total catch summed across polygon, subannual timestep and fleet is in blue.
+For all plots, the total catch derived from Catch.txt is considered correct, and is represented by the black line. The total catch summed across polygon, subannual timestep and fleet is in blue.
 
@@
 
-  # compare annual catch in tons to CATCH.txt output
+  # compare annual catch in tons to Catch.txt output
   yearcatchtons <- aggcatchtons %>%
@@
 
-Looks like we have a match to CATCH.txt output for all systems using the same function and inputs from the CATCH.nc file.
-However, it looks like there is a final value in CATCH.txt that is not present, or not being summed correctly in the CATCH.nc output in the CC and NEUS models.
+Looks like we have a match to Catch.txt output for all systems using the same function and inputs from the CATCH.nc file.
+However, it looks like there is a final value in Catch.txt that is not present, or not being summed correctly in the CATCH.nc output in the CC and NEUS models.
 Worth investigating...
 
-### 3. Incorporate new `load_nc_catchtons` function into `run_truth`
+### 3. Incorporate new `load_nc_catchtons` function into `atlantisom::run_truth`
+
+Added catchtons and disctons (same function with `select_variable = "Discards"`) to `run_truth` in the atlantisom dev branch.
+At present, catchtons and disctons are in the run_truth output and catch_all which was originally from the `Catch.txt` file have been removed from the output.
+Sections of the function that read in both the `Catch.txt` and the `CatchPerFishery.txt` files have been commented out.
+They could be reinstated if needed, but `om_init` reads in `Catch.txt` already and seems like a good place to read in text files.
+
+Here is the new function, tested by re-running the code in `SimData.Rmd` in the mskeyrun repo using the NOBA sacc38 outputs.
+The new runtruth object includes the new catchtons and disctons (0) objects.
+
+```{r, code = readLines("https://raw.githubusercontent.com/r4atlantis/atlantisom/dev/R/run_truth.R"), eval=F}
+```
+
+### 4.  Adjust `atlantisom::om_init` for the new catchtons output

### 2023-05-18 NewCatchWorkflow.Rmd
up to the last, most complex step

@@
 ### 4.  Adjust `atlantisom::om_init` for the new catchtons output
+
+Actually, no adjustment needed for `om_init`. This was run as is to produce the new truth output. I think we keep reading in `Catch.txt` here.
+
+### 5.  Adjust `atlantisom::om_species` for the new catchtons output
+
+Added the catchtons and disctons truth outputs here, subset by code rather than species name, update documentation.
+Note that code will need to be changed to species prior to applying the fishery sampling functions.
+
+New function, tested by re-running code in `SimData.Rmd` in mskeyrun as above.
+The most recent omlist_ss.rds file in the mskeyrun simulated data atlantisoutput NOBAsacc38 directory reflects the updated catch in tons at full resolution.
+
+```{r, code = readLines("https://raw.githubusercontent.com/r4atlantis/atlantisom/dev/R/om_species.R"), eval=F}
+```
+
+### 6. Adjust `atlantisom::om_index` for the new catchtons output
+
+The complex part.

### 2023-05-18 NewCatchWorkflow.Rmd
thinking through fleets

@@
 In order to get catch in tons subannually and by area, we can use the fleet-specific catch information in the `CATCH.nc` file.
-These outputs are reported in tons by species in aggregate over all age classes for each defined fleet.
+These outputs are reported in tons by species aggregated over all age classes for each defined fleet.
+Variable names in the `CATCH.nc` file for these outputs follow the format "XXX_Catch_FCN" and "XXX_Discard_FCN" where XXX is the 3 letter group Code defined in the Atlantis `groups.csv` file and N is the integer Index for the fishery defined in the Atlantis `fisheries.csv` file.
 
@@
 New function, tested by re-running code in `SimData.Rmd` in mskeyrun as above.
-The most recent omlist_ss.rds file in the mskeyrun simulated data atlantisoutput NOBAsacc38 directory reflects the updated catch in tons at full resolution.
+The most recent `omlist_ss.rds` file in the mskeyrun simulated data atlantisoutput NOBAsacc38 directory reflects the updated catch in tons at full resolution.
 
@@
 The complex part.
+This requires adjusting the `atlantisom::create_fishery_subset` to take catchtons as an input, then feeding this into the index using `atlantisom::sample_fishery_totalcatch` and applying cv at a more disaggregated point, then summing to the annual index so that subannual and annual indices are consistent.
+Then `om_index` will output the subannual index only, or both?
+It should probably just make one and users can decide how to aggregate to annual (ie.
+mskeyrun will provide both a subannual and annual index based on the same `om_index` output).
+
+#### 6a. Update `om_index`
+
+The fishery portion of this wrapper can now have parallel structure to the survey portion.
+It allows multiple fishery.R config files so the user can get a differet index for each fleet if that is desired.
+Specifications for which polygons are kept for which fleet number would be in the fishery.R config file.
+If only one file we can have it default to aggregating across fleets for a species first.
+
+```{r}
+
+```
+
+#### 6b. Update `create_fishery_subset`
+
+This needs modification to use the fleet column found in catchtons.
+
+```{r}
+
+```
+
+#### 6c. Update `sample_fishery_totalcatch`
+
+Applies cv at the subannual level for the specified fishery.
+
+```{r}
+
+```

### 2023-05-23 NewCatchWorkflow.Rmd
more notes to update

@@
 Specifications for which polygons are kept for which fleet number would be in the fishery.R config file.
-If only one file we can have it default to aggregating across fleets for a species first.
+
+If only one file we can have it default to aggregating across fleets for a species first: for each fishery.R file the first step is to select fleets identified in fishery.R and aggregate across those fleets in the input.
+Or we do this by default for each fleet and have users sum at the end?
+
+Note that allowing for multiple fishery.R files for each fleet means we need to make a parallel change to `om_comps`.
 

### 2023-06-09 NewCatchWorkflow.Rmd
all but the testing

--- 2023-05-18 NewCatchWorkflow.Rmd
@@
 The fishery portion of this wrapper can now have parallel structure to the survey portion.
-It allows multiple fishery.R config files so the user can get a differet index for each fleet if that is desired.
-Specifications for which polygons are kept for which fleet number would be in the fishery.R config file.
-If only one file we can have it default to aggregating across fleets for a species first.
+It allows multiple fishery.R config files so the user can get a different index for each fleet if that is desired.
+Specifications for which polygons are kept for which fleet number would be in the `fishery.R` config file, with an appropriate fleet.name.
 
-```{r}
+I think we want to aggregate across all fleets for a species for the given set of polygons: for each `fishery.R` file the first step is to select fleets identified in `fishery.R` and aggregate across those fleets for the polygons specified in `fishery.R`, keeping species separated out--this is done in `create_fishery_subset` below.
+
+Note that allowing for multiple `fishery.R` files for each fleet means we need to make a parallel change to `om_comps`.
+
+Note that species in here are the codes not the names, so conversion back to names needs to happen somewhere.
+The fishery.R file now specifies the species codes in fishspp instead of species names (survspp still does that).
+
+Here is the new `om_index`
+
+```{r, code = readLines("https://raw.githubusercontent.com/r4atlantis/atlantisom/dev/R/om_index.R"), eval=F}
+
+```
+
+And an example `fishery.R` file (modified `msfishery.R` file used in ms-keyrun project):
+
+```{r, code = readLines("https://raw.githubusercontent.com/NOAA-EDAB/ms-keyrun/sarah_wgsamsim/data-raw/simulated-data/config/msfishery.R"), eval=F}
 
@@
 
+Left alone, this function will aggregate across all fleets and polygons given in the input data (since it is not expecting a fleet column).
 This needs modification to use the fleet column found in catchtons.
+That way the user can control which fleets are included in the aggregated output.
+Specifying fleets=NULL means all fleets are aggreagated.
 
-```{r}
+```{r, code = readLines("https://raw.githubusercontent.com/r4atlantis/atlantisom/dev/R/create_fishery_subset.R"), eval=F}
 
@@
 
-#### 6c. Update `sample_fishery_totalcatch`
+#### 6c. Update `sample_fishery_totcatch`
 
-Applies cv at the subannual level for the specified fishery.
+Now this function determines if there are polygons in the input, and aggregates over polygon if there are. It then applies cv at the subannual level for the specified fishery.
 
-```{r}
+```{r, code = readLines("https://raw.githubusercontent.com/r4atlantis/atlantisom/dev/R/sample_fishery_totcatch.R"), eval=F}
 
 ```
+
+#### Testing
+
+Compare the previous total catch time series (annual) with the new one summed to annual and hope the mostly match (both have a cv applied but it is 0.01 for all species so shouldn't matter.)

### 2023-06-09 NewCatchWorkflow.Rmd
matches, so far so good

@@
 
+Time in this output is *timestep* not days.
+
 Here is the new `om_index`
@@
 Compare the previous total catch time series (annual) with the new one summed to annual and hope the mostly match (both have a cv applied but it is 0.01 for all species so shouldn't matter.)
+
+Read in saved files created previously in ms-keyrun and the one just created with the new functions. This is now comparing the "sampled" fishery data.
+
+```{r}
+
+mskeyrundir <- "/Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38"
+
+scenario.name <- "nordic_runresults_01"
+
+stepperyr <- 5
+
+fisherycomp <- atlantisom::read_savedfisheries(mskeyrundir, "Catch")
+
+from_nc <- fisherycomp$allfleet[[1]]
+
+from_txt <- fisherycomp$census[[1]]
+
+fgs <- atlantisom::load_fgs(mskeyrundir, "nordic_groups_v04.csv")
+
+  aggcatchtons <- from_nc %>%
+    dplyr::select(species, time, atoutput) %>%
+    dplyr::mutate(year = ceiling(time / stepperyr)) %>%
+    dplyr::group_by(species, year) %>%
+    dplyr::summarise(yeartot = sum(atoutput))
+
+  txtCatchtons <- from_txt %>%
+    dplyr::select(species, time, atoutput) %>%
+    dplyr::mutate(year = time / 365) %>%
+    dplyr::left_join(dplyr::select(fgs, Code, Name), by = c("species" = "Name"))
+
+  comparetons <- txtCatchtons %>%
+    dplyr::left_join(aggcatchtons, by = c("Code" = "species", "year" = "year")) %>%
+    ggplot2::ggplot() +
+    geom_line(aes(year, atoutput)) +
+    geom_point(aes(year, yeartot), color = "blue", alpha = 0.2) +
+    facet_wrap( ~ species, scales = "free_y")
+
+  # looks like a match!
+  comparetons
+
+```

### 2023-06-12 NewCatchWorkflow.Rmd
comps complications

@@
 ```
+
+We have a match so it seems the workflow gives the same results in aggregate. Lets call the fleet specific and subannual catch good then. It can go into `mskeyrun`.
+
+### 7. Adjust `mskeyrun` functions creating aggregated and subannual catch datasets
+
+See mskeyrun repo.
+I would like to keep life simple and not have fleet specific subannual catch indices, but since we just did fleet specific indices with the real Georges Bank data I suppose I will have to test this with simulated data.
+
+But maybe not for the WGSAM skill assessment. See below for why I don't quite have a full set of fleet specific outputs.
+
+### 8. Adjust `atlantisom::om_comps` for the new catchtons output
+
+Well this sounded easier than it is.
+We can have multiple fleets in `om_comps` using a similar structure to surveys.
+Sadly, the length comps from fisheries cannot be by fleet because they come from the CATCH.nc numbers output, which is by age class and not by fleet.
+Lengths are estimated using the age class information by numbers, structn, and resn.
+I don't think we can apply the `calcage2length` function to the true age data because we don't have the growth information from structn and resn at that level.
+
+I can rewrite `om_comps` to use fleet information on all but the lengths. Placeholder code below.
+
+```{r, code = readLines("https://raw.githubusercontent.com/r4atlantis/atlantisom/dev/R/om_comps.R"), eval=F}
+
+```

### 2023-06-14 NewBiomassPoolsWorkflow.Rmd
start biomass pools for rpath sim data

---
title: "New Biomass Pools Workflow"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output:
  html_document:
    code_fold: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(here)
library(stringr)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(atlantisom)
```

## Intro

Current `atlantisom` functions are designed primarily for age structured modeled groups, because the package was originally scoped to generate simulated data to test fishery stock assessments. It is now clear that simulated data testing can also benefit other model types, including multispecies models and food web models. Therefore, there is a need to develop functions to develop simulated data for lower trophic level biomass pools modeled in Atlantis and also included in food web models such as Rpath.

Where is biomass information for non-age structured groups output? To be parallel to survey sampling for fish, it would be good to have it by model output timestep toutinc and polygon, but depth layer is not necessary at this time.

First check content of the `.nc` output file:

```{r}

dir <- here("atlantisoutput/NOBA_sacc_30")
file_nc <- "nordic_runresults_01.nc"

file.nc <- file.path(dir, file_nc)

 # Load ATLANTIS output!
  at_out <- RNetCDF::open.nc(con = file.nc)
  on.exit(RNetCDF::close.nc(at_out))

    # Get info from netcdf file! (Filestructure and all variable names)
  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length

```

Which species are biomass pools, are they identified in the groups.csv?

No, biomass pools are defined as the combination of NumCohorts == 1 and GroupType not FISH, FISH_INVERT, SHARK, BIRD, MAMMAL, REPTILE. Biomass pools groups are among those below according to the Atlantis manual, ch 6. p 106:

>BIOMASS POOL GROUPS
Phytoplankton
7) Small Phytoplankton SM_PHY
8) Large Phytoplankton LG_PHY (used in combination with IsSiliconDep
to define diatoms)
9) Trichodesmium and Cyanobacteria TRICHO (N fixers)
Other primary producers
10) Microphtybenthos MICROPHTYBENTHOS
11) Dinoflagellates DINOFLAG
12) Phytoben PHYTOBEN (typically used to represent macroalgae) 
13) Seagrass SEAGRASS
14) Turf TURF
Zooplankton
15) Small Zooplankton SM_ZOO
16) Medium Zooplankton MED_ZOO
17) Large Zooplankton LG_ZOO
18) Jellyfish JELLIES (in the past represented as LG_ZOO)
Large pelagic invetebrates
19)Cephalopod CEP 
20) Prawns PWN
Infauna
21) Small Infauna SM_INF 
22) Large Infauna LG_INF
Epibenthic organisms
23) Sediment epibenthic filter feeders SED_EP_FF (often used for bivalves, sponges)

For NOBA Atlantis, these are the biomass pools:

```{r}

fgs <- atlantisom::load_fgs(here("atlantisoutput/NOBA_sacc_30"), "nordic_groups_v04.csv")

pools <- fgs %>%
  dplyr::filter(NumCohorts==1) %>%
  dplyr::select(Code, Name, Long.Name, NumCohorts, InvertType)

knitr::kable(pools)

```

Here are the relevant outputs in the output.nc file for these groups:

```{r}
# return everything in var_names_ncdf matching Name in pools

var_names_ncdf[str_detect(var_names_ncdf, str_c(pools$Name, collapse ="|"))]

```
So we want "N" nitrogen output. In theory this can be achieved with `atlantisom::load_nc`with `select_variable = "N"`.

```{r}

select_groups <- pools$Name
nc_out <- file_nc
bps <- load_bps(dir = dir, fgs = "nordic_groups_v04.csv", file_init = "nordic_biol_v23.nc")
  # Get the boundary boxes
  allboxes <- load_box(dir = dir, file_bgm = "Nordic02.bgm")
  boxes <- get_boundary(allboxes)

testNpools <- load_nc(dir = dir,
                  file_nc = nc_out,
                  bps = bps,
                  fgs = fgs,
                  select_groups = select_groups,
                  select_variable = "N",
                  check_acronyms = TRUE,
                  bboxes = boxes)
```

I don't think we have a function that converts N to biomass but there are pieces in other functions.

### 2023-06-14 NewBiomassPoolsWorkflow.Rmd
outputs of load_nc work for N

@@
 ```
-So we want "N" nitrogen output. In theory this can be achieved with `atlantisom::load_nc`with `select_variable = "N"`.
+So we want "N" nitrogen output.
+In theory this can be achieved with `atlantisom::load_nc`with `select_variable = "N"`.
+Looks like it works, output is N for our selected biomass pool species over all polygons, layers, and times.
 
@@
                   bboxes = boxes)
+
+unique(testNpools$species)
 ```
 
-I don't think we have a function that converts N to biomass but there are pieces in other functions.
+I don't think we have a function that aggregates over layets and converts N to biomass but there are pieces in other functions.

### 2023-06-15 NewBiomassPoolsWorkflow.Rmd
starts biomass pools sampling for Rpath

@@
   at_out <- RNetCDF::open.nc(con = file.nc)
-  on.exit(RNetCDF::close.nc(at_out))
+  #on.exit(RNetCDF::close.nc(at_out))
 
@@
 7) Small Phytoplankton SM_PHY  
-8) Large Phytoplankton LG_PHY (used in combination with IsSiliconDep
-to define diatoms)
+8) Large Phytoplankton LG_PHY (used in combination with IsSiliconDep to define diatoms)  
 9) Trichodesmium and Cyanobacteria TRICHO (N fixers)  
@@
 23) Sediment epibenthic filter feeders SED_EP_FF (often used for bivalves, sponges)  
+24) Benthic grazers SED_EP_OTHER  
+25) Mobile epibenthos MOB_EP_OTHER (often used for crabs, lobster, octopus)   
+26) Corals CORAL  
+27)Sponges SPONGE  
+Bacteria   
+28) Pelagic Bacteria PL_BACT  
+29) Sediment Bacteria SED_BACT    
+Obligatory detritus groups   
+30) Carrion CARRION   
+31) Labile detritus LAB_DET   
+32) Refractory detritus REF_DET   
+33) Additional ice and land based groups are also available.   
+Ice dwelling   
+34) ICE_BACT   
+35) ICE_MIXOTROPHS   
+36) ICE_DIATOMS   
+37) ICE_ZOOBIOTA   
+Land dwelling (vegetation only for now)    
+38) MARSH   
+39) MANGROVE
 
@@
 Looks like it works, output is N for our selected biomass pool species over all polygons, layers, and times.
+Output of `unique(testNpools$species)`:
 
@@
 
-I don't think we have a function that aggregates over layets and converts N to biomass but there are pieces in other functions.
+I don't think we have a function that aggregates over layers and converts N to biomass but there are pieces in other functions.
+
+New `atlantisom` function: `calc_biomass_pool` based on `calc_biomass_age`, but adding the expansion to volume used in `calc_pred_cons`.
+
+```{r}
+
+# we need the volume of each layer as input to this
+vol <- load_nc_physics(dir = dir,
+                         file_nc = nc_out,
+                         physic_variables = "volume",
+                         aggregate_layers = FALSE,
+                         bboxes = boxes)
+
+calc_biomass_pool <- function(pooln, vol, biolprm){
+  datalist <- list(pooln)
+
+  # Conversion factor from mg N to t wet-weight
+  # should only use conversion for non vertebrates
+  # also need volume of cell info
+  bio_conv <- biolprm$redfieldcn * biolprm$kgw2d / 1000000000
+
+  data_names <- c("species", "agecl", "polygon", "layer", "time", "atoutput")
+
+  if (all(sapply(datalist, function(x) all(is.element(names(x), data_names))))){
+
+    names(vol)[names(vol) == "atoutput"] <- "vol"
+
+    pooln <- merge(pooln, vol,
+                   by = c("time", "polygon", "layer"))
+
+    pooln$atoutput <- with(pooln, vol * atoutput * bio_conv)
+
+    # Sum over layers 
+    biomass_pools <- aggregate(atoutput ~ species + agecl + time + polygon,
+      data = pooln, sum)
+  } else {
+    stop(paste("Dataframe names do not match with", data_names))
+  }
+
+  return(biomass_pools)
+}
+
+biolprm <- load_biolprm(dir = dir, file_biolprm = "nordic_biol_incl_harv_v_011_1skg.prm")
+
+biomass_pools <- calc_biomass_pool(pooln = testNpools,
+                                   vol = vol,
+                                   biolprm = biolprm)
+
+```
+
+Do the new biomass pools match the `BiomIndx.txt` output when aggregated over polygon? If so they can be the input to the survey function for time and spatial subsetting.
+
+```{r}
+
+aggbiopools <- biomass_pools %>%
+  dplyr::group_by(species, agecl, time) %>%
+  dplyr::summarise(totpool = sum(atoutput))
+
+txtbiopools <- atlantisom::load_bioind(dir, "nordic_runresults_01BiomIndx.txt", fgs) %>%
+  dplyr::mutate(time = time/73) %>%
+  dplyr::filter(species %in% unique(aggbiopools$species)) %>%
+  dplyr::select(species, time, atoutput)
+
+ggplot(txtbiopools, aes(x=time, y=atoutput)) +
+  geom_line() +
+  geom_point(data = aggbiopools, aes(x = time, y = totpool), colour = "blue") +
+  theme_bw() +
+  facet_wrap(~species, scales = "free_y")
+
+```
+
+What is up with the phytoplankton groups? Small matches but large doesn't.
+
+```{r}
+ggplot(txtbiopools |> filter(species %in% c("Large_phytop", "Small_phytop")), 
+       aes(x=time, y=atoutput)) +
+  geom_line() +
+  geom_point(data = aggbiopools |> filter(species %in% c("Large_phytop", "Small_phytop")), 
+             aes(x = time, y = totpool), colour = "blue") +
+  theme_bw() +
+  facet_wrap(~species, scales = "free_y")
+
+```
+
+Predatory benthos is also off, Detrivore ben to a lesser extent.
+
+```{r}
+ggplot(txtbiopools |> filter(species %in% c("Predatory_ben", "Detrivore_ben")), 
+       aes(x=time, y=atoutput)) +
+  geom_line() +
+  geom_point(data = aggbiopools |> filter(species %in% c("Predatory_ben", "Detrivore_ben")), 
+             aes(x = time, y = totpool), colour = "blue") +
+  theme_bw() +
+  facet_wrap(~species, scales = "free_y")
+
+```

### 2023-06-15 docs/index.Rmd
starts biomass pools sampling for Rpath

@@
 
-### New workflow testing for WGSAM simulated datasets
+### New workflows testing for WGSAM simulated datasets
 
 <a href="https://sgaichas.github.io/poseidon-dev/NewCatchWorkflow.html" target="_blank"> Total catch in tons now by polygon, fleet, and subannual timestep, in honor of Norwegian Constitution Day, May 17, 2023</a>
+
+<a href="https://sgaichas.github.io/poseidon-dev/NewBiomassPoolsWorkflow.html" target="_blank"> Biomass pool groups (lower trophic levels, benthic invertebrates) outputs for "survey" sampling to feed food web models, June 15, 2023</a>
 

### 2023-08-22 NewBiomassPoolsWorkflow.Rmd
diagnose mismatch biopools to txt output

@@
 
-```{r}
+```{r, fig.width=10}
 
@@
 
-What is up with the phytoplankton groups? Small matches but large doesn't.
+## Diagnose mismatches between biopools and BiomInd.txt output
+
+Who matches exactly (enough).
+Patterns in absolute mismatch are interesting, some species are really close and some always way off.
+Volume wrong, depth layers wrong?
+Mine are always higher than the txt output when they are off, so maybe too much expansion somewhere.
+
+Groups that match seem more likely to be pelagic (dinoflagellates, all zooplankton, pelagic bacteria, labile and refractive detritus) although large phytoplankton is way off.
+
+```{r, warning=FALSE, fig.width=10}
+mismatch <- txtbiopools |>
+  dplyr::left_join(aggbiopools) |>
+  dplyr::mutate(mismatch = atoutput - totpool,
+                bad = ifelse(abs(mismatch) > 0.1*(atoutput), 1, 0))
+
+ggplot(mismatch, aes(x=time, y=mismatch)) +
+  geom_line() +
+  theme_bw() +
+  facet_wrap(~species, scales = 'free_y')
+```
+
+This is mismatch coded as 0 if less than 10% of txt output, 1 if more.
+Those exceeding are carrion, predatory benthos, detritivore benthos, king crab, benthic bacteria, large phytoplankton and sometimes small phytoplankton.
+
+```{r, warning=FALSE}
+
+ggplot(mismatch, aes(x=time, y=bad)) +
+  geom_line() +
+  theme_bw() +
+  facet_wrap(~species)
+```
+
+What is up with the phytoplankton groups? Small matches well visually but large doesn't. Large phytoplankton is the largest magnitude mismatch across all groups.
 
@@
 
-Predatory benthos is also off, Detrivore ben to a lesser extent.
+Predatory benthos is off by what appears to be a scalar, Detrivore ben converges as they increase.
 
@@
 ```
+
+## Remaining changes to workflow, once corrected
+
+### Add `calc_biomass_pool` to `run_truth`
+
+### Test wrapper functions with biomass pool groups

### 2023-08-23 NewBiomassPoolsWorkflow.Rmd
correction for benthos

@@
 
-calc_biomass_pool <- function(pooln, vol, biolprm){
+# we need area for the benthic inverts instead of volume
+# THANK YOU JOE for this tip
+area <- load_boxarea(dir = dir,
+                     file_bgm = "Nordic02.bgm")
+
+# we need a list of which group types to expand to area instead of volume
+benthos <- c("SED_EP_FF", "SED_EP_OTHER", "SED_BAC", 
+             "MOB_EP_OTHER", "LG_INF", "SM_INF", 
+             "MICROPHYTBENTHOS", "SEAGRASS")
+
+calc_biomass_pool <- function(pooln, vol, area, fgs, biolprm){
   datalist <- list(pooln)
@@
   bio_conv <- biolprm$redfieldcn * biolprm$kgw2d / 1000000000
+
+  # get fgs info
+    # use grouptype column to allocate
+  colnames(fgs) <- tolower(colnames(fgs))
+
+  # check for GroupType or InvertType
+  if (!("grouptype" %in% colnames(fgs) | "inverttype"%in% colnames(fgs))) {
+    stop(paste("The columns GroupType or InvertType ars not in your functional\n",
+               "groups file."))
+  }
+
+  # change inverttype to grouptype, contents should be the same
+  if("inverttype" %in% names(fgs)) names(fgs)[names(fgs) == 'inverttype'] <- 'grouptype'
+
+  fgs$grouptype <- tolower(fgs$grouptype)
+
+  pooltype <- fgs |>
+    dplyr::select(species=name, grouptype) |>
+    dplyr::mutate(pooltype = dplyr::case_when((grouptype %in% c("lg_inf",
+                                                                "sm_inf",
+                                                                "sed_bact",
+                                                                "sed_ep_ff",
+                                                                "sed_ep_other",
+                                                                "mobile_ep_other",
+                                                                "coral",
+                                                                "sponge")) ~ "benthos",
+                                              TRUE ~ "nonbenth"))
 
@@
 
-    pooln$atoutput <- with(pooln, vol * atoutput * bio_conv)
+    pooln <- merge(pooln, area,
+                   by = c("polygon"))
+
+    pooln <- merge(pooln, pooltype,
+                   by = c("species"))
+
+    #pooln$atoutput <- with(pooln, vol * atoutput * bio_conv)
+
+    pooln <- pooln |>
+      dplyr::mutate(atoutput = dplyr::case_when(pooltype == "nonbenth" ~ vol * atoutput * bio_conv,
+                                                pooltype == "benthos" ~ 
+                                                  ifelse(layer==7, area * atoutput * bio_conv, 0)))
 
@@
                                    vol = vol,
+                                   area = area,
+                                   fgs = fgs,
                                    biolprm = biolprm)
@@
   geom_line() +
-  geom_point(data = aggbiopools, aes(x = time, y = totpool), colour = "blue") +
+  geom_point(data = aggbiopools, aes(x = time, y = totpool), colour = "blue", alpha=0.1) +
   theme_bw() +
@@
 
+Joe pointed out that benthic groups should be expanded to area rather than volume. That correction has now been applied, and matches are better, but still diverge over time for some groups.
+
 Groups that match seem more likely to be pelagic (dinoflagellates, all zooplankton, pelagic bacteria, labile and refractive detritus) although large phytoplankton is way off.
@@
   dplyr::mutate(mismatch = atoutput - totpool,
-                bad = ifelse(abs(mismatch) > 0.1*(atoutput), 1, 0))
+                bad = ifelse(abs(mismatch) > 0.01*(atoutput), 1, 0))
 
@@
 
-This is mismatch coded as 0 if less than 10% of txt output, 1 if more.
-Those exceeding are carrion, predatory benthos, detritivore benthos, king crab, benthic bacteria, large phytoplankton and sometimes small phytoplankton.
+This is mismatch coded as 0 if less than 1% of txt output, 1 if more.
+Those exceeding are carrion, predatory benthos, detritivore benthos, king crab, benthic bacteria, large phytoplankton and sometimes small zooplankton.
 
@@
   geom_point(data = aggbiopools |> filter(species %in% c("Predatory_ben", "Detrivore_ben")), 
-             aes(x = time, y = totpool), colour = "blue") +
+             aes(x = time, y = totpool), colour = "blue", alpha=0.1) +
   theme_bw() +

### 2023-10-04 NewBiomassPoolsWorkflow.Rmd
fix sedlayer bug but calc still off for LG_PHY

@@
 
+    sedlayer <- max(vol$layer)
+
     pooln <- merge(pooln, vol,
@@
                                                 pooltype == "benthos" ~ 
-                                                  ifelse(layer==7, area * atoutput * bio_conv, 0)))
+                                                  ifelse(layer==sedlayer, area * atoutput * bio_conv, 0)))
 
@@
 
-```{r, warning=FALSE}
+```{r, warning=FALSE, fig.width=10}
 
@@
 
+### Try NEUS?
+
+```{r}
+dir <- here("atlantisoutput/NEUSv2.1.0")
+file_nc <- "neus_output.nc"
+
+file.nc <- file.path(dir, file_nc)
+
+ # Load ATLANTIS output!
+  at_out <- RNetCDF::open.nc(con = file.nc)
+  #on.exit(RNetCDF::close.nc(at_out))
+
+    # Get info from netcdf file! (Filestructure and all variable names)
+  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
+    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
+  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
+  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
+  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length
+
+```
+
+For NEUS Atlantis, these are the biomass pools:
+
+```{r}
+
+fgs <- atlantisom::load_fgs(here("atlantisoutput/NEUSv2.1.0"), "neus_groups.csv")
+
+pools <- fgs %>%
+  dplyr::filter(NumCohorts==1) %>%
+  dplyr::select(Code, Name, LongName, NumCohorts, GroupType)
+
+knitr::kable(pools)
+
+```
+
+Here are the relevant outputs in the output.nc file for these groups:
+
+```{r}
+# return everything in var_names_ncdf matching Name in pools
+
+var_names_ncdf[str_detect(var_names_ncdf, str_c(pools$Name, collapse ="|"))]
+
+```
+NEUS has some Cover types as well as N outputs.
+
+```{r}
+
+select_groups <- pools$Name
+nc_out <- file_nc
+bps <- load_bps(dir = dir, fgs = "neus_groups.csv", file_init = "neus_init.nc")
+  # Get the boundary boxes
+  allboxes <- load_box(dir = dir, file_bgm = "neus_tmerc_RM2.bgm")
+  boxes <- get_boundary(allboxes)
+
+testNpools <- load_nc(dir = dir,
+                  file_nc = nc_out,
+                  bps = bps,
+                  fgs = fgs,
+                  select_groups = select_groups,
+                  select_variable = "N",
+                  check_acronyms = TRUE,
+                  bboxes = boxes)
+
+unique(testNpools$species)
+```
+```{r}
+
+# we need the volume of each layer as input to this
+vol <- load_nc_physics(dir = dir,
+                         file_nc = nc_out,
+                         physic_variables = "volume",
+                         aggregate_layers = FALSE,
+                         bboxes = boxes)
+
+# we need area for the benthic inverts instead of volume
+# THANK YOU JOE for this tip
+area <- load_boxarea(dir = dir,
+                     file_bgm = "neus_tmerc_RM2.bgm")
+
+biolprm <- load_biolprm(dir = dir, file_biolprm = "at_biology.prm")
+
+biomass_pools <- calc_biomass_pool(pooln = testNpools,
+                                   vol = vol,
+                                   area = area,
+                                   fgs = fgs,
+                                   biolprm = biolprm)
+
+```
+
+Do the new biomass pools match the `BiomIndx.txt` output when aggregated over polygon? If so they can be the input to the survey function for time and spatial subsetting.
+
+```{r, fig.width=10}
+
+aggbiopools <- biomass_pools %>%
+  dplyr::group_by(species, agecl, time) %>%
+  dplyr::summarise(totpool = sum(atoutput))
+
+txtbiopools <- atlantisom::load_bioind(dir, "neus_outputBiomIndx.txt", fgs) %>%
+  dplyr::mutate(time = time/73) %>%
+  dplyr::filter(species %in% unique(aggbiopools$species)) %>%
+  dplyr::select(species, time, atoutput)
+
+ggplot(txtbiopools, aes(x=time, y=atoutput)) +
+  geom_line() +
+  geom_point(data = aggbiopools, aes(x = time, y = totpool), colour = "blue", alpha=0.1) +
+  theme_bw() +
+  facet_wrap(~species, scales = "free_y")
+
+```
+mismatch
+
+```{r, warning=FALSE, fig.width=10}
+mismatch <- txtbiopools |>
+  dplyr::left_join(aggbiopools) |>
+  dplyr::mutate(mismatch = atoutput - totpool,
+                bad = ifelse(abs(mismatch) > 0.01*(atoutput), 1, 0))
+
+ggplot(mismatch, aes(x=time, y=mismatch)) +
+  geom_line() +
+  theme_bw() +
+  facet_wrap(~species, scales = 'free_y')
+```
+This is mismatch coded as 0 if less than 1% of txt output, 1 if more.
+
+```{r, warning=FALSE, fig.width=10}
+
+ggplot(mismatch, aes(x=time, y=bad)) +
+  geom_line() +
+  theme_bw() +
+  facet_wrap(~species)
+```
+
+Learned that we should not hardcode the sediment layer! Added a sedlayer variable to the function.
+
+BUT. Diatoms are still off so this calculation may not be working for that category.
+
+What is up with the phytoplankton groups? Small matches well visually but large doesn't. Large phytoplankton is the largest magnitude mismatch across all groups.
+
+```{r}
+ggplot(txtbiopools |> filter(species %in% c("Diatom", "DinoFlag")), 
+       aes(x=time, y=atoutput)) +
+  geom_line() +
+  geom_point(data = aggbiopools |> filter(species %in% c("Diatom", "DinoFlag")), 
+             aes(x = time, y = totpool), colour = "blue") +
+  theme_bw() +
+  facet_wrap(~species, scales = "free_y")
+
+```
+
 ## Remaining changes to workflow, once corrected
@@
 
+With an if statement?
+
 ### Test wrapper functions with biomass pool groups
+
+Only need species and index here, comps don't apply.

### 2023-10-04 NewBiomassPoolsWorkflow.Rmd
fixed lg_phy, ready to go

@@
 
+With the following exceptions!
+
+Benthic groups have N expanded to area
+
+LG_PHY type has N expanded to only water column, not sediment layers.
+
 ```{r}
@@
                                                                 "sed_ep_other",
-                                                                "mobile_ep_other",
+                                                                "mob_ep_other",
                                                                 "coral",
                                                                 "sponge")) ~ "benthos",
-                                              TRUE ~ "nonbenth"))
+                                              (grouptype %in% c("lg_phy")) ~ "lg_phy",
+                                              TRUE ~ "alllayers"))
 
@@
     pooln <- pooln |>
-      dplyr::mutate(atoutput = dplyr::case_when(pooltype == "nonbenth" ~ vol * atoutput * bio_conv,
+      dplyr::mutate(atoutput = dplyr::case_when(pooltype == "alllayers" ~ vol * atoutput * bio_conv,
+                                                pooltype == "lg_phy" ~
+                                                  ifelse(layer!=sedlayer, vol * atoutput * bio_conv, 0),
                                                 pooltype == "benthos" ~
@@
 
-Joe pointed out that benthic groups should be expanded to area rather than volume. That correction has now been applied, and matches are better, but still diverge over time for some groups.
+Joe pointed out that benthic groups should be expanded to area rather than volume.
+That correction has now been applied, and matches are better, but still diverge over time for some groups.
+Not enough to worry about I think.
 
-Groups that match seem more likely to be pelagic (dinoflagellates, all zooplankton, pelagic bacteria, labile and refractive detritus) although large phytoplankton is way off.
+Originally, large phytoplankton was way off, but Joe also pointed out that LG_PHY N can be in the sediment layer but shouldn't "count" as biomass.
+Therefore, this group is expanded to all layers but the sediment layer and that creates a much better match.
 
@@
 This is mismatch coded as 0 if less than 1% of txt output, 1 if more.
-Those exceeding are carrion, predatory benthos, detritivore benthos, king crab, benthic bacteria, large phytoplankton and sometimes small zooplankton.
 
@@
 
-What is up with the phytoplankton groups? Small matches well visually but large doesn't. Large phytoplankton is the largest magnitude mismatch across all groups.
+Large phytoplankton formerly had the largest magnitude mismatch across all groups, now fixed by not expanding N in the sediment layer for this group type.
 
@@
 
-Predatory benthos is off by what appears to be a scalar, Detrivore ben converges as they increase.
+These groups seem close enough to me now.
 
@@
 
-BUT. Diatoms are still off so this calculation may not be working for that category.
-
-What is up with the phytoplankton groups? Small matches well visually but large doesn't. Large phytoplankton is the largest magnitude mismatch across all groups.
+With corrections for LG_PHY things now line up in both models.
 

### 2023-10-04 NewBiomassPoolsWorkflow.Rmd
outline next steps done and in progress

@@
 
-With an if statement?
+With an if statement? Or let the species selection take care of that.
+
+Decision: add it and if no biomass pools species selected it should return nothing
+
+Added to `run_truth`:
+
++  area using existing `load_boxarea`
++  pooln using existing `load_nc` with select_variable = "N"
++  biomass_pools using new `calc_biomass_pool`
+
+Returns biomass_pools as part of truth list
 
@@
 Only need species and index here, comps don't apply.
+
+In `om_species` need to check for biomass pools output and append that to the `truebio_ss` object? If one object can use existing `om_index` on bio pools.
+
+```{r}
+  # biomass at agecl at full resolution (all polygons and layers)
+  #truebio_ss <- omlist$truth$biomass_ages[omlist$truth$biomass_ages$species %in% species_ss,]
+```
+
+If a single true bio object is created, may have to modify selectivty definition for biomass pools in `om_index`

### 2023-10-05 NewBiomassPoolsWorkflow.Rmd
run truth works now, om_init works too

@@
 
+Testing--have to ask for N data for only biomass pool groups. If there are none we should trap this.
+
+```{r, eval=FALSE}
+
+dir <- here("atlantisoutput","NOBA_sacc_30")
+file_fgs <- "nordic_groups_v04.csv"
+biomass.pools.file <- "nordic_biol_v23.nc"
+file_biolprm <- "nordic_biol_incl_harv_v_011_1skg.prm"
+file_bgm <- "Nordic02.bgm"
+file_init <- "nordic_biol_v23.nc"
+file_runprm <- "nordic_run_v04.xml"
+scenario <- "nordic_runresults_01"
+verbose <- TRUE
+annage <- TRUE
+file_fish <- "NoBAFisheries.csv"
+
+  # Read in information
+  # Read in the functional groups csv since that is used by many functions
+  fgs <- load_fgs(dir = dir, file_fgs = file_fgs)
+
+    #Get just the names of active functional groups
+  funct.group.names <- fgs %>%
+    filter(IsTurnedOn == 1) %>%
+    select(Name) %>%
+    .$Name
+
+  select_groups <- funct.group.names
+
+  # Read in the biomass pools
+  bps <- load_bps(dir = dir, fgs = file_fgs, file_init = file_init)
+  # Read in the biological parameters
+  biol <- load_biolprm(dir = dir, file_biolprm = file_biolprm)
+  # Read in the run parameters
+  runprm <- load_runprm(dir = dir, file_runprm = file_runprm)
+
+  nc_catch <- paste0(scenario, 'CATCH.nc')
+  dietcheck <- paste0(scenario, 'DietCheck.txt')
+  nc_out <- paste0(scenario, ".nc")
+  nc_prod <- paste0(scenario, "PROD.nc")
+  file_catchfish <- file.path(dir,
+    paste0(scenario, "CatchPerFishery.txt"))
+  file_catch <- paste0(scenario, "Catch.txt")
+
+  if(annage){
+    if(!file.exists(paste0(file.path(dir,paste0(scenario, 'ANNAGEBIO.nc'))))){
+      stop("ANNAGEBIO.nc file not found")
+    }
+    if(!file.exists(paste0(file.path(dir,paste0(scenario, 'ANNAGECATCH.nc'))))){
+      stop("ANNAGECATCH.nc file not found")
+    }
+    nc_annagebio <- paste0(scenario, 'ANNAGEBIO.nc')
+    nc_annagecatch <- paste0(scenario, 'ANNAGECATCH.nc')
+  }
+
+  # Get the boundary boxes
+  allboxes <- load_box(dir = dir, file_bgm = file_bgm)
+  boxes <- get_boundary(allboxes)
+
+  # Get box area for benthic biomass pool calc
+  area <- load_boxarea(dir = dir,
+                       file_bgm = file_bgm)
+
+  #Extract from NetCDF files
+  # Need: dir, file_nc, bps, fgs, select_groups, select_variable,
+  # check_acronyms, bboxes
+
+  nums <- load_nc(dir = dir,
+                  file_nc = nc_out,
+                  bps = bps,
+                  fgs = fgs,
+                  select_groups = select_groups,
+                  select_variable = "Nums",
+                  check_acronyms = TRUE,
+                  bboxes = boxes)
+  if(verbose) message("Numbers read in.")
+
+  resn <- load_nc(dir = dir,
+                  file_nc = nc_out,
+                  bps = bps,
+                  fgs = fgs,
+                  select_groups = select_groups,
+                  select_variable = "ResN",
+                  check_acronyms = TRUE,
+                  bboxes = boxes)
+  if(verbose) message("Reserve nitrogen read in.")
+
+  structn <- load_nc(dir = dir,
+                  file_nc = nc_out,
+                  bps = bps,
+                  fgs = fgs,
+                  select_groups = select_groups,
+                  select_variable = "StructN",
+                  check_acronyms = TRUE,
+                  bboxes = boxes)
+  if(verbose) message("Structural nitrogen read in.")
+
+  eat <- load_nc(dir = dir,
+                     file_nc = nc_prod,
+                     bps = bps,
+                     fgs = fgs,
+                     select_groups = select_groups,
+                     select_variable = "Eat",
+                     check_acronyms = TRUE,
+                     bboxes = boxes)
+  if(verbose) message("Eaten read in.")
+
+  grazing <- load_nc(dir = dir,
+                 file_nc = nc_prod,
+                 bps = bps,
+                 fgs = fgs,
+                 select_groups = select_groups,
+                 select_variable = "Grazing",
+                 check_acronyms = TRUE,
+                 bboxes = boxes)
+  if(verbose) message("Grazing read in.")
+
+  vol <- load_nc_physics(dir = dir,
+                         file_nc = nc_out,
+                         physic_variables = "volume",
+                         aggregate_layers = FALSE,
+                         bboxes = boxes)
+  if(verbose) message("Volume read in.")
+
+  # biomass pools added 2023 for Rpath testing
+  pools <- fgs %>%
+    dplyr::filter(NumCohorts == 1,
+                  IsTurnedOn == 1) %>%
+    dplyr::select(Name) %>%
+    .$Name
+
+  pooln <- load_nc(dir = dir,
+                   file_nc = nc_out,
+                   bps = bps,
+                   fgs = fgs,
+                   select_groups = pools,
+                   select_variable = "N",
+                   check_acronyms = TRUE,
+                   bboxes = boxes)
+ if(verbose) message("Biomass pools N read in.")
+
+```
+
 ### Test wrapper functions with biomass pool groups
@@
 
-In `om_species` need to check for biomass pools output and append that to the `truebio_ss` object? If one object can use existing `om_index` on bio pools.
+In `om_species` need to check for biomass pools output and append that to the `truebio_ss` object?
+If one object can use existing `om_index` on bio pools.
+However, will that be a problem if trying to combine the truenums and truebio objects?
+They won't be the same length.
+Perhaps leave separate at this level and combine only if needed.
 
@@
 
-If a single true bio object is created, may have to modify selectivty definition for biomass pools in `om_index`
+If a single true bio object is created, may have to modify selectivty definition for biomass pools in `om_index`. If left separate can apply no selectivity to biomass pool groups, only efficiency.

### 2023-10-07 NewBiomassPoolsWorkflow.Rmd
got biopools integrated

@@
 
-```{r}
-  # biomass at agecl at full resolution (all polygons and layers)
-  #truebio_ss <- omlist$truth$biomass_ages[omlist$truth$biomass_ages$species %in% species_ss,]
+If a single true bio object is created, may have to modify selectivty definition for biomass pools in `om_index`. If left separate can apply no selectivity to biomass pool groups, only efficiency.
+
+Decision: leave the `truebiopool_ss` object separate and process separately with a different survey config file and a separate call to `om_index`.
+Or just bind them together for biomass based surveys if it doesn't break anything.
+
+Running `om_init` from the `mskeyrun` repo with my local NOBA atlantis outputs produced expected output:
+
 ```
 
-If a single true bio object is created, may have to modify selectivty definition for biomass pools in `om_index`. If left separate can apply no selectivity to biomass pool groups, only efficiency.
+> NOBAom <- om_init(here("data-raw/simulated-data/config/NOBA_sacc38Config.R"))
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01PROD.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01PROD.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01CATCH.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01CATCH.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01CATCH.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01ANNAGEBIO.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01ANNAGECATCH.nc successfully"
+[1] "Read /Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38/nordic_runresults_01ANNAGECATCH.nc successfully"
+> NOBAom$truth$biomass_pools
+               species agecl time polygon     atoutput
+1             Ben_bact     1    0       1 1.376317e+05
+2        Detrivore_ben     1    0       1 1.088091e+05
+3             Dinoflag     1    0       1 1.312067e+04
+4      Gel_zooplankton     1    0       1 1.569070e+05
+5              Lab_det     1    0       1 2.581300e+07
+6         Large_phytop     1    0       1 3.268895e+05
+7    Large_zooplankton     1    0       1 4.772586e+05
+8   Medium_zooplankton     1    0       1 1.118751e+06
+9             Pel_bact     1    0       1 3.268895e+04
+10       Predatory_ben     1    0       1 1.088091e+05
+11             Ref_det     1    0       1 8.904920e+06
+12        Small_phytop     1    0       1 3.280167e+05
+13   Small_zooplankton     1    0       1 9.152906e+05
+14             Sponges     1    0       1 4.621541e+05
+15            Ben_bact     1    1       1 8.875378e+05
+16     Benthic_filterf     1    1       1 1.163897e-04
+17             Carrion     1    1       1 1.700432e-10
+18              Corals     1    1       1 1.148013e-04
+19       Detrivore_ben     1    1       1 1.096034e+05
+20            Dinoflag     1    1       1 4.509383e+01
+21     Gel_zooplankton     1    1       1 1.563221e+05
+22           King_crab     1    1       1 0.000000e+00
+23             Lab_det     1    1       1 8.511963e+07
+24        Large_phytop     1    1       1 8.999719e+03
+25   Large_zooplankton     1    1       1 5.237310e+05
+26  Medium_zooplankton     1    1       1 5.390765e+06
+27            Pel_bact     1    1       1 1.230456e+06
+28       Predatory_ben     1    1       1 1.089640e+05
+29             Ref_det     1    1       1 8.142682e+06
+30        Small_phytop     1    1       1 1.177500e+03
+31   Small_zooplankton     1    1       1 7.373987e+07
+32             Sponges     1    1       1 4.647222e+05
+```
+
+Added lines to `om_species` to include the truebiopool outputs:
+
+```{r, code = readLines("https://raw.githubusercontent.com/r4atlantis/atlantisom/dev/R/om_species.R"), eval=F}
+
+```
+
+Now test modified `om_species` in `mskeyrun` repo:
+
+```
+> NOBAom_fw <- om_species(fwspp$Name, NOBAom, save = FALSE) #save by hand, don't overwrite 11 species omlist_ss
+> NOBAom_fw$truebiopool_ss
+               species agecl time polygon     atoutput
+1             Ben_bact     1    0       1 1.376317e+05
+2        Detrivore_ben     1    0       1 1.088091e+05
+3             Dinoflag     1    0       1 1.312067e+04
+4      Gel_zooplankton     1    0       1 1.569070e+05
+5              Lab_det     1    0       1 2.581300e+07
+6         Large_phytop     1    0       1 3.268895e+05
+7    Large_zooplankton     1    0       1 4.772586e+05
+8   Medium_zooplankton     1    0       1 1.118751e+06
+9             Pel_bact     1    0       1 3.268895e+04
+10       Predatory_ben     1    0       1 1.088091e+05
+11             Ref_det     1    0       1 8.904920e+06
+12        Small_phytop     1    0       1 3.280167e+05
+13   Small_zooplankton     1    0       1 9.152906e+05
+14             Sponges     1    0       1 4.621541e+05
+15            Ben_bact     1    1       1 8.875378e+05
+16     Benthic_filterf     1    1       1 1.163897e-04
+17             Carrion     1    1       1 1.700432e-10
+18              Corals     1    1       1 1.148013e-04
+19       Detrivore_ben     1    1       1 1.096034e+05
+20            Dinoflag     1    1       1 4.509383e+01
+21     Gel_zooplankton     1    1       1 1.563221e+05
+22           King_crab     1    1       1 0.000000e+00
+23             Lab_det     1    1       1 8.511963e+07
+24        Large_phytop     1    1       1 8.999719e+03
+25   Large_zooplankton     1    1       1 5.237310e+05
+26  Medium_zooplankton     1    1       1 5.390765e+06
+27            Pel_bact     1    1       1 1.230456e+06
+28       Predatory_ben     1    1       1 1.089640e+05
+29             Ref_det     1    1       1 8.142682e+06
+30        Small_phytop     1    1       1 1.177500e+03
+31   Small_zooplankton     1    1       1 7.373987e+07
+32             Sponges     1    1       1 4.647222e+05
+33            Ben_bact     1    2       1 3.080604e+06
+34     Benthic_filterf     1    2       1 1.199844e-04
+35             Carrion     1    2       1 1.698761e-10
+36              Corals     1    2       1 1.168009e-04
+```
+
+Now try a new `om_index` that rbinds the truebio and truebiopool objects together for a biomass based survey
+
+```{r, code = readLines("https://raw.githubusercontent.com/r4atlantis/atlantisom/dev/R/om_index.R"), eval=F}
+
+```
+
+It works as tested from the `mskeyrun` SimData.Rmd (both age structured and biopool groups are in the output):
+
+```
+> NOBAom_fw_ind$survObsBiomB$BTS_fall_allbox_effic1_fw[[1]] |> dplyr::group_by(species) |> dplyr::summarise(survBtot = sum(atoutput))
+# A tibble: 40 × 2
+   species         survBtot
+   <chr>              <dbl>
+ 1 Beard_seal      8.14e+ 4
+ 2 Ben_bact        9.36e+12
+ 3 Benthic_filterf 7.86e+10
+ 4 Carrion         7.46e- 7
+ 5 Corals          3.19e+11
+ 6 Demersal_large  6.47e+ 6
+ 7 Demersals_other 1.81e+ 8
+ 8 Detrivore_ben   2.59e+10
+ 9 Dinoflag        1.31e+ 6
+10 Fin_whale       3.87e+ 7
+# … with 30 more rows
+# ℹ Use `print(n = ...)` to see more rows
+```
+
+Plots of the new `mskeyrun::simSurveyIndexFW` object are [here](https://noaa-edab.github.io/RpathSim/ParametersforRpath.html).

### 2023-12-14 PerCapitaConsumption.Rmd
knits
WIP evaluate consumption estimates

@@
 source(here("config/NOBA2Config.R"))
+
+# this hardcodes a d.name in a different folder because I don't want multiple copies
+
+if(!dir.exists(d.name)){
+  d.name <- "/Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_march_2020"
+}
 
@@
 
+  # look directly at file
+  prod.nc = ncdf4::nc_open(file.path(dir, nc_prod))
+
+  names(prod.nc$var)
+
   eat <- load_nc(dir = dir,
@@
 
-This is the function code
+This is the function code. Consumption still seems low so lets not exlude sediment layer.
+
+```{r}
+calc_pred_cons <- function(eat, grazing, vol, fgs, biolprm, runprm){
+
+  # Conversion factor from mg N to tonnes wet-weight
+  bio_conv <- biolprm$redfieldcn * biolprm$kgw2d / 1000000000
+
+  # Eat and grazing are per species, agecl, polygon, and time.
+  # Therefore, we need to aggregate the vol over layers.
+  # Eat calculated in atecology/atprocess.C
+  # Eat( ) is defined here : \trunk\atlantis\atecology\atprocess.c(644)
+
+  # SKG: The main question is the UNITS of eat and grazing.
+  # manual says this is either mg/m3/day or mg/day/individ
+  # calculations here assume mg/m3/day so multiplying by box vol
+  # is this m3 assuming the entire box volume or only where groups feed?
+  # do we need to match species to layers to get appropriate volume?
+  # VERTnight_XXX and VERTday_XXX in biol.prm
+
+  # biomass pool groups may get need volume expansion while
+  # age structured are per capita by age group?
+
+  # all layers? even sediment? try filtering out layer > 7 to remove sediment volume
+  # WARNING: ensure that layer 7 (processed via atlantisom) isn't unique to NOBA
+  # layer 7 is unique to NOBA, use max of layers as sediment layer
+  # however, unclear this is necessary, try without
+
+  colnames(fgs) <- tolower(colnames(fgs))
+
+  # check for GroupType or InvertType
+  if (!("grouptype" %in% colnames(fgs) | "inverttype"%in% colnames(fgs))) {
+    stop(paste("The columns GroupType or InvertType ars not in your functional\n",
+               "groups file."))
+  }
+
+  # change inverttype to grouptype, contents should be the same
+  if("inverttype" %in% names(fgs)) names(fgs)[names(fgs) == 'inverttype'] <- 'grouptype'
+
+  fgs$grouptype <- tolower(fgs$grouptype)
+
+  # is there a different calculation for benthos? don't expand to volume?
+  # not using this yet
+  pooltype <- fgs |>
+    dplyr::select(species=name, grouptype) |>
+    dplyr::mutate(pooltype = dplyr::case_when((grouptype %in% c("lg_inf",
+                                                                "sm_inf",
+                                                                "sed_bact",
+                                                                "sed_ep_ff",
+                                                                "sed_ep_other",
+                                                                "mob_ep_other",
+                                                                "coral",
+                                                                "sponge")) ~ "benthos",
+                                              (grouptype %in% c("lg_phy")) ~ "lg_phy",
+                                              TRUE ~ "alllayers"))
+  # not using this either
+  sedlayer <- max(vol$layer)
+
+  vol <- aggregate(atoutput ~ polygon + time, data = vol, sum)
+  colnames(vol)[colnames(vol) == "atoutput"] <- "vol"
+
+  # Combine eat and grazing! Calculate eaten biomass
+  biomass_eaten <- rbind(eat, grazing)
+  biomass_eaten <- merge(biomass_eaten, vol,
+    by = c("time", "polygon"))
+
+  # todo: previously AK noted that the bio_eaten values were too high
+  # the new values need to be checked to see if they give realistic
+  # results
+  # removed layer 7 in volume which should expand mg N to only water?
+  # total consumption and per capita from this look ok
+  # snapshot so should be daily value?
+
+  biomass_eaten$bio_eaten <- with(biomass_eaten,
+    atoutput * vol * bio_conv)
+
+  biomass_eaten$atoutput <- biomass_eaten$bio_eaten
+
+  return(biomass_eaten)
+}
+
+```
 
@@
 
-This is the current result
+This is the current result compared with all layers
 
@@
 
-biomass_eaten <- calc_pred_cons(eat = eat, grazing = grazing, vol = vol, biolprm = biol, runprm = runprm)
+biomass_eaten <- calc_pred_cons(eat = eat, grazing = grazing, vol = vol, fgs = fgs, biolprm = biol, runprm = runprm)
+
+biomass_eaten_current <- atlantisom::calc_pred_cons(eat = eat, grazing = grazing, vol = vol, biolprm = biol, runprm = runprm)
 
@@
 
+Keeping the sediment layer in makes a small difference.
+
+```
+> head(biomass_eaten_current)
+  time polygon        species agecl    atoutput          vol   bio_eaten
+1    1       1 Long_rough_dab     1   0.6127621 2.865986e+13   0.6127621
+2    1       1 Long_rough_dab     8  20.7267353 2.865986e+13  20.7267353
+3    1       1 Long_rough_dab    10   6.1992620 2.865986e+13   6.1992620
+4    1       1        Haddock     2  77.4729324 2.865986e+13  77.4729324
+5    1       1         Saithe     9   9.1759148 2.865986e+13   9.1759148
+6    1       1         Saithe     3 173.4369707 2.865986e+13 173.4369707
+> head(biomass_eaten)
+  time polygon        species agecl    atoutput          vol   bio_eaten
+1    1       1 Long_rough_dab     1   0.6148762 2.875874e+13   0.6148762
+2    1       1 Long_rough_dab     8  20.7982434 2.875874e+13  20.7982434
+3    1       1 Long_rough_dab    10   6.2206497 2.875874e+13   6.2206497
+4    1       1        Haddock     2  77.7402172 2.875874e+13  77.7402172
+5    1       1         Saithe     9   9.2075721 2.875874e+13   9.2075721
+6    1       1         Saithe     3 174.0353353 2.875874e+13 174.0353353
+```
+
 Plotting a cod example total consumption by age class
